### 2022-11-11 script/manuscript_MEE.Rmd
reedme

---
title: Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
  Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
author: |
  | Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$, Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
abstract: |
  an abstract will be written to fill this space.
output:
  pdf_document: default
  toc: no
csl: ../citations/methods-in-ecology-and-evolution.csl
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

```{r Required Libraries, message=F, warning=F, results='hide', echo = F}
library(here)
library(tidyverse)
library(bookdown)
library(knitr)
library(kableExtra)
```

```{r, eval = F, echo = F}
set_here('../')

citation('usdm')
print(citation('iNEXT'), bibtex=TRUE)
```

# INTRODUCTION
The inability to reliably identify plants to the level of species often leaves our understanding of ecosystem function and interactions nebulous. Current methods to remediate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots for diagnostic material, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques. These approaches are untenable in light of the benefits offered by species in several complex genera which serve as bioindicators, and preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts. Many genera have species which are well defined based upon ecological rather than morphological properties, the identification of these taxa in degraded areas or without their mutualistic partners is fraught with difficulty, hindering an understanding of the breadth of habitat which some species occupy, and the interactions they have with other species. 
  The identification of many plant species to terminal taxon is an essential component of nearly all land management programs, where many species in the same genus  (e.g. Sagebrush - *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) serve as bioindicators, as well as in academic research (@Gage2013HistoricalRO, AIM). This endeavour is often mired by lack of diagnostic characters (e.g. flowers, fruits, or roots),  and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic). Solutions to this problem are wanting, certain programmes have relied increasingly upon revisiting field sites to identify material using morphological or chemical approaches, whereas academic research has often used high copy number plastid genes as barcodes (Rosentreter et al. 2021, MORE MORE). However, both approaches have significant downsides, the former resource intensive at the landscape scale - and often does not work, while the latter seldom works due to a lack of variability in the currently available barcodes (@liu2014identification).
  Recently barcoding, and metabarcoding, have shown much promise in all Kingdoms of life. For example …. . With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification), most studies seem to be in between this spectrum (CITE). Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence.... and... Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling.

  Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases. While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from Herbaria, (typically Governmental) survey work, and citizen science from a domain exceeding the study area. To these candidate species an approach, such as logistic regression, may be used to identify distances under which taxa are worth further exploring. To these candidate species, we generate species distribution models, which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results. This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers. Moreover, as most next-generation sequence data is deposited as raw-sequence reads, from a processing perspective, it is essential to reduce the candidate species via an approach as such.
  Concomitant to next-generation sequencing has been high-throughput sequencing (cite). Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021exploring). The dataset it creates will contain Hyb-Seq data from at least one species representing each genus in the Plant Kingdom using the popular A353 probes (). These widely available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding. 
  *A brief consideration and example on the inclusion of the temporal dimension of plant flowering, which we suspect will be of importance in sub-tropical, and tropical ecosystems - and others with high seasonality in flowering, is also discussed.* 
  We show that by combining both candidate plant species, and representative sequence data, with only a small amount of new reference sequencing, one is capable of generating ecologically realistic, and biologically relevant results. To test these assertions we utilize a long-term observational study of Bumble Bees, the most important genus of pollinators in the temperate, and one with consistently concerning conservation assessments (@cameron2020global, @goulson2008decline). By generating sequence data from corbiculae, pollen baskets, we seek to determine whether this approach is viable, and if the foraging record of pollen matches the record derived from observations.  **bee stuff here**

# METHODS

## Study System
Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Colorado, USA (see...*APPENDIX 1* for site information). Pollinator observations of *Bombus* Latreille spp. (Apidae Latreille) were conducted from June - August of 2015. This area is characterised by high-montane/subalpine Parkland vegetation communities.

### Spatial Analyses
To develop an ecologically relevant list of vascular plant species , with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from BIEN (@bien2022), and these taxa had Species Distribution Models generated to infer their suitability. This list of species served as a reference for which species to include in the Kraken2 database.

In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our field site and all plant species records were downloaded. In order to estimate the stochasticity of collections, this dataset was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing dataset. The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.

Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in (21. Southern Rockies), downloaded (n = 23,919), (@omernik1987ecoregions). These records were split into two sets, one for generating machine learning models, and the other for GLM and GAM models. The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset. To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION- need appropriate format for journal). To achieve a larger absence dataset 1000 pseudo-absence records were generated for each taxon, each of which was greater than 10km from an occurrence record. For ML models, these pseudo-absences were reduced so that the number of presence to absence records were 1:1; to achieve this, absence records inside of 10% of the mean sample value, of the presence records, for any predictor were removed.

Species abiotic niche predictors were 26 variables at 30m resolution, for linear regression models these predictors underwent both vifstep (theta = 10, max observations = 12,500) and vifcor (theta = 0.7, max observations = 12,500), and collinear features were removed leaving 16 variables (@usdm2014) (*APPENDIX 6*).

Modelling: Random Forest and Boosted Regression Trees, were sub sampled with 30% test and two replicates each before weighted ensemble based on True Skill Statistics (tss) (@sdmPackage). Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage). The results of these models were extracted to a polygon feature derived from a minimum-spanning tree which encompasses the study area, and species from either ensemble with greater than 50% habitat suitability were considered present for further purposes (@prim1957minimum). To evaluate these results GBIF (n = 61789) records which had not been assimilated to the BIEN database yet were evaluated using logistic regression (GBIF citation).

535 species were modelled using Generalized Linear Models and Generalized Additive Models. 534 species were modelled using Random Forest and Boosted Regression Trees. *Of the species modelled % were accurately classified as being present at the field station, while % were accurately classified as being absent from the field station* TABLE XX* for more results.*

To evaluate the accuracy of the Species distribution models, additional presence records from GBIF (n = 61789), and AIM (n = 12730) were used as test and training sets (n = 74,519) (CITE AIM AND GBIF). Additional novel absence records were generated from the AIM dataset to create a dataset where each species has balanced presence and absences. 11 or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (median = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).

## Morphological Pollen identification
To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single). Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (FRASER BUCK 2007), an additional 62 voucher slides were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V. 4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).  
  In order to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes, Divisive Hierarchical Clustering techniques were used. Ten readily discernible categorical traits were collected from each specimen in the image collection. These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022). Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.). This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner. 
  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples. Cerca 0.5mm2 of pollen was placed onto a ~4mm2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrats) (EMS, Hartfield, PA). The jelly was  melted until pollen grains were homogeneously spread across the microscope slide. Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA); all samples are noted in *APPENDIX 3*.
  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used. In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted. Subsequent to the first round of sampling, non-parametric Species Richness Rarefaction curves (@vegans2022), and non-parametric Species Diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage). Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.

## Molecular Lab Work
### Reference Plant Library Generation
Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), we identified the plant taxa most frequently visited by Queens across all years. We sequenced the 12 most commonly visited taxa twice using samples from one site within the Gunnison River Drainage and one individual from another population. In addition, for any of these 12 focal species which did not have a congener pair in this filtered sample, we included a congener - or a species from a closely related genus to serve as an outgroup. We also sequenced another 15 abundant taxa commonly visited by *Bombus* workers, based on the aforementioned data set (*APPENDIX 4*).

### Plant Genomic DNA Extraction
Plant genomic DNA was isolated from ~ 1 cm2 of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included  two chloroform washes. DNA was quantified using a Nanodrop 2000 (Thermo Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer (Thermo Fisher Scientific).

### Pollen Genomic DNA Extraction
Pollen genomic DNA was extracted from corbiculae using a CTAB based protocol modified from Lahlamgiahi et al. and Guertler et al. (2014, 2014). A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added followed by vortexing to allow dissolution of corbiculae. Pollen grains were then macerated with Kontes Pellet Pestles, and the tip of these washed with 130 µL of the SDS extraction buffer, samples were then incubated for 1 hour at 30°C. This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH. 8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C. After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°c for 1 hour. Samples were then incubated overnight at 40°C. 500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.

To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to 2/3 of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm. The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes. Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL) , inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded. Pellets were dried at room temperature overnight before resuspension in Nuclease free H2O.  Extractions were assessed using a Nanodrop 2000 (Thermo Fisher Scientific) and Qubit fluorometer (Thermo Fisher Scientific). DNA extracts were then cleaned using 2:1 v./v. Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions. DNA was quantified using a Qubit fluorometer.

### Fragmentation, Library Preparation & Target Enrichment
Library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.  Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37*C, with an input of 500 ng cleaned DNA. Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed with ½ volume of SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v. ethanol washes. The exception was the herbarium specimens which were not fragmented and only end repaired, with similar library preparation of all samples. Products were analysed on 4% agarose gels, and a Qubit fluorometer.

Libraries were pooled and enriched with the Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence Capture Kit) by following the manufacturer's protocol and Brewer et al. 2019. Sequencing was performed using an Illumina mi-Seq with 150-bp end reads, (NUSeq Core, Chicago, Illinois).

## Computational Processes and Analyses. 
### Reference Library Data Processing
Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic). Contigs were generated using HybPiper using target files created by M353 (@johnson2016hybpiper, @mclay2021new).

###  Sequence Identification
A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved). These sequences were processed in the same manner as our novel sequences before being placed into the database. The Kraken2 database was built using default parameters. Kraken2 was run on sequences using default parameters (*APPENDIX 5*). 
  Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken). Results from both Kraken2 and Bracken, results were reclassified manually to identify terminal taxa. For example, when only a single species of a genus was known in the study area, but our database used a representative of another taxon in the genus, this species was coded as the result. The re-coding of sequences from another representative species for the genus to the sole RMBL representative allowed the identification of XX & % more species.

### Identification of Sequence Matching Loci
A local NCBI database was built using the same processed novel and downloaded sequences (@camacho2009blast).

### Temporal Analyses

To estimate the duration of dates in which plant species were flowering Weibull estimates of all spatially modelled taxa were developed (@belitz2020accuracy). Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included. To remove temporally irrelevant herbarium records, i.e. material collected during times which flowering is impossible at the study area due to snow cover, the SnowUS dataset (@iler2021conceptual, @tran2019cloud) from 2000-2017 was analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall. Herbarium records after the 3rd quantile for melt, and the 1st quantile for snow cover of these metrics were removed. Species with > 10 records had their weibull distributions estimated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered.

# Results

## Microscopic Pollen identification
**Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 12 character states resulted in the establishment of XX morphotypes which grains could be reliably classified into. *APPENDIX 7* **. XX Samples were counted and based on rarefaction had over % of expected morphotypes found. The relative abundance of pollen grains in each sample (max % of any species, mean % of all species, min % trace amounts detected).

## Metabarcoding Pollen identification
Kraken2 was able to identify the species richness of pollen samples ($\bar{x}$ = , min = , max = ). Bracken was able to estimate the relative abundance of pollen grains in each sample (max % of any species, $\bar{x}$ % of all species, min % trace amounts detected).

## Spatial Analyses

```{r Results for distance species search, echo = F, message = F, warning = F}
citation("caret")
```

```{r Results for SA Modelling, echo = F, message = F, warning = F}

files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
ml <- files[grep( '_ml_', files)]
ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
linear <- files[grep('_glm_', files)]
linear <- unique(sub('_glm.*$', '', linear))

# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'

rm(files, ml, linear)
```

```{r Results for SA Logistic Regression Metrics, echo = F, message = F, warning = F}

logR_dat <- read.csv('/hdd/yun master/MEE_manuscript/data/PA_records_logistic_regression.csv')[,2:4] %>% 
  filter(occurrence == 0) 
# summary(logR_dat$n)
# sum(logR_dat$n)

dat <- read.csv('/hdd/yun master/MEE_manuscript/data/evaluation_table.csv')[,2:3]
data.frame(Metric  = dat[1:5,1],
                   Value   = dat[1:5,2],
                   Metric1 = dat[6:10,1],
                   Value1  = dat[6:10,2]
                   ) %>% 
  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
    col.names = c('Metric', 'Value', 'Metric', 'Value'),
    booktabs = T) #%>% 
  #kableExtra::kable_styling(position = "float_right")
```

```{r Results for SA Logistic Regression Contingency table, echo = F, message = F, warning = F}
read.csv('/hdd/yun master/MEE_manuscript/data/Pa_conting_table.csv') %>% 
    knitr::kable(caption = "SDN evaulation contingency table",
    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
    booktabs = T) %>%
    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))

rm(logR_dat, dat)
```

The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling. A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, **95% CI 0.04-1.00  ?** ) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, where true presences at the field station. Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, we declined to perform this computational task.

**overall quality of SDMS across entire range of study **

Taken together all models throughout the entire domain of modelling had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87.

**The quality of the models themselves in predicting a large component of the alpha richness of a region**

Of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%). The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.  Of the Of the 554 vascular plants with biotic pollination syndromes, the 475 ML ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%). The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA. Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database.

**The quality of the models in predicting ecologically dominant plant taxa in a pollination network**

Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2). Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which are restricted from the initial query database,

### Temporal Analyses

The first date of modelled snow melt in the Gothic area (n = 17, mean = 137.9, median = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, mean = 299.9, median = 300, 1st quantile = 291)... 332 species with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, median = 35, max = 96) had weibull estimates calculated, an additional 56 species with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, median = 13, max = 24).

Only 58 of these 388 species ($\bar{x}$ n = 34.56897, median n = 31) were able to be compared to ground truth data.

```{r phenology estimates, echo=FALSE, fig.cap="A caption", out.width = '75%'}
knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
# these are the tenth and nintey  percent non-parametric confidence intervals. 
```

# Discussion

Although we were able to use an actually fine scale flora to determine the species present at the field site, we suspect a similar approach may be accomplished via quick floristic inventories at sites, and then utilizing a bootstrap approach akin to ours.

Although our temporal results were lackluster, we note that our study area has an incredibly brief growing period. and we suspect these temporal results would be useful in sub-tropical and tropical ecosystems.

Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria.

Bayesian framework

Future Directions:

While at the time of writing this there are limited A353 sequence data, the Plant and Fungal Trees of Life  (PAFTOL) project, which is sequencing at least a species of each genera in the plant Kingdom will produce sequence data from over 14,000 species. Given the extant publicly available genomic data, we conservatively estimate that upon completion of PAFTOL there will be no fewer than 15,500 species (4.4% of all ca. 350,000 plant species) for which sequence data of a majority of these loci exist. Accordingly, projects in the near future may increase the number of metagenomics samples while decreasing the need to create their own plant sequence reference libraries. 
  As a result of PAFTOL the first ever comprehensive phylogenetic hypotheses of all plant genera will be presented. In tandem with an increased number of digitised and geo-referenced herbarium specimens, and monitoring programs in natural areas, we believe that geo-informatics, and phylogenetic inference will increase the ability of researchers applying this technique to identifying sequence reads. 
  While our approach emphasises the use of this metagenomic technique for the purpose of identifying pollen, I argue the template and resources we provide here make this approach a suitable candidate for many plant metagenomic tasks. While we did not have the resources to explore the possibility of characterising infraspecific characteristics, preliminary results from others (Wenzell et al. in prep., Loke et al. in prep) indicate a possibility for these probes to also collect data at the level of populations and individuals. **

In regards to better understanding the foraging preferences of *Bombus* feeding in subalpine ecosystems. **JANE AND PAUL SET UP FOR NEAR FUTURE RESULTS?**

**Author Contributions:**  R.C.B conducted botanical collections, conducted all molecular lab work, lead all analyses, and writing. J.E.O conceived, designed, and conducted all ecological fieldwork, assisted with analyses, and writing. E.J.W. prepared, imaged, and collected trait data on pollen reference slides, and assisted with analysis of trait data and writing a dichotomous key. S.T. assisted with spatial analyses and writing. P.J.C assisted with ecological analyses and writing. J.B.F. conceived, and designed all lab work, analyses, assisted with writing, and secured funding for molecular work.

**Acknowledgements:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC. We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles Williams at Ray J. Davis (IDS), Ernie Nelson at Rocky Mountain (RM). We thank the original collectors of these specimens: D. Knoke, L. Brummer, J. Boyd, C. Davidson, I. Gilman, M. Kirkpatrick, S. McCauley, J. Smith, K. Taylor, & C. Williams. David Giblin for sharing relevant sections of an advanced draft of FNA V. 15. Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation. The Bureau of Land Management is thanked as many plant specimens were collected by R.C.B as a partner or contractor to the agency, Sarah Burnett and Lauren Price are thanked for sharing AIM data. Sanda and NEB are gratefully acknowledged for technical support and generously sharing samples.

# References

### 2022-11-11 script/manuscript_MEE.Rmd
unstupiddid some paths, added more citations for spatial

@@
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
+```
 
@@
 library(knitr)
-library(kableExtra)
-```
+suppressWarnings(library(kableExtra))
 
-```{r, eval = F, echo = F}
-set_here('../')
-
-citation('usdm')
-print(citation('iNEXT'), bibtex=TRUE)
+#print(citation('spdep'), bibtex=TRUE)
 ```
@@
 ## Study System
-Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Colorado, USA (see...*APPENDIX 1* for site information).
+Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Gunnison County, Colorado, USA (see...*APPENDIX 1* for site information).
 Pollinator observations of *Bombus* Latreille spp.
 (Apidae Latreille) were conducted from June - August of 2015.
-This area is characterised by high-montane/subalpine Parkland vegetation communities.
+The six study sites are in areas characterised by high-montane/subalpine Parkland vegetation communities.
 
 ### Spatial Analyses
-To develop an ecologically relevant list of vascular plant species , with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from BIEN (@bien2022), and these taxa had Species Distribution Models generated to infer their suitability.
-This list of species served as a reference for which species to include in the Kraken2 database.
+To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from BIEN (@bien2022), and these taxa had Species Distribution Models generated to infer their suitability.
+This list of species served as a reference for which species to include in the genomic sequence databases.
 
@@
 
-Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in (21.
-Southern Rockies), downloaded (n = 23,919), (@omernik1987ecoregions).
-These records were split into two sets, one for generating machine learning models, and the other for GLM and GAM models.
-The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset.
+Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in (No.
+21' Southern Rockies'), downloaded (n = 23,919), (@omernik1987ecoregions).
+These records were split into two, initially identical, sets, one for generating machine learning models, and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM).
+The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset, as measured by Morans Index (@moran1950notes, @spdep).
 To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION- need appropriate format for journal).
@@
 
-```{r Results for distance species search, echo = F, message = F, warning = F}
-citation("caret")
-```
-
 ```{r Results for SA Modelling, echo = F, message = F, warning = F}
@@
 
-logR_dat <- read.csv('/hdd/yun master/MEE_manuscript/data/PA_records_logistic_regression.csv')[,2:4] %>% 
+logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
   filter(occurrence == 0) 
@@
 
-dat <- read.csv('/hdd/yun master/MEE_manuscript/data/evaluation_table.csv')[,2:3]
+dat <- read.csv('../data/evaluation_table.csv')[,2:3]
 data.frame(Metric  = dat[1:5,1],
@@
 ```{r Results for SA Logistic Regression Contingency table, echo = F, message = F, warning = F}
-read.csv('/hdd/yun master/MEE_manuscript/data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "SDN evaulation contingency table",
+read.csv('../data/Pa_conting_table.csv') %>% 
+    knitr::kable(caption = "SDM evaluation contingency table",
     col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",

### 2022-11-11 script/manuscript_MEE.Rmd
a few more references and appendix table for predictors

@@
 
-Species abiotic niche predictors were 26 variables at 30m resolution, for linear regression models these predictors underwent both vifstep (theta = 10, max observations = 12,500) and vifcor (theta = 0.7, max observations = 12,500), and collinear features were removed leaving 16 variables (@usdm2014) (*APPENDIX 6*).
+Species abiotic niche predictors were 26 variables at 30m resolution, six related to climate,  five soil, four topographic, four related to cloud cover, with the remaining ...
+(@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
+For linear regression models these predictors underwent both vifstep (theta = 10, max observations = 12,500) and vifcor (theta = 0.7, max observations = 12,500), and collinear features were removed leaving 16 variables (@usdm2014).
 
@@
 11 or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (median = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
-
-## Morphological Pollen identification
-To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
-Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (FRASER BUCK 2007), an additional 62 voucher slides were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  In order to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes, Divisive Hierarchical Clustering techniques were used.
-Ten readily discernible categorical traits were collected from each specimen in the image collection.
-These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
-This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
-  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-Cerca 0.5mm2 of pollen was placed onto a ~4mm2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrats) (EMS, Hartfield, PA).
-The jelly was  melted until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA); all samples are noted in *APPENDIX 3*.
-  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric Species Richness Rarefaction curves (@vegans2022), and non-parametric Species Diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
 
@@
 
+## Morphological Pollen identification
+To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
+Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (FRASER BUCK 2007), an additional 62 voucher slides were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
+  In order to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes, Divisive Hierarchical Clustering techniques were used.
+Ten readily discernible categorical traits were collected from each specimen in the image collection.
+These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
+Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
+This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
+  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
+Cerca 0.5mm2 of pollen was placed onto a ~4mm2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrats) (EMS, Hartfield, PA).
+The jelly was  melted until pollen grains were homogeneously spread across the microscope slide.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA); all samples are noted in *APPENDIX 3*.
+  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
+In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
+Subsequent to the first round of sampling, non-parametric Species Richness Rarefaction curves (@vegans2022), and non-parametric Species Diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
+Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
+
 ### Temporal Analyses
 
-To estimate the duration of dates in which plant species were flowering Weibull estimates of all spatially modelled taxa were developed (@belitz2020accuracy).
+To estimate the duration of dates in which plant species were flowering Weibull estimates of all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
 Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included.
@@
 # References
+
+Layer                   Description                            Source
+----    ----------------------------------------------  ---------------------
+1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
+2.           Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
+3.           Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
+4.           Cloudiness seasonality 3 - MODIS            Wilson et al. 2016
+5.          Beginning of the frost-free period               Wang et al.
+6.             Climatic moisture deficit                     Wang et al.
+7.             Degree-days above 5C from                     Wang et al.
+8.             Mean annual precipitation                     Wang et al.
+9.          Mean annual precipitation as snow                Wang et al.
+10.             Temperature seasonality                      Wang et al.
+11.       2015 Percent Grass/Herbaceous cover - MODIS         (MOD44B)
+12.      2015 Percent Tree cover from Landsat 7/8              (GLCF)
+13.       Soil probability of bedrock (R Horizon)             SoilGrids
+14.         Soil organic carbon (Tonnes / ha)  
+15.              Surface soil pH in H2O 
+16.             Surface soil percent sand 
+17.                    Soil USDA class                        SoilGrids
+18.                Topographic elevation                    EarthEnv DEM
+19.          Topographic elevation, moving window.
+20.            Topographic percent slope                    EarthEnv DEM 
+21.            Topographic wetness index                    EarthEnv DEM 
+22.        Topographic aspect from EarthEnv DEM
+23.      Annual potential solar radiation computed             r.sun.
+24.       Estimated actual (w/-cloud) solar radiation
+25.       Log-transformed distance to surface water   Global Surface Water Explorer
+26.             Percent surface water from            Global Surface Water Explorer 
+----    ----------------------------------------------  ---------------------

### 2022-11-20 script/manuscript_MEE.Rmd
updated some text and style

--- 2022-11-11 script/manuscript_MEE.Rmd
@@
 - \usepackage{lineno}
-- \linenumbers
+- \linenumbers # cannot co-exist with two columns. 
+#classoption: 
+#- twocolumn #this makes everything take up two columns, cannot exist with line numbers
 ---
@@
 
-# INTRODUCTION
-The inability to reliably identify plants to the level of species often leaves our understanding of ecosystem function and interactions nebulous.
-Current methods to remediate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots for diagnostic material, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.
-These approaches are untenable in light of the benefits offered by species in several complex genera which serve as bioindicators, and preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts.
-Many genera have species which are well defined based upon ecological rather than morphological properties, the identification of these taxa in degraded areas or without their mutualistic partners is fraught with difficulty, hindering an understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
+# 1 | INTRODUCTION
+
+The inability to reliably identify plants to the level of species often leaves our understanding of ecosystem function and interactions wanting.
+Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.
+These approaches are untenable in light of the benefits offered by: species in several morphologically difficult genera which serve as bioindicators, preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts (CITE).
+Many genera, especially with the formalized advent of integrative taxonomy, have species which are well defined based upon ecological and behavioral rather than morphological properties, the identification of these taxa in degraded areas or without their mutualistic partners is fraught with difficulty.
+Hindering an understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
   The identification of many plant species to terminal taxon is an essential component of nearly all land management programs, where many species in the same genus  (e.g.
-Sagebrush - *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) serve as bioindicators, as well as in academic research (@Gage2013HistoricalRO, AIM).
+Sagebrush - *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) serve as bioindicators (respectively for 'rangelands', streams, and wetlands), as well as in academic research (@Gage2013HistoricalRO, AIM).
 This endeavour is often mired by lack of diagnostic characters (e.g.
@@
 However, both approaches have significant downsides, the former resource intensive at the landscape scale - and often does not work, while the latter seldom works due to a lack of variability in the currently available barcodes (@liu2014identification).
-  Recently barcoding, and metabarcoding, have shown much promise in all Kingdoms of life.
+  Recently barcoding, and metabarcoding, have shown considerable promise in all Kingdoms of life.
 For example ….
 .
-With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification), most studies seem to be in between this spectrum (CITE).
+With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification), most applications lay along this spectrum (CITE).
 Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence....
 and...
-Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling.
+Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
 
-  Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases.
-While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from Herbaria, (typically Governmental) survey work, and citizen science from a domain exceeding the study area.
-To these candidate species an approach, such as logistic regression, may be used to identify distances under which taxa are worth further exploring.
-To these candidate species, we generate species distribution models, which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results.
+  Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the spatial extent of the study region.
+While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from herbaria, (typically Government) survey work, and citizen science (e.g.
+iNaturalist), from a domain exceeding the study area.
+To these candidate species modelling approaches, such as logistic regression, may be used to identify distances under which taxa warrant further exploration.
+To these candidate species, we generate species distribution models (SDM's), which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results.
 This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
 Moreover, as most next-generation sequence data is deposited as raw-sequence reads, from a processing perspective, it is essential to reduce the candidate species via an approach as such.
-  Concomitant to next-generation sequencing has been high-throughput sequencing (cite).
-Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021exploring).
-The dataset it creates will contain Hyb-Seq data from at least one species representing each genus in the Plant Kingdom using the popular A353 probes ().
-These widely available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
+  Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
+This dataset will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021exploring), resulting in over 14,000 represented species.
+These publically available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more records upon it's intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in regions e.g.
+the 'Darwin Tree of Life' are being undertaken which will contribute data for applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
+
   *A brief consideration and example on the inclusion of the temporal dimension of plant flowering, which we suspect will be of importance in sub-tropical, and tropical ecosystems - and others with high seasonality in flowering, is also discussed.*
@@
 
-# METHODS
+# 2 | METHODS
 
-## Study System
-Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Gunnison County, Colorado, USA (see...*APPENDIX 1* for site information).
+## Study System & Field Work
+
+Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Gunnison County, Colorado, USA (*APPENDIX 1* for site information).
 Pollinator observations of *Bombus* Latreille spp.
 (Apidae Latreille) were conducted from June - August of 2015.
+Observations of *Bombus* foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and place into abundance bins.
+Corbiculae loads were,  non-lethally, collected once from all Queen individuals encountered.
 The six study sites are in areas characterised by high-montane/subalpine Parkland vegetation communities.
 
-### Spatial Analyses
-To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from BIEN (@bien2022), and these taxa had Species Distribution Models generated to infer their suitability.
-This list of species served as a reference for which species to include in the genomic sequence databases.
+### 2.1 | Spatial Analyses
+
+#### 2.1.1 Candidate Species
+
+To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from the Botanical Information and Ecology Network 'BIEN' (@bien2022), and these taxa had species distribution Dodels (SDMs) generated to predict their suitability.
+The predicted plant species served as a reference for which species to include in the genomic sequence databases.
 
 In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our field site and all plant species records were downloaded.
-In order to estimate the stochasticity of collections, this dataset was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing dataset.
-The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
+In order to emulate the perceived stochasticity of collections, this dataset was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing dataset.
+The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
 
-Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in (No.
-21' Southern Rockies'), downloaded (n = 23,919), (@omernik1987ecoregions).
-These records were split into two, initially identical, sets, one for generating machine learning models, and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM).
+#### 2.1.2 Distribution Modelling
+
+Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in *(No.
+21 "Southern Rockies")*, downloaded (n = 23,919) (@omernik1987ecoregions).
+These records were copied into two, initially identical, sets, one for generating machine learning models (Random Forest, and Boosted Regression Tree's), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM).
 The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset, as measured by Morans Index (@moran1950notes, @spdep).
-To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION- need appropriate format for journal).
+To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION - need appropriate format for journal).
 To achieve a larger absence dataset 1000 pseudo-absence records were generated for each taxon, each of which was greater than 10km from an occurrence record.
-For ML models, these pseudo-absences were reduced so that the number of presence to absence records were 1:1; to achieve this, absence records inside of 10% of the mean sample value, of the presence records, for any predictor were removed.
+For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced.
+To achieve this, absence records inside of 10% of the mean sample value of the presence records, for any predictor were removed; the required number of absence records were then randomly sampled.
 
-Species abiotic niche predictors were 26 variables at 30m resolution, for linear regression models these predictors underwent both vifstep (theta = 10, max observations = 12,500) and vifcor (theta = 0.7, max observations = 12,500), and collinear features were removed leaving 16 variables (@usdm2014) (*APPENDIX 6*).
+Species abiotic niche predictors were 26 variables at 30m resolution, six related to climate,  five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
+For linear regression models these predictors underwent both vifstep (theta = 10, max observations = 12,500) and vifcor (theta = 0.7, max observations = 12,500), and collinear features were removed leaving 16 variables (@usdm2014).
 
@@
 The results of these models were extracted to a polygon feature derived from a minimum-spanning tree which encompasses the study area, and species from either ensemble with greater than 50% habitat suitability were considered present for further purposes (@prim1957minimum).
-To evaluate these results GBIF (n = 61789) records which had not been assimilated to the BIEN database yet were evaluated using logistic regression (GBIF citation).
 
@@
 534 species were modelled using Random Forest and Boosted Regression Trees.
-*Of the species modelled % were accurately classified as being present at the field station, while % were accurately classified as being absent from the field station* TABLE XX* for more results.*
-
-To evaluate the accuracy of the Species distribution models, additional presence records from GBIF (n = 61789), and AIM (n = 12730) were used as test and training sets (n = 74,519) (CITE AIM AND GBIF).
+To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (CITE AIM AND GBIF).
 Additional novel absence records were generated from the AIM dataset to create a dataset where each species has balanced presence and absences.
@@
 
-## Morphological Pollen identification
-To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
-Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (FRASER BUCK 2007), an additional 62 voucher slides were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  In order to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes, Divisive Hierarchical Clustering techniques were used.
-Ten readily discernible categorical traits were collected from each specimen in the image collection.
-These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
-This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
-  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-Cerca 0.5mm2 of pollen was placed onto a ~4mm2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrats) (EMS, Hartfield, PA).
-The jelly was  melted until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA); all samples are noted in *APPENDIX 3*.
-  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric Species Richness Rarefaction curves (@vegans2022), and non-parametric Species Diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
+## 2.2 | Molecular Lab Work
 
-## Molecular Lab Work
-### Reference Plant Library Generation
-Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), we identified the plant taxa most frequently visited by Queens across all years.
+All lab work was carried out at The Daniel F. and Ada L. Rice Plant Conservation Science Center at the Chicago Botanic Garden, Glencoe, Illinois.
+
+#### 2.2.1 | Reference Plant Library Generation
+
+Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
 We sequenced the 12 most commonly visited taxa twice using samples from one site within the Gunnison River Drainage and one individual from another population.
@@
 
-### Plant Genomic DNA Extraction
+#### 2.2.2 | Plant Genomic DNA Extraction
+
 Plant genomic DNA was isolated from ~ 1 cm2 of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included  two chloroform washes.
@@
 
-### Pollen Genomic DNA Extraction
+#### 2.2.3 | Pollen Genomic DNA Extraction
+
 Pollen genomic DNA was extracted from corbiculae using a CTAB based protocol modified from Lahlamgiahi et al.
@@
 
-### Fragmentation, Library Preparation & Target Enrichment
+#### 2.2.4 | Fragmentation, Library Preparation & Target Enrichment
+
 Library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
@@
 Products were analysed on 4% agarose gels, and a Qubit fluorometer.
-
 Libraries were pooled and enriched with the Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence Capture Kit) by following the manufacturer's protocol and Brewer et al.
@@
 
-## Computational Processes and Analyses. 
-### Reference Library Data Processing
+###  2.2.5 | Computational Processes and Analyses. 
+####  2.2.5.1 | Reference Library Data Processing
+
 Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic).
@@
 
-###  Sequence Identification
+#### 2.2.5.2 | Sequence Identification
+
 A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
@@
 
-### Identification of Sequence Matching Loci
+#### 2.2.5.3 | Identification of Sequence Matching Loci
+
 A local NCBI database was built using the same processed novel and downloaded sequences (@camacho2009blast).
 
-### Temporal Analyses
+## 2.2.5.4 | Morphological Pollen identification
 
-To estimate the duration of dates in which plant species were flowering Weibull estimates of all spatially modelled taxa were developed (@belitz2020accuracy).
+To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
+Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
+  In order to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes, Divisive Hierarchical Clustering techniques were used.
+Ten readily discernible categorical traits were collected from each specimen in the image collection.
+These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
+Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
+This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
+  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
+*Cerca* 0.5mm^2 of pollen was placed onto a ~4mm^2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrats) (EMS, Hartfield, PA).
+The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with nail polish; all samples are noted in *APPENDIX 3*.
+  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
+In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
+Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
+Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
+
+### 2.3 | Temporal Analyses
+
+To estimate the duration of dates in which plant species were flowering weibull estimates of several phenologica lparameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
 Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included.
@@
 Herbarium records after the 3rd quantile for melt, and the 1st quantile for snow cover of these metrics were removed.
-Species with > 10 records had their weibull distributions estimated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered.
+Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered.
 
-# Results
+# 3 | RESULTS
 
-## Microscopic Pollen identification
-**Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 12 character states resulted in the establishment of XX morphotypes which grains could be reliably classified into.
-*APPENDIX 7* **.
-XX Samples were counted and based on rarefaction had over % of expected morphotypes found.
-The relative abundance of pollen grains in each sample (max % of any species, mean % of all species, min % trace amounts detected).
-
-## Metabarcoding Pollen identification
-Kraken2 was able to identify the species richness of pollen samples ($\bar{x}$ = , min = , max = ).
-Bracken was able to estimate the relative abundance of pollen grains in each sample (max % of any species, $\bar{x}$ % of all species, min % trace amounts detected).
-
-## Spatial Analyses
+## 3.1 | Spatial Analyses
 
@@
 A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, **95% CI 0.04-1.00  ?** ) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, where true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, we declined to perform this computational task.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, we declined to model these taxa.
 
-**overall quality of SDMS across entire range of study **
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87.
 
-Taken together all models throughout the entire domain of modelling had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87.
+At the field site, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.
+Of the Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.
+Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
 
-**The quality of the models themselves in predicting a large component of the alpha richness of a region**
+At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2).
+Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
 
-Of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.
-Of the Of the 554 vascular plants with biotic pollination syndromes, the 475 ML ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.
-Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database.
+## 3.2 | Microscopic Pollen identification
 
-**The quality of the models in predicting ecologically dominant plant taxa in a pollination network**
+**Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 12 character states resulted in the establishment of XX morphotypes which grains could be reliably classified into.
+*APPENDIX 7* **.
+XX Samples were counted and based on rarefaction had over % of expected morphotypes found.
+The relative abundance of pollen grains in each sample (max % of any species, mean % of all species, min % trace amounts detected).
 
-Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which are restricted from the initial query database,
+## 3.3 | Metabarcoding Pollen identification
 
-### Temporal Analyses
+Kraken2 was able to identify the species richness of pollen samples ($\bar{x}$ = , min = , max = ).
+Bracken was able to estimate the relative abundance of pollen grains in each sample (max % of any species, $\bar{x}$ % of all species, min % trace amounts detected).
+BLAST, the most thorough sequence alignment algorithm,
 
-The first date of modelled snow melt in the Gothic area (n = 17, mean = 137.9, median = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, mean = 299.9, median = 300, 1st quantile = 291)...
-332 species with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, median = 35, max = 96) had weibull estimates calculated, an additional 56 species with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, median = 13, max = 24).
+## 3.4 | Temporal Analyses
 
-Only 58 of these 388 species ($\bar{x}$ n = 34.56897, median n = 31) were able to be compared to ground truth data.
+The first date of modelled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, median = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, median = 300, 1st quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+Of the 500 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, median = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, median = 13, max = 24).
+
+Only 58 of these 388 species (n = 34.56897, median n = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.
 
@@
 
-# Discussion
+# 4 | DISCUSSION
 
-Although we were able to use an actually fine scale flora to determine the species present at the field site, we suspect a similar approach may be accomplished via quick floristic inventories at sites, and then utilizing a bootstrap approach akin to ours.
+Although we were able to use an actually fine scale flora to determine the species present at the field site, we suspect a similar approach may be accomplished via quick species richness inventories at sites, and then utilizing a bootstrap approach akin to ours, to the taxa returned from databases to derive these estimates.
 
@@
 and we suspect these temporal results would be useful in sub-tropical and tropical ecosystems.
+FURTHER, the sites used for ground truthing the temporal flowering periods were not randomly selected across the study area, and cannot be used to make inference to the population across the entirety of the study area as we did here.
+Regardless both show good agreement on flower onset, peak flowering, and moderate agreement with flowering cessation.
+The disagreement in flowering cessation is perhaps due to more microclimates which retain water, rather than microclimates which allow the early accumulation of heat.
 
 Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria.
+The actual dataset which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
+We consciously chose to do this in order to showcase the possibility of this approach working in less data rich areas.
 
@@
 Given the extant publicly available genomic data, we conservatively estimate that upon completion of PAFTOL there will be no fewer than 15,500 species (4.4% of all ca.
-350,000 plant species) for which sequence data of a majority of these loci exist.
+350,000 plant species) for which sequence data of a majority of these loci exist (@govaerts2021world).
 Accordingly, projects in the near future may increase the number of metagenomics samples while decreasing the need to create their own plant sequence reference libraries.
@@
 
-**Author Contributions:**  R.C.B conducted botanical collections, conducted all molecular lab work, lead all analyses, and writing.
+**AUTHOR CONTRIBUTIONS:**  R.C.B conducted botanical collections, conducted all molecular lab work, lead all analyses, and writing.
 J.E.O conceived, designed, and conducted all ecological fieldwork, assisted with analyses, and writing.
@@
 
-**Acknowledgements:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
+**ACKNOWLEDGEMENTS:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
+Pat Herendeen for assistance with virtually all aspects of preparing pollen vouchers and the identification process.
 We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles Williams at Ray J.
@@
 
+**CONFLICT OF INTERESTS** The authors declare no conflicts of interest.
+
+**PEER REVIEW** The peer review history for this document is available at ...
+
+**DATA AVAILABILITY STATEMENT** The queries required to download all data used in this project are located in... All novel sequencing data are located at NCBI...
+
+**ORCID**
+
 # References
+
+# Supporting

### 2022-11-20 script/predictors_tables.r

@@ -1,0 +1,30 @@
+
+Layer                   Description                            Source
+----    ----------------------------------------------  ---------------------
+  1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
+2.           Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
+3.           Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
+4.           Cloudiness seasonality 3 - MODIS            Wilson et al. 2016
+5.          Beginning of the frost-free period               Wang et al.
+6.             Climatic moisture deficit                     Wang et al.
+7.             Degree-days above 5C from                     Wang et al.
+8.             Mean annual precipitation                     Wang et al.
+9.          Mean annual precipitation as snow                Wang et al.
+10.             Temperature seasonality                      Wang et al.
+11.       2015 Percent Grass/Herbaceous cover - MODIS         (MOD44B)
+12.      2015 Percent Tree cover from Landsat 7/8              (GLCF)
+13.       Soil probability of bedrock (R Horizon)             SoilGrids
+14.         Soil organic carbon (Tonnes / ha)                 SoilGrids
+15.              Surface soil pH in H2O                       SoilGrids
+16.             Surface soil percent sand                     SoilGrids
+17.                    Soil USDA class                        SoilGrids
+18.                Topographic elevation                    EarthEnv DEM
+19.          Topographic elevation, moving window           EarthEnv DEM
+20.            Topographic percent slope                    EarthEnv DEM 
+21.            Topographic wetness index                    EarthEnv DEM 
+22.        Topographic aspect from EarthEnv DEM
+23.      Annual potential solar radiation computed             r.sun.
+24.       Estimated actual (w/-cloud) solar radiation
+25.       Log-transformed distance to surface water   Global Surface Water Explorer
+26.             Percent surface water from            Global Surface Water Explorer 
+----    ----------------------------------------------  ---------------------

### 2022-11-20 script/manuscript_MEE.Rmd
updated some writing

@@
   Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
-author: |
+author: "|"
   | Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E.
@@
 abstract: |
-  an abstract will be written to fill this space.
+  an abstract will be written to fill this space. it will have from 4-6 numbered sections.
 output:
@@
   toc: no
-csl: ../citations/methods-in-ecology-and-evolution.csl
+  word_document: default
+csl: "../citations/methods-in-ecology-and-evolution.csl"
 bibliography: ../citations/citations.bib
@@
 - \usepackage{lineno}
-- \linenumbers # cannot co-exist with two columns. 
-#classoption: 
-#- twocolumn #this makes everything take up two columns, cannot exist with line numbers
+- \linenumbers
 ---
+
 ```{r setup, include=FALSE}
@@
 Hindering an understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
+
   The identification of many plant species to terminal taxon is an essential component of nearly all land management programs, where many species in the same genus  (e.g.
@@
 However, both approaches have significant downsides, the former resource intensive at the landscape scale - and often does not work, while the latter seldom works due to a lack of variability in the currently available barcodes (@liu2014identification).
+
   Recently barcoding, and metabarcoding, have shown considerable promise in all Kingdoms of life.
@@
 Moreover, as most next-generation sequence data is deposited as raw-sequence reads, from a processing perspective, it is essential to reduce the candidate species via an approach as such.
+
   Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
@@
 
-  *A brief consideration and example on the inclusion of the temporal dimension of plant flowering, which we suspect will be of importance in sub-tropical, and tropical ecosystems - and others with high seasonality in flowering, is also discussed.*
-  We show that by combining both candidate plant species, and representative sequence data, with only a small amount of new reference sequencing, one is capable of generating ecologically realistic, and biologically relevant results.
-To test these assertions we utilize a long-term observational study of Bumble Bees, the most important genus of pollinators in the temperate, and one with consistently concerning conservation assessments (@cameron2020global, @goulson2008decline).
-By generating sequence data from corbiculae, pollen baskets, we seek to determine whether this approach is viable, and if the foraging record of pollen matches the record derived from observations.
-**bee stuff here**
+  Considerable amounts of species interactions are expressed along time (@caradonna2021seeing).
+For the tropics the flowering periods of many plant species display high seasonality, and given the elevated rates of species richness relative to the temperate, this axis may provide an essential filter for identifying material in many metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
+While many pollination interactions are formed and dissolved along the temporal axis in the temperate regions (@caradonna2017interaction), the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which subjugates the utility of these to perform as filters of post-processing results, rather than distinct species assemblage for database generation.
+Nonetheless, we work through a process which seems applicable to the tropics to utilize the temporal dimension for classifying sequencing results.
+
+  We apply these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
+Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient).
+These montane ecosystems represent some of the most ecologically resilient and resistant systems in the temperate and offer unparalleled potential as refugial areas for multiple dimensions of biodiversity under climate change.
 

### 2022-11-20 script/predictors_tables.r

@@ -1,7 +1,7 @@
 
 Layer                   Description                            Source
 ----    ----------------------------------------------  ---------------------
-  1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
+1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
 2.           Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
 3.           Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
 4.           Cloudiness seasonality 3 - MODIS            Wilson et al. 2016

### 2022-11-24 script/manuscript_MEE.Rmd
just making sure we have the latest version up

@@
   Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
-author: "|"
-  | Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E.
+author: |
+  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E.
 Ogilvie $^3$, Emily J.
@@
 output:
+  word_document: default
+  toc: no
   pdf_document: default
-  toc: no
-  word_document: default
 csl: "../citations/methods-in-ecology-and-evolution.csl"
@@
 fig_caption: yes
+always_allow_html: true
 header-includes:

### 2022-11-27 script/manuscript_MEE.Rmd
some numbers to add

@@
 
+XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of XX corbiculae samples were sequenced, XX reads were recovered from sequencing.
+After trimming XX seqs remained.
+XX reads were matched using Kraken, of the reads classified by Kraken XX reads were matched using Bracken, of the reads classified by Kraken XX reads were matched using BLAST.
+The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely ().
+
 Kraken2 was able to identify the species richness of pollen samples ($\bar{x}$ = , min = , max = ).

### 2022-11-27 script/manuscript_MEE.Rmd
stuff

--- 2022-11-24 script/manuscript_MEE.Rmd
@@
 output:
+  pdf_document: default
+  toc: no
   word_document: default
-  toc: no
-  pdf_document: default
 csl: "../citations/methods-in-ecology-and-evolution.csl"
@@
 Additional novel absence records were generated from the AIM dataset to create a dataset where each species has balanced presence and absences.
-11 or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (median = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
+11 or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
 
@@
 
+### 2.4 | Floral Observations
+
 # 3 | RESULTS
@@
 
-The first date of modelled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, median = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, median = 300, 1st quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 500 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, median = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, median = 13, max = 24).
+The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1st quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+Of the 500 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, Mdn = 13, max = 24).
 
-Only 58 of these 388 species (n = 34.56897, median n = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.
+Only 58 of these 388 species (n = 34.56897, Mdn n = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.
 
-```{r phenology estimates, echo=FALSE, fig.cap="A caption", out.width = '75%'}
+```{r phenology estimates, echo=F, fig.cap="A caption", out.width = '75%'}
 knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
 # these are the tenth and nintey  percent non-parametric confidence intervals. 
+```
+
+## 3.5 | Floral Observations ?
+
+```{r Floral Obs Results, echo = F, warning = F}
+
+files <- list.files('../data', pattern = 'csv')
+
+flr_ranks <- read.csv(
+  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
+  select(site, doy, week, tot.obs.length, habitat, plant.species:abun.rank)
+obs_time <- read.csv(
+  file.path('../data', '/', files[grep('queen_observations', files)]) ) %>% 
+  select(site, week, tot.obs.length) %>% 
+  distinct(site, week, .keep_all = T)
+
+bee_obs <- read.csv(
+  file.path('../data', files[grep('queen_observations', files)]) ) %>% 
+  filter(caste == 'q') %>% 
+  mutate(species = gsub('\\..*$', '', species)) %>% 
+  select(site, doy, week, tot.obs.length, species, plant.species,
+         resource.coll:pollen.collected, fl.switch.1:fl.switch.2.resource.coll) %>% 
+  mutate(across(resource.coll:fl.switch.2.resource.coll, ~ na_if(.x, ''))) %>% 
+  pivot_longer(c(plant.species, fl.switch.1, fl.switch.2), names_to = 'observation', values_to = 'plant.species') %>% 
+  drop_na(plant.species) %>% 
+  group_by(site, week, species, plant.species) %>% 
+  count(name = 'interactions') %>% 
+  left_join(., obs_time) %>% 
+  mutate(period = case_when(
+    week %in% 3:5 ~ 'Early',
+    week %in% 6:8 ~ 'Mid',
+    week %in% 9:11 ~ 'Late'
+  ))
+
+s <- toString(format(as.Date(min(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # start obs
+e <- toString(format(as.Date(max(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # end obs
+
+t_b_wk <- bee_obs %>%  group_by(site, week, species) %>%  tally(interactions) 
+t_b_b <- bee_obs %>%  group_by(species) %>%  tally(interactions) %>%  filter(species != 'unknown') 
+t_b_wk_pl <- bee_obs %>%  group_by(site, week, plant.species) %>%  tally(interactions) 
+t_b_pl <- bee_obs %>%  group_by(plant.species) %>%  filter(species != 'unknown') %>%  tally(interactions) 
+t_spp_i <- bee_obs %>%  group_by(species) %>%  filter(species != 'unknown') %>%  distinct(plant.species) %>% 
+  count()
+
+```
+
+The six sites were surveyed for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours from `r s`-`r e`.
+A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week min = `r toString(min(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`, max = `r toString(max(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
+Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week min = `r toString(min(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, Mdn = `r toString(median(t_b_wk_pl$n))`, max = `r toString(max(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (min = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, Mdn = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
+The number of plant species which bees were observed interacting with varied more narrowly (min = `r toString(min(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`, max = `r toString(max(t_spp_i$n))`).
+
+```{r echo = F}
+rm(bee_obs, flr_ranks, obs_time, t_b_b, t_b_wk, t_b_pl, t_b_wk_pl, e, s, files, t_spp_i)
 ```

### 2022-11-28 script/manuscript_MEE.Rmd
too tired

@@
 Bracken was able to estimate the relative abundance of pollen grains in each sample (max % of any species, $\bar{x}$ % of all species, min % trace amounts detected).
-BLAST, the most thorough sequence alignment algorithm,
+BLAST, the most thorough sequence alignment algorithm...
 
@@
 
-## 3.5 | Floral Observations ?
+## 3.5 | Floral Observations
 
@@
 rm(bee_obs, flr_ranks, obs_time, t_b_b, t_b_wk, t_b_pl, t_b_wk_pl, e, s, files, t_spp_i)
+```
+
+```{r Most used resources by species, echo = F}
+
+corb <- read.csv('../data/Corbiculae_collections.csv', header = F) %>% 
+  mutate(sample.id = as.numeric(str_extract(V1, '[0-9]{1,2}')),
+         species = str_extract(V1, 'Bombus [^ ]+'),
+         caste = str_trim(str_extract(V1, '[:space:][q|w][:space:]'))) %>% 
+  filter(caste == 'q') %>% 
+  select(-V1, -caste)
+
+blst <- read.csv('../data/classified_blast.csv') %>% 
+  distinct(sample.id)
+
+blst <- read.csv('../data/classified_blast.csv') %>% 
+  mutate(across(where(is.character), ~ replace_na(., 'UNKNOWN'))) %>% 
+  arrange(sample.id, -Prcnt_seqs) %>% 
+  inner_join(., corb, by = 'sample.id')
+
+blst_cnt <- blst %>% 
+  distinct(sample.id, .keep_all = T) %>% 
+  count(species)
+
+tb <- blst %>% 
+  left_join(., blst_cnt, by = 'species') %>% 
+  group_by(species, name) %>% 
+  mutate(Mu_Seqs = sum(Prcnt_seqs)/n)  %>% 
+  distinct(species, name, .keep_all = T)
+
+tb %>% 
+  filter(species == 'Bombus appositus') %>% 
+  ungroup() %>% 
+  mutate(total = sum(Mu_Seqs))
+
+obs <- read.csv('../data/yearEndInteractionsObservations.csv')[,2:5] %>% 
+  group_by(species) %>% 
+  arrange(species, -bySpp) %>% 
+  slice_head(n = 10)
+
 ```

### 2022-11-28 script/manuscript_MEE.Rmd
added an appendix image

@@
 
-To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from the Botanical Information and Ecology Network 'BIEN' (@bien2022), and these taxa had species distribution Dodels (SDMs) generated to predict their suitability.
+To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from the Botanical Information and Ecology Network 'BIEN' (@bien2022), and these taxa had Species Distribution Models (SDMs) generated to predict their suitability.
 The predicted plant species served as a reference for which species to include in the genomic sequence databases.
@@
 XX reads were matched using Kraken, of the reads classified by Kraken XX reads were matched using Bracken, of the reads classified by Kraken XX reads were matched using BLAST.
-The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely ().
+The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.
 
@@
 # Supporting
+
+```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING}
+
+knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
+
+```

### 2022-11-29 script/manuscript_MEE.Rmd
some figures added but these missized ;

@@
 
-**Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 12 character states resulted in the establishment of XX morphotypes which grains could be reliably classified into.
-*APPENDIX 7* **.
-XX Samples were counted and based on rarefaction had over % of expected morphotypes found.
+Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be reliably classified into.
+*APPENDIX 7*.
+60 Samples were counted and based on rarefaction **had over % of expected morphotypes found**, all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
 The relative abundance of pollen grains in each sample (max % of any species, mean % of all species, min % trace amounts detected).
@@
 
-```{r Most used resources by species, echo = F}
+```{r Most used resources by species, eval = F}
 
@@
 
-```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING}
+```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F, fig.height=2.7, fig.width=9}
+knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
+```
 
-knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
+```{r APPENDIX NUMBER RMBL Pollen Key, echo = F, fig.height=2.7, fig.width=9}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key.pdf')
+```
 
+```{r APPENDIX MORPHOTYPE RICHNESS AND ABUNDANCE, echo= F, fig.height=2.7, fig.width=9}
+knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
+knitr::include_graphics('../graphics/plots/SppAbundance.png')
 ```

### 2022-11-30 script/manuscript_MEE.Rmd
have pollen morpho data brought in now

@@
 *APPENDIX 7*.
-60 Samples were counted and based on rarefaction **had over % of expected morphotypes found**, all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
-The relative abundance of pollen grains in each sample (max % of any species, mean % of all species, min % trace amounts detected).
+60 Samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, mean = 4.5, median = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
+The number of counted pollen grains in each sample range from (mean = 2788.685, median = 1453, max = 16293).
 

### 2022-12-04 script/manuscript_MEE.Rmd
more writing in intro, more metabarcoding content added

@@
 Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.
-These approaches are untenable in light of the benefits offered by: species in several morphologically difficult genera which serve as bioindicators, preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts (CITE).
+These approaches are untenable in light of the benefits offered by: species in several morphologically difficult genera which serve as bioindicators, preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts (@hebert2003biological).
 Many genera, especially with the formalized advent of integrative taxonomy, have species which are well defined based upon ecological and behavioral rather than morphological properties, the identification of these taxa in degraded areas or without their mutualistic partners is fraught with difficulty.
@@
 This endeavour is often mired by lack of diagnostic characters (e.g.
-flowers, fruits, or roots),  and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
+flowers, fruits, roots or combinations thereof),  and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
 Solutions to this problem are wanting, certain programmes have relied increasingly upon revisiting field sites to identify material using morphological or chemical approaches, whereas academic research has often used high copy number plastid genes as barcodes (Rosentreter et al.
@@
 .
-With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification), most applications lay along this spectrum (CITE).
-Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence....
-and...
-Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
+With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification, @china2011comparative), with most applications laying along this spectrum (@li2015plant, @kress2007two, @cbol2009dna).
+Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant) Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
 
@@
 This dataset will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021exploring), resulting in over 14,000 represented species.
-These publically available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
+These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
 Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more records upon it's intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in regions e.g.
@@
 ## 3.2 | Microscopic Pollen identification
+```{r, echo = F}
+a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
+b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia')
+c <- c('Jeffersonia', 'Micranthes', 'Prunus', 'Delphinium', 'Androsace', 
+       'Penstemon', 'Orthocarpus', 'Scuttelaria', 'Aquilegia', 'Castilleja', 'Draba')
+d <- c('Salix', 'Boechera')
+
+# Do we detect mertensia, asteraceae 1, asteraceae 2, in all molecular samples which they are present in the pollen samples of?
+
+```
 
@@
 
-Only 58 of these 388 species (n = 34.56897, Mdn n = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.
+Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.
 
@@
   count(name = 'interactions') %>% 
-  left_join(., obs_time) %>% 
+  left_join(., obs_time, by = c("site", "week")) %>% 
   mutate(period = case_when(
@@
 
-```{r Most used resources by species, eval = F}
+```{r Most used resources by species, eval = F, echo =F}
 

### 2022-12-04 script/manuscript_MEE.Rmd
more stat values

@@
 
-Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic).
+Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950).
 Contigs were generated using HybPiper using target files created by M353 (@johnson2016hybpiper, @mclay2021new).
@@
 
-Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be reliably classified into.
+Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into.
 *APPENDIX 7*.
-60 Samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, mean = 4.5, median = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
-The number of counted pollen grains in each sample range from (mean = 2788.685, median = 1453, max = 16293).
+60 samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, mean = 4.5, median = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
+The number of counted pollen grains in each sample range from ($\bar{x}$ = 2788.685, median = 1453, max = 16,293).
 
@@
 
-XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of XX corbiculae samples were sequenced, XX reads were recovered from sequencing.
-After trimming XX seqs remained.
-XX reads were matched using Kraken, of the reads classified by Kraken XX reads were matched using Bracken, of the reads classified by Kraken XX reads were matched using BLAST.
+```{r eval = F}
+toseq <- read.csv('../data/Benkendorf_combos_fixed.csv') |> 
+  filter(!str_detect(extract_id, '[aA-zZ]')) %>% nrow()
+
+krak <- read.csv('../data/kraken_proc_reads.csv')
+brack <- read.csv('../data/bracken_proc_reads.csv')
+blast <- read.csv('../data/blast_proc_reads.csv')
+```
+
+XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(toseq)` corbiculae samples were sequenced, XX reads were recovered from sequencing.
+After trimming XX sequences remained.
+`r toString(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r toString(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r toString(sum(blast$Total_seqs))` reads were matched using BLAST.
 The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.
@@
 
+```{r, eval = F}
+rm(toseq)
+```
+
 ## 3.4 | Temporal Analyses
@@
 
-Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.
+Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from long term (1974–2012) datasets.
 
@@
 The number of plant species which bees were observed interacting with varied more narrowly (min = `r toString(min(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`, max = `r toString(max(t_spp_i$n))`).
+
+A total of 66 corbiculae loads were collected from Bees, 64 of them from Queens.
 
@@
   While our approach emphasises the use of this metagenomic technique for the purpose of identifying pollen, I argue the template and resources we provide here make this approach a suitable candidate for many plant metagenomic tasks.
-While we did not have the resources to explore the possibility of characterising infraspecific characteristics, preliminary results from others (Wenzell et al.
-in prep., Loke et al.
+While we did not have the resources to explore the possibility of characterising infraspecific characteristics, preliminary results from others (@wenzell2021incomplete, Loke et al.
 in prep) indicate a possibility for these probes to also collect data at the level of populations and individuals.
@@
 
-```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F, fig.height=2.7, fig.width=9}
+```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F}
 knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
@@
 
+Layer                   Description                            Source
+----    ----------------------------------------------  ---------------------
+1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
+2.           Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
+3.           Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
+4.           Cloudiness seasonality 3 - MODIS            Wilson et al. 2016
+5.          Beginning of the frost-free period               Wang et al.
+6.             Climatic moisture deficit                     Wang et al.
+7.             Degree-days above 5C from                     Wang et al.
+8.             Mean annual precipitation                     Wang et al.
+9.          Mean annual precipitation as snow                Wang et al.
+10.             Temperature seasonality                      Wang et al.
+11.       2015 Percent Grass/Herbaceous cover - MODIS         (MOD44B)
+12.      2015 Percent Tree cover from Landsat 7/8              (GLCF)
+13.       Soil probability of bedrock (R Horizon)             SoilGrids
+14.         Soil organic carbon (Tonnes / ha)                 SoilGrids
+15.              Surface soil pH in H2O                       SoilGrids
+16.             Surface soil percent sand                     SoilGrids
+17.                   Soil USDA class                         SoilGrids
+18.                Topographic elevation                    EarthEnv DEM
+19.          Topographic elevation, moving window.          EarthEnv DEM
+20.            Topographic percent slope                    EarthEnv DEM 
+21.            Topographic wetness index                    EarthEnv DEM 
+22.             Topographic aspect from                     EarthEnv DEM
+23.      Annual potential solar radiation computed             r.sun
+24.       Estimated actual (w/-cloud) solar radiation  r.sun / Wilson et al. 2016
+25.       Log-transformed distance to surface water   Global Surface Water Explorer
+26.             Percent surface water                 Global Surface Water Explorer 
+----    ----------------------------------------------  ---------------------
+
 ```{r APPENDIX MORPHOTYPE RICHNESS AND ABUNDANCE, echo= F, fig.height=2.7, fig.width=9}

### 2022-12-04 script/manuscript_MEE.Rmd
some writing, need to review how many seqs are classified, it is in excess of raw

@@
 abstract: |
-  an abstract will be written to fill this space. it will have from 4-6 numbered sections.
+  an abstract will be written to fill this space.
+it will have from 4-6 numbered sections.
+like 1) A very broad overall identifying organisms is hard and we dont have time.
+2) to try and figure this out we used whoa molecular stuff and holy shit amounts of computing 3) Our results showed we were able to do some cool things 4) you should try this out, because it was pretty cool and helped us perform natural history on steroids.
 output:
@@
 .
-With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification, @china2011comparative), with most applications laying along this spectrum (@li2015plant, @kress2007two, @cbol2009dna).
+With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification, @china2011comparative), while most applications laying along this spectrum (@li2015plant, @kress2007two, @cbol2009dna).
 Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant) Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
@@
   Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the spatial extent of the study region.
-While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from herbaria, (typically Government) survey work, and citizen science (e.g.
+While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
 iNaturalist), from a domain exceeding the study area.
-To these candidate species modelling approaches, such as logistic regression, may be used to identify distances under which taxa warrant further exploration.
+To these candidate species, modelling approaches - such as logistic regression, may be used to identify distances under which taxa warrant further exploration.
 To these candidate species, we generate species distribution models (SDM's), which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results.
@@
 
+```{sh Get number of raw reads for metagenome samples, eval = F}
+
+cd ~/sequenceData/Original_zip
+files=$(find -type f -name "[0-9][0-9]_S[0-9]*_paired.fastq.gz")
+
+# seqs per each sample , need divide by 4 on import
+wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
+```
+
 ```{r eval = F}
-toseq <- read.csv('../data/Benkendorf_combos_fixed.csv') |> 
-  filter(!str_detect(extract_id, '[aA-zZ]')) %>% nrow()
+
+raw_seqs <- read.delim('../data/raw_reads_metagenome_seq.txt', header = F) %>% 
+  mutate(V1 = str_trim(V1)) %>% 
+  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
+  mutate(seq = round(as.numeric(seq)/4),
+         sample = str_extract(sampleFR, '.*L002_')) %>% 
+  filter(sampleFR != 'total') %>% 
+  group_by(sample) %>% 
+  mutate(totalReads = sum(seq)) %>% 
+  distinct(totalReads) %>% 
+  pull(totalReads)
 
@@
 
-XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(toseq)` corbiculae samples were sequenced, XX reads were recovered from sequencing.
+54 corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r toString(sum(raw_seqs))` reads were recovered from sequencing.
+The number of reads per sequence varied widely (min = `r toString(min(raw_seqs))`, $\bar{x}$ = `r toString(mean(raw_seqs))`, Mdn = `r toString(median(raw_seqs))`, max = `r toString(max(raw_seqs))`), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**..
 After trimming XX sequences remained.
 `r toString(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r toString(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r toString(sum(blast$Total_seqs))` reads were matched using BLAST.
-The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.
 
-Kraken2 was able to identify the species richness of pollen samples ($\bar{x}$ = , min = , max = ).
-Bracken was able to estimate the relative abundance of pollen grains in each sample (max % of any species, $\bar{x}$ % of all species, min % trace amounts detected).
-BLAST, the most thorough sequence alignment algorithm...
+Based upon subjective review of the three classifiers, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all analyses.
 
 ```{r, eval = F}
-rm(toseq)
+rm(toseq, krak, brack, blast)
 ```
@@
 
+Paul CaraDonna https://orcid.org/0000-0003-3517-9090
+Jeremie Fant https://orcid.org/0000-0001-9276-1111
+Jane Ogilvie https://orcid.org/0000-0001-8546-0417
+Sophie Taddeo https://orcid.org/0000-0002-7789-1417
+
 # References

### 2022-12-06 script/manuscript_MEE.Rmd
formatting some abstract ideas

@@
 ---
-title: Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
-  Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
+title: Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
+  Patterns of Queen Bumble Bees
 author: |
@@
 abstract: |
-  an abstract will be written to fill this space.
-it will have from 4-6 numbered sections.
-like 1) A very broad overall identifying organisms is hard and we dont have time.
-2) to try and figure this out we used whoa molecular stuff and holy shit amounts of computing 3) Our results showed we were able to do some cool things 4) you should try this out, because it was pretty cool and helped us perform natural history on steroids.
+  .    
+  1) DNA Barcoding has been remarkably successful in nearly all kingdoms of life and has allowed for the rapid analysis of ecological assemblies.
+Successful DNA barcoding in the plant kingdom has been more difficult than other kingdoms.
+Due to this understanding plants in ecological contexts and understandings of their synecology in some instances may begin to lag behind other kingdoms.
+  2) Here we utilize hyb-seq, museum studies, and species distribution modelling, to detect the plant species present in pollen loads collected from Queen Bumble Bees.  
+  3) We show that hyb-seq using the Angiosperms 353 probes offers significant promise to metagenomic approaches in real world scenarios.  
+  4) We conclude that these probes offer promise for the identification of plant tissue in both single sample, and metasample contexts.
+
+  **KEYWORDS**    
+   DNA barcodes, Plantae, metagenomics, molecular barcoding, species distribution modelling, Angiosperms353, bombus, queen foraging.
 output:
@@
 suppressWarnings(library(kableExtra))
-
+pNumC <- function(...) prettyNum(...,big.mark = ',')
 #print(citation('spdep'), bibtex=TRUE)
@@
 
-All lab work was carried out at The Daniel F. and Ada L. Rice Plant Conservation Science Center at the Chicago Botanic Garden, Glencoe, Illinois.
+All lab work was carried out at The Daniel F. and Ada L. Rice Plant Conservation Science Center at the Chicago Botanic Garden, Glencoe, Illinois, U.S.A.
 
@@
 
-Plant genomic DNA was isolated from ~ 1 cm2 of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included  two chloroform washes.
+Plant genomic DNA was isolated from ~ 1 cm^2^ of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included two chloroform washes.
 DNA was quantified using a Nanodrop 2000 (Thermo Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer (Thermo Fisher Scientific).
@@
 This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH.
-8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
+8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°^C^, on heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
 After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°c for 1 hour.
-Samples were then incubated overnight at 40°C.
+Samples were then incubated overnight at 40°^C^.
 500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.
@@
 The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, **95% CI 0.04-1.00  ?** ) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, where true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, we declined to model these taxa.
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, where true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences these taxa were not modelled.
 
@@
 At the field site, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.
+The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
 Of the Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754), a P VALUE IS NOT REPORTED AS THE VALUES WERE MANUALLY PARSED INTO CLASSES BASED ON SUITABILITY PER UNIT AREA.
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
 Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
 
-At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2).
+At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
 Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
@@
 *APPENDIX 7*.
-60 samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, mean = 4.5, median = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
-The number of counted pollen grains in each sample range from ($\bar{x}$ = 2788.685, median = 1453, max = 16,293).
+60 samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
+The number of counted pollen grains in each sample range from (*MIN* - 16,293, $\bar{x}$ = 2788.685, Mdn = 1453).
 
@@
 
-```{sh Get number of raw reads for metagenome samples, eval = F}
+```{sh Get number of raw and trimmed reads for metagenome samples, eval = F, echo = F}
 
 cd ~/sequenceData/Original_zip
-files=$(find -type f -name "[0-9][0-9]_S[0-9]*_paired.fastq.gz")
-
+files=$(find -type f -name "[0-9]_S[0-9]*_paired.fastq.gz") # doesnt support {2}
+files1=$(find -type f -name '[0-9][0-9]_S[0-9]*_paired.fastq.gz')
+files="${files} ${files1}"
+echo $files
 # seqs per each sample , need divide by 4 on import
 wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
+
+cd ../trimmed_metagenomic
+files=$(find  -maxdepth 1 -name '*gz')
+wc -l  $files  > '/hdd/yun master/MEE_manuscript/data/trimmed_reads_metagenome_seq.txt'
 ```
 
-```{r eval = F}
+```{r sequence numbers, echo = F}
 
@@
 
+trimmed_seqs <- read.delim('../data/trimmed_reads_metagenome_seq.txt', header = F) %>% 
+  mutate(V1 = str_trim(V1)) %>% 
+  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
+  mutate(seq = round(as.numeric(seq)/4),
+         sample = str_extract(sampleFR, '.*paired')) %>% 
+  filter(sampleFR != 'total') %>% 
+  group_by(sample) %>% 
+  mutate(totalReads = sum(seq)) %>% 
+  distinct(totalReads) %>% 
+  pull(totalReads)
+
 krak <- read.csv('../data/kraken_proc_reads.csv')
@@
 blast <- read.csv('../data/blast_proc_reads.csv')
+loci <- read.csv('../data/reads_per_loci.csv') %>% group_by(Sample)
+
+no_loci <- loci %>% group_by(Sample) %>% count(name = 'no_loci')
+no_reads <- loci %>% group_by(loci) %>% summarize(readsLoci = sum(n))
+
+rm(prettyNum)
 ```
 
-54 corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r toString(sum(raw_seqs))` reads were recovered from sequencing.
-The number of reads per sequence varied widely (min = `r toString(min(raw_seqs))`, $\bar{x}$ = `r toString(mean(raw_seqs))`, Mdn = `r toString(median(raw_seqs))`, max = `r toString(max(raw_seqs))`), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**..
-After trimming XX sequences remained.
-`r toString(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r toString(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r toString(sum(blast$Total_seqs))` reads were matched using BLAST.
+54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
+The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
+Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
+The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`.
 
-Based upon subjective review of the three classifiers, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all analyses.
+**APPENDIX X**.
 
-```{r, eval = F}
-rm(toseq, krak, brack, blast)
+After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
+`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
+
+Based upon subjective review of the three classifiers, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all subsequent analyses.
+
+```{r, eval = F, echo = F}
+rm(toseq, krak, brack, blast, loci, no_loci)
 ```
@@
 The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1st quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 500 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, Mdn = 13, max = 24).
+Of the **500** species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, Mdn = 13, max = 24).
 
-Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from long term (1974–2012) datasets.
+Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from the long term (1974–2012) dataset.
+Of these species relatively high accord was observed between the long-term ground truthed dataset, and the modelled species.
 
@@
 The six sites were surveyed for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours from `r s`-`r e`.
-A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week min = `r toString(min(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`, max = `r toString(max(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
-Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week min = `r toString(min(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, Mdn = `r toString(median(t_b_wk_pl$n))`, max = `r toString(max(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (min = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, Mdn = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
-The number of plant species which bees were observed interacting with varied more narrowly (min = `r toString(min(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`, max = `r toString(max(t_spp_i$n))`).
+A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))` - `r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
+Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week min = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, Mdn = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (min = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, Mdn = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
+The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`).
 
@@
 
-```{r APPENDIX NUMBER RMBL Pollen Key, echo = F, fig.height=2.7, fig.width=9}
+```{r APPENDIX NUMBER RMBL Pollen Key, echo = F}
 knitr::include_graphics('../graphics/assorted/RMBL_pollen_key.pdf')
@@
 
-```{r APPENDIX MORPHOTYPE RICHNESS AND ABUNDANCE, echo= F, fig.height=2.7, fig.width=9}
+```{r APPENDIX MORPHOTYPE RICHNESS AND ABUNDANCE, echo= F}
 knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')

### 2022-12-07 script/manuscript_MEE.Rmd
stylinadded results for temporal, need to checkin with paul and jane to have them clean it up and probably drop the tau for a simpler test, splitting appendixes drop em in

@@
 
-  **KEYWORDS**    
-   DNA barcodes, Plantae, metagenomics, molecular barcoding, species distribution modelling, Angiosperms353, bombus, queen foraging.
 output:
@@
 This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH.
-8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°^C^, on heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
-After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°c for 1 hour.
-Samples were then incubated overnight at 40°^C^.
+8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
+After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°C for 1 hour.
+Samples were then incubated overnight at 40°C.
 500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.
@@
   To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Cerca* 0.5mm^2 of pollen was placed onto a ~4mm^2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrats) (EMS, Hartfield, PA).
+*Cerca* 0.5mm^2 of pollen was placed onto a ~4mm^2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
 The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
@@
 
-To estimate the duration of dates in which plant species were flowering weibull estimates of several phenologica lparameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
+To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
 Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included.
@@
 The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
 The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
@@
 no_reads <- loci %>% group_by(loci) %>% summarize(readsLoci = sum(n))
-
-rm(prettyNum)
 ```
@@
 
-Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from the long term (1974–2012) dataset.
+Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) dataset.
 Of these species relatively high accord was observed between the long-term ground truthed dataset, and the modelled species.
+There was very strong evidence that the weibull estimates were positively associated with the observed onset (r^2^ = 0.72, p < 0.0001, tau = 0.61) and peak (r^2^ = 0.70, p < 0.0001, tau = 0.65) of flowering, and that the number of herbarium samples had a moderate effect on the estimates (p = 0.004 and p = 0.034 respectively).
+There was very strong evidence that the weibull estimates had a positive association with the observed cessation of flowering (r^2^ =  0.4339,  p < 0.0001, tau = 0.489), however their was no evidence that sample size had an effect (p = 0.349).
+There was moderate evidence that the weibull estimates, with an effect of sample size, had a weak positive association with the observed duration of flowering (p = 0.0401, r^2^ = 0.07, tau = 0.17).
 
@@
 
+** Mention in here that the general effect of sample size on predictions means that larger samples sizes are required for this to work!!!! **
 Although our temporal results were lackluster, we note that our study area has an incredibly brief growing period.
@@
 
-**ACKNOWLEDGEMENTS:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
+**ACKNOWLEDGMENTS:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
 Pat Herendeen for assistance with virtually all aspects of preparing pollen vouchers and the identification process.
-We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles Williams at Ray J.
-Davis (IDS), Ernie Nelson at Rocky Mountain (RM).
+Hilary Noble, Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
+We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles (Rick) Williams at Ray J.
+Davis (IDS), (B)Ernie Nelson at Rocky Mountain (RM).
 We thank the original collectors of these specimens: D.
@@
 15.
-Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
 The Bureau of Land Management is thanked as many plant specimens were collected by R.C.B as a partner or contractor to the agency, Sarah Burnett and Lauren Price are thanked for sharing AIM data.
@@
 
+\newpage
+
 ```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F}
@@
 
-```{r APPENDIX NUMBER RMBL Pollen Key, echo = F}
-knitr::include_graphics('../graphics/assorted/RMBL_pollen_key.pdf')
+\newpage
+
+```{r APPENDIX NUMBER RMBL Pollen Key 1, echo = F}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-1.pdf')
 ```
+
+\newpage
+
+```{r APPENDIX NUMBER RMBL Pollen Key 2, echo = F}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
+```
+
+\newpage
 
@@
 
+\newpage
+
 ```{r APPENDIX MORPHOTYPE RICHNESS AND ABUNDANCE, echo= F}
 knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
+```
+
+\newpage
+
+```{r, echo = F, out.width = '25%'}
 knitr::include_graphics('../graphics/plots/SppAbundance.png')
 ```
+
+\newpage
+
+```{r Voucher Reference Specimen tables 1, echo = F}
+knitr::include_graphics('../graphics/tables/specimen_tables-1.pdf')
+```
+
+\newpage
+
+```{r Voucher Reference Specimen tables 2, echo = F}
+knitr::include_graphics('../graphics/tables/specimen_tables-2.pdf')
+```
+
+\newpage

### 2022-12-07 script/manuscript_MEE.Rmd
appenddix formats added but need go back and iron out deposition and which slides were made to pollen vouchers in the end

@@
 
-```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F}
+Appendix XX - Reads Per Loci
+
+```{r READS PER LOCI, echo = F}
 knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
@@
 
-```{r APPENDIX NUMBER RMBL Pollen Key 1, echo = F}
+Appendix XX - Pollen Key
+
+```{r Pollen Key 1, echo = F}
 knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-1.pdf')
@@
 
-```{r APPENDIX NUMBER RMBL Pollen Key 2, echo = F}
+```{r Pollen Key 2, echo = F}
 knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
@@
 \newpage
+
+Appendix XX - Species Distribution Models Predictors
 
@@
 
-```{r APPENDIX MORPHOTYPE RICHNESS AND ABUNDANCE, echo= F}
+APPENDIX XX - Pollen Morphotype Richness Rarefaction Curves
+
+```{r Morphotype Richness, echo= F}
 knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
@@
 
-```{r, echo = F, out.width = '25%'}
+Appendix XX - Pollen Morphotype Abundance Rarefaction Curves
+
+```{r Morphotype Abundance, echo = F, out.width = '25%'}
 knitr::include_graphics('../graphics/plots/SppAbundance.png')
@@
 
-```{r Voucher Reference Specimen tables 1, echo = F}
+Appendix XX - Molecular Reference Specimen Table
+
+```{r Voucher Reference Specimen table 1, echo = F}
 knitr::include_graphics('../graphics/tables/specimen_tables-1.pdf')
@@
 
-```{r Voucher Reference Specimen tables 2, echo = F}
+```{r Voucher Reference Specimen table 2, echo = F}
 knitr::include_graphics('../graphics/tables/specimen_tables-2.pdf')
@@
 \newpage
+
+Appendix XX - All Species in the Sequence Databases
+
+```{r Seq DB 1, echo = F}
+knitr::include_graphics('../graphics/assorted/kraken_db_spp_table-1.pdf')
+```
+
+\newpage
+
+```{r Seq DB 2, echo = F}
+knitr::include_graphics('../graphics/assorted/kraken_db_spp_table-2.pdf')
+```
+
+\newpage
+
+```{r Seq DB 3, echo = F}
+knitr::include_graphics('../graphics/assorted/kraken_db_spp_table-3.pdf')
+```
+
+\newpage
+
+```{r Seq DB 4, echo = F}
+knitr::include_graphics('../graphics/assorted/kraken_db_spp_table-4.pdf')
+```
+
+\newpage
+
+Appendix XX - All Pollen Reference Slides Used to Establish Morphotypes
+
+```{r Pollen Reference Slides 1, echo = F}
+knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-1.pdf')
+```
+
+\newpage
+
+```{r Pollen Reference Slides 2, echo = F}
+knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-2.pdf')
+```
+
+\newpage
+
+```{r Pollen Reference Slides 3, echo = F}
+knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-3.pdf')
+```
+
+\newpage

### 2022-12-11 script/manuscript_MEE.Rmd
most of ST edits and suggestions incorporated into manuscript

@@
   2) Here we utilize hyb-seq, museum studies, and species distribution modelling, to detect the plant species present in pollen loads collected from Queen Bumble Bees.
-  3) We show that hyb-seq using the Angiosperms 353 probes offers significant promise to metagenomic approaches in real world scenarios.  
-  4) We conclude that these probes offer promise for the identification of plant tissue in both single sample, and metasample contexts.
+  3) By utilizing Species distribution modelling we allow for one to process hyb-seq data, create user specified sequence databases which may use MORE ACCURATE alignment algorithms on personal computers over realistic time periods.
+  4) We show that hyb-seq using the Angiosperms 353 probes, which are currently being used in the largest ever plant systematic endeavor, offers significant promise to metagenomic approaches in real world scenarios.
+  5) We conclude that these probes offer promise for the identification of plant tissue in both single sample, and metasample contexts.
 
@@
 
-  Recently barcoding, and metabarcoding, have shown considerable promise in all Kingdoms of life.
-For example ….
-.
-With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification, @china2011comparative), while most applications laying along this spectrum (@li2015plant, @kress2007two, @cbol2009dna).
-Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant) Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
-
-  Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the spatial extent of the study region.
-While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
-iNaturalist), from a domain exceeding the study area.
-To these candidate species, modelling approaches - such as logistic regression, may be used to identify distances under which taxa warrant further exploration.
-To these candidate species, we generate species distribution models (SDM's), which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results.
-This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
-Moreover, as most next-generation sequence data is deposited as raw-sequence reads, from a processing perspective, it is essential to reduce the candidate species via an approach as such.
+  Recently barcoding (the identification of a sample from a single organism – e.g.
+a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms – e.g.
+soil), have shown considerable promise in all Kingdoms of life (@ruppert2019past).
+With plants the identification of members of certain clades using barcoding has been quite successful (REF), whereas with others results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+Particular challenges with barcodes include the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
+Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes).
 
   Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
-This dataset will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021exploring), resulting in over 14,000 represented species.
+This dataset will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL), resulting in over 14,000 represented species.
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more records upon it's intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in regions e.g.
-the 'Darwin Tree of Life' are being undertaken which will contribute data for applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more records upon it’s intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in regions e.g.
+the ‘Darwin Tree of Life’ are being undertaken which will contribute data  applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
+These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (cite).
+However, the application of metabarcoding still face challenges relating to the enormity of the genomic datasets and the computational power required to process sequence data.
+Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
 
-  Considerable amounts of species interactions are expressed along time (@caradonna2021seeing).
-For the tropics the flowering periods of many plant species display high seasonality, and given the elevated rates of species richness relative to the temperate, this axis may provide an essential filter for identifying material in many metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
-While many pollination interactions are formed and dissolved along the temporal axis in the temperate regions (@caradonna2017interaction), the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which subjugates the utility of these to perform as filters of post-processing results, rather than distinct species assemblage for database generation.
+  To increase the quality of metabarcoding results in plants, we suggest reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the the study region and its ecological characteristics (CITE !?).
+To achieve this goal, we first create a list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
+iNaturalist), from a region exceeding the study area.
+To these candidate species, modelling approaches - such as logistic regression, may be used to identify taxa which warrant further exploration e.g.
+modelling to determine their possibility of presence in metabarcoding samples.
+We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
+Species distribution models (SDM’s) examine the ecological conditions associated with known occurrence of a species to identify where else in the study area might suitable habitats be found.
+This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
+This can also significantly reduce processing time, particularly as as most next-generation sequence data is deposited as raw-sequence reads.
+
+  Considerable amounts of species interactions vary along time (@caradonna2021seeing).
+For the tropics, contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in many metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
+In temperate regions, pollination interactions also vary temporally (@caradonna2017interaction), the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which subjugates the utility of these to perform as filters of post-processing results, rather than distinct species assemblage for database generation.
 Nonetheless, we work through a process which seems applicable to the tropics to utilize the temporal dimension for classifying sequencing results.
@@
 
-Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Gunnison County, Colorado, USA (*APPENDIX 1* for site information).
-Pollinator observations of *Bombus* Latreille spp.
-(Apidae Latreille) were conducted from June - August of 2015.
-Observations of *Bombus* foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and place into abundance bins.
+Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information).
+Pollinator observations of Bombus Latreille spp.
+(Apidae Latreille) were conducted from June - August of 2015 in six study sites characterized by high-montane/subalpine Parkland vegetation communities.
+Observations of Bombus foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and placed into abundance bins.
 Corbiculae loads were, non-lethally, collected once from all Queen individuals encountered.
-The six study sites are in areas characterised by high-montane/subalpine Parkland vegetation communities.
 
@@
 
-To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from the Botanical Information and Ecology Network 'BIEN' (@bien2022), and these taxa had Species Distribution Models (SDMs) generated to predict their suitability.
-The predicted plant species served as a reference for which species to include in the genomic sequence databases.
+We downloaded from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022) all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area .
+We then generated Species Distribution Models (SDMs) to predict their distribution throughout the study area.
+These maps of potential distribution served as a reference to reduce the list of species to include in the genomic sequence databases.
 
-In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our field site and all plant species records were downloaded.
-In order to emulate the perceived stochasticity of collections, this dataset was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing dataset.
+In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
+**_In order to emulate the stochasticity of botanical collecting, this dataset was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing dataset_**.
 The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
@@
 
-Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in *(No.
-21 "Southern Rockies")*, downloaded (n = 23,919) (@omernik1987ecoregions).
-These records were copied into two, initially identical, sets, one for generating machine learning models (Random Forest, and Boosted Regression Tree's), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM).
-The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset, as measured by Morans Index (@moran1950notes, @spdep).
-To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION - need appropriate format for journal).
-To achieve a larger absence dataset 1000 pseudo-absence records were generated for each taxon, each of which was greater than 10km from an occurrence record.
-For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced.
-To achieve this, absence records inside of 10% of the mean sample value of the presence records, for any predictor were removed; the required number of absence records were then randomly sampled.
+We used all occurrence records from BIEN (n = 23,919) within a 50km border of the Omernik level 3 ecoregion, which includes the study area *(No.
+21 “Southern Rockies”)* to construct the species distribution model  (@omernik1987ecoregions).
+These records were copied into two, initially identical, sets, one for generating machine learning models (ML; Random Forest, and Boosted Regression Tree’s), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM) (@barbet2012selecting).
+**Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble).
+No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with differing biologies and ecologies (@elith2006novel, @qiao2015no).
+In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).**
 
-Species abiotic niche predictors were 26 variables at 30m resolution, six related to climate,  five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
-For linear regression models these predictors underwent both vifstep (theta = 10, max observations = 12,500) and vifcor (theta = 0.7, max observations = 12,500), and collinear features were removed leaving 16 variables (@usdm2014).
+We then generated 4,029 absence points , locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (BLM CITATION - need appropriate format for journal).
+To achieve a larger absence dataset, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
+For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced (@barbet2012selecting).
+To achieve this, we removed absence records inside of 10% of the mean sample value of the presence records; the required number of absence records were then randomly sampled.
+
+We used 26 environmental variables at 30m resolution to predict the potential distribution of each species, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
+**These publicly available datasets, were selected as they ...** .
+For linear regression models these predictors underwent both *vifstep* (theta = 10, max observations = 12,500) and *vifcor* (theta = 0.7, max observations = 12,500) to detect highly correlated variables, and collinear features were removed leaving 16 variables (@usdm2014).
 
 Modelling: Random Forest and Boosted Regression Trees, were sub sampled with 30% test and two replicates each before weighted ensemble based on True Skill Statistics (tss) (@sdmPackage).
-Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage).
-The results of these models were extracted to a polygon feature derived from a minimum-spanning tree which encompasses the study area, and species from either ensemble with greater than 50% habitat suitability were considered present for further purposes (@prim1957minimum).
+Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage, @).
+TSS was chosen as the ensemble criterion as it has been shown to work across a range of species occurrences prevalences (@allouche2006assessing).
+The results of these models were extracted on a cell-by-cell basis to a polygon feature derived from a minimum-spanning tree which encompasses the study sites, and species from either ensemble with greater than 50% mean habitat suitability across all cells were considered present for further purposes (@prim1957minimum).
 
@@
 534 species were modelled using Random Forest and Boosted Regression Trees.
-To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (CITE AIM AND GBIF).
+To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (CITE AIM AND @gbifDL2021sdms).
 Additional novel absence records were generated from the AIM dataset to create a dataset where each species has balanced presence and absences.
@@
 
-Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
+Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
 We sequenced the 12 most commonly visited taxa twice using samples from one site within the Gunnison River Drainage and one individual from another population.
@@
 To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
-Using 5 years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
 4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  In order to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes, Divisive Hierarchical Clustering techniques were used.
+  We used Divisive Hierarchical Clustering techniques to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
 Ten readily discernible categorical traits were collected from each specimen in the image collection.
@@
 To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included.
+Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included.
 To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, the SnowUS dataset (@iler2021conceptual, @tran2019cloud) from 2000-2017 was analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
+material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS dataset (@iler2021conceptual, @tran2019cloud) from 2000-2017 was analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
 Herbarium records after the 3rd quantile for melt, and the 1st quantile for snow cover of these metrics were removed.
@@
 
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87.
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
 
@@
 Pat Herendeen for assistance with virtually all aspects of preparing pollen vouchers and the identification process.
+Ian Breckheimer for sharing the SDM predictor variables.
 Hilary Noble, Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
@@
 \newpage
+
+\newpage
+
+*Generalised Linear Models (GLM)*
+
+*Generalised Additive Models (GAM)*
+
+Ensemble learning utilizes many sets of trees, each composed of many decisions, to create a single model.
+Each independent variable ( - or *feature*, may) become a node on the tree, a location on the tree where a binary decision will move towards a predicted outcome.
+Each of the decision tree models which ensemble learning utilizes is a weak models, each of which may suffer due to high variance or bias, but which produce better outcomes than would be expected via random chance.
+When ensembled these models generate a strong model, a model which should have more appropriately balanced variance and bias and predicts outcomes which are more strongly correlated with the expected values than the individual weak models.
+
+*Random Forest (RF)* the training data are continually bootstrap re-sampled, in combination with random subsets of features, to create nodes which attempt to optimally predict a known outcome.
+A large number of trees are then aggregated, via the most common predictions, to generate a final classification prediction tree.
+Each individual prediction tree is generated independently of the others.
+
+*Boosted Regression Tree (BRT)* An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
+All response data and predictor variables are kept available to all trees.

### 2022-12-13 script/manuscript_MEE.Rmd
forget to push somewriting

@@
 
-  Recently barcoding (the identification of a sample from a single organism – e.g.
-a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms – e.g.
-soil), have shown considerable promise in all Kingdoms of life (@ruppert2019past).
-With plants the identification of members of certain clades using barcoding has been quite successful (REF), whereas with others results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of life (@ruppert2019past).
+With plants the identification of members of certain clades using barcoding has been quite successful (REF), whereas with other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
 Particular challenges with barcodes include the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
@@
   Considerable amounts of species interactions vary along time (@caradonna2021seeing).
-For the tropics, contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in many metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
-In temperate regions, pollination interactions also vary temporally (@caradonna2017interaction), the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which subjugates the utility of these to perform as filters of post-processing results, rather than distinct species assemblage for database generation.
-Nonetheless, we work through a process which seems applicable to the tropics to utilize the temporal dimension for classifying sequencing results.
+For the tropics and subtropics, contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in many metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
+In temperate regions, pollination interactions also vary temporally (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which reduces the utility of these to operate as filters in the post-processing of sequence matches.
+Nonetheless, we work through a process which seems applicable to the tropics and subtropics to utilize the temporal dimension for classifying sequencing results.
 
-  We apply these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+  We apply these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
 The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
 Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient).
-These montane ecosystems represent some of the most ecologically resilient and resistant systems in the temperate and offer unparalleled potential as refugial areas for multiple dimensions of biodiversity under climate change.
+*Heterogeneous montane ecosystems simultaneously represent one of the most ecologically resilient areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst experiencing the most abrupt changes in mean annual temperature (CITE).
+An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if these refugial ecosystemn services are to be utilized while maintaining their current species richness.*
 
@@
 
+```{r Sequence and Morphological Vouchers, out.width = "65%", echo = F,  fig.cap="A caption"}
+knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
+```
+
+*note this figure is draft mode, i reached out to C.H.
+Cole to get the official APG colors so we are gonna colour edges with that, I have also drawn phylo pics for almost all the labelled order and need to add them in !*
+
 ## 3.3 | Metabarcoding Pollen identification
@@
 The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1st quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the **500** species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the 'Sedimentary Mid-Elevation Forests', a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, Mdn = 13, max = 24).
+Of the **500** species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, Mdn = 13, max = 24).
 
@@
 
-Appendix XX - Pollen Key
-
-```{r Pollen Key 1, echo = F}
-knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-1.pdf')
-```
-
-\newpage
-
-```{r Pollen Key 2, echo = F}
-knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
-```
-
-\newpage
-
 Appendix XX - Species Distribution Models Predictors
@@
 
+Appendix XX - Molecular Reference Specimen Table (con't)
+
 ```{r Voucher Reference Specimen table 2, echo = F}
@@
 
+Appendix XX - All Species in the Sequence Databases (con't)
+
 ```{r Seq DB 2, echo = F}
@@
 
+Appendix XX - All Species in the Sequence Databases (con't)
+
 ```{r Seq DB 3, echo = F}
@@
 \newpage
+
+Appendix XX - All Species in the Sequence Databases (con't)
 
@@
 
+Appendix XX - All Pollen Reference Slides Used to Establish Morphotypes (con't)
+
 ```{r Pollen Reference Slides 2, echo = F}
@@
 \newpage
+
+Appendix XX - All Pollen Reference Slides Used to Establish Morphotypes (con't)
 
@@
 
+Appendix XX - Pollen Key
+
+```{r Pollen Key 1, echo = F}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-1.pdf')
+```
+
 \newpage
+
+```{r Pollen Key 2, echo = F}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
+```
+
+\newpage
+
+\newpage
+
+Appendix XX - Models used for Species Distribution Model Ensembles
 
@@
 All response data and predictor variables are kept available to all trees.
+
+\newpage

### 2022-12-13 script/manuscript_MEE.Rmd
j edits inserted

@@
 
-The inability to reliably identify plants to the level of species often leaves our understanding of ecosystem function and interactions wanting.
+*< UTILITY OF MASS IDENTIFICATION OF MATERIAL - FRIENDLY STAGE SET >*
+
+*< ISSUES AND COSTS WITH IDENTIFYING MATERIAL TO SPECIES >* many spp defined by interactions (<;
+
+The inability to reliably identify plants down to species can limit our understanding of ecosystem function and interactions.
+This is especially true for genera where the species are well defined based upon ecological and behavioral rather than morphological properties, and hence can serve as key habitat bioindicators(e.g.
+different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).
+The lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and the interactions they have with other species (CITE).
 Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.
-These approaches are untenable in light of the benefits offered by: species in several morphologically difficult genera which serve as bioindicators, preferred partners in ecological interactions, as well as an increasing lack of taxonomic experts (@hebert2003biological).
-Many genera, especially with the formalized advent of integrative taxonomy, have species which are well defined based upon ecological and behavioral rather than morphological properties, the identification of these taxa in degraded areas or without their mutualistic partners is fraught with difficulty.
-Hindering an understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
 
-  The identification of many plant species to terminal taxon is an essential component of nearly all land management programs, where many species in the same genus  (e.g.
-Sagebrush - *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) serve as bioindicators (respectively for 'rangelands', streams, and wetlands), as well as in academic research (@Gage2013HistoricalRO, AIM).
-This endeavour is often mired by lack of diagnostic characters (e.g.
-flowers, fruits, roots or combinations thereof),  and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
-Solutions to this problem are wanting, certain programmes have relied increasingly upon revisiting field sites to identify material using morphological or chemical approaches, whereas academic research has often used high copy number plastid genes as barcodes (Rosentreter et al.
-2021, MORE MORE).
-However, both approaches have significant downsides, the former resource intensive at the landscape scale - and often does not work, while the latter seldom works due to a lack of variability in the currently available barcodes (@liu2014identification).
+  The identification to species is often mired by lack of diagnostic characters (e.g.
+flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
+And revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often does not work.
+
+  *< BARCODING HAS ITS LIMITS CURRENTLY BUT THERE IS POTENTIAL >*
 
@@
 Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes).
+
+  *<  WHAT NOVEL APPROACES IN ADDITION TO A353 ?>*
 
@@
 Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
+
+  *< THIS P -- WHAT WE DO TO TEST THIS OUT >*
 
@@
 
-  We apply these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+  To test these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
 The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
 Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient).
-*Heterogeneous montane ecosystems simultaneously represent one of the most ecologically resilient areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst experiencing the most abrupt changes in mean annual temperature (CITE).
+*Heterogeneous montane ecosystems represent one of the most ecologically resilient areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the most abrupt changes in mean annual temperature (CITE).
 An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if these refugial ecosystemn services are to be utilized while maintaining their current species richness.*
@@
 A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
-These sequences were processed in the same manner as our novel sequences before being placed into the database.
+These sequences were processed in the same manner as our novel sequences .
 The Kraken2 database was built using default parameters.

### 2022-12-20 script/manuscript_MEE.Rmd
big moves on reducing drivel at the end to spec out a discussion hellaty skellty

@@
 
-Although we were able to use an actually fine scale flora to determine the species present at the field site, we suspect a similar approach may be accomplished via quick species richness inventories at sites, and then utilizing a bootstrap approach akin to ours, to the taxa returned from databases to derive these estimates.
+We have demonstrated how hyb-seq data may be used in for plant barcoding in a metagenomic context.
 
-** Mention in here that the general effect of sample size on predictions means that larger samples sizes are required for this to work!!!! **
-Although our temporal results were lackluster, we note that our study area has an incredibly brief growing period.
-and we suspect these temporal results would be useful in sub-tropical and tropical ecosystems.
-FURTHER, the sites used for ground truthing the temporal flowering periods were not randomly selected across the study area, and cannot be used to make inference to the population across the entirety of the study area as we did here.
-Regardless both show good agreement on flower onset, peak flowering, and moderate agreement with flowering cessation.
+These results show that the overall results between **Bumble Bee ecology  ** observational and barcoding are largely congruent. But that ...
+
+**Spatial ... & Timing ** (filters!)
+- Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria.
+- The actual dataset which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
+We consciously chose to do this in order to showcase the possibility of this approach working in less data rich areas.
+- Used fine scale flora to evaluate results of modelling, similar approaches based on quick field surveys which do not require comprehensive floristic endeavors.
+- Timing in other areas may suffer complications due to lacking as strong a signal as that induced by snowmelt, but longer durations of flowering likely to help. 
+- Sample size in phenology estimates had an effect which is likely to be maintained across all results.
+- Regardless both show good agreement on flower onset, peak flowering, and moderate agreement with flowering cessation.
 The disagreement in flowering cessation is perhaps due to more microclimates which retain water, rather than microclimates which allow the early accumulation of heat.
+- Number of botanists rapidy decreasing and little hope for much more field based information
 
-Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria.
-The actual dataset which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
-We consciously chose to do this in order to showcase the possibility of this approach working in less data rich areas.
+**Molecular ... **
+- PAFTOL project provide complete genus level reference dataset for the kingdom
+- Expected to work nearly immediately for dna barcoding of whole plant material.
+- With work expected to work well in meta contexts. 
+- Harvest both variable loci, and variable flanking regions of loci for fine taxonomic resolution
+- Potential for infraspecific inference (@wenzell2021incomplete, Loke et al. in prep, Matt JOhnson has a paper on this too.)
 
-Bayesian framework
-
-Future Directions:
-
-While at the time of writing this there are limited A353 sequence data, the Plant and Fungal Trees of Life (PAFTOL) project, which is sequencing at least a species of each genera in the plant Kingdom will produce sequence data from over 14,000 species.
-Given the extant publicly available genomic data, we conservatively estimate that upon completion of PAFTOL there will be no fewer than 15,500 species (4.4% of all ca.
-350,000 plant species) for which sequence data of a majority of these loci exist (@govaerts2021world).
-Accordingly, projects in the near future may increase the number of metagenomics samples while decreasing the need to create their own plant sequence reference libraries.
-  As a result of PAFTOL the first ever comprehensive phylogenetic hypotheses of all plant genera will be presented.
-In tandem with an increased number of digitised and geo-referenced herbarium specimens, and monitoring programs in natural areas, we believe that geo-informatics, and phylogenetic inference will increase the ability of researchers applying this technique to identifying sequence reads.
-  While our approach emphasises the use of this metagenomic technique for the purpose of identifying pollen, I argue the template and resources we provide here make this approach a suitable candidate for many plant metagenomic tasks.
-While we did not have the resources to explore the possibility of characterising infraspecific characteristics, preliminary results from others (@wenzell2021incomplete, Loke et al.
-in prep) indicate a possibility for these probes to also collect data at the level of populations and individuals.
-**
+**Final P**
+ < bayesian framework >
 
 In regards to better understanding the foraging preferences of *Bombus* feeding in subalpine ecosystems. **JANE AND PAUL SET UP FOR NEAR FUTURE RESULTS?**
+
+**CONCLUSIONS **
 

### 2022-12-20 script/manuscript_MEE.Rmd
skeletor on a halfpipe in the discussion

@@
 
-We have demonstrated how hyb-seq data may be used in for plant barcoding in a metagenomic context.
+**overall stuff and things**
+We have demonstrated how a533 data may be used for plant barcoding in a metagenomic context.
+This was exemplified in an ecologically realistic situation, centered both in a context which serves as an immediate framework for land management studies,  and which produced fundamental natural history information.
+These datasets included taxonomically difficult clades (e.g.
+Mertensia, Lupinus) .
 
@@
 The disagreement in flowering cessation is perhaps due to more microclimates which retain water, rather than microclimates which allow the early accumulation of heat.
-- Number of botanists rapidy decreasing and little hope for much more field based information
+- Number of botanists rapidly decreasing no hope for much more field based information or funding.
 
@@
 - PAFTOL project provide complete genus level reference dataset for the kingdom
-- Expected to work nearly immediately for dna barcoding of whole plant material.
+- Expected to work nearly immediately for DNA barcoding of whole plant material.
 - With work expected to work well in meta contexts. 
 - Harvest both variable loci, and variable flanking regions of loci for fine taxonomic resolution
-- Potential for infraspecific inference (@wenzell2021incomplete, Loke et al. in prep, Matt JOhnson has a paper on this too.)
+- Potential for infra-specific inference (@wenzell2021incomplete, Loke et al.,  @slimp2021potential, @beck2021palmer).
 
 **Final P**
- < bayesian framework >
-
-In regards to better understanding the foraging preferences of *Bombus* feeding in subalpine ecosystems. **JANE AND PAUL SET UP FOR NEAR FUTURE RESULTS?**
+ < bayesian framework > < harvesting loci >
 

### 2022-12-22 script/manuscript_MEE.Rmd
some discussion flex

@@
 **overall stuff and things**
-We have demonstrated how a533 data may be used for plant barcoding in a metagenomic context.
+We have demonstrated how Angiosperms533 data may be used for plant barcoding in a metagenomic context.
 This was exemplified in an ecologically realistic situation, centered both in a context which serves as an immediate framework for land management studies, and which produced fundamental natural history information.
-These datasets included taxonomically difficult clades (e.g.
-Mertensia, Lupinus) .
+These datasets included closely related taxa in notoriously taxonomically difficult clades (e.g.
+*Mertensia*, *Lupinus*).
+We utilized spatial and temporal approaches for creating custom sequence databases which are readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents.
+In conjunction with well designed and implemented field based data collection we show how these methods may be widely used to test a variety of hypotheses related to ecological interactions.
 
-These results show that the overall results between **Bumble Bee ecology  ** observational and barcoding are largely congruent. But that ...
+These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent. But that ...
+Future analyses of the long term dataset... Global studies should aim to corroborate ...
 
 **Spatial ... & Timing ** (filters!)
-- Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria.
+- Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria per application.
 - The actual dataset which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
@@
  < bayesian framework > < harvesting loci >
-
-**CONCLUSIONS **
 

### 2023-01-05 script/manuscript_MEE.Rmd
added in map and some citations on pollen metabarcoding

@@
   Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
-This dataset will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL), resulting in over 14,000 represented species.
+This data set will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL), resulting in over 14,000 represented species.
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
@@
 These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (cite).
-However, the application of metabarcoding still face challenges relating to the enormity of the genomic datasets and the computational power required to process sequence data.
+However, the application of metabarcoding still face challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
 Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
@@
   To test these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+The assessment of the plant species compositions of pollen is a desired results, with several applications, and numerous complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
 The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
@@
 In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
-**_In order to emulate the stochasticity of botanical collecting, this dataset was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing dataset_**.
+**_In order to emulate the stochasticity of botanical collecting, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set_**.
 The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
@@
 We then generated 4,029 absence points , locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (BLM CITATION - need appropriate format for journal).
-To achieve a larger absence dataset, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
+To achieve a larger absence data set, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
 For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced (@barbet2012selecting).
@@
 We used 26 environmental variables at 30m resolution to predict the potential distribution of each species, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
-**These publicly available datasets, were selected as they ...** .
+**These publicly available data sets, were selected as they ...** .
 For linear regression models these predictors underwent both *vifstep* (theta = 10, max observations = 12,500) and *vifcor* (theta = 0.7, max observations = 12,500) to detect highly correlated variables, and collinear features were removed leaving 16 variables (@usdm2014).
@@
 Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage, @).
-TSS was chosen as the ensemble criterion as it has been shown to work across a range of species occurrences prevalences (@allouche2006assessing).
+TSS was chosen as the ensemble criterion as it has been shown to work across a range of species occurrences prevalence (@allouche2006assessing).
 The results of these models were extracted on a cell-by-cell basis to a polygon feature derived from a minimum-spanning tree which encompasses the study sites, and species from either ensemble with greater than 50% mean habitat suitability across all cells were considered present for further purposes (@prim1957minimum).
@@
 To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (CITE AIM AND @gbifDL2021sdms).
-Additional novel absence records were generated from the AIM dataset to create a dataset where each species has balanced presence and absences.
+Additional novel absence records were generated from the AIM data set to create a data set where each species has balanced presence and absences.
 11 or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
@@
 To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS dataset (@iler2021conceptual, @tran2019cloud) from 2000-2017 was analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
+material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 was analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
 Herbarium records after the 3rd quantile for melt, and the 1st quantile for snow cover of these metrics were removed.
@@
 read.csv('../data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "SDM evaluation contingency table",
+    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
     col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
@@
 ## 3.2 | Microscopic Pollen identification
+
 ```{r, echo = F}
@@
 # Do we detect mertensia, asteraceae 1, asteraceae 2, in all molecular samples which they are present in the pollen samples of?
-
 ```
@@
 
-```{r Sequence and Morphological Vouchers, out.width = "65%", echo = F,  fig.cap="A caption"}
+```{r Sequence and Morphological Vouchers, out.width = "65%", echo = F,  fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
+The innermost ring indicates every genus which Queen Bee's were observed to visit.
+The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
+The outermost ring indicates that sequence data were available for at least a member of that genus.
+Branch colors follow APG 4."}
 knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
 ```
-
-*note this figure is draft mode, i reached out to C.H.
-Cole to get the official APG colors so we are gonna colour edges with that, I have also drawn phylo pics for almost all the labelled order and need to add them in !*
 
@@
 
-```{r sequence numbers, echo = F}
+```{r Summary statistics of sequence numbers, echo = F}
 
@@
 
-**APPENDIX X**.
+**APPENDIX X Reads Per Loci**.
 
@@
 
-Based upon subjective review of the three classifiers, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all subsequent analyses.
+Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all subsequent analyses.
 
-```{r, eval = F, echo = F}
+```{r clean sequence numbers from environment, eval = F, echo = F}
 rm(toseq, krak, brack, blast, loci, no_loci)
@@
 
-Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) dataset.
-Of these species relatively high accord was observed between the long-term ground truthed dataset, and the modelled species.
+Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set.
+Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
 There was very strong evidence that the weibull estimates were positively associated with the observed onset (r^2^ = 0.72, p < 0.0001, tau = 0.61) and peak (r^2^ = 0.70, p < 0.0001, tau = 0.65) of flowering, and that the number of herbarium samples had a moderate effect on the estimates (p = 0.004 and p = 0.034 respectively).
@@
 
-```{r phenology estimates, echo=F, fig.cap="A caption", out.width = '75%'}
+```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '75%'}
 knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
@@
 
-```{r Floral Obs Results, echo = F, warning = F}
+```{r Floral Observations Summary Statistics, echo = F, warning = F}
 
@@
 
+## 3.6 | Integrated Observational, Molecular, and Palynological Network
+
+... get ...
+
 # 4 | DISCUSSION
@@
 This was exemplified in an ecologically realistic situation, centered both in a context which serves as an immediate framework for land management studies, and which produced fundamental natural history information.
-These datasets included closely related taxa in notoriously taxonomically difficult clades (e.g.
+These data sets included closely related taxa in notoriously taxonomically difficult clades (e.g.
 *Mertensia*, *Lupinus*).
@@
 These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent. But that ...
-Future analyses of the long term dataset... Global studies should aim to corroborate ...
+Future analyses of the long term data set... Global studies should aim to corroborate ...
 
@@
 - Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria per application.
-- The actual dataset which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
+- The actual data set which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
 We consciously chose to do this in order to showcase the possibility of this approach working in less data rich areas.
@@
 **Molecular ... **
-- PAFTOL project provide complete genus level reference dataset for the kingdom
+- PAFTOL project provide complete genus level reference data set for the kingdom
 - Expected to work nearly immediately for DNA barcoding of whole plant material.
@@
 **Final P**
- < bayesian framework > < harvesting loci >
+ < bayesian framework > & < harvesting loci >  = Possible Application at landscape scale.
 
@@
 J.B.F.
-conceived, and designed all lab work, analyses, assisted with writing, and secured funding for molecular work.
+conceived, and designed all lab work, analyses, and integration of approaches, assisted with writing, and secured funding for molecular work.
 
@@
 We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles (Rick) Williams at Ray J.
-Davis (IDS), (B)Ernie Nelson at Rocky Mountain (RM).
-We thank the original collectors of these specimens: D.
+Davis (IDS), (B)Ernie Nelson at Rocky Mountain (RM); and the collectors: D.
 Knoke, L.
@@
 Sanda and NEB are gratefully acknowledged for technical support and generously sharing samples.
+T.C.H.
+Cole for sharing the Angiosperm Phylogeny 4 colour palette.
 
@@
 
+Appendix XX - Origins of Corbiculae Loads
+
+```{r Site Maps, echo = F}
+knitr::include_graphics('../graphics/plots/siteMaps.png')
+```
+
+\newpage
+
 Appendix XX - Reads Per Loci
 
-```{r READS PER LOCI, echo = F}
+```{r READS PER LOCI, echo = F, out.width = '98%'}
 knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
@@
 
-```{r Morphotype Richness, echo= F}
+```{r Morphotype Richness, echo= F,  out.width = '99%'}
 knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
@@
 
-```{r Morphotype Abundance, echo = F, out.width = '25%'}
+```{r Morphotype Abundance, echo = F, out.width = '99%'}
 knitr::include_graphics('../graphics/plots/SppAbundance.png')
@@
 
-\newpage
-
 Appendix XX - Models used for Species Distribution Model Ensembles

### 2023-01-06 script/manuscript_MEE.Rmd
added comparisions at taxa level for molec and obs

@@
 
+# are mertensia and Asteraceae 1 & 2 detected in all Pollen samples where the bee was COLLECTED ON THEM ?
+
+corb <- read.csv('../data/Corbiculae_collections.csv', header = F) %>% 
+  mutate(V1 = str_remove(V1, 'dark|orange')) %>% 
+  filter(!str_detect(V1, 'bifarius')) %>% 
+  separate(V1, into = c('sample.id', 'genus', 'species', 'caste',
+                        'plant.genus', 'plant.species')) %>% 
+  filter(plant.genus == 'Mertensia')
+
+pmorph <- read.csv('../data/morpho_pollen.csv')
+mertensia_samples <- pmorph %>% 
+  group_by(sample.id) %>% 
+  filter(morphotype == 'MERTENSIA') %>% 
+  select(sample.id, sum)
+
 # Do we detect mertensia, asteraceae 1, asteraceae 2, in all molecular samples which they are present in the pollen samples of?
+
+blst <- read.csv('../data/classified_blast.csv')
+
 ```
@@
 `r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-
 Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all subsequent analyses.
@@
 
-```{r Most used resources by species, eval = F, echo =F}
+```{r Most used resources by species, echo =F, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
 
@@
 
+# Need to do some kind of weighted total proportion... 
+# the number of reads per sample, and the % of reads should be a factor
+gf_lkp2 <- read.csv('../data/genus2family_lkp.csv')
+gf_lkp <- blst %>% select(Genus, Family) %>% distinct(Genus, .keep_all = T) %>% 
+  bind_rows(gf_lkp2)
+
 tb <- blst %>% 
@@
   group_by(species, name) %>% 
-  mutate(Mu_Seqs = sum(Prcnt_seqs)/n)  %>% 
-  distinct(species, name, .keep_all = T)
-
-tb %>% 
-  filter(species == 'Bombus appositus') %>% 
-  ungroup() %>% 
-  mutate(total = sum(Mu_Seqs))
+  mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
+         species = str_remove(species, 'Bombus '))  %>% 
+  distinct(species, name, .keep_all = T) %>% 
+  ungroup(name) %>% 
+  slice_max(Mu_Seqs, n = 10) %>% 
+  select(plant.species = name, species, n, Mu_Seqs, Genus, Family) %>% 
+  mutate(rank = 1:n()) %>% 
+  ungroup()
 
@@
   group_by(species) %>% 
-  arrange(species, -bySpp) %>% 
-  slice_head(n = 10)
+  arrange(-bySpp) %>% 
+  slice_head(n = 10) %>% 
+  mutate(plant.species = case_when(
+    plant.species == 'Adenolinum.lewisii' ~  'Linum lewisii', 
+    plant.species == 'Lupinus.bakeri' ~ 'Lupinus sericeus' , 
+    plant.species == 'Dugaldia.hoopesii' ~ 'Hymenoxys hoopesii', 
+    TRUE ~ plant.species 
+  )) %>% 
+  separate(plant.species, into = 'Genus', extra = 'drop', remove = F) %>% 
+  mutate(plant.species = str_replace(plant.species, '[.]', " ")) %>% 
+  left_join(., gf_lkp, by = 'Genus') %>% 
+  filter(species %in% tb$species) %>% 
+  mutate(rank = 1:n()) %>% 
+  ungroup()
+
+## Which bombus species are shared??
+
+# REMOVE KIRBIELLIUS, ONLY  A SINGLE FEEDING RECORD
+obs <- filter(obs, !species %in% c('kirbiellus', 'mixtus'))
+tb <- filter(tb, !species %in% c('kirbiellus', 'mixtus'))
+
+tblist <- split(tb, tb$species) %>% 
+  map(., ~ select(., plant.species, Genus, Family))
+oblist <- split(obs, obs$species) %>% 
+  map(., ~ select(., plant.species, Genus, Family))
+
+species <- purrr::map2_int(tblist, oblist,
+                ~ inner_join(.x, .y, by = 'plant.species') %>% nrow())
+
+#purrr::map2(tblist, oblist,  ~ inner_join(.x, .y, by = 'Genus') %>%  
+#              distinct(plant.species.y, .keep_all = T))
+
+genus <- purrr::map2_int(tblist, oblist,
+  ~ inner_join(.x, .y, by = 'Genus') %>%
+    distinct(plant.species.y) %>% nrow()
+)
+
+#purrr::map2(tblist, oblist,  ~ inner_join(.x, .y, by = 'Family') %>% 
+#              distinct(plant.species.y, .keep_all = T))
+
+family <- purrr::map2_int(tblist, oblist,
+  ~ inner_join(.x, .y, by = 'Family') %>%
+    distinct(plant.species.y) %>%
+    nrow()
+)
+
+correct <- data.frame(Bombus.Species = names(species),
+           Species = species, Genus = genus, Family = family) %>% 
+          # ObsMax = map_int(oblist, ~ nrow(.))) %>% 
+  pivot_longer(!Bombus.Species, names_to = 'Observation') %>% 
+  mutate(Observation = factor(Observation,  levels = c("Species", "Genus", "Family")))
+
+bee_cols <- c('appositus' = '#A50026FF', 'bifarius' = '#FFA319FF', 
+              'californicus' = '#9632B8', 'flavifrons' = '#EFAE90',
+              'rufocinctus' = '#155F83FF', '#8F3931FF', 
+                'mixtus' = '#350E20FF', 'kirbiellus'  =  '#8074A8', 
+              'nevadensis' = '#C96731')
+
+ggplot(data=correct, aes(x=Observation, y=value,
+                         group = Bombus.Species, color = Bombus.Species)) +
+  geom_line(lwd = 1.25)+
+  geom_point(size = 3, shape = 23, aes(fill = Bombus.Species)) +
+  scale_fill_manual('Bee species', values = bee_cols) +
+  scale_color_manual('Bee species', values = bee_cols) +
+  theme_bw() +
+  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
+  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
+       y = 'Shared Observations', x= NULL) + 
+  theme(panel.grid.major.x = element_blank(),
+        panel.border = element_blank())
+
+rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
+```
+
+```{r}
 
@@
 - Used fine scale flora to evaluate results of modelling, similar approaches based on quick field surveys which do not require comprehensive floristic endeavors.
-- Timing in other areas may suffer complications due to lacking as strong a signal as that induced by snowmelt, but longer durations of flowering likely to help. 
+- Timing in other areas may suffer complications due to lacking as strong a signal as that induced by snow melt, but longer duration of flowering likely to help. 
 - Sample size in phenology estimates had an effect which is likely to be maintained across all results.
@@
 - Harvest both variable loci, and variable flanking regions of loci for fine taxonomic resolution
+- As genes are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high.
 - Potential for infra-specific inference (@wenzell2021incomplete, Loke et al.,  @slimp2021potential, @beck2021palmer).
@@
 **Final P**
- < bayesian framework > & < harvesting loci >  = Possible Application at landscape scale.
+ < bayesian framework > & < harvesting loci >  < Possible Application at landscape scale >  < expert natural history >
 
@@
 
-```{r Morphotype Richness, echo= F,  out.width = '99%'}
+```{r Morphotype Richness, echo= F,  out.width = '95%'}
 knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
@@
 
-Appendix XX - Molecular Reference Specimen Table (con't)
-
 ```{r Voucher Reference Specimen table 2, echo = F}
@@
 \newpage
-
-Appendix XX - All Species in the Sequence Databases (con't)
 
@@
 
-Appendix XX - All Pollen Reference Slides Used to Establish Morphotypes (con't)
-
 ```{r Pollen Reference Slides 2, echo = F}
@@
 \newpage
-
-Appendix XX - All Pollen Reference Slides Used to Establish Morphotypes (con't)
 
@@
 
+```{r Molecular Networks, echo = F, out.width="110%"}
+knitr::include_graphics('../graphics/plots/Molecular_nets.pdf')
+```
+
+\newpage
+
 Appendix XX - Models used for Species Distribution Model Ensembles
@@
 
+The two machine learning models utilize Ensemble learning.
+
 Ensemble learning utilizes many sets of trees, each composed of many decisions, to create a single model.
-Each independent variable ( - or *feature*, may) become a node on the tree, a location on the tree where a binary decision will move towards a predicted outcome.
-Each of the decision tree models which ensemble learning utilizes is a weak models, each of which may suffer due to high variance or bias, but which produce better outcomes than would be expected via random chance.
+Each independent variable ( - or *feature*) may become a node on the tree - i.e.
+a location on the tree where a binary decision will move towards a predicted outcome.
+Each of the decision tree models which ensemble learning utilizes is a weak model, each of which may suffer due to high variance or bias, but which produce better outcomes than would be expected via chance.
 When ensembled these models generate a strong model, a model which should have more appropriately balanced variance and bias and predicts outcomes which are more strongly correlated with the expected values than the individual weak models.

### 2023-01-06 script/pollen_countr3.Rmd
added comparisions at taxa level for molec and obs

---
title: "pollen_countR"
author: "Reed & Emily"
date: "4/9/2021"
output: 
  pdf_document:
      fig_caption: true
      number_sections: true
      includes:
        in_header: float_free.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents 
\listoffigures
\listoftables

# Background

The purposes of this script is to provide estimates of how thorough the sampling of corbiculae loads, via light microscopy, is. As always, the goal is to sample as little as possible in order to obtain estimates of true parameters of the population.

In the first section which we be hidden in PDF report, we create data sheets for recording our data.

# Analyses

```{r library, results = 'hide', message = F, warning = F, echo = FALSE}
library(tidyverse)
library(vegan)
library(iNEXT)
library(data.table)
library(parallel)
```

```{r Import and clean field data associated with corbiculae collections, eval = F, echo = F}
field_dat <- read.csv("L:\\Pollen_Reference_Library\\pollen_countR\\Bombus_corbiculae_pollen_samples.RAW.csv")[,c(1:2,4:5,7)]

# some samples missing the DAY of month, we will say they were the 15th real quick
field_dat <- field_dat %>% 
  mutate(date = if_else(str_detect(date, "^[A-Z]"), paste('15', date, sep = "-"), date)) %>% 
  mutate(date = str_squish(date)) %>% 
  mutate(date = as.Date(date, "%d-%b-%y")) %>% 
  mutate(bombus.species = str_remove(bombus.species, "Bombus[.]")) %>% 
  rename(site = ï..site,
         caste = bombus.caste,
         species = bombus.species) %>% 
  mutate(across(c(site:caste), function(x) as.factor(x)))

```

## Create Data Sheets

Remember each slide has 20*20 2mm^2 cells on it, i.e. 400. That would be a a lot of counting. let's try to avoid that. We are going to sample 8 'transects' and each of the 20 cells.

```{r Create Data Sheets, eval = F, echo = F}
set.seed(5548) # close thine eyes, lift thy fingers high, peck ye number pad.

pollen_df <- field_dat %>% 
  dplyr::select(-date, -caste) %>% 
  slice(rep(1:n(), each = 2)) %>% # rep for the number of transects you have. 
  group_by(sample.id) %>% 
  mutate(transect = c(7, 14)) %>%  # this covers 40 or 10% of all cells

  group_by(sample.id, transect) %>% # make our rows of slides. 
  slice(rep(1:n(), each = 20)) %>% # create rows
  mutate(cell = seq(from = -20, to = -1, by = 1)) %>%

  group_by(sample.id, transect, cell) %>% 
  slice(rep(1:n(), each = 2)) %>% # space for two morphotyps per cell.

  mutate(morphotype = "") %>% # create some columns to hold the user input values. 
  mutate(count = "") %>% 
  mutate(count = as.numeric(count)) %>% 
  mutate(SpR_morpho = "") %>% 
  mutate(SpR_transect = "")

pollen_df <- as.data.frame(pollen_df)

#xlsx::write.xlsx(pollen_df, "Pollen_counting_data_sheet_2.xlsx", row.names = F)
rm(field_dat)
```

# Analyses - Rarefaction

We will use the vegan package for this I guess. I really can't spend more time messing around in R... (cries). But we should really just make this package tidy....fmwal.

First we prep our data, this chunk shows us creating dummy data which we used to code the functions. 
```{r, eval = F, echo = F}
site.id <- pollen_df %>% 
  ungroup() %>% 
  distinct(transect,cell) %>% 
  mutate(plot.id = row_number()) # we can create a unique id for each slide grid cell...

pollen_df_work <- pollen_df %>% 
  ungroup() %>% 
  mutate(morphotype = sample(letters[1:26], size = nrow(.), replace = TRUE)) %>% #dummy
  mutate(count =  sample(0:5, size = nrow(.), replace = TRUE)) %>%  # dummy

  mutate(count = replace_na(count, 0)) %>% # keeper code starts here 
  distinct() %>% # SHOULD drop the cells with no samples.
  merge(., site.id, by = c("transect", "cell"))

pollen_df_new <- pollen_df_work %>% 
  select(sample.id, morphotype, count, plot.id) %>% 
  pivot_wider(names_from = morphotype, values_from = count, values_fn = {sum}, values_fill = 0) #%>% #should only need  the values_fn term in the example due to sample producing more than one 'morphotype' per group...

rm(site.id)
```

Here we will load in the data Emily acquired via microscopy.

```{r, echo = F, message = F, warning = F}

data <- list.files(path = '../data/', pattern = 'xlsx')

pollen_df_new.1 <- readxl::read_xlsx(file.path('../data', data)) # read in the file to R. 
pollen_df_new.1 <- pollen_df_new.1[,2:11]

pollen_df_new.1 <- pollen_df_new.1 %>%
  group_by(sample.id) %>% # form groups of each microscope slide. 
  mutate(count = as.numeric(count)) %>% # ensure that the values in this columns are represented as numbers to the computer
  mutate(count = replace_na(count, 0)) %>% # remember in our sample design NA's are 0's
  filter(transect >= 4, transect <= 17) %>%  # remove the 1st and 20th transects which we stopped counting. 
  mutate(group_sum = sum(count)) %>% # see which slides have been counted...
  drop_na(morphotype) %>% 
  filter(group_sum >= 1) %>% # find only the slides which we have counted pollen grains in .
  mutate(morphotype = str_replace(morphotype, "CAN\'T_DISTINGUISH", "UNKNOWN")) %>% 
  mutate(morphotype = str_replace(morphotype, "INDISTINGUISHABLE_CLUMP", "UNKNOWN")) %>% 
  mutate(morphotype = str_replace(morphotype, "NER", "UNKNOWN")) %>% 
  mutate(morphotype = str_replace(morphotype, "DEHYDRATED", "UNKNOWN")) %>%

  mutate(morphotype = str_replace(morphotype, "SALIX", "D")) %>% 
  mutate(morphotype = str_replace(morphotype, "C\\_\\(LINUM\\)", "LINUM")) %>% 
  mutate(morphotype = str_replace_all(morphotype, " ", "_")) %>% # replace white spaces 
  mutate(morphotype = str_to_upper(morphotype)) %>% # R is case sensitive
  mutate(morphotype = str_remove_all(morphotype, "\\_\\(11B\\)")) %>% # while good to know, unknown will be it's own class. 
  mutate(morphotype = if_else(str_detect(morphotype, "\\?"), "UNKNOWN", morphotype)) %>% # make all unknown observations the same format. 
  mutate(morphotype = str_replace(morphotype, "UNK$", "UNKNOWN")) %>% # make all unknown observations the same format. 
  mutate(morphotype = str_replace_all(morphotype, "UNKNWOWN", "UNKNOWN")) %>% 
  filter(!morphotype %in% c("NO_FJ", "LOW_FJ", "LITTLE_FJ", "LESS_FJ")) %>%  # we cannot say we counted or scored these cells. 
  filter(!sample.id %in% c(28, 29, 59))  %>%  # THIS SLIDES ARE PURELY OF A SINGLE TAXON, AND CANNOT BE ANALYZED. 
  filter(!morphotype %in% c(4,8)) %>% 
  mutate(plot.id = rleid(sample.id, transect, cell))

table(pollen_df_new.1$morphotype)

pollen_df_new.1.rare <- pollen_df_new.1 %>% 
  dplyr::select(-site, -date, -species, -caste, -morpho_richness, -group_sum, -transect, -cell) %>% 
  pivot_wider(names_from = morphotype, values_from = count) %>% #, values_fill = 0
  mutate(across(everything(), ~replace(., . == "NULL", 0))) %>% 
  mutate(across(everything(), ~as.numeric(as.character(.)))) %>% 
  replace(is.na(.), 0) %>% 
  ungroup()

#pollen_df_new_test <- pollen_df_new.1 %>% 
#  distinct(sample.id, transect, cell) %>% 
#  rowid_to_column() %>% 
#  left_join(., pollen_df_new.1, by = c(sample.id = sample.id, transect = transect, cell = cell))

```

```{r, eval = F}

pollen_results <- pollen_df_new.1 %>% 
  group_by(sample.id, morphotype) %>% 
  mutate(sum = sum(count)) %>% 
  distinct(sample.id, morphotype, .keep_all = T) %>% 
  select(site:sample.id, morphotype, sum )

write.csv(pollen_results, '/hdd/yun master/MEE_manuscript/data/morpho_pollen.csv', row.names = F)
```

Now we apply the function and interpret the results.

```{r Species Richness Rarefaction, echo = F, fig.cap= "Non-parametric Species Accumulation Curves", fig.height = 11, fig.width = 8}
to_rare <- split(pollen_df_new.1.rare, f = pollen_df_new.1.rare$sample.id) # split our DF into a list of DF's by slide. 
to_rare <- map(to_rare, ~
                 (.x %>% dplyr::select(-sample.id, -plot.id))) # we need to remove these values which were required to create our list objects. We pluck them out of each DF of the list object.

pollen_df_new.1 %>% 
  group_by(sample.id) %>% 
  distinct(morphotype) %>% 
  count(name = 'MorphotypeRichness') %>% 
  ungroup() %>% 
  summarize(.groups = 'drop', mean = mean(MorphotypeRichness),
            median = median(MorphotypeRichness), min = min(MorphotypeRichness), 
            max = max(MorphotypeRichness))

pollen_df_new.1 %>% 
  group_by(sample.id) %>% 
  mutate(total_grains = sum(count)) %>% 
  ungroup() %>% 
  summarize(mean = mean(total_grains),
            median = median(total_grains), min = min(total_grains), 
            max = max(total_grains))

raw_vegan <- function(x){

  test <- specaccum(x, "random", 1000)
  values <- with(test, data.frame(sites, richness, sd))
  return(values)

}

rich_rare <- data.table::rbindlist(lapply(to_rare[c(1:41,43:44,46:60)], raw_vegan), idcol = T) # a couple still not working, 
rich_rare <- rich_rare %>% rename(sample.id = '.id') %>% 
  mutate(sample.id = as.numeric(sample.id))

ggplot(rich_rare, aes(x = sites, y = richness, colour = "purple")) +
  geom_line() +
  theme_bw() +
  facet_wrap(~sample.id, nrow = 9) +

  theme(
    legend.position = "none",
    #strip.text.x = element_blank(), 
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5)
  ) +

  labs(
  title = "Rarefaction Curves of Species Richness",
  subtitle = "Calculated as random plot order with 1000 bootstrap replicates",
  caption = "Used to assess completeness of subsampling. VEGAN package 'specaccum' function used",
  y = "Morphotype Richness",
  x = "No. of Cells ")

#   Samples 45 and 48 missing due to only a single morphotype being present per slide. 
ggsave(
   "../graphics/plots/species_richness_rarefaction.png",
   plot = last_plot(),
   scale = 1,
   width = 20,
   height =28,
   units =  "cm",
   dpi = 300,
)

rm(raw_vegan, rich_rare, to_rare, pollen_df_new.1.rare)
```

Well it is looking like something like 5 transects (100 cells) would be the best fit across all samples to date.

## Relative abundances of these grains
One way of visualizing the dispersion of count proportions (very fast and easy).

```{r Relative Abundance of Grain Morphotype, echo = F, fig.cap= "Relative Abundance of Grain Morphotypes", fig.height = 11, fig.width = 8}

abundance_metrics_df <- pollen_df_new.1 %>% # pollen.df.work in example test code. 
  dplyr::select(-transect, -cell) %>% 
  ungroup() %>% 
  group_by(sample.id, plot.id) %>% 
  mutate(rel_ab_cell = (count/sum(count))*10) %>% # in each cell, which grain is most abundant. heterogeneity as an artifact of slide preparation??
  ungroup() %>% 
  group_by(sample.id) %>% 
  mutate(count_slide = sum(count)) %>% #how many total COUNTED grains per slide...
  group_by(morphotype) %>% 
  mutate(rel_ab_slide = count_slide/sum(count)*10) %>% # what proportion of the total slide count belongs to each morphotype. 
  ungroup()

ab_to_plot <- abundance_metrics_df %>% 
  dplyr::select(sample.id, morphotype, plot.id, rel_ab_cell, rel_ab_slide)

ab_to_plot %>% 
  ungroup() %>% 
  distinct(sample.id, morphotype, .keep_all = T) %>% 
  ggplot(., aes(rel_ab_slide)) +
  geom_histogram() +
  geom_density(fill = "hotpink", alpha = 0.8) +
  theme_bw() +
  facet_wrap(~sample.id, scales = "free", nrow = 9) +
  theme(
    strip.text.x = element_blank(), 
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5), 
    axis.text.x = element_text(angle =45, vjust = 0.75)
  ) +
  labs(
  title = "Proportion of Each Morphotype as a Fraction of the Total Pollen Grains",
  subtitle = "Relative Abundance of Grain Morphotypes",
  caption = "Strictly for visualization purposes",
  y = "count of morphotypes in bin",
  x = "Relative Abundance")

rm(abundance_metrics_df, ab_to_plot)
```

## Rarefaction of incidence and diversity estimates based on sampling effort

We need to create a slightly unusual object for these tests. Shouldn't be too gnar-bar.

```{r, warning = F, echo = F, fig.cap= "Abundance Rarefaction Estimates", fig.height = 11, fig.width = 8}

rarify_abundance <- function(x){

  plot_cnt <- x %>% 
    dplyr::select(sample.id, count, plot.id, morphotype) %>% 
    group_by(sample.id) %>% 
    distinct(plot.id) %>% 
    add_count(sample.id, name = "no_plots_grid") %>% 
    distinct(sample.id, .keep_all = T) %>% 
    mutate(across(is.factor, as.character)) %>% 
    arrange(sample.id) %>% 
    ungroup() %>% 
    pull(no_plots_grid)

  morpho_cnt <- x %>% 
    dplyr::select(sample.id:count) %>% 
    group_by(sample.id, morphotype) %>% 
    add_tally(name = "no_incidences") %>% 
    mutate(total_obs = sum(count)) %>% 
    distinct(sample.id, morphotype, .keep_all = T) %>% 
    pull(total_obs)

  data <- c(plot_cnt[1], morpho_cnt) # assemble in the list iNEXT format. 
  #must specify  [1] or vector with repeat to length of the second vector...

}

to_rare_ab <- split(pollen_df_new.1, f = pollen_df_new.1$sample.id)
to_rare_ab <- lapply(to_rare_ab, rarify_abundance)

cl <- parallel::makeCluster(detectCores()) # Run in parallel
to_rare_ab <- parallel::parLapply(cl,
                      to_rare_ab,
                      iNEXT,  q = 0, datatype = "abundance", se = TRUE, conf = 0.99, nboot= 1000)
parallel::stopCluster(cl)

to_rare_ab.df <- lapply(to_rare_ab, fortify,  type = 2)
to_rare_ab.df <- rbindlist(to_rare_ab.df, idcol = 'true')
to_rare_ab.df <- to_rare_ab.df %>% 
  rename(site = true) %>% 
  mutate(site = as.numeric(site)) %>% 
  arrange(site)

df.point <- to_rare_ab.df[which(to_rare_ab.df$Method=="Observed"),]
df.line <- to_rare_ab.df[which(to_rare_ab.df$Method!="Observed"),]
df.line$method <- factor(df.line$Method, 
                         c("Interpolated", "Extrapolated"),
                         c("Interpolation", "Extrapolation"))

ggplot(to_rare_ab.df, aes(x=x, y=y)) + 
  facet_wrap(~ site, nrow = 9, scales = "free_x") +
  geom_point(size=1, data=df.point) +
  geom_line(aes(linetype = Method), lwd = 1.5, data = df.line) +
  geom_ribbon(aes(ymin = y.lwr, ymax = y.upr), alpha = 0.8, data = df.line) +

  labs(
    title = "Species Richness Abundance Estimating via Hill Numbers (q = 0)",
    subtitle = "Confidence Interval of 99% with 1000 Bootstrap replicates",
    caption = "Calculated using iNEXT",
    x="Number of Grains Counted", 
    y="Species abundance"
    ) +
  theme_bw() +

  theme(
    legend.position = "none", 
    legend.title = element_blank(),
    legend.box = "vertical",
   # strip.text.x = element_blank(), 
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5), 
    axis.text.x = element_text(angle = 45, vjust = 0.75)
    )

ggsave(
   "../graphics/plots/SppAbundance.png",
   plot = last_plot(),
   scale = 1,
   width = 21,
   height = 28,
   units =  "cm",
   dpi = 300,
)

rm(df.point, df.line, to_rare_ab)
```

```{r, warning = F, echo = F, fig.cap= "Hill-Shannon Rarefaction Estimates", fig.height = 11, fig.width = 8}

to_rare_ab <- split(pollen_df_new.1, f = pollen_df_new.1$sample.id)
to_rare_ab <- lapply(to_rare_ab, rarify_abundance)

cl <- parallel::makeCluster(detectCores())# Run parallel computation
to_rare_ab <- parallel::parLapply(cl,
                      to_rare_ab,
                      iNEXT,  q = 1, datatype = "abundance", se = TRUE, conf = 0.99, nboot= 1000)
parallel::stopCluster(cl)

to_rare_ab.df <- fortify(to_rare_ab, type=2)
to_rare_ab.df <- rbindlist(to_rare_ab.df, )
to_rare_ab.df <- to_rare_ab.df %>% 
  mutate(site = as.numeric(site)) %>% 
  arrange(site)

df.point <- to_rare_ab.df[which(to_rare_ab.df$method=="observed"),]
df.line <- to_rare_ab.df[which(to_rare_ab.df$method!="observed"),]
df.line$method <- factor(df.line$method, 
                         c("interpolated", "extrapolated"),
                         c("interpolation", "extrapolation"))

ggplot(to_rare_ab.df, aes(x=x, y=y)) + 
  facet_wrap(~site, nrow = 9, scales = "free_x") +
  geom_point(size=1, data=df.point) +
  geom_line(aes(linetype = method), lwd=1.5, data = df.line) +
  geom_ribbon(aes(ymin = y.lwr, ymax = y.upr), alpha = 0.2, data = df.line) +

  labs(
    title = "Hill-Shannon (q = 1)",
    subtitle = "CI of 99% with 1000 BS replicates",
    caption = "Calculated using iNEXT",
    x="Number of Grains Counted", 
    y="Species abundance"
    ) +
  theme_bw() +

  theme(
    legend.position = "none", 
    legend.title = element_blank(),
    text = element_text(size=10),
    legend.box = "vertical",
    strip.text.x = element_blank(), 
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5), 
    axis.text.x = element_text(angle = 45, vjust = 0.75)
    )

```

How many grains must be counted to have a satisfied CI of 95% for the relative abundance metrics ?

```{r, echo = F}
to_rare_ab.df %>%
  group_by(site) %>%
  arrange(x) %>% 
  filter(y >= 0.95) %>% 
  slice_min(y, with_ties = F)%>% 
  ungroup() %>% 
  mutate(grains_to_count = ceiling(sum(x)/nrow(.))) %>% 
  dplyr::select(grains_to_count) %>% 
  distinct() %>% 
  head()

rm(to_rare_ab.df)
```

Ok abundance is saturating much more quickly than richness. Typical. But the curves are a little strange.

# Find grains to review the identification of. We completed this in Fall of 2021.
```{r, error = F, message = F}
table(pollen_df_new.1$morphotype) #,pollen_df_new.1$count)
a <- sort(unique(pollen_df_new.1$morphotype))

a <- a[c(
  2,  # apiaceae 0
  3,  # apiaceae 1
  8,  # linum - C
  9,  # can't distinguish
  13, # indistinguishable clump
  16, # NER - could be a reeed artifact from R?
  22, #unknown
  23 #unknown ranunculus
)]

to_review <- filter(pollen_df_new.1, morphotype %in% a) %>%  
  dplyr::select(-caste, -morpho_richness, -group_sum, -plot.id) #%>%  
  arrange(sample.id, transect, cell)

to_review <- to_review[order(to_review$sample.id, to_review$transect, -to_review$cell), ]

#to_review %>% 
#  group_by(morphotype) %>% 
#  summarise(sum(count))

head(to_review)
```

## discarded

https://www.researchgate.net/post/What-is-an-appropriate-method-to-compare-species-counts-containing-many-zero-values

"you can try a generalized linear model with binomial variance function and a logit link, that worked for me. In your case the family might be poisson instead of binomial. Maybe you have to add a weight for example the total number of species present.
In R this would look like:

glm(count~site, weight=totalcount, data=speciesdata, family=binomial(link="logit"))

Hope that helps.
Otherwise here's information on zero inflated models that might help:
http://www.ats.ucla.edu/stat/r/dae/zipoisson.htm"

 ^^^^ Matthias Schumacher on Researchgate question....Don't delete this this may come in handy....

```{r, eval = F, echo =F}

table <- rhandsontable(pollen_df, width = 600, height = 300) %>% 
  hot_col(c("site", "date", "bombus.species", "bombus.caste", "sample.id", "transect"), readOnly = TRUE) %>% 
  hot_col(c("morphotype","count"), readOnly = F)

```

```{r, eval = F, echo = F}
library(DT)

datatable(data, options = list(), class = "display",
    callback = JS("return table;"), rownames, colnames, container,
    caption = NULL, filter = c("none", "bottom", "top"), escape = TRUE,
    style = "default", width = NULL, height = NULL, elementId = NULL,
    fillContainer = getOption("DT.fillContainer", NULL),
    autoHideNavigation = getOption("DT.autoHideNavigation", NULL),
    selection = c("multiple", "single", "none"), extensions = list(),
    plugins = NULL, editable = FALSE)

datatable(pollen_df, 
          editable = "cell", 
          caption = 'Pollen Grain Count Entry Data',
          rownames = FALSE)

#list(target = list(columns = c("morphotype","count"), disable = list(columns = 1:5))),
?DT::datatable
```

### 2023-01-06 script/manuscript_MEE.Rmd
possible mismatch with sequence classifications and sequence numbers

@@
   group_by(sample.id) %>% 
+  mutate(Prop = sum/sum(sum)) %>% 
   filter(morphotype == 'MERTENSIA') %>% 
-  select(sample.id, sum)
+  select(sample.id, Prop, sum)
 
@@
 blst <- read.csv('../data/classified_blast.csv') 
+blst_seqs <- distinct(blst, sample.id) 
+mertensia_samples <- filter(mertensia_samples, sample.id %in% blst_seqs$sample.id)
+
+# # SOMETHING MUST BE VERY WRONG HERE, MISSED MERTENSIA IN NEARLY PURE SAMPLES. 
+blst_mertensia <- blst %>% 
+  filter(Genus == 'Mertensia') %>% 
+  select(sample.id,  Prcnt_seqs) %>% 
+  group_by(sample.id) %>% 
+  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
+  distinct(sample.id, .keep_all = T)
+
+mertensia_detected <- mertensia_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0))
+
+M1 <- glm(DETECTED ~ Prop, mertensia_detected, family = binomial)
+M.df <- data.frame(Prop = seq(0, 100, 0.5))
+M.df$DETECTED <- predict(M1, newdata=M.df, type="response")
+
+ggplot(M.df, aes(x=Prop, y=DETECTED)) +
+  stat_smooth(method="glm", se=T, fullrange=TRUE, 
+              method.args = list(family=binomial))  +
+  theme_bw()
 

### 2023-01-06 script/manuscript_MEE.Rmd
added reporting for matching morpho to molecular

--- 2023-01-06 script/manuscript_MEE.Rmd
@@
 pNumC <- function(...) prettyNum(...,big.mark = ',')
-#print(citation('spdep'), bibtex=TRUE)
+
+prcnt_detect <- function(x, Lprop, Hprop){
+  if(missing(Lprop)){Lprop <- 0}
+
+  y <- x[x$Prop > Lprop & x$Prop < Hprop,]; total <- nrow(y)
+  detected <- round((nrow(y[y$DETECTED == 1,])/total) * 100, 1)
+  Mdn <- round(median(y$Prop),3)
+
+  return(writeLines(paste0('were detected in ',  detected, 
+                '% of samples where the proportion of target grains were between ', 
+                Lprop, '-', Hprop,  ' (n = ', total, ' Mdn = ', Mdn, ')')))
+}
 ```
@@
 ```{r, echo = F}
-a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
-b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia')
-c <- c('Jeffersonia', 'Micranthes', 'Prunus', 'Delphinium', 'Androsace', 
-       'Penstemon', 'Orthocarpus', 'Scuttelaria', 'Aquilegia', 'Castilleja', 'Draba')
-d <- c('Salix', 'Boechera')
+#a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
+#b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia')
+#c <- c('Jeffersonia', 'Micranthes', 'Prunus', 'Delphinium', 'Androsace', 
+#       'Penstemon', 'Orthocarpus', 'Scuttelaria', 'Aquilegia', 'Castilleja', 'Draba')
+#d <- c('Salix', 'Boechera')
 
@@
 
-corb <- read.csv('../data/Corbiculae_collections.csv', header = F) %>% 
-  mutate(V1 = str_remove(V1, 'dark|orange')) %>% 
-  filter(!str_detect(V1, 'bifarius')) %>% 
-  separate(V1, into = c('sample.id', 'genus', 'species', 'caste',
-                        'plant.genus', 'plant.species')) %>% 
-  filter(plant.genus == 'Mertensia')
+pmorph <- read.csv('../data/morpho_pollen.csv') 
+blst <- read.csv('../data/blast_seqs_classified.csv') 
+blst_seqs <- distinct(blst, sample.id)
 
-pmorph <- read.csv('../data/morpho_pollen.csv')
+pmorph <- filter(pmorph, sample.id %in% blst$sample.id) %>% 
+  group_by(sample.id) %>% 
+  mutate(Prop = sum/sum(sum))
+
 mertensia_samples <- pmorph %>% 
+  filter(morphotype == 'MERTENSIA') %>% 
+  select(sample.id, Prop, sum)
+
+mertensia_samples <- filter(mertensia_samples, sample.id %in% blst_seqs$sample.id)
+
+blst_mertensia <- blst %>% 
+  filter(Genus == 'Mertensia') %>% 
+  select(sample.id,  Prcnt_seqs) %>% 
   group_by(sample.id) %>% 
-  filter(morphotype == 'MERTENSIA') %>% 
-  select(sample.id, sum)
+  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
+  distinct(sample.id, .keep_all = T)
 
-# Do we detect mertensia, asteraceae 1, asteraceae 2, in all molecular samples which they are present in the pollen samples of?
+# set up as evaluation table !?
 
-blst <- read.csv('../data/classified_blast.csv')
+mertensia_detected <- mertensia_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0)) 
+ # sample 26 - very strange !!!! this needs investigation
 
+M1 <- glm(DETECTED ~ Prop, mertensia_detected, family = binomial)
+M.df <- data.frame(Prop = seq(0, 1, 0.0001))
+M.df$DETECTED <- predict(M1, newdata=M.df, type="response")
+
+ggplot(mertensia_detected, aes(x=Prop, y=DETECTED)) +
+  stat_smooth(method="glm", se=T, fullrange=TRUE, 
+              method.args = list(family=binomial))  +
+  theme_bw()
+
+# Repeat for Asteraceae 1 & 2
+
+blst_aster <- blst %>% 
+  filter(Family == 'Asteraceae') %>% 
+  select(sample.id,  Prcnt_seqs, Genus) %>% 
+  group_by(sample.id) %>% 
+  distinct(sample.id, .keep_all = T)
+
+# ASTER 2 HERE...
+aster2_samples <- pmorph %>% 
+  filter(morphotype == 'ASTERACEAE_2')
+
+blst_aster2 <- filter(blst_aster, Genus %in% 
+                            c('Taraxacum', 'Helianthella', 'Scabrethia'))
+
+aster2_detected <- aster2_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_aster2$sample.id, 1, 0))
+
+M1 <- glm(DETECTED ~ Prop, aster2_detected, family = binomial)
+M.df <- data.frame(Prop = seq(0, 1, 0.0001))
+M.df$DETECTED <- predict(M1, newdata=M.df, type="response")
+
+ggplot(aster2_detected, aes(x=Prop, y=DETECTED)) +
+  stat_smooth(method="glm", se=T, fullrange=TRUE, method.args = list(family='binomial'))  +
+  theme_bw()
+
+# ASTER 1 HERE... ? WERE ALL OF THESE SPECIES PRUNED... ?
+aster1_samples <- pmorph %>% 
+  filter(morphotype == 'ASTERACEAE_1')
+
+blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 'Scabrethia') )
+
+aster1_detected <- aster1_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_aster1$sample.id, 1, 0))
+
+ggplot(aster1_detected, aes(x=Prop, y=DETECTED)) +
+  stat_smooth(method="glm", se=T, fullrange=TRUE, method.args = list(family='binomial'))  +
+  theme_bw()
+
+# ALL HIGH NUMBER ASTERACEAE
+
+aster_samples <- pmorph %>% 
+  filter(str_detect(morphotype, 'ASTER.*'))
+aster_detected <- aster_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_aster$sample.id, 1, 0))
+
+rm(M1, M.df, M.df,  mertensia_samples, blst, blst_aster, 
+   blst_aster1, blst_aster2, aster1_samples,  aster2_samples, 
+   a, b, c, d, blast_mertensia, aster_samples)
 ```
+
+The pollen samples contained three morphotypes which could readily be identified.
+Two of these mapped to monophyletic clades (Boraginaceae & Heliantheae Alliance), and one to a paraphyletic clade (Asteraceae less Heliantheae).
+Boraginaceae grains `r prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)`.
+Asteraceae type 1, non-helianthoids, `r prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)` Asteraceae type 2, Helianthoids, `r prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)`.
+Both morphotypes of Asteraceae pollen `r prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)`.
 
@@
 The number of counted pollen grains in each sample range from (*MIN* - 16,293, $\bar{x}$ = 2788.685, Mdn = 1453).
+
+```{r}
+rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
+```
 
@@
 
-blst <- read.csv('../data/classified_blast.csv') %>% 
+blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
   distinct(sample.id)
 
-blst <- read.csv('../data/classified_blast.csv') %>% 
+blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
   mutate(across(where(is.character), ~ replace_na(., 'UNKNOWN'))) %>% 

### 2023-01-08 script/manuscript_MEE.Rmd
stable going to try lua for author formatting

@@
 Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O.
-Box 519, Crested Butte, CO 81224, USA
+Box 519, Crested Butte, CO 81224, USA]
 abstract:  |
-  .    
-  1) DNA Barcoding has been remarkably successful in nearly all kingdoms of life and has allowed for the rapid analysis of ecological assemblies.
+  1) DNA Barcoding has been remarkably successful in nearly all kingdoms of life and has allowed for the rapid analysis of ecological assemblages.
 Successful DNA barcoding in the plant kingdom has been more difficult than other kingdoms.
@@
   5) We conclude that these probes offer promise for the identification of plant tissue in both single sample, and metasample contexts.  
-
+#keywords: | 
+#"DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species #distribution model"
 output:
@@
   toc: no
+  pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
   word_document: default
@@
 
-  return(writeLines(paste0('were detected in ',  detected, 
+  r <- paste0('were detected in ',  detected, 
                 '% of samples where the proportion of target grains were between ', 
-                Lprop, '-', Hprop,  ' (n = ', total, ' Mdn = ', Mdn, ')')))
+                Lprop, '-', Hprop,  ' (n = ', total, ' Mdn = ', Mdn, ')')
+  return(r)
 }
+
 ```
@@
 
-*< UTILITY OF MASS IDENTIFICATION OF MATERIAL - FRIENDLY STAGE SET >*
-
-*< ISSUES AND COSTS WITH IDENTIFYING MATERIAL TO SPECIES >* many spp defined by interactions (<;
-
-The inability to reliably identify plants down to species can limit our understanding of ecosystem function and interactions.
-This is especially true for genera where the species are well defined based upon ecological and behavioral rather than morphological properties, and hence can serve as key habitat bioindicators(e.g.
+The inability to reliably identify plants down to species can limit our understanding of ecosystem function and interactions (@bortolus2008error).
+This is especially true for genera where species are defined based upon ecological and behavioral rather than morphological properties, and hence can serve as key habitat bioindicators (e.g.
 different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).
-The lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and the interactions they have with other species (CITE).
+The lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
 Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.
@@
 
-  *< BARCODING HAS ITS LIMITS CURRENTLY BUT THERE IS POTENTIAL >*
-
   Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of life (@ruppert2019past).
-With plants the identification of members of certain clades using barcoding has been quite successful (REF), whereas with other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas with other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
 Particular challenges with barcodes include the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
-Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes).
-
-  *<  WHAT NOVEL APPROACES IN ADDITION TO A353 ?>*
 
   Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
-This data set will contain Hyb-Seq data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL), resulting in over 14,000 represented species.
-These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
+This data set will contain hybridization capture (Hyb-Seq) data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL), resulting in over 14,000 represented species.
+These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding, and the A353 probes are currently being used in many other plant phylogenetic issues increasing the sampling depth of many clades (@baker2021exploring).
 Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more records upon it’s intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in regions e.g.
 the ‘Darwin Tree of Life’ are being undertaken which will contribute data  applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
-These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (cite).
+These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (@kress2017plant).
 However, the application of metabarcoding still face challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
@@
 
-  *< THIS P -- WHAT WE DO TO TEST THIS OUT >*
-
-  To increase the quality of metabarcoding results in plants, we suggest reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the the study region and its ecological characteristics (CITE !?).
+  To increase the quality of metabarcoding results in plants, we suggest reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the the study region and its ecological characteristics.
 To achieve this goal, we first create a list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
@@
 Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient).
-*Heterogeneous montane ecosystems represent one of the most ecologically resilient areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the most abrupt changes in mean annual temperature (CITE).
-An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if these refugial ecosystemn services are to be utilized while maintaining their current species richness.*
+Montane areas often represent the most diverse areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
+An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be utilized while maintaining their current diversity (@loarie2009velocity, @dobrowski2016climate)
 
@@
 (Apidae Latreille) were conducted from June - August of 2015 in six study sites characterized by high-montane/subalpine Parkland vegetation communities.
-Observations of Bombus foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and placed into abundance bins.
+Observations of *Bombus* foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and placed into abundance bins.
 Corbiculae loads were, non-lethally, collected once from all Queen individuals encountered.
@@
 In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
-**_In order to emulate the stochasticity of botanical collecting, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set_**.
+In order to emulate the stochasticity of botanical collecting, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
 The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
@@
 534 species were modelled using Random Forest and Boosted Regression Trees.
-To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (CITE AIM AND @gbifDL2021sdms).
+To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (@gbifDL2021sdms, @aim2019database).
 Additional novel absence records were generated from the AIM data set to create a data set where each species has balanced presence and absences.
@@
 We also sequenced another 15 abundant taxa commonly visited by *Bombus* workers, based on the aforementioned data set (*APPENDIX 4*).
+Plant collections were identified via a variety, and typically a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
 
@@
 
-A local NCBI database was built using the same processed novel and downloaded sequences (@camacho2009blast).
+A local NCBI database was built using the same processed novel and downloaded sequences as the previous databases (@camacho2009blast).
 
@@
 In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
+Subsequent to the first round osf sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
 Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
@@
 
-```{r, echo = F}
-#a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
-#b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia')
-#c <- c('Jeffersonia', 'Micranthes', 'Prunus', 'Delphinium', 'Androsace', 
-#       'Penstemon', 'Orthocarpus', 'Scuttelaria', 'Aquilegia', 'Castilleja', 'Draba')
-#d <- c('Salix', 'Boechera')
+```{r compare microscopic to molecular Pollen ID, echo = F}
+a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
+b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia')
+c <- c('Jeffersonia', 'Micranthes', 'Prunus', 'Delphinium', 'Androsace', 
+       'Penstemon', 'Orthocarpus', 'Scuttelaria', 'Aquilegia', 'Castilleja', 'Draba')
+d <- c('Salix', 'Boechera')
 
@@
 
-# set up as evaluation table !?
-
 mertensia_detected <- mertensia_samples %>% 
@@
 
-M1 <- glm(DETECTED ~ Prop, mertensia_detected, family = binomial)
-M.df <- data.frame(Prop = seq(0, 1, 0.0001))
-M.df$DETECTED <- predict(M1, newdata=M.df, type="response")
-
-ggplot(mertensia_detected, aes(x=Prop, y=DETECTED)) +
-  stat_smooth(method="glm", se=T, fullrange=TRUE, 
-              method.args = list(family=binomial))  +
-  theme_bw()
-
-# Repeat for Asteraceae 1 & 2
-
 blst_aster <- blst %>% 
   filter(Family == 'Asteraceae') %>% 
-  select(sample.id,  Prcnt_seqs, Genus) %>% 
   group_by(sample.id) %>% 
@@
 
-M1 <- glm(DETECTED ~ Prop, aster2_detected, family = binomial)
-M.df <- data.frame(Prop = seq(0, 1, 0.0001))
-M.df$DETECTED <- predict(M1, newdata=M.df, type="response")
-
-ggplot(aster2_detected, aes(x=Prop, y=DETECTED)) +
-  stat_smooth(method="glm", se=T, fullrange=TRUE, method.args = list(family='binomial'))  +
-  theme_bw()
-
 # ASTER 1 HERE... ? WERE ALL OF THESE SPECIES PRUNED... ?
@@
 
-ggplot(aster1_detected, aes(x=Prop, y=DETECTED)) +
-  stat_smooth(method="glm", se=T, fullrange=TRUE, method.args = list(family='binomial'))  +
-  theme_bw()
-
 # ALL HIGH NUMBER ASTERACEAE
@@
 
-rm(M1, M.df, M.df,  mertensia_samples, blst, blst_aster, 
-   blst_aster1, blst_aster2, aster1_samples,  aster2_samples, 
-   a, b, c, d, blast_mertensia, aster_samples)
+aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
+bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
+cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
+dd <-  prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
+
+rm(mertensia_samples, blst, blst_aster, blst_aster1,
+   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, 
+   blst_mertensia, aster_samples)
 ```
-
-The pollen samples contained three morphotypes which could readily be identified.
-Two of these mapped to monophyletic clades (Boraginaceae & Heliantheae Alliance), and one to a paraphyletic clade (Asteraceae less Heliantheae).
-Boraginaceae grains `r prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)`.
-Asteraceae type 1, non-helianthoids, `r prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)` Asteraceae type 2, Helianthoids, `r prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)`.
-Both morphotypes of Asteraceae pollen `r prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)`.
 
@@
 
-```{r}
+```{r, echo = F, warning = F, message = F}
 rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
@@
 ```{r clean sequence numbers from environment, eval = F, echo = F}
-rm(toseq, krak, brack, blast, loci, no_loci)
+rm(toseq, krak, brack, blast, loci, no_loci, aa, bb, cc, dd)
 ```
+
+To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
+The pollen samples contained three morphotypes which could readily be identified via microscopy.
+Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae.
+Boraginaceae grains `r toString(aa)`.
+Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`.
+Both morphotypes of Asteraceae pollen `r toString(dd)`.
 
@@
                          group = Bombus.Species, color = Bombus.Species)) +
-  geom_line(lwd = 1.25)+
+  geom_line(linewidth = 1.25)+
   geom_point(size = 3, shape = 23, aes(fill = Bombus.Species)) +
@@
 
-```{r}
-
-```
-
 ## 3.6 | Integrated Observational, Molecular, and Palynological Network
-
-... get ...
 
@@
 **overall stuff and things**
-We have demonstrated how Angiosperms533 data may be used for plant barcoding in a metagenomic context.
+We have demonstrated how Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context.
 This was exemplified in an ecologically realistic situation, centered both in a context which serves as an immediate framework for land management studies, and which produced fundamental natural history information.
@@
 
+*Complications* We likely had issues detecting pollen from several genera of plants, bsaed upon these and other obesrvationla studies most likely Vicia-Lathyrus and Frasera (CITE), probably due to low quality sequencing of whole plant material...
+(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)
+
 These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent. But that ...
-Future analyses of the long term data set... Global studies should aim to corroborate ...
+Future analyses of the long term data set...
 
 **Spatial ... & Timing ** (filters!)
-- Fewer modelling runs for SDM's likely to be effective for determining inclusion, elastic inclusion criteria per application.
-- The actual data set which was used for training and testing all of the models incorporated into SDM's represented only roughly one quarter of the records available for such purposes.
-We consciously chose to do this in order to showcase the possibility of this approach working in less data rich areas.
-- Used fine scale flora to evaluate results of modelling, similar approaches based on quick field surveys which do not require comprehensive floristic endeavors.
-- Timing in other areas may suffer complications due to lacking as strong a signal as that induced by snow melt, but longer duration of flowering likely to help. 
-- Sample size in phenology estimates had an effect which is likely to be maintained across all results.
-- Regardless both show good agreement on flower onset, peak flowering, and moderate agreement with flowering cessation.
-The disagreement in flowering cessation is perhaps due to more microclimates which retain water, rather than microclimates which allow the early accumulation of heat.
-- Number of botanists rapidly decreasing no hope for much more field based information or funding.
+  We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical).
+To reduce the effects of a low population density of botanists on the maintenance of and production of flora's and to foster metagenomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
+In this proof of concept example we performed several iterations of modelling runs, and several approaches (i.e.
+the 'linear models', and the 'machine learning'), which took notable amounts of compute power.
+We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicitly process to reduce spatial autocorrelation.
+However, given the time required to perform all aspects of a study, even our amount of computation was negligible.
+Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate.
+  Tandem to the lack of continued expertise required to generate and maintain species lists, is the expertise required to continue tracking when major phenological events occur in many plant species at relatively fine scales or under novel climates.
+Knowledge of these events is currently limited to general time periods of only a handful of phenological events and groups of organisms (e.g.
+flowering initiation, or trees) (@prather2004implications, @li2016responses).
+While many programs and initiatives exist to collect phenological information on subsets of easily identifiable charismatic species to detect major trends in phenology, by design these capture only a subset of the extent diversity (@betancourt2005implementing, @havens2007chicago).
+In many instances it appears that while landscapes respond similarly to environmental variables which predict phenological responses, that individual species vary widely in their responses to similar environmental cues, or respond to different cues (@augspurger2020concordance, @xie2015deciduous, @xie2018predicting, @caradonna2014shifts).
+As can be seen here, predictions of when a single, major phenological event occurs is already data limited, with sample size having an effect on the subset of species which we could even generate weibull estimates for.
 
 **Molecular ... **
-- PAFTOL project provide complete genus level reference data set for the kingdom
-- Expected to work nearly immediately for DNA barcoding of whole plant material.
-- With work expected to work well in meta contexts. 
-- Harvest both variable loci, and variable flanking regions of loci for fine taxonomic resolution
-- As genes are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high.
-- Potential for infra-specific inference (@wenzell2021incomplete, Loke et al.,  @slimp2021potential, @beck2021palmer).
+  The nearly complete Plant and Fungal Tree of Life (PAFTOL) will provide a comprehensive phylogenetic backbone of the entire plant kingdom, and the inclusion of A353 probes with lineage specific probe sets is common in producing massive genetic datasets (@baker2021exploring).
+We predict that the A353 probes which it is utilizing to work nearly immediately for DNA barcoding of whole plant material, and that more elaborate validation studies in controlled metabarcoding settings, utilizing existing experimental designs,  will have favorable results (@bell2017applying, @bell2019quantitative).
+In particular the harvesting of loci with more variation in certain lineages, and or with more variable flanking regions, will prove promising for identifying closely related plant material (CITE).
+We suspect that conserved reaches of genes resulted in the high amounts of reads in somewhat obscure species.
+Given that the A353 loci are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high, without continual PCR enrichment is possible.
+Recent evidence indicates that the potential for identifying nearly cryptic taxa and even infra-specific inference, of either whole plant material, and perhaps in metagenomic context are possible (@ottenlips2021resolving, @wenzell2021incomplete, Loke et al.
+in prep,  @slimp2021potential, @beck2021palmer).
 
-**Final P**
- < bayesian framework > & < harvesting loci >  < Possible Application at landscape scale >  < expert natural history >
+# 4 | CONCLUSION 
+We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a bayesian framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.
+Herein we crudely utilized this thinking via binary outcomes, should a species predicted be predicted present or not?
+Is it unequivocally flowering or not?
+Myriad data show biological systems and ecological interactions have more variance than can be reasonably discretely parsed.
+We expect that within a bayesian framework studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
+the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems.
+We hope that the promise of A353 probes as tools for metabarcoding play a role in these endeavors.
 
@@
 Williams.
-David Giblin for sharing relevant sections of an advanced draft of FNA V.
+David Giblin &  for sharing relevant sections of an advanced draft of FNA V.
 15.
@@
 
-Ensemble learning utilizes many sets of trees, each composed of many decisions, to create a single model.
+Decision trees, ...
+
+Ensemble learning utilizes many sets of trees, each tree being composed of many binary decisions, to create a single model.
 Each independent variable ( - or *feature*) may become a node on the tree - i.e.
@@
 
-*Boosted Regression Tree (BRT)* An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
+*Boosted Regression Tree (BRT)* (or Gradient Boosted tree) An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
 All response data and predictor variables are kept available to all trees.
 
+Random Forest have high bias and low variance, where boosted regressions trees have low bias and high variances. 
 \newpage

### 2023-01-09 script/manuscript_MEE.Rmd
finally fixed header

@@
 author:  |
-  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E.
-Ogilvie $^3$, Emily J.
-Woodworth $^1$$^,$$^2$, Sophie Taddeo $^1$$^,$$^2$, Paul J.
-CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B.
-Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O.
-Box 519, Crested Butte, CO 81224, USA]
+    |  Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ 
+    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA, 
+    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA,
+    |  $^3$Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
+    |  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$  
 abstract:  |  
@@
   5) We conclude that these probes offer promise for the identification of plant tissue in both single sample, and metasample contexts.  
-#keywords: | 
-#"DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species #distribution model"
+keywords: |  
+  DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species distribution model
 output:
-  pdf_document: default
+  word_document: default
   toc: no
@@
       - '--lua-filter=author-info-blocks.lua'
-  word_document: default
+  pdf_document: default
 csl: "../citations/methods-in-ecology-and-evolution.csl"

### 2023-01-09 script/test.Rmd
finally fixed header

---
title: "My title"
author:
  - Mario Modesto-Mata:
      email: paleomariomm@gmail.com
      institute: [astro, med]
      correspondence: "yes"
  - name: Christopher
    institute: astro
  - name: Seaghán Mhartain
    institute: med
  - name: Rita Yuri Ynoue
    institute: astro
institute:
  - astro: Instituto de Astronomía, Geofísica e Ciências Atmosféricas, Universidade de São Paulo
  - med: Faculdade de Medicina, Universidade de São Paulo
date: "1 October 2018"
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 4
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  word_document:
    toc: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
---

### 2023-01-10 script/manuscript_MEE.Rmd
added some writing on complications and mismatches

@@
 author:  |
-    |  Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ 
+    |  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie$^3$$^,$$^4$ , Emily J. Woodworth$^1$$^,$$^2$  
+    |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$ 
     |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA, 
     |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA,
-    |  $^3$Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
-    |  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$  
+    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
+    |  $^4$ Australia post fire project, ----, ----, ----, AU
 abstract:  |  
@@
 output:
+  pdf_document: default
+  toc: no
   word_document: default
-  toc: no
-  pandoc_args:
-      - '--lua-filter=scholarly-metadata.lua'
-      - '--lua-filter=author-info-blocks.lua'
-  pdf_document: default
 csl: "../citations/methods-in-ecology-and-evolution.csl"
@@
   mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0)) 
- # sample 26 - very strange !!!! this needs investigation
+ # sample 26 - very strange !!!! this needs investigation - ALSO THE ERICOID IS LIKELY ONAGRAD !?
 
@@
   distinct(sample.id, .keep_all = T)
+
+# PERFORM FOR ERICACEAE TOO - A single record maybe come out?
+ericaceae_samples <- pmorph %>% 
+  filter(morphotype == 'ERICACEAE') %>% 
+  select(sample.id, Prop, sum)
+
+blst_ericoid <- filter(blst, Family == 'Ericaceae') # 44 might be Epilobium anagallidolium
+ericaceae_detected <- ericaceae_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_ericoid$sample.id, 1, 0))
 
@@
                             c('Taraxacum', 'Helianthella', 'Scabrethia'))
-
 aster2_detected <- aster2_samples %>% 
@@
 blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 'Scabrethia') )
-
 aster1_detected <- aster1_samples %>% 
@@
 dd <-  prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
+ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)
 
@@
    blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, 
-   blst_mertensia, aster_samples)
+   blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
 ```
@@
 Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`.
+Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
 
@@
 gf_lkp2 <- read.csv('../data/genus2family_lkp.csv')
-gf_lkp <- blst %>% select(Genus, Family) %>% distinct(Genus, .keep_all = T) %>% 
-  bind_rows(gf_lkp2)
+gf_lkp <- blst %>% select(Genus, Family) %>% 
+  distinct(Genus, .keep_all = T) %>% 
+  bind_rows(gf_lkp2) %>% 
+  filter(Family != 'UNKNOWN')
 
@@
   mutate(rank = 1:n()) %>% 
-  ungroup()
+  ungroup() %>% 
+  mutate(Family = case_when(
+    Genus %in% c('Trollius', 'Caltha', 'Aquilegia', 'Delphinium') ~ 'Ranunculaceae',
+    Genus == 'Hydrophyllum' ~ 'Hydrophyllaceae',
+    Genus == 'Epilobium' ~ 'Onagraceae',
+    Genus == 'Viola' ~ 'Violaceae',
+    Genus == 'Salix' ~ 'Saliceae',
+    Genus == 'Cynoglossum' ~ 'Boraginaceae', 
+    Genus == 'Arenaria' ~ 'Caryophyllaceae',
+    Genus == 'Parnasiia' ~ 'Celastraceae',
+    TRUE ~ Family 
+  ))
 
@@
   )) %>% 
+  #mutate(Family = if_else(Family == 'UNKNOWN', 'HYDROPHYLLACEAE', Family)) %>% 
   separate(plant.species, into = 'Genus', extra = 'drop', remove = F) %>% 
@@
                          group = Bombus.Species, color = Bombus.Species)) +
-  geom_line(linewidth = 1.25)+
+  geom_line(linewidth = 1.25, linetype = 'dotted')+
   geom_point(size = 3, shape = 23, aes(fill = Bombus.Species)) +
@@
 
-rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
+n_corbiculae <- tb %>% 
+  distinct(species, .keep_all = T) %>% pull(n) 
+prop_obs <- obs %>% 
+  group_by(species) %>% 
+  slice_max(bySpp, n = 5, with_ties = F) %>% 
+  mutate(prop = sum(bySpp)/total) %>% 
+  distinct(species, .keep_all = T) %>% 
+  pull(prop)
+
+srm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
 ```
@@
 
+For example a common UNKNOWN sequence mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both
+*B.
+nevadensis* and *B.
+rufocinctus* pollen is most likely *Frasera*, failed extraction.
+A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* but where feeding was infrequently observed on *Pedicularis*, likely due to this entire order being represented by only a single molecular reference species.
+
+It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes.
+For example, no more than 10 bee corbiculae loads per species were sequenced with the Mdn = `r median(n_corbiculae)` ...
+, and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top.
+
+...
+many of our results indicate foraging on *Viola* spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads...
+..
+Or the *Epilobium sp.* results indicating that a species such as *Chamerion angustifolium* or *latifolium* is occasionally utilized, as it supported by limited paylnological data.
+
 # 4 | DISCUSSION
 
-**overall stuff and things**
 We have demonstrated how Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context.
-This was exemplified in an ecologically realistic situation, centered both in a context which serves as an immediate framework for land management studies, and which produced fundamental natural history information.
-These data sets included closely related taxa in notoriously taxonomically difficult clades (e.g.
-*Mertensia*, *Lupinus*).
-We utilized spatial and temporal approaches for creating custom sequence databases which are readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents.
-In conjunction with well designed and implemented field based data collection we show how these methods may be widely used to test a variety of hypotheses related to ecological interactions.
+This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history driven fundamental science and the applied science of land management.
+The test pollen loads contained a number of closely related taxa, some in notoriously taxonomically difficult clades (e.g.
+*Mertensia*, *Lupinus*), at naturally occurring proportions.
+We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas.
+By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.
 
-*Complications* We likely had issues detecting pollen from several genera of plants, bsaed upon these and other obesrvationla studies most likely Vicia-Lathyrus and Frasera (CITE), probably due to low quality sequencing of whole plant material...
-(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)
+We anticipate that many of the complications which we faced, using opportunistically collected pollen loads and the first implementation of this method may readily be overcome.
+It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia*, *Lathyrus*, and *Frasera* (@inouye1980effect, @pleasants1980competition), this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.
+**(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)**.
+Additional complications seem to relate to the presence of closely related false positives, e.g.
+frequent classifications of sequences as *Trollius*,  *Caltha*, and *Thalictrum* alongside a more common species in the family, e.g.
+*Delphinium.*.
+Many of our errors are known to us and multiple mnemonics are in *APPENDIX XX* to assist others in future attempts to achieve better results.
+However, the line between false and negative positives may be blurred in some of these instances and warrant further work, for example **Ericaceae pollen grains were observed in a number of samples in trace quantities...**.
 
-These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent. But that ...
+These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent.
+But that ...
+We analyzed pollen loads from all of the most common bumble bee species in the area(@pyke1982local)
 Future analyses of the long term data set...
@@
   The nearly complete Plant and Fungal Tree of Life (PAFTOL) will provide a comprehensive phylogenetic backbone of the entire plant kingdom, and the inclusion of A353 probes with lineage specific probe sets is common in producing massive genetic datasets (@baker2021exploring).
-We predict that the A353 probes which it is utilizing to work nearly immediately for DNA barcoding of whole plant material, and that more elaborate validation studies in controlled metabarcoding settings, utilizing existing experimental designs,  will have favorable results (@bell2017applying, @bell2019quantitative).
+We predict that the A353 probes which it is utilizing to work nearly immediately for DNA barcoding of whole plant material, and that more elaborate validation studies in controlled metabarcoding settings, utilizing existing experimental designs, will have favorable results (@bell2017applying, @bell2019quantitative, @bell2021comparing, @lamb2019quantitative).
 In particular the harvesting of loci with more variation in certain lineages, and or with more variable flanking regions, will prove promising for identifying closely related plant material (CITE).
 We suspect that conserved reaches of genes resulted in the high amounts of reads in somewhat obscure species.
-Given that the A353 loci are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high, without continual PCR enrichment is possible.
+Given that the A353 loci are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high, without continual PCR enrichment is possible; this would align with relatively high efficacy of WGS (@lang2019genome, @peel2019semi, @bell2021comparing).
 Recent evidence indicates that the potential for identifying nearly cryptic taxa and even infra-specific inference, of either whole plant material, and perhaps in metagenomic context are possible (@ottenlips2021resolving, @wenzell2021incomplete, Loke et al.

### 2023-01-10 script/manuscript_MEE.Rmd
typos on the cover page reduced

@@
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$ 
-    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA, 
-    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA,
-    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
+    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
+    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
+    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
     |  $^4$ Australia post fire project, ----, ----, ----, AU
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
-```
-
-```{r Required Libraries, message=F, warning=F, results='hide', echo = F}
+knitr::opts_chunk$set(echo = F, warning = F, message = F)
+```
+
+```{r Required Libraries, message=F, warning=F, results='hide'}
 library(here)
@@
 
-```{r Results for SA Modelling, echo = F, message = F, warning = F}
+```{r Results for SA Modelling}
 
@@
 
-```{r Results for SA Logistic Regression Metrics, echo = F, message = F, warning = F}
+```{r Results for SA Logistic Regression Metrics}
 
@@
 
-```{r Results for SA Logistic Regression Contingency table, echo = F, message = F, warning = F}
+```{r Results for SA Logistic Regression Contingency table}
 read.csv('../data/Pa_conting_table.csv') %>% 
@@
 
-```{r compare microscopic to molecular Pollen ID, echo = F}
+```{r compare microscopic to molecular Pollen ID, eval = T}
 a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
@@
 
-```{r, echo = F, warning = F, message = F}
+```{r}
 rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
@@
 
-```{r Sequence and Morphological Vouchers, out.width = "65%", echo = F,  fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
+```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
 The innermost ring indicates every genus which Queen Bee's were observed to visit.
@@
 
-```{sh Get number of raw and trimmed reads for metagenome samples, eval = F, echo = F}
+```{sh Get number of raw and trimmed reads for metagenome samples, eval = F}
 
@@
 
-```{r Summary statistics of sequence numbers, echo = F}
+```{r Summary statistics of sequence numbers}
 
@@
 
-```{r clean sequence numbers from environment, eval = F, echo = F}
-rm(toseq, krak, brack, blast, loci, no_loci, aa, bb, cc, dd)
-```
-
 To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
@@
 
+```{r clean sequence numbers from environment}
+rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd)
+```
+
 ## 3.4 | Temporal Analyses
@@
 
-```{r Floral Observations Summary Statistics, echo = F, warning = F}
+```{r Floral Observations Summary Statistics }
 
@@
 
-```{r Most used resources by species, echo =F, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
+```{r Most used resources by species, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
 
@@
 
-srm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
+rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
 ```
@@
 Or the *Epilobium sp.* results indicating that a species such as *Chamerion angustifolium* or *latifolium* is occasionally utilized, as it supported by limited paylnological data.
+
+```{r}
+rm(n_corbiculae, prop_obs)
+```
 

### 2023-01-12 script/manuscript_MEE.Rmd
correlation between grains counted and sequence reads classified

@@
 author:  |
-    |  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie$^3$$^,$$^4$ , Emily J. Woodworth$^1$$^,$$^2$  
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$ 
@@
     |  $^4$ Australia post fire project, ----, ----, ----, AU
+    |  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie$^3$$^,$$^4$ , Emily J. Woodworth$^1$$^,$$^2$  
 abstract:  |  
@@
 ```{r compare microscopic to molecular Pollen ID, eval = T}
-a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
-b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia')
-c <- c('Jeffersonia', 'Micranthes', 'Prunus', 'Delphinium', 'Androsace', 
-       'Penstemon', 'Orthocarpus', 'Scuttelaria', 'Aquilegia', 'Castilleja', 'Draba')
-d <- c('Salix', 'Boechera')
 
@@
 pmorph <- read.csv('../data/morpho_pollen.csv') 
-blst <- read.csv('../data/blast_seqs_classified.csv') 
+blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
+  group_by(sample.id) %>% 
+  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
+  ungroup()
 blst_seqs <- distinct(blst, sample.id)
@@
   filter(morphotype == 'MERTENSIA') %>% 
-  select(sample.id, Prop, sum)
+  select(sample.id, morphotype, Prop, sum)
 
@@
   filter(Genus == 'Mertensia') %>% 
-  select(sample.id,  Prcnt_seqs) %>% 
+  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_seqs, Prcnt_Seqs_Class) %>% 
   group_by(sample.id) %>% 
-  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
+  mutate(Prcnt_seqs = sum(Prcnt_seqs),
+         morphotype = 'MERTENSIA') %>% 
   distinct(sample.id, .keep_all = T)
@@
   filter(morphotype == 'ERICACEAE') %>% 
-  select(sample.id, Prop, sum)
-
-blst_ericoid <- filter(blst, Family == 'Ericaceae') # 44 might be Epilobium anagallidolium
+  select(sample.id, Prop, morphotype)
+
+blst_ericoid <- filter(blst, Family == 'Ericaceae') %>% # 44 might be Epilobium anagallidolium
+  mutate(morphotype = 'ERICACEAE')
 ericaceae_detected <- ericaceae_samples %>% 
@@
 blst_aster2 <- filter(blst_aster, Genus %in% 
-                            c('Taraxacum', 'Helianthella', 'Scabrethia'))
+                            c('Taraxacum', 'Helianthella', 'Scabrethia')) %>% 
+  mutate(morphotype = 'ASTERACEAE_2')
 aster2_detected <- aster2_samples %>% 
@@
 
-blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 'Scabrethia') )
+blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 
+                                                'Scabrethia') ) %>% 
+  mutate(morphotype = 'ASTERACEAE_1')
 aster1_detected <- aster1_samples %>% 
@@
 
+# Semi-quantitative accord between sequence % and Grain Count
+palyn_cnt <- bind_rows(aster1_samples, aster2_samples, ericaceae_samples, mertensia_samples) %>% 
+  select(morphotype, countProp = Prop)
+
+pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia, blst_ericoid) %>% 
+  select(morphotype, sequenceProp = Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
+  left_join(., palyn_cnt, by = c('sample.id', 'morphotype')) %>% 
+  drop_na()
+
 rm(mertensia_samples, blst, blst_aster, blst_aster1,
-   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, 
+   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
    blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
@@
 ```
+
+QUANTITATIVE EVIDENCE !?!?!
+
+```{r}
+
+pol_p <- pollen_props %>% 
+  mutate(sequenceProp = sequenceProp/100, 
+         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
+
+summary(model <-  lm(sequenceProp ~ countProp, data = pol_p) ) # better with additive term + morphotype
+summary(model)$adj.r.squared
+summary(model)$p.val
+
+newdat <- data.frame(
+  countProp = seq(from = min(pol_p$countProp), to = max(pol_p$countProp), 
+                  length.out = 100) )#, 
+#  morphotype = sample(unique(pol_p$morphotype), replace = T, size = 100,
+#       prob = as.numeric(table(pol_p$morphotype)) / nrow(pol_p)))
+
+CI <- predict(model, newdat, se.fit = TRUE, interval = "confidence", level = 0.95)
+
+par(pty="s")
+plot(x = pol_p$countProp,  pol_p$sequenceProp, 
+      xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
+     xlab = 'Grains Counted', ylab = 'Classified Sequences',
+     pch=18, slty = 3, gap = 0.025)
+axis(1)
+axis(2)
+abline(model)
+lines(newdat$countProp, CI[['fit']][,3], lty = 'dotted', col = 'grey25')
+lines(newdat$countProp, CI[['fit']][,2], lty = 'dotted', col = 'grey25')
+text(0.2, 0.8, expression("adj. r"^2*" = " ))
+text(0.4, 0.8, paste0(round(summary(model)$adj.r.squared, 3)))
+text(0.3, 0.7, "p < 0.001")
+
+rm(newdat)
+```
+
+```{r Run with correction for total sequences, eval = F}
+## doesnt really make a difference - increase r2 by about 2% - with morphotype up to 6%, multiple r2 0.8
+
+summary(model <-  lm(adjSeqProp ~ countProp, data = pol_p) )  #  + morphotype
+newdat <- data.frame(
+  countProp = seq(from = min(pol_p$countProp), to = max(pol_p$countProp), 
+                  length.out = 100) )
+CI <- predict(model, newdat, se.fit = TRUE, interval = "confidence", level = 0.95)
+
+par(pty="s")
+plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
+      xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
+     xlab = 'Grains Counted', ylab = 'Classified Sequences',
+     pch=18, slty = 3, gap = 0.025)
+
+axis(1)
+axis(2)
+abline(model)
+lines(newdat$countProp, CI[['fit']][,3], lty = 'dotted', col = 'grey25')
+lines(newdat$countProp, CI[['fit']][,2], lty = 'dotted', col = 'grey25')
+text(0.2, 0.8, expression("adj. r"^2*" = " ))
+text(0.4, 0.8, paste0(round(summary(model)$adj.r.squared, 3)))
+text(0.3, 0.7, "p < 0.001")
+
+rm(pol_p, newdat)
+```
+
+The relationship shows that grains which are often present in trace amounts have relatively high amounts of their abundances estimated, and grains which are present in high amounts have fewer sequences classified.
+This is considered to be due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021)
 

### 2023-01-13 script/manuscript_MEE.Rmd
tau for correlatoin

@@
 
-QUANTITATIVE EVIDENCE !?!?!
-
-```{r}
+```{r Correlation between Number of Grains and Number of Sequences, eval = F}
 
@@
 
-summary(model <-  lm(sequenceProp ~ countProp, data = pol_p) ) # better with additive term + morphotype
-summary(model)$adj.r.squared
-summary(model)$p.val
-
-newdat <- data.frame(
-  countProp = seq(from = min(pol_p$countProp), to = max(pol_p$countProp), 
-                  length.out = 100) )#, 
-#  morphotype = sample(unique(pol_p$morphotype), replace = T, size = 100,
-#       prob = as.numeric(table(pol_p$morphotype)) / nrow(pol_p)))
-
-CI <- predict(model, newdat, se.fit = TRUE, interval = "confidence", level = 0.95)
+ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
+# ct[["p.value"]] too small to write out. . tau > 0.4 strong
+summary(model <-  lm(sequenceProp ~ countProp, data = pol_p) )
 
 par(pty="s")
-plot(x = pol_p$countProp,  pol_p$sequenceProp, 
+plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
@@
 abline(model)
-lines(newdat$countProp, CI[['fit']][,3], lty = 'dotted', col = 'grey25')
-lines(newdat$countProp, CI[['fit']][,2], lty = 'dotted', col = 'grey25')
-text(0.2, 0.8, expression("adj. r"^2*" = " ))
-text(0.4, 0.8, paste0(round(summary(model)$adj.r.squared, 3)))
-text(0.3, 0.7, "p < 0.001")
-
-rm(newdat)
-```
-
-```{r Run with correction for total sequences, eval = F}
-## doesnt really make a difference - increase r2 by about 2% - with morphotype up to 6%, multiple r2 0.8
-
-summary(model <-  lm(adjSeqProp ~ countProp, data = pol_p) )  #  + morphotype
-newdat <- data.frame(
-  countProp = seq(from = min(pol_p$countProp), to = max(pol_p$countProp), 
-                  length.out = 100) )
-CI <- predict(model, newdat, se.fit = TRUE, interval = "confidence", level = 0.95)
-
-par(pty="s")
-plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-      xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-     xlab = 'Grains Counted', ylab = 'Classified Sequences',
-     pch=18, slty = 3, gap = 0.025)
-
-axis(1)
-axis(2)
-abline(model)
-lines(newdat$countProp, CI[['fit']][,3], lty = 'dotted', col = 'grey25')
-lines(newdat$countProp, CI[['fit']][,2], lty = 'dotted', col = 'grey25')
-text(0.2, 0.8, expression("adj. r"^2*" = " ))
-text(0.4, 0.8, paste0(round(summary(model)$adj.r.squared, 3)))
-text(0.3, 0.7, "p < 0.001")
+text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
+text(0.3, 0.8, "p < 0.001")
 
@@
 
-The relationship shows that grains which are often present in trace amounts have relatively high amounts of their abundances estimated, and grains which are present in high amounts have fewer sequences classified.
-This is considered to be due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021)
+To detect whether the sequencing reads were semi-quantitative the subset of all pollen `r ` morphotypes distinguishable by microscopy were regressed against the sequence reads.
+In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
+if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed.
+
+The relationship shows that grains which are often present in trace amounts have relatively high amounts of their abundances estimated, and grains which are present in high amounts have relatively low amounts of sequences classified.
+This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
+The correlation between
 

### 2023-01-13 script/manuscript_MEE.Rmd
some writing on correlation

@@
 ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
-# ct[["p.value"]] too small to write out. . tau > 0.4 strong
-summary(model <-  lm(sequenceProp ~ countProp, data = pol_p) )
+# ct[["p.value"]] too small to write out. tau > 0.4 strong relationship
+summary(model <-  lm(adjSeqProp ~ countProp, data = pol_p) )
 
@@
 
-To detect whether the sequencing reads were semi-quantitative the subset of all pollen `r ` morphotypes distinguishable by microscopy were regressed against the sequence reads.
+To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
 In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
 if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed.
+The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
 
@@
 This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
-The correlation between
+There was strong evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
+
+```{r}
+rm(pol_p, ct)
+```
 

### 2023-01-13 script/manuscript_MEE.Rmd
some writing on management actions

@@
 
-To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
+To detect whether the sequencing reads were semi-quantitative the proportion of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
 In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
 if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
-
-The relationship shows that grains which are often present in trace amounts have relatively high amounts of their abundances estimated, and grains which are present in high amounts have relatively low amounts of sequences classified.
+The total percentage of the ten most abundant grains per sample were then were then *corrected* to constitute the entire sample.
+
+The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
 This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
@@
 This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history driven fundamental science and the applied science of land management.
-The test pollen loads contained a number of closely related taxa, some in notoriously taxonomically difficult clades (e.g.
-*Mertensia*, *Lupinus*), at naturally occurring proportions.
+The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g.
+*Mertensia*, *Lupinus*), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic).
 We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas.
@@
 
-**Spatial ... & Timing ** (filters!)
   We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical).
-To reduce the effects of a low population density of botanists on the maintenance of and production of flora's and to foster metagenomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
+To reduce the effects of a low population density of botanists on the maintenance of and production of flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
 In this proof of concept example we performed several iterations of modelling runs, and several approaches (i.e.
@@
 
-**Molecular ... **
   The nearly complete Plant and Fungal Tree of Life (PAFTOL) will provide a comprehensive phylogenetic backbone of the entire plant kingdom, and the inclusion of A353 probes with lineage specific probe sets is common in producing massive genetic datasets (@baker2021exploring).
@@
 
-# 4 | CONCLUSION 
+  The remarkably clear foraging preferences of *Bombus*, both here and across a great many localities globally, reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae across the landscape.
+Numerous historic, and some ongoing, land management actions reduce the ability of many landscapes to support stable populations of *Bombus*.
+Historic, and occasionally current, livestock grazing may be associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle.
+In particular, the removal of locoweeds (Fabaceae: *Astragalus* & *Oxytropis*) and larkspurs (Ranunculaeae: *Delphinium*) were common across public lands administered by the United States Forest Service, and to an extent the Bureau of Land Management (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
+Further actions, generally initiated by settlers to gain water rights or to support mining operations, involved the channelization and incising of streams, and culling of beavers, reducing the extent of wetlands and associated mesic meadows which provide habitat for many species of *Mertensia* (Boraginaceae) widely distributed across Western North America ().
+Fire suppression further resulted in the succession of many Aspen groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effection habtitat for tall *Mertensia* species.
+Finally the effects of Nitrogen deposition, especially given the West's growing population still pose adverse effects on the abundance of a variety of species of Fabaceae (see @stevens2018atmospheric).
+Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitats, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used for grazing.
+The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* to improve Monarch Butterfly (*Danaus plexippus*) habitat provides an effective framework for the latter.
+
+# 5 | CONCLUSION 
 We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a bayesian framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.

### 2023-01-14 script/manuscript_MEE.Rmd
some authorities and rearrangement of appendices

@@
 --- 
-title: Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
-  Patterns of Queen Bumble Bees
+title: Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees
 author:  |
+    |  Reed Clark Benkendorf$^1$$^,$$^2$, Emily J. Woodworth$^1$$^,$$^2$, Jane E. Ogilive$^3$$^,$$^4$,   
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
 
-We then generated 4,029 absence points , locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (BLM CITATION - need appropriate format for journal).
+We then generated 4,029 absence points , locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (@aim2019database).
 To achieve a larger absence data set, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
@@
 The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae.
+Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae Berchtold & J.
+Presl less Heliantheae.
 Boraginaceae grains `r toString(aa)`.
@@
 
-```{r Correlation between Number of Grains and Number of Sequences, eval = F}
+```{r Correlation between Number of Grains and Number of Sequences}
 
@@
 # ct[["p.value"]] too small to write out. tau > 0.4 strong relationship
-summary(model <-  lm(adjSeqProp ~ countProp, data = pol_p) )
+model <-  lm(adjSeqProp ~ countProp, data = pol_p)
 
@@
 
-rm(pol_p, newdat)
+rm( newdat)
 ```
@@
 In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed.
+if both species of *Mertensia* Roth, or one species and read only classified to genus were present in a sample, the reads were summed.
 The total percentage of the ten most abundant grains per sample were then were then *corrected* to constitute the entire sample.
@@
 
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '75%'}
+```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '100%'}
 knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
@@
 *B.
-nevadensis* and *B.
-rufocinctus* pollen is most likely *Frasera*, failed extraction.
-A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* but where feeding was infrequently observed on *Pedicularis*, likely due to this entire order being represented by only a single molecular reference species.
+nevadensis* Cresson and *B.
+rufocinctus* Cresson pollen is most likely *Frasera* Walter, failed extraction.
+A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* (A.
+Heller) Rydb.
+but where feeding was infrequently observed on *Pedicularis* L., likely due to this entire order being represented by only a single molecular reference species.
 
@@
 ...
-many of our results indicate foraging on *Viola* spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads...
+many of our results indicate foraging on *Viola* L.
+spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads...
 ..
-Or the *Epilobium sp.* results indicating that a species such as *Chamerion angustifolium* or *latifolium* is occasionally utilized, as it supported by limited paylnological data.
+Or the *Epilobium* L.
+spp.
+results indicating that a species of *Chamerion* Seg.
+such as *C.
+angustifolium* (L.) Scop.
+or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited paylnological data.
 
@@
 The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g.
-*Mertensia*, *Lupinus*), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic).
+*Mertensia*, *Lupinus* L.), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic).
 We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas.
@@
 We anticipate that many of the complications which we faced, using opportunistically collected pollen loads and the first implementation of this method may readily be overcome.
-It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia*, *Lathyrus*, and *Frasera* (@inouye1980effect, @pleasants1980competition), this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.
+It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia* L., *Lathyrus* L., and *Frasera* (@inouye1980effect, @pleasants1980competition), this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.
 **(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)**.
 Additional complications seem to relate to the presence of closely related false positives, e.g.
-frequent classifications of sequences as *Trollius*,  *Caltha*, and *Thalictrum* alongside a more common species in the family, e.g.
-*Delphinium.*.
+frequent classifications of sequences as *Trollius* L.,  *Caltha* L., and *Thalictrum* L.
+alongside a more common species in the family, e.g.
+*Delphinium.* L.
 Many of our errors are known to us and multiple mnemonics are in *APPENDIX XX* to assist others in future attempts to achieve better results.
@@
 Historic, and occasionally current, livestock grazing may be associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle.
-In particular, the removal of locoweeds (Fabaceae: *Astragalus* & *Oxytropis*) and larkspurs (Ranunculaeae: *Delphinium*) were common across public lands administered by the United States Forest Service, and to an extent the Bureau of Land Management (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
+In particular, the removal of locoweeds (Fabaceae: *Astragalus* L.
+& *Oxytropis* DC.) and larkspurs (Ranunculaeae: *Delphinium*) were common across public lands administered by the United States Forest Service, and to an extent the Bureau of Land Management (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
 Further actions, generally initiated by settlers to gain water rights or to support mining operations, involved the channelization and incising of streams, and culling of beavers, reducing the extent of wetlands and associated mesic meadows which provide habitat for many species of *Mertensia* (Boraginaceae) widely distributed across Western North America ().
-Fire suppression further resulted in the succession of many Aspen groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effection habtitat for tall *Mertensia* species.
+Fire suppression further resulted in the succession of many Aspen groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effects habitat for tall *Mertensia* species.
 Finally the effects of Nitrogen deposition, especially given the West's growing population still pose adverse effects on the abundance of a variety of species of Fabaceae (see @stevens2018atmospheric).
 Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitats, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used for grazing.
-The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* to improve Monarch Butterfly (*Danaus plexippus*) habitat provides an effective framework for the latter.
+The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L.
+to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter.
 
@@
 
-Appendix XX - Origins of Corbiculae Loads
+Appendix 1 - Site Maps
 
@@
 
-Appendix XX - Reads Per Loci
-
-```{r READS PER LOCI, echo = F, out.width = '98%'}
-knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
-```
-
-\newpage
-
-Appendix XX - Species Distribution Models Predictors
-
-Layer                   Description                            Source
-----    ----------------------------------------------  ---------------------
+Appendix 2 - Species Distribution Models Predictors
+
+Layer   LM                     Description                            Source
+----   ----    ------------------------------------------     ---------------------
 1.                   Mean annual cloudiness - MODIS            Wilson et al. 2016
@@
 26.                   Percent surface water                 Global Surface Water Explorer 
-----    ----------------------------------------------  ---------------------
+----        -------------------------------------------      ---------------------
+
+\newpage
+
+Appendix 3 - Molecular Reference Specimen Table
+
+```{r Voucher Reference Specimen table 1, echo = F}
+knitr::include_graphics('../graphics/tables/specimen_tables-1.pdf')
+```
+
+\newpage
+
+```{r Voucher Reference Specimen table 2, echo = F}
+knitr::include_graphics('../graphics/tables/specimen_tables-2.pdf')
+```
+
+\newpage
+
+Appendix 4 - All Pollen Reference Slides Used to Establish Morphotypes
+
+```{r Pollen Reference Slides 1, echo = F}
+knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-1.pdf')
+```
+
+\newpage
+
+```{r Pollen Reference Slides 2, echo = F}
+knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-2.pdf')
+```
+
+\newpage
+
+```{r Pollen Reference Slides 3, echo = F}
+knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-3.pdf')
+```
+
+\newpage
+
+### POLLEN CLUSTER RESULTS SHOULD BE HERE
+
+\newpage
+
+Appendix XX - Pollen Key
+
+```{r Pollen Key 1, echo = F}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-1.pdf')
+```
+
+\newpage
+
+```{r Pollen Key 2, echo = F}
+knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
+```
 
@@
 
-Appendix XX - Molecular Reference Specimen Table
-
-```{r Voucher Reference Specimen table 1, echo = F}
-knitr::include_graphics('../graphics/tables/specimen_tables-1.pdf')
-```
-
-\newpage
-
-```{r Voucher Reference Specimen table 2, echo = F}
-knitr::include_graphics('../graphics/tables/specimen_tables-2.pdf')
-```
-
-\newpage
-
 Appendix XX - All Species in the Sequence Databases
@@
 
-Appendix XX - All Pollen Reference Slides Used to Establish Morphotypes
-
-```{r Pollen Reference Slides 1, echo = F}
-knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-1.pdf')
-```
-
-\newpage
-
-```{r Pollen Reference Slides 2, echo = F}
-knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-2.pdf')
-```
-
-\newpage
-
-```{r Pollen Reference Slides 3, echo = F}
-knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-3.pdf')
-```
-
-\newpage
-
-Appendix XX - Pollen Key
-
-```{r Pollen Key 1, echo = F}
-knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-1.pdf')
-```
-
-\newpage
-
-```{r Pollen Key 2, echo = F}
-knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
-```
-
-\newpage
+Appendix XX - Reads Per Loci
+
+```{r READS PER LOCI, echo = F, out.width = '98%'}
+knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
+```
+
+Appendix XX - Comparison of Kraken2, Bracken, and BLAST
 
@@
 Random Forest have high bias and low variance, where boosted regressions trees have low bias and high variances.
-\newpage
+
+\newpage

### 2023-01-14 script/manuscript_MEE.Rmd
some writing on land management

@@
   Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL), resulting in over 14,000 represented species.
+This data set will contain hybridization capture (Hyb-Seq) data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL, @johnson2019universal), resulting in over 14,000 represented species.
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding, and the A353 probes are currently being used in many other plant phylogenetic issues increasing the sampling depth of many clades (@baker2021exploring).
@@
 the ‘Darwin Tree of Life’ are being undertaken which will contribute data  applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
-These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (@kress2017plant).
+These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
 However, the application of metabarcoding still face challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
@@
 In temperate regions, pollination interactions also vary temporally (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which reduces the utility of these to operate as filters in the post-processing of sequence matches.
-Nonetheless, we work through a process which seems applicable to the tropics and subtropics to utilize the temporal dimension for classifying sequencing results.
+Nonetheless, we work through a general approach which seems applicable to the tropics and subtropics to utilize the temporal dimension for classifying sequencing results (but see @davis2022new)
 
@@
 The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
-Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient).
-Montane areas often represent the most diverse areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
+Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline) often represent the most diverse areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
 An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be utilized while maintaining their current diversity (@loarie2009velocity, @dobrowski2016climate)
@@
 
-We have demonstrated how Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context.
-This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history driven fundamental science and the applied science of land management.
+##### ~ What we DEMONSTRATED ~
+
+We have demonstrated how the Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context (@johnson2019universal, @hollingsworth2016telling).
+This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history driven fundamental science and land management.
 The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g.
@@
 By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.
+
+##### ~ What CHALLENGES we FACED ~
 
@@
 
+##### ~ WHAT we learned about Bee foraging (BASIC) ~
+
 These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent.
 But that ...
-We analyzed pollen loads from all of the most common bumble bee species in the area(@pyke1982local)
+We analyzed pollen loads from all of the most common bumble bee species in the area (@pyke1982local)
 Future analyses of the long term data set...
+
+##### ~ WHAT this reaffirms about Bee foraging (APPLIED) ~
+
+  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae, across Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
+Numerous historic, and some ongoing, land management practices reduce the ability of many landscapes to support stable populations of *Bombus*.
+Historic livestock grazing was often associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle.
+In particular, the removal of locoweeds (Fabaceae: *Astragalus* L.
+& *Oxytropis* DC.) and larkspurs (Ranunculaceae: *Delphinium*) were common across public lands administered by the United States Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
+Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks for significant periods of time, all processes which lower the water tables and reduced the extent of stream-associated [riverine] wetlands and the mesic meadows fringes which provide habitat for many species of tall *Mertensia* (Boraginaceae, e.g.
+*M.
+ciliata* Torr.
+G.
+Don.) widely distributed across Western North America, and to an extent *Delphinium barbeyi* (Huth) Huth and many species of native *Trifolium* L.
+(@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).
+Fire suppression further resulted in the succession of many Aspen (*Populus tremuloides* Michx.) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effects habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
+Finally the effects of Nitrogen deposition, especially given the West's rapidly growing population still pose adverse effects on the abundance of a variety of species of Fabaceae at Urban-Rural interfaces (see @stevens2018atmospheric, @fenn2003ecological).
+Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing.
+The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L., using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
+
+##### ~ WHERE we see spatial/temporal going
 
@@
 Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate.
+*Producing models which lean  towards prioritizing sensitivity rather than specificity is desirable for SDM's in these contexts, as it is most import to winnow  the possible plant species immediately.*
+
   Tandem to the lack of continued expertise required to generate and maintain species lists, is the expertise required to continue tracking when major phenological events occur in many plant species at relatively fine scales or under novel climates.
@@
 As can be seen here, predictions of when a single, major phenological event occurs is already data limited, with sample size having an effect on the subset of species which we could even generate weibull estimates for.
+
+##### ~ WHERE we see MOLECULAR going
 
@@
 
-  The remarkably clear foraging preferences of *Bombus*, both here and across a great many localities globally, reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae across the landscape.
-Numerous historic, and some ongoing, land management actions reduce the ability of many landscapes to support stable populations of *Bombus*.
-Historic, and occasionally current, livestock grazing may be associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle.
-In particular, the removal of locoweeds (Fabaceae: *Astragalus* L.
-& *Oxytropis* DC.) and larkspurs (Ranunculaeae: *Delphinium*) were common across public lands administered by the United States Forest Service, and to an extent the Bureau of Land Management (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
-Further actions, generally initiated by settlers to gain water rights or to support mining operations, involved the channelization and incising of streams, and culling of beavers, reducing the extent of wetlands and associated mesic meadows which provide habitat for many species of *Mertensia* (Boraginaceae) widely distributed across Western North America ().
-Fire suppression further resulted in the succession of many Aspen groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effects habitat for tall *Mertensia* species.
-Finally the effects of Nitrogen deposition, especially given the West's growing population still pose adverse effects on the abundance of a variety of species of Fabaceae (see @stevens2018atmospheric).
-Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitats, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used for grazing.
-The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L.
-to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter.
-
 # 5 | CONCLUSION
+
 We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a bayesian framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.
@@
 \newpage
+
+APPENDIX XX - Tips and tricks for implementing plant metagenomic sequencing

### 2023-01-16 script/manuscript_MEE.Rmd
some writing on sdms in discussion

@@
 
-The inability to reliably identify plants down to species can limit our understanding of ecosystem function and interactions (@bortolus2008error).
-This is especially true for genera where species are defined based upon ecological and behavioral rather than morphological properties, and hence can serve as key habitat bioindicators (e.g.
+The inability to reliably identify plants down to terminal taxon can limit our understanding of ecosystem function and interactions (@bortolus2008error).
+This is especially true for genera where many species are defined upon ecological and behavioral rather than morphological properties, and hence often serve as bioindicators of habitat (e.g.
 different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).
 The lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
-Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques.
-
-  The identification to species is often mired by lack of diagnostic characters (e.g.
+Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques (CITE).
+The identification of organisms to terminal taxon is often mired by lack of diagnostic characters (e.g.
 flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
@@
 
-  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of life (@ruppert2019past).
-With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas with other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
-Particular challenges with barcodes include the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
-
-  Currently the largest plant systematic endeavor ever undertaken, the Kew Plant and Fungal Tree of Life (PAFTOL), is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species representing each genus in the plant kingdom using the popular A353 probes (@baker2021PAFTOL, @johnson2019universal), resulting in over 14,000 represented species.
+  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
+With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas with many other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+Particular challenges with the high copy number barcodes (e.g.
+ITS2, *rbcL*, *matK*, *trnH-psbA*) include the utilization their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
+
+  Currently the largest plant systematic endeavor ever undertaken, the Plant and Fungal Tree of Life (PAFTOL) undertaken by the Royal Botanic Gardens Kew, is approaching completion (@baker2021PAFTOL).
+This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom using the popular Angiosperms353 (A353) probes, at 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal), resulting in over 14,000 represented species.
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding, and the A353 probes are currently being used in many other plant phylogenetic issues increasing the sampling depth of many clades (@baker2021exploring).
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more records upon it’s intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in regions e.g.
-the ‘Darwin Tree of Life’ are being undertaken which will contribute data  applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more species upon it’s intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in geographic regions e.g.
+the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, are being undertaken which will contribute data sets applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
 These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
 However, the application of metabarcoding still face challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
-Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
-
-  To increase the quality of metabarcoding results in plants, we suggest reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the the study region and its ecological characteristics.
+Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
+
+  To increase the quality of metabarcoding results in plants, we suggest reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the the region of study and its ecological characteristics.
 To achieve this goal, we first create a list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
@@
 This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
-This can also significantly reduce processing time, particularly as as most next-generation sequence data is deposited as raw-sequence reads.
-
-  Considerable amounts of species interactions vary along time (@caradonna2021seeing).
-For the tropics and subtropics, contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in many metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
-In temperate regions, pollination interactions also vary temporally (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks in these systems which reduces the utility of these to operate as filters in the post-processing of sequence matches.
-Nonetheless, we work through a general approach which seems applicable to the tropics and subtropics to utilize the temporal dimension for classifying sequencing results (but see @davis2022new)
-
-  To test these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's is consistent across direct observations and the pollen record, an incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-The assessment of the plant species compositions of pollen is a desired results, with several applications, and numerous complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
-The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
-Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous populations declines, while simultaneously serving as the most effective pollinators in temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline) often represent the most diverse areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
-An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be utilized while maintaining their current diversity (@loarie2009velocity, @dobrowski2016climate)
+This approach can also significantly reduce processing time, particularly as as most next-generation sequence data is deposited as raw-sequence reads.
+
+  Considerable amounts of species interactions vary both in space and time (@caradonna2021seeing).
+In many regions of the world, contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
+In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks,  which reduces the utility of these to operate as filters in the post-processing of sequence matches.
+Nonetheless, we work develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new).
+
+  We test these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's are consistent across direct observations and the pollen record, an incongruency noted in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+The assessment of the plant species compositions of pollen is a desired result, with several applications, and numerous complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
+The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble, ...).
+Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline) often represent the most diverse areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
+An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be incorporated into their current diversity (@loarie2009velocity, @dobrowski2016climate).
 
@@
 Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information).
-Pollinator observations of Bombus Latreille spp.
+Pollinator observations of *Bombus* Latreille spp.
 (Apidae Latreille) were conducted from June - August of 2015 in six study sites characterized by high-montane/subalpine Parkland vegetation communities.
@@
 In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
-In order to emulate the stochasticity of botanical collecting, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
+In order to emulate the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
 The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
@@
 
-We then generated 4,029 absence points , locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (@aim2019database).
+We then generated 4,029 absence points, locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (@aim2019database).
 To achieve a larger absence data set, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
 For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced (@barbet2012selecting).
-To achieve this, we removed absence records inside of 10% of the mean sample value of the presence records; the required number of absence records were then randomly sampled.
-
-We used 26 environmental variables at 30m resolution to predict the potential distribution of each species, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
+To achieve this, we removed absence records inside of 10% of the mean sample value of any predictor variable the presence records; the required number of absence records were then randomly sampled.
+
+To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
 **These publicly available data sets, were selected as they ...** .
@@
 Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage, @).
-TSS was chosen as the ensemble criterion as it has been shown to work across a range of species occurrences prevalence (@allouche2006assessing).
+TSS was chosen as the ensemble criterion as it has been shown to work across a wide range of species occurrences prevalence (@allouche2006assessing).
 The results of these models were extracted on a cell-by-cell basis to a polygon feature derived from a minimum-spanning tree which encompasses the study sites, and species from either ensemble with greater than 50% mean habitat suitability across all cells were considered present for further purposes (@prim1957minimum).
@@
 Additional novel absence records were generated from the AIM data set to create a data set where each species has balanced presence and absences.
-11 or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
+Eleven or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
 
@@
 
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
-We sequenced the 12 most commonly visited taxa twice using samples from one site within the Gunnison River Drainage and one individual from another population.
+Using five years (2015-2020) of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
+We sequenced the 12 most commonly visited taxa twice using samples from one site within the Gunnison Basin River Drainage and one individual from another more distal population.
 In addition, for any of these 12 focal species which did not have a congener pair in this filtered sample, we included a congener - or a species from a closely related genus to serve as an outgroup.
-We also sequenced another 15 abundant taxa commonly visited by *Bombus* workers, based on the aforementioned data set (*APPENDIX 4*).
+We also sequenced another 15 abundant taxa commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set (*APPENDIX 4*).
 Plant collections were identified via a variety, and typically a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
@@
 
-To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to 2/3 of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm.
+To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to $\frac{2}{3}$ of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm.
 The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes.
 Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL) , inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
-Pellets were dried at room temperature overnight before resuspension in Nuclease free H2O.
+Pellets were dried at room temperature overnight before resuspension in nuclease free H~2~O.
 Extractions were assessed using a Nanodrop 2000 (Thermo Fisher Scientific) and Qubit fluorometer (Thermo Fisher Scientific).
 DNA extracts were then cleaned using 2:1 v./v.
-Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions.
+Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute ratio following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions.
 DNA was quantified using a Qubit fluorometer.
@@
 Library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
-Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37*C, with an input of 500 ng cleaned DNA.
+Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
 Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed with ½ volume of SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v.
@@
 A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
-These sequences were processed in the same manner as our novel sequences .
+These sequences were processed in the same manner as our novel sequences.
 The Kraken2 database was built using default parameters.
@@
 For example, when only a single species of a genus was known in the study area, but our database used a representative of another taxon in the genus, this species was coded as the result.
-The re-coding of sequences from another representative species for the genus to the sole RMBL representative allowed the identification of XX & % more species.
+The re-coding of sequences from another representative species for the genus to the sole RMBL representative allowed the identification of *XX & %* more species.
 
@@
 
-A local NCBI database was built using the same processed novel and downloaded sequences as the previous databases (@camacho2009blast).
+A local NCBI database was built using the same processed novel and downloaded sequences as the previous database (@camacho2009blast).
 
@@
   To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Cerca* 0.5mm^2 of pollen was placed onto a ~4mm^2 fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
+*Cerca* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
 The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with nail polish; all samples are noted in *APPENDIX 3*.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with nail polish to prevent oxidation; all samples are noted in *APPENDIX 3*.
   To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
 In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round osf sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
+Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
 Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
@@
 To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 if enough records not be found in the nearest 5), and which were from herbarium records were included.
+Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not be found in the nearest 5), and which were from herbarium records were included.
 To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 was analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
-Herbarium records after the 3rd quantile for melt, and the 1st quantile for snow cover of these metrics were removed.
-Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered.
+material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
+Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
+Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
 
@@
 The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, where true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-At the field site, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+
+In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
 The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
@@
 `r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all subsequent analyses.
-
-To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
-The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae Berchtold & J.
-Presl less Heliantheae.
-Boraginaceae grains `r toString(aa)`.
-Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
+Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and it's values were used for all subsequent analyses.
 
@@
 
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3rd quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1st quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the **500** species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.01657, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion in general just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.86885, Mdn = 13, max = 24).
-
-Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set.
+The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+Of the **500** species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
+
+Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (CITE).
 Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
@@
 
-The six sites were surveyed for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours from `r s`-`r e`.
-A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))` - `r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
+The six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours from .
+A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
 Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week min = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, Mdn = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (min = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, Mdn = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
@@
 
-For example a common UNKNOWN sequence mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both
-*B.
+While the spatial results were used to declare the taxonomic composition of the sequence database, temporal results were used in consideration with plant phylogeny to retroactively, reassign the assignment of sequences to taxa.
+Essentially, if a sequence was identified to a taxon which was not known from the field site
+
+For example a common UNKNOWN sequence mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
 nevadensis* Cresson and *B.
@@
 spp.
-results indicating that a species of *Chamerion* Seg.
+results likely indicating that a species of *Chamerion* Seg.
 such as *C.
 angustifolium* (L.) Scop.
-or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited paylnological data.
+or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.
 
@@
 We anticipate that many of the complications which we faced, using opportunistically collected pollen loads and the first implementation of this method may readily be overcome.
-It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia* L., *Lathyrus* L., and *Frasera* (@inouye1980effect, @pleasants1980competition), this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.
+It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia* L., *Lathyrus* L., and *Frasera* (@inouye1980effect, @pleasants1980competition); this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.
 **(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)**.
@@
 However, the line between false and negative positives may be blurred in some of these instances and warrant further work, for example **Ericaceae pollen grains were observed in a number of samples in trace quantities...**.
+We used only a relatively few amount of species records for both the spatial and temporal modelling, we feel this was an appropriate measure to show that these methods would work with data sets which were readily available without considerable data 'scraping'.
+We anticipate that the number of records available to all researchers have grown rapidly since the inception of our modelling (2019) to date, largely as a result of herbarium digitization efforts, and that speak of a massive 'one stop shop' of an integrated database is just over the horizon (@feng2022review, @soltis2017digitization).
+  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area.
+However, we had difficulties in evaluating our predictions in an operational context.
+We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local).
+While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult.
+We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data.
+Further given the, size of the minimum spanning tree (AREA???) which we extracted points to, a formal floristic inventory would still be a time intensive process.
+Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding.
+We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.
+  Difficulties in temporal models related to variability in drivers of flowering phenology.
 
@@
 
-##### ~ WHAT this reaffirms about Bee foraging (APPLIED) ~
-
-  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae, across Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
+##### ~ WHAT this tells us about Bee foraging (APPLIED) ~
+
+  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
 Numerous historic, and some ongoing, land management practices reduce the ability of many landscapes to support stable populations of *Bombus*.
@@
 Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate.
-*Producing models which lean  towards prioritizing sensitivity rather than specificity is desirable for SDM's in these contexts, as it is most import to winnow  the possible plant species immediately.*
+In certain scenarios modelling of predicted species via more formally tailored S(tacked)-SDM or J(oint)-SDM approaches may be beneficial (@wilkinson2021defining, @pinto2021predicting, @schmitt2017ssdm).
 
@@
 flowering initiation, or trees) (@prather2004implications, @li2016responses).
-While many programs and initiatives exist to collect phenological information on subsets of easily identifiable charismatic species to detect major trends in phenology, by design these capture only a subset of the extent diversity (@betancourt2005implementing, @havens2007chicago).
+While many programs and initiatives exist to collect phenological information on subsets of easily identifiable charismatic species to detect major trends in phenology, these capture only a subset of the extent diversity (@betancourt2005implementing, @havens2007chicago).
 In many instances it appears that while landscapes respond similarly to environmental variables which predict phenological responses, that individual species vary widely in their responses to similar environmental cues, or respond to different cues (@augspurger2020concordance, @xie2015deciduous, @xie2018predicting, @caradonna2014shifts).
-As can be seen here, predictions of when a single, major phenological event occurs is already data limited, with sample size having an effect on the subset of species which we could even generate weibull estimates for.
+**As can be seen here, predictions of when a single, major phenological event occurs is already data limited, with sample size having an effect on the subset of species which we could even generate weibull estimates for.
+; check assumptions of model again and just do a quasi something** A more promising approach for the tropics may lay in circular statistics (@park2022herbarium).
 
@@
 in prep,  @slimp2021potential, @beck2021palmer).
+We further believe that in synthetic phylogenetic trees - with incorporation of NGS backbones - will allow in automatic reassignment of reads as a function of phylogenetic distance with measures of uncertainty (@hinchliff2015synthesis, @smith2018constructing, @baker2021PAFTOL).
 
@@
 Williams.
-David Giblin &  for sharing relevant sections of an advanced draft of FNA V.
+David Giblin & Mare Nazaire for sharing relevant sections of an advanced draft of FNA V.
 15.
-The Bureau of Land Management is thanked as many plant specimens were collected by R.C.B as a partner or contractor to the agency, Sarah Burnett and Lauren Price are thanked for sharing AIM data.
-Sanda and NEB are gratefully acknowledged for technical support and generously sharing samples.
+The Bureau of Land Management is thanked as many plant specimens were collected by R.C.B as a partner or contractor to the agency; Sarah Burnett and Lauren Price are thanked for sharing AIM data.
+Sanda and New England Biotech are gratefully acknowledged for technical support and generously sharing samples.
 T.C.H.
 Cole for sharing the Angiosperm Phylogeny 4 colour palette.
+The Program in Plant Biology and Conservation is thanked for funding.
 
@@
 14.               Soil organic carbon (Tonnes / ha)                 SoilGrids
-15.                    Surface soil pH in H2O                       SoilGrids
+15.                    Surface soil pH in H~2~O                     SoilGrids
 16.                   Surface soil percent sand                     SoilGrids
@@
 APPENDIX XX - Tips and tricks for implementing plant metagenomic sequencing
+
+\newpage
+
+```{r}
+knitr::include_graphics('../graphics/assorted/overall_metagenomic_workflow.pdf')
+```
+
+\newpage
+
+```{r Tips and tricks Morphological Workflow}
+knitr::include_graphics('../graphics/assorted/microscopy_workflow.pdf')
+```
+
+\newpage
+
+```{r Tips and tricks Molecular Workflow}
+knitr::include_graphics('../graphics/assorted/molecular_workflow.pdf')
+```
+
+\newpage
+
+```{r Tips and tricks Extraction Buffer}
+knitr::include_graphics('../graphics/assorted/pollen_ctab.pdf')
+```
+
+\newpage
+
+```{r Tips and tricks Extraction Buffer back}
+knitr::include_graphics('../graphics/assorted/pollen_ctab-1.pdf')
+```
+
+\newpage
+
+THIS SHOULD BE TURNED INTO A SMALLER PNG, AND HAVE THE NUMBER OF SEQUENCED METAGENOMIC SAMPLES PLACED INTO THAT COLUMN AND INCLUDED IN TEXT ~~~ NEED THIS !!!!
+
+```{r Samples per Bee}
+knitr::include_graphics('../graphics/tables/bombus_samples.pdf')
+```

### 2023-01-16 script/manuscript_MEE.Rmd
number of species modelled presence in a concensus manner!

@@
 The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the **500** species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
+Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
 

### 2023-01-20 graphics/tables/reclassify_reads_table.Rmd
adding a chart to clarify thinking on post processing reclassification

---
title: NULL
author: NULL
date: NULL
output: 
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}

library(tidyverse)
m <- expand.grid(Spat_Present = c(0,1), 
            Temp_Present = c(0,1),
            Congener_1 = c(0,1),
            Fam_Congener_1 = c(0,1),
            Congener_2 = c(0,1),
            Fam_Congener_2 = c(0,1)
            ) %>% 
  filter(Congener_2 <= Congener_1, Fam_Congener_2 <=  Fam_Congener_1,
         Temp_Present <= Spat_Present) %>%
  filter(rowSums(.) > 0) %>% 
  arrange(-Spat_Present, -Temp_Present, -Congener_1, -Fam_Congener_1) %>%

  mutate(Condition = case_when(
    Spat_Present == 1 & Temp_Present == 1 ~ 'A',
    Spat_Present == 1 & Temp_Present == 0 &
    Congener_1 == 1 & Congener_2 == 0 ~ 'B',
    Spat_Present == 1 & Congener_2 == 1 ~ 'C',
    Spat_Present == 1 & Fam_Congener_1 == 1 & Fam_Congener_2 == 0 ~ 'D',
    Spat_Present == 1 & Fam_Congener_1 == 1 & Fam_Congener_2 == 1 ~ 'E',
    Spat_Present == 1 ~ 'F',
    Spat_Present == 0 & Congener_1 == 1 & Congener_2 == 0 ~ 'G',
    Spat_Present == 0 & Congener_2 == 1 ~ 'H',
    Spat_Present == 0 & Fam_Congener_1 == 1 & Fam_Congener_2 == 0 ~ 'I',
    Spat_Present == 0 & Fam_Congener_1 == 1 & Fam_Congener_2 == 1 ~ 'J',
  ),
  Return = case_when(
    Spat_Present == 1 & Temp_Present == 1 ~ 'Input',
    Spat_Present == 1 & Temp_Present == 0 &
      Congener_1 == 1 & Congener_2 == 0 ~ 'Congener',
    Spat_Present == 1 & Congener_2 == 1 ~ 'Congener',
    Spat_Present == 1 & Fam_Congener_1 == 1 & Fam_Congener_2 == 0 ~ 'Confamilial',
    Spat_Present == 1 & Fam_Congener_1 == 1 & Fam_Congener_2 == 1 ~ 'Confamilial',
    Spat_Present == 1 ~ 'Input',
    Spat_Present == 0 & Congener_1 == 1 & Congener_2 == 0 ~ 'Congener',
    Spat_Present == 0 & Congener_2 == 1 ~ 'Congener',
    Spat_Present == 0 & Fam_Congener_1 == 1 & Fam_Congener_2 == 0 ~ 'Confamilial',
    Spat_Present == 0 & Fam_Congener_1 == 1 & Fam_Congener_2 == 1 ~ 'Confamilial',
  ),
  Rank = case_when(
    Spat_Present == 1 & Temp_Present == 1 ~ 'Species',
    Spat_Present == 1 & Temp_Present == 0 &
      Congener_1 == 1 & Congener_2 == 0 ~ 'Species',
    Spat_Present == 1 & Congener_2 == 1 ~ 'Genus',
    Spat_Present == 1 & Fam_Congener_1 == 1 & Fam_Congener_2 == 0 ~ 'Species',
    Spat_Present == 1 & Fam_Congener_1 == 1 & Fam_Congener_2 == 1 ~ 'Family',
    Spat_Present == 1 ~ 'Species',
    Spat_Present == 0 & Congener_1 == 1 & Congener_2 == 0 ~ 'Species',
    Spat_Present == 0 & Congener_2 == 1 ~ 'Genus',
    Spat_Present == 0 & Fam_Congener_1 == 1 & Fam_Congener_2 == 0 ~ 'Species',
    Spat_Present == 0 & Fam_Congener_1 == 1 & Fam_Congener_2 == 1 ~ 'Family',
  )) %>% 
  rename(Spatial = Spat_Present, Temporal = Temp_Present, Congener = Congener_1,
         Congeners = Congener_2, Confamilial = Fam_Congener_1, 
         Confamilials = Fam_Congener_2) %>% 
  arrange(Condition) %>% 
  group_by(Condition) %>% 
  mutate(Condition = paste(Condition, 1:n(), sep = '.'))

knitr::kable(m, align = "c", 
             caption = "Subset of Possible Combinations for re-classifying Sequences by Incorporating Ecological Factors") %>% 
  kableExtra::kable_styling(
      latex_options = "striped",
      stripe_color = "gray!10"
)

```

Note, for both 'Congener' and 'Confamilial' (*in the singular*) '1' denotes that a species is present; in a sense the genus is monotypic in space and time. For both 'Congener**s**' and 'Confamilial**s**' (*in the plural*), '1' denotes that two or more species are present; 'Confamilial' again representing a monotypic entity in space and time.

\center
Spatial == 1 & Temporal == 1  ~ *__A__*  
Spatial == 1 & Temporal == 0 & Congener = 1  ~ *__B__*

*The temporal dimension is now buffered and a form of __A__ is employed*    
Spatial == 1 & Temporal +/- Buffer == 1 ~ *__X__*

Spatial == 1 & Temporal == 0 & Congener**s** >= 2  ~ *__C__*  
Spatial == 1 & Temporal == 0 & Congener**s** == 0 & Confamilial == 1 ~ *__D__*  
Spatial == 1 & Temporal == 0 & Congener**s** == 0 & Confamilial >= 2 ~ *__E__*   
Spatial == 1 & Temporal == 0 & Congener|**s** == 0 & Confamilial|**s** == 0 ~ *__F__*

Spatial == 0 & Temporal == 0 & Congener == 1  ~ *__G__*    
Spatial == 0 & Temporal == 0 & Congener**s** == 1 ~ *__H__*    
Spatial == 0 & Temporal == 0 & Confamilial == 1  ~ *__I__*   
Spatial == 0 & Temporal == 0 & Confamilial**s** == 1 ~ *__J__*

\center

While the overall order matters, *__X__* in particular may significantly alter conclusions.

### 2023-01-21 script/manuscript_MEE.Rmd
citations for use of alpha taxonomy literature for classing seqs

@@
 
+Sequences which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
+These were manually assigned to a variety of ranks, occasionally to genus #COUNTS# and family #COUNTS#,  by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
+
 ```{r}

### 2023-01-22 script/manuscript_MEE.Rmd
adding in more cleanly processed post class seqs

@@
 pmorph <- read.csv('../data/morpho_pollen.csv') 
-blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
+blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  rename(sample.id = Sample, Family = FAMILY, Genus = GENUS) %>% 
   group_by(sample.id) %>% 
@@
 blst_mertensia <- blst %>% 
-  filter(Genus == 'Mertensia') %>% 
-  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_seqs, Prcnt_Seqs_Class) %>% 
+  filter(Family == 'Boraginaceae') %>% 
+  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
   group_by(sample.id) %>% 
@@
 
+```{r Read Post Processing}
+
+pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') )
+
+knitr::kable(pp_seqs_t, align = "c", 
+             col.names = c('Condition', 'No. Class.', 'Prcnt. Class.', 'Total Seqs', 'Rank'),
+             caption = "Post classification of Sequences via Taxonomy and Ecology") %>% 
+  kableExtra::kable_styling(
+      latex_options = "striped",
+      stripe_color = "gray!10")
+
+seqs_sp <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records'] )
+seqs_ge <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records'] )
+seqs_fa <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records'] )
+
+sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs'] )
+sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs'] )
+fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )
+```
+
+The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
+A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix XX).
+Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
+Each of the top ten taxa which were identified by BLAST from the aforementioned process composed `r sum(pp_seqs_t$Samples_Classified)` distinct records (CONFUSING WORDS), of these `r seqs_sp`% of the reads were classified to species `r sp_total_seqs`% which represent of the total classified sequences, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of the classified reads, and `r seqs_fa`% of the records were classified to genus `r fa_total_seqs`% of total sequences.
+
+*Sequences which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
+These were manually assigned to a variety of ranks, occasionally to genus #COUNTS# and family #COUNTS#, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
+* <- went back on this as new function captures most of it, but don't throw out.
+
+To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
+The pollen samples contained three morphotypes which could readily be identified via microscopy.
+Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae.
+Boraginaceae grains `r toString(aa)`.
+Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`.
+Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
+
 ```{r clean sequence numbers from environment}
-rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd)
+rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, 
+    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa)
 ```
@@
 
-To detect whether the sequencing reads were semi-quantitative the proportion of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
+To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
 In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia* Roth, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then *corrected* to constitute the entire sample.
+if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed.
+The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
 
@@
 There was strong evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
-
-Sequences which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were manually assigned to a variety of ranks, occasionally to genus #COUNTS# and family #COUNTS#,  by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
 
@@
 
-blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
+blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  rename(sample.id = Sample, Genus = GENUS, Family = FAMILY) %>% 
   distinct(sample.id)
@@
     Genus == 'Arenaria' ~ 'Caryophyllaceae',
-    Genus == 'Parnasiia' ~ 'Celastraceae',
+    Genus == 'Parnassia' ~ 'Celastraceae',
+    Genus %in% c('Erigeron', 'Symphyotrichum') ~ 'Asteraceae',
     TRUE ~ Family 
@@
 
+APPENDIX XX - Review Classified Reads and Reassign
+
+```{r}
+knitr::include_graphics('../graphics/tables/reclassify_reads_table.pdf')
+```
+
+\newpage
+
 APPENDIX XX - Tips and tricks for implementing plant metagenomic sequencing
-
-\newpage
 

### 2023-01-22 script/manuscript_MEE.Rmd
added co-occurrence metrics to convey splashout in families

@@
 library(knitr)
+#devtools::install_github("kpmainali/CooccurrenceAffinity")
 suppressWarnings(library(kableExtra))
@@
 
+```{r Co-occurrence of splash drop sequences, eval = F}
+
+e <- blst %>% 
+  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae')) %>% 
+  select(sample.id, Genus) %>% 
+  mutate(Present = 1) %>% 
+  group_by(Genus) %>% 
+  distinct() %>% 
+  pivot_wider(names_from = Genus, values_from = Present) %>% 
+  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
+  column_to_rownames('sample.id')
+
+myout <- CooccurrenceAffinity::affinity(data = e, row.or.col = "col", squarematrix = c("all"))
+CooccurrenceAffinity::plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
+```
+
+To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard,  $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
+Numerous taxa from the family Ranunculaceae (*Caltha* L.
+sp., *Thalictrum* L.
+spp., *Trollius* L.
+sp., *Aquilegia* L.
+spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
+were recorded.
+A similar relationship was observed in the Hydrophyllaceae with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
+species were present.
+
 ## 3.4 | Temporal Analyses
@@
 
-Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (CITE).
+Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
 Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
@@
     Genus == 'Cynoglossum' ~ 'Boraginaceae', 
-    Genus == 'Arenaria' ~ 'Caryophyllaceae',
+    Genus == c('Arenaria', 'Odontostemma') ~ 'Caryophyllaceae',
     Genus == 'Parnassia' ~ 'Celastraceae',
@@
 
-#purrr::map2(tblist, oblist,  ~ inner_join(.x, .y, by = 'Genus') %>%  
-#              distinct(plant.species.y, .keep_all = T))
-
 genus <- purrr::map2_int(tblist, oblist,
@@
 )
-
-#purrr::map2(tblist, oblist,  ~ inner_join(.x, .y, by = 'Family') %>% 
-#              distinct(plant.species.y, .keep_all = T))
 
@@
 
+Results from palynological analyses show that several species of bee show near perfect fidelity to the genus Hydrophyllum on a per visit basis.
+
 ##### ~ WHAT this tells us about Bee foraging (APPLIED) ~
@@
 G.
-Don.) widely distributed across Western North America, and to an extent *Delphinium barbeyi* (Huth) Huth and many species of native *Trifolium* L.
+Don.) widely distributed across Western North America, and to an extent *Delphinium barbeyi*  and many species of native *Trifolium* L.
 (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).

### 2023-01-22 script/manuscript_MEE.Rmd
patched up writing with properly run blast classification!

@@
 ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)
+ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)
 
@@
 
-rm(mertensia_samples, blst, blst_aster, blst_aster1,
+rm(mertensia_samples, blst_aster, blst_aster1,
    blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
@@
 The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`.
-
 **APPENDIX X Reads Per Loci**.
+
+... with samples 22, and 41 both having < 100 reads indicates, virtual failure of these records (REMOVE from analyses).
 
@@
 fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )
+
+byhand <- blst %>% 
+  filter(TAXON_NEW != TAXON_NEW2)
+
+j <- toString(paste0(names(table(byhand$Condition)), ": ",
+                     as.vector (table(byhand$Condition))))
+gen <- byhand[byhand$TAXON_NEW2==c('Salix'),]
+spp <- byhand[byhand$TAXON_NEW2==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
 ```
@@
 Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
-Each of the top ten taxa which were identified by BLAST from the aforementioned process composed `r sum(pp_seqs_t$Samples_Classified)` distinct records (CONFUSING WORDS), of these `r seqs_sp`% of the reads were classified to species `r sp_total_seqs`% which represent of the total classified sequences, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of the classified reads, and `r seqs_fa`% of the records were classified to genus `r fa_total_seqs`% of total sequences.
-
-*Sequences which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were manually assigned to a variety of ranks, occasionally to genus #COUNTS# and family #COUNTS#, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
-* <- went back on this as new function captures most of it, but don't throw out.
+Each of the top ten taxa which were identified by BLAST from the aforementioned process composed `r sum(pp_seqs_t$Samples_Classified)` distinct records, of these `r seqs_sp`% of the reads were classified to species `r sp_total_seqs`% representing of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to genus `r fa_total_seqs`%  of all total sequences.
+
+`r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
+These were all assigned via post-processing conditions (`r j`).
+These were manually assigned to a variety of ranks, occasionally to genus `r nrow(gen)` and species `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
 
@@
 Boraginaceae grains `r toString(aa)`.
-Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`.
+Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
 Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
@@
 ```{r clean sequence numbers from environment}
-rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, 
-    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa)
+rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
+    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
 ```
@@
 In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia*, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
+if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
+The total percentage of the ten most abundant grains per sample were then were then relativize to constitute the entire sample.
 
@@
 This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
-There was strong evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
+There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
 
@@
 e <- blst %>% 
-  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae')) %>% 
+  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
+                       'Celastraceae')) %>% 
   select(sample.id, Genus) %>% 
@@
 To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
-Numerous taxa from the family Ranunculaceae (*Caltha* L.
+Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
 sp., *Thalictrum* L.
@@
 sp., *Aquilegia* L.
-spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
+spp.), had ...
+$\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
 were recorded.
-A similar relationship was observed in the Hydrophyllaceae with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
+A similar relationship was observed in the Hydrophyllaceae R.Br.
+with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
 species were present.
@@
   rename(sample.id = Sample, Genus = GENUS, Family = FAMILY) %>% 
-  distinct(sample.id)
-
-blst <- read.csv('../data/blast_seqs_classified.csv') %>% 
-  mutate(across(where(is.character), ~ replace_na(., 'UNKNOWN'))) %>% 
   arrange(sample.id, -Prcnt_seqs) %>% 
@@
   left_join(., blst_cnt, by = 'species') %>% 
-  group_by(species, name) %>% 
+  group_by(species, TAXON_NEW2) %>% 
   mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
          species = str_remove(species, 'Bombus '))  %>% 
-  distinct(species, name, .keep_all = T) %>% 
-  ungroup(name) %>% 
+  distinct(species, TAXON_NEW2, .keep_all = T) %>% 
+  ungroup(TAXON_NEW2) %>% 
   slice_max(Mu_Seqs, n = 10) %>% 
-  select(plant.species = name, species, n, Mu_Seqs, Genus, Family) %>% 
+  select(plant.species = TAXON_NEW2, species, n, Mu_Seqs, Genus, Family) %>% 
   mutate(rank = 1:n()) %>% 
@@
                          group = Bombus.Species, color = Bombus.Species)) +
-  geom_line(linewidth = 1.25, linetype = 'dotted')+
-  geom_point(size = 3, shape = 23, aes(fill = Bombus.Species)) +
+  geom_line(linewidth = 1.25, linetype = 'dotted', alpha = 0.3)+
+  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
+              height = 0.05, width = 0.05) +
   scale_fill_manual('Bee species', values = bee_cols) +
@@
 
+```{r}
+
+blst_final <- blst %>% 
+  mutate(TAXON_EXPERTS = case_when(
+      TAXON_NEW2 %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
+      'Cynoglossum officinale') ~ 'Mertensia',
+      TAXON_NEW2 %in% c('Caltha leptosepala', 'Trollius europaeus', 'Trollius') ~ 'Delphinium',
+    # 'Thalictrum thalictroides' ~ 'Delphinium',
+      TAXON_NEW2 == 'Erigeron' ~ 'Erigeron grandiflorus', 
+      TAXON_NEW2 == 'Scabrethia scabra' ~ 'Wyethia arizonica',
+      TAXON_NEW2 == 'Nemophila breviflora' ~ 'Hydrophyllum',
+      TAXON_NEW2 == 'Erigeron grandiflorus' ~ 'Erigeron grandiflorus',
+      TAXON_NEW2 == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
+      TAXON_NEW2 == 'Paxistima myrsinites' ~ 'Parnassia palustris', # original was probably right !!! shows the consequence of bad time estimates.#, 
+    # TRUE ~ as.character(TAXON_NEW2)
+  )) %>% 
+  select(TAXON, TAXON_NEW2, TAXON_EXPERTS)
+
+# no TARAXACUM OFICINALE, shows a limitation of the data sources.
+
+```
+
+Situations where SDM's led to incorrect results at the species level are evident with classification to *Scabrethia Scabra* (Hooker) W.A.
+Weber, this match almost certainly representing *Wyethia arizonica* A.
+Gray (@weber1998new), a taxon known to be visited by Queen bee's via our floral observations.
+
 It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes.
@@
 spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads...
-..
-Or the *Epilobium* L.
-spp.
-results likely indicating that a species of *Chamerion* Seg.
-such as *C.
+(IS FORREST PAPER WORTH CITING ?
+IS THIS EVEN WORTH HAVING?)
+
+An expected inaccuracy of the classification scheme is in genus level placements, e.g.
+were *Epilobium* L.
+(Onagraceae Juss.) spp.
+were classified.
+However, given the small size of their flowers in the study area, these results more likely indicate that a species of *Chamaenerion* Seg.
+(a segregate genus) such as *C.
 angustifolium* (L.) Scop.
 or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.
+
+Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed reclassifications of the contents of pollen grains...
+
+<final integrative results here>
 
@@
 
-We anticipate that many of the complications which we faced, using opportunistically collected pollen loads and the first implementation of this method may readily be overcome.
-It seems apparent that we had issues detecting pollen from several genera of plants, based upon these and other observational studies most likely *Vicia* L., *Lathyrus* L., and *Frasera* (@inouye1980effect, @pleasants1980competition); this is most likely related to user error in obtaining high quality DNA during the plant reference library generation period.
-**(REED SHOULD HUNT FOR LOCI RETURNS FROM BLAST LIKE HE DID POLLEN)**.
-Additional complications seem to relate to the presence of closely related false positives, e.g.
-frequent classifications of sequences as *Trollius* L.,  *Caltha* L., and *Thalictrum* L.
-alongside a more common species in the family, e.g.
-*Delphinium.* L.
-Many of our errors are known to us and multiple mnemonics are in *APPENDIX XX* to assist others in future attempts to achieve better results.
-However, the line between false and negative positives may be blurred in some of these instances and warrant further work, for example **Ericaceae pollen grains were observed in a number of samples in trace quantities...**.
-We used only a relatively few amount of species records for both the spatial and temporal modelling, we feel this was an appropriate measure to show that these methods would work with data sets which were readily available without considerable data 'scraping'.
-We anticipate that the number of records available to all researchers have grown rapidly since the inception of our modelling (2019) to date, largely as a result of herbarium digitization efforts, and that speak of a massive 'one stop shop' of an integrated database is just over the horizon (@feng2022review, @soltis2017digitization).
   The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area.
@@
 
-Results from palynological analyses show that several species of bee show near perfect fidelity to the genus Hydrophyllum on a per visit basis.
+Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Hydrophyllum* on a per visit basis...
+General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persit with *Bombus* studies...
 

### 2023-02-01 script/manuscript_MEE.Rmd
added herbaria to acknowledgements
incorporating big batch of introduction edits from JBF

@@
 abstract:  |  
-  1) DNA Barcoding has been remarkably successful in nearly all kingdoms of life and has allowed for the rapid analysis of ecological assemblages.
-Successful DNA barcoding in the plant kingdom has been more difficult than other kingdoms.
-Due to this understanding plants in ecological contexts and understandings of their synecology in some instances may begin to lag behind other kingdoms.
-  2) Here we utilize hyb-seq, museum studies, and species distribution modelling, to detect the plant species present in pollen loads collected from Queen Bumble Bees.  
-  3) By utilizing Species distribution modelling we allow for one to process hyb-seq data, create user specified sequence databases which may use MORE ACCURATE alignment algorithms on personal computers over realistic time periods.
-  4) We show that hyb-seq using the Angiosperms 353 probes, which are currently being used in the largest ever plant systematic endeavor, offers significant promise to metagenomic approaches in real world scenarios.
-  5) We conclude that these probes offer promise for the identification of plant tissue in both single sample, and metasample contexts.  
+  1) DNA Barcoding has been successful for the rapid analysis of ecological assemblages.
+Although barcoding in the plant kingdom barcoding has been more difficult than others, and hence may begin to lag behind other kingdoms.
+  2) Here we test the utilization of Angiosperms 353 probes to barcode plant species found in pollen loads collected from Queen Bumble Bees.  
+  3) To verify the accuracy for this barcoding system we compared the data to museum species, observation studies, and species distribution modelling to identify likely candidate species. 
+  4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which may use increase the alignment algorithms minimizing need for large computational power, and run time.
+  5) We show that the Angiosperms 353 probes, which are currently being used in the largest ever plant systematic endeavor, offers significant promise to metagenomic approaches.    
+  6) Understanding plants in ecological contexts and understandings of their synecology.  
 keywords: |
@@
 
-The inability to reliably identify plants down to terminal taxon can limit our understanding of ecosystem function and interactions (@bortolus2008error).
-This is especially true for genera where many species are defined upon ecological and behavioral rather than morphological properties, and hence often serve as bioindicators of habitat (e.g.
-different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).
-The lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and the interactions they have with other species.
-Current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, assistance from taxonomic specialists, or the use of barcoding or other molecular techniques (CITE).
-The identification of organisms to terminal taxon is often mired by lack of diagnostic characters (e.g.
-flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and increasingly the description of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
-And revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often does not work.
-
-  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
-With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas with many other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
-Particular challenges with the high copy number barcodes (e.g.
-ITS2, *rbcL*, *matK*, *trnH-psbA*) include the utilization their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
-
-  Currently the largest plant systematic endeavor ever undertaken, the Plant and Fungal Tree of Life (PAFTOL) undertaken by the Royal Botanic Gardens Kew, is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom using the popular Angiosperms353 (A353) probes, at 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal), resulting in over 14,000 represented species.
-These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding, and the A353 probes are currently being used in many other plant phylogenetic issues increasing the sampling depth of many clades (@baker2021exploring).
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more species upon it’s intended completion, now slated to be by 2030, similar projects which seek to sequence high amounts of genomes in geographic regions e.g.
-the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, are being undertaken which will contribute data sets applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
-These data will promote the ability to apply metabarcoding to resolve a diversity of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
+The inability to reliably identify plants to terminal taxon can limit our understanding of ecosystem function and interactions (@bortolus2008error).
+This is especially true for genera where identification down to species can be useful as specific bioindicators; defining ecological and behavioral properties  (e.g.
+different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).In these instances the lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and their interactions with other species.
+This can be further complicated by the fact that the identification of organisms to terminal taxon is also often mired by lack of diagnostic characters(e.g.
+flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and the presence of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
+Taxonomic verification can also be limited by the fact that revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often does not work.
+The current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, seeking the assistance from taxonomic specialists, or the use of barcoding using molecular techniques (CITE)
+
+  Recently molecular barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), or metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in many taxa (@ruppert2019past).
+For plants the success is a little more mix, with the identification of certain clades using barcoding being quite successful (@kress2017plant), however for many other clades the results have been more elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), while metabarcoding has incurred additional challenges for the currently available barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+Particular challenges for the utilization of the high copy number barcodes (e.g.
+ITS2, *rbcL*, *matK*, *trnH-psbA*) include their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
+
+  Currently the largest plant systematic endeavor ever undertaken,by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).
+This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom,14,000 represented species, using the popular Angiosperms353 (A353) probes, which includes 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal).
+These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
+The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more species upon it’s intended completion, slated for 2030.
+Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high amounts of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
+These data will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
 However, the application of metabarcoding still face challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
+
   Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
 
-  To increase the quality of metabarcoding results in plants, we suggest reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the the region of study and its ecological characteristics.
-To achieve this goal, we first create a list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
+  To increase the quality of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest.
+To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
 iNaturalist), from a region exceeding the study area.
-To these candidate species, modelling approaches - such as logistic regression, may be used to identify taxa which warrant further exploration e.g.
-modelling to determine their possibility of presence in metabarcoding samples.
+For these candidate species, a modelling approach, such as logistic regression, may be used to identify taxa which warrant further exploration e.g.
+determine their possibility of presence in metabarcoding samples.
 We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
-Species distribution models (SDM’s) examine the ecological conditions associated with known occurrence of a species to identify where else in the study area might suitable habitats be found.
-This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
-This approach can also significantly reduce processing time, particularly as as most next-generation sequence data is deposited as raw-sequence reads.
-
-  Considerable amounts of species interactions vary both in space and time (@caradonna2021seeing).
-In many regions of the world, contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
+Species distribution models (SDM’s) examine the ecological conditions associated with known occurrence of a species to identify where in the study area might suitable habitats.
+This approach has the benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
+This approach can significantly reduce processing time, particularly as as most next-generation sequence data is deposited as raw-sequence reads.
+
+  As species interactions vary both in space and time (@~THOMSON1995 MOSIAC?, @caradonna2021seeing).
+Contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
 In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks,  which reduces the utility of these to operate as filters in the post-processing of sequence matches.
@@
 
-  We test these metagenomic and informatics approaches to determine whether the foraging record of Queen Bumble Bee's are consistent across direct observations and the pollen record, an incongruency noted in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-The assessment of the plant species compositions of pollen is a desired result, with several applications, and numerous complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
-The two foraging phases of the Queen Bumble Bee life cycle is essential to 1) increase their weight before diapause, 2) increase their ovary weights while establishing their recently found nests, both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble, ...).
-Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline) often represent the most diverse areas in the temperate and oftentimes offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
+  We tested this metagenomic and informatics approach to determine to examine the foraging behaviour of queen bumblebee, and compare this approach with direct observations and the pollen record, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+The assessment of the plant species compositions from pollen is a desirable approach, with several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
+The two foraging phases of the queen bumblebee life cycle are essential to 1) increase their weight before diapause, and 2) increase their ovary weights while establishing their recently found nests.
+Both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble, ...).
+Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline).
+These areas often represent the most diverse areas in the temperate and often offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
 An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be incorporated into their current diversity (@loarie2009velocity, @dobrowski2016climate).
@@
 
-Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information).
+Observations and bee sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information), characterized by high-montane/subalpine Parkland vegetation communities.
 Pollinator observations of *Bombus* Latreille spp.
-(Apidae Latreille) were conducted from June - August of 2015 in six study sites characterized by high-montane/subalpine Parkland vegetation communities.
+(Apidae Latreille) were conducted from June - August of 2015 in six study sites.
 Observations of *Bombus* foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and placed into abundance bins.
@@
 
+```{r}
+a <- read.csv('../data/herbaria_4_acknowledgements.csv') %>% 
+  pull() 
+```
+
 **ACKNOWLEDGMENTS:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
 Pat Herendeen for assistance with virtually all aspects of preparing pollen vouchers and the identification process.
+Hilary Noble, Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
 Ian Breckheimer for sharing the SDM predictor variables.
-Hilary Noble, Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
 We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles (Rick) Williams at Ray J.
@@
 The Program in Plant Biology and Conservation is thanked for funding.
+The holdings of the following herbaria were essential for this project: `r toString(a)`.
 
@@
 Layer   LM                     Description                              Source
-----   ----    ------------------------------------------     ---------------------
-1.                   Mean annual cloudiness - MODIS            Wilson et al. 2016
-2.                 Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
-3.                 Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
-4.                 Cloudiness seasonality 3 - MODIS            Wilson et al. 2016
-5.                Beginning of the frost-free period               Wang et al.
-6.                   Climatic moisture deficit                     Wang et al.
-7.                   Degree-days above 5C from                     Wang et al.
-8.                   Mean annual precipitation                     Wang et al.
-9.                Mean annual precipitation as snow                Wang et al.
-10.                   Temperature seasonality                      Wang et al.
-11.             2015 Percent Grass/Herbaceous cover - MODIS         (MOD44B)
-12.            2015 Percent Tree cover from Landsat 7/8              (GLCF)
-13.             Soil probability of bedrock (R Horizon)             SoilGrids
-14.               Soil organic carbon (Tonnes / ha)                 SoilGrids
-15.                    Surface soil pH in H~2~O                     SoilGrids
-16.                   Surface soil percent sand                     SoilGrids
-17.                         Soil USDA class                         SoilGrids
-18.                       Topographic elevation                    EarthEnv DEM
-19.                 Topographic elevation, moving window.          EarthEnv DEM
-20.                   Topographic percent slope                    EarthEnv DEM 
-21.                   Topographic wetness index                    EarthEnv DEM 
-22.                    Topographic aspect from                     EarthEnv DEM
-23.             Annual potential solar radiation computed             r.sun
-24.            Estimated actual (w/-cloud) solar radiation  r.sun / Wilson et al. 2016
-25.             Log-transformed distance to surface water   Global Surface Water Explorer
-26.                   Percent surface water                 Global Surface Water Explorer 
-----        -------------------------------------------      ---------------------
+----   ----      ------------------------------------------    -------------------------
+1.      N              Mean annual cloudiness - MODIS              Wilson et al. 2016
+2.      Y            Cloudiness seasonality 1 - MODIS              Wilson et al. 2016 
+3.      N            Cloudiness seasonality 2 - MODIS              Wilson et al. 2016 
+4.      Y            Cloudiness seasonality 3 - MODIS              Wilson et al. 2016
+5.      N           Beginning of the frost-free period                 Wang et al.
+6.      N              Climatic moisture deficit                       Wang et al.
+7.      N              Degree-days above 5C from                       Wang et al.
+8.      N              Mean annual precipitation                       Wang et al.
+9.      Y           Mean annual precipitation as snow                  Wang et al.
+10.     Y               Temperature seasonality                        Wang et al.
+11.     Y         2015 Percent Grass/Herbaceous cover - MODIS           (MOD44B)
+12.     Y          2015 Percent Tree cover from Landsat 7/8              (GLCF)
+13.     Y         Soil probability of bedrock (R Horizon)               SoilGrids
+14.     N           Soil organic carbon (Tonnes / ha)                   SoilGrids
+15.     N                Surface soil pH in H~2~O                       SoilGrids
+16.     Y               Surface soil percent sand                       SoilGrids
+17.     Y                     Soil USDA class                           SoilGrids
+18.     N                   Topographic elevation                      EarthEnv DEM
+19.     Y             Topographic elevation, moving window.            EarthEnv DEM
+20.     Y               Topographic percent slope                      EarthEnv DEM 
+21.     Y               Topographic wetness index                      EarthEnv DEM 
+22.     Y                  Topographic aspect                          EarthEnv DEM
+23.     Y         Annual potential solar radiation computed               r.sun
+24.     N        Estimated actual (w/-cloud) solar radiation      r.sun / Wilson et al. 2016
+25.     Y        Log-transformed distance to surface water      Global Surface Water Explorer
+26.     Y                Percent surface water                  Global Surface Water Explorer 
+----   ----     -------------------------------------------    ------------------------------
 
@@
 
-```{r READS PER LOCI, echo = F, out.width = '98%'}
+```{r READS PER LOCI - REFERENCE, echo = F, out.width = '98%'}
+knitr::include_graphics('../graphics/plots/Loci_Reference_Samples.png')
+```
+
+```{r READS PER LOCI - METAGENOMICS, echo = F, out.width = '98%'}
 knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
@@
 
-*Generalised Linear Models (GLM)*
-
-*Generalised Additive Models (GAM)*
-
 The two machine learning models utilize Ensemble learning.
 
-Decision trees, ...
-
-Ensemble learning utilizes many sets of trees, each tree being composed of many binary decisions, to create a single model.
+__Ensemble learning__ utilizes many sets of trees, each tree being composed of many binary decisions, to create a single model.
 Each independent variable ( - or *feature*) may become a node on the tree - i.e.
@@
 
-*Random Forest (RF)* the training data are continually bootstrap re-sampled, in combination with random subsets of features, to create nodes which attempt to optimally predict a known outcome.
+*__Random Forest (RF)__* the training data are continually bootstrap re-sampled, in combination with random subsets of features, to create nodes which attempt to optimally predict a known outcome.
 A large number of trees are then aggregated, via the most common predictions, to generate a final classification prediction tree.
@@
 
-*Boosted Regression Tree (BRT)* (or Gradient Boosted tree) An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
+*__Boosted Regression Tree (BRT)__* (or Gradient Boosted tree) An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
 All response data and predictor variables are kept available to all trees.
 
-Random Forest have high bias and low variance, where boosted regressions trees have low bias and high variances.
+*__Bias__* predictions from an algorithm are systematically in error due to being prejudiced for or against certain results, due to assumptions during learning.
+
+*__Variance__* errors in models due to an over-reliance and sensitivity of training to outliers in training data.
+
+In general, Random Forest  models have high bias and low variance, where boosted regressions trees have lower bias and higher variance.
+Theoretically, the weaknesses and strengths of bootstrap aggregation (bagging) as implemented by Random Forests are supplemented by the boosting.
+
+\newpage
+
+APPENDIX XX - Time Spent Generating Species Distribution Models
+
+```{r}
+knitr::include_graphics('../graphics/assorted/individual_model_runs.pdf')
+```
 

### 2023-02-02 script/manuscript_MEE.Rmd
incorporated most of jbf positional edits

@@
 
-## Study System & Field Work
+## 2.1 Study System & Field Work
 
@@
 
-### 2.1 | Spatial Analyses
-
-#### 2.1.1 Candidate Species
-
-We downloaded from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022) all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area .
-We then generated Species Distribution Models (SDMs) to predict their distribution throughout the study area.
-These maps of potential distribution served as a reference to reduce the list of species to include in the genomic sequence databases.
-
-In order to minimise the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
-In order to emulate the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
+## 2.2 | Floral Visitation
+
+...
+
+## 2.3 | Pollen Morphological identification
+
+### 2.3.1 | Pollen Reference Library
+
+To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
+Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
+  We used Divisive Hierarchical Clustering techniques to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
+Ten readily discernible categorical traits were collected from each specimen in the image collection.
+These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
+Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
+This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
+
+### 2.3.2 | Pollen Corbiculae Loads
+
+  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
+*Cerca* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
+The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with nail polish to prevent oxidation; all samples are noted in *APPENDIX 3*.
+  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
+In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
+Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
+Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
+
+### 2.4 | Molecular Barcoding
+
+#### 2.4.1 | Species reference list
+
+#### 2.4.1.1 Spatial Analyses
+
+We generate a short list of potential candidate species we downloaded from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022) all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
+To reduce the list of species to include in the genomic sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.
+
+In order to minimize the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
+To account for the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
 The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
 
-#### 2.1.2 Distribution Modelling
+#### 2.4.1.2 Distribution Modelling
 
@@
 These records were copied into two, initially identical, sets, one for generating machine learning models (ML; Random Forest, and Boosted Regression Tree’s), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM) (@barbet2012selecting).
-**Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble).
+Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble).
 No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with differing biologies and ecologies (@elith2006novel, @qiao2015no).
-In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).**
+In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).
 
@@
 
-535 species were modelled using Generalized Linear Models and Generalized Additive Models.
-534 species were modelled using Random Forest and Boosted Regression Trees.
+A total of 535 species were modelled using Generalized Linear Models and Generalized Additive Models and 534 species were modelled using Random Forest and Boosted Regression Trees.
 To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (@gbifDL2021sdms, @aim2019database).
@@
 
-## 2.2 | Molecular Lab Work
+### 2.4.2 | Temporal Analyses
+
+To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
+Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not be found in the nearest 5), and which were from herbarium records were included.
+To remove temporally irrelevant herbarium records, i.e.
+material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
+Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
+Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
+
+### 2.5.2 | Barcode references library
 
@@
 
-#### 2.2.1 | Reference Plant Library Generation
+#### 2.5.2.1 | Sampling Species for Barcoding
 
 Using five years (2015-2020) of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
-We sequenced the 12 most commonly visited taxa twice using samples from one site within the Gunnison Basin River Drainage and one individual from another more distal population.
-In addition, for any of these 12 focal species which did not have a congener pair in this filtered sample, we included a congener - or a species from a closely related genus to serve as an outgroup.
-We also sequenced another 15 abundant taxa commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set (*APPENDIX 4*).
-Plant collections were identified via a variety, and typically a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
-
-#### 2.2.2 | Plant Genomic DNA Extraction
+We sequenced the 12 most commonly visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
+In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
+In addition we sequenced another 15 taxa commonly visited by Bombus workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (*APPENDIX 4*).
+Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
+
+#### 2.5.2.2 | Plant Genomic DNA Extraction
 
@@
 
-#### 2.2.3 | Pollen Genomic DNA Extraction
+#### 2.5.2.3 | Pollen Genomic DNA Extraction
 
@@
 
-#### 2.2.4 | Fragmentation, Library Preparation & Target Enrichment
+#### 2.5.2.4 | Fragmentation, Library Preparation & Target Enrichment
 
@@
 Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
-Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed with ½ volume of SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v.
+Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed using SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v.
 ethanol washes.
@@
 
-###  2.2.5 | Computational Processes and Analyses. 
-####  2.2.5.1 | Reference Library Data Processing
+###  2.6 | Computational Processes and Analyses. 
+####  2.6.1 | Reference Library Data Processing
 
 Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950).
-Contigs were generated using HybPiper using target files created by M353 (@johnson2016hybpiper, @mclay2021new).
-
-#### 2.2.5.2 | Sequence Identification
-
-A custom Kraken2 database was created by downloading representative species of each genus indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
+Contigs generated were mapped to a reference with HybPiper with using target files created by M353 (@johnson2016hybpiper, @mclay2021new).
+
+#### 2.6.2 | Sequence Identification
+
+A custom Kraken2 database was created by downloading representative species indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
 These sequences were processed in the same manner as our novel sequences.
@@
 
-#### 2.2.5.3 | Identification of Sequence Matching Loci
+#### 2.6.3 | Identification of Sequence Matching Loci
 
@@
 
-## 2.2.5.4 | Morphological Pollen identification
-
-To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  We used Divisive Hierarchical Clustering techniques to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
-Ten readily discernible categorical traits were collected from each specimen in the image collection.
-These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
-This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
-  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Cerca* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
-The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with nail polish to prevent oxidation; all samples are noted in *APPENDIX 3*.
-  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
-
-### 2.3 | Temporal Analyses
-
-To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not be found in the nearest 5), and which were from herbarium records were included.
-To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
-Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
-Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
-
-### 2.4 | Floral Observations
-
 # 3 | RESULTS
 
-## 3.1 | Spatial Analyses
-
-```{r Results for SA Modelling}
-
-files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
-ml <- files[grep( '_ml_', files)]
-ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
-linear <- files[grep('_glm_', files)]
-linear <- unique(sub('_glm.*$', '', linear))
-
-# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
-# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
-
-rm(files, ml, linear)
-```
-
-```{r Results for SA Logistic Regression Metrics}
-
-logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
-  filter(occurrence == 0) 
-# summary(logR_dat$n)
-# sum(logR_dat$n)
-
-dat <- read.csv('../data/evaluation_table.csv')[,2:3]
-data.frame(Metric  = dat[1:5,1],
-                   Value   = dat[1:5,2],
-                   Metric1 = dat[6:10,1],
-                   Value1  = dat[6:10,2]
-                   ) %>% 
-  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
-    col.names = c('Metric', 'Value', 'Metric', 'Value'),
-    booktabs = T) #%>% 
-  #kableExtra::kable_styling(position = "float_right")
-```
-
-```{r Results for SA Logistic Regression Contingency table}
-read.csv('../data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
-    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
-    booktabs = T) %>%
-    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
-
-rm(logR_dat, dat)
-```
-
-The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
-Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
-
-At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
-
-## 3.2 | Microscopic Pollen identification
-
-```{r compare microscopic to molecular Pollen ID, eval = T}
-
-# are mertensia and Asteraceae 1 & 2 detected in all Pollen samples where the bee was COLLECTED ON THEM ?
-
-pmorph <- read.csv('../data/morpho_pollen.csv') 
-blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
-  rename(sample.id = Sample, Family = FAMILY, Genus = GENUS) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
-  ungroup()
-blst_seqs <- distinct(blst, sample.id)
-
-pmorph <- filter(pmorph, sample.id %in% blst$sample.id) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prop = sum/sum(sum))
-
-mertensia_samples <- pmorph %>% 
-  filter(morphotype == 'MERTENSIA') %>% 
-  select(sample.id, morphotype, Prop, sum)
-
-mertensia_samples <- filter(mertensia_samples, sample.id %in% blst_seqs$sample.id)
-
-blst_mertensia <- blst %>% 
-  filter(Family == 'Boraginaceae') %>% 
-  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_seqs = sum(Prcnt_seqs),
-         morphotype = 'MERTENSIA') %>% 
-  distinct(sample.id, .keep_all = T)
-
-mertensia_detected <- mertensia_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0)) 
- # sample 26 - very strange !!!! this needs investigation - ALSO THE ERICOID IS LIKELY ONAGRAD !?
-
-blst_aster <- blst %>% 
-  filter(Family == 'Asteraceae') %>% 
-  group_by(sample.id) %>% 
-  distinct(sample.id, .keep_all = T)
-
-# PERFORM FOR ERICACEAE TOO - A single record maybe come out?
-ericaceae_samples <- pmorph %>% 
-  filter(morphotype == 'ERICACEAE') %>% 
-  select(sample.id, Prop, morphotype)
-
-blst_ericoid <- filter(blst, Family == 'Ericaceae') %>% # 44 might be Epilobium anagallidolium
-  mutate(morphotype = 'ERICACEAE')
-ericaceae_detected <- ericaceae_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_ericoid$sample.id, 1, 0))
-
-# ASTER 2 HERE...
-aster2_samples <- pmorph %>% 
-  filter(morphotype == 'ASTERACEAE_2')
-
-blst_aster2 <- filter(blst_aster, Genus %in% 
-                            c('Taraxacum', 'Helianthella', 'Scabrethia')) %>% 
-  mutate(morphotype = 'ASTERACEAE_2')
-aster2_detected <- aster2_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster2$sample.id, 1, 0))
-
-# ASTER 1 HERE... ? WERE ALL OF THESE SPECIES PRUNED... ?
-aster1_samples <- pmorph %>% 
-  filter(morphotype == 'ASTERACEAE_1')
-
-blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 
-                                                'Scabrethia') ) %>% 
-  mutate(morphotype = 'ASTERACEAE_1')
-aster1_detected <- aster1_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster1$sample.id, 1, 0))
-
-# ALL HIGH NUMBER ASTERACEAE
-
-aster_samples <- pmorph %>% 
-  filter(str_detect(morphotype, 'ASTER.*'))
-aster_detected <- aster_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster$sample.id, 1, 0))
-
-aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
-bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
-cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
-dd <- prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
-ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)
-ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)
-
-# Semi-quantitative accord between sequence % and Grain Count
-palyn_cnt <- bind_rows(aster1_samples, aster2_samples, ericaceae_samples, mertensia_samples) %>% 
-  select(morphotype, countProp = Prop)
-
-pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia, blst_ericoid) %>% 
-  select(morphotype, sequenceProp = Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
-  left_join(., palyn_cnt, by = c('sample.id', 'morphotype')) %>% 
-  drop_na()
-
-rm(mertensia_samples, blst_aster, blst_aster1,
-   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
-   blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
-```
-
-Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into.
-*APPENDIX 7*.
-60 samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
-The number of counted pollen grains in each sample range from (*MIN* - 16,293, $\bar{x}$ = 2788.685, Mdn = 1453).
-
-```{r}
-rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
-```
-
-```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
-The innermost ring indicates every genus which Queen Bee's were observed to visit.
-The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
-The outermost ring indicates that sequence data were available for at least a member of that genus.
-Branch colors follow APG 4."}
-knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
-```
-
-## 3.3 | Metabarcoding Pollen identification
-
-```{sh Get number of raw and trimmed reads for metagenome samples, eval = F}
-
-cd ~/sequenceData/Original_zip
-files=$(find -type f -name "[0-9]_S[0-9]*_paired.fastq.gz") # doesnt support {2}
-files1=$(find -type f -name '[0-9][0-9]_S[0-9]*_paired.fastq.gz')
-files="${files} ${files1}"
-echo $files
-# seqs per each sample , need divide by 4 on import
-wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
-
-cd ../trimmed_metagenomic
-files=$(find  -maxdepth 1 -name '*gz')
-wc -l  $files  > '/hdd/yun master/MEE_manuscript/data/trimmed_reads_metagenome_seq.txt'
-```
-
-```{r Summary statistics of sequence numbers}
-
-raw_seqs <- read.delim('../data/raw_reads_metagenome_seq.txt', header = F) %>% 
-  mutate(V1 = str_trim(V1)) %>% 
-  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
-  mutate(seq = round(as.numeric(seq)/4),
-         sample = str_extract(sampleFR, '.*L002_')) %>% 
-  filter(sampleFR != 'total') %>% 
-  group_by(sample) %>% 
-  mutate(totalReads = sum(seq)) %>% 
-  distinct(totalReads) %>% 
-  pull(totalReads)
-
-trimmed_seqs <- read.delim('../data/trimmed_reads_metagenome_seq.txt', header = F) %>% 
-  mutate(V1 = str_trim(V1)) %>% 
-  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
-  mutate(seq = round(as.numeric(seq)/4),
-         sample = str_extract(sampleFR, '.*paired')) %>% 
-  filter(sampleFR != 'total') %>% 
-  group_by(sample) %>% 
-  mutate(totalReads = sum(seq)) %>% 
-  distinct(totalReads) %>% 
-  pull(totalReads)
-
-krak <- read.csv('../data/kraken_proc_reads.csv')
-brack <- read.csv('../data/bracken_proc_reads.csv')
-blast <- read.csv('../data/blast_proc_reads.csv')
-loci <- read.csv('../data/reads_per_loci.csv') %>% group_by(Sample)
-
-no_loci <- loci %>% group_by(Sample) %>% count(name = 'no_loci')
-no_reads <- loci %>% group_by(loci) %>% summarize(readsLoci = sum(n))
-```
-
-54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
-The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
-Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
-The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`.
-**APPENDIX X Reads Per Loci**.
-
-... with samples 22, and 41 both having < 100 reads indicates, virtual failure of these records (REMOVE from analyses).
-
-After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
-`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and it's values were used for all subsequent analyses.
-
-```{r Read Post Processing}
-
-pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') )
-
-knitr::kable(pp_seqs_t, align = "c", 
-             col.names = c('Condition', 'No. Class.', 'Prcnt. Class.', 'Total Seqs', 'Rank'),
-             caption = "Post classification of Sequences via Taxonomy and Ecology") %>% 
-  kableExtra::kable_styling(
-      latex_options = "striped",
-      stripe_color = "gray!10")
-
-seqs_sp <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records'] )
-seqs_ge <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records'] )
-seqs_fa <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records'] )
-
-sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs'] )
-sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs'] )
-fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )
-
-byhand <- blst %>% 
-  filter(TAXON_NEW != TAXON_NEW2)
-
-j <- toString(paste0(names(table(byhand$Condition)), ": ",
-                     as.vector (table(byhand$Condition))))
-gen <- byhand[byhand$TAXON_NEW2==c('Salix'),]
-spp <- byhand[byhand$TAXON_NEW2==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
-```
-
-The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
-A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix XX).
-Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
-Each of the top ten taxa which were identified by BLAST from the aforementioned process composed `r sum(pp_seqs_t$Samples_Classified)` distinct records, of these `r seqs_sp`% of the reads were classified to species `r sp_total_seqs`% representing of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to genus `r fa_total_seqs`%  of all total sequences.
-
-`r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were all assigned via post-processing conditions (`r j`).
-These were manually assigned to a variety of ranks, occasionally to genus `r nrow(gen)` and species `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
-
-To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
-The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae.
-Boraginaceae grains `r toString(aa)`.
-Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
-
-```{r clean sequence numbers from environment}
-rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
-    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
-```
-
-```{r Correlation between Number of Grains and Number of Sequences}
-
-pol_p <- pollen_props %>% 
-  mutate(sequenceProp = sequenceProp/100, 
-         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
-
-ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
-# ct[["p.value"]] too small to write out. tau > 0.4 strong relationship
-model <-  lm(adjSeqProp ~ countProp, data = pol_p)
-
-par(pty="s")
-plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-      xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-     xlab = 'Grains Counted', ylab = 'Classified Sequences',
-     pch=18, slty = 3, gap = 0.025)
-axis(1)
-axis(2)
-abline(model)
-text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
-text(0.3, 0.8, "p < 0.001")
-
-rm( newdat)
-```
-
-To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
-In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then relativize to constitute the entire sample.
-
-The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
-This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
-There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
-
-```{r}
-rm(pol_p, ct)
-```
-
-```{r Co-occurrence of splash drop sequences, eval = F}
-
-e <- blst %>% 
-  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
-                       'Celastraceae')) %>% 
-  select(sample.id, Genus) %>% 
-  mutate(Present = 1) %>% 
-  group_by(Genus) %>% 
-  distinct() %>% 
-  pivot_wider(names_from = Genus, values_from = Present) %>% 
-  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
-  column_to_rownames('sample.id')
-
-myout <- CooccurrenceAffinity::affinity(data = e, row.or.col = "col", squarematrix = c("all"))
-CooccurrenceAffinity::plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
-```
-
-To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
-Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
-sp., *Thalictrum* L.
-spp., *Trollius* L.
-sp., *Aquilegia* L.
-spp.), had ...
-$\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
-were recorded.
-A similar relationship was observed in the Hydrophyllaceae R.Br.
-with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
-species were present.
-
-## 3.4 | Temporal Analyses
-
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
-
-Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
-Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
-There was very strong evidence that the weibull estimates were positively associated with the observed onset (r^2^ = 0.72, p < 0.0001, tau = 0.61) and peak (r^2^ = 0.70, p < 0.0001, tau = 0.65) of flowering, and that the number of herbarium samples had a moderate effect on the estimates (p = 0.004 and p = 0.034 respectively).
-There was very strong evidence that the weibull estimates had a positive association with the observed cessation of flowering (r^2^ =  0.4339,  p < 0.0001, tau = 0.489), however their was no evidence that sample size had an effect (p = 0.349).
-There was moderate evidence that the weibull estimates, with an effect of sample size, had a weak positive association with the observed duration of flowering (p = 0.0401, r^2^ = 0.07, tau = 0.17).
-
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
-# these are the tenth and nintey  percent non-parametric confidence intervals. 
-```
-
-## 3.5 | Floral Observations
+## 3.1 | Floral Observations
 
@@
 ```
+
+**NEED (AT LEAST) TABLE OF TOP VISITED PLANTS**
 
@@
 
+## 3.1 | Spatial Analyses
+
+```{r Results for SA Modelling}
+
+files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
+ml <- files[grep( '_ml_', files)]
+ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
+linear <- files[grep('_glm_', files)]
+linear <- unique(sub('_glm.*$', '', linear))
+
+# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
+# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
+
+rm(files, ml, linear)
+```
+
+```{r Results for SA Logistic Regression Metrics}
+
+logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
+  filter(occurrence == 0) 
+# summary(logR_dat$n)
+# sum(logR_dat$n)
+
+dat <- read.csv('../data/evaluation_table.csv')[,2:3]
+data.frame(Metric  = dat[1:5,1],
+                   Value   = dat[1:5,2],
+                   Metric1 = dat[6:10,1],
+                   Value1  = dat[6:10,2]
+                   ) %>% 
+  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
+    col.names = c('Metric', 'Value', 'Metric', 'Value'),
+    booktabs = T) #%>% 
+  #kableExtra::kable_styling(position = "float_right")
+```
+
+```{r Results for SA Logistic Regression Contingency table}
+read.csv('../data/Pa_conting_table.csv') %>% 
+    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
+    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
+    booktabs = T) %>%
+    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
+
+rm(logR_dat, dat)
+```
+
+## 3.2 | Microscopic Pollen identification
+
+```{r compare microscopic to molecular Pollen ID, eval = T}
+
+# are mertensia and Asteraceae 1 & 2 detected in all Pollen samples where the bee was COLLECTED ON THEM ?
+
+pmorph <- read.csv('../data/morpho_pollen.csv') 
+blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  rename(sample.id = Sample, Family = FAMILY, Genus = GENUS) %>% 
+  group_by(sample.id) %>% 
+  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
+  ungroup()
+blst_seqs <- distinct(blst, sample.id)
+
+pmorph <- filter(pmorph, sample.id %in% blst$sample.id) %>% 
+  group_by(sample.id) %>% 
+  mutate(Prop = sum/sum(sum))
+
+mertensia_samples <- pmorph %>% 
+  filter(morphotype == 'MERTENSIA') %>% 
+  select(sample.id, morphotype, Prop, sum)
+
+mertensia_samples <- filter(mertensia_samples, sample.id %in% blst_seqs$sample.id)
+
+blst_mertensia <- blst %>% 
+  filter(Family == 'Boraginaceae') %>% 
+  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
+  group_by(sample.id) %>% 
+  mutate(Prcnt_seqs = sum(Prcnt_seqs),
+         morphotype = 'MERTENSIA') %>% 
+  distinct(sample.id, .keep_all = T)
+
+mertensia_detected <- mertensia_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0)) 
+ # sample 26 - very strange !!!! this needs investigation - ALSO THE ERICOID IS LIKELY ONAGRAD !?
+
+blst_aster <- blst %>% 
+  filter(Family == 'Asteraceae') %>% 
+  group_by(sample.id) %>% 
+  distinct(sample.id, .keep_all = T)
+
+# PERFORM FOR ERICACEAE TOO - A single record maybe come out?
+ericaceae_samples <- pmorph %>% 
+  filter(morphotype == 'ERICACEAE') %>% 
+  select(sample.id, Prop, morphotype)
+
+blst_ericoid <- filter(blst, Family == 'Ericaceae') %>% # 44 might be Epilobium anagallidolium
+  mutate(morphotype = 'ERICACEAE')
+ericaceae_detected <- ericaceae_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_ericoid$sample.id, 1, 0))
+
+# ASTER 2 HERE...
+aster2_samples <- pmorph %>% 
+  filter(morphotype == 'ASTERACEAE_2')
+
+blst_aster2 <- filter(blst_aster, Genus %in% 
+                            c('Taraxacum', 'Helianthella', 'Scabrethia')) %>% 
+  mutate(morphotype = 'ASTERACEAE_2')
+aster2_detected <- aster2_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_aster2$sample.id, 1, 0))
+
+# ASTER 1 HERE... ? WERE ALL OF THESE SPECIES PRUNED... ?
+aster1_samples <- pmorph %>% 
+  filter(morphotype == 'ASTERACEAE_1')
+
+blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 
+                                                'Scabrethia') ) %>% 
+  mutate(morphotype = 'ASTERACEAE_1')
+aster1_detected <- aster1_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_aster1$sample.id, 1, 0))
+
+# ALL HIGH NUMBER ASTERACEAE
+
+aster_samples <- pmorph %>% 
+  filter(str_detect(morphotype, 'ASTER.*'))
+aster_detected <- aster_samples %>% 
+  mutate(DETECTED = if_else(sample.id %in% blst_aster$sample.id, 1, 0))
+
+aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
+bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
+cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
+dd <- prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
+ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)
+ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)
+
+# Semi-quantitative accord between sequence % and Grain Count
+palyn_cnt <- bind_rows(aster1_samples, aster2_samples, ericaceae_samples, mertensia_samples) %>% 
+  select(morphotype, countProp = Prop)
+
+pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia, blst_ericoid) %>% 
+  select(morphotype, sequenceProp = Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
+  left_join(., palyn_cnt, by = c('sample.id', 'morphotype')) %>% 
+  drop_na()
+
+rm(mertensia_samples, blst_aster, blst_aster1,
+   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
+   blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
+```
+
+Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into.
+*APPENDIX 7*.
+From the 60 samples that were counted and based on rarefaction we **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
+The number of counted pollen grains in each sample range from (*MIN* - 16,293, $\bar{x}$ = 2788.685, Mdn = 1453).
+
+**SUMMARY REQUIRED - Number of species from number of families, and how many species are identified to species versus number only to genera **
+
+```{r}
+rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
+```
+
+```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
+The innermost ring indicates every genus which Queen Bee's were observed to visit.
+The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
+The outermost ring indicates that sequence data were available for at least a member of that genus.
+Branch colors follow APG 4."}
+knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
+```
+
+### 3.3 | Metabarcoding Pollen Indentification
+
+### 3.3.1 | Spatial Analyses to identify candidate taxa
+
+The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+
+In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
+Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
+Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
+
+At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
+
+### 3.3.2 | Temporal Analysis
+
+The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
+
+Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
+Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
+There was very strong evidence that the weibull estimates were positively associated with the observed onset (r^2^ = 0.72, p < 0.0001, tau = 0.61) and peak (r^2^ = 0.70, p < 0.0001, tau = 0.65) of flowering, and that the number of herbarium samples had a moderate effect on the estimates (p = 0.004 and p = 0.034 respectively).
+There was very strong evidence that the weibull estimates had a positive association with the observed cessation of flowering (r^2^ =  0.4339,  p < 0.0001, tau = 0.489), however their was no evidence that sample size had an effect (p = 0.349).
+There was moderate evidence that the weibull estimates, with an effect of sample size, had a weak positive association with the observed duration of flowering (p = 0.0401, r^2^ = 0.07, tau = 0.17).
+
+```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '100%'}
+knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
+# these are the tenth and nintey  percent non-parametric confidence intervals. 
+```
+
+```{sh Get number of raw and trimmed reads for metagenome samples, eval = F}
+
+cd ~/sequenceData/Original_zip
+files=$(find -type f -name "[0-9]_S[0-9]*_paired.fastq.gz") # doesnt support {2}
+files1=$(find -type f -name '[0-9][0-9]_S[0-9]*_paired.fastq.gz')
+files="${files} ${files1}"
+echo $files
+# seqs per each sample , need divide by 4 on import
+wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
+
+cd ../trimmed_metagenomic
+files=$(find  -maxdepth 1 -name '*gz')
+wc -l  $files  > '/hdd/yun master/MEE_manuscript/data/trimmed_reads_metagenome_seq.txt'
+```
+
+```{r Summary statistics of sequence numbers}
+
+raw_seqs <- read.delim('../data/raw_reads_metagenome_seq.txt', header = F) %>% 
+  mutate(V1 = str_trim(V1)) %>% 
+  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
+  mutate(seq = round(as.numeric(seq)/4),
+         sample = str_extract(sampleFR, '.*L002_')) %>% 
+  filter(sampleFR != 'total') %>% 
+  group_by(sample) %>% 
+  mutate(totalReads = sum(seq)) %>% 
+  distinct(totalReads) %>% 
+  pull(totalReads)
+
+trimmed_seqs <- read.delim('../data/trimmed_reads_metagenome_seq.txt', header = F) %>% 
+  mutate(V1 = str_trim(V1)) %>% 
+  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
+  mutate(seq = round(as.numeric(seq)/4),
+         sample = str_extract(sampleFR, '.*paired')) %>% 
+  filter(sampleFR != 'total') %>% 
+  group_by(sample) %>% 
+  mutate(totalReads = sum(seq)) %>% 
+  distinct(totalReads) %>% 
+  pull(totalReads)
+
+krak <- read.csv('../data/kraken_proc_reads.csv')
+brack <- read.csv('../data/bracken_proc_reads.csv')
+blast <- read.csv('../data/blast_proc_reads.csv')
+loci <- read.csv('../data/reads_per_loci.csv') %>% group_by(Sample)
+
+no_loci <- loci %>% group_by(Sample) %>% count(name = 'no_loci')
+no_reads <- loci %>% group_by(loci) %>% summarize(readsLoci = sum(n))
+```
+
+## 3.3.1 | Molecular analysis of corbiculae loads
+
+The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
+The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
+Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
+The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`.
+**APPENDIX X Reads Per Loci**.
+
+... with samples 22, and 41 both having < 100 reads indicates, virtual failure of these records (REMOVE from analyses).
+
+After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
+`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
+Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and it's values were used for all subsequent analyses.
+
+```{r Read Post Processing}
+
+pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') )
+
+knitr::kable(pp_seqs_t, align = "c", 
+             col.names = c('Condition', 'No. Class.', 'Prcnt. Class.', 'Total Seqs', 'Rank'),
+             caption = "Post classification of Sequences via Taxonomy and Ecology") %>% 
+  kableExtra::kable_styling(
+      latex_options = "striped",
+      stripe_color = "gray!10")
+
+seqs_sp <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records'] )
+seqs_ge <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records'] )
+seqs_fa <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records'] )
+
+sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs'] )
+sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs'] )
+fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )
+
+byhand <- blst %>% 
+  filter(TAXON_NEW != TAXON_NEW2)
+
+j <- toString(paste0(names(table(byhand$Condition)), ": ",
+                     as.vector (table(byhand$Condition))))
+gen <- byhand[byhand$TAXON_NEW2==c('Salix'),]
+spp <- byhand[byhand$TAXON_NEW2==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
+```
+
+The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
+A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix XX).
+Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
+Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species `r sp_total_seqs`% representing of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family `r fa_total_seqs`%  of all total sequences.
+
+Of the `r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
+These were all assigned via post-processing conditions (`r j`, APPENDIX XX).
+These were manually assigned to a variety of ranks, occasionally to genus - `r nrow(gen)`, and species - `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
+
+To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
+The pollen samples contained three morphotypes which could readily be identified via microscopy.
+Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae.
+Boraginaceae grains `r toString(aa)`.
+Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
+Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
+
+```{r clean sequence numbers from environment}
+rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
+    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
+```
+
+```{r Correlation between Number of Grains and Number of Sequences}
+
+pol_p <- pollen_props %>% 
+  mutate(sequenceProp = sequenceProp/100, 
+         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
+
+ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
+# ct[["p.value"]] too small to write out. tau > 0.4 strong relationship
+model <-  lm(adjSeqProp ~ countProp, data = pol_p)
+
+par(pty="s")
+plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
+      xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
+     xlab = 'Grains Counted', ylab = 'Classified Sequences',
+     pch=18, slty = 3, gap = 0.025)
+axis(1)
+axis(2)
+abline(model)
+text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
+text(0.3, 0.8, "p < 0.001")
+
+rm( newdat)
+```
+
+To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
+In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
+if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
+The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
+
+The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
+This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
+There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
+
+```{r}
+rm(pol_p, ct)
+```
+
+```{r Co-occurrence of splash drop sequences, eval = F}
+
+e <- blst %>% 
+  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
+                       'Celastraceae')) %>% 
+  select(sample.id, Genus) %>% 
+  mutate(Present = 1) %>% 
+  group_by(Genus) %>% 
+  distinct() %>% 
+  pivot_wider(names_from = Genus, values_from = Present) %>% 
+  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
+  column_to_rownames('sample.id')
+
+myout <- CooccurrenceAffinity::affinity(data = e, row.or.col = "col", squarematrix = c("all"))
+CooccurrenceAffinity::plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
+```
+
+To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
+Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
+sp., *Thalictrum* L.
+spp., *Trollius* L.
+sp., *Aquilegia* L.
+spp.), had ...
+$\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
+were recorded.
+A similar relationship was observed in the Hydrophyllaceae R.Br.
+with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
+species were present.
+
 ## 3.6 | Integrated Observational, Molecular, and Palynological Network
@@
 
-For example a common UNKNOWN sequence mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
+For example a many sequences which mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
 nevadensis* Cresson and *B.
-rufocinctus* Cresson pollen is most likely *Frasera* Walter, failed extraction.
+rufocinctus* Cresson pollen is most likely *Frasera* Walter, which failed extractions for the reference library failed (APPENDIX XX).
 A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* (A.
@@
 
+\newpage
+
 Appendix XX - Comparison of Kraken2, Bracken, and BLAST

### 2023-02-03 script/manuscript_MEE.Rmd
modified header a bit

@@
 author:  |
-    |  Reed Clark Benkendorf$^1$$^,$$^2$, Emily J. Woodworth$^1$$^,$$^2$, Jane E. Ogilive$^3$$^,$$^4$,   
+    |  Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
     |  $^4$ Australia post fire project, ----, ----, ----, AU  
-    |  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie$^3$$^,$$^4$ , Emily J. Woodworth$^1$$^,$$^2$  
 abstract:  |  
+  |
   1) DNA Barcoding has been successful for the rapid analysis of ecological assemblages.
@@
 
-  As species interactions vary both in space and time (@~THOMSON1995 MOSIAC?, @caradonna2021seeing).
+  As species interactions vary both in space and time (@~THOMSON1994 GEO?, @caradonna2021seeing).
 Contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
@@
 Pollinator observations of *Bombus* Latreille spp.
-(Apidae Latreille) were conducted from June - August of 2015 in six study sites.
-Observations of *Bombus* foraging took place for one hour at each field site in three 100m transects, where all flowers were also counted and placed into abundance bins.
+(Apidae Latreille) were conducted from May 29^th^ - July 23^rd^ of 2015 in six study sites.
+Observations of *Bombus* foraging took place for one hour at each field site in three 50m transects, each transect was in a major vegetation type (dry, and wet meadows, and Aspen forest), where all flowers abundances were estimated and placed into binds of abundance.
 Corbiculae loads were, non-lethally, collected once from all Queen individuals encountered.
@@
 
-...
+*... does this need more info and warrant a section independent of the above?*
 
@@
 For example, when only a single species of a genus was known in the study area, but our database used a representative of another taxon in the genus, this species was coded as the result.
-The re-coding of sequences from another representative species for the genus to the sole RMBL representative allowed the identification of *XX & %* more species.
 
@@
 ```
-
-**NEED (AT LEAST) TABLE OF TOP VISITED PLANTS**
 
@@
 ggplot(data=correct, aes(x=Observation, y=value,
-                         group = Bombus.Species, color = Bombus.Species)) +
-  geom_line(linewidth = 1.25, linetype = 'dotted', alpha = 0.3)+
+                         group = Bombus.Species, fill = Bombus.Species)) +
+ # geom_line(linewidth = 1.25, linetype = 'dotted', alpha = 0.3)+
   geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
-              height = 0.05, width = 0.05) +
+              height = 0.05, width = 0.3) +
   scale_fill_manual('Bee species', values = bee_cols) +
@@
 
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quantile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quantile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
 Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
@@
   filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
-                       'Celastraceae')) %>% 
+                       'Celastraceae', 'Fabaceae')) %>% 
   select(sample.id, Genus) %>% 
@@
 
-To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, $\alpha$ index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
+To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
 Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
@@
 sp., *Aquilegia* L.
-spp.), had ...
-$\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
+spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
 were recorded.
@@
 
-...
-many of our results indicate foraging on *Viola* L.
-spp, zygomorphic flowers with architecture which would require subtle handling and strength to reach the pollen and nectar loads...
-(IS FORREST PAPER WORTH CITING ?
-IS THIS EVEN WORTH HAVING?)
-
 An expected inaccuracy of the classification scheme is in genus level placements, e.g.
@@
 
-Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed reclassifications of the contents of pollen grains...
+Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed re-classifications of the contents of pollen grains...
 
@@
 ```
+
+\newpage
+
+# References

### 2023-02-04 script/manuscript_MEE.Rmd
margins of ref seq plots wider

@@
 author:  |
-    |  Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
     |  $^4$ Australia post fire project, ----, ----, ----, AU  
+    |  Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
 abstract:  |  
@@
 ```
+
+The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+
+In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
+Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
+Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
+
+At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
 

### 2023-02-12 script/flowchart.R
added code for flowchart

@@ -1,0 +1,95 @@
+library(DiagrammeR)
+
+setwd('/home/sagesteppe/Documents/ManuSCript/script')
+
+g <- grViz("
+
+digraph boxes_and_circle {
+  
+  node [shape = rectangle, fontsize = 18]
+  
+      # fieldwork operations
+      subgraph cluster_fieldwork {
+        style=dashed; color= '#ffa600'; pendwidth = 3
+        node [fontname = arial, style = filled, fontcolor = White, color = '#ffa600']
+        
+          'field work' -> 'plant survey'
+          'plant survey' -> 'herbarium samples' [label = 'vouchers', fontsize = 18]
+          'field work' -> 'pollinator obs.'
+          'pollinator obs.' -> 'pollen samples' [label = 'capture insects', fontsize = 18]
+      }
+        
+      # Main spatial operations
+      subgraph cluster_spatial {
+      style=dashed; color= '#ff764a'; pendwidth = 3
+      node [fontname = arial, style = filled, fontcolor = White, color = '#ff764a']
+      
+      'distance query' -> 'SDM' [label = 'log. regression',  fontsize = 18]
+      }
+      
+      subgraph cluster3 {
+      style=dashed; color= '#ef5675'; pendwidth = 3
+      node [fontname = arial,  style = filled, fontcolor = White, color = '#ef5675']
+
+        'herbarium samples' -> 'DNA extraction' 
+        'DNA extraction' -> 'hyb-seq'  [label = 'CTAB',  fontsize = 18]
+        'hyb-seq' -> 'mi-Seq' [label = 'A353', fontsize = 18]
+      }
+      
+      subgraph cluster_morphology {
+      style=dashed; color= '#bc5090'; pendwidth = 3
+      node [fontname = arial, style = filled, fontcolor = White, color = '#bc5090']
+      
+        'herbarium samples' -> 'pollen morpho'
+        'pollen morpho' -> 'dichotomous key' [label = 'cluster', fontsize = 18]
+        'dichotomous key' -> 'count grains' [label = 'microscopy', fontsize = 18]
+        'pollen samples' -> 'count grains'
+      }
+      
+      subgraph cluster_phenology {
+      style=dashed; color= '#7a5195'; pendwidth = 3
+      node [fontname = arial, style = filled, fontcolor = White, color = '#7a5195']
+      
+        'SDM' -> 'museum records' 
[…]

### 2023-02-15 script/manuscript_MEE.Rmd
new results in, lower detection?

@@
 fig_caption: yes
-always_allow_html: true
+always_allow_html: yes
 header-includes:
@@
   mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
-  ungroup()
+  slice_max( n = 10, with_ties = F, order_by = Prcnt_seqs)
 blst_seqs <- distinct(blst, sample.id)
@@
 
-# PERFORM FOR ERICACEAE TOO - A single record maybe come out?
-ericaceae_samples <- pmorph %>% 
-  filter(morphotype == 'ERICACEAE') %>% 
-  select(sample.id, Prop, morphotype)
-
-blst_ericoid <- filter(blst, Family == 'Ericaceae') %>% # 44 might be Epilobium anagallidolium
-  mutate(morphotype = 'ERICACEAE')
-ericaceae_detected <- ericaceae_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_ericoid$sample.id, 1, 0))
-
 # ASTER 2 HERE...
@@
 dd <- prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
-ee <- prcnt_detect(ericaceae_detected, Lprop = 0.001, Hprop = 0.1)
 ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)
@@
 # Semi-quantitative accord between sequence % and Grain Count
-palyn_cnt <- bind_rows(aster1_samples, aster2_samples, ericaceae_samples, mertensia_samples) %>% 
+palyn_cnt <- bind_rows(aster1_samples, aster2_samples, mertensia_samples) %>% 
   select(morphotype, countProp = Prop)
 
-pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia, blst_ericoid) %>% 
+pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia) %>% 
   select(morphotype, sequenceProp = Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
@@
    blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
-   blst_mertensia, aster_samples, ericaceae_samples, blst_ericoid)
+   blst_mertensia, aster_samples)
 ```
@@
              col.names = c('Condition', 'No. Class.', 'Prcnt. Class.', 'Total Seqs', 'Rank'),
-             caption = "Post classification of Sequences via Taxonomy and Ecology") %>% 
+             caption = "Post classification of Sequences via Taxonomy and Ecology,\ntop 15 most abundant reads") %>% 
   kableExtra::kable_styling(
@@
 Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`, and Ericaceae `r toString(ee)`.
+Both morphotypes of Asteraceae pollen `r toString(dd)`.
 
@@
 
-par(pty="s")
-plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-      xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-     xlab = 'Grains Counted', ylab = 'Classified Sequences',
-     pch=18, slty = 3, gap = 0.025)
-axis(1)
-axis(2)
-abline(model)
-text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
-text(0.3, 0.8, "p < 0.001")
-
-rm( newdat)
+# par(pty="s")
+# plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
+#       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
+#      xlab = 'Grains Counted', ylab = 'Classified Sequences',
+#      pch=18, slty = 3, gap = 0.025)
+# axis(1)
+# axis(2)
+# abline(model)
+# text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
+#text(0.3, 0.8, "p < 0.001")
+
+# rm( newdat)
 ```
@@
 6.      N              Climatic moisture deficit                       Wang et al.
-7.      N              Degree-days above 5C from                       Wang et al.
+7.      N               Degree-days above 5C                           Wang et al.
 8.      N              Mean annual precipitation                       Wang et al.

### 2023-02-15 script/pollen_countr3.Rmd
new results in, lower detection?

@@
   mutate(count = replace_na(count, 0)) %>% # remember in our sample design NA's are 0's
-  filter(transect >= 4, transect <= 17) %>%  # remove the 1st and 20th transects which we stopped counting. 
   mutate(group_sum = sum(count)) %>% # see which slides have been counted...
@@
   filter(group_sum >= 1) %>% # find only the slides which we have counted pollen grains in .
-  mutate(morphotype = str_replace(morphotype, "CAN\'T_DISTINGUISH", "UNKNOWN")) %>% 
-  mutate(morphotype = str_replace(morphotype, "INDISTINGUISHABLE_CLUMP", "UNKNOWN")) %>% 
-  mutate(morphotype = str_replace(morphotype, "NER", "UNKNOWN")) %>% 
-  mutate(morphotype = str_replace(morphotype, "DEHYDRATED", "UNKNOWN")) %>%
-
-  mutate(morphotype = str_replace(morphotype, "SALIX", "D")) %>% 
-  mutate(morphotype = str_replace(morphotype, "C\\_\\(LINUM\\)", "LINUM")) %>% 
-  mutate(morphotype = str_replace_all(morphotype, " ", "_")) %>% # replace white spaces 
-  mutate(morphotype = str_to_upper(morphotype)) %>% # R is case sensitive
-  mutate(morphotype = str_remove_all(morphotype, "\\_\\(11B\\)")) %>% # while good to know, unknown will be it's own class. 
-  mutate(morphotype = if_else(str_detect(morphotype, "\\?"), "UNKNOWN", morphotype)) %>% # make all unknown observations the same format. 
-  mutate(morphotype = str_replace(morphotype, "UNK$", "UNKNOWN")) %>% # make all unknown observations the same format. 
-  mutate(morphotype = str_replace_all(morphotype, "UNKNWOWN", "UNKNOWN")) %>% 
-  filter(!morphotype %in% c("NO_FJ", "LOW_FJ", "LITTLE_FJ", "LESS_FJ")) %>%  # we cannot say we counted or scored these cells. 
-  filter(!sample.id %in% c(28, 29, 59))  %>%  # THIS SLIDES ARE PURELY OF A SINGLE TAXON, AND CANNOT BE ANALYZED. 
-  filter(!morphotype %in% c(4,8)) %>% 
+  mutate(
+    morphotype = str_to_upper(morphotype),
+    morphotype = str_replace(morphotype, "CAN\'T_DISTINGUISH", "UNKNOWN"),
+    morphotype = str_replace(morphotype, "INDISTINGUISHABLE_CLUMP", "UNKNOWN"),
+    morphotype = str_replace(morphotype, "NER", "UNKNOWN"),
+    morphotype = str_replace(morphotype, "DEHYDRATED", "UNKNOWN"),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]5B[)]', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, '5B', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, 'ASTERACEAE-2', 'ASTERACEAE_2'),
+    morphotype = str_replace(morphotype, "SALIX", "D"),
+    morphotype = str_replace(morphotype, 'C4', 'C'),
+    morphotype = str_replace(morphotype, 'OENOTHERA', 'ONAGRACEAE'),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]D[)]', 'D'),
+    morphotype = str_replace(morphotype, 'OTHER_B', 'B'),
+    morphotype = str_replace(morphotype, 'OTHER_C', 'C'),
+    morphotype = str_replace(morphotype, "C\\_\\(LINUM\\)", "LINUM"),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]RANUNCULUS[)]', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]11B[/]D[)]', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]4_APERTURES[)]', 'UNKNOWN'),
+    morphotype = str_replace_all(morphotype, " ", "_"),
+    morphotype = str_remove(morphotype, "\\_\\(11B\\)"),
+    morphotype = str_replace_all(morphotype, "UNKNWOWN", "UNKNOWN")) %>% 
+  filter(!morphotype %in% c("NO_FJ", "LOW_FJ", "LITTLE_FJ", "LESS_FJ", 'LILTTLE_FJ')) %>%  # we cannot say we counted or scored these cells. 
+  filter(!morphotype %in% c(4,8, 11)) %>% 
   mutate(plot.id = rleid(sample.id, transect, cell))
@@
 
-write.csv(pollen_results, '/hdd/yun master/MEE_manuscript/data/morpho_pollen.csv', row.names = F)
+write.csv(pollen_results, '../data/morpho_pollen.csv', row.names = F)
 ```
@@
 
-rich_rare <- data.table::rbindlist(lapply(to_rare[c(1:41,43:44,46:60)], raw_vegan), idcol = T) # a couple still not working, 
+#[c(1:41,43:44,46:60)]
+# 45 does not work
+u <- lapply(lapply(to_rare, rowSums), sum)
+
+to_rare <- to_rare[u > 1]
+rich_rare <- data.table::rbindlist(lapply(to_rare, raw_vegan), idcol = T)
+
 rich_rare <- rich_rare %>% rename(sample.id = '.id') %>% 
@@
 
-#   Samples 45 and 48 missing due to only a single morphotype being present per slide. 
+#   Samples 45 iz missing due to only a single morphotype being present per slide. 
 ggsave(

### 2023-02-19 script/manuscript_MEE.Rmd
new data in

@@
 author:  |
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
     |  $^4$ Australia post fire project, ----, ----, ----, AU  
-    |  Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
 abstract:  |  
@@
 
-  To increase the quality of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest.
+  To increase the quality of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
 To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
@@
 
+no_plant_sp <- bee_obs %>% 
+  ungroup() %>% 
+  distinct(plant.species) %>%
+  nrow()
+
 s <- toString(format(as.Date(min(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # start obs
@@
 
-The six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours from .
+The six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
 A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
 Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week min = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, Mdn = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (min = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, Mdn = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
-The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`).
+The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
 
@@
   arrange(sample.id, -Prcnt_seqs) %>% 
-  inner_join(., corb, by = 'sample.id')
+  inner_join(., corb, by = 'sample.id')  %>% 
+  distinct() %>% 
+  group_by(sample.id) %>% 
+  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
+  ungroup()
+
+# combine reassigned TAXON_NEW
+
+blst <- blst %>% 
+  group_by(TAXON_NEW, sample.id) %>% 
+  mutate(across(.cols = Seqs_pr_taxon:Prcnt_seqs, ~ sum(.x))) %>% 
+  distinct(TAXON_NEW, sample.id, .keep_all = T) %>%
+
+  # slice top ten records
+  group_by(sample.id) %>% 
+  slice_max(order_by = Prcnt_seqs, n = 10) %>% 
+  ungroup()
 
@@
 gf_lkp2 <- read.csv('../data/genus2family_lkp.csv')
-gf_lkp <- blst %>% select(Genus, Family) %>% 
+gf_lkp <- blst %>% 
+  select(Genus, Family) %>% 
   distinct(Genus, .keep_all = T) %>% 
@@
   left_join(., blst_cnt, by = 'species') %>% 
-  group_by(species, TAXON_NEW2) %>% 
+  group_by(species, TAXON_NEW) %>% 
   mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
          species = str_remove(species, 'Bombus '))  %>% 
-  distinct(species, TAXON_NEW2, .keep_all = T) %>% 
-  ungroup(TAXON_NEW2) %>% 
+  distinct(species, TAXON_NEW, .keep_all = T) %>% 
+  ungroup(TAXON_NEW) %>% 
   slice_max(Mu_Seqs, n = 10) %>% 
-  select(plant.species = TAXON_NEW2, species, n, Mu_Seqs, Genus, Family) %>% 
+  select(plant.species = TAXON_NEW, species, n, Mu_Seqs, Genus, Family) %>% 
   mutate(rank = 1:n()) %>% 
@@
   left_join(., gf_lkp, by = 'Genus') %>% 
+  distinct() %>% 
   filter(species %in% tb$species) %>% 
@@
 species <- purrr::map2_int(tblist, oblist,
-                ~ inner_join(.x, .y, by = 'plant.species') %>% nrow())
+                ~ inner_join(.x, .y, by = 'plant.species') %>%
+                   nrow())
 
@@
 A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+Includinfg these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
 
@@
 pmorph <- read.csv('../data/morpho_pollen.csv') 
-blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
-  rename(sample.id = Sample, Family = FAMILY, Genus = GENUS) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
-  slice_max( n = 10, with_ties = F, order_by = Prcnt_seqs)
 blst_seqs <- distinct(blst, sample.id)
@@
 
-rm(mertensia_samples, blst_aster, blst_aster1,
-   blst_aster2, aster1_samples,  aster2_samples, a, b, c, d, palyn_cnt,
-   blst_mertensia, aster_samples)
+rm(mertensia_samples, blst_aster, blst_aster1, blst_aster2, aster1_samples, 
+   aster2_samples, palyn_cnt, blst_mertensia, aster_samples)
 ```
@@
 
-pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') )
+pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') ) %>% 
+  mutate(across(where(is.numeric), ~ round(.x, 1)))
 
@@
 
-seqs_sp <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records'] )
-seqs_ge <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records'] )
-seqs_fa <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records'] )
-
-sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs'] )
-sp_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs'] )
-fa_total_seqs <- sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs'] )
+seqs_sp <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records']), 1)
+seqs_ge <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records']), 1)
+seqs_fa <- round(sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records']), 1)
+
+sp_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs']),1)
+sp_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs']),1)
+fa_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs']),1)
 
 byhand <- blst %>% 
-  filter(TAXON_NEW != TAXON_NEW2)
+  filter(TAXON_NEW != TAXON_NEW)
 
@@
                      as.vector (table(byhand$Condition))))
-gen <- byhand[byhand$TAXON_NEW2==c('Salix'),]
-spp <- byhand[byhand$TAXON_NEW2==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
+gen <- byhand[byhand$TAXON_NEW==c('Salix'),]
+spp <- byhand[byhand$TAXON_NEW==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
 ```
@@
 Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
-Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species `r sp_total_seqs`% representing of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family `r fa_total_seqs`%  of all total sequences.
+Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species representing `r sp_total_seqs`% of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family.
 
@@
 rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
-    ge_total_seqs, sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
+    sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
 ```
@@
 ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
-# ct[["p.value"]] too small to write out. tau > 0.4 strong relationship
+# ct[["p.value"]] #too small to write out. tau > 0.4 strong relationship
 model <-  lm(adjSeqProp ~ countProp, data = pol_p)
 
-# par(pty="s")
-# plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-#       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-#      xlab = 'Grains Counted', ylab = 'Classified Sequences',
-#      pch=18, slty = 3, gap = 0.025)
-# axis(1)
-# axis(2)
-# abline(model)
-# text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
-#text(0.3, 0.8, "p < 0.001")
+# png('../../thesis_talk/graphics/plots/CorrelationCountedSequences.png', )
+par(pty="s")
+plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
+       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
+      xlab = 'Grains Counted', ylab = 'Classified Sequences',
+      pch=18, slty = 3, gap = 0.025)
+axis(1)
+axis(2)
+abline(model)
+text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
+text(0.3, 0.8, "p < 0.001") 
+# dev.off()
 
@@
 The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
-This is likely due to the proportion of high false positives which occur in the classification process with NGS (BELL NOVEMEBER 2021).
+This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing).
 There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
@@
   mutate(TAXON_EXPERTS = case_when(
-      TAXON_NEW2 %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
+      TAXON_NEW %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
       'Cynoglossum officinale') ~ 'Mertensia',
-      TAXON_NEW2 %in% c('Caltha leptosepala', 'Trollius europaeus', 'Trollius') ~ 'Delphinium',
+      TAXON_NEW %in% c('Caltha leptosepala', 'Trollius europaeus', 'Trollius') ~ 'Delphinium',
     # 'Thalictrum thalictroides' ~ 'Delphinium',
-      TAXON_NEW2 == 'Erigeron' ~ 'Erigeron grandiflorus', 
-      TAXON_NEW2 == 'Scabrethia scabra' ~ 'Wyethia arizonica',
-      TAXON_NEW2 == 'Nemophila breviflora' ~ 'Hydrophyllum',
-      TAXON_NEW2 == 'Erigeron grandiflorus' ~ 'Erigeron grandiflorus',
-      TAXON_NEW2 == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
-      TAXON_NEW2 == 'Paxistima myrsinites' ~ 'Parnassia palustris', # original was probably right !!! shows the consequence of bad time estimates.#, 
-    # TRUE ~ as.character(TAXON_NEW2)
+      TAXON_NEW == 'Erigeron' ~ 'Erigeron grandiflorus', 
+      TAXON_NEW == 'Scabrethia scabra' ~ 'Wyethia arizonica',
+      TAXON_NEW == 'Nemophila breviflora' ~ 'Hydrophyllum',
+      TAXON_NEW == 'Erigeron grandiflorus' ~ 'Erigeron grandiflorus',
+      TAXON_NEW == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
+      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', # original was probably right !!! shows the consequence of bad time estimates.#, 
+    # TRUE ~ as.character(TAXON_NEW)
   )) %>% 
-  select(TAXON, TAXON_NEW2, TAXON_EXPERTS)
+  select(TAXON, TAXON_NEW, TAXON_EXPERTS)
 
@@
 
+```{r Overview of entire process, out.height="125%"}
+knitr::include_graphics('../graphics/plots/flowchart.png')
+```
+
+\newpage
+
 # References

### 2023-02-19 script/manuscript_MEE.Rmd
added new graphs, reporting on phenology, now working on "expert" recoding of results =-\

@@
 Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
+These estimates were compared to groundtruthed data from a long term experiment 1974-2012, and the field data from 2015, using kendall's tau.
 
@@
                          group = Bombus.Species, fill = Bombus.Species)) +
- # geom_line(linewidth = 1.25, linetype = 'dotted', alpha = 0.3)+
   geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
@@
 
-Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
+Only 58 of these 388 species (n = 34.6, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
 Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
-There was very strong evidence that the weibull estimates were positively associated with the observed onset (r^2^ = 0.72, p < 0.0001, tau = 0.61) and peak (r^2^ = 0.70, p < 0.0001, tau = 0.65) of flowering, and that the number of herbarium samples had a moderate effect on the estimates (p = 0.004 and p = 0.034 respectively).
-There was very strong evidence that the weibull estimates had a positive association with the observed cessation of flowering (r^2^ =  0.4339,  p < 0.0001, tau = 0.489), however their was no evidence that sample size had an effect (p = 0.349).
-There was moderate evidence that the weibull estimates, with an effect of sample size, had a weak positive association with the observed duration of flowering (p = 0.0401, r^2^ = 0.07, tau = 0.17).
-
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '100%'}
+There was very strong evidence that the weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61),  peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
+There was moderate evidence that the weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
+
+Of the previous 58 species compared, 47 of these could be compared to plot based data from the six sites observed in 2015.
+Due to methodological differences, the peak flowering was not compared, and due to the low performance of attempts to model 'duration' in the previous step it was also not compared.
+There was very strong evidence that the weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).
+
+```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
 knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
 # these are the tenth and nintey  percent non-parametric confidence intervals. 
+```
+
+```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
+knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
 ```
@@
 **APPENDIX X Reads Per Loci**.
-
-... with samples 22, and 41 both having < 100 reads indicates, virtual failure of these records (REMOVE from analyses).
 
@@
 
-The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
+The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly linear, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
 This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing).
@@
   filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
-                       'Celastraceae', 'Fabaceae')) %>% 
+                       'Celastraceae')) %>% 
   select(sample.id, Genus) %>% 
@@
 species were present.
+The size of flower of *Nemophila breviflora* A.
+Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
+The floral morphology of *Thalictrum*  spp.
+also makes it unlikely to be visited (BELEIVE JANE SAID THIS- IF NOT STRIKE OUT), and while evidence of visits to Caltha and Trollius are lacking, due to the association between the reads these results appear unlikely.
 
@@
 
-```{r}
-
-blst_final <- blst %>% 
+```{r Build reference dataset for reclassification process}
+
+pollen_sample_info <- readxl::read_xlsx(file.path('../data', 
+  list.files(path = '../data', pattern = 'xlsx') )) %>% 
+  select(site, date, sample.id) %>% 
+  distinct() %>% 
+  filter(!date %in% c('42170', '42169'))
+
+files <- list.files(path = '../data')
+
+flr_date_ranges <- read.csv(
+  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
+  select(site, doy, week, plant.species:abun.rank) %>% 
+  group_by(plant.species, site) %>% 
+  select(site, week, doy, plant.species) %>% 
+  mutate(Genus = str_extract(plant.species, '^.+?(?=[.]|$)')) %>% 
+  mutate(Genus = case_when(
+    Genus == 'Erythrocoma' ~ 'Geum', 
+    Genus == 'Amerosedum' ~ 'Sedum', 
+    Genus == 'Pentaphylloides' ~ 'Dasiphora',
+    TRUE ~ Genus
+    )) %>% 
+  mutate(plant.species = case_when(
+    plant.species == 'Lathyrus.leucanthus' ~ 'Lathyrus.lanszwertii',
+    plant.species == 'Lupinus.bakeri' ~ 'Lupinus.sericeus',
+    TRUE ~ plant.species
+  )) %>% distinct()
+ # this provides us with a data.frame of the ACTUAL flowering dates!
+
+week_tab <- flr_date_ranges %>% 
+  ungroup() %>% 
+  select(doy, week) %>% 
+  distinct()
+
+flr_date_ranges <- select(flr_date_ranges, -doy)
+
+blst_expert <- blst %>% 
+  select(species, sample.id, TAXON_NEW, Genus, Family, Prcnt_seqs, Prcnt_Seqs_Class) %>% 
+  mutate(RecordID = 1:n()) %>% 
+  left_join(., pollen_sample_info,  by = 'sample.id') %>% 
+  relocate(site:date, .before = TAXON_NEW) %>% 
+  mutate(doy = lubridate::yday(date)) %>% 
+  left_join(., week_tab, by = 'doy') %>% 
+  left_join(., flr_date_ranges, by = c('site', 'Genus', 'week')) %>% 
+  relocate(sample.id, .before = plant.species)
+
+  # reduce to temporally similar results
+
+ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia')
+
+# all samples with duplicate genera in them # be careful these are not easy mappings
+duplicate_reps_in_sample <- blst_expert %>%
+
+  # identify the duplicate samples
+  drop_na(plant.species) %>% 
+  group_by(sample.id, TAXON_NEW) %>% 
+  ungroup(TAXON_NEW)  %>% 
+  distinct(plant.species, .keep_all = T) %>% 
+  group_by(sample.id, TAXON_NEW) %>% 
+  filter(n() >= 2) %>%
+
+  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Soloman'
+  ungroup() %>% 
+  group_by(RecordID) %>% 
+  mutate(Prcnt_seqs_Reclass = Prcnt_seqs/n())
+
+blst_reclass_expert <- blst_expert %>% 
+  filter(!RecordID %in% duplicate_reps_in_sample$RecordID) %>% 
+  rowwise() %>% 
+  mutate(Genus = str_trim(Genus))  %>%
+
+  mutate(TAXON_EXPERT = case_when(
+    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
+    sample.id == 4 ~ plant.species, # 4 has a viola split
+    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 6 ~ plant.species, # has a delphinium split
+    sample.id == 6 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
+
+    sample.id == 9 ~ plant.species, # has both Delphinium
+    sample.id == 10 ~ plant.species,
+
+    sample.id == 14 ~ plant.species,
+
+    sample.id == 18 ~ plant.species,
+    sample.id == 19 ~ plant.species,
+    sample.id == 20 ~ plant.species,
+    sample.id == 21 ~ plant.species,
+    sample.id == 23 ~ plant.species, # has both Viola
+    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 24 ~ plant.species,
+    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 25 ~ plant.species,
+    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 26 ~ plant.species,
+    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+  #  sample.id == 27 ~ plant.species, # has both Hydrophyllum
+    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 28 ~ plant.species,
+    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 29 ~ plant.species,
+    sample.id == 30 ~ plant.species,
+    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
+    sample.id == 32 ~ plant.species,
+    sample.id == 32 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 33 ~ plant.species, # has both Viola
+
+    sample.id == 35 ~ plant.species, # has both Viola
+    sample.id == 35 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 36 ~ plant.species,
+
+    sample.id == 39 ~ plant.species, # has both hydrophyllum
+    sample.id == 39 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 41 ~ plant.species,
+    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata',
+
+    sample.id == 42 ~ plant.species,
+    sample.id == 43 ~ plant.species,
+    #sample.id == 44 ~ plant.species, # 44 has both delphinium
+
+    sample.id == 46 ~ plant.species,
+    sample.id == 49 ~ plant.species,
+    sample.id == 49 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.barbeyi',
+  #  sample.id == 51 ~ plant.species, # 51 has a viola split
+    sample.id == 51 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 52 ~ plant.species,
+    sample.id == 54 ~ plant.species,
+    sample.id == 55 ~ plant.species,
+
+    sample.id == 56 ~ plant.species,
+    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
+
+    sample.id == 60 ~ plant.species,
+    sample.id == 65 ~ plant.species, # 65 has a mertensia split
+
+   TRUE ~ TAXON_NEW
+  )) %>%
+
+  # use this to sum up the total sequences by original read record
+  distinct(sample.id, TAXON_NEW, .keep_all = T) %>% 
+  group_by(sample.id, plant.species) %>% 
+  mutate(Prcnt_seqs_Reclass = sum(Prcnt_seqs)) %>%
+
+  # now reduce to the amount of rows of the proper classification
+  ungroup() %>% 
+  distinct(sample.id, TAXON_EXPERT, .keep_all = T)
+
   mutate(TAXON_EXPERTS = case_when(
       TAXON_NEW %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
-      'Cynoglossum officinale') ~ 'Mertensia',
-      TAXON_NEW %in% c('Caltha leptosepala', 'Trollius europaeus', 'Trollius') ~ 'Delphinium',
-    # 'Thalictrum thalictroides' ~ 'Delphinium',
+      'Cynoglossum officinale', 'Cynoglossum pringlei') ~ 'Mertensia',
+      TAXON_NEW %in% c('Caltha leptosepala', 'Caltha palustris', 'Trollius europaeus', 'Trollius', 
+                       'Thalictrum thalictroides', 'Aquilegia ecalcarata', 
+                       'Thalictrum tuberosum') ~ 'Delphinium',
       TAXON_NEW == 'Erigeron' ~ 'Erigeron grandiflorus', 
@@
       TAXON_NEW == 'Nemophila breviflora' ~ 'Hydrophyllum',
-      TAXON_NEW == 'Erigeron grandiflorus' ~ 'Erigeron grandiflorus',
       TAXON_NEW == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
-      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', # original was probably right !!! shows the consequence of bad time estimates.#, 
+      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', #, 
+      TAXON_NEW == 'Taraxacum cucullatum' ~ 'Taraxaum oficinale'
+    # original was probably right !!! shows the consequence of bad time estimates.#, 
   #   TRUE ~ as.character(TAXON_NEW)
@@
 
-# no TARAXACUM OFICINALE, shows a limitation of the data sources.
-
+rm(pollen_sample_info)
 ```
@@
 Gray (@weber1998new), a taxon known to be visited by Queen bee's via our floral observations.
-
-It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes.
-For example, no more than 10 bee corbiculae loads per species were sequenced with the Mdn = `r median(n_corbiculae)` ...
-, and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top.
-
 An expected inaccuracy of the classification scheme is in genus level placements, e.g.
@@
 or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.
-
+An issue with reclassification within the family level in combination with time included reclassifying *Parnassia palustris* L.
+to *Paxistima myrsinites* (Pursh) Rafinesque.
+However, based on flower size it is more likely that the visited taxon was *P.
+palustris*.
+
+It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes.
+For example, no more than 10 bee corbiculae loads per species were sequenced with the Mdn = `r median(n_corbiculae)` ...
+, and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top interactions.
 Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed re-classifications of the contents of pollen grains...
@@
 
-##### ~ What we DEMONSTRATED ~
-
 We have demonstrated how the Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context (@johnson2019universal, @hollingsworth2016telling).
-This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history driven fundamental science and land management.
+This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history guided fundamental science and land management.
 The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g.
@@
 By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.
-
-##### ~ What CHALLENGES we FACED ~
 
@@
 We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data.
-Further given the, size of the minimum spanning tree (AREA???) which we extracted points to, a formal floristic inventory would still be a time intensive process.
+Further given the, size of the minimum spanning tree  which we extracted points to, a formal floristic inventory would still be a time intensive process.
 Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding.
 We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.
-  Difficulties in temporal models related to variability in drivers of flowering phenology.
-
-##### ~ WHAT we learned about Bee foraging (BASIC) ~
+  Difficulties in temporal models related to variability in drivers of flowering phenology. AND SMALL SAMPLE SIZE
 
@@
 
-Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Hydrophyllum* on a per visit basis...
+Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Mertensia* on a per visit basis...
 General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persit with *Bombus* studies...
-
-##### ~ WHAT this tells us about Bee foraging (APPLIED) ~
 
@@
 
-##### ~ WHERE we see spatial/temporal going
-
   We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical).
-To reduce the effects of a low population density of botanists on the maintenance of and production of flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
+To reduce the effects of a low population density of botanists on the maintenance of and production of Flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
 In this proof of concept example we performed several iterations of modelling runs, and several approaches (i.e.
@@
 ; check assumptions of model again and just do a quasi something** A more promising approach for the tropics may lay in circular statistics (@park2022herbarium).
-
-##### ~ WHERE we see MOLECULAR going
 

### 2023-02-20 script/manuscript_MEE.Rmd
working on expert reclass

@@
 
+rm(corb, week_tab, flr_date_ranges, pollen_sample_info, blst)
+
   # reduce to temporally similar results
 
-ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia')
-
-# all samples with duplicate genera in them # be careful these are not easy mappings
+######## SAMPLE 55 SHOULD BE SOMEWHERE HERE. 
 duplicate_reps_in_sample <- blst_expert %>%
@@
   drop_na(plant.species) %>% 
-  group_by(sample.id, TAXON_NEW) %>% 
-  ungroup(TAXON_NEW)  %>% 
+  group_by(RecordID) %>% 
   distinct(plant.species, .keep_all = T) %>% 
   group_by(sample.id, TAXON_NEW) %>% 
-  filter(n() >= 2) %>%
-
-  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Soloman'
+  filter(n() >= 2)
+
+dupe_ids <- duplicate_reps_in_sample %>% pull(RecordID) %>% unique()
+dupe_samples <- duplicate_reps_in_sample %>% pull(sample.id) %>% unique()
+
+  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Solomon'
+duplicate_reps_in_sample <- duplicate_reps_in_sample %>% 
   ungroup() %>% 
-  group_by(RecordID) %>% 
-  mutate(Prcnt_seqs_Reclass = Prcnt_seqs/n())
+  group_by(sample.id, Genus) %>% 
+  mutate(Prcnt_seqs_Reclass = 
+           ifelse(n() == 2, Prcnt_seqs/2, sum(Prcnt_seqs)/n_distinct(plant.species)/2 ),
+         TAXON_EXPERT = plant.species) %>% 
+  distinct(sample.id, TAXON_EXPERT, .keep_all = T)
+
+ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia')
 
 blst_reclass_expert <- blst_expert %>% 
-  filter(!RecordID %in% duplicate_reps_in_sample$RecordID) %>% 
+  filter(!RecordID %in% dupe_ids) %>% 
+  distinct() %>% 
   rowwise() %>% 
@@
     sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
+    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 4 ~ plant.species, # 4 has a viola split
-    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
     sample.id == 6 ~ plant.species, # has a delphinium split
-    sample.id == 6 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
 
@@
 
+    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',
     sample.id == 18 ~ plant.species,
+
+    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', 
     sample.id == 19 ~ plant.species,
+
+    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
     sample.id == 20 ~ plant.species,
     sample.id == 21 ~ plant.species,
+
+    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 23 ~ plant.species, # has both Viola
-    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 24 ~ plant.species,
-    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 25 ~ plant.species,
-    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 26 ~ plant.species,
-    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-  #  sample.id == 27 ~ plant.species, # has both Hydrophyllum
+
     sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 27 ~ plant.species, # has both Hydrophyllum
+
+    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 28 ~ plant.species,
-    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
     sample.id == 29 ~ plant.species,
+
+    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
     sample.id == 30 ~ plant.species,
-    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
+
+    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 32 ~ plant.species,
-    sample.id == 32 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
     sample.id == 33 ~ plant.species, # has both Viola
 
+    sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  # 7.99
     sample.id == 35 ~ plant.species, # has both Viola
-    sample.id == 35 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-
     sample.id == 36 ~ plant.species,
 
+    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 39 ~ plant.species, # has both hydrophyllum
-    sample.id == 39 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-
+    sample.id == 40 ~ plant.species,
+
+    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
     sample.id == 41 ~ plant.species,
-    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata',
 
@@
     sample.id == 43 ~ plant.species,
-    #sample.id == 44 ~ plant.species, # 44 has both delphinium
-
+    sample.id == 44 ~ plant.species, # 44 has both delphinium
+
+    sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
     sample.id == 46 ~ plant.species,
+
+    sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
     sample.id == 49 ~ plant.species,
-    sample.id == 49 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.barbeyi',
-  #  sample.id == 51 ~ plant.species, # 51 has a viola split
-    sample.id == 51 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+
+    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
+    sample.id == 51 ~ plant.species, # 51 has a viola split
 
@@
     sample.id == 54 ~ plant.species,
+
+    ######### 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? ####
+    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
     sample.id == 55 ~ plant.species,
 
+    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
     sample.id == 56 ~ plant.species,
-    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
 
@@
 
-   TRUE ~ TAXON_NEW
+   TRUE ~ as.character(TAXON_NEW)
   )) %>% 
+  mutate(
+    TAXON_EXPERT = str_replace_all(TAXON_EXPERT, '[.]', ' '),
+    TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Arenaria globiflora', 'Eremogone congesta'),
+    TAXON_EXPERT = if_else(doy < 190 & TAXON_EXPERT == 'Lupinus argenteus', 
+                           'Lupinus sericeus', TAXON_EXPERT),) %>%
 
@@
   distinct(sample.id, TAXON_NEW, .keep_all = T) %>% 
-  group_by(sample.id, plant.species) %>% 
+  group_by(sample.id, TAXON_EXPERT) %>% 
   mutate(Prcnt_seqs_Reclass = sum(Prcnt_seqs)) %>%
@@
   ungroup() %>% 
-  distinct(sample.id, TAXON_EXPERT, .keep_all = T)
-
-  mutate(TAXON_EXPERTS = case_when(
-      TAXON_NEW %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
-      'Cynoglossum officinale', 'Cynoglossum pringlei') ~ 'Mertensia',
-      TAXON_NEW %in% c('Caltha leptosepala', 'Caltha palustris', 'Trollius europaeus', 'Trollius', 
-                       'Thalictrum thalictroides', 'Aquilegia ecalcarata', 
-                       'Thalictrum tuberosum') ~ 'Delphinium',
-      TAXON_NEW == 'Erigeron' ~ 'Erigeron grandiflorus', 
-      TAXON_NEW == 'Scabrethia scabra' ~ 'Wyethia arizonica',
-      TAXON_NEW == 'Nemophila breviflora' ~ 'Hydrophyllum',
-      TAXON_NEW == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
-      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', #, 
-      TAXON_NEW == 'Taraxacum cucullatum' ~ 'Taraxaum oficinale'
-    # original was probably right !!! shows the consequence of bad time estimates.#, 
-  #   TRUE ~ as.character(TAXON_NEW)
-  )) %>% 
-  select(TAXON, TAXON_NEW, TAXON_EXPERTS)
+  distinct(sample.id, TAXON_EXPERT, .keep_all = T) %>% # what is going on with sample 56 then?
+
+  # add back on the multiple groups
+  bind_rows(., duplicate_reps_in_sample)  %>% 
+  arrange(sample.id, -Prcnt_seqs_Reclass)
+
+# now we split up the reads assigned to conflicted groups  # e.g. Cynoglossum amplifolium in sample 65
+b <- blst_reclass_expert %>% 
+  filter(sample.id %in% dupe_samples) %>% 
+  group_by(sample.id, Family) %>% 
+  filter(n() >= 3)
 

### 2023-02-20 script/manuscript_MEE.Rmd
catching weirds

@@
          TAXON_EXPERT = plant.species) %>% 
-  distinct(sample.id, TAXON_EXPERT, .keep_all = T)
+  distinct(sample.id, plant.species, .keep_all = T)
 
@@
   mutate(TAXON_EXPERT = case_when(
+
     sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
@@
     sample.id == 19 ~ plant.species,
-
     sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
@@
     sample.id == 23 ~ plant.species, # has both Viola
-
     sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 24 ~ plant.species,
-
     sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 25 ~ plant.species,
-
     sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 26 ~ plant.species,
-
     sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 27 ~ plant.species, # has both Hydrophyllum
-
     sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 28 ~ plant.species,
-
     sample.id == 29 ~ plant.species,
-
     sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
     sample.id == 30 ~ plant.species,
-
     sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 32 ~ plant.species,
-
     sample.id == 33 ~ plant.species, # has both Viola
@@
     sample.id == 40 ~ plant.species,
-
     sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
     sample.id == 41 ~ plant.species,
-
     sample.id == 42 ~ plant.species,
@@
     sample.id == 51 ~ plant.species, # 51 has a viola split
-
     sample.id == 52 ~ plant.species,
@@
 
-    ######### 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? ####
     sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
@@
     TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Arenaria globiflora', 'Eremogone congesta'),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Geum ternatum', 'Geum triflorum'),
+    TAXON_EXPERT = str_replace(
+      TAXON_EXPERT, 'Arenaria globiflora|Minuartia recurva|Sagina procumbens', 'Eremogone congesta'),
+    TAXON_EXPERT = if_else(doy < 171 & Genus == 'Cynoglossum', 
+                           'Mertensia fusiformis', TAXON_EXPERT), 
     TAXON_EXPERT = if_else(doy < 190 & TAXON_EXPERT == 'Lupinus argenteus', 
-                           'Lupinus sericeus', TAXON_EXPERT),) %>%
+                           'Lupinus sericeus', TAXON_EXPERT)) %>%
 
@@
 
+rm(ranunc_splash, blst_expert)
+
 # now we split up the reads assigned to conflicted groups  # e.g. Cynoglossum amplifolium in sample 65
@@
   group_by(sample.id, Family) %>% 
-  filter(n() >= 3)
+  filter(n() >= 3) %>% 
+  mutate(plant.species = na_if(plant.species, "Thalictrum.fendleri")) %>% 
+  mutate(Seqs2distribute = sum(Prcnt_seqs_Reclass))
+
+records2drop <- b %>% pull(RecordID)
+
+b <- b %>% 
+  drop_na(plant.species) %>% 
+  mutate(Prcnt_seqs_Reclass = Seqs2distribute/n_distinct(TAXON_EXPERT)) %>% 
+  select(-Seqs2distribute)
+
+blst_reclass_expert <- blst_reclass_expert %>% 
+  filter(!RecordID %in% records2drop) %>% 
+  bind_rows(., b) %>% 
+  arrange(sample.id, -Prcnt_seqs_Reclass)
+
+rm(records2drop, b)
+
+# 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? #
+# seems i may not have either delphinium present in that week? and  if falls between two. 
+# deal with sample 55 one off, split the delphinium gracile
+
+s55 <- blst_reclass_expert %>% 
+  filter(sample.id == 55) %>% 
+  group_by(Genus) %>% 
+  mutate(Seqs2distribute = sum(Prcnt_seqs_Reclass)) %>% 
+  filter(TAXON_EXPERT != 'Delphinium gracile') %>% 
+  mutate(Prcnt_seqs_Reclass = Seqs2distribute/n_distinct(TAXON_EXPERT)) %>% 
+  select(-Seqs2distribute)
+
+blst_reclass_expert <- blst_reclass_expert %>% 
+  filter(sample.id != 55) %>% 
+  bind_rows(., s55)  %>% 
+  arrange(sample.id, -Prcnt_seqs_Reclass)
+
+rm(s55)
+
+# re class the totals to sum to 100%
+
+blst_reclass_expert <- blst_reclass_expert %>% 
+  group_by(sample.id) %>% 
+  mutate(multiplier = 100/ sum(Prcnt_seqs_Reclass),
+         Prcnt_seqs_Reclass2 = Prcnt_seqs_Reclass * multiplier)
 

### 2023-02-20 script/manuscript_MEE.Rmd
more cleaning

@@
 author:  |
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
     |  $^4$ Australia post fire project, ----, ----, ----, AU  
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$,
 abstract:  |  
@@
 
-rm(corb, week_tab, flr_date_ranges, pollen_sample_info, blst)
+rm(corb, week_tab, flr_date_ranges, pollen_sample_info)
 
@@
     sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 4 ~ plant.species, # 4 has a viola split
+    sample.id == 4 ~ plant.species,                               ## COMPLETE ##
 
     sample.id == 6 ~ plant.species, # has a delphinium split
-
-    sample.id == 9 ~ plant.species, # has both Delphinium
-    sample.id == 10 ~ plant.species,
-
-    sample.id == 14 ~ plant.species,
-
-    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',
-    sample.id == 18 ~ plant.species,
-
-    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', 
-    sample.id == 19 ~ plant.species,
+    # Symphyotrichum foliaceum check on day 173
+
+    sample.id == 9 ~ plant.species, # has both Delphinium # Symphyotrichum foliaceum check on day 176
+
+    sample.id == 10 ~ plant.species, # has many late season taxa
+
+    sample.id == 14 ~ plant.species,  # # Symphyotrichum foliaceum check on day 165
+
+    sample.id == 15 ~ plant.species, ## has many mysterious Rosids @ Gothic Spires  - call potentilla?
+
+    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', # & Ericaceae
+    sample.id == 18 ~ plant.species, # Erigeron grandiflorus on day 166 @ Frogs Pond
+    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', # multiple Onagraceae
+    sample.id == 19 ~ plant.species, # 
     sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
-    sample.id == 20 ~ plant.species,
-    sample.id == 21 ~ plant.species,
+    sample.id == 20 ~ plant.species, # Erigeron grandiflorus doy 165 @ Avery Wash
+    sample.id == 21 ~ plant.species, # Erigeron grandiflorus & cirsium parryi doy 165 @ Gothic Spires
 
     sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 23 ~ plant.species, # has both Viola
+    sample.id == 23 ~ plant.species, # has both Viola             ## COMPLETE ##
     sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 24 ~ plant.species,
+    sample.id == 24 ~ plant.species,                              ## COMPLETE ##
     sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 25 ~ plant.species,
+    sample.id == 25 ~ plant.species,                              ## COMPELTE ##
     sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 26 ~ plant.species,
+    sample.id == 26 ~ plant.species,                              ## COMPLETE ##
     sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 27 ~ plant.species, # has both Hydrophyllum
+    sample.id == 27 ~ plant.species, # has both Hydrophyllum      ## COMPLETE ##
     sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 28 ~ plant.species,
-    sample.id == 29 ~ plant.species,
+    sample.id == 28 ~ plant.species,                              ## COMPLETE ##
+    sample.id == 29 ~ plant.species,                              ## COMPLETE ##
     sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 30 ~ plant.species,
+    sample.id == 30 ~ plant.species,                              ## COMPLETE ##
+    sample.id == 30 & Genus == 'Erigeron' ~ 'Erigeron elatior',
     sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 32 ~ plant.species,
+    sample.id == 32 ~ plant.species,                              ## COMPLETE ##
     sample.id == 33 ~ plant.species, # has both Viola
+    # Symphyotrichum foliaceum check on day 166
 
     sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  # 7.99
-    sample.id == 35 ~ plant.species, # has both Viola
-    sample.id == 36 ~ plant.species,
+    sample.id == 35 ~ plant.species, # has both Viola             ## COMPLETE ##
+
+    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  
+    sample.id == 36 ~ plant.species,                              ## COMPLETE ##
 
     sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 39 ~ plant.species, # has both hydrophyllum
-    sample.id == 40 ~ plant.species,
+    sample.id == 39 ~ plant.species, # has both hydrophyllum      ## COMPLETE ##
+    sample.id == 40 ~ plant.species,                              ## COMPLETE ##
     sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
-    sample.id == 41 ~ plant.species,
+    sample.id == 41 & Genus == 'Viola' ~ 'Viola adunca',          ## COMPLETE ##
+    sample.id == 41 ~ plant.species,                              ## COMPLETE ##
     sample.id == 42 ~ plant.species,
-    sample.id == 43 ~ plant.species,
+    sample.id == 43 ~ plant.species,                             ## COMPLETE  ##
     sample.id == 44 ~ plant.species, # 44 has both delphinium
 
+    sample.id == 45 ~ plant.species,         ## EARLY ASTER DAY 163 @ FROGS.POND ##
     sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 46 ~ plant.species,
+    sample.id == 46 ~ plant.species,         ## EARLY ASTER DAY 163 @ FROGS.POND ##
 
     sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
-    sample.id == 49 ~ plant.species,
+    sample.id == 49 & Genus == 'Viola' ~ 'Viola.adunca',
+    sample.id == 49 ~ plant.species,                              ## COMPLETE ##
 
     sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 51 ~ plant.species, # 51 has a viola split
-    sample.id == 52 ~ plant.species,
-    sample.id == 54 ~ plant.species,
+    sample.id == 51 ~ plant.species,  ## Helianthella quinquenervis @ DOY 166 ##
+    sample.id == 52 ~ plant.species,    ## Erigeron elatior  on DOY 163 @ AVERY WASH ##
+    sample.id == 54 ~ plant.species,                             ## COMPLETE ##
 
     sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 55 ~ plant.species,
+    sample.id == 55 ~ plant.species,                              ## COMPLETE ##
 
     sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
-    sample.id == 56 ~ plant.species,
-
+    sample.id == 56 ~ plant.species,          ## DOY 169 FROGS.POND Symphyotrichum foliaceum
+                                              ## & Erigeron elatior ##
+    sample.id == 60 & Genus == 'Viola' ~ 'Viola.adunca',          ## COMPLETE ##
     sample.id == 60 ~ plant.species,
-    sample.id == 65 ~ plant.species, # 65 has a mertensia split
+    sample.id == 60 & Genus == 'Viola' ~ 'Viola',                 ## COMPLETE ##
+    sample.id == 65 ~ plant.species,
 
@@
     TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Holodiscus argenteus', 'Potentilla pulcherrima'),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Scabrethia scabra', 'Wyethia arizonica'),
     TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Geum ternatum', 'Geum triflorum'),
-    TAXON_EXPERT = str_replace(
-      TAXON_EXPERT, 'Arenaria globiflora|Minuartia recurva|Sagina procumbens', 'Eremogone congesta'),
-    TAXON_EXPERT = if_else(doy < 171 & Genus == 'Cynoglossum', 
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Erigeron grandiflorus', 'Erigeron elatior'),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Agastache pallidiflora', 'Pedicularis'),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Lithophragma parviflorum', 'Lithophragma glabrum'),
+    TAXON_EXPERT = str_replace(TAXON_EXPERT, 
+                               'Arenaria globiflora|Minuartia recurva|Sagina procumbens', 'Eremogone congesta'),
+    TAXON_EXPERT = if_else(doy <= 171 & Genus == 'Cynoglossum', 
                            'Mertensia fusiformis', TAXON_EXPERT), 
@@
 
-# re class the totals to sum to 100%
+# re class the totals percent of all sequences to sum to 100%
 
@@
          Prcnt_seqs_Reclass2 = Prcnt_seqs_Reclass * multiplier)
+
+yargh <- split(blst_reclass_expert, f = blst_reclass_expert$sample.id)
 
@@
 In many instances it appears that while landscapes respond similarly to environmental variables which predict phenological responses, that individual species vary widely in their responses to similar environmental cues, or respond to different cues (@augspurger2020concordance, @xie2015deciduous, @xie2018predicting, @caradonna2014shifts).
-**As can be seen here, predictions of when a single, major phenological event occurs is already data limited, with sample size having an effect on the subset of species which we could even generate weibull estimates for.
-; check assumptions of model again and just do a quasi something** A more promising approach for the tropics may lay in circular statistics (@park2022herbarium).
+As can be seen here, predictions of when a single, major phenological event occurs is already data limited.
+A more promising approach for the tropics may lay in utilizing circular statistics (@park2022herbarium).
 

### 2023-02-21 script/manuscript_MEE.Rmd
fully integrated reads!

@@
   filter(occurrence == 0)
-# summary(logR_dat$n)
-# sum(logR_dat$n)
 
@@
 
-pmorph <- read.csv('../data/morpho_pollen.csv') 
+pmorph <- read.csv('../data/morpho_pollen.csv') %>% 
+  filter(!sample.id %in% c(45, 54))
 blst_seqs <- distinct(blst, sample.id)
@@
    aster2_samples, palyn_cnt, blst_mertensia, aster_samples)
+
+grains_counted <- pmorph %>% 
+  group_by(sample.id) %>% 
+  summarize(grains_counted = sum(sum)) %>% 
+  pull(grains_counted)
+
 ```
@@
 *APPENDIX 7*.
-From the 60 samples that were counted and based on rarefaction we **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.
-The number of counted pollen grains in each sample range from (*MIN* - 16,293, $\bar{x}$ = 2788.685, Mdn = 1453).
+From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), **all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.** The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, Mdn = `r median(grains_counted)`).
 
@@
 ```{r}
-rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected)
+rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected, grains_counted)
 ```
@@
 ```{r Build reference dataset for reclassification process}
+
+# write.csv(blst, '../data/blst4experts.csv' , row.names = F)
 
@@
 
-    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
+    sample.id == 3 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
+    sample.id == 3 & Family %in% c('Rosaceae', 'Onagraceae') ~  'Potentilla.pulcherrima',
+    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,       ## COMPLETE ##
     sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 4 ~ plant.species,                               ## COMPLETE ##
-
-    sample.id == 6 ~ plant.species, # has a delphinium split
-    # Symphyotrichum foliaceum check on day 173
-
-    sample.id == 9 ~ plant.species, # has both Delphinium # Symphyotrichum foliaceum check on day 176
-
+    sample.id == 6 ~ plant.species,                               ## COMPLETE ##
+    sample.id == 9 & TAXON_NEW == 'Symphyotrichum foliaceum' ~ 'Erigeron speciosus',
+    sample.id == 9 ~ plant.species,                               ## COMPLETE ##
+
+    sample.id == 10 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
+  #  sample.id == 10 & Family == 'Onagraceae' ~ ''
     sample.id == 10 ~ plant.species, # has many late season taxa
 
-    sample.id == 14 ~ plant.species,  # # Symphyotrichum foliaceum check on day 165
-
-    sample.id == 15 ~ plant.species, ## has many mysterious Rosids @ Gothic Spires  - call potentilla?
-
-    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', # & Ericaceae
-    sample.id == 18 ~ plant.species, # Erigeron grandiflorus on day 166 @ Frogs Pond
-    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri', # multiple Onagraceae
-    sample.id == 19 ~ plant.species, # 
+    sample.id == 14 & Genus == 'Symphyotrichum' ~ 'Senecio.integerrimus',
+    sample.id == 14 ~ plant.species,                              ## COMPLETE ##
+    sample.id == 15 & Family %in% c('Onagraceae', 'Rosaceae')  ~ 'Potential pulcherrima',
+    sample.id == 15 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
+    sample.id == 15 ~ plant.species,                              ## COMPLETE ## 
+    sample.id == 18 & Family == 'Ericaceae' ~ 'Ericaceae', # ONLY TO FAMILY
+    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
+    sample.id == 18 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
+    sample.id == 18 ~ plant.species,                              ## COMPLETE ##
+
+    sample.id == 19 & Family == 'Onagraceae' ~ 'Epilobium',
+    sample.id == 19 & Genus == 'Pseudognaphalium' ~ 'Senecio.integerrimus',
+    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri',
+    sample.id == 19 ~ plant.species,                              ## COMPLETE ##
     sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
-    sample.id == 20 ~ plant.species, # Erigeron grandiflorus doy 165 @ Avery Wash
-    sample.id == 21 ~ plant.species, # Erigeron grandiflorus & cirsium parryi doy 165 @ Gothic Spires
-
+    sample.id == 20 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
+    sample.id == 20 ~ plant.species,                              ## COMPLETE ##
+    sample.id == 21 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
+    sample.id == 21 ~ plant.species,                              ## COMPLETE ##
     sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 23 ~ plant.species, # has both Viola             ## COMPLETE ##
+    sample.id == 23 ~ plant.species,                              ## COMPLETE ##
     sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
@@
     sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 27 ~ plant.species, # has both Hydrophyllum      ## COMPLETE ##
+    sample.id == 27 ~ plant.species,                              ## COMPLETE ##
     sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
@@
     sample.id == 32 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 33 ~ plant.species, # has both Viola
-    # Symphyotrichum foliaceum check on day 166
+
+    sample.id == 33 ~ plant.species, # Symphyotrichum foliaceum check on day 166
 
     sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 35 ~ plant.species, # has both Viola             ## COMPLETE ##
-
+    sample.id == 35 ~ plant.species,                              ## COMPLETE ##
     sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  
     sample.id == 36 ~ plant.species,                              ## COMPLETE ##
-
     sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 39 ~ plant.species, # has both hydrophyllum      ## COMPLETE ##
+    sample.id == 39 ~ plant.species,                              ## COMPLETE ##
     sample.id == 40 ~ plant.species,                              ## COMPLETE ##
@@
     sample.id == 43 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 44 ~ plant.species, # 44 has both delphinium
-
-    sample.id == 45 ~ plant.species,         ## EARLY ASTER DAY 163 @ FROGS.POND ##
+    sample.id == 44 ~ plant.species,
+    sample.id == 45 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
+    sample.id == 45 ~ plant.species,                              ## COMPLETE ##
+    sample.id == 46 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
     sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 46 ~ plant.species,         ## EARLY ASTER DAY 163 @ FROGS.POND ##
-
+    sample.id == 46 ~ plant.species,                              ## COMPLETE ##
     sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
@@
     sample.id == 52 ~ plant.species,    ## Erigeron elatior  on DOY 163 @ AVERY WASH ##
+
     sample.id == 54 ~ plant.species,                              ## COMPLETE ##
-
     sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
     sample.id == 55 ~ plant.species,                              ## COMPLETE ##
-
     sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
-    sample.id == 56 ~ plant.species,          ## DOY 169 FROGS.POND Symphyotrichum foliaceum
-                                              ## & Erigeron elatior ##
+    sample.id == 56 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
+    sample.id == 56 ~ plant.species,                              ## COMPLETE ##
     sample.id == 60 & Genus == 'Viola' ~ 'Viola.adunca',          ## COMPLETE ##
     sample.id == 60 ~ plant.species,
-    sample.id == 60 & Genus == 'Viola' ~ 'Viola',                 ## COMPLETE ##
+    sample.id == 65 & Genus == 'Viola' ~ 'Viola.praemorsa',       ## COMPLETE ##
     sample.id == 65 ~ plant.species,
@@
 
-rm(ranunc_splash, blst_expert)
+rm(ranunc_splash)
 
@@
 rm(s55)
+
+# we will split the reads for the Cynoglossum detected in sample 3.
+
+blst_reclass_expert <- blst_reclass_expert %>% 
+  filter(sample.id == 3 & TAXON_EXPERT == 'Cynoglossum pringlei') %>% 
+  slice(rep(1:n(), each=2)) %>% 
+  mutate(Prcnt_seqs_Reclass = Prcnt_seqs_Reclass/2,
+         dummyID = 1:n(),
+         TAXON_EXPERT = if_else(dummyID == 1, 'Mertensia fusiformis', 'Mertensia cilita')) %>% 
+  select(-dummyID) %>% 
+  bind_rows(., filter(
+              blst_reclass_expert, sample.id == 3 & TAXON_EXPERT != 'Cynoglossum pringlei'),
+            filter(blst_reclass_expert, sample.id != 3))
 
@@
 palustris*.
+
+Regarding limitations of morphological data we suspect that there were two morphotypes of pollen identified as Ericaceae Juss. were actually Onagraceae (Samples 19 & 44), based on molecular results.
 

### 2023-02-21 script/pollen_countr3.Rmd
fully integrated reads!

@@
     morphotype = str_replace(morphotype, 'UNKNOWN_[(]D[)]', 'D'),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]GROSSULARIACEAE[)]', 'GROSSULARIACEAE'),
     morphotype = str_replace(morphotype, 'OTHER_B', 'B'),
@@
     morphotype = str_replace(morphotype, "C\\_\\(LINUM\\)", "LINUM"),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]AMSONIA[?)]', 'UNKNOWN'),
     morphotype = str_replace(morphotype, 'UNKNOWN_[(]RANUNCULUS[)]', 'UNKNOWN'),
@@
     morphotype = str_replace(morphotype, 'UNKNOWN_[(]4_APERTURES[)]', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, '^UNKNOWN.*$', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, 'UNKNOWN_[(]UNKNOWN[)]', 'UNKNOWN'),
+    morphotype = str_replace(morphotype, 'PEDICULARISUNKNOWN', 'PEDICULARIS'),
     morphotype = str_replace_all(morphotype, " ", "_"),

### 2023-02-21 script/expert_opinion_reclassification.R

@@ -1,0 +1,462 @@
+library(tidyverse)
+
+setwd('~/Documents/ManuSCript/script')
+
+pmorpho <- read.csv('../data/morpho_pollen.csv')
+
+# this is an export of the variable 'blst' as would be expected on line 717  of the
+# manuscript 
+
+blst <- read.csv('../data/blst4experts.csv')
+
+################################################################################
+##############       RECLASIFY READS USING EXPERT DATA SET      ################
+################################################################################
+
+## Load in the microscopy count datasheets to gather information on sample provenance
+pollen_sample_info <- readxl::read_xlsx(file.path('../data', 
+                                                  list.files(path = '../data', pattern = 'xlsx') )) %>% 
+  select(site, date, sample.id) %>% 
+  distinct() %>% 
+  filter(!date %in% c('42170', '42169'))
+
+
+files <- list.files(path = '../data')
+
+## load in the floral observations to provide high resolution spatial and temporal information
+flr_date_ranges <- read.csv(
+  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
+  select(site, doy, week, plant.species:abun.rank) %>% 
+  group_by(plant.species, site) %>% 
+  select(site, week, doy, plant.species) %>% 
+  mutate(Genus = str_extract(plant.species, '^.+?(?=[.]|$)')) %>% 
+  mutate(Genus = case_when(
+    Genus == 'Erythrocoma' ~ 'Geum', 
+    Genus == 'Amerosedum' ~ 'Sedum', 
+    Genus == 'Pentaphylloides' ~ 'Dasiphora',
+    TRUE ~ Genus
+  )) %>% 
+  mutate(plant.species = case_when(
+    plant.species == 'Lathyrus.leucanthus' ~ 'Lathyrus.lanszwertii',
+    plant.species == 'Lupinus.bakeri' ~ 'Lupinus.sericeus',
+    TRUE ~ plant.species
+  )) %>% distinct()
+
+rm(files)
+## create a look up table to be able to join the known plant locations to 
+## sequence data by WEEK
+week_tab <- flr_date_ranges %>% 
+  ungroup() %>% 
+  select(doy, week) %>% 
+  distinct()
+
[…]

### 2023-02-21 script/manuscript_MEE.Rmd
final touches for the night

@@
 
+### 2.7 | Integrated Observational, Molecular, and Palynological Corbiculae
+
+In order to precisely classify the contents of each corbiculae load the sequences classified by molecular methods were compared with the fieldwork which at a very fine resolution, recorded the presence and absence of species and their duration of flowering, and was interpreted ala the computer derived temporal and spatial data sets.
+The quantitative counts of grains from microscopy, were combined with the semi-quantitative sequencing results, to estimate the abundance of each identified species in each corbiculae load.
+
+To reclassify the sequence reads, these data were combined with the flora observation data, and mapped by genus.
+If more than one species in the genus was flowering at that time and site, than the reads were split evenly between the taxa.
+For sequence data which did not match at the genus level, a user subjectively scored them based on the species composition and phenological activity at each plot, the Queen interaction data, and pollen assignments.
+To estimate the abundance of each of these species in the corbiculae loads, these data were combined with the microscopy data.
+For each morphotype detected in pollen, and each classified sequence read which was not detected via microscopy, they were given a value of 0.5% to indicate there trace presences.
+When more than a single species belonged to a morphotype group in a single sample, the quantitative values from the morphological work were multiplied by the relative sequence abundance of each species in the load.
+All final compositions were standardized to a sum of 100%, by adding or subtracting the differences (induced by classifying records as 'trace') to all species with abundances > 1%.
+
 # 3 | RESULTS
@@
          resource.coll:pollen.collected, fl.switch.1:fl.switch.2.resource.coll) %>% 
-  mutate(across(resource.coll:fl.switch.2.resource.coll, ~ na_if(.x, ''))) %>% 
+  mutate(across(where(is.character), ~ na_if(., ''))) %>% 
   pivot_longer(c(plant.species, fl.switch.1, fl.switch.2), names_to = 'observation', values_to = 'plant.species') %>% 
@@
 The floral morphology of *Thalictrum*  spp.
-also makes it unlikely to be visited (BELEIVE JANE SAID THIS- IF NOT STRIKE OUT), and while evidence of visits to Caltha and Trollius are lacking, due to the association between the reads these results appear unlikely.
+also makes it unlikely to be visited (BELEIVE JANE SAID THIS- IF NOT STRIKE OUT), and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
 
@@
 ```{r Build reference dataset for reclassification process}
-
-# write.csv(blst, '../data/blst4experts.csv' , row.names = F)
-
-pollen_sample_info <- readxl::read_xlsx(file.path('../data', 
-  list.files(path = '../data', pattern = 'xlsx') )) %>% 
-  select(site, date, sample.id) %>% 
-  distinct() %>% 
-  filter(!date %in% c('42170', '42169'))
-
-files <- list.files(path = '../data')
-
-flr_date_ranges <- read.csv(
-  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
-  select(site, doy, week, plant.species:abun.rank) %>% 
-  group_by(plant.species, site) %>% 
-  select(site, week, doy, plant.species) %>% 
-  mutate(Genus = str_extract(plant.species, '^.+?(?=[.]|$)')) %>% 
-  mutate(Genus = case_when(
-    Genus == 'Erythrocoma' ~ 'Geum', 
-    Genus == 'Amerosedum' ~ 'Sedum', 
-    Genus == 'Pentaphylloides' ~ 'Dasiphora',
-    TRUE ~ Genus
-    )) %>% 
-  mutate(plant.species = case_when(
-    plant.species == 'Lathyrus.leucanthus' ~ 'Lathyrus.lanszwertii',
-    plant.species == 'Lupinus.bakeri' ~ 'Lupinus.sericeus',
-    TRUE ~ plant.species
-  )) %>% distinct()
- # this provides us with a data.frame of the ACTUAL flowering dates!
-
-week_tab <- flr_date_ranges %>% 
-  ungroup() %>% 
-  select(doy, week) %>% 
-  distinct()
-
-flr_date_ranges <- select(flr_date_ranges, -doy)
-
-blst_expert <- blst %>% 
-  select(species, sample.id, TAXON_NEW, Genus, Family, Prcnt_seqs, Prcnt_Seqs_Class) %>% 
-  mutate(RecordID = 1:n()) %>% 
-  left_join(., pollen_sample_info,  by = 'sample.id') %>% 
-  relocate(site:date, .before = TAXON_NEW) %>% 
-  mutate(doy = lubridate::yday(date)) %>% 
-  left_join(., week_tab, by = 'doy') %>% 
-  left_join(., flr_date_ranges, by = c('site', 'Genus', 'week')) %>% 
-  relocate(sample.id, .before = plant.species)
-
-rm(corb, week_tab, flr_date_ranges, pollen_sample_info)
-
-  # reduce to temporally similar results
-
-######## SAMPLE 55 SHOULD BE SOMEWHERE HERE. 
-duplicate_reps_in_sample <- blst_expert %>%
-
-  # identify the duplicate samples
-  drop_na(plant.species) %>% 
-  group_by(RecordID) %>% 
-  distinct(plant.species, .keep_all = T) %>% 
-  group_by(sample.id, TAXON_NEW) %>% 
-  filter(n() >= 2)
-
-dupe_ids <- duplicate_reps_in_sample %>% pull(RecordID) %>% unique()
-dupe_samples <- duplicate_reps_in_sample %>% pull(sample.id) %>% unique()
-
-  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Solomon'
-duplicate_reps_in_sample <- duplicate_reps_in_sample %>% 
-  ungroup() %>% 
-  group_by(sample.id, Genus) %>% 
-  mutate(Prcnt_seqs_Reclass = 
-           ifelse(n() == 2, Prcnt_seqs/2, sum(Prcnt_seqs)/n_distinct(plant.species)/2 ),
-         TAXON_EXPERT = plant.species) %>% 
-  distinct(sample.id, plant.species, .keep_all = T)
-
-ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia')
-
-blst_reclass_expert <- blst_expert %>% 
-  filter(!RecordID %in% dupe_ids) %>% 
-  distinct() %>% 
-  rowwise() %>% 
-  mutate(Genus = str_trim(Genus)) %>%
-
-  mutate(TAXON_EXPERT = case_when(
-
-    sample.id == 3 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
-    sample.id == 3 & Family %in% c('Rosaceae', 'Onagraceae') ~  'Potentilla.pulcherrima',
-    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,       ## COMPLETE ##
-    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 4 ~ plant.species,                               ## COMPLETE ##
-    sample.id == 6 ~ plant.species,                               ## COMPLETE ##
-    sample.id == 9 & TAXON_NEW == 'Symphyotrichum foliaceum' ~ 'Erigeron speciosus',
-    sample.id == 9 ~ plant.species,                               ## COMPLETE ##
-
-    sample.id == 10 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
-  #  sample.id == 10 & Family == 'Onagraceae' ~ ''
-    sample.id == 10 ~ plant.species, # has many late season taxa
-
-    sample.id == 14 & Genus == 'Symphyotrichum' ~ 'Senecio.integerrimus',
-    sample.id == 14 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 15 & Family %in% c('Onagraceae', 'Rosaceae')  ~ 'Potential pulcherrima',
-    sample.id == 15 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 15 ~ plant.species,                              ## COMPLETE ## 
-    sample.id == 18 & Family == 'Ericaceae' ~ 'Ericaceae', # ONLY TO FAMILY
-    sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 18 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 18 ~ plant.species,                              ## COMPLETE ##
-
-    sample.id == 19 & Family == 'Onagraceae' ~ 'Epilobium',
-    sample.id == 19 & Genus == 'Pseudognaphalium' ~ 'Senecio.integerrimus',
-    sample.id == 19 & Genus == 'Nemophila' ~ 'Hydrophyllum.fendleri',
-    sample.id == 19 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 20 & TAXON_NEW == 'Hydrophyllum canadense' ~ 'Hydrophyllum fendleri',
-    sample.id == 20 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 20 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 21 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 21 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 23 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 24 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 25 ~ plant.species,                              ## COMPELTE ##
-    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 26 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 27 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 28 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 29 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 30 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 30 & Genus == 'Erigeron' ~ 'Erigeron elatior',
-    sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 32 ~ plant.species,                              ## COMPLETE ##
-
-    sample.id == 33 ~ plant.species, # Symphyotrichum foliaceum check on day 166
-
-    sample.id == 35 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 35 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 36 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis',  
-    sample.id == 36 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 39 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 39 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 40 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata', 
-    sample.id == 41 & Genus == 'Viola' ~ 'Viola adunca',          ## COMPLETE ##
-    sample.id == 41 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 42 ~ plant.species,
-    sample.id == 43 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 44 ~ plant.species,
-    sample.id == 45 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 45 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 46 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 46 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 46 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 49 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',
-    sample.id == 49 & Genus == 'Viola' ~ 'Viola.adunca',
-    sample.id == 49 ~ plant.species,                              ## COMPLETE ##
-
-    sample.id == 51 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
-    sample.id == 51 ~ plant.species,  ## Helianthella quinquenervis @ DOY 166 ##
-    sample.id == 52 ~ plant.species,    ## Erigeron elatior  on DOY 163 @ AVERY WASH ##
-
-    sample.id == 54 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 55 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
-    sample.id == 55 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',
-    sample.id == 56 & Family == 'Asteraceae' ~ 'Senecio.integerrimus',
-    sample.id == 56 ~ plant.species,                              ## COMPLETE ##
-    sample.id == 60 & Genus == 'Viola' ~ 'Viola.adunca',          ## COMPLETE ##
-    sample.id == 60 ~ plant.species,
-    sample.id == 65 & Genus == 'Viola' ~ 'Viola.praemorsa',       ## COMPLETE ##
-    sample.id == 65 ~ plant.species,
-
-   TRUE ~ as.character(TAXON_NEW)
-  ))  %>% 
-  mutate( 
-    TAXON_EXPERT = str_replace_all(TAXON_EXPERT, '[.]', ' '),
-    TAXON_EXPERT = if_else(is.na(TAXON_EXPERT), TAXON_NEW, TAXON_EXPERT),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Holodiscus argenteus', 'Potentilla pulcherrima'),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Scabrethia scabra', 'Wyethia arizonica'),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Geum ternatum', 'Geum triflorum'),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Erigeron grandiflorus', 'Erigeron elatior'),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Agastache pallidiflora', 'Pedicularis'),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 'Lithophragma parviflorum', 'Lithophragma glabrum'),
-    TAXON_EXPERT = str_replace(TAXON_EXPERT, 
-                               'Arenaria globiflora|Minuartia recurva|Sagina procumbens', 'Eremogone congesta'),
-    TAXON_EXPERT = if_else(doy <= 171 & Genus == 'Cynoglossum', 
-                           'Mertensia fusiformis', TAXON_EXPERT), 
-    TAXON_EXPERT = if_else(doy < 190 & TAXON_EXPERT == 'Lupinus argenteus', 
-                           'Lupinus sericeus', TAXON_EXPERT))  %>%
-
-  # use this to sum up the total sequences by original read record
-  distinct(sample.id, TAXON_NEW, .keep_all = T) %>% 
-  group_by(sample.id, TAXON_EXPERT) %>% 
-  mutate(Prcnt_seqs_Reclass = sum(Prcnt_seqs)) %>%
-
-  # now reduce to the amount of rows of the proper classification
-  ungroup() %>% 
-  distinct(sample.id, TAXON_EXPERT, .keep_all = T) %>% # what is going on with sample 56 then?
-
-  # add back on the multiple groups
-  bind_rows(., duplicate_reps_in_sample)  %>% 
-  arrange(sample.id, -Prcnt_seqs_Reclass)
-
-rm(ranunc_splash)
-
-# now we split up the reads assigned to conflicted groups  # e.g. Cynoglossum amplifolium in sample 65
-b <- blst_reclass_expert %>% 
-  filter(sample.id %in% dupe_samples) %>% 
-  group_by(sample.id, Family) %>% 
-  filter(n() >= 3) %>% 
-  mutate(plant.species = na_if(plant.species, "Thalictrum.fendleri")) %>% 
-  mutate(Seqs2distribute = sum(Prcnt_seqs_Reclass))
-
-records2drop <- b %>% pull(RecordID)
-
-b <- b %>% 
-  drop_na(plant.species) %>% 
-  mutate(Prcnt_seqs_Reclass = Seqs2distribute/n_distinct(TAXON_EXPERT)) %>% 
-  select(-Seqs2distribute)
-
-blst_reclass_expert <- blst_reclass_expert %>% 
-  filter(!RecordID %in% records2drop) %>% 
-  bind_rows(., b) %>% 
-  arrange(sample.id, -Prcnt_seqs_Reclass)
-
-rm(records2drop, b)
-
-# 55 SHOULD NOT BE IN HERE, HOW IS NOT CAUGHT IN THE ABOVE EXCEPTION???? #
-# seems i may not have either delphinium present in that week? and  if falls between two. 
-# deal with sample 55 one off, split the delphinium gracile
-
-s55 <- blst_reclass_expert %>% 
-  filter(sample.id == 55) %>% 
-  group_by(Genus) %>% 
-  mutate(Seqs2distribute = sum(Prcnt_seqs_Reclass)) %>% 
-  filter(TAXON_EXPERT != 'Delphinium gracile') %>% 
-  mutate(Prcnt_seqs_Reclass = Seqs2distribute/n_distinct(TAXON_EXPERT)) %>% 
-  select(-Seqs2distribute)
-
-blst_reclass_expert <- blst_reclass_expert %>% 
-  filter(sample.id != 55) %>% 
-  bind_rows(., s55)  %>% 
-  arrange(sample.id, -Prcnt_seqs_Reclass)
-
-rm(s55)
-
-# we will split the reads for the Cynoglossum detected in sample 3.
-
-blst_reclass_expert <- blst_reclass_expert %>% 
-  filter(sample.id == 3 & TAXON_EXPERT == 'Cynoglossum pringlei') %>% 
-  slice(rep(1:n(), each=2)) %>% 
-  mutate(Prcnt_seqs_Reclass = Prcnt_seqs_Reclass/2,
-         dummyID = 1:n(),
-         TAXON_EXPERT = if_else(dummyID == 1, 'Mertensia fusiformis', 'Mertensia cilita')) %>% 
-  select(-dummyID) %>% 
-  bind_rows(., filter(
-              blst_reclass_expert, sample.id == 3 & TAXON_EXPERT != 'Cynoglossum pringlei'),
-            filter(blst_reclass_expert, sample.id != 3))
-
-# re class the totals percent of all sequences to sum to 100%
-
-blst_reclass_expert <- blst_reclass_expert %>% 
-  group_by(sample.id) %>% 
-  mutate(multiplier = 100/ sum(Prcnt_seqs_Reclass),
-         Prcnt_seqs_Reclass2 = Prcnt_seqs_Reclass * multiplier)
-
-yargh <- split(blst_reclass_expert, f = blst_reclass_expert$sample.id)
-
-rm(pollen_sample_info)
+reclassified <- read.csv('../data/Fully_Integrated_Corbiculae.csv')
 ```
@@
 ```{r}
-a <- read.csv('../data/herbaria_4_acknowledgements.csv') %>% 
-  pull() 
+a <- pull(read.csv('../data/herbaria_4_acknowledgements.csv'))
 ```

### 2023-02-21 script/expert_opinion_reclassification.R

@@ -335,7 +335,6 @@ pmorpho <- pmorpho %>%
   mutate(Percent = (sum / sum(sum)) * 100) %>% 
   ungroup()
 
-
 a <- c('Trifolium', 'Lupinus', 'Glyccrrhiza', 'Mitella', 'Geum')
 b <- c('Lupinus', 'Lathyrus', 'Potentilla', 'Androsace', 'Bistorta', 'Vicia',
        'Dasiphora') # while not prepared, we include this with Potentilla
@@ -436,6 +435,8 @@ quant_read_final <- bind_rows(quant_read_low, quant_read_high) %>%
 
 rm(quant_read_low, quant_read_high, quant_read)
 # PROCESS
+
+write.csv(quant_read_final, '../data/Fully_Integrated_Corbiculae.csv', row.names = F)
 
 # 1) Round up all counted morphotypes from < 0.5 to 0.5 to serve as TRACE quantities
 # 2) Join Pollen count data with sequence data

### 2023-02-25 script/expert_opinion_reclassification.R
had old csvs in for BLAST it never ends

@@ -7,7 +7,10 @@ pmorpho <- read.csv('../data/morpho_pollen.csv')
 # this is an export of the variable 'blst' as would be expected on line 717  of the
 # manuscript 
 
-blst <- read.csv('../data/blst4experts.csv')
+blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  group_by(Sample) %>% 
+  slice_max(Prcnt_seqs, n = 10) %>% 
+  ungroup()
 
 ################################################################################
 ##############       RECLASIFY READS USING EXPERT DATA SET      ################
@@ -55,7 +58,7 @@ flr_date_ranges <- select(flr_date_ranges, -doy)
 ## Create a data frame which contains the sample provenance and sequencing data
 ## information
 blst_expert <- blst %>% 
-  select(species, sample.id, TAXON_NEW, Genus, Family, Prcnt_seqs, Prcnt_Seqs_Class) %>% 
+  select(species = TAXON, sample.id = Sample, TAXON_NEW, Genus = GENUS, Family = FAMILY, Prcnt_seqs) %>% 
   mutate(RecordID = 1:n()) %>% 
   left_join(., pollen_sample_info,  by = 'sample.id') %>% 
   relocate(site:date, .before = TAXON_NEW) %>% 
@@ -94,8 +97,7 @@ duplicate_reps_in_sample <- duplicate_reps_in_sample %>%
 
 ## define a list of splash out sequences, note the lsat 2 are not obligates with
 ## delphinium, and will not always be substituted out!!
-ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia')
-
+ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia', 'Corydalis')
 
 # we will now manually reconsider EVERY single sequence classification in 
 # light of the field based data. AND floral interactions observations
@@ -109,38 +111,50 @@ blst_reclass_expert <- blst_expert %>%
     
     sample.id == 3 & Family == 'Asteraceae' ~  'Taraxacum.officinale',
     sample.id == 3 & Family %in% c('Rosaceae', 'Onagraceae') ~  'Potentilla.pulcherrima',
[…]

### 2023-02-26 script/manuscript_MEE.Rmd
importing final data for comparisions

@@
 
-```{r Build reference dataset for reclassification process}
-reclassified <- read.csv('../data/Fully_Integrated_Corbiculae.csv')
+```{r compare results}
+
+blast <- read.csv('../data/tally_blast_classifications.csv') %>% 
+  group_by(Sample) %>% 
+  slice_max(Prcnt_seqs, n = 10) %>% 
+  select(Sample, Taxon = taxid, Percent_Sample = Prcnt_seqs) %>% 
+  mutate(
+    Taxon = str_replace(Taxon, 'Arenaria', 'Eremogone'),
+    Genus = str_remove(str_extract(Taxon, '^.* '), ' '),
+    Method = 'BLAST')
+
+automated <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  group_by(Sample, TAXON_NEW) %>%
+  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
+  drop_na() %>% 
+  distinct(Sample, TAXON_NEW, .keep_all = T) %>% 
+  ungroup(TAXON_NEW) %>% 
+  slice_max(Prcnt_seqs, n = 10) %>% 
+  select(Sample, Genus = GENUS, Taxon = TAXON_NEW, Percent_Sample = Prcnt_seqs) %>% 
+  mutate(across(where(is.character), ~ str_replace(., 'Arenaria', 'Eremogone')),
+         Method = 'Automated')
+
+expert <- read.csv('../data/Fully_Integrated_Corbiculae.csv') %>% 
+  rename(Sample = sample.id) %>% 
+  mutate(across(where(is.character), ~ str_trim(.)), 
+         Method = 'Expert') %>% 
+  drop_na() %>% 
+  relocate()
+
+genera2query <- na.omit(unique(c(blast$Genus, automated$Genus, expert$Genus))) 
+genera2query <- genera2query[genera2query != 'Ericaceae'] 
+genera2query <- taxizedb::name2taxid(genera2query, out_type = 'summary')
+
+genera_out <- taxizedb::classification(genera2query$id, db='ncbi', rank = 'family') 
+genera_out <- map(genera_out, ~ 
+                    data.frame( 
+                      'Family' = filter(.x, rank == "family") %>% select(name), 
+                      'Genus' = filter(.x, rank == 'genus') %>%  select(name) 
+                      )) %>% 
+  bind_rows() %>% 
+  rename(Family = 1, Genus = 2)
+
+all_classifications <- bind_rows(
+  automated, blast, expert
+) %>% 
+  left_join(., genera_out, by = 'Genus') %>% 
+  mutate(Rank = case_when(
+    is.na(Family) ~ 'Family',
+    Taxon == Family ~ 'Family',
+    Taxon == Genus ~ 'Genus',
+    TRUE ~ 'Species'
+  ),
+  Family = if_else(Taxon == 'Ericaceae', 'Eriaceae', Family))
+
+rm(genera2query, genera_out, automated, blast, expert)
 ```

### 2023-02-26 script/expert_opinion_reclassification.R

@@ -137,6 +137,7 @@ blst_reclass_expert <- blst_expert %>%
     sample.id == 15 & Genus %in% c('Cirsium', 'Pseudognaphalium') ~ 'Senecio.integerrimus',
     sample.id == 15 ~ plant.species,                              ## COMPLETE ## 
     
+    sample.id == 17 & Genus == 'Lupinus' ~ 'Lupinus.sericeus',
     
     sample.id == 18 & Family == 'Ericaceae' ~ 'Ericaceae', # ONLY TO FAMILY
     sample.id == 18 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
@@ -156,6 +157,7 @@ blst_reclass_expert <- blst_expert %>%
     sample.id == 23 & Genus == 'Delphinium' ~ 'Delphinium barbeyi',
     sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium barbeyi',
     sample.id == 23 ~ plant.species,                              ## COMPLETE ##
+    sample.id == 23 & Genus == 'Delphinium' ~ 'Delphinium.nuttallianum',
     sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 24 ~ plant.species,                              ## COMPLETE ##
     sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
@@ -172,6 +174,10 @@ blst_reclass_expert <- blst_expert %>%
     sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
     sample.id == 30 ~ plant.species,                              ## COMPLETE ##
     sample.id == 30 & Genus == 'Erigeron' ~ 'Erigeron elatior',
+    
+    sample.id == 31 & Genus == 'Delphinium' ~ 'Delphinium barbeyi',
+    sample.id == 31 ~ plant.species,                              ## COMPLETE ##
+    
     sample.id == 32 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
     sample.id == 32 ~ plant.species,                              ## COMPLETE ##
     
@@ -256,6 +262,9 @@ blst_reclass_expert <- blst_expert %>%
 
 rm(ranunc_splash)
 
+# now hold the 
+
+
[…]

### 2023-03-02 script/conceptual-tables-boxes.rmd
many edits from JEO PJC, and initiation of table

---
title: "Conceptual Boxes"
author: "steppe"
date: "2023-03-01"
output: html_document
---

Here we create three text boxes for the adjacent manuscript. These serve to illustrate three main ideas which we found difficult to incorporate into the written text. We also produce one diagram, which serves to outline the entirety of the process we engaged in.

1) The Applications of Meta-barcoding with Plants

```{r Applications of metabarcoding in plants}

applications <- data.frame(
  'Application' = c(
    'Plant-Animal Interactions',
    'Noxious Weed Detection',
    'Detection of Rare Species', 
    'Forensic Science',
    'Pharmaceutical/Nutraceutical'
  ),
  'Example' = c(

    'Dietary Preferences of herbivores',
    'Presence of species by hydrologic Basin',
    'Presence of Rare Aquatic Species',
    'Identifying the provenance of materials found at crime scene',
    'Identifying adulterants in wholesale products'
  ),
  'Citation' = c(
    'Soininen et al. 2009',
    'Coghlan et al. 2021',
    'Tsukamoto et al. 2021',
    'Allwood et al. 2020',
    'Bell et al. 2022'
  )
)
```

2) Major Issues and bottlenecks currently related to plant Meta-barcoding.

```{r Major Issues}

bottlenecks <- tibble(

  'Issue' = c(
    'Taxonomic Resolution',
    'Reference Library Coverage - Phylogenetic',
    'Reference Library Coverage - Spatial',
    'Reference Library Generation', 
    'Uncertainty with Matches',
    'Species Surrogates',
    'False Positives'
  ),

  'Approach' = c(
    'A353', 
    'Kew PAFTOL, no phylogenetic biases', 
    'Some bias persists towards Europe, North America',
    'Spatial Modelling; Code within', 
    'Temporal Filter System',
    'Temporal Filter System',
    'Spatial & Temporal Modelling, Jaccard Index, high quality reference loci'
  ),

  'Citation' = c(
    c('Coissoc et al. 2016, Kress 2017'),  #taxonomic resolution
    c('Kress 2017, Bell et al. 2022'),  # reference library phylogenetic
    c('Cheng et al. 2018, Darwin Tree of Life 2022, Lewin et al. 2020, Bell et al. 2021'),  # 'Reference Library spatial
    c('Bell et al. 2022'), #  'Reference Library Generation', 
    c('Bell et al. 2022'), #   'Uncertainty with Matches',
    c('?'), #   'Species Surrogates',
    c('Bell et al. 2021') #   'False Positives'
  )

)

```

3) The Summary of Limitations which each method we utilized contains

```{r Summary of Limitations}

limitations <- data.frame(

  'Component' = c(
    'Stage 1 Species Filter',
    'Species Distribution Modelling',
    'Phenological Modelling',
    'Database Generation',
    'Read Re-assignment',
    'False Positives',
    'Semi-Quantitative Inference'
  ),

  'Limitation' = c(
    'Test Data', # stage 1 species filter
    'Number of Records; Taxonomically Difficult Groups', # SDM generation
    'Post-Initiation of Climate Change Records', # phenological modelling
    'Adequate Phylogenetic/Spatial Representation', # database generation
    'Discrete Frequentist Data', # read re-assignment
    'Which True Species?', # false positives
    'Genome Size, Pollen Grain Size?' # semi-quantitative inference
  ),

  'Paths_Forward' = c(
    'Flash Plant Species Surveys on Plot', 
    'Develop and Disseminate Education Materials; Herbaria Collections', 
    'Advocate Herbarium Collections', 
    'Plant and Fungal Tree of Life; ', 
    'Posterior-Probabilities; Floral Abundance, Nectar/Pollen Nutrition', 
    'Jaccard Index, Plot Abundance',
    'Spike Samples with Reference Materials; Several Genome Sizes'
  )
)
```

4) Diagram of Major conceptual 'filtering steps'

### 2023-03-02 script/manuscript_MEE.Rmd
many edits from JEO PJC, and initiation of table

@@
 author:  |
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^, Emily J. Woodworth$^1$$^,$$^2$, 
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
     |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
-    |  $^4$ Australia post fire project, ----, ----, ----, AU  
-    |  Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$, 
 abstract:  |  
   |  
-  1) DNA Barcoding has been successful for the rapid analysis of ecological assemblages.
-Although barcoding in the plant kingdom barcoding has been more difficult than others, and hence may begin to lag behind other kingdoms.
-  2) Here we test the utilization of Angiosperms 353 probes to barcode plant species found in pollen loads collected from Queen Bumble Bees. 
-  3) To verify the accuracy for this barcoding system we compared the data to museum species, observation studies, and species distribution modelling to identify likely candidate species. 
+  1) DNA Barcoding has been successful for the rapid analysis of complex ecological assemblages.
+Barcoding in the plant kingdom barcoding has been more difficult than in others, and hence may begin to lag behind other kingdoms.
+  2) Here we test the utilization of Angiosperms 353 probes to barcode plant species found in corbiculae pollen loads collected from Queen Bumble Bees. 
+  3) To verify the accuracy of this barcoding system we compared the barcoding data to museum plant specimens, observations of plant flowering and bee flower visitation in the field, and species distribution modelling to identify likely candidate plant species found within the corbiculae pollen loads.
   4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which may use increase the alignment algorithms minimizing need for large computational power, and run time.
-  5) We show that the Angiosperms 353 probes, which are currently being used in the largest ever plant systematic endeavor, offers significant promise to metagenomic approaches. 
-  6) Understanding plants in ecological contexts and understandings of their synecology.
+  5) We show that the Angiosperms 353 probes, developed for phylogenmoics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches.
+  6) The DNA barcoding of bumble bee corbiculae pollen loads was most accurate when combined with knowledge of what plant species were flowering in the plant community when they were collected.
+Thus, supplementing DNA barcoding data with ecological context is most accurate and powerful.
 keywords: |
@@
 different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).In these instances the lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and their interactions with other species.
-This can be further complicated by the fact that the identification of organisms to terminal taxon is also often mired by lack of diagnostic characters(e.g.
+This can be further complicated by the fact that the identification of organisms to species is also often mired by lack of diagnostic characters(e.g.
 flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and the presence of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
-Taxonomic verification can also be limited by the fact that revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often does not work.
-The current methods to ameliorate this situation include: ignoring these ecologically relevant levels of detail, revisiting plots as diagnostic material becomes temporally available, seeking the assistance from taxonomic specialists, or the use of barcoding using molecular techniques (CITE)
-
-  Recently molecular barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), or metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in many taxa (@ruppert2019past).
-For plants the success is a little more mix, with the identification of certain clades using barcoding being quite successful (@kress2017plant), however for many other clades the results have been more elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), while metabarcoding has incurred additional challenges for the currently available barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
-Particular challenges for the utilization of the high copy number barcodes (e.g.
-ITS2, *rbcL*, *matK*, *trnH-psbA*) include their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
+Taxonomic verification can also be limited by the fact that revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often ineffective.
+The current methods to ameliorate this situation include: (i) ignoring these ecologically relevant levels of detail, (ii) revisiting plots as diagnostic material becomes temporally available, (iii) seeking the assistance from taxonomic specialists, or (iv) the use of barcoding using molecular techniques (CITE).
 
@@
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
-The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants will contribute many more species upon it’s intended completion, slated for 2030.
-Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high amounts of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@cheng2018tenkp, @darwin2022project, @lewin2022biogenome).
+The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants, will contribute many more species by 2030 (@cheng2018tenkp).
+Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high numbers of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@darwin2022project, @lewin2022biogenome).
 These data will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
-However, the application of metabarcoding still face challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
+However, the application of metabarcoding still faces challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
 
@@
 We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
-Species distribution models (SDM’s) examine the ecological conditions associated with known occurrence of a species to identify where in the study area might suitable habitats.
-This approach has the benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers.
-This approach can significantly reduce processing time, particularly as as most next-generation sequence data is deposited as raw-sequence reads.
+Species distribution models examine the ecological conditions associated with the known occurrence of a species to identify suitable habitats in the study area.
+This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers.
+This approach can significantly reduce processing time, particularly as most next-generation sequence data is deposited as raw-sequence reads.
 
@@
 
+@ogilvie2022shifting
+
 # 2 | METHODS
 
-## 2.1 Study System & Field Work
-
-Observations and bee sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information), characterized by high-montane/subalpine Parkland vegetation communities.
-Pollinator observations of *Bombus* Latreille spp.
-(Apidae Latreille) were conducted from May 29^th^ - July 23^rd^ of 2015 in six study sites.
-Observations of *Bombus* foraging took place for one hour at each field site in three 50m transects, each transect was in a major vegetation type (dry, and wet meadows, and Aspen forest), where all flowers abundances were estimated and placed into binds of abundance.
-Corbiculae loads were, non-lethally, collected once from all Queen individuals encountered.
-
-## 2.2 | Floral Visitation
-
-*... does this need more info and warrant a section independent of the above?*
+## 2.1 Case Study: Bee-Flower Observations and Pollen Load Collection
+
+Bee and flower observations and bee corbiculae pollen collection was conducted around the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information).
+The area is characterized by subalpine meadow vegetation communities.
+Pollinator observations of Bombus Latreille spp.
+(Apidae Latreille) were conducted from May 29th – July 23rd of 2015 in six study sites as a part of a larger study (described in Ogilvie and CaraDonna 2022).
+Observations of Bombus foraging took place for one hour at each field site, with equal time spent searching for bee in the major vegetation types (dry, and wet meadows, and aspen forest).
+Corbiculae pollens loads were, non-lethally, collected from queens encountered by capturing them in an insect net and transferring them into a restraining device (“bee squeezer”, @kearns2001natural).
+We then collected a single pollen load (i.e., from one leg) from the bee and then released it.
+At weekly intervals at each site, we also recorded the abundances of flowers visited by bumble bees within belt transects spread over the three vegetation types (0.5 x 40 m transects in each vegetation type, 60 m^2^ total area per site).
 
@@
 Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
+4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
   We used Divisive Hierarchical Clustering techniques to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
@@
 The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with nail polish to prevent oxidation; all samples are noted in *APPENDIX 3*.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in *APPENDIX 3*.
   To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained an unique morphotype than all grains in that transect were also identified and counted.
+In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained a unique morphotype than all grains in that transect were also identified and counted.
 Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
@@
 
-We generate a short list of potential candidate species we downloaded from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022) all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
+We first generated a short list of potential candidate species.
+We downloaded from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022) all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
 To reduce the list of species to include in the genomic sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.
 
-In order to minimize the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
+To minimize the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
 To account for the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
@@
 
-We used all occurrence records from BIEN (n = 23,919) within a 50km border of the Omernik level 3 ecoregion, which includes the study area *(No.
+To determine which clades to include in the reference sequence database we used Species Distribution Modelling.
+We used all occurrence records from BIEN (n = 23,919) within a 50km border of the ecoregion, Omernik level 3, which includes the study area *(No.
 21 “Southern Rockies”)* to construct the species distribution model  (@omernik1987ecoregions).
@@
 Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble).
-No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with differing biologies and ecologies (@elith2006novel, @qiao2015no).
+No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with different biology and ecology (@elith2006novel, @qiao2015no).
 In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).
@@
 To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
-**These publicly available data sets, were selected as they ...** .
+These publicly available data sets, were selected as they pertain to a wide range of variables interacting with plant physiology.
 For linear regression models these predictors underwent both *vifstep* (theta = 10, max observations = 12,500) and *vifcor* (theta = 0.7, max observations = 12,500) to detect highly correlated variables, and collinear features were removed leaving 16 variables (@usdm2014).
@@
 
-To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not be found in the nearest 5), and which were from herbarium records were included.
+For assignment of reads to ecologically probabilistic species subsequent to BLAST, flowering time was used as a filter.
+To estimate the duration of dates in which plant species were flowering Weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
+Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not found in the nearest 5), and which were from herbarium records were included.
 To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall.
+material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in fall.
 Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
-Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
-These estimates were compared to groundtruthed data from a long term experiment 1974-2012, and the field data from 2015, using kendall's tau.
+Species with > 10 records had their Weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
+These estimates were compared to a long-term observational study of flowering phenology 1974-2012 (@caradonna2014shifts), and the floral abundance data from 2015, using Kendall's tau.
 
@@
 
-Using five years (2015-2020) of observational data on *Bombus* Queen Bee foraging at these studies sites, we identified the plant taxa most frequently visited by Queens across all years.
-We sequenced the 12 most commonly visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
+Using five years (2015-2020) of observational data on *Bombus* queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years.
+We sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
 In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
-In addition we sequenced another 15 taxa commonly visited by Bombus workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (*APPENDIX 4*).
+We also sequenced another 15 taxa commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (*APPENDIX 4*).
 Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
@@
 This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH.
-8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
+8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on a heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
 After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°C for 1 hour.
@@
 The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes.
-Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL) , inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
+Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
 Pellets were dried at room temperature overnight before resuspension in nuclease free H~2~O.
@@
 
-In order to precisely classify the contents of each corbiculae load the sequences classified by molecular methods were compared with the fieldwork which at a very fine resolution, recorded the presence and absence of species and their duration of flowering, and was interpreted ala the computer derived temporal and spatial data sets.
+To precisely classify the contents of each corbiculae load the sequences classified by molecular methods were compared with the fieldwork which at a very fine resolution, recorded the presence and absence of species and their duration of flowering, and was interpreted ala the computer derived temporal and spatial data sets.
 The quantitative counts of grains from microscopy, were combined with the semi-quantitative sequencing results, to estimate the abundance of each identified species in each corbiculae load.
@@
 If more than one species in the genus was flowering at that time and site, than the reads were split evenly between the taxa.
-For sequence data which did not match at the genus level, a user subjectively scored them based on the species composition and phenological activity at each plot, the Queen interaction data, and pollen assignments.
+For sequence data which did not match at the genus level, a user subjectively scored them based on the species composition and phenological activity at each site, the queen interaction data, and pollen assignments.
 To estimate the abundance of each of these species in the corbiculae loads, these data were combined with the microscopy data.
-For each morphotype detected in pollen, and each classified sequence read which was not detected via microscopy, they were given a value of 0.5% to indicate there trace presences.
+For each morphotype detected in pollen, and each classified sequence read which was not detected via microscopy, they were given a value of 0.5% to indicate their trace presences.
 When more than a single species belonged to a morphotype group in a single sample, the quantitative values from the morphological work were multiplied by the relative sequence abundance of each species in the load.
@@
 The six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
-A total of `r toString(sum(bee_obs$interactions))` queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, Mdn = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (min = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, Mdn = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
-Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week min = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, Mdn = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (min = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, Mdn = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
-The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, Mdn = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
-
-A total of 66 corbiculae loads were collected from Bees, 64 of them from Queens.
+A total of `r toString(sum(bee_obs$interactions))` overwinter queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
+Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week minimum = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, median = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (minimum = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, median = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
+The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, median = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
+
+A total of 66 corbiculae loads were collected from bees, 64 of them from queens.
 
@@
 
-The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
+The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
 A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Includinfg these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
 
@@
 The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
 The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
@@
 
-At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+At the six study sites, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
 Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
@@
 
-### 3.3 | Metabarcoding Pollen Indentification
-
-### 3.3.1 | Spatial Analyses to identify candidate taxa
-
-The median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance to include species for distribution modelling.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the 554 vascular plants with biotic pollination syndromes, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
-Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
-
-At the six study plots, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
+### 3.3 | Metabarcoding Pollen Identification
 
@@
 The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
+Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
 
@@
 Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
-There was very strong evidence that the weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61),  peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
-There was moderate evidence that the weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
-
-Of the previous 58 species compared, 47 of these could be compared to plot based data from the six sites observed in 2015.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
+There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
+
+Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
 Due to methodological differences, the peak flowering was not compared, and due to the low performance of attempts to model 'duration' in the previous step it was also not compared.
-There was very strong evidence that the weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).
 
@@
 `r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and it's values were used for all subsequent analyses.
+Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
 
@@
 
-To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
+To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala Jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
 Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
@@
 Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
-The floral morphology of *Thalictrum*  spp.
-also makes it unlikely to be visited (BELEIVE JANE SAID THIS- IF NOT STRIKE OUT), and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
+The floral morphology and orientation of flower of *Thalictrum*  spp.
+also makes them unlikely to be visited, and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
 
@@
   rename(Family = 1, Genus = 2)
+
+genera_out <- filter(genera_out, !Family %in% c('Mertensiidae', 'Hesperiidae'))
 
@@
 rm(genera2query, genera_out, automated, blast, expert)
+
+expert <- all_classifications %>% filter(Method == 'Expert')
+blast <- all_classifications %>% 
+  filter(Method == 'BLAST' & Sample %in% expert$Sample)
+automated <- all_classifications %>% 
+  filter(Method == 'Automated' & Sample %in% expert$Sample)
+
+## now determine which proportion of records are correct/incorrect
+
+expert_g <- expert %>% 
+  select(Sample, Genus) %>% 
+  distinct()
+expert_f <- expert %>% 
+  select(Sample, Family) %>% 
+  distinct()
+
+# CORRECT RECORDS FIRST
+
+true_species_blast <- left_join(expert, blast, by = c('Sample', 'Taxon')) %>% 
+  group_by(Sample) %>% 
+  drop_na(Method.y) %>% 
+  summarize(Species_T = n())
+true_genus_blast <- left_join(expert_g, blast, by = c('Sample', 'Genus')) %>% 
+  group_by(Sample) %>% 
+  drop_na(Method) %>% 
+  distinct(Sample, Genus) %>% 
+  summarize(Genera_T = n())
+true_family_blast <- inner_join(expert_f, blast, by = c('Sample', 'Family')) %>% 
+  group_by(Sample) %>% 
+  drop_na(Method) %>% 
+  distinct(Sample, Family) %>% 
+  summarize(Family_T = n())
+
+blast_correct <- left_join(true_species_blast, true_genus_blast, by = 'Sample') %>% 
+  left_join(., true_family_blast, by = 'Sample')
+
+rm(true_species_blast, true_genus_blast, true_family_blast)
+
+true_species_computer <- left_join(expert, 
+                                    filter(automated, Rank == 'Species'),
+                                    by = c('Sample', 'Taxon')) %>% 
+  group_by(Sample) %>% 
+  drop_na(Method.y) %>% 
+  summarize(Species_T = n())
+
+true_genus_computer <- left_join(expert_g,  filter(automated, Rank != 'Family'),
+                                    by = c('Sample', 'Genus')) %>% 
+  group_by(Sample) %>% 
+  drop_na(Method) %>% 
+  distinct(Sample, Genus) %>% 
+  summarize(Genera_T = n())
+
+true_family_computer <- left_join(expert_f, automated, by = c('Sample', 'Family')) %>% 
+  group_by(Sample) %>% 
+  drop_na(Method) %>% 
+  distinct(Sample, Family) %>% 
+  summarize(Family_T = n())
+
+computer_correct <- left_join(true_species_computer, true_genus_computer, by = 'Sample') %>% 
+  left_join(., true_family_computer, by = 'Sample')
+
+rm(true_species_computer, true_genus_computer, true_family_computer)
+
+# FALSE records second
+
+false_species_blast <- anti_join(blast, expert,  by = c('Sample', 'Taxon')) %>% 
+  group_by(Sample) %>% 
+  summarize(Species_F = n())
+
+false_genus_blast <- anti_join(blast, expert_g, by = c('Sample', 'Genus')) %>% 
+  group_by(Sample) %>% 
+  distinct(Sample, Genus) %>% 
+  summarize(Genera_F = n())
+
+false_family_blast <-anti_join(blast, expert_f, by = c('Sample', 'Family')) %>% 
+  group_by(Sample) %>% 
+  distinct(Sample, Family) %>% 
+  summarize(Family_F = n())
+
+blast_incorrect <- left_join(false_species_blast, false_genus_blast, by = 'Sample') %>% 
+  left_join(., false_family_blast, by = 'Sample')
+
+blast_results <- left_join(blast_correct, blast_incorrect, by = 'Sample') %>% 
+  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))
+
+rm(false_species_blast, false_genus_blast, false_family_blast, blast_correct, blast_incorrect)
+
+false_species_computer <- anti_join(automated, expert,  by = c('Sample', 'Taxon')) %>% 
+  group_by(Sample) %>% 
+  summarize(Species_F = n())
+
+false_genus_computer <- anti_join(automated, expert_g, by = c('Sample', 'Genus')) %>% 
+  group_by(Sample) %>% 
+  distinct(Sample, Genus) %>% 
+  summarize(Genera_F = n())
+
+false_family_computer <-anti_join(automated, expert_f, by = c('Sample', 'Family')) %>% 
+  group_by(Sample) %>% 
+  distinct(Sample, Family) %>% 
+  summarize(Family_F = n())
+
+computer_incorrect <- left_join(false_species_computer, false_genus_computer, by = 'Sample') %>% 
+  left_join(., false_family_computer, by = 'Sample')
+
+computer_results <- left_join(computer_correct, computer_incorrect, by = 'Sample') %>% 
+  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))
+
+rm(false_species_computer, false_genus_compuer, false_family_computer, 
+   computer_correct, computer_incorrect)
+
+### calculate accuracy
+
+# number of assignemnts
+
+automated_cnts <- automated %>% 
+  group_by(Sample) %>% 
+  count(Rank) %>% 
+  pivot_wider(names_from = Rank, values_from = n, names_prefix = 'NO_') %>% 
+  mutate(Method = 'COMP')
+blast_cnts <- blast %>% 
+  group_by(Sample) %>% 
+  count(Rank) %>% 
+  pivot_wider(names_from = Rank, values_from = n, names_prefix = 'NO_') %>% 
+  mutate(Method = 'BLAST')
+cnts <- bind_rows(automated_cnts, blast_cnts)
+
+spp_count <- expert %>% 
+  group_by(Sample) %>% 
+  count(name = 'Spp_No')
+gen_count <- expert_g %>% 
+  group_by(Sample) %>% 
+  count(name = 'Gen_No')
+fam_count <- expert_f %>% 
+  group_by(Sample) %>% 
+  count(name = 'Fam_No')
+
+ACC_sens_Spec <- bind_rows(
+  mutate(blast_results, Method = 'BLAST'),
+  mutate(computer_results, Method  = 'COMP')
+) %>% 
+  left_join(., spp_count) %>% 
+  left_join(., gen_count) %>% 
+  left_join(., fam_count) %>% 
+  left_join(., cnts, by = c('Sample', 'Method')) %>% 
+  rowwise() %>% 
+  mutate(
+    Species_false = NO_Species - Species_T,
+
+    Sp_Acc = (Species_T / NO_Species) * 100,
+    Gen_Acc = (Genera_T / sum(NO_Genus, NO_Species, na.rm = T) ) * 100,
+
+    Sp_Sens = (Species_T)  / (Species_T + (Spp_No - Species_T)) * 100
+  )
+
+# Spp_No - Species_match is my TRUE NEGATIVE
+# species_false is MY FALSE POSITIVE
+
+ggplot(ACC_sens_Spec, aes(y = Sp_Acc,  x= Method, color = Method)) +
+  geom_boxplot() +
+  geom_jitter(width = 0.1) +
+  theme_bw()
+
+wilcox.test(
+  pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'COMP', 'Sp_Acc']) , 
+  pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'BLAST', 'Sp_Acc']),
+  alternative = "greater"
+  )
+
+my_means <- ggpubr::compare_means(Sp_Acc ~ Method,  data = ACC_sens_Spec)
+my_comparisons <- my_means %>% 
+    nest(groups = c(group1, group2)) %>% 
+    pull(groups) %>% 
+    map(., as.character)
+sample_sizes <- ACC_sens_Spec %>% 
+    dplyr::group_by(Method) %>%  
+    dplyr::tally() %>% 
+    dplyr::mutate(n = paste0('n = ', n)) 
+min_v <- dplyr::summarise(ACC_sens_Spec, mean_mpg = floor(min(Sp_Acc))) |>
+    pull() |> min()
+
+ggplot(ACC_sens_Spec, aes(y = Sp_Acc,  x= Method, color = Method),  
+                        alpha = 0.5) + 
+  stat_boxplot(notch = T, notchwidth = 0.75, 
+               varwidth = T, 
+               outlier.shape = F, outlier.alpha = 0.8, outlier.colour = 'black') +
+  geom_jitter(width = 0.1, shape = 1) +
+
+  geom_text(data = sample_sizes,
+            aes(Method, Inf, label = n), color = 'black', 
+            vjust = "inward", size = 4, 
+            y = min_v * 0.5) +
+  ggpubr::stat_compare_means(comparisons = my_comparisons, 
+                             aes(label = ..p.signif..),
+                             tip.length = 0, vjust = 0.25, size = 4) +
+
+  expand_limits(y= min_v - 5) +
+  theme_bw() +
+  scale_colour_manual(values = c('#F4AFB4', '#94A89A')) +
+  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
+  scale_x_discrete(labels = c('BLAST', 'Post Classification')) + 
+  labs(y = 'Accuracy', title = 'Accuracy of "naive" BLAST and post classification process') +
+  theme(legend.position = 'none', 
+        plot.title = element_text(hjust = 0.5))
+
 ```
@@
 Weber, this match almost certainly representing *Wyethia arizonica* A.
-Gray (@weber1998new), a taxon known to be visited by Queen bee's via our floral observations.
+Gray (@weber1998new), a taxon known to be visited by queen bee's via our floral observations.
 An expected inaccuracy of the classification scheme is in genus level placements, e.g.
@@
 , and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top interactions.
+
 Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed re-classifications of the contents of pollen grains...
@@
 Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Mertensia* on a per visit basis...
-General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persit with *Bombus* studies...
+General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persist with *Bombus* studies...
 
@@
 To reduce the effects of a low population density of botanists on the maintenance of and production of Flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
-In this proof of concept example we performed several iterations of modelling runs, and several approaches (i.e.
+In this proof-of-concept example we performed several iterations of modelling runs, and several approaches (i.e.
 the 'linear models', and the 'machine learning'), which took notable amounts of compute power.

### 2023-03-02 script/compare_results_BLAST.R

@@ -1,0 +1,268 @@
+library(tidyverse)
+
+setwd('~/Documents/ManuSCript/script')
+
+blast <- read.csv('../data/tally_blast_classifications.csv') %>% 
+  group_by(Sample) %>% 
+  slice_max(Prcnt_seqs, n = 10) %>% 
+  select(Sample, Taxon = taxid, Percent_Sample = Prcnt_seqs) %>% 
+  mutate(
+    Taxon = str_replace(Taxon, 'Arenaria', 'Eremogone'),
+    Genus = str_remove(str_extract(Taxon, '^.* '), ' '),
+    Method = 'BLAST')
+
+automated <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  group_by(Sample, TAXON_NEW) %>%
+  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
+  drop_na() %>% 
+  distinct(Sample, TAXON_NEW, .keep_all = T) %>% 
+  ungroup(TAXON_NEW) %>% 
+  slice_max(Prcnt_seqs, n = 10) %>% 
+  select(Sample, Genus = GENUS, Taxon = TAXON_NEW, Percent_Sample = Prcnt_seqs) %>% 
+  mutate(across(where(is.character), ~ str_replace(., 'Arenaria', 'Eremogone')),
+         Method = 'Automated')
+
+expert <- read.csv('../data/Fully_Integrated_Corbiculae.csv') %>% 
+  rename(Sample = sample.id) %>% 
+  mutate(across(where(is.character), ~ str_trim(.)), 
+         Method = 'Expert') %>% 
+  drop_na() %>% 
+  relocate()
+
+genera2query <- na.omit(unique(c(blast$Genus, automated$Genus, expert$Genus))) 
+genera2query <- genera2query[genera2query != 'Ericaceae'] 
+genera2query <- taxizedb::name2taxid(genera2query, out_type = 'summary') 
+
+genera_out <- taxizedb::classification(genera2query$id, db='ncbi', rank = 'family') 
+genera_out <- map(genera_out, ~ 
+                    data.frame( 
+                      'Family' = filter(.x, rank == "family") %>% select(name), 
+                      'Genus' = filter(.x, rank == 'genus') %>%  select(name) 
+                    )) %>% 
+  bind_rows() %>% 
+  rename(Family = 1, Genus = 2) 
+
+genera_out <- filter(genera_out, !Family %in% c('Mertensiidae', 'Hesperiidae')) 
+
[…]

### 2023-03-03 script/conceptual-tables-boxes.rmd
need to check diff, copied in a funny character

@@
 date: "2023-03-01"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 
-```{r Applications of metabarcoding in plants}
+```{r Applications of metabarcoding in plants, echo = F}
 
@@
     'Forensic Science',
-    'Pharmaceutical/Nutraceutical'
+    'Pharmaceutical'
   ),
@@
 )
+
+write.csv(applications, '../data/Applications_Table.csv', row.names = F)
+read.csv('../data/Applications_Table.csv') |>
+  knitr::kable(caption = 'Applications of Plant Metabarcoding',
+             col.names = c('Application', 'Example', 'Citations'))
+
+rm(applications)
 ```
@@
 
-```{r Major Issues}
+```{r Major Issues, echo = F}
 
-bottlenecks <- tibble(
+bottlenecks <- data.frame(
 
@@
     'Taxonomic Resolution',
-    'Reference Library Coverage - Phylogenetic',
-    'Reference Library Coverage - Spatial',
+    'Reference Library - Phylogenetic',
+    'Reference Library - Spatial',
     'Reference Library Generation', 
@@
     'Kew PAFTOL, no phylogenetic biases', 
-    'Some bias persists towards Europe, North America',
+    'Some bias persists towards Europe',
     'Spatial Modelling; Code within', 
@@
   )
-
 )
 
+write.csv(bottlenecks, '../data/Bottlenecks_Table.csv', row.names = F)
+
+read.csv('../data/Bottlenecks_Table.csv') |>
+  knitr::kable(caption = 'Current Issues Facing Plant Metagenomics',
+             col.names = c('Issue', 'Our Approach', 'Issue Citation'))
+
+rm(bottlenecks)
 ```
@@
 
-```{r Summary of Limitations}
+```{r Summary of Limitations, echo = F}
 
@@
     'Advocate Herbarium Collections', 
-    'Plant and Fungal Tree of Life; ', 
+    'Plant and Fungal Tree of Life; 10kP', 
     'Posterior-Probabilities; Floral Abundance, Nectar/Pollen Nutrition', 
@@
 )
+
+write.csv(limitations, '../data/Limitations_Table.csv', row.names = F)
+
+read.csv('../data/Limitations_Table.csv') |>
+  knitr::kable(caption = 'Limitations Faced and Possible Solutions',
+             col.names = c('Method Component', 'Limitations', 'Paths Forward'))
+
+rm(limitations)
 ```

### 2023-03-03 script/manuscript_MEE.Rmd
need to check diff, copied in a funny character

@@
 author:  |
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$$^, Emily J. Woodworth$^1$$^,$$^2$, 
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive^$$^3$^, Emily J. Woodworth$^1$$^,$$^2$, 
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
 abstract:  |  
-  |  
   1) DNA Barcoding has been successful for the rapid analysis of complex ecological assemblages.
@@
 }
-
 ```
@@
 This is especially true for genera where identification down to species can be useful as specific bioindicators; defining ecological and behavioral properties  (e.g.
-different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).In these instances the lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and their interactions with other species.
+different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).
+In these instances the lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and their interactions with other species.
 This can be further complicated by the fact that the identification of organisms to species is also often mired by lack of diagnostic characters(e.g.
@@
 Taxonomic verification can also be limited by the fact that revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often ineffective.
-The current methods to ameliorate this situation include: (i) ignoring these ecologically relevant levels of detail, (ii) revisiting plots as diagnostic material becomes temporally available, (iii) seeking the assistance from taxonomic specialists, or (iv) the use of barcoding using molecular techniques (CITE).
+The current methods to ameliorate this situation include: (i) ignoring these ecologically relevant levels of detail, (ii) revisiting plots as diagnostic material becomes temporally available, (iii) seeking the assistance from taxonomic specialists, or (iv) the use of barcoding using molecular techniques (@hebert2003biological).
+
+  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
+With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas with many other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+Particular challenges with the high copy number barcodes (e.g.
+ITS2, *rbcL*, *matK*, *trnH-psbA*) include the utilization their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
+Currently, most plant metabarcoding endeavors only allow the identification of material to the level of family or genus.
+
+```{r Applications of Metabarcoding}
+read.csv('../data/Applications_Table.csv') |>
+  knitr::kable(caption = 'Applications of Plant Metabarcoding',
+             col.names = c('Application', 'Example', 'Citations'))
+```
 
@@
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
-The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants, will contribute many more species by 2030 (@cheng2018tenkp).
+The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants, will contribute many more species by 2030 (@cheng2018tenkp).
 Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high numbers of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@darwin2022project, @lewin2022biogenome).
@@
 
-  Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
-
-  To increase the quality of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
+  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
+To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
 To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
@@
 This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers.
-This approach can significantly reduce processing time, particularly as most next-generation sequence data is deposited as raw-sequence reads.
-
-  As species interactions vary both in space and time (@~THOMSON1994 GEO?, @caradonna2021seeing).
-Contrasts in the flowering periods of many plant species, can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
+This approach can significantly reduce processing time, increasing the a projects efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.
+
+```{r Current Bottlenecks in Plant Metabarcoding}
+read.csv('../data/Bottlenecks_Table.csv') |>
+  knitr::kable(caption = 'Current Issues Facing Plant Metagenomics',
+             col.names = c('Issue', 'Our Approach', 'Issue Citation'))
+```
+
+  As species interactions vary both in space and time (@thompson1994coevolutionary, @caradonna2021seeing).
+Contrasts in the flowering periods of many plant species, can provide an additional filter for increasing the accuracy of identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
 In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks, which reduces the utility of these to operate as filters in the post-processing of sequence matches.
@@
 
-  We tested this metagenomic and informatics approach to determine to examine the foraging behaviour of queen bumblebee, and compare this approach with direct observations and the pollen record, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+  We tested this metagenomic and informatics approach to determine to examine the foraging of queen bumblebee, and compare this approach with direct observations and the pollen record, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
 The assessment of the plant species compositions from pollen is a desirable approach, with several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
 The two foraging phases of the queen bumblebee life cycle are essential to 1) increase their weight before diapause, and 2) increase their ovary weights while establishing their recently found nests.
-Both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble, ...).
+Both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
 Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline).
@@
 An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be incorporated into their current diversity (@loarie2009velocity, @dobrowski2016climate).
-
-@ogilvie2022shifting
 
@@
 Observations of Bombus foraging took place for one hour at each field site, with equal time spent searching for bee in the major vegetation types (dry, and wet meadows, and aspen forest).
-Corbiculae pollens loads were, non-lethally, collected from queens encountered by capturing them in an insect net and transferring them into a restraining device (“bee squeezer”, @kearns2001natural).
+Corbiculae pollens loads were, non-lethally, collected from queens encountered by capturing them in an insect net and transferring them into a restraining device ("bee squeezer", @kearns2001natural).
 We then collected a single pollen load (i.e., from one leg) from the bee and then released it.
@@
 Future analyses of the long term data set...
+
+```{r Table of Limitations Faced}
+read.csv('../data/Limitations_Table.csv') |>
+  knitr::kable(caption = 'Limitations Faced and Possible Solutions',
+             col.names = c('Method Component', 'Limitations', 'Paths Forward')) 
+```
 
@@
 ```{r Molecular Networks, echo = F, out.width="110%"}
-knitr::include_graphics('../graphics/plots/Molecular_nets.pdf')
+knitr::include_graphics('../graphics/plots/Molecular_nets-crop.pdf')
 ```

### 2023-03-03 script/manuscript_MEE.tex
need to check diff, copied in a funny character

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{endfloat}
\usepackage{setspace}\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees},
  pdfkeywords={DNA baroding, metagenomics, hyb-seq, Angiosperms353,
plant metagenomics, bombus queens, subalpine, species distribution
model},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Metagenomic Barcoding of Pollen Loads Offers Insights on the
Foraging Patterns of Queen Bumble Bees}
\author{Reed Clark Benkendorf\(^1\)\(^,\)\(^2\)\footnote{Correspondence:
  \href{mailto:reedbenkendorf2021@u.northwestern.edu}{\nolinkurl{reedbenkendorf2021@u.northwestern.edu}}},
Jane E.
Ogilive\textsuperscript{\[^3$^, Emily J. Woodworth$^1\]},\$\(^2\),\\
\hspace*{0.333em}Sophie Taddeo\(^1\)\(^,\)\(^2\), Paul J.
CaraDonna\(^1\)\(^,\)\(^2\)\(^,\)\(^3\), Jeremie B.
Fant\(^1\)\(^,\)\(^2\)\\
\hspace*{0.333em}\(^1\)Chicago Botanic Garden, 1000 Lake Cook Road,
Glencoe, Illinois 60022, USA\\
\hspace*{0.333em}\(^2\)Plant Biology and Conservation, Northwestern
University, Evanston, Illinois 60208, USA\\
\hspace*{0.333em}\(^3\) Rocky Mountain Biological Laboratory, P.O. Box
519, Crested Butte, Colorado 81224, USA}
\date{}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{1 \textbar{} INTRODUCTION}\label{introduction}}

The inability to reliably identify plants to terminal taxon can limit
our understanding of ecosystem function and interactions (Bortolus
(\protect\hyperlink{ref-bortolus2008error}{2008})). This is especially
true for genera where identification down to species can be useful as
specific bioindicators; defining ecological and behavioral properties
(e.g.~different species of Sagebrush- \emph{Artemisia} L., Willows -
\emph{Salix} L., and Sedges - \emph{Carex} L.) (Gage \& Cooper
(\protect\hyperlink{ref-Gage2013HistoricalRO}{2013})). In these
instances the lack of species level data can hinder our understanding of
the breadth of habitat which some species occupy, and their interactions
with other species. This can be further complicated by the fact that the
identification of organisms to species is also often mired by lack of
diagnostic characters(e.g.~flowers, fruits, roots or combinations
thereof), an increasing lack of taxonomic experts (Hebert \emph{et al.}
(\protect\hyperlink{ref-hebert2003biological}{2003})) and the presence
of cryptic species (Janzen \emph{et al.}
(\protect\hyperlink{ref-janzen2017nuclear}{2017}), Oliver \emph{et al.}
(\protect\hyperlink{ref-oliver2009cryptic}{2009})). Taxonomic
verification can also be limited by the fact that revisiting field sites
to identify material using morphological or chemical approaches, can be
resource intensive and often ineffective. The current methods to
ameliorate this situation include: (i) ignoring these ecologically
relevant levels of detail, (ii) revisiting plots as diagnostic material
becomes temporally available, (iii) seeking the assistance from
taxonomic specialists, or (iv) the use of barcoding using molecular
techniques (Hebert \emph{et al.}
(\protect\hyperlink{ref-hebert2003biological}{2003})).

Recently barcoding (the identification of a sample from a single
organism \emph{e.g.} a piece of leaf), and metabarcoding (the
identification of a sample containing a mix of organisms \emph{e.g.}
soil), have shown considerable promise in all Kingdoms of Life (Ruppert
\emph{et al.} (\protect\hyperlink{ref-ruppert2019past}{2019})). With
plants the identification of members of certain clades using barcoding
has been quite successful (Kress
(\protect\hyperlink{ref-kress2017plant}{2017})), whereas with many other
clades results have been elusive (Liu \emph{et al.}
(\protect\hyperlink{ref-liu2014identification}{2014}), Group \emph{et
al.} (\protect\hyperlink{ref-china2011comparative}{2011}), Coissac
\emph{et al.} (\protect\hyperlink{ref-coissac2012bioinformatic}{2012})),
however metabarcoding incurs additional challenges to those which exist
for the currently implemented barcodes (Li \emph{et al.}
(\protect\hyperlink{ref-li2015plant}{2015}), Kress \& Erickson
(\protect\hyperlink{ref-kress2007two}{2007}), Group \emph{et al.}
(\protect\hyperlink{ref-cbol2009dna}{2009}), Coissac \emph{et al.}
(\protect\hyperlink{ref-coissac2012bioinformatic}{2012})). Particular
challenges with the high copy number barcodes (e.g.~ITS2, \emph{rbcL},
\emph{matK}, \emph{trnH-psbA}) include the utilization their rates of
divergence, gene tree conflict, and hybridization (Coissac \emph{et al.}
(\protect\hyperlink{ref-coissac2016barcodes}{2016}), Fazekas \emph{et
al.} (\protect\hyperlink{ref-fazekas2009plant}{2009})). Currently, most
plant metabarcoding endeavors only allow the identification of material
to the level of family or genus.

Currently the largest plant systematic endeavor ever undertaken,by the
Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is
approaching completion (Baker \emph{et al.}
(\protect\hyperlink{ref-baker2021PAFTOL}{2021a})). This data set will
contain hybridization capture (Hyb-Seq) data from at least one species
in each genus of the plant kingdom,14,000 represented species, using the
popular Angiosperms353 (A353) probes, which includes 353 single-copy
orthologous loci, (Baker \emph{et al.}
(\protect\hyperlink{ref-baker2021PAFTOL}{2021a}), Johnson \emph{et al.}
(\protect\hyperlink{ref-johnson2019universal}{2019})). These publicly
available data serve to provide a taxonomically comprehensive backbone
for plant metabarcoding. The A353 probes are currently being used in
many other plant phylogenetic studies increasing the sampling depth of
many clades (Baker \emph{et al.}
(\protect\hyperlink{ref-baker2021exploring}{2021b})). Data from the 10kP
project, which seeks to develop reference genomes from a
phylogenetically diverse suite of plants, will contribute many more
species by 2030 (Cheng \emph{et al.}
(\protect\hyperlink{ref-cheng2018tenkp}{2018})). Similar projects such
as the `Darwin Tree of Life' which will sequence all described taxa in
Britain and Ireland, seek to sequence high numbers of genomes in
geographic regions will contribute data sets applicable to enormous
spatial domains (Life Project Consortium \emph{et al.}
(\protect\hyperlink{ref-darwin2022project}{2022}), Lewin \emph{et al.}
(\protect\hyperlink{ref-lewin2022biogenome}{2022})). These data will
promote the ability to apply metabarcoding to resolve a diverse array of
questions relevant to theoretical and applied ecology (Kress
(\protect\hyperlink{ref-kress2017plant}{2017}), Hollingsworth \emph{et
al.} (\protect\hyperlink{ref-hollingsworth2016telling}{2016})). However,
the application of metabarcoding still faces challenges relating to the
enormity of the genomic data sets and the computational power required
to process sequence data.

Herein we have resolved major components of the problems of accurately
and effectively identifying plant material without diagnostic
morphological character states using the A353 Hyb-Seq probes (Johnson
\emph{et al.} (\protect\hyperlink{ref-johnson2019universal}{2019})), and
custom species sequence databases derived via species distribution
modelling, and temporal filtering. To increase the accuracy and
efficiency of metabarcoding results in plants, we are proposing reducing
the number of possible candidate species by generating a user specific
databases relevant to the region of study and the ecological
characteristics of interest (Bell \emph{et al.}
(\protect\hyperlink{ref-bell2022plants}{2022})). To achieve this goal,
we first create a regional list of candidate species using digital
collections gleaned from herbaria, survey work, and citizen science
(e.g.~iNaturalist), from a region exceeding the study area. For these
candidate species, a modelling approach, such as logistic regression,
may be used to identify taxa which warrant further exploration
e.g.~determine their possibility of presence in metabarcoding samples.
We then use species distribution models to create potential distribution
maps for the candidate species to limit the impact of spatial and
taxonomic biases in the species list and account for spatial variations
in niche availability throughout the study area. Species distribution
models examine the ecological conditions associated with the known
occurrence of a species to identify suitable habitats in the study area.
This approach has the benefit of greatly reducing the size of a sequence
database, which allows for the use of genomic data on personal
computers. This approach can significantly reduce processing time,
increasing the a projects efficiency, particularly as most
next-generation sequence data is deposited as raw-sequence reads.

As species interactions vary both in space and time (Thompson
(\protect\hyperlink{ref-thompson1994coevolutionary}{1994}), CaraDonna
\emph{et al.} (\protect\hyperlink{ref-caradonna2021seeing}{2021})).
Contrasts in the flowering periods of many plant species, can provide an
additional filter for increasing the accuracy of identifying material in
certain types of metagenomic samples (Janzen
(\protect\hyperlink{ref-janzen1967synchronization}{1967}), Newstrom
\emph{et al.} (\protect\hyperlink{ref-Newstrom1994ANC}{1994})). In high
elevation temperate regions, pollination interactions vary temporally
and are characterized by high turnover in active periods of species
(CaraDonna \emph{et al.}
(\protect\hyperlink{ref-caradonna2017interaction}{2017})), however the
overall shorter extent of the active growing season in these systems
results in the presence of few to any natural breaks, which reduces the
utility of these to operate as filters in the post-processing of
sequence matches. Nonetheless, we work develop a general approach which
seems applicable to many areas which utilize the temporal dimension for
classifying sequences in metagenomic samples (but see Davis \emph{et
al.} (\protect\hyperlink{ref-davis2022new}{2022})).

We tested this metagenomic and informatics approach to determine to
examine the foraging of queen bumblebee, and compare this approach with
direct observations and the pollen record, which has shown some
incongruency in several floral visitation networks involving smaller
bodied fauna (Barker \& Arceo-Gomez
(\protect\hyperlink{ref-barker2021pollen}{2021}), Zhao \emph{et al.}
(\protect\hyperlink{ref-zhao2019topology}{2019}), Alarcón
(\protect\hyperlink{ref-alarcon2010congruence}{2010})). The assessment
of the plant species compositions from pollen is a desirable approach,
with several potential applications, despite numerous potential
complications (Pornon \emph{et al.}
(\protect\hyperlink{ref-pornon2017dna}{2017}), Bell \emph{et al.}
(\protect\hyperlink{ref-bell2017applying}{2017}), Sickel \emph{et al.}
(\protect\hyperlink{ref-sickel2015increased}{2015}), Bell \emph{et al.}
(\protect\hyperlink{ref-bell2019quantitative}{2019}), Suchan \emph{et
al.} (\protect\hyperlink{ref-suchan2019pollen}{2019}), Johnson \emph{et
al.} (\protect\hyperlink{ref-johnson2021airborne}{2021})). The two
foraging phases of the queen bumblebee life cycle are essential to 1)
increase their weight before diapause, and 2) increase their ovary
weights while establishing their recently found nests. Both of these
time periods represent potential demographic bottlenecks in bumble bee
populations (Sarro \emph{et al.}
(\protect\hyperlink{ref-sarro2022bumble}{2022})). Bumblebees are one of
the only groups of insects with unequivocal quantitative evidence for
numerous population declines, while simultaneously serving as the most
effective pollinators in many temperate montane ecosystems (Cameron \&
Sadd (\protect\hyperlink{ref-cameron2020global}{2020}), Goulson \emph{et
al.} (\protect\hyperlink{ref-goulson2008decline}{2008}), Williams
(\protect\hyperlink{ref-williams1982distribution}{1982}), Colla \emph{et
al.} (\protect\hyperlink{ref-colla2012assessing}{2012}), Bergman
\emph{et al.}
(\protect\hyperlink{ref-bergman1996micrometeorological}{1996}), Bingham
\& Orthner (\protect\hyperlink{ref-bingham1998efficient}{1998}), Grixti
\emph{et al.} (\protect\hyperlink{ref-grixti2009decline}{2009})). These
areas often represent the most diverse areas in the temperate and often
offer the sole potential refugia for multiple dimensions of biodiversity
under climate change, whilst simultaneously experiencing the greatest
proportional changes in mean annual temperature (Brito-Morales \emph{et
al.} (\protect\hyperlink{ref-brito2018climate}{2018}), Pepin \emph{et
al.} (\protect\hyperlink{ref-pepin2022climate}{2022})). An immediate
understanding of how to manage previously overlooked keystone insect
species, such as bumble bees, is essential if the refugial potential of
the temperate mountains are to be incorporated into their current
diversity (Loarie \emph{et al.}
(\protect\hyperlink{ref-loarie2009velocity}{2009}), Dobrowski \& Parks
(\protect\hyperlink{ref-dobrowski2016climate}{2016})).

\hypertarget{methods}{%
\section{2 \textbar{} METHODS}\label{methods}}

\hypertarget{case-study-bee-flower-observations-and-pollen-load-collection}{%
\subsection{2.1 Case Study: Bee-Flower Observations and Pollen Load
Collection}\label{case-study-bee-flower-observations-and-pollen-load-collection}}

Bee and flower observations and bee corbiculae pollen collection was
conducted around the Rocky Mountain Biological Laboratory (RMBL;
38°57.5'' N, 106°59.3'' W (WGS 84), 2900 m.a.s.l.), Colorado, USA
(APPENDIX 1 for site information). The area is characterized by
subalpine meadow vegetation communities. Pollinator observations of
Bombus Latreille spp. (Apidae Latreille) were conducted from May 29th --
July 23rd of 2015 in six study sites as a part of a larger study
(described in Ogilvie and CaraDonna 2022). Observations of Bombus
foraging took place for one hour at each field site, with equal time
spent searching for bee in the major vegetation types (dry, and wet
meadows, and aspen forest). Corbiculae pollens loads were, non-lethally,
collected from queens encountered by capturing them in an insect net and
transferring them into a restraining device (``bee squeezer'', Kearns
\emph{et al.} (\protect\hyperlink{ref-kearns2001natural}{2001})). We
then collected a single pollen load (i.e., from one leg) from the bee
and then released it. At weekly intervals at each site, we also recorded
the abundances of flowers visited by bumble bees within belt transects
spread over the three vegetation types (0.5 x 40 m transects in each
vegetation type, 60 m\textsuperscript{2} total area per site).

\hypertarget{pollen-morphological-identification}{%
\subsection{2.3 \textbar{} Pollen Morphological
identification}\label{pollen-morphological-identification}}

\hypertarget{pollen-reference-library}{%
\subsubsection{2.3.1 \textbar{} Pollen Reference
Library}\label{pollen-reference-library}}

To develop a reference library of pollen grains which may be present in
corbiculae loads, an image reference collection of fuchsin-jelly stained
(Beattie (\protect\hyperlink{ref-beattie1971technique}{1971})) slides
was assembled from slides previously prepared by the authors (n = 21),
and other researchers (n = 38) (Brosi \& Briggs
(\protect\hyperlink{ref-brosi2013single}{2013})). Using five years of
observational data on \emph{Bombus} Queen Bee foraging at these studies
sites (Ogilvie unpublished), as well as the Vascular Plant Checklist
(Frase \& Buck (\protect\hyperlink{ref-fraser2007vpc}{2007})), an
additional 62 voucher slides for species were prepared and imaged at
400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
4.13.0) from non-accessioned herbarium collections to supplement the
number of species and clades covered (Appendix 3).\\
We used Divisive Hierarchical Clustering techniques to determine which
plant taxa were distinguishable via light microscopy, and to develop a
dichotomous key to pollen morphotypes. Ten readily discernible
categorical traits were collected from each specimen in the image
collection. These traits were transformed using Gower distances, and
clustered using Divisive Hierarchical clustering techniques (Maechler
\emph{et al.} (\protect\hyperlink{ref-cluster2022}{2022})). Using the
cluster dendrogram, elbow plot, and heatmaps (Hennig
(\protect\hyperlink{ref-fpc2022}{2020})), of these results morphological
groups of pollen which could not be resolved via microscopy were
delineated, and a dichotomous key was prepared (APPENDIX NO.). This key
was then used to identify the pollen grains sampled from corbiculae
loads to morphotypes in a consistent manner.

\hypertarget{pollen-corbiculae-loads}{%
\subsubsection{2.3.2 \textbar{} Pollen Corbiculae
Loads}\label{pollen-corbiculae-loads}}

To prepare the pollen slides from corbiculae, all corbiculae loads were
broken apart and rolled using dissection needlepoints to increase
heterogeneity of samples. \emph{Cerca} 0.5mm\textsuperscript{2} of
pollen was placed onto a \textasciitilde4mm\textsuperscript{2} fuchsin
jelly cube (Beattie
(\protect\hyperlink{ref-beattie1971technique}{1971})) atop a
graticulated microscope slide, with 20 transects and 20 rows (400
quadrants) (EMS, Hartfield, PA). The jelly was melted, with stirring,
until pollen grains were homogeneously spread across the microscope
slide. Slides were sealed with Canada Balsam (Rublev Colours, Willits,
CA) followed by sealing with clear nail polish to prevent oxidation; all
samples are noted in \emph{APPENDIX 3}. To identify the pollen present
in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was
used. In initial sampling in three transects, each pollen grain was
identified to morphotype and counted; an additional two transects were
scanned for morphotypes unique to that slide, if either transect
contained a unique morphotype than all grains in that transect were also
identified and counted. Subsequent to the first round of sampling,
non-parametric species richness rarefaction curves (Oksanen \emph{et
al.} (\protect\hyperlink{ref-vegans2022}{2022})), and non-parametric
species diversity rarefaction curves were used to assess the
completeness of sampling (Chao \emph{et al.}
(\protect\hyperlink{ref-inextArticle}{2014}), Hsieh \emph{et al.}
(\protect\hyperlink{ref-inextPackage}{2020})). Slides not approaching
the asymptote of the rarefaction curve were then re-sampled, and
analysed iteratively for up to a total of seven transects \emph{APPENDIX
2}.

\hypertarget{molecular-barcoding}{%
\subsubsection{2.4 \textbar{} Molecular
Barcoding}\label{molecular-barcoding}}

\hypertarget{species-reference-list}{%
\paragraph{2.4.1 \textbar{} Species reference
list}\label{species-reference-list}}

\hypertarget{spatial-analyses}{%
\paragraph{2.4.1.1 Spatial Analyses}\label{spatial-analyses}}

We first generated a short list of potential candidate species. We
downloaded from the Botanical Information and Ecology Network `BIEN'
(Maitner (\protect\hyperlink{ref-bien2022}{2022})) all records adjacent
to the field sites to develop an ecologically relevant list of vascular
plant species, with expected biotic pollination, which may be present at
the study area. To reduce the list of species to include in the genomic
sequence databases, we then generated Species Distribution Models (SDMs)
for these taxa to predict their distribution throughout the study area.

To minimize the number of species for which SDM's were to be generated,
BIEN was queried at a distance of up to 100km from our study area and
all plant species records were downloaded. To account for the
stochasticity of botanical collecting and offset the number of records
associated with the research station, this data set was bootstrap
re-sampled 250 times, with 90\% of samples selected, to create a testing
data set. The median of the logistic regression assessing the
probability of occurrence of a species record as a function of distance
from the study area was used as a threshold distance, under which, to
include species as candidates for distribution modelling.

\hypertarget{distribution-modelling}{%
\paragraph{2.4.1.2 Distribution
Modelling}\label{distribution-modelling}}

To determine which clades to include in the reference sequence database
we used Species Distribution Modelling. We used all occurrence records
from BIEN (n = 23,919) within a 50km border of the ecoregion, Omernik
level 3, which includes the study area \emph{(No.~21 ``Southern
Rockies'')} to construct the species distribution model (Omernik
(\protect\hyperlink{ref-omernik1987ecoregions}{1987})). These records
were copied into two, initially identical, sets, one for generating
machine learning models (ML; Random Forest, and Boosted Regression
Tree's), and the other for Generalised Linear (GLM) and Generalized
Additive Models (GAM) (Barbet-Massin \emph{et al.}
(\protect\hyperlink{ref-barbet2012selecting}{2012})). Ensembled
predictions have been shown to outperform their constituent models, on
average, and to reduce the ecological signal to the analytical noise of
individual runs (Araujo \& New
(\protect\hyperlink{ref-araujo2007ensemble}{2007})). No single method of
producing SDMs has been shown to universally outperform others when
faced with a large and diverse number of applications, in our case a
great number of species with different biology and ecology (Elith*
\emph{et al.} (\protect\hyperlink{ref-elith2006novel}{2006}), Qiao
\emph{et al.} (\protect\hyperlink{ref-qiao2015no}{2015})). In the spirit
of these findings, multiple families of models, which can be generated
together as they have similar requirements regarding the number and
ratios of Presence to Absence records were ensembled together
(Barbet-Massin \emph{et al.}
(\protect\hyperlink{ref-barbet2012selecting}{2012})).

We then generated 4,029 absence points, locations where the focal taxon
is anticipated missing, through a random stratification of 19\% of the
land cover in the area and included them in (Land Management
(\protect\hyperlink{ref-aim2019database}{2019})). To achieve a larger
absence data set, we generated 1,000 pseudo-absence records for each
taxon by randomly selecting coordinates located at least 10km away from
an occurrence record. For ML models, these pseudo-absences were reduced
so that the ratio of presence to absence records were balanced
(Barbet-Massin \emph{et al.}
(\protect\hyperlink{ref-barbet2012selecting}{2012})). To achieve this,
we removed absence records inside of 10\% of the mean sample value of
any predictor variable the presence records; the required number of
absence records were then randomly sampled.

To predict the potential distribution of each species we used 26
environmental variables at 30m resolution, six related to climate, five
soil, four topographic, four related to cloud cover, with the remaining
reflecting assorted abiotic parameters (Wilson \& Jetz
(\protect\hyperlink{ref-wilson2016remotely}{2016}), Wang \emph{et al.}
(\protect\hyperlink{ref-wang2016locally}{2016}), Hengl \emph{et al.}
(\protect\hyperlink{ref-hengl2017soilgrids250m}{2017}), Robinson
\emph{et al.} (\protect\hyperlink{ref-robinson2014earthenv}{2014}))
(\emph{APPENDIX 6}). These publicly available data sets, were selected
as they pertain to a wide range of variables interacting with plant
physiology. For linear regression models these predictors underwent both
\emph{vifstep} (theta = 10, max observations = 12,500) and \emph{vifcor}
(theta = 0.7, max observations = 12,500) to detect highly correlated
variables, and collinear features were removed leaving 16 variables
(Naimi \emph{et al.} (\protect\hyperlink{ref-usdm2014}{2014})).

Modelling: Random Forest and Boosted Regression Trees, were sub sampled
with 30\% test and two replicates each before weighted ensemble based on
True Skill Statistics (tss) (Naimi \& Araujo
(\protect\hyperlink{ref-sdmPackage}{2016})). Generalised linear models
(GLM) and Generalised additive models (GAM) with 30\% sub sampling and
three replicates each were also ensembled using the tss (Naimi \& Araujo
(\protect\hyperlink{ref-sdmPackage}{2016}), @). TSS was chosen as the
ensemble criterion as it has been shown to work across a wide range of
species occurrences prevalence (Allouche \emph{et al.}
(\protect\hyperlink{ref-allouche2006assessing}{2006})). The results of
these models were extracted on a cell-by-cell basis to a polygon feature
derived from a minimum-spanning tree which encompasses the study sites,
and species from either ensemble with greater than 50\% mean habitat
suitability across all cells were considered present for further
purposes (Prim (\protect\hyperlink{ref-prim1957minimum}{1957})).

A total of 535 species were modelled using Generalized Linear Models and
Generalized Additive Models and 534 species were modelled using Random
Forest and Boosted Regression Trees. To evaluate the accuracy of the
species distribution models, additional presence records from GBIF (n =
61,789), and AIM (n = 12,730) were used as test and training sets (n =
74,519) for logistic regression (Occdownload Gbif.Org
(\protect\hyperlink{ref-gbifDL2021sdms}{2021}), Land Management
(\protect\hyperlink{ref-aim2019database}{2019})). Additional novel
absence records were generated from the AIM data set to create a data
set where each species has balanced presence and absences. Eleven or
more paired presence and absence records were required for this testing,
resulting in 334 species being included in the logistic regression (Mdn
= 110.0, \(\bar{x}\) = 223.1, max = 1568 record pairs used) with a 70\%
test split (Kuhn (\protect\hyperlink{ref-caret}{2022})).

\hypertarget{temporal-analyses}{%
\subsubsection{2.4.2 \textbar{} Temporal
Analyses}\label{temporal-analyses}}

For assignment of reads to ecologically probabilistic species subsequent
to BLAST, flowering time was used as a filter. To estimate the duration
of dates in which plant species were flowering Weibull estimates of
several phenological parameters all spatially modelled taxa were
developed (Belitz \emph{et al.}
(\protect\hyperlink{ref-belitz2020accuracy}{2020}), Pearse \emph{et al.}
(\protect\hyperlink{ref-pearse2017statistical}{2017})). Only BIEN
records which occurred in the Omernik Level 4 Ecoregions within 15km of
the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions
if enough records were not found in the nearest 5), and which were from
herbarium records were included. To remove temporally irrelevant
herbarium records, i.e.~material collected during times which flowering
is impossible at the study area due to snow cover, we used the SnowUS
data set (Iler \emph{et al.}
(\protect\hyperlink{ref-iler2021conceptual}{2021}), Tran \emph{et al.}
(\protect\hyperlink{ref-tran2019cloud}{2019})) from 2000-2017 were
analyzed for the first three days of contiguous snow absence, and the
first three days of contiguous snow cover in fall. Herbarium records
after the 3\textsuperscript{rd} quantile for melt, and the
1\textsuperscript{st} quantile for snow cover of these metrics were
removed. Species with \textgreater{} 10 records had their Weibull
distributions generated for the date when 10\% of individuals had begun
flowering, when 50\% were flowering, and when 90\% of individuals had
flowered, we used the initiation and cessation dates, respectively, as
effective start and ends of flowering. These estimates were compared to
a long-term observational study of flowering phenology 1974-2012
(CaraDonna \emph{et al.}
(\protect\hyperlink{ref-caradonna2014shifts}{2014})), and the floral
abundance data from 2015, using Kendall's tau.

\hypertarget{barcode-references-library}{%
\subsubsection{2.5.2 \textbar{} Barcode references
library}\label{barcode-references-library}}

All lab work was carried out at The Daniel F. and Ada L. Rice Plant
Conservation Science Center at the Chicago Botanic Garden, Glencoe,
Illinois, U.S.A.

\hypertarget{sampling-species-for-barcoding}{%
\paragraph{2.5.2.1 \textbar{} Sampling Species for
Barcoding}\label{sampling-species-for-barcoding}}

Using five years (2015-2020) of observational data on \emph{Bombus}
queen interactions with flowering plants at these studies sites, we
identified the plant taxa most frequently visited by queens across all
years. We sequenced the 12 most visited taxa twice using samples
collected from one site within the Gunnison Basin River Drainage and one
individual collected from another more distal population. In addition we
included a congener - or a species from a closely related genus to serve
as an outgroup for all 12 taxa. We also sequenced another 15 taxa
commonly visited by \emph{Bombus} workers, based on the abundances, and
immediate access to plant tissue, in the aforementioned data set
(\emph{APPENDIX 4}). Plant collections were identified typically using a
combination, of dichotomous keys and primary literature as required
(Flora of North America Editorial Committee
(\protect\hyperlink{ref-flora1993flora}{1993+}), Hitchcock \& Cronquist
(\protect\hyperlink{ref-hitchcock2018flora}{2018}), Ackerfield
(\protect\hyperlink{ref-ackerfield2015flora}{2015}), Lesica \emph{et
al.} (\protect\hyperlink{ref-lesica2012manual}{2012}), Cronquist
\emph{et al.}
(\protect\hyperlink{ref-cronquist1977intermountain}{1977+}), Allred \&
Ivey (\protect\hyperlink{ref-allred2012flora}{2012}), \emph{Jepson flora
project} (\protect\hyperlink{ref-jepson2022online}{2020}), Mohlenbrock
(\protect\hyperlink{ref-mohlenbrock2002vascular}{2002})).

\hypertarget{plant-genomic-dna-extraction}{%
\paragraph{2.5.2.2 \textbar{} Plant Genomic DNA
Extraction}\label{plant-genomic-dna-extraction}}

Plant genomic DNA was isolated from \textasciitilde{} 1
cm\textsuperscript{2} of leaf tissue from silica-gel dried or herbarium
material using a modified cetyltrimethylammonium (CTAB) protocol (Doyle
\& Doyle (\protect\hyperlink{ref-doylesCTAB}{1987})) that included two
chloroform washes. DNA was quantified using a Nanodrop 2000 (Thermo
Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer
(Thermo Fisher Scientific).

\hypertarget{pollen-genomic-dna-extraction}{%
\paragraph{2.5.2.3 \textbar{} Pollen Genomic DNA
Extraction}\label{pollen-genomic-dna-extraction}}

Pollen genomic DNA was extracted from corbiculae using a CTAB based
protocol modified from Lahlamgiahi et al.~and Guertler et al.~(2014,
2014). A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50
mM NaCl, 10\% SDS v/v., pH 7.5) was added followed by vortexing to allow
dissolution of corbiculae. Pollen grains were then macerated with Kontes
Pellet Pestles, and the tip of these washed with 130 µL of the SDS
extraction buffer, samples were then incubated for 1 hour at 30°C. This
was followed by the addition of 10\% CTAB solution (450ul, of 20 mM
Tris-Cl pH. 8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10\% CTAB, 5\% PVP,
\textasciitilde85\% Deionized water) and RNAse (10 uL of 10 mg/mL) and
samples were incubated for 40 minutes at 37°C, on a heat block
(Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to
40°C. After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and
DTT (12.5 µL of 1M in water) were added, and the samples were further
incubated at 60°C for 1 hour. Samples were then incubated overnight at
40°C. 500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added,
vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous
phase was pipetted to a 1.5 ml centrifuge tube.

To precipitate the DNA, chilled Isopropyl alcohol \& 3 mM Sodium acetate
(5:1) equivalent to \(\frac{2}{3}\) of the volume of sample were added,
with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging
at 13,000 rpm. The supernatant was pipetted to a new 1.5 ml centrifuge
tube, and 70\% EtOH (400 µL) were added before chilling at -20°C for 20
minutes followed by centrifugation at 13,000 rpm for 10 minutes. Both
tubes were then washed with 75\% EtOH (400 µL), inverted, centrifuged at
13,000 rpm for 4 minutes, and the solution discarded, then washed with
95\% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes,
and the solution discarded. Pellets were dried at room temperature
overnight before resuspension in nuclease free H\textsubscript{2}O.
Extractions were assessed using a Nanodrop 2000 (Thermo Fisher
Scientific) and Qubit fluorometer (Thermo Fisher Scientific). DNA
extracts were then cleaned using 2:1 v./v. Sera-Mag beads (Cytiva,
Little Chalfont, UK) to solute ratio following the manufacturer's
protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half
volume in ambient conditions. DNA was quantified using a Qubit
fluorometer.

\hypertarget{fragmentation-library-preparation-target-enrichment}{%
\paragraph{2.5.2.4 \textbar{} Fragmentation, Library Preparation \&
Target
Enrichment}\label{fragmentation-library-preparation-target-enrichment}}

Library preparation was performed using the NEBNext Ultra II FS-DNA
Library Prep Kit for Illumina (New England BioLabs, Ipswich,
Massachusetts, USA) using slightly modified manufacturers
recommendation. Fragmentation was performed at ½ volume of reagents and
¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned
DNA. Adapter Ligation and PCR enrichment were performed with ½ volumes,
while cleanup of products was performed using SPRI beads (Beckman
Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80\%
v./v. ethanol washes. The exception was the herbarium specimens which
were not fragmented and only end repaired, with similar library
preparation of all samples. Products were analysed on 4\% agarose gels,
and a Qubit fluorometer. Libraries were pooled and enriched with the
Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence
Capture Kit) by following the manufacturer's protocol and Brewer et
al.~2019. Sequencing was performed using an Illumina mi-Seq with 150-bp
end reads, (NUSeq Core, Chicago, Illinois).

\hypertarget{computational-processes-and-analyses.}{%
\subsubsection{2.6 \textbar{} Computational Processes and
Analyses.}\label{computational-processes-and-analyses.}}

\hypertarget{reference-library-data-processing}{%
\paragraph{2.6.1 \textbar{} Reference Library Data
Processing}\label{reference-library-data-processing}}

Sequences were processed using Trimmomatic, which removed sequence
adapters, clipped the first 3 bp, discarding reads less than 36 bp, and
removing reads if their average PHRED score dropped beneath 20 over a
window of 5 bp (Bolger \& Giorgi
(\protect\hyperlink{ref-bolger2014trimmomatic}{2014}), Tange
(\protect\hyperlink{ref-tange_2022_6377950}{2021})). Contigs generated
were mapped to a reference with HybPiper with using target files created
by M353 (Johnson \emph{et al.}
(\protect\hyperlink{ref-johnson2016hybpiper}{2016}), McLay \emph{et al.}
(\protect\hyperlink{ref-mclay2021new}{2021})).

\hypertarget{sequence-identification}{%
\paragraph{2.6.2 \textbar{} Sequence
Identification}\label{sequence-identification}}

A custom Kraken2 database was created by downloading representative
species indicated as being present in the study area by the spatial
analyses from the Sequence Read Archive (SRA) NCBI (Wood \emph{et al.}
(\protect\hyperlink{ref-wood2019improved}{2019})). These sequences were
processed in the same manner as our novel sequences. The Kraken2
database was built using default parameters. Kraken2 was run on
sequences using default parameters (\emph{APPENDIX 5}). Following
Kraken2, Bracken was used to classify sequences to terminal taxa (Lu
\emph{et al.} (\protect\hyperlink{ref-lu2017bracken}{2017})). Results
from both Kraken2 and Bracken, results were reclassified manually to
identify terminal taxa. For example, when only a single species of a
genus was known in the study area, but our database used a
representative of another taxon in the genus, this species was coded as
the result.

\hypertarget{identification-of-sequence-matching-loci}{%
\paragraph{2.6.3 \textbar{} Identification of Sequence Matching
Loci}\label{identification-of-sequence-matching-loci}}

A local NCBI database was built using the same processed novel and
downloaded sequences as the previous database (Camacho \emph{et al.}
(\protect\hyperlink{ref-camacho2009blast}{2009})).

\hypertarget{integrated-observational-molecular-and-palynological-corbiculae}{%
\subsubsection{2.7 \textbar{} Integrated Observational, Molecular, and
Palynological
Corbiculae}\label{integrated-observational-molecular-and-palynological-corbiculae}}

To precisely classify the contents of each corbiculae load the sequences
classified by molecular methods were compared with the fieldwork which
at a very fine resolution, recorded the presence and absence of species
and their duration of flowering, and was interpreted ala the computer
derived temporal and spatial data sets. The quantitative counts of
grains from microscopy, were combined with the semi-quantitative
sequencing results, to estimate the abundance of each identified species
in each corbiculae load.

To reclassify the sequence reads, these data were combined with the
flora observation data, and mapped by genus. If more than one species in
the genus was flowering at that time and site, than the reads were split
evenly between the taxa. For sequence data which did not match at the
genus level, a user subjectively scored them based on the species
composition and phenological activity at each site, the queen
interaction data, and pollen assignments. To estimate the abundance of
each of these species in the corbiculae loads, these data were combined
with the microscopy data. For each morphotype detected in pollen, and
each classified sequence read which was not detected via microscopy,
they were given a value of 0.5\% to indicate their trace presences. When
more than a single species belonged to a morphotype group in a single
sample, the quantitative values from the morphological work were
multiplied by the relative sequence abundance of each species in the
load. All final compositions were standardized to a sum of 100\%, by
adding or subtracting the differences (induced by classifying records as
`trace') to all species with abundances \textgreater{} 1\%.

\hypertarget{results}{%
\section{3 \textbar{} RESULTS}\label{results}}

\hypertarget{floral-observations}{%
\subsection{3.1 \textbar{} Floral
Observations}\label{floral-observations}}

The six sites were surveyed once weekly from May 27-July 27 for a total
of 52 hours. A total of 723 overwinter queen-pollen foraging
interactions were observed (range per bee species by week range = 1-18,
\(\bar{x}\) = 3.46, median = 2), with a range of total observed
interactions per bee species across this time period (minimum = 1,
\(\bar{x}\) = 59.08, median = 19, max = 184). Plants varied widely in
the number of interactions which they partook in with each species of
bee (range per plant species by week minimum = 1 - 20, \(\bar{x}\) =
3.51, median = 2), with a range of total observed interactions per plant
species over this time period (minimum = 1, \(\bar{x}\) = 20.26, median
= 4, max = 141). The number of plant species which bees were observed
interacting with varied more narrowly (range = 1 - 18, \(\bar{x}\) = 8,
median = 6), interactions were observed with a total of 36 plant
species.

A total of 66 corbiculae loads were collected from bees, 64 of them from
queens.

\begin{figure}
\includegraphics[width=0.45\linewidth]{manuscript_MEE_files/figure-latex/Most used resources by species-1} \caption{Number of the ten most commonly visited plants which are also in the top ten most common sequences}\label{fig:Most used resources by species}
\end{figure}

\hypertarget{spatial-analyses-1}{%
\subsection{3.1 \textbar{} Spatial Analyses}\label{spatial-analyses-1}}

\begin{table}

\caption{\label{tab:Results for SA Logistic Regression Metrics}Logistic regression assessing accuracy of SDMs}
\centering
\begin{tabular}[t]{lrlr}
\toprule
Metric & Value & Metric & Value\\
\midrule
Accuracy (Training) & 83.75 & F-Score & 0.84\\
Accuracy (Test) & 84.00 & AUC & 0.92\\
Recall & 81.03 & Concordance & 0.92\\
True Neg. Rate & 86.97 & Discordance & 0.08\\
Precision & 88.04 & Tied & 0.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:Results for SA Logistic Regression Contingency table}Species Distribution Modeling evaluation contingency table}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Training} & \multicolumn{2}{c}{Testing} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
 &  Absence  &   Presence   &  Absence  &  Presence \\
\midrule
Absence & 25620 & 3838 & 11130 & 1653\\
Presence & 6614 & 28248 & 2758 & 12024\\
\bottomrule
\end{tabular}
\end{table}

The threshold distance under which a species would undergo species
distribution modelling was the median (25.009 km) of the logistic
regression assessing the probability of occurrence of a species record
as a function of distance from the study area. A 2-sample test for
equality of proportions with continuity correction (X-squared = 13.254,
df = 1, p-value = 0.000136, 95\% CI 0.04-1.00) was used to test whether
more of the records located in the broad ecological sites present at the
field station, between the distance of the median (25.009 km) to the
third quantile (ca 43.830 km) of the regression distance, were true
presences at the field station. Including these records would have
resulted in modelling an additional 222 species distributions of which
30 are true presences, these taxa were not modelled.

Across the entire spatial domain of modelling all ensembled models (n =
968) had an accuracy of 0.84 (95\% CI 0.8356 - 0.8443), kappa = 0.68,
p-value \textless{} 0.001, sensitivity = 0.80, specificity = 0.87, AUC =
0.92.

In the area of the minimum-spanning tree encompassing the field sites,
of the 554 vascular plants with biotic pollination syndromes, the 493 ML
ensembles accurately predicted the presence of 362 (65.3\%), incorrectly
predicted the presence of 64 (11.6\%), incorrectly predicted 34 true
presences (6.1\%) as being absent, and correctly predicted the true
absence of 33 (6.0\%). The balanced accuracy of the ensembled models is
0.627 (Sensitivity = 0.340, Specificity 0.914). Of the 554 vascular
plants with biotic pollination, the 475 LM ensembles accurately
predicted the presence of 286 (51.6\%), incorrectly predicted the
presence of 41 (14.3\%), incorrectly predicted 93 true presences
(16.8\%) as being absent, and correctly predicted the true absence of 55
(9.9\%). The balanced accuracy of the ensembled models is 0.664
(Sensitivity = 0.573, Specificity 0.754). Of the 554 vascular plants
with biotic pollination syndromes in the flora 13 (2.3\%) were in the
Orchid family and 41 (7.4\%) are non-natives, both of which are
restricted from the database, and can only reduce the number of true
predicted presences by roughly 10\%.

At the six study sites, of the 117 plant species identified to the
species level across the spatial extents of all plots and duration of
queen bee activity, the ML ensembles predicted the presence of 105
(89.7\%) of them, and LM ensembles 102 (87.2\%). Of the missing species
two (1.7\%) are Orchids, six (5.1\%) are non-native, and one (0.85\%) is
of contested taxonomic standing, all of which (7.65\%) are restricted
from the initial query database.

\hypertarget{microscopic-pollen-identification}{%
\subsection{3.2 \textbar{} Microscopic Pollen
identification}\label{microscopic-pollen-identification}}

Using the fuchsin jelly preparation and light microscopic analyses of
grains and scoring of 10 character states resulted in the establishment
of 28 morphotypes which grains could be classified into. \emph{APPENDIX
7}. From the 37 samples that were counted and based on rarefaction we
\textbf{had over \% of expected morphotypes found} (morphotype richness,
\(\bar{x}\) = 4.5, Mdn = 4, min = 1, max = 9), \textbf{all samples had
expected morphotype diversity reach the asymptote \emph{APPENDIX 8}.}
The number of counted pollen grains in each sample range from (514 -
19924, \(\bar{x}\) = 3319, Mdn = 1891).

\textbf{SUMMARY REQUIRED - Number of species from number of families,
and how many species are identified to species versus number only to
genera }

\begin{figure}
\includegraphics[width=0.65\linewidth]{../graphics/plots/rmbl_draft_tree} \caption{Phylogenetic tree of all biotically pollinated plant genera in the study area. The innermost ring indicates every genus which Queen Bee's were observed to visit. The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus. The outermost ring indicates that sequence data were available for at least a member of that genus. Branch colors follow APG 4.}\label{fig:Sequence and Morphological Vouchers}
\end{figure}

\hypertarget{metabarcoding-pollen-identification}{%
\subsubsection{3.3 \textbar{} Metabarcoding Pollen
Identification}\label{metabarcoding-pollen-identification}}

\hypertarget{temporal-analysis}{%
\subsubsection{3.3.2 \textbar{} Temporal
Analysis}\label{temporal-analysis}}

The first date of modeled snow melt in the Gothic area (n = 17,
\(\bar{x}\) = 137.9, Mdn = 135, 3\textsuperscript{rd} quartile = 151),
and the first date of a consistent winter snow base (n = 17, \(\bar{x}\)
= 299.9, Mdn = 300, 1\textsuperscript{st} quartile = 291) from
2000-2017, were used as delimiters for the inclusions of herbarium
records in modelling. Of the 439 species predicted likely present in the
area via logistic regression, 332 species (64.4\%) with more than 10
records in the focal level 4 ecoregions (\(\bar{x}\) = 35.016, Mdn = 35,
max = 96) had Weibull estimates calculated, an additional 56 species
(11.2\%) with enough contributing records from the ``Sedimentary
Mid-Elevation Forests'', a large ecoregion generally just beneath the
elevation bands occupied by the five ecoregions around the study area
had Weibull estimates also calculated (\(\bar{x}\) = 13.868, Mdn = 13,
max = 24).

Only 58 of these 388 species (n = 34.6, Mdn = 31) were able to be
compared to plot based observational data from the long term
(1974--2012) data set (CaraDonna \emph{et al.}
(\protect\hyperlink{ref-caradonna2014shifts}{2014})). Of these species
relatively high accord was observed between the long-term ground truthed
data set, and the modelled species. There was very strong evidence that
the Weibull estimates were positively associated with the observed onset
(p \textless{} 0.0001, tau = 0.61), peak (p \textless{} 0.0001, tau =
0.65), and cessation of flowering (p \textless{} 0.0001, tau = 0.49).
There was moderate evidence that the Weibull estimates had a weak
positive association with the observed duration of flowering (p = 0.58,
tau = 0.17).

Of the previous 58 species compared, 47 of these could be compared to
transect based data from the six sites observed in 2015. Due to
methodological differences, the peak flowering was not compared, and due
to the low performance of attempts to model `duration' in the previous
step it was also not compared. There was very strong evidence that the
Weibull estimates were positively associated with the observed onset (p
\textless{} 0.0001, tau = 0.58), and cessation of flowering (p
\textless{} 0.0001, tau = 0.40).

\begin{figure}
\includegraphics[width=1\linewidth]{../graphics/plots/phenology_estimates_10_90} \caption{Modelled dates of when major flowering events occurred compared between long term and modelled data}\label{fig:plot phenology estimates}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{../graphics/plots/phenology_estimates-jeo_10_90} \caption{Modelled dates of when major flowering events occurred compared between 2015 and modelled data}\label{fig:plot phenology estimates JEO}
\end{figure}

\hypertarget{molecular-analysis-of-corbiculae-loads}{%
\subsection{3.3.1 \textbar{} Molecular analysis of corbiculae
loads}\label{molecular-analysis-of-corbiculae-loads}}

The 54 corbiculae loads had DNA extracted and underwent various steps
towards hyb-seq, in the end a total of 44 corbiculae samples were
sequenced, 7,752,353 reads were recovered from sequencing. The number of
reads per sequence varied widely (range = 76 - 508,795, \(\bar{x}\) =
176,189.8, Mdn = 138,395). Of the possible 353 loci, the number which
were recovered from each sample, and informative to BLAST were range =
24 - 353, \(\bar{x}\) = 305.5, Mdn = 331. The number of reads per loci
from across all samples had a range of 178 - 506,653, \(\bar{x}\) =
20,688, Mdn = 12,616. \textbf{APPENDIX X Reads Per Loci}.

After trimming 7,865,680 sequences remained. 10,682,538 reads were
matched using Kraken, of the reads classified by Kraken 10,160,768 reads
were matched using Bracken, of the reads classified by Kraken 7,549,608
reads were matched using BLAST. Based upon subjective review of the
three classifiers \textbf{APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT
ONES}, BLAST was chosen as the classification method which yielded the
most probable results by the field ecologist, and its values were used
for all subsequent analyses.

\begin{table}

\caption{\label{tab:Read Post Processing}Post classification of Sequences via Taxonomy and Ecology,
top 15 most abundant reads}
\centering
\begin{tabular}[t]{c|c|c|c|c}
\hline
Condition & No. Class. & Prcnt. Class. & Total Seqs & Rank\\
\hline
\cellcolor{gray!10}{A} & \cellcolor{gray!10}{143} & \cellcolor{gray!10}{21.0} & \cellcolor{gray!10}{32.0} & \cellcolor{gray!10}{Species}\\
\hline
B & 205 & 30.1 & 10.5 & Species\\
\hline
\cellcolor{gray!10}{C} & \cellcolor{gray!10}{5} & \cellcolor{gray!10}{0.7} & \cellcolor{gray!10}{0.4} & \cellcolor{gray!10}{Genus}\\
\hline
G & 29 & 4.3 & 7.8 & Species\\
\hline
\cellcolor{gray!10}{H} & \cellcolor{gray!10}{280} & \cellcolor{gray!10}{41.2} & \cellcolor{gray!10}{47.9} & \cellcolor{gray!10}{Genus}\\
\hline
None met & 18 & 2.6 & 1.4 & Multiple\\
\hline
\end{tabular}
\end{table}

The initial classification of sequences which were made by BLAST were
reviewed programmatically, using predicted presence of the species (from
spatial modelling), modelled flowering time (from temporal modelling),
and taxonomy (from existing sources). A sequential process was utilized
which reassigned sequences based on binary combinations of the factors
above (Appendix XX). Given the relative sparsity of the number, and
relatedness, of species represented in the sequence database this was
performed to: 1) Identify locally present species represented by
surrogates in the DB 2) Reduce false classifications of focal species 3)
Identify high confidence sequence matches. Of the top ten taxa which
were identified by BLAST for the 680 distinct records, 55.4\% of the
reads were classified to a species representing 48.3\% of all classified
reads, 41.9\% of the reads were classified to genus representing 48.3\%
of all classified reads, and 0\% of the records were classified to
family.

Of the 0 classifications which were assigned to genera without any
species predicted by spatial analyses, were investigated by hand after
post-processing steps. These were all assigned via post-processing
conditions (: , APPENDIX XX). These were manually assigned to a variety
of ranks, occasionally to genus - 0, and species - 0, by consultation of
the alpha-taxonomic literature (Sadeghian \emph{et al.}
(\protect\hyperlink{ref-sadeghian2015molecular}{2015}), Sennikov \&
Kurtto (\protect\hyperlink{ref-sennikov2017phylogenetic}{2017}), Rabeler
\& Wagner (\protect\hyperlink{ref-rabeler2016new}{2016}), Pusalkar \&
Singh (\protect\hyperlink{ref-pusalkar2015taxonomic}{2015}), Moore \&
Bohs (\protect\hyperlink{ref-moore2003its}{2003}), Weber
(\protect\hyperlink{ref-weber1998new}{1998})).

To determine at which level species in pollen loads could be detected
the results of light microscopy were compared to the molecular results.
The pollen samples contained three morphotypes which could readily be
identified via microscopy. Two of these mapped to the clades
(Boraginaceae \& Heliantheae Alliance), and one to a Asteraceae less
Heliantheae. Boraginaceae grains were detected in 92.3\% of samples
where the proportion of target grains were between 0.01-1 (n = 13 Mdn =
0.663). Asteraceae type 1, non-helianthoids, were detected in 50\% of
samples where the proportion of target grains were between 0.001-0.01 (n
= 4 Mdn = 0.001) Asteraceae type 2, Helianthoids, were detected in
33.3\% of samples where the proportion of target grains were between
0.001-0.01 (n = 6 Mdn = 0.005); however, Asteraceae were detected in
80\% of samples where the proportion of target grains were between
0.001-0.01 (n = 10 Mdn = 0.003). Both morphotypes of Asteraceae pollen
were detected in 100\% of samples where the proportion of target grains
were between 0.01-1 (n = 2 Mdn = 0.338).

\includegraphics{manuscript_MEE_files/figure-latex/Correlation between Number of Grains and Number of Sequences-1.pdf}

To detect whether the sequencing reads were semi-quantitative the subset
of all pollen morphotypes distinguishable by microscopy were compared to
the sequence reads. In all instances sequence reads were pooled to the
highest taxonomic rank associated with the morphotype, e.g.~if both
species of \emph{Mertensia} Huth, or one species and read only
classified to genus were present in a sample, the reads were summed. The
total percentage of the ten most abundant grains per sample were then
were then relativized to constitute the entire sample.

The relationship between the number of pollen grains in a sample and the
number of sequence reads is roughly linear, where grains which are
present in trace amounts are overestimated by sequence counts, while
grains present in high amounts are underestimated. This is likely due to
the proportion of high false positives which occur in the classification
process with next-generation sequencing (Bell \emph{et al.}
(\protect\hyperlink{ref-bell2021comparing}{2021})). There was evidence
of a strong correlation between the proportion of grains per morphotype
and the number of sequences per group (0.426, p \textless{} 0.0001, n =
32).

To ascertain the extent to which records of multiple species in a
family, which were suspected to be sampling artefacts occurred in
molecular samples an index of similarity, ala Jaccard, the affinity
index was used to assess co-occurrence (Mainali \emph{et al.}
(\protect\hyperlink{ref-mainali2022better}{2022}), Mainali \& Slud
(\protect\hyperlink{ref-mainali2022alpha}{2022})). Numerous taxa from
the family Ranunculaceae Jussieu (\emph{Caltha} L. sp.,
\emph{Thalictrum} L. spp., \emph{Trollius} L. sp., \emph{Aquilegia} L.
spp.), had \(\alpha\) scores which indicated that they are only present
when a more common confamilial taxa \emph{Delphinium barbeyi} (Huth)
Huth \emph{nuttallianum} Pritz. were recorded. A similar relationship
was observed in the Hydrophyllaceae R.Br. with samples placed in
\emph{Nemophila} Nutt., which only occurred when the more abundant
\emph{Hydrophyllum} L. species were present. The size of flower of
\emph{Nemophila breviflora} A. Gray make it unlikely to be visited by
Bumble Bees, and it is a false positive. The floral morphology and
orientation of flower of \emph{Thalictrum} spp. also makes them unlikely
to be visited, and while evidence of visits to \emph{Caltha} and
\emph{Trollius} are lacking, due to the association between the reads
these results appear unlikely.

\hypertarget{integrated-observational-molecular-and-palynological-network}{%
\subsection{3.6 \textbar{} Integrated Observational, Molecular, and
Palynological
Network}\label{integrated-observational-molecular-and-palynological-network}}

While the spatial results were used to declare the taxonomic composition
of the sequence database, temporal results were used in consideration
with plant phylogeny to retroactively, reassign the assignment of
sequences to taxa. Essentially, if a sequence was identified to a taxon
which was not known from the field site

For example a many sequences which mapped to the Asteraceae family, but
which was flagged by temporal filters and is present in both \emph{B.
nevadensis} Cresson and \emph{B. rufocinctus} Cresson pollen is most
likely \emph{Frasera} Walter, which failed extractions for the reference
library failed (APPENDIX XX). A similar likely mismatch could be between
what was fide molecular evidence as \emph{Agastache pallidiflora} (A.
Heller) Rydb. but where feeding was infrequently observed on
\emph{Pedicularis} L., likely due to this entire order being represented
by only a single molecular reference species.

\includegraphics{manuscript_MEE_files/figure-latex/compare results-1.pdf}

\begin{verbatim}
## 
##  Wilcoxon rank sum test with continuity correction
## 
## data:  pull(ACC_sens_Spec[ACC_sens_Spec$Method == "COMP", "Sp_Acc"]) and pull(ACC_sens_Spec[ACC_sens_Spec$Method == "BLAST", "Sp_Acc"])
## W = 1254.5, p-value = 5.367e-06
## alternative hypothesis: true location shift is greater than 0
\end{verbatim}

\includegraphics{manuscript_MEE_files/figure-latex/compare results-2.pdf}

Situations where SDM's led to incorrect results at the species level are
evident with classification to \emph{Scabrethia Scabra} (Hooker) W.A.
Weber, this match almost certainly representing \emph{Wyethia arizonica}
A. Gray (Weber (\protect\hyperlink{ref-weber1998new}{1998})), a taxon
known to be visited by queen bee's via our floral observations. An
expected inaccuracy of the classification scheme is in genus level
placements, e.g.~were \emph{Epilobium} L. (Onagraceae Juss.) spp. were
classified. However, given the small size of their flowers in the study
area, these results more likely indicate that a species of
\emph{Chamaenerion} Seg. (a segregate genus) such as \emph{C.
angustifolium} (L.) Scop. or \emph{latifolium} (L.) Sweet is
occasionally utilized, as it supported by limited palynology data. An
issue with reclassification within the family level in combination with
time included reclassifying \emph{Parnassia palustris} L. to
\emph{Paxistima myrsinites} (Pursh) Rafinesque. However, based on flower
size it is more likely that the visited taxon was \emph{P. palustris}.

Regarding limitations of morphological data we suspect that there were
two morphotypes of pollen identified as Ericaceae Juss. were actually
Onagraceae (Samples 19 \& 44), based on molecular results.

It is not unlikely that much of the difference in the results between
the observational and molecular work are attributable to the challenges
in detecting rare events in these smaller sizes. For example, no more
than 10 bee corbiculae loads per species were sequenced with the Mdn =
5.5 \ldots{} , and the median of interactions with the top 5 plant sizes
constituted 0.8283385 of the top interactions.

Accordingly, combining the results of floral observations, and
palynology, molecular sequencing - both pre and post processing, we
subjectively developed re-classifications of the contents of pollen
grains\ldots{}

\hypertarget{discussion}{%
\section{4 \textbar{} DISCUSSION}\label{discussion}}

We have demonstrated how the Angiosperms533 hyb-seq probes may be used
for plant barcoding in a metagenomic context (Johnson \emph{et al.}
(\protect\hyperlink{ref-johnson2019universal}{2019}), Hollingsworth
\emph{et al.} (\protect\hyperlink{ref-hollingsworth2016telling}{2016})).
This was exemplified in an ecologically relevant scenario, where the
results have immediate implications for natural history guided
fundamental science and land management. The test pollen loads contained
a number of closely related taxa, some in notoriously morphologically
difficult clades with rapid rates of diversification
(e.g.~\emph{Mertensia}, \emph{Lupinus} L.), at naturally occurring
proportions (Nevado \emph{et al.}
(\protect\hyperlink{ref-nevado2016widespread}{2016}), Nazaire \& Hufford
(\protect\hyperlink{ref-nazaire2014phylogenetic}{2014})). We
incorporated spatial and temporal approaches for creating custom
sequence databases an approach which is readily applicable to any lab
group with the capacity to perform next-generation sequencing across the
entirety of multiple continents, and which we expect to be highly
beneficial in many study areas. By combining insights from these novel
approaches with an extensive observational field based study we show how
these methods may be applied to test a variety of hypotheses related to
ecological interactions.

The SDM's which we generated, with relatively few occurrence records and
few modelling iterations, performed beyond expectations, likely due to
the utility of the predictor variables and strong alignment of
vegetation by orographic precipitation in the study area. However, we
had difficulties in evaluating our predictions in an operational
context. We utilized the database query approach, to only model species
with a high probability of not being dispersal limited to the focal
area, and focused on a relevant subset of many of these species ranges
to reduce the contributions of range wide adaptions on habitat (Sork
(\protect\hyperlink{ref-sork2018genomic}{2018}), Joshi \emph{et al.}
(\protect\hyperlink{ref-joshi2001local}{2001})). While the models worked
well compared to both test, and validation with external point data,
moving from points to polygon features was more difficult. We were able
to compare our results to 1) a Flora, 2) lists of plants used by Bumble
Bees at plots; the former inappropriate in that it contained a great
number of species which we sought to use modelling to reduce \emph{e.g.}
all strictly alpine species, and the latter inappropriate in that it
contained only species relevant to \emph{Bombus} but had no official
`absence' data. Further given the, size of the minimum spanning tree
which we extracted points to, a formal floristic inventory would still
be a time intensive process. Accordingly, we expect the real results of
our data lay somewhere in between these two evaluations; with an excess
of species predicted present (Dubuis \emph{et al.}
(\protect\hyperlink{ref-dubuis2011predicting}{2011}), Calabrese \emph{et
al.} (\protect\hyperlink{ref-calabrese2014stacking}{2014}),
Pinto-Ledezma \& Cavender-Bares
(\protect\hyperlink{ref-pinto2021predicting}{2021})), but few enough
that they lend themselves to metabarcoding. We observe that our models
seemed very capable of effectively identifying alpine species and
removing them in binomial contexts. Difficulties in temporal models
related to variability in drivers of flowering phenology. AND SMALL
SAMPLE SIZE

These results show that the overall results between \textbf{Bumble Bee
ecology} observational and barcoding are largely congruent. But that
\ldots{} We analyzed pollen loads from all of the most common bumble bee
species in the area (Pyke (\protect\hyperlink{ref-pyke1982local}{1982}))
Future analyses of the long term data set\ldots{}

Results from palynological analyses show that several species of bee
show near perfect fidelity to the genus \emph{Mertensia} on a per visit
basis\ldots{} General results show high congruence between foraging and
molecular results, indicating that concerns regarding mismatch between
observational networks need not persist with \emph{Bombus}
studies\ldots{}

Some foraging preferences of \emph{Bombus}, both at this field site and
across a great many localities globally emerge from this work, which
reiterates the needs for land managers to maintain relatively high
amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae, in
Western North American montane landscapes (Goulson \emph{et al.}
(\protect\hyperlink{ref-goulson2005causes}{2005}), Goulson
(\protect\hyperlink{ref-goulson2010bumblebees}{2010}), Liang \emph{et
al.} (\protect\hyperlink{ref-liang2021evolutionary}{2021}), Bontsutsnaja
\emph{et al.} (\protect\hyperlink{ref-bontvsutvsnaja2021bumble}{2021})).
Numerous historic, and some ongoing, land management practices reduce
the ability of many landscapes to support stable populations of
\emph{Bombus}. Historic livestock grazing was often associated with the
targeted removal of many species of plants which are known to have
compounds toxic to cattle. In particular, the removal of locoweeds
(Fabaceae: \emph{Astragalus} L. \& \emph{Oxytropis} DC.) and larkspurs
(Ranunculaceae: \emph{Delphinium}) were common across public lands
administered by the United States Forest Service (Ralphs \& Ueckert
(\protect\hyperlink{ref-ralphs1988herbicide}{1988}), Aldous
(\protect\hyperlink{ref-aldous1919eradicating}{1919}), Ralphs \emph{et
al.} (\protect\hyperlink{ref-ralphs2003mechanism}{2003})). Further
actions, generally initiated by early settlers, involved the
channelization and incising of streams, culling of beavers, and leaving
cattle concentrated on higher order stream banks for significant periods
of time, all processes which lower the water tables and reduced the
extent of stream-associated {[}riverine{]} wetlands and the mesic
meadows fringes which provide habitat for many species of tall
\emph{Mertensia} (Boraginaceae, e.g.~\emph{M. ciliata} Torr. G. Don.)
widely distributed across Western North America, and to an extent
\emph{Delphinium barbeyi} and many species of native \emph{Trifolium} L.
(Dahl (\protect\hyperlink{ref-dahl1990wetlands}{1990}), Naiman \emph{et
al.} (\protect\hyperlink{ref-naiman1988alteration}{1988}), Belsky
\emph{et al.} (\protect\hyperlink{ref-belsky1999survey}{1999}), Cooke \&
Reeves (\protect\hyperlink{ref-cooke1976arroyos}{1976})). Fire
suppression further resulted in the succession of many Aspen
(\emph{Populus tremuloides} Michx.) groves to Conifer stands, decreasing
the mosaic of age structured habitats in many landscapes, adversely
effects habitat for tall \emph{Mertensia} species and several species of
\emph{Delphinium} (Brewen \emph{et al.}
(\protect\hyperlink{ref-brewen202176}{2021}), Keane
(\protect\hyperlink{ref-keane2002cascading}{2002})). Finally the effects
of Nitrogen deposition, especially given the West's rapidly growing
population still pose adverse effects on the abundance of a variety of
species of Fabaceae at Urban-Rural interfaces (see Stevens \emph{et al.}
(\protect\hyperlink{ref-stevens2018atmospheric}{2018}), Fenn \emph{et
al.} (\protect\hyperlink{ref-fenn2003ecological}{2003})). Current
solutions to these issues, involve targeted burns, reintroduction of
beavers and beaver habitat analogs, and the possibility of re-seeding a
variety of `locoweeds' and `larkspurs' in areas now seldom used, or only
used for early, grazing. The highly enthusiastic response of land
managers, and homeowners, to plant \emph{Ascelpias} L., using
genetically appropriate materials, to improve Monarch Butterfly
(\emph{Danaus plexippus} L.) habitat provides an effective framework for
the latter (Oberhauser \emph{et al.}
(\protect\hyperlink{ref-oberhauser2015monarchs}{2015}), Basey \emph{et
al.} (\protect\hyperlink{ref-basey2015producing}{2015})).

We have concerns regarding the number of persons training to become and
practice botany, and grave concerns regarding the funding mechanisms for
floristic and field based botanical research and for centralized
authorities to produce consensus opinions on alpha taxonomy (Prather
\emph{et al.} (\protect\hyperlink{ref-prather2004decline}{2004b}),
Kramer \& Havens (\protect\hyperlink{ref-kramer2015report}{2015}),
Prather \emph{et al.}
(\protect\hyperlink{ref-prather2004implications}{2004a}), Crisci
\emph{et al.} (\protect\hyperlink{ref-crisci2020end}{2020}), Manzano
(\protect\hyperlink{ref-manzano2021flippant}{2021}), Stroud \emph{et
al.} (\protect\hyperlink{ref-stroud2022botanical}{2022})). To reduce the
effects of a low population density of botanists on the maintenance of
and production of Flora's and to foster meta-genomics across landscapes
without field stations we utilized Species Distribution Modelling to
generate predictive species lists. In this proof-of-concept example we
performed several iterations of modelling runs, and several approaches
(i.e.~the `linear models', and the `machine learning'), which took
notable amounts of compute power. We suspect the possible deleterious
nature of this endeavor may be reduced by: 1) more field surveying by
crews will reduce the need to generate as many species 2) fewer runs of
models, 3) only running machine learning models which do not require an
explicitly process to reduce spatial autocorrelation. However, given the
time required to perform all aspects of a study, even our amount of
computation was negligible. Further, we are very optimistic about the
possibility for persons to perform these tasks, as mentioned we utilized
roughly only one quarter of the records which were digitally available
for presence, and we suspect others will have enough records to perform
this process nearly anywhere else in the temperate. In certain scenarios
modelling of predicted species via more formally tailored S(tacked)-SDM
or J(oint)-SDM approaches may be beneficial (Wilkinson \emph{et al.}
(\protect\hyperlink{ref-wilkinson2021defining}{2021}), Pinto-Ledezma \&
Cavender-Bares (\protect\hyperlink{ref-pinto2021predicting}{2021}),
Schmitt \emph{et al.} (\protect\hyperlink{ref-schmitt2017ssdm}{2017})).

Tandem to the lack of continued expertise required to generate and
maintain species lists, is the expertise required to continue tracking
when major phenological events occur in many plant species at relatively
fine scales or under novel climates. Knowledge of these events is
currently limited to general time periods of only a handful of
phenological events and groups of organisms (e.g.~flowering initiation,
or trees) (Prather \emph{et al.}
(\protect\hyperlink{ref-prather2004implications}{2004a}), Li \emph{et
al.} (\protect\hyperlink{ref-li2016responses}{2016})). While many
programs and initiatives exist to collect phenological information on
subsets of easily identifiable charismatic species to detect major
trends in phenology, these capture only a subset of the extent diversity
(Betancourt \emph{et al.}
(\protect\hyperlink{ref-betancourt2005implementing}{2005}), Havens
\emph{et al.} (\protect\hyperlink{ref-havens2007chicago}{2007})). In
many instances it appears that while landscapes respond similarly to
environmental variables which predict phenological responses, that
individual species vary widely in their responses to similar
environmental cues, or respond to different cues (Augspurger \& Zaya
(\protect\hyperlink{ref-augspurger2020concordance}{2020}), Xie \emph{et
al.} (\protect\hyperlink{ref-xie2015deciduous}{2015}), Xie \emph{et al.}
(\protect\hyperlink{ref-xie2018predicting}{2018}), CaraDonna \emph{et
al.} (\protect\hyperlink{ref-caradonna2014shifts}{2014})). As can be
seen here, predictions of when a single, major phenological event occurs
is already data limited. A more promising approach for the tropics may
lay in utilizing circular statistics (Park \emph{et al.}
(\protect\hyperlink{ref-park2022herbarium}{2022})).

The nearly complete Plant and Fungal Tree of Life (PAFTOL) will provide
a comprehensive phylogenetic backbone of the entire plant kingdom, and
the inclusion of A353 probes with lineage specific probe sets is common
in producing massive genetic datasets (Baker \emph{et al.}
(\protect\hyperlink{ref-baker2021exploring}{2021b})). We predict that
the A353 probes which it is utilizing to work nearly immediately for DNA
barcoding of whole plant material, and that more elaborate validation
studies in controlled metabarcoding settings, utilizing existing
experimental designs, will have favorable results (Bell \emph{et al.}
(\protect\hyperlink{ref-bell2017applying}{2017}), Bell \emph{et al.}
(\protect\hyperlink{ref-bell2019quantitative}{2019}), Bell \emph{et al.}
(\protect\hyperlink{ref-bell2021comparing}{2021}), Lamb \emph{et al.}
(\protect\hyperlink{ref-lamb2019quantitative}{2019})). In particular the
harvesting of loci with more variation in certain lineages, and or with
more variable flanking regions, will prove promising for identifying
closely related plant material (CITE). We suspect that conserved reaches
of genes resulted in the high amounts of reads in somewhat obscure
species. Given that the A353 loci are nuclear, single copy, and a
variety are present the possibility of identifying target loci for
quantitative purposes is high, without continual PCR enrichment is
possible; this would align with relatively high efficacy of WGS (Lang
\emph{et al.} (\protect\hyperlink{ref-lang2019genome}{2019}), Peel
\emph{et al.} (\protect\hyperlink{ref-peel2019semi}{2019}), Bell
\emph{et al.} (\protect\hyperlink{ref-bell2021comparing}{2021})). Recent
evidence indicates that the potential for identifying nearly cryptic
taxa and even infra-specific inference, of either whole plant material,
and perhaps in metagenomic context are possible (Ottenlips \emph{et al.}
(\protect\hyperlink{ref-ottenlips2021resolving}{2021}), Wenzell \emph{et
al.} (\protect\hyperlink{ref-wenzell2021incomplete}{2021}), Loke et
al.~in prep, Slimp \emph{et al.}
(\protect\hyperlink{ref-slimp2021potential}{2021}), Beck \emph{et al.}
(\protect\hyperlink{ref-beck2021palmer}{2021})). We further believe that
in synthetic phylogenetic trees - with incorporation of NGS backbones -
will allow in automatic reassignment of reads as a function of
phylogenetic distance with measures of uncertainty (Hinchliff \emph{et
al.} (\protect\hyperlink{ref-hinchliff2015synthesis}{2015}), Smith \&
Brown (\protect\hyperlink{ref-smith2018constructing}{2018}), Baker
\emph{et al.} (\protect\hyperlink{ref-baker2021PAFTOL}{2021a})).

\hypertarget{conclusion}{%
\section{5 \textbar{} CONCLUSION}\label{conclusion}}

We believe that the combination of spatial and temporal models, united
and guided by localized natural history knowledge, provides the
essential components of a bayesian framework for approaching the coarse
elucidation of ecological interactions using DNA Barcoding. Herein we
crudely utilized this thinking via binary outcomes, should a species
predicted be predicted present or not? Is it unequivocally flowering or
not? Myriad data show biological systems and ecological interactions
have more variance than can be reasonably discretely parsed. We expect
that within a bayesian framework studies of pollinator behavior may be
enacted via this approach at a landscape level, e.g.~the scale of an
entire drainage basin such as the Gunnison which is quickly becoming one
of the worlds few model ecosystems. We hope that the promise of A353
probes as tools for metabarcoding play a role in these endeavors.

\textbf{AUTHOR CONTRIBUTIONS:} R.C.B conducted botanical collections,
conducted all molecular lab work, lead all analyses, and writing. J.E.O
conceived, designed, and conducted all ecological fieldwork, assisted
with analyses, and writing. E.J.W. prepared, imaged, and collected trait
data on pollen reference slides, and assisted with analysis of trait
data and writing a dichotomous key. S.T. assisted with spatial analyses
and writing. P.J.C assisted with ecological analyses and writing. J.B.F.
conceived, and designed all lab work, analyses, and integration of
approaches, assisted with writing, and secured funding for molecular
work.

\textbf{ACKNOWLEDGMENTS:} Nyree Zerega for assistance obtaining herbaria
loans and accessioning our collections at CHIC. Pat Herendeen for
assistance with virtually all aspects of preparing pollen vouchers and
the identification process. Hilary Noble, Zoe Diaz-Martinez, Angela
McDonnell, \& Elena Loke for assistance with genomic library
preparation. Ian Breckheimer for sharing the SDM predictor variables. We
thank the curators at the following herbaria for supplying tissue: Ben
Legler at Stillinger (ID), Charles (Rick) Williams at Ray J. Davis
(IDS), (B)Ernie Nelson at Rocky Mountain (RM); and the collectors: D.
Knoke, L. Brummer, J. Boyd, C. Davidson, I. Gilman, M. Kirkpatrick, S.
McCauley, J. Smith, K. Taylor, \& C. Williams. David Giblin \& Mare
Nazaire for sharing relevant sections of an advanced draft of FNA V. 15.
The Bureau of Land Management is thanked as many plant specimens were
collected by R.C.B as a partner or contractor to the agency; Sarah
Burnett and Lauren Price are thanked for sharing AIM data. Sanda and New
England Biotech are gratefully acknowledged for technical support and
generously sharing samples. T.C.H. Cole for sharing the Angiosperm
Phylogeny 4 colour palette. The Program in Plant Biology and
Conservation is thanked for funding. The holdings of the following
herbaria were essential for this project: AK, ALTA, ASU, BABY, BC, BM,
BMO, BOON, BRIT, CANB, CAS, CHSC, CM, CMN, CNS, COLO, CONN, CS, CSU,
DAV, DBG, DES, ENCB, F, FR, G, GH, GZU, IAC, K, KR, KSP, KSTC, KU, LD,
LOB, LSU, MA, MACF, MEL, MICH, MIL, MIN, MNHN, MO, MO, MT, MW, NCSC,
NSW, NY, NYBG, O, OBI, PI, RBG, RSA, SD, SDSU, SFV, TENN, TRT, UA, UAC,
UAM, UAZ, UBC, UBC, UCR, UCS, UCSB, UMO, UNM, UPS, US, USCH, USF, USU,
UTEP, UWBM, V, VT, W, WSCO, WU, XAL, YPM, Z.

\textbf{CONFLICT OF INTERESTS} The authors declare no conflicts of
interest.

\textbf{PEER REVIEW} The peer review history for this document is
available at \ldots{}

\textbf{DATA AVAILABILITY STATEMENT} The queries required to download
all data used in this project are located in\ldots{} All novel
sequencing data are located at NCBI\ldots{}

\textbf{ORCID}

Paul CaraDonna \url{https://orcid.org/0000-0003-3517-9090}\\
Jeremie Fant \url{https://orcid.org/0000-0001-9276-1111}\\
Jane Ogilvie \url{https://orcid.org/0000-0001-8546-0417}\\
Sophie Taddeo \url{https://orcid.org/0000-0002-7789-1417}

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{supporting}{%
\section{Supporting}\label{supporting}}

\newpage

Appendix 1 - Site Maps

\includegraphics[width=27.01in]{../graphics/plots/siteMaps}

\newpage

Appendix 2 - Species Distribution Models Predictors

\begin{longtable}[]{@{}lccc@{}}
\toprule()
Layer & LM & Description & Source \\
\midrule()
\endhead
1. & N & Mean annual cloudiness - MODIS & Wilson et al.~2016 \\
2. & Y & Cloudiness seasonality 1 - MODIS & Wilson et al.~2016 \\
3. & N & Cloudiness seasonality 2 - MODIS & Wilson et al.~2016 \\
4. & Y & Cloudiness seasonality 3 - MODIS & Wilson et al.~2016 \\
5. & N & Beginning of the frost-free period & Wang et al. \\
6. & N & Climatic moisture deficit & Wang et al. \\
7. & N & Degree-days above 5C & Wang et al. \\
8. & N & Mean annual precipitation & Wang et al. \\
9. & Y & Mean annual precipitation as snow & Wang et al. \\
10. & Y & Temperature seasonality & Wang et al. \\
11. & Y & 2015 Percent Grass/Herbaceous cover - MODIS & (MOD44B) \\
12. & Y & 2015 Percent Tree cover from Landsat 7/8 & (GLCF) \\
13. & Y & Soil probability of bedrock (R Horizon) & SoilGrids \\
14. & N & Soil organic carbon (Tonnes / ha) & SoilGrids \\
15. & N & Surface soil pH in H\textsubscript{2}O & SoilGrids \\
16. & Y & Surface soil percent sand & SoilGrids \\
17. & Y & Soil USDA class & SoilGrids \\
18. & N & Topographic elevation & EarthEnv DEM \\
19. & Y & Topographic elevation, moving window. & EarthEnv DEM \\
20. & Y & Topographic percent slope & EarthEnv DEM \\
21. & Y & Topographic wetness index & EarthEnv DEM \\
22. & Y & Topographic aspect & EarthEnv DEM \\
23. & Y & Annual potential solar radiation computed & r.sun \\
24. & N & Estimated actual (w/-cloud) solar radiation & r.sun / Wilson
et al.~2016 \\
25. & Y & Log-transformed distance to surface water & Global Surface
Water Explorer \\
26. & Y & Percent surface water & Global Surface Water Explorer \\
\bottomrule()
\end{longtable}

\newpage

Appendix 3 - Molecular Reference Specimen Table

\includegraphics{../graphics/tables/specimen_tables-1.pdf}

\newpage

\includegraphics{../graphics/tables/specimen_tables-2.pdf}

\newpage

Appendix 4 - All Pollen Reference Slides Used to Establish Morphotypes

\includegraphics{../graphics/assorted/pollen_slide_table_reduced-1.pdf}

\newpage

\includegraphics{../graphics/assorted/pollen_slide_table_reduced-2.pdf}

\newpage

\includegraphics{../graphics/assorted/pollen_slide_table_reduced-3.pdf}

\newpage

\hypertarget{pollen-cluster-results-should-be-here}{%
\subsubsection{POLLEN CLUSTER RESULTS SHOULD BE
HERE}\label{pollen-cluster-results-should-be-here}}

\newpage

Appendix XX - Pollen Key

\includegraphics{../graphics/assorted/RMBL_pollen_key-1.pdf}

\newpage

\includegraphics{../graphics/assorted/RMBL_pollen_key-2.pdf}

\newpage

APPENDIX XX - Pollen Morphotype Richness Rarefaction Curves

\includegraphics[width=0.95\linewidth]{../graphics/plots/species_richness_rarefaction}

\newpage

Appendix XX - Pollen Morphotype Abundance Rarefaction Curves

\includegraphics[width=0.99\linewidth]{../graphics/plots/SppAbundance}

\newpage

Appendix XX - All Species in the Sequence Databases

\includegraphics{../graphics/assorted/kraken_db_spp_table-1.pdf}

\newpage

\includegraphics{../graphics/assorted/kraken_db_spp_table-2.pdf}

\newpage

Appendix XX - All Species in the Sequence Databases (con't)

\includegraphics{../graphics/assorted/kraken_db_spp_table-3.pdf}

\newpage

Appendix XX - All Species in the Sequence Databases (con't)

\includegraphics{../graphics/assorted/kraken_db_spp_table-4.pdf}

\newpage

Appendix XX - Reads Per Loci

\includegraphics[width=0.98\linewidth]{../graphics/plots/Loci_Reference_Samples}

\includegraphics[width=0.98\linewidth]{../graphics/plots/Percent_loci_matched_by_sample}

\newpage

Appendix XX - Comparison of Kraken2, Bracken, and BLAST

\includegraphics[width=1.1\linewidth]{../graphics/plots/Molecular_nets-crop}

\newpage

Appendix XX - Models used for Species Distribution Model Ensembles

The two machine learning models utilize Ensemble learning.

\textbf{Ensemble learning} utilizes many sets of trees, each tree being
composed of many binary decisions, to create a single model. Each
independent variable ( - or \emph{feature}) may become a node on the
tree - i.e.~a location on the tree where a binary decision will move
towards a predicted outcome. Each of the decision tree models which
ensemble learning utilizes is a weak model, each of which may suffer due
to high variance or bias, but which produce better outcomes than would
be expected via chance. When ensembled these models generate a strong
model, a model which should have more appropriately balanced variance
and bias and predicts outcomes which are more strongly correlated with
the expected values than the individual weak models.

\emph{\textbf{Random Forest (RF)}} the training data are continually
bootstrap re-sampled, in combination with random subsets of features, to
create nodes which attempt to optimally predict a known outcome. A large
number of trees are then aggregated, via the most common predictions, to
generate a final classification prediction tree. Each individual
prediction tree is generated independently of the others.

\emph{\textbf{Boosted Regression Tree (BRT)}} (or Gradient Boosted tree)
An initial tree is grown, and all other trees are derived sequentially
from it, as each new tree is grown the errors in responses from the last
tree are weighed more heavily so that the model focuses on selecting
dependent variables which refine predictions. All response data and
predictor variables are kept available to all trees.

\emph{\textbf{Bias}} predictions from an algorithm are systematically in
error due to being prejudiced for or against certain results, due to
assumptions during learning.

\emph{\textbf{Variance}} errors in models due to an over-reliance and
sensitivity of training to outliers in training data.

In general, Random Forest models have high bias and low variance, where
boosted regressions trees have lower bias and higher variance.
Theoretically, the weaknesses and strengths of bootstrap aggregation
(bagging) as implemented by Random Forests are supplemented by the
boosting.

\newpage

APPENDIX XX - Time Spent Generating Species Distribution Models

\includegraphics{../graphics/assorted/individual_model_runs.pdf}

\newpage

APPENDIX XX - Review Classified Reads and Reassign

\includegraphics{../graphics/tables/reclassify_reads_table.pdf}

\newpage

APPENDIX XX - Tips and tricks for implementing plant metagenomic
sequencing

\includegraphics{../graphics/assorted/overall_metagenomic_workflow.pdf}

\newpage

\includegraphics{../graphics/assorted/microscopy_workflow.pdf}

\newpage

\includegraphics{../graphics/assorted/molecular_workflow.pdf}

\newpage

\includegraphics{../graphics/assorted/pollen_ctab.pdf}

\newpage

\includegraphics{../graphics/assorted/pollen_ctab-1.pdf}

\newpage

THIS SHOULD BE TURNED INTO A SMALLER PNG, AND HAVE THE NUMBER OF
SEQUENCED METAGENOMIC SAMPLES PLACED INTO THAT COLUMN AND INCLUDED IN
TEXT \textasciitilde\textasciitilde\textasciitilde{} NEED THIS !!!!

\includegraphics{../graphics/tables/bombus_samples.pdf}

\newpage

\includegraphics[width=5.83in,height=1.25\textheight]{../graphics/plots/flowchart}

\newpage

\hypertarget{references-1}{%
\section*{References}\label{references-1}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-ackerfield2015flora}{}}%
Ackerfield, J. (2015). \emph{Flora of colorado}. BRIT Press Fort Worth.

\leavevmode\vadjust pre{\hypertarget{ref-alarcon2010congruence}{}}%
Alarcón, R. (2010). Congruence between visitation and pollen-transport
networks in a california plant--pollinator community. \emph{Oikos},
\textbf{119}, 35--44. Retrieved from
\url{https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1600-0706.2009.17694.x}

\leavevmode\vadjust pre{\hypertarget{ref-aldous1919eradicating}{}}%
Aldous, A.E. (1919). \emph{Eradicating tall larkspur on cattle ranges in
the national forest}. US Department of Agriculture.

\leavevmode\vadjust pre{\hypertarget{ref-allouche2006assessing}{}}%
Allouche, O., Tsoar, A. \& Kadmon, R. (2006). Assessing the accuracy of
species distribution models: Prevalence, kappa and the true skill
statistic (TSS). \emph{Journal of applied ecology}, \textbf{43},
1223--1232.

\leavevmode\vadjust pre{\hypertarget{ref-allred2012flora}{}}%
Allred, K.W. \& Ivey, R. (2012). Flora neomexicana III: An illustrated
identification manual. \emph{Lulu. com}.

\leavevmode\vadjust pre{\hypertarget{ref-araujo2007ensemble}{}}%
Araujo, M.B. \& New, M. (2007). Ensemble forecasting of species
distributions. \emph{Trends in ecology \& evolution}, \textbf{22},
42--47.

\leavevmode\vadjust pre{\hypertarget{ref-augspurger2020concordance}{}}%
Augspurger, C.K. \& Zaya, D.N. (2020). Concordance of long-term shifts
with climate warming varies among phenological events and herbaceous
species. \emph{Ecological Monographs}, \textbf{90}, e01421.

\leavevmode\vadjust pre{\hypertarget{ref-baker2021PAFTOL}{}}%
Baker, W.J., Bailey, P., Barber, V., Barker, A., Bellot, S., Bishop, D.,
Botigué, L.R., Brewer, G., Carruthers, T., Clarkson, J.J., Cook, J.,
Cowan, R.S., Dodsworth, S., Epitawalage, N., Françoso, E., Gallego, B.,
Johnson, M.G., Kim, J.T., Leempoel, K., Maurin, O., Mcginnie, C.,
Pokorny, L., Roy, S., Stone, M., Toledo, E., Wickett, N.J., Zuntini,
A.R., Eiserhardt, W.L., Kersey, P.J., Leitch, I.J. \& Forest, F.
(2021a). {A Comprehensive Phylogenomic Platform for Exploring the
Angiosperm Tree of Life}. \emph{Systematic Biology}, \textbf{71},
301--319. Retrieved from \url{https://doi.org/10.1093/sysbio/syab035}

\leavevmode\vadjust pre{\hypertarget{ref-baker2021exploring}{}}%
Baker, W., Dodsworth, S., Forest, F., Graham, S., Johnson, M.,
McDonnell, A., Pokorny, L., Tate, J., Wicke, S. \& Wickett, N. (2021b).
\href{https://doi.org/10.1002/ajb2.1703}{Exploring Angiosperms353: An
open, community toolkit for collaborative phylogenomic research on
flowering plants}. \emph{American Journal of Botany}, \textbf{108}.

\leavevmode\vadjust pre{\hypertarget{ref-barbet2012selecting}{}}%
Barbet-Massin, M., Jiguet, F., Albert, C.H. \& Thuiller, W. (2012).
Selecting pseudo-absences for species distribution models: How, where
and how many? \emph{Methods in ecology and evolution}, \textbf{3},
327--338.

\leavevmode\vadjust pre{\hypertarget{ref-barker2021pollen}{}}%
Barker, D.A. \& Arceo-Gomez, G. (2021). {Pollen transport networks
reveal highly diverse and temporally stable plant--pollinator
interactions in an Appalachian floral community}. \emph{AoB PLANTS},
\textbf{13}. Retrieved from \url{https://doi.org/10.1093/aobpla/plab062}

\leavevmode\vadjust pre{\hypertarget{ref-basey2015producing}{}}%
Basey, A.C., Fant, J.B. \& Kramer, A.T. (2015). Producing native plant
materials for restoration: 10 rules to collect and maintain genetic
diversity. \emph{Native Plants Journal}, \textbf{16}, 37--53.

\leavevmode\vadjust pre{\hypertarget{ref-beattie1971technique}{}}%
Beattie, A. (1971). A technique for the study of insect-borne pollen.
\emph{The Pan-Pacific Entomologist}, \textbf{47}, 82.

\leavevmode\vadjust pre{\hypertarget{ref-beck2021palmer}{}}%
Beck, J.B., Markley, M.L., Zielke, M.G., Thomas, J.R., Hale, H.J.,
Williams, L.D. \& Johnson, M.G. (2021). Are palmer's elm-leaf goldenrod
and the smooth elm-leaf goldenrod real? The Angiosperms353 kit provides
within-species signal in solidago ulmifolia sl. \emph{Systematic
Botany}, \textbf{46}, 1107--1113.

\leavevmode\vadjust pre{\hypertarget{ref-belitz2020accuracy}{}}%
Belitz, M.W., Larsen, E.A., Ries, L. \& Guralnick, R.P. (2020). The
accuracy of phenology estimators for use with sparsely sampled
presence-only observations. \emph{Methods in Ecology and Evolution},
\textbf{11}, 1273--1285.

\leavevmode\vadjust pre{\hypertarget{ref-bell2019quantitative}{}}%
Bell, K.L., Burgess, K.S., Botsch, J.C., Dobbs, E.K., Read, T.D. \&
Brosi, B.J. (2019). Quantitative and qualitative assessment of pollen
DNA metabarcoding using constructed species mixtures. \emph{Molecular
Ecology}, \textbf{28}, 431--455.

\leavevmode\vadjust pre{\hypertarget{ref-bell2017applying}{}}%
Bell, K.L., Fowler, J., Burgess, K.S., Dobbs, E.K., Gruenewald, D.,
Lawley, B., Morozumi, C. \& Brosi, B.J. (2017). Applying pollen DNA
metabarcoding to the study of plant--pollinator interactions.
\emph{Applications in plant sciences}, \textbf{5}, 1600124.

\leavevmode\vadjust pre{\hypertarget{ref-bell2021comparing}{}}%
Bell, K.L., Petit III, R.A., Cutler, A., Dobbs, E.K., Macpherson, J.M.,
Read, T.D., Burgess, K.S. \& Brosi, B.J. (2021). Comparing whole-genome
shotgun sequencing and DNA metabarcoding approaches for species
identification and quantification of pollen species mixtures.
\emph{Ecology and Evolution}, \textbf{11}, 16082--16098.

\leavevmode\vadjust pre{\hypertarget{ref-bell2022plants}{}}%
Bell, K.L., Turo, K.J., Lowe, A., Nota, K., Keller, A., Encinas-Viso,
F., Parducci, L., Richardson, R.T., Leggett, R.M., Brosi, B.J. \&
others. (2022). Plants, pollinators and their interactions under global
ecological change: The role of pollen DNA metabarcoding. \emph{Molecular
ecology}.

\leavevmode\vadjust pre{\hypertarget{ref-belsky1999survey}{}}%
Belsky, A.J., Matzke, A. \& Uselman, S. (1999). Survey of livestock
influences on stream and riparian ecosystems in the western united
states. \emph{Journal of Soil and water Conservation}, \textbf{54},
419--431.

\leavevmode\vadjust pre{\hypertarget{ref-bergman1996micrometeorological}{}}%
Bergman, P., Molau, U. \& Holmgren, B. (1996). Micrometeorological
impacts on insect activity and plant reproductive success in an alpine
environment, swedish lapland. \emph{Arctic and alpine research},
\textbf{28}, 196--202.

\leavevmode\vadjust pre{\hypertarget{ref-betancourt2005implementing}{}}%
Betancourt, J.L., Schwartz, M.D., Breshears, D.D., Cayan, D.R.,
Dettinger, M.D., Inouye, D.W., Post, E. \& Reed, B.C. (2005).
Implementing a US national phenology network.

\leavevmode\vadjust pre{\hypertarget{ref-bingham1998efficient}{}}%
Bingham, R.A. \& Orthner, A.R. (1998). Efficient pollination of alpine
plants. \emph{Nature}, \textbf{391}, 238--239.

\leavevmode\vadjust pre{\hypertarget{ref-bolger2014trimmomatic}{}}%
Bolger, A. \& Giorgi, F. (2014). Trimmomatic: A flexible read trimming
tool for illumina NGS data. \emph{Bioinformatics}, \textbf{30},
2114--2120.

\leavevmode\vadjust pre{\hypertarget{ref-bontvsutvsnaja2021bumble}{}}%
Bontsutsnaja, A., Karise, R., Mand, M. \& Smagghe, G. (2021). Bumble bee
foraged pollen analyses in spring time in southern estonia shows
abundant food sources. \emph{Insects}, \textbf{12}, 922.

\leavevmode\vadjust pre{\hypertarget{ref-bortolus2008error}{}}%
Bortolus, A. (2008). Error cascades in the biological sciences: The
unwanted consequences of using bad taxonomy in ecology. \emph{AMBIO: A
journal of the human environment}, \textbf{37}, 114--118.

\leavevmode\vadjust pre{\hypertarget{ref-brewen202176}{}}%
Brewen, C.J., Berrill, J.-P., Ritchie, M.W., Boston, K., Dagley, C.M.,
Jones, B., Coppoletta, M. \& Burnett, C.L. (2021). 76-year decline and
recovery of aspen mediated by contrasting fire regimes: Long-unburned,
infrequent and frequent mixed-severity wildfire. \emph{Plos one},
\textbf{16}, e0232995.

\leavevmode\vadjust pre{\hypertarget{ref-brito2018climate}{}}%
Brito-Morales, I., Molinos, J.G., Schoeman, D.S., Burrows, M.T.,
Poloczanska, E.S., Brown, C.J., Ferrier, S., Harwood, T.D., Klein, C.J.,
McDonald-Madden, E. \& others. (2018). Climate velocity can inform
conservation in a warming world. \emph{Trends in ecology \& evolution},
\textbf{33}, 441--457.

\leavevmode\vadjust pre{\hypertarget{ref-brosi2013single}{}}%
Brosi, B.J. \& Briggs, H.M. (2013). Single pollinator species losses
reduce floral fidelity and plant reproductive function.
\emph{Proceedings of the National Academy of Sciences}, \textbf{110},
13044--13048.

\leavevmode\vadjust pre{\hypertarget{ref-calabrese2014stacking}{}}%
Calabrese, J.M., Certain, G., Kraan, C. \& Dormann, C.F. (2014).
Stacking species distribution models and adjusting bias by linking them
to macroecological models. \emph{Global Ecology and Biogeography},
\textbf{23}, 99--112.

\leavevmode\vadjust pre{\hypertarget{ref-camacho2009blast}{}}%
Camacho, C., Coulouris, G., Avagyan, V., Ma, N., Papadopoulos, J.,
Bealer, K. \& Madden, T.L. (2009). BLAST+: Architecture and
applications. \emph{BMC bioinformatics}, \textbf{10}, 1--9.

\leavevmode\vadjust pre{\hypertarget{ref-cameron2020global}{}}%
Cameron, S.A. \& Sadd, B.M. (2020). Global trends in bumble bee health.
\emph{Annual review of entomology}, \textbf{65}, 209--232.

\leavevmode\vadjust pre{\hypertarget{ref-caradonna2021seeing}{}}%
CaraDonna, P.J., Burkle, L.A., Schwarz, B., Resasco, J., Knight, T.M.,
Benadi, G., Bluthgen, N., Dormann, C.F., Fang, Q., Frund, J. \& others.
(2021). Seeing through the static: The temporal dimension of
plant--animal mutualistic interactions. \emph{Ecology Letters},
\textbf{24}, 149--161.

\leavevmode\vadjust pre{\hypertarget{ref-caradonna2014shifts}{}}%
CaraDonna, P.J., Iler, A.M. \& Inouye, D.W. (2014). Shifts in flowering
phenology reshape a subalpine plant community. \emph{Proceedings of the
National Academy of Sciences}, \textbf{111}, 4916--4921.

\leavevmode\vadjust pre{\hypertarget{ref-caradonna2017interaction}{}}%
CaraDonna, P.J., Petry, W.K., Brennan, R.M., Cunningham, J.L.,
Bronstein, J.L., Waser, N.M. \& Sanders, N.J. (2017). Interaction
rewiring and the rapid turnover of plant--pollinator networks.
\emph{Ecology letters}, \textbf{20}, 385--394.

\leavevmode\vadjust pre{\hypertarget{ref-inextArticle}{}}%
Chao, A., Gotelli, N.J., Hsieh, T.C., Sande, E.L., Ma, K.H., Colwell,
R.K. \& Ellison, A.M. (2014). Rarefaction and extrapolation with hill
numbers: A framework for sampling and estimation in species diversity
studies. \emph{Ecological Monographs}, \textbf{84}, 45--67.

\leavevmode\vadjust pre{\hypertarget{ref-cheng2018tenkp}{}}%
Cheng, S., Melkonian, M., Smith, S.A., Brockington, S., Archibald, J.M.,
Delaux, P.-M., Li, F.-W., Melkonian, B., Mavrodiev, E.V., Sun, W., Fu,
Y., Yang, H., Soltis, D.E., Graham, S.W., Soltis, P.S., Liu, X., Xu, X.
\& Wong, G.K.-S. (2018). {10KP: A phylodiverse genome sequencing plan}.
\emph{GigaScience}, \textbf{7}. Retrieved from
\url{https://doi.org/10.1093/gigascience/giy013}

\leavevmode\vadjust pre{\hypertarget{ref-coissac2016barcodes}{}}%
Coissac, E., Hollingsworth, P.M., Lavergne, S. \& Taberlet, P. (2016).
From barcodes to genomes: Extending the concept of DNA barcoding.

\leavevmode\vadjust pre{\hypertarget{ref-coissac2012bioinformatic}{}}%
Coissac, E., Riaz, T. \& Puillandre, N. (2012). Bioinformatic challenges
for DNA metabarcoding of plants and animals. \emph{Molecular ecology},
\textbf{21}, 1834--1847.

\leavevmode\vadjust pre{\hypertarget{ref-colla2012assessing}{}}%
Colla, S.R., Gadallah, F., Richardson, L., Wagner, D. \& Gall, L.
(2012). Assessing declines of north american bumble bees (bombus spp.)
Using museum specimens. \emph{Biodiversity and Conservation},
\textbf{21}, 3585--3595.

\leavevmode\vadjust pre{\hypertarget{ref-cooke1976arroyos}{}}%
Cooke, R.U. \& Reeves, R.W. (1976). \emph{Arroyos and environmental
change in the american south-west}. Clarendon Press.

\leavevmode\vadjust pre{\hypertarget{ref-crisci2020end}{}}%
Crisci, J.V., Katinas, L., Apodaca, M.J. \& Hoch, P.C. (2020). The end
of botany. \emph{Trends in Plant Science}, \textbf{25}, 1173--1176.

\leavevmode\vadjust pre{\hypertarget{ref-cronquist1977intermountain}{}}%
Cronquist, A., Holmgren, A.H., Holmgren, N.H., Reveal, J.L., Holmgren,
P.K., Barneby, R \& others. (1977+). \emph{Intermountain flora. Vascular
plants of the intermountain west, USA volume six. The monocotyledons.}
Columbia University.

\leavevmode\vadjust pre{\hypertarget{ref-dahl1990wetlands}{}}%
Dahl, T.E. (1990). \emph{Wetlands losses in the united states, 1780's to
1980's}. US Department of the Interior, Fish; Wildlife Service.

\leavevmode\vadjust pre{\hypertarget{ref-davis2022new}{}}%
Davis, C.C., Lyra, G.M., Park, D.S., Asprino, R., Maruyama, R.,
Torquato, D., Cook, B.I. \& Ellison, A.M. (2022). New directions in
tropical phenology. \emph{Trends in Ecology \& Evolution}.

\leavevmode\vadjust pre{\hypertarget{ref-dobrowski2016climate}{}}%
Dobrowski, S.Z. \& Parks, S.A. (2016). Climate change velocity
underestimates climate change exposure in mountainous regions.
\emph{Nature Communications}, \textbf{7}, 1--8.

\leavevmode\vadjust pre{\hypertarget{ref-doylesCTAB}{}}%
Doyle, J.J. \& Doyle, J.L. (1987). A rapid DNA isolation procedure for
small quantities of fresh leaf tissue. \emph{Phytochemical Bulletin},
\textbf{19}, 11--15.

\leavevmode\vadjust pre{\hypertarget{ref-dubuis2011predicting}{}}%
Dubuis, A., Pottier, J., Rion, V., Pellissier, L., Theurillat, J.-P. \&
Guisan, A. (2011). Predicting spatial patterns of plant species
richness: A comparison of direct macroecological and species stacking
modelling approaches. \emph{Diversity and Distributions}, \textbf{17},
1122--1131.

\leavevmode\vadjust pre{\hypertarget{ref-elith2006novel}{}}%
Elith*, J., H. Graham*, C., P. Anderson, R., Dudik, M., Ferrier, S.,
Guisan, A., J. Hijmans, R., Huettmann, F., R. Leathwick, J., Lehmann, A.
\& others. (2006). Novel methods improve prediction of species'
distributions from occurrence data. \emph{Ecography}, \textbf{29},
129--151.

\leavevmode\vadjust pre{\hypertarget{ref-fazekas2009plant}{}}%
Fazekas, A.J., Kesanakurti, P.R., Burgess, K.S., Percy, D.M., Graham,
S.W., Barrett, S.C., Newmaster, S.G., Hajibabaei, M. \& Husband, B.C.
(2009). Are plant species inherently harder to discriminate than animal
species using DNA barcoding markers? \emph{Molecular Ecology Resources},
\textbf{9}, 130--139.

\leavevmode\vadjust pre{\hypertarget{ref-fenn2003ecological}{}}%
Fenn, M.E., Baron, J.S., Allen, E.B., Rueth, H.M., Nydick, K.R., Geiser,
L., Bowman, W.D., Sickman, J.O., Meixner, T., Johnson, D.W. \& others.
(2003). Ecological effects of nitrogen deposition in the western united
states. \emph{BioScience}, \textbf{53}, 404--420.

\leavevmode\vadjust pre{\hypertarget{ref-flora1993flora}{}}%
Flora of North America Editorial Committee, eds. (1993+). \emph{Flora of
north america north of mexico {[}online{]}}. Oxford University Press on
Demand.

\leavevmode\vadjust pre{\hypertarget{ref-fraser2007vpc}{}}%
Frase, Barbara A. \& Buck, P. (2007). {Vascular Plants of the Gothic
Area}. Retrieved from
\url{https://www.digitalrmbl.org/wp-content/uploads/2016/05/vascularplantlist_20071.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-Gage2013HistoricalRO}{}}%
Gage, E. \& Cooper, D.J. (2013). Historical range of variation
assessment for wetland and riparian ecosystems, u.s. Forest service
rocky mountain region

\leavevmode\vadjust pre{\hypertarget{ref-goulson2010bumblebees}{}}%
Goulson, D. (2010). \emph{Bumblebees: Behaviour, ecology, and
conservation}. Oxford University Press on Demand.

\leavevmode\vadjust pre{\hypertarget{ref-goulson2005causes}{}}%
Goulson, D., Hanley, M.E., Darvill, B., Ellis, J. \& Knight, M.E.
(2005). Causes of rarity in bumblebees. \emph{Biological conservation},
\textbf{122}, 1--8.

\leavevmode\vadjust pre{\hypertarget{ref-goulson2008decline}{}}%
Goulson, D., Lye, G. \& Darvill, B. (2008). The decline and conservation
of bumblebees. \emph{Annual review of entomology}, \textbf{53},
191--208.

\leavevmode\vadjust pre{\hypertarget{ref-grixti2009decline}{}}%
Grixti, J.C., Wong, L.T., Cameron, S.A. \& Favret, C. (2009). Decline of
bumble bees (bombus) in the north american midwest. \emph{Biological
conservation}, \textbf{142}, 75--84.

\leavevmode\vadjust pre{\hypertarget{ref-cbol2009dna}{}}%
Group, C.P.W., Hollingsworth, P.M., Forrest, L.L., Spouge, J.L.,
Hajibabaei, M., Ratnasingham, S., Bank, M. van der, Chase, M.W., Cowan,
R.S., Erickson, D.L. \& others. (2009). A DNA barcode for land plants.
\emph{Proceedings of the National Academy of Sciences}, \textbf{106},
12794--12797.

\leavevmode\vadjust pre{\hypertarget{ref-china2011comparative}{}}%
Group, C.P.B., Li, D.-Z., Gao, L.-M., Li, H.-T., Wang, H., Ge, X.-J.,
Liu, J.-Q., Chen, Z.-D., Zhou, S.-L., Chen, S.-L. \& others. (2011).
Comparative analysis of a large dataset indicates that internal
transcribed spacer (ITS) should be incorporated into the core barcode
for seed plants. \emph{Proceedings of the National Academy of Sciences},
\textbf{108}, 19641--19646.

\leavevmode\vadjust pre{\hypertarget{ref-havens2007chicago}{}}%
Havens, K., Vitt, P., Schwarz, J., Orr, B. \& Crimmins, T. (2007).
Chicago botanic garden's conservation and outreach efforts on climate
change. \emph{BGjournal}, \textbf{4}, 13--16.

\leavevmode\vadjust pre{\hypertarget{ref-hebert2003biological}{}}%
Hebert, P.D., Cywinska, A., Ball, S.L. \& DeWaard, J.R. (2003).
Biological identifications through DNA barcodes. \emph{Proceedings of
the Royal Society of London. Series B: Biological Sciences},
\textbf{270}, 313--321.

\leavevmode\vadjust pre{\hypertarget{ref-hengl2017soilgrids250m}{}}%
Hengl, T., Mendes de Jesus, J., Heuvelink, G.B., Ruiperez Gonzalez, M.,
Kilibarda, M., Blagotić, A., Shangguan, W., Wright, M.N., Geng, X.,
Bauer-Marschallinger, B. \& others. (2017). SoilGrids250m: Global
gridded soil information based on machine learning. \emph{PLoS one},
\textbf{12}, e0169748.

\leavevmode\vadjust pre{\hypertarget{ref-fpc2022}{}}%
Hennig, C. (2020). \emph{Fpc: Flexible procedures for clustering}.
Retrieved from \url{https://CRAN.R-project.org/package=fpc}

\leavevmode\vadjust pre{\hypertarget{ref-hinchliff2015synthesis}{}}%
Hinchliff, C.E., Smith, S.A., Allman, J.F., Burleigh, J.G., Chaudhary,
R., Coghill, L.M., Crandall, K.A., Deng, J., Drew, B.T., Gazis, R. \&
others. (2015). Synthesis of phylogeny and taxonomy into a comprehensive
tree of life. \emph{Proceedings of the National Academy of Sciences},
\textbf{112}, 12764--12769.

\leavevmode\vadjust pre{\hypertarget{ref-hitchcock2018flora}{}}%
Hitchcock, C.L. \& Cronquist, A. (2018). \emph{Flora of the pacific
northwest: An illustrated manual}. University of Washington Press.

\leavevmode\vadjust pre{\hypertarget{ref-hollingsworth2016telling}{}}%
Hollingsworth, P.M., Li, D.-Z., Bank, M. van der \& Twyford, A.D.
(2016). Telling plant species apart with DNA: From barcodes to genomes.
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \textbf{371}, 20150338.

\leavevmode\vadjust pre{\hypertarget{ref-inextPackage}{}}%
Hsieh, T.C., Ma, K.H. \& Chao, A. (2020). \emph{iNEXT: Interpolation and
extrapolation for species diversity}. Retrieved from
\url{http://chao.stat.nthu.edu.tw/wordpress/software_download/}

\leavevmode\vadjust pre{\hypertarget{ref-iler2021conceptual}{}}%
Iler, A.M., Humphrey, P.T., Ogilvie, J.E. \& CaraDonna, P.J. (2021).
Conceptual and practical issues limit the utility of statistical
estimators of phenological events. \emph{Ecosphere}, \textbf{12},
e03828.

\leavevmode\vadjust pre{\hypertarget{ref-janzen1967synchronization}{}}%
Janzen, D.H. (1967). Synchronization of sexual reproduction of trees
within the dry season in central america. \emph{Evolution}, \textbf{21},
620--637.

\leavevmode\vadjust pre{\hypertarget{ref-janzen2017nuclear}{}}%
Janzen, D.H., Burns, J.M., Cong, Q., Hallwachs, W., Dapkey, T.,
Manjunath, R., Hajibabaei, M., Hebert, P.D. \& Grishin, N.V. (2017).
Nuclear genomes distinguish cryptic species suggested by their DNA
barcodes and ecology. \emph{Proceedings of the National Academy of
Sciences}, \textbf{114}, 8313--8318.

\leavevmode\vadjust pre{\hypertarget{ref-jepson2022online}{}}%
\emph{Jepson flora project}. (2020).

\leavevmode\vadjust pre{\hypertarget{ref-johnson2021airborne}{}}%
Johnson, M.D., Fokar, M., Cox, R.D. \& Barnes, M.A. (2021). Airborne
environmental DNA metabarcoding detects more diversity, with less
sampling effort, than a traditional plant community survey. \emph{BMC
Ecology and Evolution}, \textbf{21}, 1--15.

\leavevmode\vadjust pre{\hypertarget{ref-johnson2016hybpiper}{}}%
Johnson, M.G., Gardner, E.M., Liu, Y., Medina, R., Goffinet, B., Shaw,
A.J., Zerega, N.J. \& Wickett, N.J. (2016). HybPiper: Extracting coding
sequence and introns for phylogenetics from high-throughput sequencing
reads using target enrichment. \emph{Applications in plant sciences},
\textbf{4}, 1600016.

\leavevmode\vadjust pre{\hypertarget{ref-johnson2019universal}{}}%
Johnson, M.G., Pokorny, L., Dodsworth, S., Botigue, L.R., Cowan, R.S.,
Devault, A., Eiserhardt, W.L., Epitawalage, N., Forest, F., Kim, J.T. \&
others. (2019). A universal probe set for targeted sequencing of 353
nuclear genes from any flowering plant designed using k-medoids
clustering. \emph{Systematic biology}, \textbf{68}, 594--606.

\leavevmode\vadjust pre{\hypertarget{ref-joshi2001local}{}}%
Joshi, J., Schmid, B., Caldeira, M., Dimitrakopoulos, P., Good, J.,
Harris, R., Hector, A., Huss-Danell, K., Jumpponen, A., Minns, A. \&
others. (2001). Local adaptation enhances performance of common plant
species. \emph{Ecology Letters}, \textbf{4}, 536--544.

\leavevmode\vadjust pre{\hypertarget{ref-keane2002cascading}{}}%
Keane, R.E. (2002). Cascading effects of fire exclusion in rocky
mountain ecosystems: A literature review.

\leavevmode\vadjust pre{\hypertarget{ref-kearns2001natural}{}}%
Kearns, C.A., Thomson, J.D. \& others. (2001). \emph{Natural history of
bumblebees}. University Press of Colorado.

\leavevmode\vadjust pre{\hypertarget{ref-kramer2015report}{}}%
Kramer, A.T. \& Havens, K. (2015). Report in brief: Assessing botanical
capacity to address grand challenges in the united states. \emph{Natural
Areas Journal}, \textbf{35}, 83--89.

\leavevmode\vadjust pre{\hypertarget{ref-kress2017plant}{}}%
Kress, W.J. (2017). Plant DNA barcodes: Applications today and in the
future. \emph{Journal of systematics and evolution}, \textbf{55},
291--307.

\leavevmode\vadjust pre{\hypertarget{ref-kress2007two}{}}%
Kress, W.J. \& Erickson, D.L. (2007). A two-locus global DNA barcode for
land plants: The coding rbcL gene complements the non-coding trnH-psbA
spacer region. \emph{PLoS one}, \textbf{2}, e508.

\leavevmode\vadjust pre{\hypertarget{ref-caret}{}}%
Kuhn, M. (2022). \emph{Caret: Classification and regression training}.
Retrieved from \url{https://CRAN.R-project.org/package=caret}

\leavevmode\vadjust pre{\hypertarget{ref-lamb2019quantitative}{}}%
Lamb, P.D., Hunter, E., Pinnegar, J.K., Creer, S., Davies, R.G. \&
Taylor, M.I. (2019). How quantitative is metabarcoding: A
meta-analytical approach. \emph{Molecular ecology}, \textbf{28},
420--430.

\leavevmode\vadjust pre{\hypertarget{ref-aim2019database}{}}%
Land Management, B. of. (2019). U.s. Department of interior bureau of
land management, BLM - assessment, inventory, and monitoring (AIM)
terrestrial indicators raw dataset. Retrieved from
\url{https://gbp-blm-egis.hub.arcgis.com/pages/aim}

\leavevmode\vadjust pre{\hypertarget{ref-lang2019genome}{}}%
Lang, D., Tang, M., Hu, J. \& Zhou, X. (2019). Genome-skimming provides
accurate quantification for pollen mixtures. \emph{Molecular Ecology
Resources}, \textbf{19}, 1433--1446.

\leavevmode\vadjust pre{\hypertarget{ref-lesica2012manual}{}}%
Lesica, P., Lavin, M. \& Stickney, P.F. (2012). \emph{Manual of montana
vascular plants}. Brit Press.

\leavevmode\vadjust pre{\hypertarget{ref-lewin2022biogenome}{}}%
Lewin, H.A., Richards, S., Aiden, E.L., Allende, M.L., Archibald, J.M.,
Bálint, M., Barker, K.B., Baumgartner, B., Belov, K., Bertorelle, G.,
Blaxter, M.L., Cai, J., Caperello, N.D., Carlson, K., Castilla-Rubio,
J.C., Chaw, S.-M., Chen, L., Childers, A.K., Coddington, J.A., Conde,
D.A., Corominas, M., Crandall, K.A., Crawford, A.J., DiPalma, F.,
Durbin, R., Ebenezer, T.E., Edwards, S.V., Fedrigo, O., Flicek, P.,
Formenti, G., Gibbs, R.A., Gilbert, M.T.P., Goldstein, M.M., Graves,
J.M., Greely, H.T., Grigoriev, I.V., Hackett, K.J., Hall, N., Haussler,
D., Helgen, K.M., Hogg, C.J., Isobe, S., Jakobsen, K.S., Janke, A.,
Jarvis, E.D., Johnson, W.E., Jones, S.J.M., Karlsson, E.K., Kersey,
P.J., Kim, J.-H., Kress, W.J., Kuraku, S., Lawniczak, M.K.N.,
Leebens-Mack, J.H., Li, X., Lindblad-Toh, K., Liu, X., Lopez, J.V.,
Marques-Bonet, T., Mazard, S., Mazet, J.A.K., Mazzoni, C.J., Myers,
E.W., O'Neill, R.J., Paez, S., Park, H., Robinson, G.E., Roquet, C.,
Ryder, O.A., Sabir, J.S.M., Shaffer, H.B., Shank, T.M., Sherkow, J.S.,
Soltis, P.S., Tang, B., Tedersoo, L., Uliano-Silva, M., Wang, K., Wei,
X., Wetzer, R., Wilson, J.L., Xu, X., Yang, H., Yoder, A.D. \& Zhang, G.
(2022). The earth BioGenome project 2020: Starting the clock.
\emph{Proceedings of the National Academy of Sciences}, \textbf{119},
e2115635118. Retrieved from
\url{https://www.pnas.org/doi/abs/10.1073/pnas.2115635118}

\leavevmode\vadjust pre{\hypertarget{ref-liang2021evolutionary}{}}%
Liang, H., Zhao, Y.-H., Rafferty, N.E., Ren, Z.-X., Zhong, L., Li,
H.-D., Li, D.-Z. \& Wang, H. (2021). Evolutionary and ecological factors
structure a plant--bumblebee network in a biodiversity hotspot, the
himalaya--hengduan mountains. \emph{Functional Ecology}, \textbf{35},
2523--2535.

\leavevmode\vadjust pre{\hypertarget{ref-darwin2022project}{}}%
Life Project Consortium, D.T. of, Blaxter, M., Mieszkowska, N., Palma,
F.D., Holland, P., Durbin, R., Richards, T., Berriman, M., Kersey, P.,
Hollingsworth, P., Wilson, W., Twyford, A., Gaya, E., Lawniczak, M.,
Lewis, O., Broad, G., Howe, K., Hart, M., Flicek, P. \& Barnes, I.
(2022). Sequence locally, think globally: The darwin tree of life
project. \emph{Proceedings of the National Academy of Sciences},
\textbf{119}, e2115642118. Retrieved from
\url{https://www.pnas.org/doi/abs/10.1073/pnas.2115642118}

\leavevmode\vadjust pre{\hypertarget{ref-li2016responses}{}}%
Li, X., Jiang, L., Meng, F., Wang, S., Niu, H., Iler, A.M., Duan, J.,
Zhang, Z., Luo, C., Cui, S. \& others. (2016). Responses of sequential
and hierarchical phenological events to warming and cooling in alpine
meadows. \emph{Nature Communications}, \textbf{7}, 1--8.

\leavevmode\vadjust pre{\hypertarget{ref-liu2014identification}{}}%
Liu, J., Shi, L., Han, J., Li, G., Lu, H., Hou, J., Zhou, X., Meng, F.
\& Downie, S.R. (2014). Identification of species in the angiosperm
family apiaceae using DNA barcodes. \emph{Molecular ecology resources},
\textbf{14}, 1231--1238.

\leavevmode\vadjust pre{\hypertarget{ref-li2015plant}{}}%
Li, X., Yang, Y., Henry, R.J., Rossetto, M., Wang, Y. \& Chen, S.
(2015). Plant DNA barcoding: From gene to genome. \emph{Biological
Reviews}, \textbf{90}, 157--166.

\leavevmode\vadjust pre{\hypertarget{ref-loarie2009velocity}{}}%
Loarie, S.R., Duffy, P.B., Hamilton, H., Asner, G.P., Field, C.B. \&
Ackerly, D.D. (2009). The velocity of climate change. \emph{Nature},
\textbf{462}, 1052--1055.

\leavevmode\vadjust pre{\hypertarget{ref-lu2017bracken}{}}%
Lu, J., Breitwieser, F.P., Thielen, P. \& Salzberg, S.L. (2017).
Bracken: Estimating species abundance in metagenomics data. \emph{PeerJ
Computer Science}, \textbf{3}, e104.

\leavevmode\vadjust pre{\hypertarget{ref-cluster2022}{}}%
Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M. \& Hornik, K.
(2022). \emph{Cluster: Cluster analysis basics and extensions}.
Retrieved from \url{https://CRAN.R-project.org/package=cluster}

\leavevmode\vadjust pre{\hypertarget{ref-mainali2022alpha}{}}%
Mainali, K. \& Slud, E. (2022). \emph{CooccurrenceAffinity: Affinity in
cooccrrence data}.

\leavevmode\vadjust pre{\hypertarget{ref-mainali2022better}{}}%
Mainali, K.P., Slud, E., Singer, M.C. \& Fagan, W.F. (2022). A better
index for analysis of co-occurrence and similarity. \emph{Science
Advances}, \textbf{8}, eabj9204.

\leavevmode\vadjust pre{\hypertarget{ref-bien2022}{}}%
Maitner, B. (2022). \emph{BIEN: Tools for accessing the botanical
information and ecology network database}. Retrieved from
\url{https://CRAN.R-project.org/package=BIEN}

\leavevmode\vadjust pre{\hypertarget{ref-manzano2021flippant}{}}%
Manzano, S. (2021). Flippant attitudes towards plant identification
jeopardize early career botanists. \emph{Trends in Plant Science},
\textbf{26}, 987--988.

\leavevmode\vadjust pre{\hypertarget{ref-mclay2021new}{}}%
McLay, T.G., Birch, J.L., Gunn, B.F., Ning, W., Tate, J.A., Nauheimer,
L., Joyce, E.M., Simpson, L., Schmidt-Lebuhn, A.N., William J \& others.
(2021). New targets acquired: Improving locus recovery from the
Angiosperms353 probe set. \emph{Applications in plant sciences},
\textbf{9}.

\leavevmode\vadjust pre{\hypertarget{ref-mohlenbrock2002vascular}{}}%
Mohlenbrock, R.H. (2002). \emph{Vascular flora of illinois}. SIU Press.

\leavevmode\vadjust pre{\hypertarget{ref-moore2003its}{}}%
Moore, A.J. \& Bohs, L. (2003). An ITS phylogeny of balsamorhiza and
wyethia (asteraceae: heliantheae). \emph{American Journal of Botany},
\textbf{90}, 1653--1660.

\leavevmode\vadjust pre{\hypertarget{ref-naiman1988alteration}{}}%
Naiman, R.J., Johnston, C.A. \& Kelley, J.C. (1988). Alteration of north
american streams by beaver. \emph{BioScience}, \textbf{38}, 753--762.

\leavevmode\vadjust pre{\hypertarget{ref-sdmPackage}{}}%
Naimi, B. \& Araujo, M.B. (2016).
\href{https://doi.org/10.1111/ecog.01881}{Sdm: A reproducible and
extensible r platform for species distribution modelling}.
\emph{Ecography}, \textbf{39}, 368--375.

\leavevmode\vadjust pre{\hypertarget{ref-usdm2014}{}}%
Naimi, B., Hamm, N. a.s., Groen, T.A., Skidmore, A.K. \& Toxopeus, A.G.
(2014). \href{https://doi.org/10.1111/j.1600-0587.2013.00205.x}{Where is
positional uncertainty a problem for species distribution modelling}.
\emph{Ecography}, \textbf{37}, 191--203.

\leavevmode\vadjust pre{\hypertarget{ref-nazaire2014phylogenetic}{}}%
Nazaire, M. \& Hufford, L. (2014). Phylogenetic systematics of the genus
mertensia (boraginaceae). \emph{Systematic Botany}, \textbf{39},
268--303.

\leavevmode\vadjust pre{\hypertarget{ref-nevado2016widespread}{}}%
Nevado, B., Atchison, G.W., Hughes, C.E. \& Filatov, D.A. (2016).
Widespread adaptive evolution during repeated evolutionary radiations in
new world lupins. \emph{Nature communications}, \textbf{7}, 1--9.

\leavevmode\vadjust pre{\hypertarget{ref-Newstrom1994ANC}{}}%
Newstrom, L.E., Frankie, G.W. \& Baker, H.G. (1994). A new
classification for plant phenology based on flowering patterns in
lowland tropical rain forest trees at la selva, costa rica.
\emph{Biotropica}, \textbf{26}, 141--159.

\leavevmode\vadjust pre{\hypertarget{ref-oberhauser2015monarchs}{}}%
Oberhauser, K.S., Nail, K.R. \& Altizer, S. (2015). \emph{Monarchs in a
changing world: Biology and conservation of an iconic butterfly}.
Cornell University Press.

\leavevmode\vadjust pre{\hypertarget{ref-gbifDL2021sdms}{}}%
Occdownload Gbif.Org. (2021). Occurrence download. Retrieved from
\url{https://www.gbif.org/occurrence/download/0206948-200613084148143}

\leavevmode\vadjust pre{\hypertarget{ref-vegans2022}{}}%
Oksanen, J., Simpson, G.L., Blanchet, F.G., Kindt, R., Legendre, P.,
Minchin, P.R., O'Hara, R.B., Solymos, P., Stevens, M.H.H., Szoecs, E.,
Wagner, H., Barbour, M., Bedward, M., Bolker, B., Borcard, D., Carvalho,
G., Chirico, M., De Caceres, M., Durand, S., Evangelista, H.B.A.,
FitzJohn, R., Friendly, M., Furneaux, B., Hannigan, G., Hill, M.O.,
Lahti, L., McGlinn, D., Ouellette, M.-H., Ribeiro Cunha, E., Smith, T.,
Stier, A., Ter Braak, C.J.F. \& Weedon, J. (2022). \emph{Vegan:
Community ecology package}. Retrieved from
\url{https://CRAN.R-project.org/package=vegan}

\leavevmode\vadjust pre{\hypertarget{ref-oliver2009cryptic}{}}%
Oliver, P.M., Adams, M., Lee, M.S., Hutchinson, M.N. \& Doughty, P.
(2009). Cryptic diversity in vertebrates: Molecular data double
estimates of species diversity in a radiation of australian lizards
(diplodactylus, gekkota). \emph{Proceedings of the Royal Society B:
Biological Sciences}, \textbf{276}, 2001--2007.

\leavevmode\vadjust pre{\hypertarget{ref-omernik1987ecoregions}{}}%
Omernik, J.M. (1987). Ecoregions of the conterminous united states.
\emph{Annals of the Association of American geographers}, \textbf{77},
118--125.

\leavevmode\vadjust pre{\hypertarget{ref-ottenlips2021resolving}{}}%
Ottenlips, M.V., Mansfield, D.H., Buerki, S., Feist, M.A.E., Downie,
S.R., Dodsworth, S., Forest, F., Plunkett, G.M. \& Smith, J.F. (2021).
Resolving species boundaries in a recent radiation with the
Angiosperms353 probe set: The lomatium packardiae/l. Anomalum clade of
the l. Triternatum (apiaceae) complex. \emph{American journal of
botany}, \textbf{108}, 1217--1233.

\leavevmode\vadjust pre{\hypertarget{ref-park2022herbarium}{}}%
Park, D.S., Lyra, G.M., Ellison, A.M., Maruyama, R.K.B., Torquato, D.
dos R., Asprino, R.C., Cook, B.I. \& Davis, C.C. (2022). Herbarium
records provide reliable phenology estimates in the understudied
tropics. \emph{Journal of Ecology}.

\leavevmode\vadjust pre{\hypertarget{ref-pearse2017statistical}{}}%
Pearse, W.D., Davis, C.C., Inouye, D.W., Primack, R.B. \& Davies, T.J.
(2017). A statistical estimator for determining the limits of
contemporary and historic phenology. \emph{Nature Ecology \& Evolution},
\textbf{1}, 1876--1882.

\leavevmode\vadjust pre{\hypertarget{ref-peel2019semi}{}}%
Peel, N., Dicks, L.V., Clark, M.D., Heavens, D., Percival-Alwyn, L.,
Cooper, C., Davies, R.G., Leggett, R.M. \& Yu, D.W. (2019).
Semi-quantitative characterisation of mixed pollen samples using MinION
sequencing and reverse metagenomics (RevMet). \emph{Methods in Ecology
and Evolution}, \textbf{10}, 1690--1701.

\leavevmode\vadjust pre{\hypertarget{ref-pepin2022climate}{}}%
Pepin, N., Arnone, E., Gobiet, A., Haslinger, K., Kotlarski, S.,
Notarnicola, C., Palazzi, E., Seibert, P., Serafin, S., Schöner, W. \&
others. (2022). Climate changes and their elevational patterns in the
mountains of the world. \emph{Reviews of geophysics}, \textbf{60},
e2020RG000730.

\leavevmode\vadjust pre{\hypertarget{ref-pinto2021predicting}{}}%
Pinto-Ledezma, J.N. \& Cavender-Bares, J. (2021). Predicting species
distributions and community composition using satellite remote sensing
predictors. \emph{Scientific Reports}, \textbf{11}, 1--12.

\leavevmode\vadjust pre{\hypertarget{ref-pornon2017dna}{}}%
Pornon, A., Andalo, C., Burrus, M. \& Escaravage, N. (2017). DNA
metabarcoding data unveils invisible pollination networks.
\emph{Scientific Reports}, \textbf{7}, 1--11.

\leavevmode\vadjust pre{\hypertarget{ref-prather2004implications}{}}%
Prather, L.A., Alvarez-Fuentes, O., Mayfield, M.H. \& Ferguson, C.J.
(2004a). Implications of the decline in plant collecting for systematic
and floristic research. \emph{Systematic Botany}, \textbf{29}, 216--220.

\leavevmode\vadjust pre{\hypertarget{ref-prather2004decline}{}}%
Prather, L.A., Alvarez-Fuentes, O., Mayfield, M.H. \& Ferguson, C.J.
(2004b). The decline of plant collecting in the united states: A threat
to the infrastructure of biodiversity studies. \emph{Systematic Botany},
\textbf{29}, 15--28.

\leavevmode\vadjust pre{\hypertarget{ref-prim1957minimum}{}}%
Prim, R.C. (1957). Shortest connection networks and some
generalisations. \emph{Bell System Technical Journal}, \textbf{36},
1389--1401.

\leavevmode\vadjust pre{\hypertarget{ref-pusalkar2015taxonomic}{}}%
Pusalkar, P.K. \& Singh, D.K. (2015). Taxonomic rearrangement of
arenaria (caryophyllaceae) in indian western himalaya. \emph{Journal of
Japanese Botany}, \textbf{90}, 77--91.

\leavevmode\vadjust pre{\hypertarget{ref-pyke1982local}{}}%
Pyke, G.H. (1982). Local geographic distributions of bumblebees near
crested butte, colorado: Competition and community structure.
\emph{Ecology}, \textbf{63}, 555--573.

\leavevmode\vadjust pre{\hypertarget{ref-qiao2015no}{}}%
Qiao, H., Soberon, J. \& Peterson, A.T. (2015). No silver bullets in
correlative ecological niche modelling: Insights from testing among many
potential algorithms for niche estimation. \emph{Methods in Ecology and
Evolution}, \textbf{6}, 1126--1136.

\leavevmode\vadjust pre{\hypertarget{ref-rabeler2016new}{}}%
Rabeler, R.K. \& Wagner, W.L. (2016). New combinations in odontostemma
(caryophyllaceae). \emph{PhytoKeys}, 77.

\leavevmode\vadjust pre{\hypertarget{ref-ralphs1988herbicide}{}}%
Ralphs, M.H. \& Ueckert, D.N. (1988). Herbicide control of locoweeds: A
review. \emph{Weed Technology}, \textbf{2}, 460--465.

\leavevmode\vadjust pre{\hypertarget{ref-ralphs2003mechanism}{}}%
Ralphs, M., Woolsey, L. \& Bowns, J. (2003). Mechanism by which ammonium
fertilizers kill tall larkspur. \emph{Rangeland Ecology \&
Management/Journal of Range Management Archives}, \textbf{56}, 524--528.

\leavevmode\vadjust pre{\hypertarget{ref-robinson2014earthenv}{}}%
Robinson, N., Regetz, J. \& Guralnick, R.P. (2014). EarthEnv-DEM90: A
nearly-global, void-free, multi-scale smoothed, 90m digital elevation
model from fused ASTER and SRTM data. \emph{ISPRS Journal of
Photogrammetry and Remote Sensing}, \textbf{87}, 57--67.

\leavevmode\vadjust pre{\hypertarget{ref-ruppert2019past}{}}%
Ruppert, K.M., Kline, R.J. \& Rahman, M.S. (2019). Past, present, and
future perspectives of environmental DNA (eDNA) metabarcoding: A
systematic review in methods, monitoring, and applications of global
eDNA. \emph{Global Ecology and Conservation}, \textbf{17}, e00547.

\leavevmode\vadjust pre{\hypertarget{ref-sadeghian2015molecular}{}}%
Sadeghian, S., Zarre, S., Rabeler, R.K. \& Heubl, G. (2015). Molecular
phylogenetic analysis of arenaria (caryophyllaceae: Tribe arenarieae)
and its allies inferred from nuclear DNA internal transcribed spacer and
plastid DNA rps16 sequences. \emph{Botanical Journal of the Linnean
Society}, \textbf{178}, 648--669.

\leavevmode\vadjust pre{\hypertarget{ref-sarro2022bumble}{}}%
Sarro, E., Tripodi, A. \& Woodard, S.H. (2022). Bumble bee (bombus
vosnesenskii) queen nest searching occurs independent of ovary
developmental status. \emph{Integrative Organismal Biology}, \textbf{4},
obac007.

\leavevmode\vadjust pre{\hypertarget{ref-schmitt2017ssdm}{}}%
Schmitt, S., Pouteau, R., Justeau, D., De Boissieu, F. \& Birnbaum, P.
(2017). Ssdm: An r package to predict distribution of species richness
and composition based on stacked species distribution models.
\emph{Methods in Ecology and Evolution}, \textbf{8}, 1795--1803.

\leavevmode\vadjust pre{\hypertarget{ref-sennikov2017phylogenetic}{}}%
Sennikov, A.N. \& Kurtto, A. (2017). A phylogenetic checklist of sorbus
sl (rosaceae) in europe.

\leavevmode\vadjust pre{\hypertarget{ref-sickel2015increased}{}}%
Sickel, W., Ankenbrand, M.J., Grimmer, G., Holzschuh, A., Hartel, S.,
Lanzen, J., Steffan-Dewenter, I. \& Keller, A. (2015). Increased
efficiency in identifying mixed pollen samples by meta-barcoding with a
dual-indexing approach. \emph{BMC ecology}, \textbf{15}, 1--9.

\leavevmode\vadjust pre{\hypertarget{ref-slimp2021potential}{}}%
Slimp, M., Williams, L.D., Hale, H. \& Johnson, M.G. (2021). On the
potential of Angiosperms353 for population genomic studies.
\emph{Applications in Plant Sciences}, \textbf{9}.

\leavevmode\vadjust pre{\hypertarget{ref-smith2018constructing}{}}%
Smith, S.A. \& Brown, J.W. (2018). Constructing a broadly inclusive seed
plant phylogeny. \emph{American journal of botany}, \textbf{105},
302--314.

\leavevmode\vadjust pre{\hypertarget{ref-sork2018genomic}{}}%
Sork, V.L. (2018). Genomic studies of local adaptation in natural plant
populations. \emph{Journal of Heredity}, \textbf{109}, 3--15.

\leavevmode\vadjust pre{\hypertarget{ref-stevens2018atmospheric}{}}%
Stevens, C.J., David, T.I. \& Storkey, J. (2018). Atmospheric nitrogen
deposition in terrestrial ecosystems: Its impact on plant communities
and consequences across trophic levels. \emph{Functional Ecology},
\textbf{32}, 1757--1769.

\leavevmode\vadjust pre{\hypertarget{ref-stroud2022botanical}{}}%
Stroud, S., Fennell, M., Mitchley, J., Lydon, S., Peacock, J. \& Bacon,
K.L. (2022). The botanical education extinction and the fall of plant
awareness. \emph{Ecology and Evolution}, \textbf{12}, e9019.

\leavevmode\vadjust pre{\hypertarget{ref-suchan2019pollen}{}}%
Suchan, T., Talavera, G., Saez, L., Ronikier, M. \& Vila, R. (2019).
Pollen metabarcoding as a tool for tracking long-distance insect
migrations. \emph{Molecular Ecology Resources}, \textbf{19}, 149--162.

\leavevmode\vadjust pre{\hypertarget{ref-tange_2022_6377950}{}}%
Tange, O. (2021). GNU parallel 20220322 (savannah). Retrieved from
\url{https://doi.org/10.5281/zenodo.6377950}

\leavevmode\vadjust pre{\hypertarget{ref-thompson1994coevolutionary}{}}%
Thompson, J.N. (1994). \emph{The coevolutionary process}. University of
Chicago press.

\leavevmode\vadjust pre{\hypertarget{ref-tran2019cloud}{}}%
Tran, H., Nguyen, P., Ombadi, M., Hsu, K., Sorooshian, S. \& Qing, X.
(2019). A cloud-free MODIS snow cover dataset for the contiguous united
states from 2000 to 2017. \emph{Scientific data}, \textbf{6}, 1--13.

\leavevmode\vadjust pre{\hypertarget{ref-wang2016locally}{}}%
Wang, T., Hamann, A., Spittlehouse, D. \& Carroll, C. (2016). Locally
downscaled and spatially customizable climate data for historical and
future periods for north america. \emph{PloS one}, \textbf{11},
e0156720.

\leavevmode\vadjust pre{\hypertarget{ref-weber1998new}{}}%
Weber, W. (1998). New names and combinations in asteraceae:
Heliantheae-ecliptinae. \emph{Phytologia}, \textbf{85}, 19--21.

\leavevmode\vadjust pre{\hypertarget{ref-wenzell2021incomplete}{}}%
Wenzell, K.E., McDonnell, A.J., Wickett, N.J., Fant, J.B. \& Skogen,
K.A. (2021). Incomplete reproductive isolation and low genetic
differentiation despite floral divergence across varying geographic
scales in castilleja. \emph{American Journal of Botany}, \textbf{108},
1270--1288.

\leavevmode\vadjust pre{\hypertarget{ref-wilkinson2021defining}{}}%
Wilkinson, D.P., Golding, N., Guillera-Arroita, G., Tingley, R. \&
McCarthy, M.A. (2021). Defining and evaluating predictions of joint
species distribution models. \emph{Methods in Ecology and Evolution},
\textbf{12}, 394--404.

\leavevmode\vadjust pre{\hypertarget{ref-williams1982distribution}{}}%
Williams, P.H. (1982). The distribution and decline of british bumble
bees (bombus latr.). \emph{Journal of Apicultural Research},
\textbf{21}, 236--245. Retrieved from
\url{https://doi.org/10.1080/00218839.1982.11100549}

\leavevmode\vadjust pre{\hypertarget{ref-wilson2016remotely}{}}%
Wilson, A.M. \& Jetz, W. (2016). Remotely sensed high-resolution global
cloud dynamics for predicting ecosystem and biodiversity distributions.
\emph{PLoS biology}, \textbf{14}, e1002415.

\leavevmode\vadjust pre{\hypertarget{ref-wood2019improved}{}}%
Wood, D.E., Lu, J. \& Langmead, B. (2019). Improved metagenomic analysis
with kraken 2. \emph{Genome biology}, \textbf{20}, 1--13.

\leavevmode\vadjust pre{\hypertarget{ref-xie2015deciduous}{}}%
Xie, Y., Wang, X. \& Silander Jr, J.A. (2015). Deciduous forest
responses to temperature, precipitation, and drought imply complex
climate change impacts. \emph{Proceedings of the National Academy of
Sciences}, \textbf{112}, 13585--13590.

\leavevmode\vadjust pre{\hypertarget{ref-xie2018predicting}{}}%
Xie, Y., Wang, X., Wilson, A.M. \& Silander Jr, J.A. (2018). Predicting
autumn phenology: How deciduous tree species respond to weather
stressors. \emph{Agricultural and Forest Meteorology}, \textbf{250},
127--137.

\leavevmode\vadjust pre{\hypertarget{ref-zhao2019topology}{}}%
Zhao, Y.-H., Lázaro, A., Ren, Z.-X., Zhou, W., Li, H.-D., Tao, Z.-B.,
Xu, K., Wu, Z.-K., Wolfe, L.M., Li, D.-Z. \& Wang, H. (2019). The
topological differences between visitation and pollen transport
networks: A comparison in species rich communities of the
himalaya--hengduan mountains. \emph{Oikos}, \textbf{128}, 551--562.
Retrieved from
\url{https://onlinelibrary.wiley.com/doi/abs/10.1111/oik.05262}

\end{CSLReferences}

\end{document}

### 2023-03-03 script/manuscript_MEE.Rmd
refined text, mapped appendices

@@
 author:  |
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive^$$^3$^, Emily J. Woodworth$^1$$^,$$^2$, 
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$, Emily J. Woodworth$^1$$^,$$^2$, 
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
 library(knitr)
+library(kableExtra)
 #devtools::install_github("kpmainali/CooccurrenceAffinity")
@@
 
-Bee and flower observations and bee corbiculae pollen collection was conducted around the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (APPENDIX 1 for site information).
+Bee and flower observations and bee corbiculae pollen collection was conducted around the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (Appendix 1 for site information).
 The area is characterized by subalpine meadow vegetation communities.
@@
 These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (APPENDIX NO.).
+Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (Appendix 6).
 This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
@@
 The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in *APPENDIX 3*.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in Appendix 4.
   To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
@@
 Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects *APPENDIX 2*.
+Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects (Appendix 7 & 8).
 
@@
 
-To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (*APPENDIX 6*).
+To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (Appendix 2).
 These publicly available data sets, were selected as they pertain to a wide range of variables interacting with plant physiology.
@@
 In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
-We also sequenced another 15 taxa commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (*APPENDIX 4*).
+We also sequenced another 15 taxa commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (Appendix 3).
 Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
@@
 The Kraken2 database was built using default parameters.
-Kraken2 was run on sequences using default parameters (*APPENDIX 5*).
+Kraken2 was run on sequences using default parameters (Appendix 9).
   Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken).
@@
 
-Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into.
-*APPENDIX 7*.
-From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we **had over % of expected morphotypes found** (morphotype richness, $\bar{x}$ = 4.5, Mdn = 4, min = 1, max = 9), **all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*.** The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, Mdn = `r median(grains_counted)`).
-
-**SUMMARY REQUIRED - Number of species from number of families, and how many species are identified to species versus number only to genera **
+Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into (Appendix 6).
+From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we identified substantial amounts of the abundance and morphotype richness of the samples (morphotype richness, $\bar{x}$ = 4.5, median = 4, min = 1, max = 9) (Appendix  7 & 8).
+The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, median = `r  median(grains_counted)`).
 
@@
 Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
-The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))`.
-**APPENDIX X Reads Per Loci**.
+The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
 
@@
 `r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers **APPENDIX X MOLECULAR NETWORKS - 3 DIFFERENT ONES**, BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
+Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
 
@@
 The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
-A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix XX).
+A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix 15).
 Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
@@
 Of the `r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were all assigned via post-processing conditions (`r j`, APPENDIX XX).
+These were all assigned via post-processing conditions (`r j`, Appendix 15).
 These were manually assigned to a variety of ranks, occasionally to genus - `r nrow(gen)`, and species - `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
@@
 nevadensis* Cresson and *B.
-rufocinctus* Cresson pollen is most likely *Frasera* Walter, which failed extractions for the reference library failed (APPENDIX XX).
+rufocinctus* Cresson pollen is most likely *Frasera* Walter, which failed extractions for the reference library failed (Appendix 11).
 A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* (A.
@@
 
-```{r compare results}
+```{r compare results, out.width="50%"}
 
@@
 
-ggplot(ACC_sens_Spec, aes(y = Sp_Acc,  x= Method, color = Method)) +
-  geom_boxplot() +
-  geom_jitter(width = 0.1) +
-  theme_bw()
-
-wilcox.test(
+wt_res <- wilcox.test(
   pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'COMP', 'Sp_Acc']) , 
@@
 
-<final integrative results here>
-
 ```{r}
@@
 the 'linear models', and the 'machine learning'), which took notable amounts of compute power.
-We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicitly process to reduce spatial autocorrelation.
+We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicit process to reduce spatial autocorrelation.
 However, given the time required to perform all aspects of a study, even our amount of computation was negligible.
@@
 We predict that the A353 probes which it is utilizing to work nearly immediately for DNA barcoding of whole plant material, and that more elaborate validation studies in controlled metabarcoding settings, utilizing existing experimental designs, will have favorable results (@bell2017applying, @bell2019quantitative, @bell2021comparing, @lamb2019quantitative).
-In particular the harvesting of loci with more variation in certain lineages, and or with more variable flanking regions, will prove promising for identifying closely related plant material (CITE).
+In particular the harvesting of loci with more variation in certain lineages, and or with more variable flanking regions, will prove promising for identifying closely related plant material.
 We suspect that conserved reaches of genes resulted in the high amounts of reads in somewhat obscure species.
@@
 We expect that within a bayesian framework studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
-the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems.
+the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems (.
 We hope that the promise of A353 probes as tools for metabarcoding play a role in these endeavors.
@@
 
-### POLLEN CLUSTER RESULTS SHOULD BE HERE
-
-\newpage
-
-Appendix XX - Pollen Key
+Appendix 5 - Pollen Dendrogram
+```{r}
+
+```
+
+\newpage
+
+Appendix 6 - Pollen Key
 
@@
 
-APPENDIX XX - Pollen Morphotype Richness Rarefaction Curves
+Appendix 7 - Pollen Morphotype Richness Rarefaction Curves
 
@@
 
-Appendix XX - Pollen Morphotype Abundance Rarefaction Curves
+Appendix 8 - Pollen Morphotype Abundance Rarefaction Curves
 
@@
 
-Appendix XX - All Species in the Sequence Databases
+Appendix 9 - All Species in the Sequence Databases
 
@@
 
-Appendix XX - All Species in the Sequence Databases (con't)
+Appendix 10 - All Species in the Sequence Databases (con't)
 
@@
 
-Appendix XX - Reads Per Loci
+Appendix 11 - Reads Per Loci
 
@@
 
-Appendix XX - Comparison of Kraken2, Bracken, and BLAST
+Appendix 12 - Comparison of Kraken2, Bracken, and BLAST
 
@@
 
-Appendix XX - Models used for Species Distribution Model Ensembles
+Appendix 13 - Models used for Species Distribution Model Ensembles
 
@@
 
-APPENDIX XX - Time Spent Generating Species Distribution Models
+APPENDIX 14 - Time Spent Generating Species Distribution Models
 
@@
 
-APPENDIX XX - Review Classified Reads and Reassign
+APPENDIX 15 - Review Classified Reads and Reassign
 
@@
 
-APPENDIX XX - Tips and tricks for implementing plant metagenomic sequencing
+APPENDIX 16 - Tips and tricks for implementing plant metagenomic sequencing
 
@@
 
-THIS SHOULD BE TURNED INTO A SMALLER PNG, AND HAVE THE NUMBER OF SEQUENCED METAGENOMIC SAMPLES PLACED INTO THAT COLUMN AND INCLUDED IN TEXT ~~~ NEED THIS !!!!
+Appendix 17 Samples Per Bee
 
@@
 
+Appendix 18 - Overview of the Whole Process
+
 ```{r Overview of entire process, out.height="125%"}

### 2023-03-05 script/manuscript_MEE.Rmd
resized flowchart, added figure captions to paper

@@
 --- 
-title: Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees
+title: Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from Queen Bumble Bees
 author:  |
@@
   knitr::kable(caption = 'Applications of Plant Metabarcoding',
-             col.names = c('Application', 'Example', 'Citations'))
+             col.names = c('Application', 'Example', 'Citations')) |>
+  kable_styling(full_width = TRUE)
 ```
@@
   knitr::kable(caption = 'Current Issues Facing Plant Metagenomics',
-             col.names = c('Issue', 'Our Approach', 'Issue Citation'))
+             col.names = c('Issue', 'Our Approach', 'Possible Advances')) |>
+  kable_styling(full_width = TRUE)
 ```
@@
 In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks, which reduces the utility of these to operate as filters in the post-processing of sequence matches.
-Nonetheless, we work develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new).
+Nonetheless, we work develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new, Figure 1).
+
+```{r Filter Diagram, out.width = "35%", fig.cap="Simplified Conceptual Diagram of three approaches leading towards classification of sequencing results, and the number of species associated with them in our area.
+The upper three boxes indicate a common coarse approach, assuming one has a digitizewd Flora, which is not always the case.
+The center two boxes indicate the computational approach illustrated here.
+The final box indicates the use of the expert field data in the case study.
+The stem of the final applies to both Computational and Expert Survey results, and should be thought of as using time ala chromatography, in the post-classification process."}
+knitr::include_graphics('../graphics/plots/filter_diagram.png')
+```
 
@@
 Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken).
-Results from both Kraken2 and Bracken, results were reclassified manually to identify terminal taxa.
-For example, when only a single species of a genus was known in the study area, but our database used a representative of another taxon in the genus, this species was coded as the result.
+FInally all reads which could be classified by these databases were passed to a local BLAST database.
 
@@
 
-For example a many sequences which mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
+For example many sequences which mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
 nevadensis* Cresson and *B.
@@
 Heller) Rydb.
-but where feeding was infrequently observed on *Pedicularis* L., likely due to this entire order being represented by only a single molecular reference species.
-
-```{r compare results, out.width="50%"}
+but where feeding was infrequently observed on *Pedicularis* L., likely due to the Lamiales order being represented by only a single molecular reference species.
+
+```{r compare results, out.width="50%", fig.cap= 'Comparision of Accuracy between the initial output data from BLAST, and these same data subjected to the post-classification process which removes surrogate, and temporally restricted species'}
 
@@
 We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.
-  Difficulties in temporal models related to variability in drivers of flowering phenology. AND SMALL SAMPLE SIZE
-
-These results show that the overall results between **Bumble Bee ecology** observational and barcoding are largely congruent.
-But that ...
-We analyzed pollen loads from all of the most common bumble bee species in the area (@pyke1982local)
-Future analyses of the long term data set...
+  Difficulties in temporal models related to variability in drivers of flowering phenology, and relatively small appropriate sample sizes per species.
+
+Based on our results, we believe the overall results between observations of queen Bumble Bee and barcoding are largely congruent.
+Which relaxes concerns regarding differences between the insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+This indicates that results from fine resolution survey's, may serve as calibration standards to projects which expand geographically outward from them, and seek to make inference to larger areas.
+We analyzed pollen loads from all of the most common bumble bee species in the area and were able to ...
+(@pyke1982local).
+The species which appeared most often in loads via the molecular records, yet seldom noted via observations, mostly had affinities to wetlands or stream banks.
+These habitats are not often sampled by terrestrial ecologists due to logistics and concerns regarding degrading them.
+This indicates that metabarcoding may be applied in studies where areas, or administrative units, prohibit the disturbances associated sampling.
 
@@
   knitr::kable(caption = 'Limitations Faced and Possible Solutions',
-             col.names = c('Method Component', 'Limitations', 'Paths Forward')) 
+             col.names = c('Method Component', 'Limitations', 'Paths Forward')) |>
+  kable_styling(full_width = TRUE)
 ```
@@
 We expect that within a bayesian framework studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
-the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems (.
-We hope that the promise of A353 probes as tools for metabarcoding play a role in these endeavors.
+the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems.
+We hope that the A353 probes as tools for metabarcoding play a role in these endeavors.
 
@@
 
-```{r Overview of entire process, out.height="125%"}
+```{r Overview of entire process, out.height="90%"}
 knitr::include_graphics('../graphics/plots/flowchart.png')

### 2023-03-05 script/Filter.R

@@ -1,0 +1,87 @@
+library(tidyverse)
+library(sf)
+library(terra)
+library(RColorBrewer)
+library(ggtext)
+library(ggnewscale)
+
+setwd('~/Documents/ManuSCript/script')
+
+filt <- st_polygon( # complete
+    list(
+      rbind(
+        c(43, 1), c(57, 15), # bottom portion right
+        c(57, 60), # stem portion right
+        c(85, 100), c(15, 100), # very top of filter
+        c(43, 60), # stem portion left
+        c(43, 1)  # bottom portion left
+      )
+    )
+  )  %>% 
+  st_geometry()  %>% 
+  st_as_sf(crs = 26913) %>% 
+  mutate(Attribute = 'Filter') %>% 
+  rename('geometry' = x) %>% 
+  st_as_sf()  
+  
+filt_v <-   vect(filt)
+r <- rast(filt_v, nrow = 1000, ncol = 1000)
+r <- setValues(r, 1:1000000)
+r <- crop(r, filt_v, mask = T)
+r.df <- as.data.frame(r, xy = T)
+rm(r, filt_v)
+
+
+filt_text <- data.frame(
+  'Value' = c(
+    'Angiosperms 353: 14,000+', 'Floristic Unit: 3226', 'Insect Pollinated: 2721', 
+    'Spatial Search: 501', 'SDM: 426', 'Survey: 117' ),
+  'Approach' = c(rep('Coarse', times = 3), 
+                 rep('Computational', times = 2), 
+                 rep('Survey', times = 1)),
+  y = c(96, 90, 84, 78, 72, 66),
+  x = c(50, 50, 50, 50, 50, 50)
+) 
+
+time <- data.frame(
+  'Value' = 'temporal gradient',
+  'Approach' = 'Computational/Survey',
+  x = 50, 
+  y = 35
+)
+
+
+myco <- setNames(c("#A50F15", "#CB181D",  "#EF3B2C", "#FB6A4A"),
+         c("Computational/Survey", "Survey", "Computational",'Coarse') )
+
+base <- ggplot() +
+  geom_raster(data = r.df, aes(x = x, y = y, fill = lyr.1)) + 
+  scale_fill_distiller(palette = 'YlGn', direction = -1) +
+  guides(fill = 'none') +
+  theme_void() 
+
+base +
+  ggnewscale::new_scale_fill() +
+  geom_label(data = filt_text, lwd = 4, 
+             aes(x = x, y = y, label = Value, fill = Approach)) +
[…]

### 2023-03-05 script/flowchart.R

@@ -2,94 +2,100 @@ library(DiagrammeR)
 
 setwd('/home/sagesteppe/Documents/ManuSCript/script')
 
-g <- grViz("
+grViz("
 
 digraph boxes_and_circle {
   
-  node [shape = rectangle, fontsize = 18]
+  node [shape = rectangle, fontsize = 36]
   
       # fieldwork operations
       subgraph cluster_fieldwork {
-        style=dashed; color= '#ffa600'; pendwidth = 3
+        style=dashed; color= '#ffa600'; 
         node [fontname = arial, style = filled, fontcolor = White, color = '#ffa600']
         
-          'field work' -> 'plant survey'
-          'plant survey' -> 'herbarium samples' [label = 'vouchers', fontsize = 18]
-          'field work' -> 'pollinator obs.'
-          'pollinator obs.' -> 'pollen samples' [label = 'capture insects', fontsize = 18]
+          'field work' -> 'plant survey' [penwidth = 6]
+          'plant survey' -> 'herbarium samples' [label = '  vouchers',
+              fontsize = 36, penwidth = 6]
+          'field work' -> 'pollinator obs.' [penwidth = 6]
+          'pollinator obs.' -> 'pollen samples' [label = '  capture insects', 
+             fontsize = 36, penwidth = 6]
       }
         
       # Main spatial operations
       subgraph cluster_spatial {
-      style=dashed; color= '#ff764a'; pendwidth = 3
+      style=dashed; color= '#ff764a'; 
       node [fontname = arial, style = filled, fontcolor = White, color = '#ff764a']
       
-      'distance query' -> 'SDM' [label = 'log. regression',  fontsize = 18]
+      'distance query' -> 'Species Distribution Models' [label = '  log. regression', 
+          fontsize = 36, penwidth = 6]
       }
       
       subgraph cluster3 {
-      style=dashed; color= '#ef5675'; pendwidth = 3
+      style=dashed; color= '#ef5675';
       node [fontname = arial,  style = filled, fontcolor = White, color = '#ef5675']
 
-        'herbarium samples' -> 'DNA extraction' 
-        'DNA extraction' -> 'hyb-seq'  [label = 'CTAB',  fontsize = 18]
-        'hyb-seq' -> 'mi-Seq' [label = 'A353', fontsize = 18]
[…]

### 2023-03-05 script/conceptual-tables-boxes.rmd
text

@@
   pdf_document: default
-  html_document: default
 ---
 
-Here we create three text boxes for the adjacent manuscript.
-These serve to illustrate three main ideas which we found difficult to incorporate into the written text.
-We also produce one diagram, which serves to outline the entirety of the process we engaged in.
+Here we create three text boxes for the adjacent manuscript. These serve to illustrate three main ideas which we found difficult to incorporate into the written text.
 
@@
     'Presence of species by hydrologic Basin',
-    'Presence of Rare Aquatic Species',
+    'Presence of Rare Aquatic Species; Others',
     'Identifying the provenance of materials found at crime scene',
@@
   'Citation' = c(
-    'Soininen et al. 2009',
-    'Coghlan et al. 2021',
-    'Tsukamoto et al. 2021',
+    'Soininen et al. 2009; reviewed in Banerjee et al. 2022',
+    'Coghlan et al. 2021; Xu et al. 2018',
+    'Tsukamoto et al. 2021, reviewed in Banerjee et al. 2022',
     'Allwood et al. 2020',
@@
   'Citation' = c(
-    c('Coissoc et al. 2016, Kress 2017'),  #taxonomic resolution
-    c('Kress 2017, Bell et al. 2022'),  # reference library phylogenetic
+    c('Coissoc et al. 2016, Kress 2017, Johnson et al. 2023'),  #taxonomic resolution
+    c('Kress 2017, Bell et al. 2022, Johnson et al. 2023'),  # reference library phylogenetic
     c('Cheng et al. 2018, Darwin Tree of Life 2022, Lewin et al. 2020, Bell et al. 2021'),  # 'Reference Library spatial
@@
 ```
-
-4) Diagram of Major conceptual 'filtering steps'

### 2023-03-05 script/manuscript_MEE.Rmd
text

@@
 --- 
-title: Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from Queen Bumble Bees
+title: Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees
 author:  |
@@
   1) DNA Barcoding has been successful for the rapid analysis of complex ecological assemblages.
-Barcoding in the plant kingdom barcoding has been more difficult than in others, and hence may begin to lag behind other kingdoms.
-  2) Here we test the utilization of Angiosperms 353 probes to barcode plant species found in corbiculae pollen loads collected from Queen Bumble Bees. 
-  3) To verify the accuracy of this barcoding system we compared the barcoding data to museum plant specimens, observations of plant flowering and bee flower visitation in the field, and species distribution modelling to identify likely candidate plant species found within the corbiculae pollen loads.
-  4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which may use increase the alignment algorithms minimizing need for large computational power, and run time.
-  5) We show that the Angiosperms 353 probes, developed for phylogenmoics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches.
+However, barcoding in the plant kingdom has been more difficult than in others limiting the promises of it's use in eDNA applications.
+  2) Here we test the use of Angiosperms 353 probes to barcode plant species found in corbiculae pollen loads collected from wild foraging bumble bees. 
+  3) Using a high resolution long term observation study of wild bumble bees, we develop a framework for the application of metagenomics in plant-pollinator interactions to increase the efficiency and accuracy of applications using species distribution models, phenological modelling, and simple procedural decision making.
+  4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which are poised to increase accuracy of the sequence assignments and  which minimize the need for large computational power, and run time.
+  5) We show that the Angiosperms 353 probes, developed for phylogenomics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches around the globe.
   6) The DNA barcoding of bumble bee corbiculae pollen loads was most accurate when combined with knowledge of what plant species were flowering in the plant community when they were collected.
@@
 
-The inability to reliably identify plants to terminal taxon can limit our understanding of ecosystem function and interactions (@bortolus2008error).
-This is especially true for genera where identification down to species can be useful as specific bioindicators; defining ecological and behavioral properties  (e.g.
-different species of Sagebrush- *Artemisia* L., Willows - *Salix* L., and Sedges - *Carex* L.) (@Gage2013HistoricalRO).
-In these instances the lack of species level data can hinder our understanding of the breadth of habitat which some species occupy, and their interactions with other species.
-This can be further complicated by the fact that the identification of organisms to species is also often mired by lack of diagnostic characters(e.g.
-flowers, fruits, roots or combinations thereof), an increasing lack of taxonomic experts (@hebert2003biological) and the presence of cryptic species (@janzen2017nuclear, @oliver2009cryptic).
-Taxonomic verification can also be limited by the fact that revisiting field sites to identify material using morphological or chemical approaches, can be resource intensive and often ineffective.
-The current methods to ameliorate this situation include: (i) ignoring these ecologically relevant levels of detail, (ii) revisiting plots as diagnostic material becomes temporally available, (iii) seeking the assistance from taxonomic specialists, or (iv) the use of barcoding using molecular techniques (@hebert2003biological).
+The plurality of Earths biodiversity may be attributed to the interactions between species (CITE).
+These interactions not only lead to the the origin of species but are essential to the maintenance of virtually all ecosystems (CITE).
+In order to understand and rationally conserve both the species and the ecosystems which their interactions compose integrative approaches, with the potential for scaling are required immediately.
+Major limitations, imposed by a lack of taxonomic expertise, impede our initial ability to identify organisms let alone more complex phenomena (@hebert2003biological).
+The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a bio-monitoring role they are pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic).
+The ability to identify fragments of organisms (e.g.
+leaf tissue) increases our ability to understand the interactions of a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
+those for restoration processes.
 
   Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
-With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas with many other clades results have been elusive (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas many other clades have proven more problematic (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
 Particular challenges with the high copy number barcodes (e.g.
@@
   Currently the largest plant systematic endeavor ever undertaken,by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom,14,000 represented species, using the popular Angiosperms353 (A353) probes, which includes 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal).
+This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000 represented species, using the popular Angiosperms353 (A353) probes, which includes 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal).
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
 The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of plants, will contribute many more species by 2030 (@cheng2018tenkp).
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute many more species by 2030 (@cheng2018tenkp).
 Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high numbers of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@darwin2022project, @lewin2022biogenome).
@@
 ```{r Co-occurrence of splash drop sequences, eval = F}
+library(CooccurrenceAffinity)
 
@@
 
-myout <- CooccurrenceAffinity::affinity(data = e, row.or.col = "col", squarematrix = c("all"))
-CooccurrenceAffinity::plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
+myout <- affinity(data = e, row.or.col = "col", squarematrix = c("all"))
+plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
 ```

### 2023-03-05 script/Filter.R

@@ -7,7 +7,12 @@ library(ggnewscale)
 
 setwd('~/Documents/ManuSCript/script')
 
-filt <- st_polygon( # complete
+# first we will create the shape of the filter itself. We will turn into an 
+# explicitly spatial object using the 'sf' package. We will do this so that we
+# are able to color the inside of the Filter shape with a gradient of colors, 
+# which is defined within a geom_raster, created by the Terra package. 
+
+filt <- st_polygon( 
     list(
       rbind(
         c(43, 1), c(57, 15), # bottom portion right
@@ -24,6 +29,10 @@ filt <- st_polygon( # complete
   rename('geometry' = x) %>% 
   st_as_sf()  
   
+# we create an empty raster the size and extent of the filter, and mask it
+# to the shape to keep only pixels within the lines in; we are drawing between 
+# the lines! FINALLY after all these years!!
+
 filt_v <-   vect(filt)
 r <- rast(filt_v, nrow = 1000, ncol = 1000)
 r <- setValues(r, 1:1000000)
@@ -31,6 +40,8 @@ r <- crop(r, filt_v, mask = T)
 r.df <- as.data.frame(r, xy = T)
 rm(r, filt_v)
 
+## we create labels and place these at positions to represent how many species 
+# exist at each stage of the filter 'head' (...the top part!!!) 
 
 filt_text <- data.frame(
   'Value' = c(
@@ -43,6 +54,7 @@ filt_text <- data.frame(
   x = c(50, 50, 50, 50, 50, 50)
 ) 
 
+# we create this label seperately so that we can readily rotate it 90 degrees. 
 time <- data.frame(
   'Value' = 'temporal gradient',
   'Approach' = 'Computational/Survey',
@@ -50,16 +62,19 @@ time <- data.frame(
   y = 35
 )
 
-
+# define our colors, they run counter to the background colors... 
 myco <- setNames(c("#A50F15", "#CB181D",  "#EF3B2C", "#FB6A4A"),
          c("Computational/Survey", "Survey", "Computational",'Coarse') )
 
+# created a base plot due to the second scale label being weird, probably 
+# unnecessary after some code re-factoring, but too scared to investigate. 
 base <- ggplot() +
   geom_raster(data = r.df, aes(x = x, y = y, fill = lyr.1)) + 
   scale_fill_distiller(palette = 'YlGn', direction = -1) +
   guides(fill = 'none') +
[…]

### 2023-03-05 script/flowchart.R

@@ -1,5 +1,4 @@
 library(DiagrammeR)
-
 setwd('/home/sagesteppe/Documents/ManuSCript/script')
 
 grViz("

### 2023-03-06 script/manuscript_MEE.Rmd
citations for new first paragraph, and incorporated some more paul edits

@@
 ```{r Required Libraries, message=F, warning=F, results='hide'}
-library(here)
 library(tidyverse)
@@
 
-The plurality of Earths biodiversity may be attributed to the interactions between species (CITE).
-These interactions not only lead to the the origin of species but are essential to the maintenance of virtually all ecosystems (CITE).
-In order to understand and rationally conserve both the species and the ecosystems which their interactions compose integrative approaches, with the potential for scaling are required immediately.
+An enormous amount of Earths biodiversity may be attributed to the interactions between species (@soltis2019darwin, @futuyma2009macroevolution, @voje2015role, @weber2017evolution, @hembry2020ecological).
+These interactions not only leading to the the origin of many species but are essential to the maintenance of virtually all ecosystems (CITE).
+In order to understand and rationally conserve both the species and the ecosystems which their interactions compose - Darwins 'Entangled Bank' integrative approaches, with the potential for scaling are required immediately (@darwin2004origin, @thompson1994coevolutionary, @banerjee2022environmental).
 Major limitations, imposed by a lack of taxonomic expertise, impede our initial ability to identify organisms let alone more complex phenomena (@hebert2003biological).
-The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a bio-monitoring role they are pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic).
+The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a role they are generally pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic).
 The ability to identify fragments of organisms (e.g.
 leaf tissue) increases our ability to understand the interactions of a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
-those for restoration processes.
+those for restoration processes (@banerjee2022environmental, @johnson2023environmental).
 
@@
 
-  As species interactions vary both in space and time (@thompson1994coevolutionary, @caradonna2021seeing).
-Contrasts in the flowering periods of many plant species, can provide an additional filter for increasing the accuracy of identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC).
-In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks, which reduces the utility of these to operate as filters in the post-processing of sequence matches.
+  As species interactions vary both in space and time contrasts in the flowering periods of many plant species, can provide an additional filter for increasing the accuracy of identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC, @thompson1994coevolutionary, @caradonna2021seeing).
+In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction, @caradonna2020temporal), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks, which reduces the utility of these to operate as filters in the post-processing of sequence matches.
 Nonetheless, we work develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new, Figure 1).
@@
 To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (Ogilvie unpublished), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
 4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
@@
 
-### 2.3.2 | Pollen Corbiculae Loads
+### 2.3.2 | Preperation of Pollen Corbiculae Loads
 
   To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Cerca* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
+*Circa* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
 The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
@@
 Using five years (2015-2020) of observational data on *Bombus* queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years.
-We sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
+In order to capture more variability inherit in the 353 loci we sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
 In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
-We also sequenced another 15 taxa commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (Appendix 3).
+We also sequenced another 15 taxa of plants commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (Appendix 3).
 Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).

### 2023-03-07 script/manuscript_MEE.Rmd
added citations to first paragraph

@@
 An enormous amount of Earths biodiversity may be attributed to the interactions between species (@soltis2019darwin, @futuyma2009macroevolution, @voje2015role, @weber2017evolution, @hembry2020ecological).
-These interactions not only leading to the the origin of many species but are essential to the maintenance of virtually all ecosystems (CITE).
-In order to understand and rationally conserve both the species and the ecosystems which their interactions compose - Darwins 'Entangled Bank' integrative approaches, with the potential for scaling are required immediately (@darwin2004origin, @thompson1994coevolutionary, @banerjee2022environmental).
+These interactions not only lead to the the origin of many species but appear essential to the maintenance of virtually all ecosystems (@agrawal2007filling, @valiente2015beyond, @bascompte2006asymmetric).
+In order to understand and rationally conserve both the species and the ecosystems which their interactions compose - Darwins 'Entangled Bank'- integrative approaches, with the potential for scaling are required immediately (@darwin2004origin, @thompson1994coevolutionary, @agrawal2007filling, @banerjee2022environmental, @blanchet2020co, @jordano2016sampling).
 Major limitations, imposed by a lack of taxonomic expertise, impede our initial ability to identify organisms let alone more complex phenomena (@hebert2003biological).
-The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a role they are generally pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic).
+The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a role they are typically pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic).
 The ability to identify fragments of organisms (e.g.
-leaf tissue) increases our ability to understand the interactions of a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
+leaf tissue) increases our ability to understand the interactions of  not only entire ecosystems but also a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
 those for restoration processes (@banerjee2022environmental, @johnson2023environmental).
@@
 These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
-The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring).
 Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute many more species by 2030 (@cheng2018tenkp).

### 2023-03-07 script/manuscript_MEE.Rmd
added citations for academic conclusions

@@
 
-  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.
+  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), within a framework which utilizes custom species sequence databases derived via species distribution modelling, and temporal filtering.
 To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
@@
 We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.
-  Difficulties in temporal models related to variability in drivers of flowering phenology, and relatively small appropriate sample sizes per species.
-
-Based on our results, we believe the overall results between observations of queen Bumble Bee and barcoding are largely congruent.
-Which relaxes concerns regarding differences between the insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-This indicates that results from fine resolution survey's, may serve as calibration standards to projects which expand geographically outward from them, and seek to make inference to larger areas.
-We analyzed pollen loads from all of the most common bumble bee species in the area and were able to ...
-(@pyke1982local).
-The species which appeared most often in loads via the molecular records, yet seldom noted via observations, mostly had affinities to wetlands or stream banks.
-These habitats are not often sampled by terrestrial ecologists due to logistics and concerns regarding degrading them.
-This indicates that metabarcoding may be applied in studies where areas, or administrative units, prohibit the disturbances associated sampling.
+
+Our results indicate the overall information gleaned from observations of queen Bumble Bee foraging and analysis of pollen records are largely congruent.
+Relaxing concerns regarding differences between the broad insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+In general when interaction networks are considered at coarse levels, such as the duration of a season, our perceptions regarding the generality of interactions at smaller time scales may be inflated relative to the actualized interactions within them, e.g.
+a week (@caradonna2020temporal).
+These results indicate a possibility that at even finer levels *Bombus* display high amounts of floral fidelity within foraging bouts, an observation which implies that part of the reason for the high efficiency of *Bombus* as a pollinator might partially be related to their lack of movement of hetero-specific pollen (@brosi2013single, @ashman2013toward, @galloni2008visitor, @brosi2016pollinator).
+The mechanisms behind this observed fidelity are likely related to pollen nutritional values, specifically high concentrations of protein, and the absence of particular amino acids required for larval development in other flower more commonly used by workers (@genissel2002influence, @tasei2008nutritive, @goulson2005causes, @goulson2008diet, @hanley2008breeding).
 
@@
 
-Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Mertensia* on a per visit basis...
-General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persist with *Bombus* studies...
-
-  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
+  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the famillies Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
 Numerous historic, and some ongoing, land management practices reduce the ability of many landscapes to support stable populations of *Bombus*.

### 2023-03-08 script/conceptual-tables-boxes.rmd
added a table of most visited plants

@@
     'Jaccard Index, Plot Abundance',
-    'Spike Samples with Reference Materials; Several Genome Sizes'
+    'Spike Samples with Reference Materials; Several C Sizes'
   )

### 2023-03-08 script/manuscript_MEE.Rmd
added a table of most visited plants

@@
 
-```{r Correlation between Number of Grains and Number of Sequences}
+```{r Correlation between Number of Grains and Number of Sequences, fig.cap = 'Relationship between morphological count data and sequence reads', out.height="50%", out.width="50%"}
 
@@
 
-# number of assignemnts
+# number of assignments
 
@@
 
+```{r Most Used Floral Resources Table}
+
+fam <- all_classifications %>% 
+  ungroup() %>% 
+  select(Family, Genus) %>% 
+  distinct()
+
+expert %>% 
+  group_by(Taxon) %>% 
+  add_count(name = 'No.Corb') %>% 
+  mutate(median_percent_corb = round( mean(Percent_Sample), 1), 
+         Taxon = str_remove(Taxon, '^.* '), 
+         Taxon = if_else(Taxon %in% c('Salix', 'Erigeron'), 'sp.', Taxon)) %>% 
+  select(Genus, Species = Taxon, Percent_Sample, No.Corb, median_percent_corb) %>% 
+  left_join(., fam, by = 'Genus') %>% 
+  select(-Percent_Sample) %>% 
+  relocate(Family, .before = Genus) %>% 
+  filter(No.Corb > 4) %>% 
+  distinct() %>% 
+  arrange(Family, Genus)  %>% 
+  knitr::kable(
+    escape = F,
+    align = c('l', 'l', 'l', 'c', 'c'),
+    caption = 'Plant species detected in five or more corbiculae loads',
+    col.names = c('Family', 'Genus', 'Species', 'No. Samples', 'Mean Prop.\\textsuperscript{a}')
+  ) %>% 
+  add_footnote(
+    'The mean only calculated across the samples where the species was detected') %>% 
+  collapse_rows()
+
+```
+
 ```{r}
@@
 In particular, the removal of locoweeds (Fabaceae: *Astragalus* L.
-& *Oxytropis* DC.) and larkspurs (Ranunculaceae: *Delphinium*) were common across public lands administered by the United States Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
+& *Oxytropis* DC.) and larkspurs (Ranunculaceae: *Delphinium*) were common across public lands administered by the U.S.
+Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
 Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks for significant periods of time, all processes which lower the water tables and reduced the extent of stream-associated [riverine] wetlands and the mesic meadows fringes which provide habitat for many species of tall *Mertensia* (Boraginaceae, e.g.

### 2023-03-21 script/conceptual-tables-boxes.rmd
added second round of PJC comment edits

@@
 ```
+
+4) Summary of errors from the framework approach
+
+```{r}
+
+errors <- data.frame(
+
+  'Component' = c(
+    'Spatial', 'Molecular (Reference Library)',
+    'Molecular Reference (Loci)', 'DNA difficult to extract',
+    'Reclassification (Genus)', 'Reclassification (Family)'
+  ),
+
+  'Example' = c(
+    '"Scabrethia Scabra" (false postiive)', '"Asteraceae" (wrong order)',
+    '"Trollium" (false positives, same family as target)', '"Agastache" (Hemiparasites)',
+    '"Epilobium"', '"Paxistima myrsinites"'
+  ),
+
+  'Correction'  = c(
+    'Wyethia arizonica', 'Frasera', 
+    'Lupinus sp.', 'Pedicularis',
+    'Chamaenerion latifolium?', 'Parnassia palustris'
+  )
+)
+
+write.csv(errors, '../data/Errors_Table.csv', row.names = F)
+
+read.csv('../data/Errors_Table.csv') |>
+  knitr::kable(caption = 'Errors faced with components of the framework',
+             col.names = c('Component', 'Example', 'Correction'))
+
+rm(errors)
+```
+
+For example many sequences which mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
+nevadensis* Cresson and *B.
+rufocinctus* Cresson pollen is most likely *Frasera* Walter, which failed extractions for the reference library failed (Appendix 11).
+A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* (A.
+Heller) Rydb.
+but where feeding was infrequently observed on *Pedicularis* L., likely due to the Lamiales order being represented by only a single molecular reference species.
+
+Situations where SDM's led to incorrect results at the species level are evident with classification to *Scabrethia Scabra* (Hooker) W.A.
+Weber, this match almost certainly representing *Wyethia arizonica* A.
+Gray (@weber1998new), a species known to be visited by queen bee's via our floral observations.
+An expected inaccuracy of the classification scheme is in genus level placements, e.g.
+were *Epilobium* L.
+(Onagraceae Juss.) spp.
+were classified.
+However, given the small size of their flowers in the study area, these results more likely indicate that a species of *Chamaenerion* Seg.
+(a segregate genus) such as *C.
+angustifolium* (L.) Scop.
+or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.
+An issue with reclassification within the family level in combination with time included reclassifying *Parnassia palustris* L.
+to *Paxistima myrsinites* (Pursh) Rafinesque.
+However, based on flower size it is more likely that the visited taxon was *P.
+palustris*.

### 2023-03-21 script/manuscript_MEE.Rmd
added second round of PJC comment edits

@@
 author:  |
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$, Emily J. Woodworth$^1$$^,$$^2$, 
     |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
@@
     |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$, Emily J. Woodworth$^1$$^,$$^2$, 
 abstract:  |  
   1) DNA Barcoding has been successful for the rapid analysis of complex ecological assemblages.
-However, barcoding in the plant kingdom has been more difficult than in others limiting the promises of it's use in eDNA applications.
+However, barcoding in the plant kingdom has been more difficult than in others, limiting the applications of eDNA and metagenomics in ecology.
   2) Here we test the use of Angiosperms 353 probes to barcode plant species found in corbiculae pollen loads collected from wild foraging bumble bees. 
-  3) Using a high resolution long term observation study of wild bumble bees, we develop a framework for the application of metagenomics in plant-pollinator interactions to increase the efficiency and accuracy of applications using species distribution models, phenological modelling, and simple procedural decision making.
-  4) By utilizing Species distribution modelling we allow users to create a regionally appropriate sequence databases which are poised to increase accuracy of the sequence assignments and  which minimize the need for large computational power, and run time.
+  3) Using a high resolution long term observation study of wild bumble bees, we develop a simple framework for the application of metagenomics in plant-pollinator interactions to increase the efficiency and accuracy of applications using species distribution models, phenological modelling, and simple procedural decision making.
+  4) By utilizing Species distribution modelling we allow users to create regionally appropriate sequence databases which are poised to increase accuracy of the sequence assignments and  which minimize the need for large computational power, and run time.
   5) We show that the Angiosperms 353 probes, developed for phylogenomics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches around the globe.
@@
 
-  Currently the largest plant systematic endeavor ever undertaken,by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000 represented species, using the popular Angiosperms353 (A353) probes, which includes 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal).
-These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding.
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute many more species by 2030 (@cheng2018tenkp).
-Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high numbers of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@darwin2022project, @lewin2022biogenome).
-These data will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
+  Currently the largest plant systematic endeavor ever undertaken, by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).
+This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000+ represented species, using the popular Angiosperms353 (A353) probes (@baker2021PAFTOL, @johnson2019universal).
+Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute more species by 2030 (@cheng2018tenkp).
+These publicly available data serve to provide a phylogenetically comprehensive backbone for plant metabarcoding.
+Similar projects, focused on whole genome sequencing of all organisms in certain geographic areas, such as the ‘Darwin Tree of Life’, which aims to sequence all described taxa in Britain and Ireland, will provide geographically dense data sets to seed projects globally (@darwin2022project, @lewin2022biogenome).
+These novel genomic data sets will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
 However, the application of metabarcoding still faces challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
@@
   Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), within a framework which utilizes custom species sequence databases derived via species distribution modelling, and temporal filtering.
-To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
+To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific sequence database relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
 To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
-iNaturalist), from a region exceeding the study area.
-For these candidate species, a modelling approach, such as logistic regression, may be used to identify taxa which warrant further exploration e.g.
-determine their possibility of presence in metabarcoding samples.
+iNaturalist).
 We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
-Species distribution models examine the ecological conditions associated with the known occurrence of a species to identify suitable habitats in the study area.
 This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers.
@@
 
-  We tested this metagenomic and informatics approach to determine to examine the foraging of queen bumblebee, and compare this approach with direct observations and the pollen record, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-The assessment of the plant species compositions from pollen is a desirable approach, with several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
+  We tested the proposed metagenomic and informatics approach to examine the diets of wild bumblebee pollinators, and compare this approach with direct observations and the pollen record, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
 The two foraging phases of the queen bumblebee life cycle are essential to 1) increase their weight before diapause, and 2) increase their ovary weights while establishing their recently found nests.
@@
 
-To develop a reference library of  pollen grains which may be present in corbiculae loads, an image reference collection of fuchsin-jelly stained (@beattie1971technique) slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@brosi2013single).
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+To qualitatively identify, and quantitatively note, the plant species present in corbiculae loads microscopy was used.
+A pollen reference library of fuchsin-jelly stained grains which may be present in corbiculae loads of slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@beattie1971technique, @brosi2013single).
+Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the RMBL Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
 4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
@@
 We first generated a short list of potential candidate species.
-We downloaded from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022) all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
-To reduce the list of species to include in the genomic sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.
+We downloaded, from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022), all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
+To reduce the number of species to include in the reference sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.
 
@@
 Modelling: Random Forest and Boosted Regression Trees, were sub sampled with 30% test and two replicates each before weighted ensemble based on True Skill Statistics (tss) (@sdmPackage).
-Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage, @).
+Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage).
 TSS was chosen as the ensemble criterion as it has been shown to work across a wide range of species occurrences prevalence (@allouche2006assessing).
@@
 
-Pollen genomic DNA was extracted from corbiculae using a CTAB based protocol modified from Lahlamgiahi et al.
+To extract genomic DNA from pollen a CTAB based protocol was modified from Lahlamgiahi et al.
 and Guertler et al.
 (2014, 2014).
-A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added followed by vortexing to allow dissolution of corbiculae.
+A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added to the sampled followed by vortexing to allow dissolution of corbiculae.
 Pollen grains were then macerated with Kontes Pellet Pestles, and the tip of these washed with 130 µL of the SDS extraction buffer, samples were then incubated for 1 hour at 30°C.
@@
 
-Library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
+Sequence library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
 Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
@@
 
-The six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
+In 2015 the six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
 A total of `r toString(sum(bee_obs$interactions))` overwinter queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
@@
 
-In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination syndromes, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
 The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
@@
 
-At the six study sites, of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
 Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
@@
 
-## 3.6 | Integrated Observational, Molecular, and Palynological Network
-
-While the spatial results were used to declare the taxonomic composition of the sequence database, temporal results were used in consideration with plant phylogeny to retroactively, reassign the assignment of sequences to taxa.
-Essentially, if a sequence was identified to a taxon which was not known from the field site
-
-For example many sequences which mapped to the Asteraceae family, but which was flagged by temporal filters and is present in both *B.
-nevadensis* Cresson and *B.
-rufocinctus* Cresson pollen is most likely *Frasera* Walter, which failed extractions for the reference library failed (Appendix 11).
-A similar likely mismatch could be between what was fide molecular evidence as *Agastache pallidiflora* (A.
-Heller) Rydb.
-but where feeding was infrequently observed on *Pedicularis* L., likely due to the Lamiales order being represented by only a single molecular reference species.
+## 3.6 | Integrated Observational, Molecular, and Palynological Results
 
@@
 ```
-
-Situations where SDM's led to incorrect results at the species level are evident with classification to *Scabrethia Scabra* (Hooker) W.A.
-Weber, this match almost certainly representing *Wyethia arizonica* A.
-Gray (@weber1998new), a taxon known to be visited by queen bee's via our floral observations.
-An expected inaccuracy of the classification scheme is in genus level placements, e.g.
-were *Epilobium* L.
-(Onagraceae Juss.) spp.
-were classified.
-However, given the small size of their flowers in the study area, these results more likely indicate that a species of *Chamaenerion* Seg.
-(a segregate genus) such as *C.
-angustifolium* (L.) Scop.
-or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.
-An issue with reclassification within the family level in combination with time included reclassifying *Parnassia palustris* L.
-to *Paxistima myrsinites* (Pursh) Rafinesque.
-However, based on flower size it is more likely that the visited taxon was *P.
-palustris*.
-
-Regarding limitations of morphological data we suspect that there were two morphotypes of pollen identified as Ericaceae Juss. were actually Onagraceae (Samples 19 & 44), based on molecular results.
-
-It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes.
-For example, no more than 10 bee corbiculae loads per species were sequenced with the Mdn = `r median(n_corbiculae)` ...
-, and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top interactions.
-
-Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed re-classifications of the contents of pollen grains...
 
@@
 
-Our results indicate the overall information gleaned from observations of queen Bumble Bee foraging and analysis of pollen records are largely congruent.
+In regards to the case study, our results indicate the overall information gleaned from observations of queen Bumble Bee foraging and analysis of pollen records are largely congruent.
 Relaxing concerns regarding differences between the broad insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
@@
 
-  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the famillies Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
-Numerous historic, and some ongoing, land management practices reduce the ability of many landscapes to support stable populations of *Bombus*.
-Historic livestock grazing was often associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle.
+Also regarding the case study, some foraging preferences of *Bombus*, at this field site and across several localities globally emerge.
+These suggest the need for land managers to maintain relatively high amounts of members of the plant families Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
+Numerous historic, and some ongoing, practices reduce the ability of many landscapes to support stable populations of *Bombus*.
+Historic livestock grazing was often associated with the targeted removal of many species known to have compounds toxic to cattle.
 In particular, the removal of locoweeds (Fabaceae: *Astragalus* L.
@@
 Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
-Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks for significant periods of time, all processes which lower the water tables and reduced the extent of stream-associated [riverine] wetlands and the mesic meadows fringes which provide habitat for many species of tall *Mertensia* (Boraginaceae, e.g.
+Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks, processes which lowered water tables and reduced the extent of stream-associated wetlands and the mesic meadows fringes which provide habitat for many species of 'tall' *Mertensia* (Boraginaceae, e.g.
 *M.
@@
 G.
-Don.) widely distributed across Western North America, and to an extent *Delphinium barbeyi*  and many species of native *Trifolium* L.
+Don.), to an extent *Delphinium barbeyi*  and many species of native *Trifolium* L.
 (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).
-Fire suppression further resulted in the succession of many Aspen (*Populus tremuloides* Michx.) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effects habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
-Finally the effects of Nitrogen deposition, especially given the West's rapidly growing population still pose adverse effects on the abundance of a variety of species of Fabaceae at Urban-Rural interfaces (see @stevens2018atmospheric, @fenn2003ecological).
-Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing.
+Fire suppression resulted in the succession of many Aspen (*Populus tremuloides* Michx.) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effecting habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
+Finally the effects of Nitrogen deposition, given the West's rapidly growing population still pose adverse effects on the abundance of species of Fabaceae at urban-rural interfaces (see @stevens2018atmospheric, @fenn2003ecological).
+Current solutions to the above issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing.
 The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L., using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
+
+```{r}
+read.csv('../data/Errors_Table.csv') |>
+  knitr::kable(caption = 'Errors faced with components of the framework',
+             col.names = c('Component', 'Example', 'Correction')) 
+```
 
@@
 **AUTHOR CONTRIBUTIONS:**  R.C.B conducted botanical collections, conducted all molecular lab work, lead all analyses, and writing.
-J.E.O conceived, designed, and conducted all ecological fieldwork, assisted with analyses, and writing.
+J.E.O conceived, designed, and conducted ecological fieldwork, assisted with analyses, and writing.
+P.J.C conducted ecological fieldwork, assisted with ecological analyses, and writing.
 E.J.W.
@@
 assisted with spatial analyses and writing.
-P.J.C assisted with ecological analyses and writing.
 J.B.F.
-conceived, and designed all lab work, analyses, and integration of approaches, assisted with writing, and secured funding for molecular work.
+conceived, and designed lab work, analyses, and integration of approaches, assisted with writing, and secured funding for molecular work.
 

### 2023-05-23 script/manuscript_MEE.Rmd
mit
adding JBF ST intro edits

@@
 abstract:  |  
-  1) DNA Barcoding has been successful for the rapid analysis of complex ecological assemblages.
-However, barcoding in the plant kingdom has been more difficult than in others, limiting the applications of eDNA and metagenomics in ecology.
-  2) Here we test the use of Angiosperms 353 probes to barcode plant species found in corbiculae pollen loads collected from wild foraging bumble bees. 
-  3) Using a high resolution long term observation study of wild bumble bees, we develop a simple framework for the application of metagenomics in plant-pollinator interactions to increase the efficiency and accuracy of applications using species distribution models, phenological modelling, and simple procedural decision making.
-  4) By utilizing Species distribution modelling we allow users to create regionally appropriate sequence databases which are poised to increase accuracy of the sequence assignments and  which minimize the need for large computational power, and run time.
-  5) We show that the Angiosperms 353 probes, developed for phylogenomics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches around the globe.
-  6) The DNA barcoding of bumble bee corbiculae pollen loads was most accurate when combined with knowledge of what plant species were flowering in the plant community when they were collected.
-Thus, supplementing DNA barcoding data with ecological context is most accurate and powerful.
+  1) DNA Barcoding has been successful for the rapid identification of species in complex ecological assemblages.
+However, metabarcoding in the plant kingdom has been limited due to a lack of universal gene regions that works across all taxa, limiting the applications of eDNA and metagenomics in ecology.
+  2) To circumvent these limitations, we propose using more holistic approach which combines a multiple gene approach that incorporates existing plant occurrence databases, species distribution models, and phenology analyses to generate a list of likely candidate species to increase accuracy.
+Such an effort can help guide plant sample collection and library construction, particularly for projects with limited resources or local taxonomic knowledge.
+In addition, we demonstrate that building a custom candidate plant list based on known occurrences, ecological requirements, and phenology can improve computing efficiency, speed, and accuracy of metabarcoding.
+We tested this approach to identify the plant species contributing to the corbiculae pollen loads of bumble bees.
+To validate the efficiency of this approach we compare results of the DNA barcoding to expert-led pollen identification and field observations of plant-pollinator interactions that generated these pollen loads.
+  3) We show that the Angiosperms 353 probes, developed for phylogenomics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches around the globe.
+The DNA barcoding of bumble bee corbiculae pollen loads was most accurate when combined with knowledge of what plant species were flowering in the plant community when they were collected.
+  4) Supplementing DNA barcoding data with ecological context is most accurate and powerful.  
 keywords: |
@@
 
-An enormous amount of Earths biodiversity may be attributed to the interactions between species (@soltis2019darwin, @futuyma2009macroevolution, @voje2015role, @weber2017evolution, @hembry2020ecological).
-These interactions not only lead to the the origin of many species but appear essential to the maintenance of virtually all ecosystems (@agrawal2007filling, @valiente2015beyond, @bascompte2006asymmetric).
-In order to understand and rationally conserve both the species and the ecosystems which their interactions compose - Darwins 'Entangled Bank'- integrative approaches, with the potential for scaling are required immediately (@darwin2004origin, @thompson1994coevolutionary, @agrawal2007filling, @banerjee2022environmental, @blanchet2020co, @jordano2016sampling).
-Major limitations, imposed by a lack of taxonomic expertise, impede our initial ability to identify organisms let alone more complex phenomena (@hebert2003biological).
-The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a role they are typically pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic).
-The ability to identify fragments of organisms (e.g.
-leaf tissue) increases our ability to understand the interactions of  not only entire ecosystems but also a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
-those for restoration processes (@banerjee2022environmental, @johnson2023environmental).
+Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogeneization (REF), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity).
+Monitoring ecological communities structure and interaction will be critical to informing and prioritizing conservation efforts (REF), particularly considering the large proportion of know species currently threaten or likely to be threaten in the near future (@pimm2014biodiversity; @pereira2010scenarios).
+However, ecosystem monitoring remains inconsistent in time and space (@yoccoz2001monitoring) as it requires substantial resources, thereby leaving several regions, ecosystems, and even species under-observed (@pereira2012global; @collen2008tropical; @meyer2015global; @ruete2015displaying).
+
+Species interactions are a critical component of ecosystem stability and therefore biodiversity conservation (Soltis et al.
+(2019), @futuyma2009macroevolution, Voje et al.
+(2015), Weber et al.
+(2017), @hembry2020ecological).
+Species interactions, which span from positive (e.g., symbiosis, mutualistic) to negative (e.g., predation, competition), shape species distribution (@bascompte2009mutualistic; Wisz et al.
+2013) and tolerance to environmental stressors (REF).
+They impact ecosystem stability and diversity (@agrawal2007filling, @valiente2015beyond, @bascompte2006asymmetric, and the relationships between biodiversity and ecosystem functions (@duffy2007functional).
+However, species interactions are challenging to observe (REFs).
+They are typically measured through shorter-term observational studies yet are highly variable in time and space (@liu2022environmental).
+Studying species interactions also requires extensive taxonomic expertise, especially those from diverse clades (@hebert2003biological), further challenging their frequent assessment.
+As a result, few ecological networks have been fully mapped (REFs).
+Improving our capacity to efficiently study species interactions consistently across several ecosystems, landscapes, and management context is thus pivotal to enacting conservation interventions (@tylianakis2010conservation).
+This is particularly urgent as increased biological introductions and climate change are impacting species interactions at a rapid rate, thus calling for more frequent and consistent monitoring (REFs).
 
   Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
-With plants the identification of members of certain clades using barcoding has been quite successful (@kress2017plant), whereas many other clades have proven more problematic (@liu2014identification, @china2011comparative, @coissac2012bioinformatic), however metabarcoding incurs additional challenges to those which  exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
-Particular challenges with the high copy number barcodes (e.g.
+The ability to identify species from a fragments of organisms (e.g.
+hair, scat, leaf tissue) increases our ability to understand the interactions of not only entire ecosystems but also a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
+those for restoration processes (Banerjee et al.
+(2022), Johnson et al.
+(2023)).
+With plants the identification of members of certain clades to species using barcoding has been successful in some groups (@kress2017plant), whereas many other clades have proven more problematic (@liu2014identification, @china2011comparative, @coissac2012bioinformatic).
+Metabarcoding incurs additional challenges to those which exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
+Particular limitations of the traditional high copy number barcodes (e.g.
 ITS2, *rbcL*, *matK*, *trnH-psbA*) include the utilization their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
@@
 
-  As species interactions vary both in space and time contrasts in the flowering periods of many plant species, can provide an additional filter for increasing the accuracy of identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC, @thompson1994coevolutionary, @caradonna2021seeing).
-In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction, @caradonna2020temporal), however the overall shorter extent of the active growing season in these systems results in the presence of few to any natural breaks, which reduces the utility of these to operate as filters in the post-processing of sequence matches.
-Nonetheless, we work develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new, Figure 1).
+  Plant flowering varies over space and also in time, and in particular, contrasts in the flowering periods of many plant species can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC, @thompson1994coevolutionary, @caradonna2021seeing).
+In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction, @caradonna2020temporal).
+We work to develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new, Figure 1).
 
@@
 
-  We tested the proposed metagenomic and informatics approach to examine the diets of wild bumblebee pollinators, and compare this approach with direct observations and the pollen record, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+  We tested this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study.
+In particular, we use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
 The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
 The two foraging phases of the queen bumblebee life cycle are essential to 1) increase their weight before diapause, and 2) increase their ovary weights while establishing their recently found nests.
-Both of these time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
+Both time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
 Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline).
@@
 4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  We used Divisive Hierarchical Clustering techniques to determine which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
+  We used clustering techniques to supplement our subjective opinions of which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
 Ten readily discernible categorical traits were collected from each specimen in the image collection.
@@
 In 2015 the six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
-A total of `r toString(sum(bee_obs$interactions))` overwinter queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
+A total of `r toString(sum(bee_obs$interactions))` overwintered queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
 Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week minimum = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, median = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (minimum = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, median = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
@@
 
-## 3.1 | Spatial Analyses
+## 3.2 | Spatial Analyses
 
@@
 The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less Heliantheae.
-Boraginaceae grains `r toString(aa)`.
+Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less the Heliantheae Aliiance (Sunflower Tribe) (Boraginaceae grains `r toString(aa)`.
 Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
@@
 
+A number of suspected false positives were returned by the sequence classified, in all instances these were species present in low abundances at the field station which were also returned with a common species in the same family.
 To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala Jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
@@
 species were present.
-The size of flower of *Nemophila breviflora* A.
+The flower size  and sparsity of *Nemophila breviflora* A.
 Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
@@
 These suggest the need for land managers to maintain relatively high amounts of members of the plant families Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
-Numerous historic, and some ongoing, practices reduce the ability of many landscapes to support stable populations of *Bombus*.
-Historic livestock grazing was often associated with the targeted removal of many species known to have compounds toxic to cattle.
-In particular, the removal of locoweeds (Fabaceae: *Astragalus* L.
-& *Oxytropis* DC.) and larkspurs (Ranunculaceae: *Delphinium*) were common across public lands administered by the U.S.
+Numerous historic  practices reduce the ability of many landscapes to support stable populations of *Bombus*.
+Historic livestock grazing involved the removal of many species known to have compounds toxic to cattle.
+In particular, the removal of locoweeds (*Astragalus* & *Oxytropis*) and larkspurs (*Delphinium*) were common across public lands administered by the U.S.
 Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
-Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks, processes which lowered water tables and reduced the extent of stream-associated wetlands and the mesic meadows fringes which provide habitat for many species of 'tall' *Mertensia* (Boraginaceae, e.g.
+Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks, processes which lowered water tables and reduced the extent of stream-associated wetlands and the mesic meadows fringes which provide habitat for many species of 'tall' *Mertensia* (e.g.
 *M.
-ciliata* Torr.
-G.
-Don.), to an extent *Delphinium barbeyi*  and many species of native *Trifolium* L.
-(@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).
-Fire suppression resulted in the succession of many Aspen (*Populus tremuloides* Michx.) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effecting habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
+ciliata*), to an extent *Delphinium barbeyi*  and many species of native *Trifolium* (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).
+Fire suppression resulted in the succession of many Aspen (*Populus tremuloides*) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effecting habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
 Finally the effects of Nitrogen deposition, given the West's rapidly growing population still pose adverse effects on the abundance of species of Fabaceae at urban-rural interfaces (see @stevens2018atmospheric, @fenn2003ecological).
 Current solutions to the above issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing.
-The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L., using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
+The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias*, using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus*) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
 
@@
 
-We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a bayesian framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.
+We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.
 Herein we crudely utilized this thinking via binary outcomes, should a species predicted be predicted present or not?
@@
 Myriad data show biological systems and ecological interactions have more variance than can be reasonably discretely parsed.
-We expect that within a bayesian framework studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
+We expect that within a framework developed from our preliminary works studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
 the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems.

### 2023-05-27 script/manuscript_MEE.Rmd
JBF ST edits throughout intro added

@@
 
-Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogeneization (REF), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity).
+Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogenization (@olden2004ecological, @wang2021biotic), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity).
 Monitoring ecological communities structure and interaction will be critical to informing and prioritizing conservation efforts (REF), particularly considering the large proportion of know species currently threaten or likely to be threaten in the near future (@pimm2014biodiversity; @pereira2010scenarios).
@@
 
-  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), within a framework which utilizes custom species sequence databases derived via species distribution modelling, and temporal filtering.
-To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific sequence database relevant to the region of study and the ecological characteristics of interest (@bell2022plants).
-To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g.
-iNaturalist).
-We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
-This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers.
-This approach can significantly reduce processing time, increasing the a projects efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.
+  To address these challenges, we are proposing the generation of a smaller, and more appropriate reference list for identifying candidate taxa from barcode data.
+By reducing the list of likely species in a study site it can increase the accuracy and efficiency of metabarcoding results in plants, which will better inform data collection for projects with limited access to taxonomic expertise.
+Specifically, we propose using current distribution data to generate a list of candidate species relevant to the study area and its ecological characteristics (@bell2022plants).
+Initially such a species list can be leveraged to identify a target collections list of plant tissues to be included in genetic studies.
+It can subsequently help to reduce the size of a reference sequence database, which increases computation speed and efficiency, allowing for the use of genomic data on personal computers.
+This approach can significantly reduce processing time, increasing  project efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.
+We test this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study.
+We use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
+The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen, @johnson2021airborne).
 
@@
 
-  Plant flowering varies over space and also in time, and in particular, contrasts in the flowering periods of many plant species can provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC, @thompson1994coevolutionary, @caradonna2021seeing).
-In high elevation temperate regions, pollination interactions vary temporally and are characterized by high turnover in active periods of species (@caradonna2017interaction, @caradonna2020temporal).
-We work to develop a general approach which seems applicable to many areas which utilize the temporal dimension for classifying sequences in metagenomic samples (but see @davis2022new, Figure 1).
-
 ```{r Filter Diagram, out.width = "35%", fig.cap="Simplified Conceptual Diagram of three approaches leading towards classification of sequencing results, and the number of species associated with them in our area.
-The upper three boxes indicate a common coarse approach, assuming one has a digitizewd Flora, which is not always the case.
+The upper three boxes indicate a common coarse approach, assuming one has a digitized Flora, which is not always the case.
 The center two boxes indicate the computational approach illustrated here.
@@
 
-  We tested this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study.
-In particular, we use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @bell2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen,  @johnson2021airborne).
-The two foraging phases of the queen bumblebee life cycle are essential to 1) increase their weight before diapause, and 2) increase their ovary weights while establishing their recently found nests.
-Both time periods represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
-Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines, while simultaneously serving as the most effective pollinators in many temperate montane ecosystems (@cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline).
-These areas often represent the most diverse areas in the temperate and often offer the sole potential refugia for multiple dimensions of biodiversity under climate change, whilst simultaneously experiencing the greatest proportional changes in mean annual temperature (@brito2018climate, @pepin2022climate).
-An immediate understanding of how to manage previously overlooked keystone insect species, such as bumble bees, is essential if the refugial potential of the temperate mountains are to be incorporated into their current diversity (@loarie2009velocity, @dobrowski2016climate).
-
 # 2 | METHODS
+
+To improve metabarcoding reliability and efficiency, we suggest creating a regional list of candidate species using digital collections gleaned from herbaria, survey work, and community science  (Figure 1).
+This list can further be refined using species distribution models and temporal filtering to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
+The final list is then used to inform collection of plant samples to create a library and inform metabarcoding .
+We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations.
 

### 2023-05-27 script/manuscript_MEE.Rmd
working on some results sections

@@
 The final list is then used to inform collection of plant samples to create a library and inform metabarcoding .
-We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations.
-
-## 2.1 Case Study: Bee-Flower Observations and Pollen Load Collection
-
-Bee and flower observations and bee corbiculae pollen collection was conducted around the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (Appendix 1 for site information).
-The area is characterized by subalpine meadow vegetation communities.
+We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations (@pereira2006towards, )
+
+## Create Species List
+
+### Survey Databases
+
+The first step of our methodological approach consists in collating existing species occurrence data to create  a preliminary list of candidate species likely to occur in the study area.
+Such occurrence data can be retrieved from databases including the Botanical Information and Ecology Network ‘BIEN’ (Maitner (2022)) or the Global Biodiversity Information Facility (GBIF; REF) which aggregate occurrence records from various sources including community science datasets (e.g., iNaturalist) and herbarium records.
+These databases can be queried to extract a list of ecologically relevant vascular plant species with occurrence records in the administrative (e.g.
+county, region) or ecological unit (e.g.
+ecoregion) including the study site or within a search radius surrounding the cite.
+Focusing on a broader geographic region than the study area alone can account for inconsistent sampling effort and undetected species, which are both common in herbarium records and community science datasets
+
+###  Distribution modeling to see which taxa are likely to be within his area
+
+While increasingly used in biodiversity assessments (REFs), datasets like BIEN and GBIF can nonetheless include occurrence records with erroneous taxonomic identification, geographic inaccuracies, or historical records (@freitas2020reliable, SMITH????) Hence additional filtering might be needed to develop a more robust list of candidate species.
+Once a regional list of potential species is established, Species Distribution Models can be used to further reduce the species list to only keep species with ecological requirements found within study sites (Figure 1).
+This step can also help filter for species adapted to the local ecological condition of the study sites, which can be particularly useful in highly heterogeneous landscapes in which a simple survey of databases  may overestimate the list of potential species or when limited taxonomic expertise is available to filter the preliminary species list.
+Species Distribution Models examine relationships between known species occurrences data and a set of environmental variables to identify potentially suitable habitats with similar ecological conditions elsewhere in the landscape (@guisan2000predictive, ...
+).
+They are often used to inform conservation strategies, evaluate invasion risks, or estimate species richness (@thuiller2006using; @pineda2009assessing).
+Environmental variables used as predictors in SDMs typically describe the fundamental ecological niche of the species (e.g., climate, elevation, soil types) or constraints to its distribution (e.g., land uses/land covers, canopy openness) and commonly rely on publicly available geospatial datasets (see Table 1 for examples).
+Once relationships between known occurrence records and these environmental variables are established, SDM algorithms then score remaining portions of the study area on a suitability scale ranging from 0 (highly unsuitable) to 1 (highly suitable) on the basis of their similarity to habitats currently sheltering species populations.
+Once SDMs are generated for the entire pool of species identified through a database search (section 2.3), these suitability maps can be aggregated to refine the list of species likely to occur in study sites.
+
+### Temporal filtering
+
+When studying species interactions, a potentially helpful additional step is to query the candidate species list to retain plant species with phenology that aligns with the study species preferences.
+For example, when studying plant-pollinator interactions, considering that plant flowering varies over space and time, the project may want to focus only on the species flowering during the pollinators’ active period,.
+These contrasts in the phenological periods can thus provide an additional filter for identifying material in certain types of metagenomic samples (@janzen1967synchronization, @Newstrom1994ANC, @thompson1994coevolutionary, @caradonna2017interaction).
+Community science efforts focusing on plant phenology can provide helpful information on the phenology of commonly observed species.
+Examples of such datasets include: the National Phenology Network in the United States (Denny et al.
+2014), SeasonWatch in India, PlantWatch in Canada, and the Pan European Phenology Project.
+Local Floras can also provide useful information on the phenological characteristics of species.
+
+### Collecting representative samples for reference library
+
+GOES HERE
+
+### Using baits to pull out specific regions (barcode) of the genome
+
+For barcoding the environmental samples and leaf samples collected from candidate taxa, we are proposing using a Next generation sequencing approach which incorporates a target capture probes (@jones2016targeted).
+There are several approaches available for preparing genomic libraries that vary in cost, ease of use (CITE), and the quality and quantity of sample DNA required that are all appropriate for target capture approach.
+The choice of best library preparation approach will depend on the type of samples, the lab and sequencing platform preferred.
+Once a library has been generated, and size selected to generate appropriate sized fragments, and appropriately barcoded to distinguish individuals, a synthetic probes, often referred to as baits, is used to preferably retain fragments within the library which contain sequences associated with desired loci, while remainder of fragments which do not are washed away, maximizing sequencing efficiency.
+There are a number of probe sets which have been developed, some are taxon specific, but a handful are more general and appropriate for barcoding multiple taxa, including Angiosperms353 probe set (@johnson2019universal), which is appropriate for most flowering species, and goFlag (@breinholt2021target) might be more appropriate for non-flowering species.
+These approaches  capture data from multiple loci across genome, which are both expressed, therefore have lower mutation rates, and are found to occur in single copies in most taxa, making them appropriate for taxonomic studies.
+
+### Compare barcode to reference library
+
+The data generated is filtered and trimmed using pipelines such as TRIMMOMATIC @bolger2014trimmomatic to identify and remove sequences added during library preparation process and not associated with sample.
+These sequence fragments (contigs) are then compared back to the reference probe sequences to sort them by loci of origin @johnson2016hybpiper.
+For libraries which contain only a single taxa, the contigs from the same loci then be aligned to generates a consensus sequence @katoh2013mafft.
+This consensus sequence can then be used as a reference for which to compare the multiple species samples and identify taxa.
+
+## Case Study: Measuring bumblebee visitation based on corbiculae load
+
+### 2.5.1 Bee-Flower Observations and Pollen Load Collection
+
+To test this metagenomic and informatics approach, we used the identification of plant species within corbiculae pollen loads of bumble bees collected at the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (Appendix 1 for site information).
+Bumblebees are one of the only groups of insects with unequivocal quantitative evidence for numerous population declines.
+They are also the most effective pollinators in many temperate montane ecosystems ( @cameron2020global, @goulson2008decline, @williams1982distribution, @colla2012assessing, @bergman1996micrometeorological, @bingham1998efficient, @grixti2009decline), which are experiencing steep changes in annual temperature under climate change (@brito2018climate, @pepin2022climate).
+As such, the effective management of these ecosystems relies on the study of keystone species, such as bumble bees, that play a critical role in maintaining their current diversity (@loarie2009velocity, @dobrowski2016climate).
+The queen bumblebees emerge early in the season to identify nesting sites.
+Floral resources are essential at these foraging stages for increasing their weight before diapause and increasing their ovary weights while establishing their found nests, and therefore this period represent potential demographic bottlenecks in bumble bee populations (@sarro2022bumble).
+
+### Species list
+
+content goes here
+
+### Pollen and plant material collection
+
+Observations of visitation and bee corbiculae pollen collection was conducted at the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W, 2900 m.a.s.l.), Colorado, USA (Appendix 1 for site information).
 Pollinator observations of Bombus Latreille spp.
-(Apidae Latreille) were conducted from May 29th – July 23rd of 2015 in six study sites as a part of a larger study (described in Ogilvie and CaraDonna 2022).
-Observations of Bombus foraging took place for one hour at each field site, with equal time spent searching for bee in the major vegetation types (dry, and wet meadows, and aspen forest).
-Corbiculae pollens loads were, non-lethally, collected from queens encountered by capturing them in an insect net and transferring them into a restraining device ("bee squeezer", @kearns2001natural).
+(Apidae Latreille) were conducted from May 29th – July 23rd of 2015 in six study sites as a part of a larger study (described in @ogilvie2022shifting).
+Observations of Bombus foraging took place for one hour at each field site.
+Corbiculae pollens loads were non-lethally collected from queens encountered by capturing them in an insect net and transferring them into a restraining device (@kearns2001natural).
 We then collected a single pollen load (i.e., from one leg) from the bee and then released it.
-At weekly intervals at each site, we also recorded the abundances of flowers visited by bumble bees within belt transects spread over the three vegetation types (0.5 x 40 m transects in each vegetation type, 60 m^2^ total area per site).
+At weekly intervals at each site, we also recorded the abundances of flowers visited by bumble bees within belt transects spread over three vegetation types (0.5 x 40 m transects in each vegetation type, 60 m2 total area per site).
+Using five years (2015-2020) of observational data on Bombus queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years.
+In order to capture more variability inherit in the 353 loci we sampled the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
+In addition, we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa to confirm the efficiency of 353 baits for species level identification.
+We also sequenced another 15 taxa of plants commonly visited by Bombus workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set (Appendix 3).
+Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
+
+### Validation process
+
+To precisely classify the contents of each corbiculae load, the sequences classified by molecular methods were compared with the fieldwork which recorded the presence and absence of species and their duration of flowering at a very fine resolution, .
+The quantitative counts of grains from microscopy were combined with the semi-quantitative sequencing results to estimate the abundance of each identified species in each corbiculae load.
+
+### Validation using morphological pollen identification
+
+Five years of observational data on Bombus Queen Bee foraging at these studies sites @ogilvie2022shifting was used as baseline of what we should expect for visitation within this system.
+This was supplemented with the RMBL Vascular Plant Checklist (@fraser2007vpc).
+Microscopy was use to qualitatively identify, and quantitatively note, the plant species present in corbiculae loads.
+A pollen reference library of know taxa visited by Bumbleebees and therefore expected to be present in the corbiculae loads was made using a fuchsin-jelly stained grains from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@beattie1971technique, @brosi2013single).
+A total of 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
+4.13.0) from previously prepared slides and both non-accessioned and RMBL herbarium collections to supplement the number of species and clades covered (Appendix 3).
+To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
+Circa 0.5mm2 of pollen was placed onto a ~4mm2 fuchsin jelly cube @beattie1971technique atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
+The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
+Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in Appendix 4.
+
+To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
+In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained a unique morphotype then all grains in that transect were also identified and counted.
+We used clustering techniques to supplement our subjective opinions of which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
+Ten readily discernible categorical traits were collected from each specimen in the image collection.
+These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
+Using the cluster dendrogram, elbow plot, and heatmaps @fpc2022, of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (Appendix 6).
+This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
+
+Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
+Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects (Appendix 7 & 8).
+
+To reclassify the sequence reads, these data were combined with the floral observation data, and mapped by genus.
+If more than one species in the genus was flowering at that time and site, then the reads were split evenly between the taxa.
+For sequence data which did not match at the genus level, a user subjectively scored them based on the species composition and phenological activity at each site, the queen interaction data, and pollen assignments.
+To estimate the abundance of each of these species in the corbiculae loads, these data were combined with the microscopy data.
+For each morphotype detected in pollen, and each classified sequence read which was not detected via microscopy, they were given a value of 0.5% to indicate their trace presences.
+When more than a single species belonged to a morphotype group in a single sample, the quantitative values from the morphological work were multiplied by the relative sequence abundance of each species in the load.
+All final compositions were standardized to a sum of 100%, by adding or subtracting the differences (induced by classifying records as ‘trace’) to all species with abundances > 1%.
 
@@
 # 3 | RESULTS
+
+### Spatial Analyses
+
+```{r Results for SA Modelling}
+
+files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
+ml <- files[grep( '_ml_', files)]
+ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
+linear <- files[grep('_glm_', files)]
+linear <- unique(sub('_glm.*$', '', linear))
+
+# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
+# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
+
+rm(files, ml, linear)
+```
+
+```{r Results for SA Logistic Regression Metrics}
+
+logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
+  filter(occurrence == 0)
+
+dat <- read.csv('../data/evaluation_table.csv')[,2:3]
+data.frame(Metric  = dat[1:5,1],
+                   Value   = dat[1:5,2],
+                   Metric1 = dat[6:10,1],
+                   Value1  = dat[6:10,2]
+                   ) %>% 
+  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
+    col.names = c('Metric', 'Value', 'Metric', 'Value'),
+    booktabs = T) #%>% 
+  #kableExtra::kable_styling(position = "float_right")
+```
+
+```{r Results for SA Logistic Regression Contingency table}
+read.csv('../data/Pa_conting_table.csv') %>% 
+    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
+    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
+    booktabs = T) %>%
+    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
+
+rm(logR_dat, dat)
+```
+
+The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+
+Across the entire spatial domain of modelling, all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+The 493 ML ensembles accurately predicted the presence of 362 occurrences (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
+Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
+
+### Temporal Analysis
+
+The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
+We could only compare 58 of these 388 species to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts), but this revealed high accord between the long-term ground truthed data set and the modelled species.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
+There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
+
+Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58), and cessation of flowering (p < 0.0001, tau = 0.40).
+
+```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
+knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
+# these are the tenth and nintey  percent non-parametric confidence intervals. 
+```
+
+```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
+knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
+```
+
+### Molecular analysis 
+### Plant data
+
+Need some text here… how well did plant capture go?
+
+### Corbiculae loads
+
+The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
+The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
+Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
+The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
+After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
+`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
+Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
 
@@
 
-## 3.2 | Spatial Analyses
-
-```{r Results for SA Modelling}
-
-files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
-ml <- files[grep( '_ml_', files)]
-ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
-linear <- files[grep('_glm_', files)]
-linear <- unique(sub('_glm.*$', '', linear))
-
-# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
-# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
-
-rm(files, ml, linear)
-```
-
-```{r Results for SA Logistic Regression Metrics}
-
-logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
-  filter(occurrence == 0)
-
-dat <- read.csv('../data/evaluation_table.csv')[,2:3]
-data.frame(Metric  = dat[1:5,1],
-                   Value   = dat[1:5,2],
-                   Metric1 = dat[6:10,1],
-                   Value1  = dat[6:10,2]
-                   ) %>% 
-  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
-    col.names = c('Metric', 'Value', 'Metric', 'Value'),
-    booktabs = T) #%>% 
-  #kableExtra::kable_styling(position = "float_right")
-```
-
-```{r Results for SA Logistic Regression Contingency table}
-read.csv('../data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
-    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
-    booktabs = T) %>%
-    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
-
-rm(logR_dat, dat)
-```
-
-The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
-Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
-
-Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
-
 ## 3.2 | Microscopic Pollen identification
@@
 ### 3.3 | Metabarcoding Pollen Identification
-
-### 3.3.2 | Temporal Analysis
-
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
-
-Only 58 of these 388 species (n = 34.6, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
-Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
-There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
-
-Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
-Due to methodological differences, the peak flowering was not compared, and due to the low performance of attempts to model 'duration' in the previous step it was also not compared.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).
-
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
-# these are the tenth and nintey  percent non-parametric confidence intervals. 
-```
-
-```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
-```
 
@@
 ```
-
-## 3.3.1 | Molecular analysis of corbiculae loads
-
-The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
-The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
-Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
-The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
-
-After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
-`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
 

### 2023-05-29 script/manuscript_MEE.Rmd
writing in word

@@
 --- 
-title: Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees
+title: Protocol to increase accuracy and fidelity of Pollen Meta-genomic Barcoding using Angiosperms353; Case study using Corbiculae loads from wild Bumble Bees
 author:  |
-    |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
+    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$, Jane E. Ogilive$^3$, 
+    |  Sophie Taddeo$^1$$^,$$^2$$^,$$^4$, Emily J. Woodworth$^1$$^,$$^2$
     |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
@@
     |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$, Emily J. Woodworth$^1$$^,$$^2$, 
+    |  $^4$ The University of California at San Diego, 9500 Gilman Dr, La Jolla, California 92093, USA
 abstract:  |  
@@
 library(kableExtra)
+# devtools::install_github("kupietz/kableExtra") you may need this patch if error persists on render may. 27. 2023
 #devtools::install_github("kpmainali/CooccurrenceAffinity")
@@
 
-GOES HERE
+In order to acquire references genomes in a cost-effective manner, we suggest combining existing resources which are readily available at several databases, with *de novo* sequencing of species which are likely to be involved in the interactions of interest.
+
+Although several online genomic databases do contain numerous taxa, these often contain only a single representative of each genus, and do not always account for geographic variability.
+For this reason, we are recommending using the reduced list of potential candidate plant species as a target list to collect leaf samples to generate a reference library.
+This will increase efficiency and accuracy of analyses down the line.
+We recognize that this might not be feasible for all study systems, or for extensively long candidate species list, however using current knowledge of the system and key players in an interaction will allow for more accurate hypothesis testing.
+For example, if interested in pollination, you can focus on known candidate species that represent flower morphologies associated with specific pollinator guilds (Fenster et al 2004).
+DNA extraction and processing of these should be done alongside the samples collected for this system.
 
@@
 
-### 2.5.1 Bee-Flower Observations and Pollen Load Collection
+### Bee-Flower Observations and Pollen Load Collection
 
@@
 
-content goes here
+#### Spatial Analyses
+
+The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+
+Across the entire spatial domain of modelling, all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+The 493 ML ensembles accurately predicted the presence of 362 occurrences (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
+Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
 
@@
 
-## 2.3 | Pollen Morphological identification
-
-### 2.3.1 | Pollen Reference Library
-
-To qualitatively identify, and quantitatively note, the plant species present in corbiculae loads microscopy was used.
-A pollen reference library of fuchsin-jelly stained grains which may be present in corbiculae loads of slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@beattie1971technique, @brosi2013single).
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the RMBL Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  We used clustering techniques to supplement our subjective opinions of which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
-Ten readily discernible categorical traits were collected from each specimen in the image collection.
-These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (Appendix 6).
-This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
-
-### 2.3.2 | Preperation of Pollen Corbiculae Loads
-
-  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Circa* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
-The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in Appendix 4.
-  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained a unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects (Appendix 7 & 8).
-
-### 2.4 | Molecular Barcoding
-
 #### 2.4.1 | Species reference list
@@
 
-### 2.4.2 | Temporal Analyses
-
-For assignment of reads to ecologically probabilistic species subsequent to BLAST, flowering time was used as a filter.
-To estimate the duration of dates in which plant species were flowering Weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not found in the nearest 5), and which were from herbarium records were included.
-To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in fall.
-Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
-Species with > 10 records had their Weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
-These estimates were compared to a long-term observational study of flowering phenology 1974-2012 (@caradonna2014shifts), and the floral abundance data from 2015, using Kendall's tau.
-
-### 2.5.2 | Barcode references library
-
-All lab work was carried out at The Daniel F. and Ada L. Rice Plant Conservation Science Center at the Chicago Botanic Garden, Glencoe, Illinois, U.S.A.
-
-#### 2.5.2.1 | Sampling Species for Barcoding
-
-Using five years (2015-2020) of observational data on *Bombus* queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years.
-In order to capture more variability inherit in the 353 loci we sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
-In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
-We also sequenced another 15 taxa of plants commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (Appendix 3).
-Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
-
-#### 2.5.2.2 | Plant Genomic DNA Extraction
-
-Plant genomic DNA was isolated from ~ 1 cm^2^ of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included two chloroform washes.
-DNA was quantified using a Nanodrop 2000 (Thermo Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer (Thermo Fisher Scientific).
-
-#### 2.5.2.3 | Pollen Genomic DNA Extraction
-
-To extract genomic DNA from pollen a CTAB based protocol was modified from Lahlamgiahi et al.
-and Guertler et al.
-(2014, 2014).
-A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added to the sampled followed by vortexing to allow dissolution of corbiculae.
-Pollen grains were then macerated with Kontes Pellet Pestles, and the tip of these washed with 130 µL of the SDS extraction buffer, samples were then incubated for 1 hour at 30°C.
-This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH.
-8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on a heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
-After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°C for 1 hour.
-Samples were then incubated overnight at 40°C.
-500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.
-
-To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to $\frac{2}{3}$ of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm.
-The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes.
-Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
-Pellets were dried at room temperature overnight before resuspension in nuclease free H~2~O.
-Extractions were assessed using a Nanodrop 2000 (Thermo Fisher Scientific) and Qubit fluorometer (Thermo Fisher Scientific).
-DNA extracts were then cleaned using 2:1 v./v.
-Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute ratio following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions.
-DNA was quantified using a Qubit fluorometer.
-
-#### 2.5.2.4 | Fragmentation, Library Preparation & Target Enrichment
-
-Sequence library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
-Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
-Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed using SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v.
-ethanol washes.
-The exception was the herbarium specimens which were not fragmented and only end repaired, with similar library preparation of all samples.
-Products were analysed on 4% agarose gels, and a Qubit fluorometer.
-Libraries were pooled and enriched with the Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence Capture Kit) by following the manufacturer's protocol and Brewer et al.
-2019.
-Sequencing was performed using an Illumina mi-Seq with 150-bp end reads, (NUSeq Core, Chicago, Illinois).
-
-###  2.6 | Computational Processes and Analyses. 
-####  2.6.1 | Reference Library Data Processing
-
-Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950).
-Contigs generated were mapped to a reference with HybPiper with using target files created by M353 (@johnson2016hybpiper, @mclay2021new).
-
-#### 2.6.2 | Sequence Identification
-
-A custom Kraken2 database was created by downloading representative species indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
-These sequences were processed in the same manner as our novel sequences.
-The Kraken2 database was built using default parameters.
-Kraken2 was run on sequences using default parameters (Appendix 9).
-Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken).
-FInally all reads which could be classified by these databases were passed to a local BLAST database.
-
-#### 2.6.3 | Identification of Sequence Matching Loci
-
-A local NCBI database was built using the same processed novel and downloaded sequences as the previous database (@camacho2009blast).
-
 ### 2.7 | Integrated Observational, Molecular, and Palynological Corbiculae
@@
 
-### Spatial Analyses
-
-```{r Results for SA Modelling}
-
-files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
-ml <- files[grep( '_ml_', files)]
-ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
-linear <- files[grep('_glm_', files)]
-linear <- unique(sub('_glm.*$', '', linear))
-
-# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
-# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
-
-rm(files, ml, linear)
-```
-
-```{r Results for SA Logistic Regression Metrics}
-
-logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
-  filter(occurrence == 0)
-
-dat <- read.csv('../data/evaluation_table.csv')[,2:3]
-data.frame(Metric  = dat[1:5,1],
-                   Value   = dat[1:5,2],
-                   Metric1 = dat[6:10,1],
-                   Value1  = dat[6:10,2]
-                   ) %>% 
-  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
-    col.names = c('Metric', 'Value', 'Metric', 'Value'),
-    booktabs = T) #%>% 
-  #kableExtra::kable_styling(position = "float_right")
-```
-
-```{r Results for SA Logistic Regression Contingency table}
-read.csv('../data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
-    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
-    booktabs = T) %>%
-    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
-
-rm(logR_dat, dat)
-```
-
-The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-Across the entire spatial domain of modelling, all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-The 493 ML ensembles accurately predicted the presence of 362 occurrences (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
-Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
-
-### Temporal Analysis
-
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
-We could only compare 58 of these 388 species to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts), but this revealed high accord between the long-term ground truthed data set and the modelled species.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
-There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
-
-Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58), and cessation of flowering (p < 0.0001, tau = 0.40).
-
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
-# these are the tenth and nintey  percent non-parametric confidence intervals. 
-```
-
-```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
-```
-
-### Molecular analysis 
-### Plant data
-
-Need some text here… how well did plant capture go?
-
-### Corbiculae loads
-
-The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
-The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
-Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
-The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
-After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
-`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
-
-## 3.1 | Floral Observations
-
-```{r Floral Observations Summary Statistics }
-
-files <- list.files('../data', pattern = 'csv')
-
-flr_ranks <- read.csv(
-  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
-  select(site, doy, week, tot.obs.length, habitat, plant.species:abun.rank)
-obs_time <- read.csv(
-  file.path('../data', '/', files[grep('queen_observations', files)]) ) %>% 
-  select(site, week, tot.obs.length) %>% 
-  distinct(site, week, .keep_all = T)
-
-bee_obs <- read.csv(
-  file.path('../data', files[grep('queen_observations', files)]) ) %>% 
-  filter(caste == 'q') %>% 
-  mutate(species = gsub('\\..*$', '', species)) %>% 
-  select(site, doy, week, tot.obs.length, species, plant.species,
-         resource.coll:pollen.collected, fl.switch.1:fl.switch.2.resource.coll) %>% 
-  mutate(across(where(is.character), ~ na_if(., ''))) %>% 
-  pivot_longer(c(plant.species, fl.switch.1, fl.switch.2), names_to = 'observation', values_to = 'plant.species') %>% 
-  drop_na(plant.species) %>% 
-  group_by(site, week, species, plant.species) %>% 
-  count(name = 'interactions') %>% 
-  left_join(., obs_time, by = c("site", "week")) %>% 
-  mutate(period = case_when(
-    week %in% 3:5 ~ 'Early',
-    week %in% 6:8 ~ 'Mid',
-    week %in% 9:11 ~ 'Late'
-  ))
-
-no_plant_sp <- bee_obs %>% 
-  ungroup() %>% 
-  distinct(plant.species) %>%
-  nrow()
-
-s <- toString(format(as.Date(min(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # start obs
-e <- toString(format(as.Date(max(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # end obs
-
-t_b_wk <- bee_obs %>%  group_by(site, week, species) %>%  tally(interactions) 
-t_b_b <- bee_obs %>%  group_by(species) %>%  tally(interactions) %>%  filter(species != 'unknown') 
-t_b_wk_pl <- bee_obs %>%  group_by(site, week, plant.species) %>%  tally(interactions) 
-t_b_pl <- bee_obs %>%  group_by(plant.species) %>%  filter(species != 'unknown') %>%  tally(interactions) 
-t_spp_i <- bee_obs %>%  group_by(species) %>%  filter(species != 'unknown') %>%  distinct(plant.species) %>% 
-  count()
-
-```
-
-In 2015 the six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
-A total of `r toString(sum(bee_obs$interactions))` overwintered queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
-Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week minimum = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, median = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (minimum = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, median = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
-The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, median = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
-
-A total of 66 corbiculae loads were collected from bees, 64 of them from queens.
-
-```{r echo = F}
-rm(bee_obs, flr_ranks, obs_time, t_b_b, t_b_wk, t_b_pl, t_b_wk_pl, e, s, files, t_spp_i)
-```
-
-```{r Most used resources by species, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
+```{r Most used resources by species, out.width = "45%"}
 
@@
 
-bee_cols <- c('appositus' = '#A50026FF', 'bifarius' = '#FFA319FF', 
-              'californicus' = '#9632B8', 'flavifrons' = '#EFAE90',
-              'rufocinctus' = '#155F83FF', '#8F3931FF', 
-                'mixtus' = '#350E20FF', 'kirbiellus'  =  '#8074A8', 
-              'nevadensis' = '#C96731')
-
-ggplot(data=correct, aes(x=Observation, y=value,
-                         group = Bombus.Species, fill = Bombus.Species)) +
-  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
-              height = 0.05, width = 0.3) +
-  scale_fill_manual('Bee species', values = bee_cols) +
-  scale_color_manual('Bee species', values = bee_cols) +
-  theme_bw() +
-  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
-  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
-       y = 'Shared Observations', x= NULL) + 
-  theme(panel.grid.major.x = element_blank(),
-        panel.border = element_blank())
-
 n_corbiculae <- tb %>% 
@@
 
-rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
-```
-
-## 3.2 | Microscopic Pollen identification
+```
 
@@
 ```
-
-Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into (Appendix 6).
-From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we identified substantial amounts of the abundance and morphotype richness of the samples (morphotype richness, $\bar{x}$ = 4.5, median = 4, min = 1, max = 9) (Appendix  7 & 8).
-The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, median = `r  median(grains_counted)`).
-
-```{r}
-rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected, grains_counted)
-```
-
-```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
-The innermost ring indicates every genus which Queen Bee's were observed to visit.
-The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
-The outermost ring indicates that sequence data were available for at least a member of that genus.
-Branch colors follow APG 4."}
-knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
-```
-
-### 3.3 | Metabarcoding Pollen Identification
 
@@
 
-The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
-A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix 15).
-Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
-Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species representing `r sp_total_seqs`% of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family.
-
-Of the `r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were all assigned via post-processing conditions (`r j`, Appendix 15).
-These were manually assigned to a variety of ranks, occasionally to genus - `r nrow(gen)`, and species - `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
-
-To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
-The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less the Heliantheae Aliiance (Sunflower Tribe) (Boraginaceae grains `r toString(aa)`.
-Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`.
-
-```{r clean sequence numbers from environment}
-rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
-    sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
-```
-
-```{r Correlation between Number of Grains and Number of Sequences, fig.cap = 'Relationship between morphological count data and sequence reads', out.height="50%", out.width="50%"}
-
-pol_p <- pollen_props %>% 
-  mutate(sequenceProp = sequenceProp/100, 
-         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
-
-ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
-# ct[["p.value"]] #too small to write out. tau > 0.4 strong relationship
-model <-  lm(adjSeqProp ~ countProp, data = pol_p)
-
-# png('../../thesis_talk/graphics/plots/CorrelationCountedSequences.png', )
-par(pty="s")
-plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-      xlab = 'Grains Counted', ylab = 'Classified Sequences',
-      pch=18, slty = 3, gap = 0.025)
-axis(1)
-axis(2)
-abline(model)
-text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
-text(0.3, 0.8, "p < 0.001") 
-# dev.off()
-
-# rm( newdat)
-```
-
-To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
-In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
-
-The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly linear, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
-This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing).
-There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
-
-```{r}
-rm(pol_p, ct)
-```
-
 ```{r Co-occurrence of splash drop sequences, eval = F}
@@
 ```
-
-A number of suspected false positives were returned by the sequence classified, in all instances these were species present in low abundances at the field station which were also returned with a common species in the same family.
-To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala Jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
-Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
-sp., *Thalictrum* L.
-spp., *Trollius* L.
-sp., *Aquilegia* L.
-spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
-were recorded.
-A similar relationship was observed in the Hydrophyllaceae R.Br.
-with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
-species were present.
-The flower size  and sparsity of *Nemophila breviflora* A.
-Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
-The floral morphology and orientation of flower of *Thalictrum*  spp.
-also makes them unlikely to be visited, and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
-
-## 3.6 | Integrated Observational, Molecular, and Palynological Results
 
@@
   ) %>% 
+  collapse_rows() %>% 
   add_footnote(
-    'The mean only calculated across the samples where the species was detected') %>% 
-  collapse_rows()
-
-```
+    'The mean only calculated across the samples where the species was detected')
+
+```
+
+### Spatial Analyses
+
+```{r Results for SA Modelling}
+
+files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
+ml <- files[grep( '_ml_', files)]
+ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
+linear <- files[grep('_glm_', files)]
+linear <- unique(sub('_glm.*$', '', linear))
+
+# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
+# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
+
+rm(files, ml, linear)
+```
+
+```{r Results for SA Logistic Regression Metrics}
+
+logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
+  filter(occurrence == 0)
+
+dat <- read.csv('../data/evaluation_table.csv')[,2:3]
+data.frame(Metric  = dat[1:5,1],
+                   Value   = dat[1:5,2],
+                   Metric1 = dat[6:10,1],
+                   Value1  = dat[6:10,2]
+                   ) %>% 
+  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
+    col.names = c('Metric', 'Value', 'Metric', 'Value'),
+    booktabs = T) #%>% 
+  #kableExtra::kable_styling(position = "float_right")
+```
+
+```{r Results for SA Logistic Regression Contingency table}
+read.csv('../data/Pa_conting_table.csv') %>% 
+    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
+    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
+    booktabs = T) %>%
+    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
+
+rm(logR_dat, dat)
+```
+
+The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
+A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
+Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
+
+Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+
+Across the entire spatial domain of modelling, all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
+The 493 ML ensembles accurately predicted the presence of 362 occurrences (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
+Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
+The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
+Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
+Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
+
+### Temporal Analysis
+
+The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
+Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
+We could only compare 58 of these 388 species to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts), but this revealed high accord between the long-term ground truthed data set and the modelled species.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
+There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
+
+Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
+There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58), and cessation of flowering (p < 0.0001, tau = 0.40).
+
+```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
+knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
+# these are the tenth and nintey  percent non-parametric confidence intervals. 
+```
+
+```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
+knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
+```
+
+### Molecular analysis 
+### Plant data
+
+Need some text here… how well did plant capture go?
+
+### Corbiculae loads
+
+The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
+The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
+Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
+The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
+After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
+`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
+Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
+
+### Metabarcoding Pollen Identification
+
+content should be here
+
+### Validation
+
+```{r Floral Observations Summary Statistics }
+
+files <- list.files('../data', pattern = 'csv')
+
+flr_ranks <- read.csv(
+  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
+  select(site, doy, week, tot.obs.length, habitat, plant.species:abun.rank)
+obs_time <- read.csv(
+  file.path('../data', '/', files[grep('queen_observations', files)]) ) %>% 
+  select(site, week, tot.obs.length) %>% 
+  distinct(site, week, .keep_all = T)
+
+bee_obs <- read.csv(
+  file.path('../data', files[grep('queen_observations', files)]) ) %>% 
+  filter(caste == 'q') %>% 
+  mutate(species = gsub('\\..*$', '', species)) %>% 
+  select(site, doy, week, tot.obs.length, species, plant.species,
+         resource.coll:pollen.collected, fl.switch.1:fl.switch.2.resource.coll) %>% 
+  mutate(across(where(is.character), ~ na_if(., ''))) %>% 
+  pivot_longer(c(plant.species, fl.switch.1, fl.switch.2), names_to = 'observation', values_to = 'plant.species') %>% 
+  drop_na(plant.species) %>% 
+  group_by(site, week, species, plant.species) %>% 
+  count(name = 'interactions') %>% 
+  left_join(., obs_time, by = c("site", "week")) %>% 
+  mutate(period = case_when(
+    week %in% 3:5 ~ 'Early',
+    week %in% 6:8 ~ 'Mid',
+    week %in% 9:11 ~ 'Late'
+  ))
+
+no_plant_sp <- bee_obs %>% 
+  ungroup() %>% 
+  distinct(plant.species) %>%
+  nrow()
+
+s <- toString(format(as.Date(min(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # start obs
+e <- toString(format(as.Date(max(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # end obs
+
+t_b_wk <- bee_obs %>%  group_by(site, week, species) %>%  tally(interactions) 
+t_b_b <- bee_obs %>%  group_by(species) %>%  tally(interactions) %>%  filter(species != 'unknown') 
+t_b_wk_pl <- bee_obs %>%  group_by(site, week, plant.species) %>%  tally(interactions) 
+t_b_pl <- bee_obs %>%  group_by(plant.species) %>%  filter(species != 'unknown') %>%  tally(interactions) 
+t_spp_i <- bee_obs %>%  group_by(species) %>%  filter(species != 'unknown') %>%  distinct(plant.species) %>% 
+  count()
+
+```
+
+In 2015 the six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
+A total of `r toString(sum(bee_obs$interactions))` overwintered queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
+Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week minimum = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, median = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (minimum = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, median = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
+The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, median = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
+
+A total of 66 corbiculae loads were collected from bees, 64 of them from queens.
+
+```{r echo = F}
+rm(bee_obs, flr_ranks, obs_time, t_b_b, t_b_wk, t_b_pl, t_b_wk_pl, e, s, files, t_spp_i)
+```
+
+```{r Most used resources by species figure, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
+
+bee_cols <- c('appositus' = '#A50026FF', 'bifarius' = '#FFA319FF', 
+              'californicus' = '#9632B8', 'flavifrons' = '#EFAE90',
+              'rufocinctus' = '#155F83FF', '#8F3931FF', 
+                'mixtus' = '#350E20FF', 'kirbiellus'  =  '#8074A8', 
+              'nevadensis' = '#C96731')
+
+ggplot(data=correct, aes(x=Observation, y=value,
+                         group = Bombus.Species, fill = Bombus.Species)) +
+  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
+              height = 0.05, width = 0.3) +
+  scale_fill_manual('Bee species', values = bee_cols) +
+  scale_color_manual('Bee species', values = bee_cols) +
+  theme_bw() +
+  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
+  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
+       y = 'Shared Observations', x= NULL) + 
+  theme(panel.grid.major.x = element_blank(),
+        panel.border = element_blank())
+
+rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
+```
+
+## Microscopic Pollen identification
+
+Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into (Appendix 6).
+From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we identified substantial amounts of the abundance and morphotype richness of the samples (morphotype richness, $\bar{x}$ = 4.5, median = 4, min = 1, max = 9) (Appendix  7 & 8).
+The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, median = `r  median(grains_counted)`).
+
+```{r}
+rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected, grains_counted)
+```
+
+```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
+The innermost ring indicates every genus which Queen Bee's were observed to visit.
+The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
+The outermost ring indicates that sequence data were available for at least a member of that genus.
+Branch colors follow APG 4."}
+knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
+```
+
+### 3.3 | Metabarcoding Pollen Identification
+
+The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
+A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix 15).
+Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
+Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species representing `r sp_total_seqs`% of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family.
+
+Of the `r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
+These were all assigned via post-processing conditions (`r j`, Appendix 15).
+These were manually assigned to a variety of ranks, occasionally to genus - `r nrow(gen)`, and species - `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
+
+To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
+The pollen samples contained three morphotypes which could readily be identified via microscopy.
+Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less the Heliantheae Aliiance (Sunflower Tribe) (Boraginaceae grains `r toString(aa)`.
+Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
+Both morphotypes of Asteraceae pollen `r toString(dd)`.
+
+```{r clean sequence numbers from environment}
+rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
+    sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
+```
+
+```{r Correlation between Number of Grains and Number of Sequences, fig.cap = 'Relationship between morphological count data and sequence reads', out.height="50%", out.width="50%"}
+
+pol_p <- pollen_props %>% 
+  mutate(sequenceProp = sequenceProp/100, 
+         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
+
+ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
+# ct[["p.value"]] #too small to write out. tau > 0.4 strong relationship
+model <-  lm(adjSeqProp ~ countProp, data = pol_p)
+
+# png('../../thesis_talk/graphics/plots/CorrelationCountedSequences.png', )
+par(pty="s")
+plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
+       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
+      xlab = 'Grains Counted', ylab = 'Classified Sequences',
+      pch=18, slty = 3, gap = 0.025)
+axis(1)
+axis(2)
+abline(model)
+text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
+text(0.3, 0.8, "p < 0.001") 
+# dev.off()
+
+# rm( newdat)
+```
+
+To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
+In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
+if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
+The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
+
+The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly linear, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
+This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing).
+There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
+
+```{r}
+rm(pol_p, ct)
+```
+
+A number of suspected false positives were returned by the sequence classified, in all instances these were species present in low abundances at the field station which were also returned with a common species in the same family.
+To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala Jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
+Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
+sp., *Thalictrum* L.
+spp., *Trollius* L.
+sp., *Aquilegia* L.
+spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
+were recorded.
+A similar relationship was observed in the Hydrophyllaceae R.Br.
+with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
+species were present.
+The flower size  and sparsity of *Nemophila breviflora* A.
+Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
+The floral morphology and orientation of flower of *Thalictrum*  spp.
+also makes them unlikely to be visited, and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
 

### 2023-07-07 script/expert_opinion_reclassification.R
added file writeout for values JBF csv;

@@ -397,6 +397,11 @@ morpho_lkp <- data.frame(
 
 rm(a, b, c, d, assorted, Asteraceae_1, Asteraceae_2)
 
+###############################################################################
+###      write out file detailing pollen percentages for JBF July 6th       ###
+###############################################################################
+write.csv(morpho_lkp, '../data/pollen_morphotype_grp_lookup.csv', row.names = F)
+write.csv()
 
 ## read in the sample collection csv, for species not noted add them. 
 

### 2023-07-07 script/collate_interaction_values.R
collating records for jbf and chart

@@ -1,0 +1,51 @@
+setwd('~/Documents/ManuSCript/script')
+
+library(tidyverse)
+
+pmorpho <- read.csv('../data/morpho_pollen.csv') %>% 
+  filter(caste == 'q') %>%  # reduce to queens only here, use this as the filtering join later
+  select(Sample = sample.id, Bee_spp = species, Site = site, Date = date,
+         Plant_spp_morpho = morphotype, Count = sum ) %>% 
+  mutate(Observation = 'Microscopy')
+
+site_info <- select(pmorpho,   Sample, Bee_spp, Site, Date) %>% 
+  distinct()
+  
+morpho_lkp <- read.csv('../data/pollen_morphotype_grp_lookup.csv') %>% 
+  mutate(morphotype = str_to_upper(morphotype))
+
+blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
+  group_by(Sample) %>% 
+  slice_max(Prcnt_seqs, n = 10) %>% # reduce to the top 10 records
+  ungroup() %>% 
+  select(Sample, Plant_spp_morpho = TAXON, Count = Seqs_pr_taxon) %>% 
+  mutate(Observation = 'Molecular') %>% 
+  inner_join(., site_info, by = 'Sample', relationship = "many-to-many")
+# use inner join to drop the unmatched records
+
+
+queen_obs <- read.csv('../data/Bombus_queen_observations_2015.csv') #%>% 
+  select(Bee_spp = species, Site = site, Date = date,
+         Plant_spp_morpho = plant.species ) %>% 
+  mutate(Bee_spp = str_remove(Bee_spp, '[.]orange|[.]dark'), 
+         Plant_spp_morpho = str_replace(Plant_spp_morpho, '[.]', ' '),
+         Date = as.character(lubridate::dmy(Date)))  %>% 
+  count(Bee_spp, Site, Date, Plant_spp_morpho, name = 'Count') %>% 
+  mutate(Observation = 'Field', Sample = NA) %>% 
+  
+  mutate(
+    Plant_spp_morpho = str_replace(Plant_spp_morpho, 'leucanthus', 'lanszwertii'),
+    Plant_spp_morpho = str_replace(Plant_spp_morpho, 'bakeri', 'sericeus'),
+    Plant_spp_morpho = str_replace(Plant_spp_morpho, 'Adenolinum', 'Linum'),
[…]

### 2023-07-07 script/expert_opinion_reclassification.R

@@ -401,7 +401,6 @@ rm(a, b, c, d, assorted, Asteraceae_1, Asteraceae_2)
 ###      write out file detailing pollen percentages for JBF July 6th       ###
 ###############################################################################
 write.csv(morpho_lkp, '../data/pollen_morphotype_grp_lookup.csv', row.names = F)
-write.csv()
 
 ## read in the sample collection csv, for species not noted add them. 
 

### 2024-01-03 script/compare_results_BLAST2.R
citations again

@@ -56,6 +56,7 @@ all_classifications <- bind_rows(
   ), 
   Family = if_else(Taxon == 'Ericaceae', 'Eriaceae', Family))
 
+
 rm(genera2query, genera_out, automated, blast, expert)
 
 expert <- all_classifications %>% filter(Method == 'Expert')
@@ -210,12 +211,35 @@ ACC_sens_Spec <- bind_rows(
   mutate(
     Species_false = NO_Species - Species_T,
     
-    Sp_Acc = (Species_T / NO_Species) * 100,
-    Gen_Acc = (Genera_T / sum(NO_Genus, NO_Species, na.rm = T) ) * 100,
-    Sp_Sens = (Species_T)  / (Species_T + (Spp_No - Species_T)) * 100
-  )
+    Sp_Sens = (Species_T / Spp_No) * 100, 
+    Sp_Prec = (Species_T / (Species_T + Species_F)) * 100,
+    FDR = (Species_F / (Species_F + Species_T)) * 100, 
+    F1 = ((2*Species_T) / (2*Species_T + Species_F + FN)) * 100
+  ) #%>% 
+  select(Sample, Method, NO_Species, Species_T, Sp_Acc )
 
 rm(spp_count, gen_count, fam_count, blast_results, computer_results, cnts)
+
+
+#or# Sensitivity
+Sensitivity = (TP / P)  * 100
+Sensitivty = (TP / (TP + FN)) * 100
+
+# Precision
+Precision = (TP / (TP + FP)) * 100
+
+# False Discovery Rate
+FDR = (FP / (FP + TP)) * 100
+
+# F1 Score
+F1 = ((2*TP) / (2*TP + FP + FN)) * 100
+  
+# Condition Positive P = the number of species in the 'expert' list
+# True Positive TP = The number of species in the target 'naive|reclass' which are in the 'expert' list
+# False Positive FP = The # of species in the target 'naive|reclass' which are NOT in the 'expert' list
+# False Negative FN = The # of species in the 'expert' list which are NOT in the 'naive|reclass' list
+# True Negative TN = 
+
 # Spp_No - Species_match is my TRUE NEGATIVE
 # species_false is MY FALSE POSITIVE
 
@@ -223,8 +247,8 @@ write.csv(ACC_sens_Spec, '../data/AccuracyOfPostClassification.csv', row.names =
 
 
 wilcox.test(
[…]

### 2024-02-11 script/SupplementaryInformation.Rmd
subset supplementals

--- 2023-05-27 script/manuscript_MEE.Rmd
@@
 ---
-title: Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees
-author:  |
-    |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
-    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
-    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
-    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$, Emily J. Woodworth$^1$$^,$$^2$, 
-abstract:  |  
-  1) DNA Barcoding has been successful for the rapid identification of species in complex ecological assemblages.
-However, metabarcoding in the plant kingdom has been limited due to a lack of universal gene regions that works across all taxa, limiting the applications of eDNA and metagenomics in ecology.
-  2) To circumvent these limitations, we propose using more holistic approach which combines a multiple gene approach that incorporates existing plant occurrence databases, species distribution models, and phenology analyses to generate a list of likely candidate species to increase accuracy.
-Such an effort can help guide plant sample collection and library construction, particularly for projects with limited resources or local taxonomic knowledge.
-In addition, we demonstrate that building a custom candidate plant list based on known occurrences, ecological requirements, and phenology can improve computing efficiency, speed, and accuracy of metabarcoding.
-We tested this approach to identify the plant species contributing to the corbiculae pollen loads of bumble bees.
-To validate the efficiency of this approach we compare results of the DNA barcoding to expert-led pollen identification and field observations of plant-pollinator interactions that generated these pollen loads.
-  3) We show that the Angiosperms 353 probes, developed for phylogenomics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches around the globe.
-The DNA barcoding of bumble bee corbiculae pollen loads was most accurate when combined with knowledge of what plant species were flowering in the plant community when they were collected.
-  4) Supplementing DNA barcoding data with ecological context is most accurate and powerful.  
-keywords: |
-  DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species distribution model
-output:
-  word_document: default
-  toc: no
-  pdf_document: default
-csl: "../citations/methods-in-ecology-and-evolution.csl"
-bibliography: ../citations/citations.bib
-link-citations: yes
-fig_caption: yes
-always_allow_html: yes
-header-includes:
-- \usepackage{endfloat}
-- \usepackage{setspace}\doublespacing
-- \usepackage{lineno}
-- \linenumbers
+title: "Supplemental Information"
+author: "steppe"
+output: pdf_document
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, warning = F, message = F)
+```{r, include=FALSE}
+knitr::opts_chunk$set(
+  comment = "#>", echo = FALSE, fig.width = 6
+)
 ```
 
-```{r Required Libraries, message=F, warning=F, results='hide'}
-library(tidyverse)
-library(bookdown)
-library(knitr)
-library(kableExtra)
-#devtools::install_github("kpmainali/CooccurrenceAffinity")
-suppressWarnings(library(kableExtra))
-pNumC <- function(...) prettyNum(...,big.mark = ',')
+```{r Create Table 1}
 
-prcnt_detect <- function(x, Lprop, Hprop){
-  if(missing(Lprop)){Lprop <- 0}
+variable <- c(
+  rep('Climate', times = 4),
+  rep('Elevation', times = 4), 
+  rep('Land Uses / Land Covers', times = 4), 
+  rep('Vegetation Cover', times = 1), 
+  rep('Soils & Geomorphology', times = 4)
+)
 
-  y <- x[x$Prop > Lprop & x$Prop < Hprop,]; total <- nrow(y)
-  detected <- round((nrow(y[y$DETECTED == 1,])/total) * 100, 1)
-  Mdn <- round(median(y$Prop),3)
+dataSrc <- c(
+  'WorldClim', 
+  'CHELSA',
+  'Gridded Surface Meterological (GRIDMET)',
+  'PRISM',
 
-  r <- paste0('were detected in ',  detected, 
-                '% of samples where the proportion of target grains were between ', 
-                Lprop, '-', Hprop,  ' (n = ', total, ' Mdn = ', Mdn, ')')
-  return(r)
-}
+  'Multi-Error-Removed Improved-Terrain DEM', 
+  'Global Multi-resolution Terrain Elevation Data (GMTED2010)',
+  'Aster Global Digital Elevation Model (ASTCTM)', 
+  'United States National Elevation Dataset',
+
+  'Dynamic World', 
+  'Global Land Cover Characterization (GLCC)', 
+  'Moderate Resolution Imaging Spectroradiometer (MODIS) Land Cover', 
+  'National Land Cover Dataset (NLCD)',
+
+  'Remote Sensing Products (e.g. Landsat, MODIS)',
+
+  'SoilGrids', 
+  'Harmonized World Soil Database', 
+  'GeoMorpho90m',
+  'Web Soil Survey'
+)
+
+coverage <- c(
+  rep('Global', times = 2),
+  rep('CONUS', times = 2),
+
+  rep('Global', times = 3),
+  'CONUS',
+
+  rep('Global', times = 3), 
+  'CONUS',
+
+  'Global',
+
+  rep('Global', times = 3), 
+  'CONUS'
+)
+
+df <- data.frame(variable, dataSrc, coverage)
+
+info <- 'Example of environmental variables and data sources commonly used in species distribution models that could be used during the spatial filtering step of our approach.'
 ```
 
-# 1 | INTRODUCTION
-
-Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogenization (@olden2004ecological, @wang2021biotic), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity).
-Monitoring ecological communities structure and interaction will be critical to informing and prioritizing conservation efforts (REF), particularly considering the large proportion of know species currently threaten or likely to be threaten in the near future (@pimm2014biodiversity; @pereira2010scenarios).
-However, ecosystem monitoring remains inconsistent in time and space (@yoccoz2001monitoring) as it requires substantial resources, thereby leaving several regions, ecosystems, and even species under-observed (@pereira2012global; @collen2008tropical; @meyer2015global; @ruete2015displaying).
-
-Species interactions are a critical component of ecosystem stability and therefore biodiversity conservation (Soltis et al.
-(2019), @futuyma2009macroevolution, Voje et al.
-(2015), Weber et al.
-(2017), @hembry2020ecological).
-Species interactions, which span from positive (e.g., symbiosis, mutualistic) to negative (e.g., predation, competition), shape species distribution (@bascompte2009mutualistic; Wisz et al.
-2013) and tolerance to environmental stressors (REF).
-They impact ecosystem stability and diversity (@agrawal2007filling, @valiente2015beyond, @bascompte2006asymmetric, and the relationships between biodiversity and ecosystem functions (@duffy2007functional).
-However, species interactions are challenging to observe (REFs).
-They are typically measured through shorter-term observational studies yet are highly variable in time and space (@liu2022environmental).
-Studying species interactions also requires extensive taxonomic expertise, especially those from diverse clades (@hebert2003biological), further challenging their frequent assessment.
-As a result, few ecological networks have been fully mapped (REFs).
-Improving our capacity to efficiently study species interactions consistently across several ecosystems, landscapes, and management context is thus pivotal to enacting conservation interventions (@tylianakis2010conservation).
-This is particularly urgent as increased biological introductions and climate change are impacting species interactions at a rapid rate, thus calling for more frequent and consistent monitoring (REFs).
-
-  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
-The ability to identify species from a fragments of organisms (e.g.
-hair, scat, leaf tissue) increases our ability to understand the interactions of not only entire ecosystems but also a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
-those for restoration processes (Banerjee et al.
-(2022), Johnson et al.
-(2023)).
-With plants the identification of members of certain clades to species using barcoding has been successful in some groups (@kress2017plant), whereas many other clades have proven more problematic (@liu2014identification, @china2011comparative, @coissac2012bioinformatic).
-Metabarcoding incurs additional challenges to those which exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
-Particular limitations of the traditional high copy number barcodes (e.g.
-ITS2, *rbcL*, *matK*, *trnH-psbA*) include the utilization their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
-Currently, most plant metabarcoding endeavors only allow the identification of material to the level of family or genus.
-
-```{r Applications of Metabarcoding}
-read.csv('../data/Applications_Table.csv') |>
-  knitr::kable(caption = 'Applications of Plant Metabarcoding',
-             col.names = c('Application', 'Example', 'Citations')) |>
-  kable_styling(full_width = TRUE)
-```
-
-  Currently the largest plant systematic endeavor ever undertaken, by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000+ represented species, using the popular Angiosperms353 (A353) probes (@baker2021PAFTOL, @johnson2019universal).
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute more species by 2030 (@cheng2018tenkp).
-These publicly available data serve to provide a phylogenetically comprehensive backbone for plant metabarcoding.
-Similar projects, focused on whole genome sequencing of all organisms in certain geographic areas, such as the ‘Darwin Tree of Life’, which aims to sequence all described taxa in Britain and Ireland, will provide geographically dense data sets to seed projects globally (@darwin2022project, @lewin2022biogenome).
-These novel genomic data sets will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
-However, the application of metabarcoding still faces challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
-
-  To address these challenges, we are proposing the generation of a smaller, and more appropriate reference list for identifying candidate taxa from barcode data.
-By reducing the list of likely species in a study site it can increase the accuracy and efficiency of metabarcoding results in plants, which will better inform data collection for projects with limited access to taxonomic expertise.
-Specifically, we propose using current distribution data to generate a list of candidate species relevant to the study area and its ecological characteristics (@bell2022plants).
-Initially such a species list can be leveraged to identify a target collections list of plant tissues to be included in genetic studies.
-It can subsequently help to reduce the size of a reference sequence database, which increases computation speed and efficiency, allowing for the use of genomic data on personal computers.
-This approach can significantly reduce processing time, increasing  project efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.
-We test this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study.
-We use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen, @johnson2021airborne).
-
-```{r Current Bottlenecks in Plant Metabarcoding}
-read.csv('../data/Bottlenecks_Table.csv') |>
-  knitr::kable(caption = 'Current Issues Facing Plant Metagenomics',
-             col.names = c('Issue', 'Our Approach', 'Possible Advances')) |>
-  kable_styling(full_width = TRUE)
-```
-
-```{r Filter Diagram, out.width = "35%", fig.cap="Simplified Conceptual Diagram of three approaches leading towards classification of sequencing results, and the number of species associated with them in our area.
-The upper three boxes indicate a common coarse approach, assuming one has a digitized Flora, which is not always the case.
-The center two boxes indicate the computational approach illustrated here.
-The final box indicates the use of the expert field data in the case study.
-The stem of the final applies to both Computational and Expert Survey results, and should be thought of as using time ala chromatography, in the post-classification process."}
-knitr::include_graphics('../graphics/plots/filter_diagram.png')
-```
-
-# 2 | METHODS
-
-To improve metabarcoding reliability and efficiency, we suggest creating a regional list of candidate species using digital collections gleaned from herbaria, survey work, and community science  (Figure 1).
-This list can further be refined using species distribution models and temporal filtering to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
-The final list is then used to inform collection of plant samples to create a library and inform metabarcoding .
-We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations.
-
-## 2.1 Case Study: Bee-Flower Observations and Pollen Load Collection
-
-Bee and flower observations and bee corbiculae pollen collection was conducted around the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (Appendix 1 for site information).
-The area is characterized by subalpine meadow vegetation communities.
-Pollinator observations of Bombus Latreille spp.
-(Apidae Latreille) were conducted from May 29th – July 23rd of 2015 in six study sites as a part of a larger study (described in Ogilvie and CaraDonna 2022).
-Observations of Bombus foraging took place for one hour at each field site, with equal time spent searching for bee in the major vegetation types (dry, and wet meadows, and aspen forest).
-Corbiculae pollens loads were, non-lethally, collected from queens encountered by capturing them in an insect net and transferring them into a restraining device ("bee squeezer", @kearns2001natural).
-We then collected a single pollen load (i.e., from one leg) from the bee and then released it.
-At weekly intervals at each site, we also recorded the abundances of flowers visited by bumble bees within belt transects spread over the three vegetation types (0.5 x 40 m transects in each vegetation type, 60 m^2^ total area per site).
-
-## 2.3 | Pollen Morphological identification
-
-### 2.3.1 | Pollen Reference Library
-
-To qualitatively identify, and quantitatively note, the plant species present in corbiculae loads microscopy was used.
-A pollen reference library of fuchsin-jelly stained grains which may be present in corbiculae loads of slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@beattie1971technique, @brosi2013single).
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the RMBL Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  We used clustering techniques to supplement our subjective opinions of which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
-Ten readily discernible categorical traits were collected from each specimen in the image collection.
-These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (Appendix 6).
-This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
-
-### 2.3.2 | Preperation of Pollen Corbiculae Loads
-
-  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Circa* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
-The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in Appendix 4.
-  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained a unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects (Appendix 7 & 8).
-
-### 2.4 | Molecular Barcoding
-
-#### 2.4.1 | Species reference list
-
-#### 2.4.1.1 Spatial Analyses
-
-We first generated a short list of potential candidate species.
-We downloaded, from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022), all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
-To reduce the number of species to include in the reference sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.
-
-To minimize the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
-To account for the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
-The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
-
-#### 2.4.1.2 Distribution Modelling
-
-To determine which clades to include in the reference sequence database we used Species Distribution Modelling.
-We used all occurrence records from BIEN (n = 23,919) within a 50km border of the ecoregion, Omernik level 3, which includes the study area *(No.
-21 “Southern Rockies”)* to construct the species distribution model  (@omernik1987ecoregions).
-These records were copied into two, initially identical, sets, one for generating machine learning models (ML; Random Forest, and Boosted Regression Tree’s), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM) (@barbet2012selecting).
-Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble).
-No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with different biology and ecology (@elith2006novel, @qiao2015no).
-In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).
-
-We then generated 4,029 absence points, locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (@aim2019database).
-To achieve a larger absence data set, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
-For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced (@barbet2012selecting).
-To achieve this, we removed absence records inside of 10% of the mean sample value of any predictor variable the presence records; the required number of absence records were then randomly sampled.
-
-To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (Appendix 2).
-These publicly available data sets, were selected as they pertain to a wide range of variables interacting with plant physiology.
-For linear regression models these predictors underwent both *vifstep* (theta = 10, max observations = 12,500) and *vifcor* (theta = 0.7, max observations = 12,500) to detect highly correlated variables, and collinear features were removed leaving 16 variables (@usdm2014).
-
-Modelling: Random Forest and Boosted Regression Trees, were sub sampled with 30% test and two replicates each before weighted ensemble based on True Skill Statistics (tss) (@sdmPackage).
-Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage).
-TSS was chosen as the ensemble criterion as it has been shown to work across a wide range of species occurrences prevalence (@allouche2006assessing).
-The results of these models were extracted on a cell-by-cell basis to a polygon feature derived from a minimum-spanning tree which encompasses the study sites, and species from either ensemble with greater than 50% mean habitat suitability across all cells were considered present for further purposes (@prim1957minimum).
-
-A total of 535 species were modelled using Generalized Linear Models and Generalized Additive Models and 534 species were modelled using Random Forest and Boosted Regression Trees.
-To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (@gbifDL2021sdms, @aim2019database).
-Additional novel absence records were generated from the AIM data set to create a data set where each species has balanced presence and absences.
-Eleven or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
-
-### 2.4.2 | Temporal Analyses
-
-For assignment of reads to ecologically probabilistic species subsequent to BLAST, flowering time was used as a filter.
-To estimate the duration of dates in which plant species were flowering Weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not found in the nearest 5), and which were from herbarium records were included.
-To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in fall.
-Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
-Species with > 10 records had their Weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
-These estimates were compared to a long-term observational study of flowering phenology 1974-2012 (@caradonna2014shifts), and the floral abundance data from 2015, using Kendall's tau.
-
-### 2.5.2 | Barcode references library
-
-All lab work was carried out at The Daniel F. and Ada L. Rice Plant Conservation Science Center at the Chicago Botanic Garden, Glencoe, Illinois, U.S.A.
-
-#### 2.5.2.1 | Sampling Species for Barcoding
-
-Using five years (2015-2020) of observational data on *Bombus* queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years.
-In order to capture more variability inherit in the 353 loci we sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
-In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
-We also sequenced another 15 taxa of plants commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (Appendix 3).
-Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
-
-#### 2.5.2.2 | Plant Genomic DNA Extraction
-
-Plant genomic DNA was isolated from ~ 1 cm^2^ of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included two chloroform washes.
-DNA was quantified using a Nanodrop 2000 (Thermo Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer (Thermo Fisher Scientific).
-
-#### 2.5.2.3 | Pollen Genomic DNA Extraction
-
-To extract genomic DNA from pollen a CTAB based protocol was modified from Lahlamgiahi et al.
-and Guertler et al.
-(2014, 2014).
-A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added to the sampled followed by vortexing to allow dissolution of corbiculae.
-Pollen grains were then macerated with Kontes Pellet Pestles, and the tip of these washed with 130 µL of the SDS extraction buffer, samples were then incubated for 1 hour at 30°C.
-This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH.
-8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on a heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
-After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°C for 1 hour.
-Samples were then incubated overnight at 40°C.
-500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.
-
-To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to $\frac{2}{3}$ of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm.
-The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes.
-Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
-Pellets were dried at room temperature overnight before resuspension in nuclease free H~2~O.
-Extractions were assessed using a Nanodrop 2000 (Thermo Fisher Scientific) and Qubit fluorometer (Thermo Fisher Scientific).
-DNA extracts were then cleaned using 2:1 v./v.
-Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute ratio following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions.
-DNA was quantified using a Qubit fluorometer.
-
-#### 2.5.2.4 | Fragmentation, Library Preparation & Target Enrichment
-
-Sequence library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
-Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
-Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed using SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v.
-ethanol washes.
-The exception was the herbarium specimens which were not fragmented and only end repaired, with similar library preparation of all samples.
-Products were analysed on 4% agarose gels, and a Qubit fluorometer.
-Libraries were pooled and enriched with the Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence Capture Kit) by following the manufacturer's protocol and Brewer et al.
-2019.
-Sequencing was performed using an Illumina mi-Seq with 150-bp end reads, (NUSeq Core, Chicago, Illinois).
-
-###  2.6 | Computational Processes and Analyses. 
-####  2.6.1 | Reference Library Data Processing
-
-Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950).
-Contigs generated were mapped to a reference with HybPiper with using target files created by M353 (@johnson2016hybpiper, @mclay2021new).
-
-#### 2.6.2 | Sequence Identification
-
-A custom Kraken2 database was created by downloading representative species indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
-These sequences were processed in the same manner as our novel sequences.
-The Kraken2 database was built using default parameters.
-Kraken2 was run on sequences using default parameters (Appendix 9).
-Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken).
-FInally all reads which could be classified by these databases were passed to a local BLAST database.
-
-#### 2.6.3 | Identification of Sequence Matching Loci
-
-A local NCBI database was built using the same processed novel and downloaded sequences as the previous database (@camacho2009blast).
-
-### 2.7 | Integrated Observational, Molecular, and Palynological Corbiculae
-
-To precisely classify the contents of each corbiculae load the sequences classified by molecular methods were compared with the fieldwork which at a very fine resolution, recorded the presence and absence of species and their duration of flowering, and was interpreted ala the computer derived temporal and spatial data sets.
-The quantitative counts of grains from microscopy, were combined with the semi-quantitative sequencing results, to estimate the abundance of each identified species in each corbiculae load.
-
-To reclassify the sequence reads, these data were combined with the flora observation data, and mapped by genus.
-If more than one species in the genus was flowering at that time and site, than the reads were split evenly between the taxa.
-For sequence data which did not match at the genus level, a user subjectively scored them based on the species composition and phenological activity at each site, the queen interaction data, and pollen assignments.
-To estimate the abundance of each of these species in the corbiculae loads, these data were combined with the microscopy data.
-For each morphotype detected in pollen, and each classified sequence read which was not detected via microscopy, they were given a value of 0.5% to indicate their trace presences.
-When more than a single species belonged to a morphotype group in a single sample, the quantitative values from the morphological work were multiplied by the relative sequence abundance of each species in the load.
-All final compositions were standardized to a sum of 100%, by adding or subtracting the differences (induced by classifying records as 'trace') to all species with abundances > 1%.
-
-# 3 | RESULTS
-
-## 3.1 | Floral Observations
-
-```{r Floral Observations Summary Statistics }
-
-files <- list.files('../data', pattern = 'csv')
-
-flr_ranks <- read.csv(
-  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
-  select(site, doy, week, tot.obs.length, habitat, plant.species:abun.rank)
-obs_time <- read.csv(
-  file.path('../data', '/', files[grep('queen_observations', files)]) ) %>% 
-  select(site, week, tot.obs.length) %>% 
-  distinct(site, week, .keep_all = T)
-
-bee_obs <- read.csv(
-  file.path('../data', files[grep('queen_observations', files)]) ) %>% 
-  filter(caste == 'q') %>% 
-  mutate(species = gsub('\\..*$', '', species)) %>% 
-  select(site, doy, week, tot.obs.length, species, plant.species,
-         resource.coll:pollen.collected, fl.switch.1:fl.switch.2.resource.coll) %>% 
-  mutate(across(where(is.character), ~ na_if(., ''))) %>% 
-  pivot_longer(c(plant.species, fl.switch.1, fl.switch.2), names_to = 'observation', values_to = 'plant.species') %>% 
-  drop_na(plant.species) %>% 
-  group_by(site, week, species, plant.species) %>% 
-  count(name = 'interactions') %>% 
-  left_join(., obs_time, by = c("site", "week")) %>% 
-  mutate(period = case_when(
-    week %in% 3:5 ~ 'Early',
-    week %in% 6:8 ~ 'Mid',
-    week %in% 9:11 ~ 'Late'
-  ))
-
-no_plant_sp <- bee_obs %>% 
-  ungroup() %>% 
-  distinct(plant.species) %>%
-  nrow()
-
-s <- toString(format(as.Date(min(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # start obs
-e <- toString(format(as.Date(max(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # end obs
-
-t_b_wk <- bee_obs %>%  group_by(site, week, species) %>%  tally(interactions) 
-t_b_b <- bee_obs %>%  group_by(species) %>%  tally(interactions) %>%  filter(species != 'unknown') 
-t_b_wk_pl <- bee_obs %>%  group_by(site, week, plant.species) %>%  tally(interactions) 
-t_b_pl <- bee_obs %>%  group_by(plant.species) %>%  filter(species != 'unknown') %>%  tally(interactions) 
-t_spp_i <- bee_obs %>%  group_by(species) %>%  filter(species != 'unknown') %>%  distinct(plant.species) %>% 
-  count()
+# Table S1
+```{r}
+knitr::kable(df, caption = info, col.names = c('Variable', 'Data Source', 'Geographic Extent'), booktabs = TRUE) |> 
+  kableExtra::row_spec(0,bold=TRUE) |>
+  kableExtra::collapse_rows(1) |>
+  kableExtra::kable_styling(latex_options="scale_down") |>
+  kableExtra::add_footnote('CONUS; CONtinental United States')
 
@@
 
-In 2015 the six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
-A total of `r toString(sum(bee_obs$interactions))` overwintered queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
-Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week minimum = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, median = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (minimum = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, median = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
-The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, median = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
-
-A total of 66 corbiculae loads were collected from bees, 64 of them from queens.
-
-```{r echo = F}
-rm(bee_obs, flr_ranks, obs_time, t_b_b, t_b_wk, t_b_pl, t_b_wk_pl, e, s, files, t_spp_i)
-```
-
-```{r Most used resources by species, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
-
-corb <- read.csv('../data/Corbiculae_collections.csv', header = F) %>% 
-  mutate(sample.id = as.numeric(str_extract(V1, '[0-9]{1,2}')),
-         species = str_extract(V1, 'Bombus [^ ]+'),
-         caste = str_trim(str_extract(V1, '[:space:][q|w][:space:]'))) %>% 
-  filter(caste == 'q') %>% 
-  select(-V1, -caste)
-
-blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
-  rename(sample.id = Sample, Genus = GENUS, Family = FAMILY) %>% 
-  arrange(sample.id, -Prcnt_seqs) %>% 
-  inner_join(., corb, by = 'sample.id')  %>% 
-  distinct() %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
-  ungroup()
-
-# combine reassigned TAXON_NEW
-
-blst <- blst %>% 
-  group_by(TAXON_NEW, sample.id) %>% 
-  mutate(across(.cols = Seqs_pr_taxon:Prcnt_seqs, ~ sum(.x))) %>% 
-  distinct(TAXON_NEW, sample.id, .keep_all = T) %>%
-
-  # slice top ten records
-  group_by(sample.id) %>% 
-  slice_max(order_by = Prcnt_seqs, n = 10) %>% 
-  ungroup()
-
-blst_cnt <- blst %>% 
-  distinct(sample.id, .keep_all = T) %>% 
-  count(species)
-
-# Need to do some kind of weighted total proportion... 
-# the number of reads per sample, and the % of reads should be a factor
-gf_lkp2 <- read.csv('../data/genus2family_lkp.csv')
-gf_lkp <- blst %>% 
-  select(Genus, Family) %>% 
-  distinct(Genus, .keep_all = T) %>% 
-  bind_rows(gf_lkp2) %>% 
-  filter(Family != 'UNKNOWN')
-
-tb <- blst %>% 
-  left_join(., blst_cnt, by = 'species') %>% 
-  group_by(species, TAXON_NEW) %>% 
-  mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
-         species = str_remove(species, 'Bombus '))  %>% 
-  distinct(species, TAXON_NEW, .keep_all = T) %>% 
-  ungroup(TAXON_NEW) %>% 
-  slice_max(Mu_Seqs, n = 10) %>% 
-  select(plant.species = TAXON_NEW, species, n, Mu_Seqs, Genus, Family) %>% 
-  mutate(rank = 1:n()) %>% 
-  ungroup() %>% 
-  mutate(Family = case_when(
-    Genus %in% c('Trollius', 'Caltha', 'Aquilegia', 'Delphinium') ~ 'Ranunculaceae',
-    Genus == 'Hydrophyllum' ~ 'Hydrophyllaceae',
-    Genus == 'Epilobium' ~ 'Onagraceae',
-    Genus == 'Viola' ~ 'Violaceae',
-    Genus == 'Salix' ~ 'Saliceae',
-    Genus == 'Cynoglossum' ~ 'Boraginaceae', 
-    Genus == c('Arenaria', 'Odontostemma') ~ 'Caryophyllaceae',
-    Genus == 'Parnassia' ~ 'Celastraceae',
-    Genus %in% c('Erigeron', 'Symphyotrichum') ~ 'Asteraceae',
-    TRUE ~ Family 
-  ))
-
-obs <- read.csv('../data/yearEndInteractionsObservations.csv')[,2:5] %>% 
-  group_by(species) %>% 
-  arrange(-bySpp) %>% 
-  slice_head(n = 10)  %>% 
-  mutate(plant.species = case_when(
-    plant.species == 'Adenolinum.lewisii' ~  'Linum lewisii', 
-    plant.species == 'Lupinus.bakeri' ~ 'Lupinus sericeus' , 
-    plant.species == 'Dugaldia.hoopesii' ~ 'Hymenoxys hoopesii', 
-    TRUE ~ plant.species 
-  )) %>% 
-  #mutate(Family = if_else(Family == 'UNKNOWN', 'HYDROPHYLLACEAE', Family)) %>% 
-  separate(plant.species, into = 'Genus', extra = 'drop', remove = F) %>% 
-  mutate(plant.species = str_replace(plant.species, '[.]', " ")) %>% 
-  left_join(., gf_lkp, by = 'Genus') %>% 
-  distinct() %>% 
-  filter(species %in% tb$species) %>% 
-  mutate(rank = 1:n()) %>% 
-  ungroup()
-
-## Which bombus species are shared??
-
-# REMOVE KIRBIELLIUS, ONLY  A SINGLE FEEDING RECORD
-obs <- filter(obs, !species %in% c('kirbiellus', 'mixtus'))
-tb <- filter(tb, !species %in% c('kirbiellus', 'mixtus'))
-
-tblist <- split(tb, tb$species) %>% 
-  map(., ~ select(., plant.species, Genus, Family))
-oblist <- split(obs, obs$species) %>% 
-  map(., ~ select(., plant.species, Genus, Family))
-
-species <- purrr::map2_int(tblist, oblist,
-                ~ inner_join(.x, .y, by = 'plant.species') %>%
-                   nrow())
-
-genus <- purrr::map2_int(tblist, oblist,
-  ~ inner_join(.x, .y, by = 'Genus') %>%
-    distinct(plant.species.y) %>% nrow()
-)
-
-family <- purrr::map2_int(tblist, oblist,
-  ~ inner_join(.x, .y, by = 'Family') %>%
-    distinct(plant.species.y) %>%
-    nrow()
-)
-
-correct <- data.frame(Bombus.Species = names(species),
-           Species = species, Genus = genus, Family = family) %>% 
-          # ObsMax = map_int(oblist, ~ nrow(.))) %>% 
-  pivot_longer(!Bombus.Species, names_to = 'Observation') %>% 
-  mutate(Observation = factor(Observation,  levels = c("Species", "Genus", "Family")))
-
-bee_cols <- c('appositus' = '#A50026FF', 'bifarius' = '#FFA319FF', 
-              'californicus' = '#9632B8', 'flavifrons' = '#EFAE90',
-              'rufocinctus' = '#155F83FF', '#8F3931FF', 
-                'mixtus' = '#350E20FF', 'kirbiellus'  =  '#8074A8', 
-              'nevadensis' = '#C96731')
-
-ggplot(data=correct, aes(x=Observation, y=value,
-                         group = Bombus.Species, fill = Bombus.Species)) +
-  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
-              height = 0.05, width = 0.3) +
-  scale_fill_manual('Bee species', values = bee_cols) +
-  scale_color_manual('Bee species', values = bee_cols) +
-  theme_bw() +
-  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
-  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
-       y = 'Shared Observations', x= NULL) + 
-  theme(panel.grid.major.x = element_blank(),
-        panel.border = element_blank())
-
-n_corbiculae <- tb %>% 
-  distinct(species, .keep_all = T) %>% pull(n) 
-prop_obs <- obs %>% 
-  group_by(species) %>% 
-  slice_max(bySpp, n = 5, with_ties = F) %>% 
-  mutate(prop = sum(bySpp)/total) %>% 
-  distinct(species, .keep_all = T) %>% 
-  pull(prop)
-
-rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
-```
-
-## 3.2 | Spatial Analyses
-
-```{r Results for SA Modelling}
-
-files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
-ml <- files[grep( '_ml_', files)]
-ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
-linear <- files[grep('_glm_', files)]
-linear <- unique(sub('_glm.*$', '', linear))
-
-# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
-# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
-
-rm(files, ml, linear)
-```
-
-```{r Results for SA Logistic Regression Metrics}
-
-logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
-  filter(occurrence == 0)
-
-dat <- read.csv('../data/evaluation_table.csv')[,2:3]
-data.frame(Metric  = dat[1:5,1],
-                   Value   = dat[1:5,2],
-                   Metric1 = dat[6:10,1],
-                   Value1  = dat[6:10,2]
-                   ) %>% 
-  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
-    col.names = c('Metric', 'Value', 'Metric', 'Value'),
-    booktabs = T) #%>% 
-  #kableExtra::kable_styling(position = "float_right")
-```
-
-```{r Results for SA Logistic Regression Contingency table}
-read.csv('../data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
-    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
-    booktabs = T) %>%
-    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
-
-rm(logR_dat, dat)
-```
-
-The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
-Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
-
-Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
-
-## 3.2 | Microscopic Pollen identification
-
-```{r compare microscopic to molecular Pollen ID, eval = T}
-
-# are mertensia and Asteraceae 1 & 2 detected in all Pollen samples where the bee was COLLECTED ON THEM ?
-
-pmorph <- read.csv('../data/morpho_pollen.csv') %>% 
-  filter(!sample.id %in% c(45, 54))
-blst_seqs <- distinct(blst, sample.id)
-
-pmorph <- filter(pmorph, sample.id %in% blst$sample.id) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prop = sum/sum(sum))
-
-mertensia_samples <- pmorph %>% 
-  filter(morphotype == 'MERTENSIA') %>% 
-  select(sample.id, morphotype, Prop, sum)
-
-mertensia_samples <- filter(mertensia_samples, sample.id %in% blst_seqs$sample.id)
-
-blst_mertensia <- blst %>% 
-  filter(Family == 'Boraginaceae') %>% 
-  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_seqs = sum(Prcnt_seqs),
-         morphotype = 'MERTENSIA') %>% 
-  distinct(sample.id, .keep_all = T)
-
-mertensia_detected <- mertensia_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0)) 
- # sample 26 - very strange !!!! this needs investigation - ALSO THE ERICOID IS LIKELY ONAGRAD !?
-
-blst_aster <- blst %>% 
-  filter(Family == 'Asteraceae') %>% 
-  group_by(sample.id) %>% 
-  distinct(sample.id, .keep_all = T)
-
-# ASTER 2 HERE...
-aster2_samples <- pmorph %>% 
-  filter(morphotype == 'ASTERACEAE_2')
-
-blst_aster2 <- filter(blst_aster, Genus %in% 
-                            c('Taraxacum', 'Helianthella', 'Scabrethia')) %>% 
-  mutate(morphotype = 'ASTERACEAE_2')
-aster2_detected <- aster2_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster2$sample.id, 1, 0))
-
-# ASTER 1 HERE... ? WERE ALL OF THESE SPECIES PRUNED... ?
-aster1_samples <- pmorph %>% 
-  filter(morphotype == 'ASTERACEAE_1')
-
-blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 
-                                                'Scabrethia') ) %>% 
-  mutate(morphotype = 'ASTERACEAE_1')
-aster1_detected <- aster1_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster1$sample.id, 1, 0))
-
-# ALL HIGH NUMBER ASTERACEAE
-
-aster_samples <- pmorph %>% 
-  filter(str_detect(morphotype, 'ASTER.*'))
-aster_detected <- aster_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster$sample.id, 1, 0))
-
-aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
-bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
-cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
-dd <- prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
-ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)
-
-# Semi-quantitative accord between sequence % and Grain Count
-palyn_cnt <- bind_rows(aster1_samples, aster2_samples, mertensia_samples) %>% 
-  select(morphotype, countProp = Prop)
-
-pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia) %>% 
-  select(morphotype, sequenceProp = Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
-  left_join(., palyn_cnt, by = c('sample.id', 'morphotype')) %>% 
-  drop_na()
-
-rm(mertensia_samples, blst_aster, blst_aster1, blst_aster2, aster1_samples, 
-   aster2_samples, palyn_cnt, blst_mertensia, aster_samples)
-
-grains_counted <- pmorph %>% 
-  group_by(sample.id) %>% 
-  summarize(grains_counted = sum(sum)) %>% 
-  pull(grains_counted)
-
-```
-
-Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into (Appendix 6).
-From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we identified substantial amounts of the abundance and morphotype richness of the samples (morphotype richness, $\bar{x}$ = 4.5, median = 4, min = 1, max = 9) (Appendix  7 & 8).
-The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, median = `r  median(grains_counted)`).
-
-```{r}
-rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected, grains_counted)
-```
-
-```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
-The innermost ring indicates every genus which Queen Bee's were observed to visit.
-The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
-The outermost ring indicates that sequence data were available for at least a member of that genus.
-Branch colors follow APG 4."}
-knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
-```
-
-### 3.3 | Metabarcoding Pollen Identification
-
-### 3.3.2 | Temporal Analysis
-
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
-
-Only 58 of these 388 species (n = 34.6, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
-Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
-There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
-
-Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
-Due to methodological differences, the peak flowering was not compared, and due to the low performance of attempts to model 'duration' in the previous step it was also not compared.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).
-
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
-# these are the tenth and nintey  percent non-parametric confidence intervals. 
-```
-
-```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
-```
-
-```{sh Get number of raw and trimmed reads for metagenome samples, eval = F}
-
-cd ~/sequenceData/Original_zip
-files=$(find -type f -name "[0-9]_S[0-9]*_paired.fastq.gz") # doesnt support {2}
-files1=$(find -type f -name '[0-9][0-9]_S[0-9]*_paired.fastq.gz')
-files="${files} ${files1}"
-echo $files
-# seqs per each sample , need divide by 4 on import
-wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
-
-cd ../trimmed_metagenomic
-files=$(find  -maxdepth 1 -name '*gz')
-wc -l  $files  > '/hdd/yun master/MEE_manuscript/data/trimmed_reads_metagenome_seq.txt'
-```
-
-```{r Summary statistics of sequence numbers}
-
-raw_seqs <- read.delim('../data/raw_reads_metagenome_seq.txt', header = F) %>% 
-  mutate(V1 = str_trim(V1)) %>% 
-  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
-  mutate(seq = round(as.numeric(seq)/4),
-         sample = str_extract(sampleFR, '.*L002_')) %>% 
-  filter(sampleFR != 'total') %>% 
-  group_by(sample) %>% 
-  mutate(totalReads = sum(seq)) %>% 
-  distinct(totalReads) %>% 
-  pull(totalReads)
-
-trimmed_seqs <- read.delim('../data/trimmed_reads_metagenome_seq.txt', header = F) %>% 
-  mutate(V1 = str_trim(V1)) %>% 
-  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
-  mutate(seq = round(as.numeric(seq)/4),
-         sample = str_extract(sampleFR, '.*paired')) %>% 
-  filter(sampleFR != 'total') %>% 
-  group_by(sample) %>% 
-  mutate(totalReads = sum(seq)) %>% 
-  distinct(totalReads) %>% 
-  pull(totalReads)
-
-krak <- read.csv('../data/kraken_proc_reads.csv')
-brack <- read.csv('../data/bracken_proc_reads.csv')
-blast <- read.csv('../data/blast_proc_reads.csv')
-loci <- read.csv('../data/reads_per_loci.csv') %>% group_by(Sample)
-
-no_loci <- loci %>% group_by(Sample) %>% count(name = 'no_loci')
-no_reads <- loci %>% group_by(loci) %>% summarize(readsLoci = sum(n))
-```
-
-## 3.3.1 | Molecular analysis of corbiculae loads
-
-The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
-The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
-Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
-The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
-
-After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
-`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
-
-```{r Read Post Processing}
-
-pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') ) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 1)))
-
-knitr::kable(pp_seqs_t, align = "c", 
-             col.names = c('Condition', 'No. Class.', 'Prcnt. Class.', 'Total Seqs', 'Rank'),
-             caption = "Post classification of Sequences via Taxonomy and Ecology,\ntop 15 most abundant reads") %>% 
-  kableExtra::kable_styling(
-      latex_options = "striped",
-      stripe_color = "gray!10")
-
-seqs_sp <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records']), 1)
-seqs_ge <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records']), 1)
-seqs_fa <- round(sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records']), 1)
-
-sp_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs']),1)
-sp_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs']),1)
-fa_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs']),1)
-
-byhand <- blst %>% 
-  filter(TAXON_NEW != TAXON_NEW)
-
-j <- toString(paste0(names(table(byhand$Condition)), ": ",
-                     as.vector (table(byhand$Condition))))
-gen <- byhand[byhand$TAXON_NEW==c('Salix'),]
-spp <- byhand[byhand$TAXON_NEW==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
-```
-
-The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
-A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix 15).
-Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
-Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species representing `r sp_total_seqs`% of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family.
-
-Of the `r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were all assigned via post-processing conditions (`r j`, Appendix 15).
-These were manually assigned to a variety of ranks, occasionally to genus - `r nrow(gen)`, and species - `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
-
-To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
-The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less the Heliantheae Aliiance (Sunflower Tribe) (Boraginaceae grains `r toString(aa)`.
-Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`.
-
-```{r clean sequence numbers from environment}
-rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
-    sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
-```
-
-```{r Correlation between Number of Grains and Number of Sequences, fig.cap = 'Relationship between morphological count data and sequence reads', out.height="50%", out.width="50%"}
-
-pol_p <- pollen_props %>% 
-  mutate(sequenceProp = sequenceProp/100, 
-         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
-
-ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
-# ct[["p.value"]] #too small to write out. tau > 0.4 strong relationship
-model <-  lm(adjSeqProp ~ countProp, data = pol_p)
-
-# png('../../thesis_talk/graphics/plots/CorrelationCountedSequences.png', )
-par(pty="s")
-plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-      xlab = 'Grains Counted', ylab = 'Classified Sequences',
-      pch=18, slty = 3, gap = 0.025)
-axis(1)
-axis(2)
-abline(model)
-text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
-text(0.3, 0.8, "p < 0.001") 
-# dev.off()
-
-# rm( newdat)
-```
-
-To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
-In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
-
-The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly linear, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
-This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing).
-There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
-
-```{r}
-rm(pol_p, ct)
-```
-
-```{r Co-occurrence of splash drop sequences, eval = F}
-library(CooccurrenceAffinity)
-
-e <- blst %>% 
-  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
-                       'Celastraceae')) %>% 
-  select(sample.id, Genus) %>% 
-  mutate(Present = 1) %>% 
-  group_by(Genus) %>% 
-  distinct() %>% 
-  pivot_wider(names_from = Genus, values_from = Present) %>% 
-  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
-  column_to_rownames('sample.id')
-
-myout <- affinity(data = e, row.or.col = "col", squarematrix = c("all"))
-plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
-```
-
-A number of suspected false positives were returned by the sequence classified, in all instances these were species present in low abundances at the field station which were also returned with a common species in the same family.
-To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala Jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
-Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
-sp., *Thalictrum* L.
-spp., *Trollius* L.
-sp., *Aquilegia* L.
-spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
-were recorded.
-A similar relationship was observed in the Hydrophyllaceae R.Br.
-with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
-species were present.
-The flower size  and sparsity of *Nemophila breviflora* A.
-Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
-The floral morphology and orientation of flower of *Thalictrum*  spp.
-also makes them unlikely to be visited, and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
-
-## 3.6 | Integrated Observational, Molecular, and Palynological Results
-
-```{r compare results, out.width="50%", fig.cap= 'Comparision of Accuracy between the initial output data from BLAST, and these same data subjected to the post-classification process which removes surrogate, and temporally restricted species'}
-
-blast <- read.csv('../data/tally_blast_classifications.csv') %>% 
-  group_by(Sample) %>% 
-  slice_max(Prcnt_seqs, n = 10) %>% 
-  select(Sample, Taxon = taxid, Percent_Sample = Prcnt_seqs) %>% 
-  mutate(
-    Taxon = str_replace(Taxon, 'Arenaria', 'Eremogone'),
-    Genus = str_remove(str_extract(Taxon, '^.* '), ' '),
-    Method = 'BLAST')
-
-automated <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
-  group_by(Sample, TAXON_NEW) %>%
-  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
-  drop_na() %>% 
-  distinct(Sample, TAXON_NEW, .keep_all = T) %>% 
-  ungroup(TAXON_NEW) %>% 
-  slice_max(Prcnt_seqs, n = 10) %>% 
-  select(Sample, Genus = GENUS, Taxon = TAXON_NEW, Percent_Sample = Prcnt_seqs) %>% 
-  mutate(across(where(is.character), ~ str_replace(., 'Arenaria', 'Eremogone')),
-         Method = 'Automated')
-
-expert <- read.csv('../data/Fully_Integrated_Corbiculae.csv') %>% 
-  rename(Sample = sample.id) %>% 
-  mutate(across(where(is.character), ~ str_trim(.)), 
-         Method = 'Expert') %>% 
-  drop_na() %>% 
-  relocate()
-
-genera2query <- na.omit(unique(c(blast$Genus, automated$Genus, expert$Genus))) 
-genera2query <- genera2query[genera2query != 'Ericaceae'] 
-genera2query <- taxizedb::name2taxid(genera2query, out_type = 'summary')
-
-genera_out <- taxizedb::classification(genera2query$id, db='ncbi', rank = 'family') 
-genera_out <- map(genera_out, ~ 
-                    data.frame( 
-                      'Family' = filter(.x, rank == "family") %>% select(name), 
-                      'Genus' = filter(.x, rank == 'genus') %>%  select(name) 
-                      )) %>% 
-  bind_rows() %>% 
-  rename(Family = 1, Genus = 2)
-
-genera_out <- filter(genera_out, !Family %in% c('Mertensiidae', 'Hesperiidae'))
-
-all_classifications <- bind_rows(
-  automated, blast, expert
-) %>% 
-  left_join(., genera_out, by = 'Genus') %>% 
-  mutate(Rank = case_when(
-    is.na(Family) ~ 'Family',
-    Taxon == Family ~ 'Family',
-    Taxon == Genus ~ 'Genus',
-    TRUE ~ 'Species'
-  ), 
-  Family = if_else(Taxon == 'Ericaceae', 'Eriaceae', Family))
-
-rm(genera2query, genera_out, automated, blast, expert)
-
-expert <- all_classifications %>% filter(Method == 'Expert')
-blast <- all_classifications %>% 
-  filter(Method == 'BLAST' & Sample %in% expert$Sample)
-automated <- all_classifications %>% 
-  filter(Method == 'Automated' & Sample %in% expert$Sample)
-
-## now determine which proportion of records are correct/incorrect
-
-expert_g <- expert %>% 
-  select(Sample, Genus) %>% 
-  distinct()
-expert_f <- expert %>% 
-  select(Sample, Family) %>% 
-  distinct()
-
-# CORRECT RECORDS FIRST
-
-true_species_blast <- left_join(expert, blast, by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method.y) %>% 
-  summarize(Species_T = n())
-true_genus_blast <- left_join(expert_g, blast, by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_T = n())
-true_family_blast <- inner_join(expert_f, blast, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_T = n())
-
-blast_correct <- left_join(true_species_blast, true_genus_blast, by = 'Sample') %>% 
-  left_join(., true_family_blast, by = 'Sample')
-
-rm(true_species_blast, true_genus_blast, true_family_blast)
-
-true_species_computer <- left_join(expert, 
-                                    filter(automated, Rank == 'Species'),
-                                    by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method.y) %>% 
-  summarize(Species_T = n())
-
-true_genus_computer <- left_join(expert_g,  filter(automated, Rank != 'Family'),
-                                    by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_T = n())
-
-true_family_computer <- left_join(expert_f, automated, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_T = n())
-
-computer_correct <- left_join(true_species_computer, true_genus_computer, by = 'Sample') %>% 
-  left_join(., true_family_computer, by = 'Sample')
-
-rm(true_species_computer, true_genus_computer, true_family_computer)
-
-# FALSE records second
-
-false_species_blast <- anti_join(blast, expert,  by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  summarize(Species_F = n())
-
-false_genus_blast <- anti_join(blast, expert_g, by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_F = n())
-
-false_family_blast <-anti_join(blast, expert_f, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_F = n())
-
-blast_incorrect <- left_join(false_species_blast, false_genus_blast, by = 'Sample') %>% 
-  left_join(., false_family_blast, by = 'Sample')
-
-blast_results <- left_join(blast_correct, blast_incorrect, by = 'Sample') %>% 
-  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))
-
-rm(false_species_blast, false_genus_blast, false_family_blast, blast_correct, blast_incorrect)
-
-false_species_computer <- anti_join(automated, expert,  by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  summarize(Species_F = n())
-
-false_genus_computer <- anti_join(automated, expert_g, by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_F = n())
-
-false_family_computer <-anti_join(automated, expert_f, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_F = n())
-
-computer_incorrect <- left_join(false_species_computer, false_genus_computer, by = 'Sample') %>% 
-  left_join(., false_family_computer, by = 'Sample')
-
-computer_results <- left_join(computer_correct, computer_incorrect, by = 'Sample') %>% 
-  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))
-
-rm(false_species_computer, false_genus_compuer, false_family_computer, 
-   computer_correct, computer_incorrect)
-
-### calculate accuracy
-
-# number of assignments
-
-automated_cnts <- automated %>% 
-  group_by(Sample) %>% 
-  count(Rank) %>% 
-  pivot_wider(names_from = Rank, values_from = n, names_prefix = 'NO_') %>% 
-  mutate(Method = 'COMP')
-blast_cnts <- blast %>% 
-  group_by(Sample) %>% 
-  count(Rank) %>% 
-  pivot_wider(names_from = Rank, values_from = n, names_prefix = 'NO_') %>% 
-  mutate(Method = 'BLAST')
-cnts <- bind_rows(automated_cnts, blast_cnts)
-
-spp_count <- expert %>% 
-  group_by(Sample) %>% 
-  count(name = 'Spp_No')
-gen_count <- expert_g %>% 
-  group_by(Sample) %>% 
-  count(name = 'Gen_No')
-fam_count <- expert_f %>% 
-  group_by(Sample) %>% 
-  count(name = 'Fam_No')
-
-ACC_sens_Spec <- bind_rows(
-  mutate(blast_results, Method = 'BLAST'),
-  mutate(computer_results, Method  = 'COMP')
-) %>% 
-  left_join(., spp_count) %>% 
-  left_join(., gen_count) %>% 
-  left_join(., fam_count) %>% 
-  left_join(., cnts, by = c('Sample', 'Method')) %>% 
-  rowwise() %>% 
-  mutate(
-    Species_false = NO_Species - Species_T,
-
-    Sp_Acc = (Species_T / NO_Species) * 100,
-    Gen_Acc = (Genera_T / sum(NO_Genus, NO_Species, na.rm = T) ) * 100,
-
-    Sp_Sens = (Species_T)  / (Species_T + (Spp_No - Species_T)) * 100
-  )
-
-# Spp_No - Species_match is my TRUE NEGATIVE
-# species_false is MY FALSE POSITIVE
-
-wt_res <- wilcox.test(
-  pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'COMP', 'Sp_Acc']) , 
-  pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'BLAST', 'Sp_Acc']),
-  alternative = "greater"
-  )
-
-my_means <- ggpubr::compare_means(Sp_Acc ~ Method,  data = ACC_sens_Spec)
-my_comparisons <- my_means %>% 
-    nest(groups = c(group1, group2)) %>% 
-    pull(groups) %>% 
-    map(., as.character)
-sample_sizes <- ACC_sens_Spec %>% 
-    dplyr::group_by(Method) %>%  
-    dplyr::tally() %>% 
-    dplyr::mutate(n = paste0('n = ', n)) 
-min_v <- dplyr::summarise(ACC_sens_Spec, mean_mpg = floor(min(Sp_Acc))) |>
-    pull() |> min()
-
-ggplot(ACC_sens_Spec, aes(y = Sp_Acc,  x= Method, color = Method),  
-                        alpha = 0.5) + 
-  stat_boxplot(notch = T, notchwidth = 0.75, 
-               varwidth = T, 
-               outlier.shape = F, outlier.alpha = 0.8, outlier.colour = 'black') +
-  geom_jitter(width = 0.1, shape = 1) +
-
-  geom_text(data = sample_sizes,
-            aes(Method, Inf, label = n), color = 'black', 
-            vjust = "inward", size = 4, 
-            y = min_v * 0.5) +
-  ggpubr::stat_compare_means(comparisons = my_comparisons, 
-                             aes(label = ..p.signif..),
-                             tip.length = 0, vjust = 0.25, size = 4) +
-
-  expand_limits(y= min_v - 5) +
-  theme_bw() +
-  scale_colour_manual(values = c('#F4AFB4', '#94A89A')) +
-  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
-  scale_x_discrete(labels = c('BLAST', 'Post Classification')) + 
-  labs(y = 'Accuracy', title = 'Accuracy of "naive" BLAST and post classification process') +
-  theme(legend.position = 'none', 
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-```{r Most Used Floral Resources Table}
-
-fam <- all_classifications %>% 
-  ungroup() %>% 
-  select(Family, Genus) %>% 
-  distinct()
-
-expert %>% 
-  group_by(Taxon) %>% 
-  add_count(name = 'No.Corb') %>% 
-  mutate(median_percent_corb = round( mean(Percent_Sample), 1), 
-         Taxon = str_remove(Taxon, '^.* '), 
-         Taxon = if_else(Taxon %in% c('Salix', 'Erigeron'), 'sp.', Taxon)) %>% 
-  select(Genus, Species = Taxon, Percent_Sample, No.Corb, median_percent_corb) %>% 
-  left_join(., fam, by = 'Genus') %>% 
-  select(-Percent_Sample) %>% 
-  relocate(Family, .before = Genus) %>% 
-  filter(No.Corb > 4) %>% 
-  distinct() %>% 
-  arrange(Family, Genus)  %>% 
-  knitr::kable(
-    escape = F,
-    align = c('l', 'l', 'l', 'c', 'c'),
-    caption = 'Plant species detected in five or more corbiculae loads',
-    col.names = c('Family', 'Genus', 'Species', 'No. Samples', 'Mean Prop.\\textsuperscript{a}')
-  ) %>% 
-  add_footnote(
-    'The mean only calculated across the samples where the species was detected') %>% 
-  collapse_rows()
-
-```
-
-```{r}
-rm(n_corbiculae, prop_obs)
-```
-
-# 4 | DISCUSSION
-
-We have demonstrated how the Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context (@johnson2019universal, @hollingsworth2016telling).
-This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history guided fundamental science and land management.
-The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g.
-*Mertensia*, *Lupinus* L.), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic).
-We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas.
-By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.
-
-  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area.
-However, we had difficulties in evaluating our predictions in an operational context.
-We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local).
-While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult.
-We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data.
-Further given the, size of the minimum spanning tree  which we extracted points to, a formal floristic inventory would still be a time intensive process.
-Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding.
-We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.
-
-In regards to the case study, our results indicate the overall information gleaned from observations of queen Bumble Bee foraging and analysis of pollen records are largely congruent.
-Relaxing concerns regarding differences between the broad insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-In general when interaction networks are considered at coarse levels, such as the duration of a season, our perceptions regarding the generality of interactions at smaller time scales may be inflated relative to the actualized interactions within them, e.g.
-a week (@caradonna2020temporal).
-These results indicate a possibility that at even finer levels *Bombus* display high amounts of floral fidelity within foraging bouts, an observation which implies that part of the reason for the high efficiency of *Bombus* as a pollinator might partially be related to their lack of movement of hetero-specific pollen (@brosi2013single, @ashman2013toward, @galloni2008visitor, @brosi2016pollinator).
-The mechanisms behind this observed fidelity are likely related to pollen nutritional values, specifically high concentrations of protein, and the absence of particular amino acids required for larval development in other flower more commonly used by workers (@genissel2002influence, @tasei2008nutritive, @goulson2005causes, @goulson2008diet, @hanley2008breeding).
-
-```{r Table of Limitations Faced}
-read.csv('../data/Limitations_Table.csv') |>
-  knitr::kable(caption = 'Limitations Faced and Possible Solutions',
-             col.names = c('Method Component', 'Limitations', 'Paths Forward')) |>
-  kable_styling(full_width = TRUE)
-```
-
-Also regarding the case study, some foraging preferences of *Bombus*, at this field site and across several localities globally emerge.
-These suggest the need for land managers to maintain relatively high amounts of members of the plant families Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
-Numerous historic  practices reduce the ability of many landscapes to support stable populations of *Bombus*.
-Historic livestock grazing involved the removal of many species known to have compounds toxic to cattle.
-In particular, the removal of locoweeds (*Astragalus* & *Oxytropis*) and larkspurs (*Delphinium*) were common across public lands administered by the U.S.
-Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
-Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks, processes which lowered water tables and reduced the extent of stream-associated wetlands and the mesic meadows fringes which provide habitat for many species of 'tall' *Mertensia* (e.g.
-*M.
-ciliata*), to an extent *Delphinium barbeyi*  and many species of native *Trifolium* (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).
-Fire suppression resulted in the succession of many Aspen (*Populus tremuloides*) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effecting habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
-Finally the effects of Nitrogen deposition, given the West's rapidly growing population still pose adverse effects on the abundance of species of Fabaceae at urban-rural interfaces (see @stevens2018atmospheric, @fenn2003ecological).
-Current solutions to the above issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing.
-The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias*, using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus*) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
-
-```{r}
-read.csv('../data/Errors_Table.csv') |>
-  knitr::kable(caption = 'Errors faced with components of the framework',
-             col.names = c('Component', 'Example', 'Correction')) 
-```
-
-  We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical).
-To reduce the effects of a low population density of botanists on the maintenance of and production of Flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
-In this proof-of-concept example we performed several iterations of modelling runs, and several approaches (i.e.
-the 'linear models', and the 'machine learning'), which took notable amounts of compute power.
-We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicit process to reduce spatial autocorrelation.
-However, given the time required to perform all aspects of a study, even our amount of computation was negligible.
-Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate.
-In certain scenarios modelling of predicted species via more formally tailored S(tacked)-SDM or J(oint)-SDM approaches may be beneficial (@wilkinson2021defining, @pinto2021predicting, @schmitt2017ssdm).
-
-  Tandem to the lack of continued expertise required to generate and maintain species lists, is the expertise required to continue tracking when major phenological events occur in many plant species at relatively fine scales or under novel climates.
-Knowledge of these events is currently limited to general time periods of only a handful of phenological events and groups of organisms (e.g.
-flowering initiation, or trees) (@prather2004implications, @li2016responses).
-While many programs and initiatives exist to collect phenological information on subsets of easily identifiable charismatic species to detect major trends in phenology, these capture only a subset of the extent diversity (@betancourt2005implementing, @havens2007chicago).
-In many instances it appears that while landscapes respond similarly to environmental variables which predict phenological responses, that individual species vary widely in their responses to similar environmental cues, or respond to different cues (@augspurger2020concordance, @xie2015deciduous, @xie2018predicting, @caradonna2014shifts).
-As can be seen here, predictions of when a single, major phenological event occurs is already data limited.
-A more promising approach for the tropics may lay in utilizing circular statistics (@park2022herbarium).
-
-  The nearly complete Plant and Fungal Tree of Life (PAFTOL) will provide a comprehensive phylogenetic backbone of the entire plant kingdom, and the inclusion of A353 probes with lineage specific probe sets is common in producing massive genetic datasets (@baker2021exploring).
-We predict that the A353 probes which it is utilizing to work nearly immediately for DNA barcoding of whole plant material, and that more elaborate validation studies in controlled metabarcoding settings, utilizing existing experimental designs, will have favorable results (@bell2017applying, @bell2019quantitative, @bell2021comparing, @lamb2019quantitative).
-In particular the harvesting of loci with more variation in certain lineages, and or with more variable flanking regions, will prove promising for identifying closely related plant material.
-We suspect that conserved reaches of genes resulted in the high amounts of reads in somewhat obscure species.
-Given that the A353 loci are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high, without continual PCR enrichment is possible; this would align with relatively high efficacy of WGS (@lang2019genome, @peel2019semi, @bell2021comparing).
-Recent evidence indicates that the potential for identifying nearly cryptic taxa and even infra-specific inference, of either whole plant material, and perhaps in metagenomic context are possible (@ottenlips2021resolving, @wenzell2021incomplete, Loke et al.
-in prep,  @slimp2021potential, @beck2021palmer).
-We further believe that in synthetic phylogenetic trees - with incorporation of NGS backbones - will allow in automatic reassignment of reads as a function of phylogenetic distance with measures of uncertainty (@hinchliff2015synthesis, @smith2018constructing, @baker2021PAFTOL).
-
-# 5 | CONCLUSION
-
-We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.
-Herein we crudely utilized this thinking via binary outcomes, should a species predicted be predicted present or not?
-Is it unequivocally flowering or not?
-Myriad data show biological systems and ecological interactions have more variance than can be reasonably discretely parsed.
-We expect that within a framework developed from our preliminary works studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
-the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems.
-We hope that the A353 probes as tools for metabarcoding play a role in these endeavors.
-
-**AUTHOR CONTRIBUTIONS:**  R.C.B conducted botanical collections, conducted all molecular lab work, lead all analyses, and writing.
-J.E.O conceived, designed, and conducted ecological fieldwork, assisted with analyses, and writing.
-P.J.C conducted ecological fieldwork, assisted with ecological analyses, and writing.
-E.J.W.
-prepared, imaged, and collected trait data on pollen reference slides, and assisted with analysis of trait data and writing a dichotomous key.
-S.T.
-assisted with spatial analyses and writing.
-J.B.F.
-conceived, and designed lab work, analyses, and integration of approaches, assisted with writing, and secured funding for molecular work.
-
-```{r}
-a <- pull(read.csv('../data/herbaria_4_acknowledgements.csv'))
-```
-
-**ACKNOWLEDGMENTS:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
-Pat Herendeen for assistance with virtually all aspects of preparing pollen vouchers and the identification process.
-Hilary Noble, Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
-Ian Breckheimer for sharing the SDM predictor variables.
-We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles (Rick) Williams at Ray J.
-Davis (IDS), (B)Ernie Nelson at Rocky Mountain (RM); and the collectors: D.
-Knoke, L.
-Brummer, J.
-Boyd, C.
-Davidson, I.
-Gilman, M.
-Kirkpatrick, S.
-McCauley, J.
-Smith, K.
-Taylor, & C.
-Williams.
-David Giblin & Mare Nazaire for sharing relevant sections of an advanced draft of FNA V.
-15.
-The Bureau of Land Management is thanked as many plant specimens were collected by R.C.B as a partner or contractor to the agency; Sarah Burnett and Lauren Price are thanked for sharing AIM data.
-Sanda and New England Biotech are gratefully acknowledged for technical support and generously sharing samples.
-T.C.H.
-Cole for sharing the Angiosperm Phylogeny 4 colour palette.
-The Program in Plant Biology and Conservation is thanked for funding.
-The holdings of the following herbaria were essential for this project: `r toString(a)`.
-
-**CONFLICT OF INTERESTS** The authors declare no conflicts of interest.
-
-**PEER REVIEW** The peer review history for this document is available at ...
-
-**DATA AVAILABILITY STATEMENT** The queries required to download all data used in this project are located in... All novel sequencing data are located at NCBI...
-
-**ORCID**
-
-Paul CaraDonna https://orcid.org/0000-0003-3517-9090  
-Jeremie Fant https://orcid.org/0000-0001-9276-1111  
-Jane Ogilvie https://orcid.org/0000-0001-8546-0417  
-Sophie Taddeo https://orcid.org/0000-0002-7789-1417
-
-# References
-
-# Supporting
-
 \newpage
 
-Appendix 1 - Site Maps
+# Site Info
 
@@
 
-Appendix 2 - Species Distribution Models Predictors
+# Table S2 - Models used for Species Distribution Model Ensembles
 
-Layer   LM                     Description                              Source
-----   ----      ------------------------------------------    -------------------------
-1.      N              Mean annual cloudiness - MODIS              Wilson et al. 2016
-2.      Y            Cloudiness seasonality 1 - MODIS              Wilson et al. 2016 
-3.      N            Cloudiness seasonality 2 - MODIS              Wilson et al. 2016 
-4.      Y            Cloudiness seasonality 3 - MODIS              Wilson et al. 2016
-5.      N           Beginning of the frost-free period                 Wang et al.
-6.      N              Climatic moisture deficit                       Wang et al.
-7.      N               Degree-days above 5C                           Wang et al.
-8.      N              Mean annual precipitation                       Wang et al.
-9.      Y           Mean annual precipitation as snow                  Wang et al.
-10.     Y               Temperature seasonality                        Wang et al.
+The two machine learning models utilize Ensemble learning.
+
+__Ensemble learning__ utilizes many sets of trees, each tree being composed of many binary decisions, to create a single model.
+Each independent variable ( - or *feature*) may become a node on the tree - i.e.
+a location on the tree where a binary decision will move towards a predicted outcome.
+Each of the decision tree models which ensemble learning utilizes is a weak model, each of which may suffer due to high variance or bias, but which produce better outcomes than would be expected via chance.
+When ensembled these models generate a strong model, a model which should have more appropriately balanced variance and bias and predicts outcomes which are more strongly correlated with the expected values than the individual weak models.
+
+*__Random Forest (RF)__* the training data are continually bootstrap re-sampled, in combination with random subsets of features, to create nodes which attempt to optimally predict a known outcome.
+A large number of trees are then aggregated, via the most common predictions, to generate a final classification prediction tree.
+Each individual prediction tree is generated independently of the others.
+
+*__Boosted Regression Tree (BRT)__* (or Gradient Boosted tree) An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
+All response data and predictor variables are kept available to all trees.
+
+*__Bias__* predictions from an algorithm are systematically in error due to being prejudiced for or against certain results, due to assumptions during learning.
+
+*__Variance__* errors in models due to an over-reliance and sensitivity of training to outliers in training data.
+
+In general, Random Forest  models have high bias and low variance, where boosted regressions trees have lower bias and higher variance.
+Theoretically, the weaknesses and strengths of bootstrap aggregation (bagging) as implemented by Random Forests are supplemented by the boosting.
+
+\newpage
+
+Table S3 - Species Distribution Models Predictors
+
+Layer   LM                     Description                              Name                    Source
+----   ----      ------------------------------------------    -------------------------     ----------------
+1.      N              Mean annual cloudiness - MODIS           Cloud Cover (EarthEnv)       Wilson et al. 2016
+2.      Y            Cloudiness seasonality 1 - MODIS           Cloud Cover (EarthEnv)       Wilson et al. 2016 
+3.      N            Cloudiness seasonality 2 - MODIS           Cloud Cover (EarthEnv)       Wilson et al. 2016 
+4.      Y            Cloudiness seasonality 3 - MODIS           Cloud Cover (EarthEnv)       Wilson et al. 2016
+5.      N           Beginning of the frost-free period                 ClimateNA               Wang et al. 2016
+6.      N              Climatic moisture deficit                       ClimateNA               Wang et al. 2016
+7.      N               Degree-days above 5C                           ClimateNA               Wang et al. 2016
+8.      N              Mean annual precipitation                       ClimateNA               Wang et al. 2016
+9.      Y           Mean annual precipitation as snow                  ClimateNA               Wang et al. 2016
+10.     Y               Temperature seasonality                        ClimateNA               Wang et al. 2016
 11.     Y         2015 Percent Grass/Herbaceous cover - MODIS           (MOD44B)
-12.     Y          2015 Percent Tree cover from Landsat 7/8              (GLCF)
-13.     Y         Soil probability of bedrock (R Horizon)               SoilGrids
-14.     N           Soil organic carbon (Tonnes / ha)                   SoilGrids
-15.     N                Surface soil pH in H~2~O                       SoilGrids
-16.     Y               Surface soil percent sand                       SoilGrids
-17.     Y                     Soil USDA class                           SoilGrids
-18.     N                   Topographic elevation                      EarthEnv DEM
-19.     Y           Topographic elevation, moving window.              EarthEnv DEM
-20.     Y               Topographic percent slope                      EarthEnv DEM 
-21.     Y               Topographic wetness index                      EarthEnv DEM 
-22.     Y                  Topographic aspect                          EarthEnv DEM
+12.     Y          2015 Percent Tree cover from Landsat 7/8              (GLCC)
+13.     Y         Soil probability of bedrock (R Horizon)               SoilGrids              Hengl et al. 2017
+14.     N           Soil organic carbon (Tonnes / ha)                   SoilGrids              Hengl et al. 2017
+15.     N                Surface soil pH in H~2~O                       SoilGrids              Hengl et al. 2017
+16.     Y               Surface soil percent sand                       SoilGrids              Hengl et al. 2017
+17.     Y                     Soil USDA class                           SoilGrids              Hengl et al. 2017
+18.     N                   Topographic elevation               Topography (EarthEnv)         Amatulli et al. 2018   
+19.     Y           Topographic elevation, moving window.       Topography (EarthEnv)         Amatulli et al. 2018   
+20.     Y               Topographic percent slope               Topography (EarthEnv)         Amatulli et al. 2018    
+21.     Y               Topographic wetness index               Topography (EarthEnv)         Amatulli et al. 2018    
+22.     Y                  Topographic aspect                   Topography (EarthEnv)         Amatulli et al. 2018   
 23.     Y         Annual potential solar radiation computed               r.sun
-24.     N        Estimated actual (w/-cloud) solar radiation      r.sun / Wilson et al. 2016
-25.     Y        Log-transformed distance to surface water      Global Surface Water Explorer
-26.     Y                Percent surface water                  Global Surface Water Explorer 
-----   ----     -------------------------------------------    ------------------------------
+24.     N        Estimated actual (w/-cloud) solar radiation              r.sun                Wilson et al. 2016
+25.     Y        Log-transformed distance to surface water      Global Surface Water Explorer   Pekel et al. 2016
+26.     Y                Percent surface water                  Global Surface Water Explorer   Pekel et al. 2016
+----   ----     -------------------------------------------    ------------------------------   ----------------
+
+Amatulli, G., Domisch, S., Tuanmu, M.-N., Parmentier, B., Ranipeta, A., Malczyk, J., and Jetz, W.
+(2018) A suite of global, cross-scale topographic variables for environmental and biodiversity modeling.
+Scientific Data volume 5, Article number: 180040.
+DOI: doi:10.1038/sdata.2018.40.
+
+Hengl T, Mendes de Jesus J, Heuvelink GBM, Ruiperez Gonzalez M, Kilibarda M, Blagotić A, et al.
+(2017) SoilGrids250m: Global gridded soil information based on machine learning.
+PLoS ONE 12(2): e0169748.
+https://doi.org/10.1371/journal.pone.0169748
+
+Pekel, JF., Cottam, A., Gorelick, N. et al. High-resolution mapping of global surface water and its long-term changes. Nature 540, 418–422 (2016). https://doi.org/10.1038/nature20584
+
+Wang T, Hamann A, Spittlehouse D, Carroll C (2016) Locally Downscaled and Spatially Customizable Climate Data for Historical and Future Periods for North America.
+PLoS ONE 11(6): e0156720.
+https://doi.org/10.1371/journal.pone.0156720
+
+Wilson AM, Jetz W (2016) Remotely Sensed High-Resolution Global Cloud Dynamics for Predicting Ecosystem and Biodiversity Distributions.
+PLoS Biol 14(3): e1002415.
+https://doi.org/10.1371/journal.pbio.1002415
 
@@
 
-\newpage
-
-Appendix 4 - All Pollen Reference Slides Used to Establish Morphotypes
-
-```{r Pollen Reference Slides 1, echo = F}
-knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-1.pdf')
-```
-
-\newpage
-
-```{r Pollen Reference Slides 2, echo = F}
-knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-2.pdf')
-```
-
-\newpage
-
-```{r Pollen Reference Slides 3, echo = F}
-knitr::include_graphics('../graphics/assorted/pollen_slide_table_reduced-3.pdf')
-```
-
-\newpage
-
-Appendix 5 - Pollen Dendrogram
-```{r}
-
-```
-
-\newpage
-
 Appendix 6 - Pollen Key
@@
 knitr::include_graphics('../graphics/assorted/RMBL_pollen_key-2.pdf')
-```
-
-\newpage
-
-Appendix 7 - Pollen Morphotype Richness Rarefaction Curves
-
-```{r Morphotype Richness, echo= F,  out.width = '95%'}
-knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
-```
-
-\newpage
-
-Appendix 8 - Pollen Morphotype Abundance Rarefaction Curves
-
-```{r Morphotype Abundance, echo = F, out.width = '99%'}
-knitr::include_graphics('../graphics/plots/SppAbundance.png')
 ```
@@
 
-Appendix 10 - All Species in the Sequence Databases (con't)
-
 ```{r Seq DB 3, echo = F}
@@
 \newpage
-
-Appendix 10 - All Species in the Sequence Databases (con't)
 
@@
 \newpage
-
-Appendix 13 - Models used for Species Distribution Model Ensembles
-
-The two machine learning models utilize Ensemble learning.
-
-__Ensemble learning__ utilizes many sets of trees, each tree being composed of many binary decisions, to create a single model.
-Each independent variable ( - or *feature*) may become a node on the tree - i.e.
-a location on the tree where a binary decision will move towards a predicted outcome.
-Each of the decision tree models which ensemble learning utilizes is a weak model, each of which may suffer due to high variance or bias, but which produce better outcomes than would be expected via chance.
-When ensembled these models generate a strong model, a model which should have more appropriately balanced variance and bias and predicts outcomes which are more strongly correlated with the expected values than the individual weak models.
-
-*__Random Forest (RF)__* the training data are continually bootstrap re-sampled, in combination with random subsets of features, to create nodes which attempt to optimally predict a known outcome.
-A large number of trees are then aggregated, via the most common predictions, to generate a final classification prediction tree.
-Each individual prediction tree is generated independently of the others.
-
-*__Boosted Regression Tree (BRT)__* (or Gradient Boosted tree) An initial tree is grown, and all other trees are derived sequentially from it, as each new tree is grown the errors in responses from the last tree are weighed more heavily so that the model focuses on selecting dependent variables which refine predictions.
-All response data and predictor variables are kept available to all trees.
-
-*__Bias__* predictions from an algorithm are systematically in error due to being prejudiced for or against certain results, due to assumptions during learning.
-
-*__Variance__* errors in models due to an over-reliance and sensitivity of training to outliers in training data.
-
-In general, Random Forest  models have high bias and low variance, where boosted regressions trees have lower bias and higher variance.
-Theoretically, the weaknesses and strengths of bootstrap aggregation (bagging) as implemented by Random Forests are supplemented by the boosting.
-
-\newpage
-
-APPENDIX 14 - Time Spent Generating Species Distribution Models
-
-```{r}
-knitr::include_graphics('../graphics/assorted/individual_model_runs.pdf')
-```
-
-\newpage
-
-APPENDIX 15 - Review Classified Reads and Reassign
-
-```{r}
-knitr::include_graphics('../graphics/tables/reclassify_reads_table.pdf')
-```
-
-\newpage
-
-APPENDIX 16 - Tips and tricks for implementing plant metagenomic sequencing
-
-```{r}
-knitr::include_graphics('../graphics/assorted/overall_metagenomic_workflow.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Morphological Workflow}
-knitr::include_graphics('../graphics/assorted/microscopy_workflow.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Molecular Workflow}
-knitr::include_graphics('../graphics/assorted/molecular_workflow.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Extraction Buffer}
-knitr::include_graphics('../graphics/assorted/pollen_ctab.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Extraction Buffer back}
-knitr::include_graphics('../graphics/assorted/pollen_ctab-1.pdf')
-```
-
-\newpage
-
-Appendix 17 Samples Per Bee
-
-```{r Samples per Bee}
-knitr::include_graphics('../graphics/tables/bombus_samples.pdf')
-```
-
-\newpage
-
-Appendix 18 - Overview of the Whole Process
-
-```{r Overview of entire process, out.height="90%"}
-knitr::include_graphics('../graphics/plots/flowchart.png')
-```
-
-\newpage
-
-# References

### 2024-02-11 script/SupplementaryMaterial.Rmd
subset supplementals

--- 2023-05-27 script/manuscript_MEE.Rmd
@@
 ---
-title: Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees
-author:  |
-    |  Sophie Taddeo$^1$$^,$$^2$, Paul J. CaraDonna$^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant$^1$$^,$$^2$  
-    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
-    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
-    |  $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA
-    | Reed Clark Benkendorf$^1$$^,$$^2$^[Correspondence: reedbenkendorf2021@u.northwestern.edu], Jane E. Ogilive$^3$, Emily J. Woodworth$^1$$^,$$^2$, 
-abstract:  |  
-  1) DNA Barcoding has been successful for the rapid identification of species in complex ecological assemblages.
-However, metabarcoding in the plant kingdom has been limited due to a lack of universal gene regions that works across all taxa, limiting the applications of eDNA and metagenomics in ecology.
-  2) To circumvent these limitations, we propose using more holistic approach which combines a multiple gene approach that incorporates existing plant occurrence databases, species distribution models, and phenology analyses to generate a list of likely candidate species to increase accuracy.
-Such an effort can help guide plant sample collection and library construction, particularly for projects with limited resources or local taxonomic knowledge.
-In addition, we demonstrate that building a custom candidate plant list based on known occurrences, ecological requirements, and phenology can improve computing efficiency, speed, and accuracy of metabarcoding.
-We tested this approach to identify the plant species contributing to the corbiculae pollen loads of bumble bees.
-To validate the efficiency of this approach we compare results of the DNA barcoding to expert-led pollen identification and field observations of plant-pollinator interactions that generated these pollen loads.
-  3) We show that the Angiosperms 353 probes, developed for phylogenomics, and which are currently being used in the largest ever plant systematic endeavor, offer significant promise to metagenomic approaches around the globe.
-The DNA barcoding of bumble bee corbiculae pollen loads was most accurate when combined with knowledge of what plant species were flowering in the plant community when they were collected.
-  4) Supplementing DNA barcoding data with ecological context is most accurate and powerful.  
-keywords: |
-  DNA baroding, metagenomics, hyb-seq, Angiosperms353, plant metagenomics, bombus queens, subalpine, species distribution model
-output:
-  word_document: default
-  toc: no
-  pdf_document: default
-csl: "../citations/methods-in-ecology-and-evolution.csl"
-bibliography: ../citations/citations.bib
-link-citations: yes
-fig_caption: yes
-always_allow_html: yes
-header-includes:
-- \usepackage{endfloat}
-- \usepackage{setspace}\doublespacing
-- \usepackage{lineno}
-- \linenumbers
+title: "Supplemental Information"
+author: "steppe"
+date: "2024-02-11"
+output: pdf_document
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, warning = F, message = F)
-```
-
-```{r Required Libraries, message=F, warning=F, results='hide'}
-library(tidyverse)
-library(bookdown)
-library(knitr)
-library(kableExtra)
-#devtools::install_github("kpmainali/CooccurrenceAffinity")
-suppressWarnings(library(kableExtra))
-pNumC <- function(...) prettyNum(...,big.mark = ',')
-
-prcnt_detect <- function(x, Lprop, Hprop){
-  if(missing(Lprop)){Lprop <- 0}
-
-  y <- x[x$Prop > Lprop & x$Prop < Hprop,]; total <- nrow(y)
-  detected <- round((nrow(y[y$DETECTED == 1,])/total) * 100, 1)
-  Mdn <- round(median(y$Prop),3)
-
-  r <- paste0('were detected in ',  detected, 
-                '% of samples where the proportion of target grains were between ', 
-                Lprop, '-', Hprop,  ' (n = ', total, ' Mdn = ', Mdn, ')')
-  return(r)
-}
-```
-
-# 1 | INTRODUCTION
-
-Large scale species loss (@joppa2011biodiversity; @pimm2014biodiversity) and biotic homogenization (@olden2004ecological, @wang2021biotic), and the impacts of these processes on ecosystem functions and human well-being (@cardinale2012biodiversity), have inspired numerous calls for a more consistent monitoring of ecosystems and their diversity (@pereira2006towards; @pecl2017biodiversity).
-Monitoring ecological communities structure and interaction will be critical to informing and prioritizing conservation efforts (REF), particularly considering the large proportion of know species currently threaten or likely to be threaten in the near future (@pimm2014biodiversity; @pereira2010scenarios).
-However, ecosystem monitoring remains inconsistent in time and space (@yoccoz2001monitoring) as it requires substantial resources, thereby leaving several regions, ecosystems, and even species under-observed (@pereira2012global; @collen2008tropical; @meyer2015global; @ruete2015displaying).
-
-Species interactions are a critical component of ecosystem stability and therefore biodiversity conservation (Soltis et al.
-(2019), @futuyma2009macroevolution, Voje et al.
-(2015), Weber et al.
-(2017), @hembry2020ecological).
-Species interactions, which span from positive (e.g., symbiosis, mutualistic) to negative (e.g., predation, competition), shape species distribution (@bascompte2009mutualistic; Wisz et al.
-2013) and tolerance to environmental stressors (REF).
-They impact ecosystem stability and diversity (@agrawal2007filling, @valiente2015beyond, @bascompte2006asymmetric, and the relationships between biodiversity and ecosystem functions (@duffy2007functional).
-However, species interactions are challenging to observe (REFs).
-They are typically measured through shorter-term observational studies yet are highly variable in time and space (@liu2022environmental).
-Studying species interactions also requires extensive taxonomic expertise, especially those from diverse clades (@hebert2003biological), further challenging their frequent assessment.
-As a result, few ecological networks have been fully mapped (REFs).
-Improving our capacity to efficiently study species interactions consistently across several ecosystems, landscapes, and management context is thus pivotal to enacting conservation interventions (@tylianakis2010conservation).
-This is particularly urgent as increased biological introductions and climate change are impacting species interactions at a rapid rate, thus calling for more frequent and consistent monitoring (REFs).
-
-  Recently barcoding (the identification of a sample from a single organism *e.g.* a piece of leaf), and metabarcoding (the identification of a sample containing a mix of organisms *e.g.* soil), have shown considerable promise in all Kingdoms of Life (@ruppert2019past).
-The ability to identify species from a fragments of organisms (e.g.
-hair, scat, leaf tissue) increases our ability to understand the interactions of not only entire ecosystems but also a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g.
-those for restoration processes (Banerjee et al.
-(2022), Johnson et al.
-(2023)).
-With plants the identification of members of certain clades to species using barcoding has been successful in some groups (@kress2017plant), whereas many other clades have proven more problematic (@liu2014identification, @china2011comparative, @coissac2012bioinformatic).
-Metabarcoding incurs additional challenges to those which exist for the currently implemented barcodes (@li2015plant, @kress2007two, @cbol2009dna, @coissac2012bioinformatic).
-Particular limitations of the traditional high copy number barcodes (e.g.
-ITS2, *rbcL*, *matK*, *trnH-psbA*) include the utilization their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant).
-Currently, most plant metabarcoding endeavors only allow the identification of material to the level of family or genus.
-
-```{r Applications of Metabarcoding}
-read.csv('../data/Applications_Table.csv') |>
-  knitr::kable(caption = 'Applications of Plant Metabarcoding',
-             col.names = c('Application', 'Example', 'Citations')) |>
-  kable_styling(full_width = TRUE)
-```
-
-  Currently the largest plant systematic endeavor ever undertaken, by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).
-This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000+ represented species, using the popular Angiosperms353 (A353) probes (@baker2021PAFTOL, @johnson2019universal).
-Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute more species by 2030 (@cheng2018tenkp).
-These publicly available data serve to provide a phylogenetically comprehensive backbone for plant metabarcoding.
-Similar projects, focused on whole genome sequencing of all organisms in certain geographic areas, such as the ‘Darwin Tree of Life’, which aims to sequence all described taxa in Britain and Ireland, will provide geographically dense data sets to seed projects globally (@darwin2022project, @lewin2022biogenome).
-These novel genomic data sets will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling).
-However, the application of metabarcoding still faces challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
-
-  To address these challenges, we are proposing the generation of a smaller, and more appropriate reference list for identifying candidate taxa from barcode data.
-By reducing the list of likely species in a study site it can increase the accuracy and efficiency of metabarcoding results in plants, which will better inform data collection for projects with limited access to taxonomic expertise.
-Specifically, we propose using current distribution data to generate a list of candidate species relevant to the study area and its ecological characteristics (@bell2022plants).
-Initially such a species list can be leveraged to identify a target collections list of plant tissues to be included in genetic studies.
-It can subsequently help to reduce the size of a reference sequence database, which increases computation speed and efficiency, allowing for the use of genomic data on personal computers.
-This approach can significantly reduce processing time, increasing  project efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.
-We test this metagenomic and informatics approach by examining the plant species composition within corbiculae pollen loads of bumble bees as a case study.
-We use DNA barcoding to identify the plant species in corbiculae pollen loads collected from overwintered queen bumble bees, and compare this approach with direct observations of bee flower visits and microscopically-examined pollen slides, which has shown some incongruency in several floral visitation networks involving smaller bodied fauna (@olesen2011missing, @barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-The assessment of the plant species in pollen is a commonly desired, across several potential applications, despite numerous potential complications (@pornon2017dna, @2017applying, @sickel2015increased, @bell2019quantitative, @suchan2019pollen, @johnson2021airborne).
-
-```{r Current Bottlenecks in Plant Metabarcoding}
-read.csv('../data/Bottlenecks_Table.csv') |>
-  knitr::kable(caption = 'Current Issues Facing Plant Metagenomics',
-             col.names = c('Issue', 'Our Approach', 'Possible Advances')) |>
-  kable_styling(full_width = TRUE)
-```
-
-```{r Filter Diagram, out.width = "35%", fig.cap="Simplified Conceptual Diagram of three approaches leading towards classification of sequencing results, and the number of species associated with them in our area.
-The upper three boxes indicate a common coarse approach, assuming one has a digitized Flora, which is not always the case.
-The center two boxes indicate the computational approach illustrated here.
-The final box indicates the use of the expert field data in the case study.
-The stem of the final applies to both Computational and Expert Survey results, and should be thought of as using time ala chromatography, in the post-classification process."}
-knitr::include_graphics('../graphics/plots/filter_diagram.png')
-```
-
-# 2 | METHODS
-
-To improve metabarcoding reliability and efficiency, we suggest creating a regional list of candidate species using digital collections gleaned from herbaria, survey work, and community science  (Figure 1).
-This list can further be refined using species distribution models and temporal filtering to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area.
-The final list is then used to inform collection of plant samples to create a library and inform metabarcoding .
-We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations.
-
-## 2.1 Case Study: Bee-Flower Observations and Pollen Load Collection
-
-Bee and flower observations and bee corbiculae pollen collection was conducted around the Rocky Mountain Biological Laboratory (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.), Colorado, USA (Appendix 1 for site information).
-The area is characterized by subalpine meadow vegetation communities.
-Pollinator observations of Bombus Latreille spp.
-(Apidae Latreille) were conducted from May 29th – July 23rd of 2015 in six study sites as a part of a larger study (described in Ogilvie and CaraDonna 2022).
-Observations of Bombus foraging took place for one hour at each field site, with equal time spent searching for bee in the major vegetation types (dry, and wet meadows, and aspen forest).
-Corbiculae pollens loads were, non-lethally, collected from queens encountered by capturing them in an insect net and transferring them into a restraining device ("bee squeezer", @kearns2001natural).
-We then collected a single pollen load (i.e., from one leg) from the bee and then released it.
-At weekly intervals at each site, we also recorded the abundances of flowers visited by bumble bees within belt transects spread over the three vegetation types (0.5 x 40 m transects in each vegetation type, 60 m^2^ total area per site).
-
-## 2.3 | Pollen Morphological identification
-
-### 2.3.1 | Pollen Reference Library
-
-To qualitatively identify, and quantitatively note, the plant species present in corbiculae loads microscopy was used.
-A pollen reference library of fuchsin-jelly stained grains which may be present in corbiculae loads of slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@beattie1971technique, @brosi2013single).
-Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the RMBL Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V.
-4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered (Appendix 3).
-  We used clustering techniques to supplement our subjective opinions of which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes.
-Ten readily discernible categorical traits were collected from each specimen in the image collection.
-These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022).
-Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared (Appendix 6).
-This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.
-
-### 2.3.2 | Preperation of Pollen Corbiculae Loads
-
-  To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples.
-*Circa* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA).
-The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide.
-Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in Appendix 4.
-  To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used.
-In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained a unique morphotype than all grains in that transect were also identified and counted.
-Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage).
-Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects (Appendix 7 & 8).
-
-### 2.4 | Molecular Barcoding
-
-#### 2.4.1 | Species reference list
-
-#### 2.4.1.1 Spatial Analyses
-
-We first generated a short list of potential candidate species.
-We downloaded, from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022), all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area.
-To reduce the number of species to include in the reference sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.
-
-To minimize the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded.
-To account for the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set.
-The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.
-
-#### 2.4.1.2 Distribution Modelling
-
-To determine which clades to include in the reference sequence database we used Species Distribution Modelling.
-We used all occurrence records from BIEN (n = 23,919) within a 50km border of the ecoregion, Omernik level 3, which includes the study area *(No.
-21 “Southern Rockies”)* to construct the species distribution model  (@omernik1987ecoregions).
-These records were copied into two, initially identical, sets, one for generating machine learning models (ML; Random Forest, and Boosted Regression Tree’s), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM) (@barbet2012selecting).
-Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble).
-No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with different biology and ecology (@elith2006novel, @qiao2015no).
-In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).
-
-We then generated 4,029 absence points, locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (@aim2019database).
-To achieve a larger absence data set, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record.
-For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced (@barbet2012selecting).
-To achieve this, we removed absence records inside of 10% of the mean sample value of any predictor variable the presence records; the required number of absence records were then randomly sampled.
-
-To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv) (Appendix 2).
-These publicly available data sets, were selected as they pertain to a wide range of variables interacting with plant physiology.
-For linear regression models these predictors underwent both *vifstep* (theta = 10, max observations = 12,500) and *vifcor* (theta = 0.7, max observations = 12,500) to detect highly correlated variables, and collinear features were removed leaving 16 variables (@usdm2014).
-
-Modelling: Random Forest and Boosted Regression Trees, were sub sampled with 30% test and two replicates each before weighted ensemble based on True Skill Statistics (tss) (@sdmPackage).
-Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage).
-TSS was chosen as the ensemble criterion as it has been shown to work across a wide range of species occurrences prevalence (@allouche2006assessing).
-The results of these models were extracted on a cell-by-cell basis to a polygon feature derived from a minimum-spanning tree which encompasses the study sites, and species from either ensemble with greater than 50% mean habitat suitability across all cells were considered present for further purposes (@prim1957minimum).
-
-A total of 535 species were modelled using Generalized Linear Models and Generalized Additive Models and 534 species were modelled using Random Forest and Boosted Regression Trees.
-To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (@gbifDL2021sdms, @aim2019database).
-Additional novel absence records were generated from the AIM data set to create a data set where each species has balanced presence and absences.
-Eleven or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).
-
-### 2.4.2 | Temporal Analyses
-
-For assignment of reads to ecologically probabilistic species subsequent to BLAST, flowering time was used as a filter.
-To estimate the duration of dates in which plant species were flowering Weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical).
-Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not found in the nearest 5), and which were from herbarium records were included.
-To remove temporally irrelevant herbarium records, i.e.
-material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in fall.
-Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed.
-Species with > 10 records had their Weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.
-These estimates were compared to a long-term observational study of flowering phenology 1974-2012 (@caradonna2014shifts), and the floral abundance data from 2015, using Kendall's tau.
-
-### 2.5.2 | Barcode references library
-
-All lab work was carried out at The Daniel F. and Ada L. Rice Plant Conservation Science Center at the Chicago Botanic Garden, Glencoe, Illinois, U.S.A.
-
-#### 2.5.2.1 | Sampling Species for Barcoding
-
-Using five years (2015-2020) of observational data on *Bombus* queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years.
-In order to capture more variability inherit in the 353 loci we sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
-In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa.
-We also sequenced another 15 taxa of plants commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set  (Appendix 3).
-Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).
-
-#### 2.5.2.2 | Plant Genomic DNA Extraction
-
-Plant genomic DNA was isolated from ~ 1 cm^2^ of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included two chloroform washes.
-DNA was quantified using a Nanodrop 2000 (Thermo Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer (Thermo Fisher Scientific).
-
-#### 2.5.2.3 | Pollen Genomic DNA Extraction
-
-To extract genomic DNA from pollen a CTAB based protocol was modified from Lahlamgiahi et al.
-and Guertler et al.
-(2014, 2014).
-A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added to the sampled followed by vortexing to allow dissolution of corbiculae.
-Pollen grains were then macerated with Kontes Pellet Pestles, and the tip of these washed with 130 µL of the SDS extraction buffer, samples were then incubated for 1 hour at 30°C.
-This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH.
-8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on a heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C.
-After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°C for 1 hour.
-Samples were then incubated overnight at 40°C.
-500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.
-
-To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to $\frac{2}{3}$ of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm.
-The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes.
-Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
-Pellets were dried at room temperature overnight before resuspension in nuclease free H~2~O.
-Extractions were assessed using a Nanodrop 2000 (Thermo Fisher Scientific) and Qubit fluorometer (Thermo Fisher Scientific).
-DNA extracts were then cleaned using 2:1 v./v.
-Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute ratio following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions.
-DNA was quantified using a Qubit fluorometer.
-
-#### 2.5.2.4 | Fragmentation, Library Preparation & Target Enrichment
-
-Sequence library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation.
-Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
-Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed using SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v.
-ethanol washes.
-The exception was the herbarium specimens which were not fragmented and only end repaired, with similar library preparation of all samples.
-Products were analysed on 4% agarose gels, and a Qubit fluorometer.
-Libraries were pooled and enriched with the Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence Capture Kit) by following the manufacturer's protocol and Brewer et al.
-2019.
-Sequencing was performed using an Illumina mi-Seq with 150-bp end reads, (NUSeq Core, Chicago, Illinois).
-
-###  2.6 | Computational Processes and Analyses. 
-####  2.6.1 | Reference Library Data Processing
-
-Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950).
-Contigs generated were mapped to a reference with HybPiper with using target files created by M353 (@johnson2016hybpiper, @mclay2021new).
-
-#### 2.6.2 | Sequence Identification
-
-A custom Kraken2 database was created by downloading representative species indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved).
-These sequences were processed in the same manner as our novel sequences.
-The Kraken2 database was built using default parameters.
-Kraken2 was run on sequences using default parameters (Appendix 9).
-Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken).
-FInally all reads which could be classified by these databases were passed to a local BLAST database.
-
-#### 2.6.3 | Identification of Sequence Matching Loci
-
-A local NCBI database was built using the same processed novel and downloaded sequences as the previous database (@camacho2009blast).
-
-### 2.7 | Integrated Observational, Molecular, and Palynological Corbiculae
-
-To precisely classify the contents of each corbiculae load the sequences classified by molecular methods were compared with the fieldwork which at a very fine resolution, recorded the presence and absence of species and their duration of flowering, and was interpreted ala the computer derived temporal and spatial data sets.
-The quantitative counts of grains from microscopy, were combined with the semi-quantitative sequencing results, to estimate the abundance of each identified species in each corbiculae load.
-
-To reclassify the sequence reads, these data were combined with the flora observation data, and mapped by genus.
-If more than one species in the genus was flowering at that time and site, than the reads were split evenly between the taxa.
-For sequence data which did not match at the genus level, a user subjectively scored them based on the species composition and phenological activity at each site, the queen interaction data, and pollen assignments.
-To estimate the abundance of each of these species in the corbiculae loads, these data were combined with the microscopy data.
-For each morphotype detected in pollen, and each classified sequence read which was not detected via microscopy, they were given a value of 0.5% to indicate their trace presences.
-When more than a single species belonged to a morphotype group in a single sample, the quantitative values from the morphological work were multiplied by the relative sequence abundance of each species in the load.
-All final compositions were standardized to a sum of 100%, by adding or subtracting the differences (induced by classifying records as 'trace') to all species with abundances > 1%.
-
-# 3 | RESULTS
-
-## 3.1 | Floral Observations
-
-```{r Floral Observations Summary Statistics }
-
-files <- list.files('../data', pattern = 'csv')
-
-flr_ranks <- read.csv(
-  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
-  select(site, doy, week, tot.obs.length, habitat, plant.species:abun.rank)
-obs_time <- read.csv(
-  file.path('../data', '/', files[grep('queen_observations', files)]) ) %>% 
-  select(site, week, tot.obs.length) %>% 
-  distinct(site, week, .keep_all = T)
-
-bee_obs <- read.csv(
-  file.path('../data', files[grep('queen_observations', files)]) ) %>% 
-  filter(caste == 'q') %>% 
-  mutate(species = gsub('\\..*$', '', species)) %>% 
-  select(site, doy, week, tot.obs.length, species, plant.species,
-         resource.coll:pollen.collected, fl.switch.1:fl.switch.2.resource.coll) %>% 
-  mutate(across(where(is.character), ~ na_if(., ''))) %>% 
-  pivot_longer(c(plant.species, fl.switch.1, fl.switch.2), names_to = 'observation', values_to = 'plant.species') %>% 
-  drop_na(plant.species) %>% 
-  group_by(site, week, species, plant.species) %>% 
-  count(name = 'interactions') %>% 
-  left_join(., obs_time, by = c("site", "week")) %>% 
-  mutate(period = case_when(
-    week %in% 3:5 ~ 'Early',
-    week %in% 6:8 ~ 'Mid',
-    week %in% 9:11 ~ 'Late'
-  ))
-
-no_plant_sp <- bee_obs %>% 
-  ungroup() %>% 
-  distinct(plant.species) %>%
-  nrow()
-
-s <- toString(format(as.Date(min(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # start obs
-e <- toString(format(as.Date(max(flr_ranks$doy), origin = "2016-01-01"), "%B %d")) # end obs
-
-t_b_wk <- bee_obs %>%  group_by(site, week, species) %>%  tally(interactions) 
-t_b_b <- bee_obs %>%  group_by(species) %>%  tally(interactions) %>%  filter(species != 'unknown') 
-t_b_wk_pl <- bee_obs %>%  group_by(site, week, plant.species) %>%  tally(interactions) 
-t_b_pl <- bee_obs %>%  group_by(plant.species) %>%  filter(species != 'unknown') %>%  tally(interactions) 
-t_spp_i <- bee_obs %>%  group_by(species) %>%  filter(species != 'unknown') %>%  distinct(plant.species) %>% 
-  count()
-
-```
-
-In 2015 the six sites were surveyed once weekly from `r s`-`r e` for a total of `r toString(sum(obs_time$tot.obs.length)/60)` hours.
-A total of `r toString(sum(bee_obs$interactions))` overwintered queen-pollen foraging interactions were observed (range per bee species by week range = `r toString(min(t_b_wk$n))`-`r toString(max(t_b_wk$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk$n), 2))`, median = `r toString(median(t_b_wk$n))`), with a range of total observed interactions per bee species across this time period (minimum = `r toString(min(t_b_b$n))`, $\bar{x}$  = `r toString(round(mean(t_b_b$n), 2))`, median = `r toString(median(t_b_b$n))`, max = `r toString(max(t_b_b$n))`).
-Plants varied widely in the number of interactions which they partook in with each species of bee (range per plant species by week minimum = `r toString(min(t_b_wk_pl$n))` - `r toString(max(t_b_wk_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_wk_pl$n), 2))`, median = `r toString(median(t_b_wk_pl$n))`), with a range of total observed interactions per plant species over this time period (minimum = `r toString(min(t_b_pl$n))`, $\bar{x}$  = `r toString(round(mean(t_b_pl$n), 2))`, median = `r toString(median(t_b_pl$n))`, max = `r toString(max(t_b_pl$n))`).
-The number of plant species which bees were observed interacting with varied more narrowly (range = `r toString(min(t_spp_i$n))` - `r toString(max(t_spp_i$n))`, $\bar{x}$  = `r toString(round(mean(t_spp_i$n), 2))`, median = `r toString(median(t_spp_i$n))`), interactions were observed with a total of `r no_plant_sp` plant species.
-
-A total of 66 corbiculae loads were collected from bees, 64 of them from queens.
-
-```{r echo = F}
-rm(bee_obs, flr_ranks, obs_time, t_b_b, t_b_wk, t_b_pl, t_b_wk_pl, e, s, files, t_spp_i)
-```
-
-```{r Most used resources by species, fig.cap = 'Number of the ten most commonly visited plants which are also in the top ten most common sequences', out.width = "45%"}
-
-corb <- read.csv('../data/Corbiculae_collections.csv', header = F) %>% 
-  mutate(sample.id = as.numeric(str_extract(V1, '[0-9]{1,2}')),
-         species = str_extract(V1, 'Bombus [^ ]+'),
-         caste = str_trim(str_extract(V1, '[:space:][q|w][:space:]'))) %>% 
-  filter(caste == 'q') %>% 
-  select(-V1, -caste)
-
-blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
-  rename(sample.id = Sample, Genus = GENUS, Family = FAMILY) %>% 
-  arrange(sample.id, -Prcnt_seqs) %>% 
-  inner_join(., corb, by = 'sample.id')  %>% 
-  distinct() %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_Seqs_Class = sum(Prcnt_seqs)) %>% 
-  ungroup()
-
-# combine reassigned TAXON_NEW
-
-blst <- blst %>% 
-  group_by(TAXON_NEW, sample.id) %>% 
-  mutate(across(.cols = Seqs_pr_taxon:Prcnt_seqs, ~ sum(.x))) %>% 
-  distinct(TAXON_NEW, sample.id, .keep_all = T) %>%
-
-  # slice top ten records
-  group_by(sample.id) %>% 
-  slice_max(order_by = Prcnt_seqs, n = 10) %>% 
-  ungroup()
-
-blst_cnt <- blst %>% 
-  distinct(sample.id, .keep_all = T) %>% 
-  count(species)
-
-# Need to do some kind of weighted total proportion... 
-# the number of reads per sample, and the % of reads should be a factor
-gf_lkp2 <- read.csv('../data/genus2family_lkp.csv')
-gf_lkp <- blst %>% 
-  select(Genus, Family) %>% 
-  distinct(Genus, .keep_all = T) %>% 
-  bind_rows(gf_lkp2) %>% 
-  filter(Family != 'UNKNOWN')
-
-tb <- blst %>% 
-  left_join(., blst_cnt, by = 'species') %>% 
-  group_by(species, TAXON_NEW) %>% 
-  mutate(Mu_Seqs = sum(Prcnt_seqs)/n,
-         species = str_remove(species, 'Bombus '))  %>% 
-  distinct(species, TAXON_NEW, .keep_all = T) %>% 
-  ungroup(TAXON_NEW) %>% 
-  slice_max(Mu_Seqs, n = 10) %>% 
-  select(plant.species = TAXON_NEW, species, n, Mu_Seqs, Genus, Family) %>% 
-  mutate(rank = 1:n()) %>% 
-  ungroup() %>% 
-  mutate(Family = case_when(
-    Genus %in% c('Trollius', 'Caltha', 'Aquilegia', 'Delphinium') ~ 'Ranunculaceae',
-    Genus == 'Hydrophyllum' ~ 'Hydrophyllaceae',
-    Genus == 'Epilobium' ~ 'Onagraceae',
-    Genus == 'Viola' ~ 'Violaceae',
-    Genus == 'Salix' ~ 'Saliceae',
-    Genus == 'Cynoglossum' ~ 'Boraginaceae', 
-    Genus == c('Arenaria', 'Odontostemma') ~ 'Caryophyllaceae',
-    Genus == 'Parnassia' ~ 'Celastraceae',
-    Genus %in% c('Erigeron', 'Symphyotrichum') ~ 'Asteraceae',
-    TRUE ~ Family 
-  ))
-
-obs <- read.csv('../data/yearEndInteractionsObservations.csv')[,2:5] %>% 
-  group_by(species) %>% 
-  arrange(-bySpp) %>% 
-  slice_head(n = 10)  %>% 
-  mutate(plant.species = case_when(
-    plant.species == 'Adenolinum.lewisii' ~  'Linum lewisii', 
-    plant.species == 'Lupinus.bakeri' ~ 'Lupinus sericeus' , 
-    plant.species == 'Dugaldia.hoopesii' ~ 'Hymenoxys hoopesii', 
-    TRUE ~ plant.species 
-  )) %>% 
-  #mutate(Family = if_else(Family == 'UNKNOWN', 'HYDROPHYLLACEAE', Family)) %>% 
-  separate(plant.species, into = 'Genus', extra = 'drop', remove = F) %>% 
-  mutate(plant.species = str_replace(plant.species, '[.]', " ")) %>% 
-  left_join(., gf_lkp, by = 'Genus') %>% 
-  distinct() %>% 
-  filter(species %in% tb$species) %>% 
-  mutate(rank = 1:n()) %>% 
-  ungroup()
-
-## Which bombus species are shared??
-
-# REMOVE KIRBIELLIUS, ONLY  A SINGLE FEEDING RECORD
-obs <- filter(obs, !species %in% c('kirbiellus', 'mixtus'))
-tb <- filter(tb, !species %in% c('kirbiellus', 'mixtus'))
-
-tblist <- split(tb, tb$species) %>% 
-  map(., ~ select(., plant.species, Genus, Family))
-oblist <- split(obs, obs$species) %>% 
-  map(., ~ select(., plant.species, Genus, Family))
-
-species <- purrr::map2_int(tblist, oblist,
-                ~ inner_join(.x, .y, by = 'plant.species') %>%
-                   nrow())
-
-genus <- purrr::map2_int(tblist, oblist,
-  ~ inner_join(.x, .y, by = 'Genus') %>%
-    distinct(plant.species.y) %>% nrow()
-)
-
-family <- purrr::map2_int(tblist, oblist,
-  ~ inner_join(.x, .y, by = 'Family') %>%
-    distinct(plant.species.y) %>%
-    nrow()
-)
-
-correct <- data.frame(Bombus.Species = names(species),
-           Species = species, Genus = genus, Family = family) %>% 
-          # ObsMax = map_int(oblist, ~ nrow(.))) %>% 
-  pivot_longer(!Bombus.Species, names_to = 'Observation') %>% 
-  mutate(Observation = factor(Observation,  levels = c("Species", "Genus", "Family")))
-
-bee_cols <- c('appositus' = '#A50026FF', 'bifarius' = '#FFA319FF', 
-              'californicus' = '#9632B8', 'flavifrons' = '#EFAE90',
-              'rufocinctus' = '#155F83FF', '#8F3931FF', 
-                'mixtus' = '#350E20FF', 'kirbiellus'  =  '#8074A8', 
-              'nevadensis' = '#C96731')
-
-ggplot(data=correct, aes(x=Observation, y=value,
-                         group = Bombus.Species, fill = Bombus.Species)) +
-  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
-              height = 0.05, width = 0.3) +
-  scale_fill_manual('Bee species', values = bee_cols) +
-  scale_color_manual('Bee species', values = bee_cols) +
-  theme_bw() +
-  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
-  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
-       y = 'Shared Observations', x= NULL) + 
-  theme(panel.grid.major.x = element_blank(),
-        panel.border = element_blank())
-
-n_corbiculae <- tb %>% 
-  distinct(species, .keep_all = T) %>% pull(n) 
-prop_obs <- obs %>% 
-  group_by(species) %>% 
-  slice_max(bySpp, n = 5, with_ties = F) %>% 
-  mutate(prop = sum(bySpp)/total) %>% 
-  distinct(species, .keep_all = T) %>% 
-  pull(prop)
-
-rm(bee_cols, correct, family, genus, tblist, oblist, gf_lkp, gf_lkp2)
-```
-
-## 3.2 | Spatial Analyses
-
-```{r Results for SA Modelling}
-
-files <- list.files('/hdd/MS_SDM_RMBL/SDMS_RMBL/results/maps', pattern = 'tif$')
-ml <- files[grep( '_ml_', files)]
-ml <- unique(sub('_ml.*$', '', ml)) # ensure we do not have dupes
-linear <- files[grep('_glm_', files)]
-linear <- unique(sub('_glm.*$', '', linear))
-
-# setdiff(linear, ml) # Barbarea vulgaris, epilobium lactiflorum not modelled with ml
-# setdiff(ml, linear) # ericameria parryi not modelled with 'linear regression'
-
-rm(files, ml, linear)
-```
-
-```{r Results for SA Logistic Regression Metrics}
-
-logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
-  filter(occurrence == 0)
-
-dat <- read.csv('../data/evaluation_table.csv')[,2:3]
-data.frame(Metric  = dat[1:5,1],
-                   Value   = dat[1:5,2],
-                   Metric1 = dat[6:10,1],
-                   Value1  = dat[6:10,2]
-                   ) %>% 
-  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
-    col.names = c('Metric', 'Value', 'Metric', 'Value'),
-    booktabs = T) #%>% 
-  #kableExtra::kable_styling(position = "float_right")
-```
-
-```{r Results for SA Logistic Regression Contingency table}
-read.csv('../data/Pa_conting_table.csv') %>% 
-    knitr::kable(caption = "Species Distribution Modeling evaluation contingency table",
-    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
-    booktabs = T) %>%
-    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
-
-rm(logR_dat, dat)
-```
-
-The threshold distance under which a species would undergo species distribution modelling was the median (25.009 km) of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area.
-A 2-sample test for equality of proportions with continuity correction (X-squared = 13.254, df = 1, p-value = 0.000136, 95% CI 0.04-1.00) was used to test whether more of the records located in the broad ecological sites present at the field station, between the distance of the median (25.009 km) to the third quantile (ca 43.830 km) of the regression distance, were true presences at the field station.
-Including these records would have resulted in modelling an additional 222 species distributions of which 30 are true presences, these taxa were not modelled.
-
-Across the entire spatial domain of modelling all ensembled models (n = 968) had an accuracy of 0.84 (95% CI 0.8356 - 0.8443), kappa = 0.68, p-value < 0.001, sensitivity = 0.80, specificity = 0.87, AUC = 0.92.
-
-In the area of the minimum-spanning tree encompassing the field sites, of the 554 vascular plants with biotic pollination, the 493 ML ensembles accurately predicted the presence of 362 (65.3%), incorrectly predicted the presence of 64 (11.6%), incorrectly predicted 34 true presences (6.1%) as being absent, and correctly predicted the true absence of 33 (6.0%).
-The balanced accuracy of the ensembled models is 0.627 (Sensitivity = 0.340, Specificity 0.914).
-Of the 554 vascular plants with biotic pollination, the 475 LM ensembles accurately predicted the presence of 286 (51.6%), incorrectly predicted the presence of 41 (14.3%), incorrectly predicted 93 true presences (16.8%) as being absent, and correctly predicted the true absence of 55 (9.9%).
-The balanced accuracy of the ensembled models is 0.664 (Sensitivity = 0.573, Specificity 0.754).
-Of the 554 vascular plants with biotic pollination syndromes in the flora 13 (2.3%) were in the Orchid family and 41 (7.4%) are non-natives, both of which are restricted from the database, and can only reduce the number of true predicted presences by roughly 10%.
-
-Of the 117 plant species identified to the species level across the spatial extents of all plots and duration of queen bee activity, the ML ensembles predicted the presence of 105 (89.7%) of them, and LM ensembles 102 (87.2%).
-Of the missing species two (1.7%) are Orchids, six (5.1%) are non-native, and one (0.85%) is of contested taxonomic standing, all of which (7.65%) are restricted from the initial query database.
-
-## 3.2 | Microscopic Pollen identification
-
-```{r compare microscopic to molecular Pollen ID, eval = T}
-
-# are mertensia and Asteraceae 1 & 2 detected in all Pollen samples where the bee was COLLECTED ON THEM ?
-
-pmorph <- read.csv('../data/morpho_pollen.csv') %>% 
-  filter(!sample.id %in% c(45, 54))
-blst_seqs <- distinct(blst, sample.id)
-
-pmorph <- filter(pmorph, sample.id %in% blst$sample.id) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prop = sum/sum(sum))
-
-mertensia_samples <- pmorph %>% 
-  filter(morphotype == 'MERTENSIA') %>% 
-  select(sample.id, morphotype, Prop, sum)
-
-mertensia_samples <- filter(mertensia_samples, sample.id %in% blst_seqs$sample.id)
-
-blst_mertensia <- blst %>% 
-  filter(Family == 'Boraginaceae') %>% 
-  select(sample.id,  Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
-  group_by(sample.id) %>% 
-  mutate(Prcnt_seqs = sum(Prcnt_seqs),
-         morphotype = 'MERTENSIA') %>% 
-  distinct(sample.id, .keep_all = T)
-
-mertensia_detected <- mertensia_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_mertensia$sample.id, 1, 0)) 
- # sample 26 - very strange !!!! this needs investigation - ALSO THE ERICOID IS LIKELY ONAGRAD !?
-
-blst_aster <- blst %>% 
-  filter(Family == 'Asteraceae') %>% 
-  group_by(sample.id) %>% 
-  distinct(sample.id, .keep_all = T)
-
-# ASTER 2 HERE...
-aster2_samples <- pmorph %>% 
-  filter(morphotype == 'ASTERACEAE_2')
-
-blst_aster2 <- filter(blst_aster, Genus %in% 
-                            c('Taraxacum', 'Helianthella', 'Scabrethia')) %>% 
-  mutate(morphotype = 'ASTERACEAE_2')
-aster2_detected <- aster2_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster2$sample.id, 1, 0))
-
-# ASTER 1 HERE... ? WERE ALL OF THESE SPECIES PRUNED... ?
-aster1_samples <- pmorph %>% 
-  filter(morphotype == 'ASTERACEAE_1')
-
-blst_aster1 <- filter(blst_aster, !Genus %in% c('Taraxacum', 'Helianthella', 
-                                                'Scabrethia') ) %>% 
-  mutate(morphotype = 'ASTERACEAE_1')
-aster1_detected <- aster1_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster1$sample.id, 1, 0))
-
-# ALL HIGH NUMBER ASTERACEAE
-
-aster_samples <- pmorph %>% 
-  filter(str_detect(morphotype, 'ASTER.*'))
-aster_detected <- aster_samples %>% 
-  mutate(DETECTED = if_else(sample.id %in% blst_aster$sample.id, 1, 0))
-
-aa <- prcnt_detect(mertensia_detected, Lprop = 0.01, Hprop = 1)
-bb <- prcnt_detect(aster1_detected, Lprop = 0.001, Hprop = 0.01)
-cc <- prcnt_detect(aster2_detected, Lprop = 0.001, Hprop = 0.01)
-dd <- prcnt_detect(aster_detected, Lprop = 0.01, Hprop = 1)
-ff <- prcnt_detect(aster_detected, Lprop = 0.001, Hprop = 0.01)
-
-# Semi-quantitative accord between sequence % and Grain Count
-palyn_cnt <- bind_rows(aster1_samples, aster2_samples, mertensia_samples) %>% 
-  select(morphotype, countProp = Prop)
-
-pollen_props <- bind_rows(blst_aster1, blst_aster2, blst_mertensia) %>% 
-  select(morphotype, sequenceProp = Prcnt_seqs, Seqs_pr_taxon, Total_seqs, Prcnt_Seqs_Class) %>% 
-  left_join(., palyn_cnt, by = c('sample.id', 'morphotype')) %>% 
-  drop_na()
-
-rm(mertensia_samples, blst_aster, blst_aster1, blst_aster2, aster1_samples, 
-   aster2_samples, palyn_cnt, blst_mertensia, aster_samples)
-
-grains_counted <- pmorph %>% 
-  group_by(sample.id) %>% 
-  summarize(grains_counted = sum(sum)) %>% 
-  pull(grains_counted)
-
-```
-
-Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into (Appendix 6).
-From the `r length(unique(pmorph$sample.id))` samples that were counted and based on rarefaction we identified substantial amounts of the abundance and morphotype richness of the samples (morphotype richness, $\bar{x}$ = 4.5, median = 4, min = 1, max = 9) (Appendix  7 & 8).
-The number of counted pollen grains in each sample range from (`r min(grains_counted)`  - `r max(grains_counted)`, $\bar{x}$ = `r round(mean(grains_counted), 2)`, median = `r  median(grains_counted)`).
-
-```{r}
-rm(mertensia_detected, aster1_detected, aster2_detected, aster_detected, grains_counted)
-```
-
-```{r Sequence and Morphological Vouchers, out.width = "65%", fig.cap="Phylogenetic tree of all biotically pollinated plant genera in the study area.
-The innermost ring indicates every genus which Queen Bee's were observed to visit.
-The intermediate ring indicates that at least a single morphological pollen voucher slide was prepared for a member of the genus.
-The outermost ring indicates that sequence data were available for at least a member of that genus.
-Branch colors follow APG 4."}
-knitr::include_graphics("../graphics/plots/rmbl_draft_tree.png")
-```
-
-### 3.3 | Metabarcoding Pollen Identification
-
-### 3.3.2 | Temporal Analysis
-
-The first date of modeled snow melt in the Gothic area (n = 17, $\bar{x}$ = 137.9, Mdn = 135, 3^rd^ quartile = 151), and the first date of a consistent winter snow base (n = 17, $\bar{x}$ = 299.9, Mdn = 300, 1^st^ quartile = 291) from 2000-2017, were used as delimiters for the inclusions of herbarium records in modelling.
-Of the 439 species predicted likely present in the area via logistic regression, 332 species (64.4%) with more than 10 records in the focal level 4 ecoregions ($\bar{x}$ = 35.016, Mdn = 35, max = 96) had Weibull estimates calculated, an additional 56 species (11.2%) with enough contributing records from the "Sedimentary Mid-Elevation Forests", a large ecoregion generally just beneath the elevation bands occupied by the five ecoregions around the study area had Weibull estimates also calculated ($\bar{x}$ = 13.868, Mdn = 13, max = 24).
-
-Only 58 of these 388 species (n = 34.6, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts).
-Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61), peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49).
-There was moderate evidence that the Weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).
-
-Of the previous 58 species compared, 47 of these could be compared to transect based data from the six sites observed in 2015.
-Due to methodological differences, the peak flowering was not compared, and due to the low performance of attempts to model 'duration' in the previous step it was also not compared.
-There was very strong evidence that the Weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).
-
-```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
-# these are the tenth and nintey  percent non-parametric confidence intervals. 
-```
-
-```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
-knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
-```
-
-```{sh Get number of raw and trimmed reads for metagenome samples, eval = F}
-
-cd ~/sequenceData/Original_zip
-files=$(find -type f -name "[0-9]_S[0-9]*_paired.fastq.gz") # doesnt support {2}
-files1=$(find -type f -name '[0-9][0-9]_S[0-9]*_paired.fastq.gz')
-files="${files} ${files1}"
-echo $files
-# seqs per each sample , need divide by 4 on import
-wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
-
-cd ../trimmed_metagenomic
-files=$(find  -maxdepth 1 -name '*gz')
-wc -l  $files  > '/hdd/yun master/MEE_manuscript/data/trimmed_reads_metagenome_seq.txt'
-```
-
-```{r Summary statistics of sequence numbers}
-
-raw_seqs <- read.delim('../data/raw_reads_metagenome_seq.txt', header = F) %>% 
-  mutate(V1 = str_trim(V1)) %>% 
-  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
-  mutate(seq = round(as.numeric(seq)/4),
-         sample = str_extract(sampleFR, '.*L002_')) %>% 
-  filter(sampleFR != 'total') %>% 
-  group_by(sample) %>% 
-  mutate(totalReads = sum(seq)) %>% 
-  distinct(totalReads) %>% 
-  pull(totalReads)
-
-trimmed_seqs <- read.delim('../data/trimmed_reads_metagenome_seq.txt', header = F) %>% 
-  mutate(V1 = str_trim(V1)) %>% 
-  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
-  mutate(seq = round(as.numeric(seq)/4),
-         sample = str_extract(sampleFR, '.*paired')) %>% 
-  filter(sampleFR != 'total') %>% 
-  group_by(sample) %>% 
-  mutate(totalReads = sum(seq)) %>% 
-  distinct(totalReads) %>% 
-  pull(totalReads)
-
-krak <- read.csv('../data/kraken_proc_reads.csv')
-brack <- read.csv('../data/bracken_proc_reads.csv')
-blast <- read.csv('../data/blast_proc_reads.csv')
-loci <- read.csv('../data/reads_per_loci.csv') %>% group_by(Sample)
-
-no_loci <- loci %>% group_by(Sample) %>% count(name = 'no_loci')
-no_reads <- loci %>% group_by(loci) %>% summarize(readsLoci = sum(n))
-```
-
-## 3.3.1 | Molecular analysis of corbiculae loads
-
-The 54 corbiculae loads had DNA extracted and underwent various steps towards hyb-seq, in the end a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r pNumC(sum(raw_seqs))` reads were recovered from sequencing.
-The number of reads per sequence varied widely (range = `r pNumC(min(raw_seqs))` - `r pNumC(max(raw_seqs))`, $\bar{x}$ = `r pNumC(round(mean(raw_seqs),1))`, Mdn = `r pNumC(median(raw_seqs))`).
-Of the possible 353 loci, the number which were recovered from each sample, and informative to BLAST were range = `r toString(min(no_loci$no_loci))` - `r toString(max(no_loci$no_loci))`, $\bar{x}$ = `r toString(round(mean(no_loci$no_loci), 1))`, Mdn = `r toString(median(no_loci$no_loci))`.
-The number of reads per loci from across all samples had a range of `r min(no_reads$readsLoci)` - `r pNumC(max(no_reads$readsLoci))`, $\bar{x}$ = `r pNumC(round(mean(no_reads$readsLoci), 1))`, Mdn = `r pNumC(median(no_reads$readsLoci))` (Appendix 11).
-
-After trimming `r pNumC(sum(trimmed_seqs))` sequences remained.
-`r pNumC(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r pNumC(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r pNumC(sum(blast$Total_seqs))` reads were matched using BLAST.
-Based upon subjective review of the three classifiers (Appendix 12) BLAST was chosen as the classification method which yielded the most probable results by the field ecologist, and its values were used for all subsequent analyses.
-
-```{r Read Post Processing}
-
-pp_seqs_t <- read.csv( file.path('../data', 'Sequence_Reclass_table.csv') ) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 1)))
-
-knitr::kable(pp_seqs_t, align = "c", 
-             col.names = c('Condition', 'No. Class.', 'Prcnt. Class.', 'Total Seqs', 'Rank'),
-             caption = "Post classification of Sequences via Taxonomy and Ecology,\ntop 15 most abundant reads") %>% 
-  kableExtra::kable_styling(
-      latex_options = "striped",
-      stripe_color = "gray!10")
-
-seqs_sp <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Species','Prcnt_Records']), 1)
-seqs_ge <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Genus','Prcnt_Records']), 1)
-seqs_fa <- round(sum (pp_seqs_t[pp_seqs_t$Rank=='Family','Prcnt_Records']), 1)
-
-sp_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Species','Total_Prcnt_Seqs']),1)
-sp_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Genus','Total_Prcnt_Seqs']),1)
-fa_total_seqs <- round(sum(pp_seqs_t[pp_seqs_t$Rank=='Family','Total_Prcnt_Seqs']),1)
-
-byhand <- blst %>% 
-  filter(TAXON_NEW != TAXON_NEW)
-
-j <- toString(paste0(names(table(byhand$Condition)), ": ",
-                     as.vector (table(byhand$Condition))))
-gen <- byhand[byhand$TAXON_NEW==c('Salix'),]
-spp <- byhand[byhand$TAXON_NEW==c('Sorbus scopulina', 'Pseudostellaria jamesiana'),]
-```
-
-The initial classification of sequences which were made by BLAST were reviewed programmatically, using predicted presence of the species (from spatial modelling), modelled flowering time (from temporal modelling), and taxonomy (from existing sources).
-A sequential process was utilized which reassigned sequences based on binary combinations of the factors above (Appendix 15).
-Given the relative sparsity of the number, and relatedness, of species represented in the sequence database this was performed to: 1) Identify locally present species represented by surrogates in the DB 2) Reduce false classifications of focal species 3) Identify high confidence sequence matches.
-Of the top ten taxa which were identified by BLAST for the `r sum(pp_seqs_t$Samples_Classified)` distinct records, `r seqs_sp`% of the reads were classified to a species representing `r sp_total_seqs`% of all classified reads, `r seqs_ge`% of the reads were classified to genus representing `r sp_total_seqs`% of all classified reads, and `r seqs_fa`% of the records were classified to family.
-
-Of the `r nrow(byhand)` classifications which were assigned to genera without any species predicted by spatial analyses, were investigated by hand after post-processing steps.
-These were all assigned via post-processing conditions (`r j`, Appendix 15).
-These were manually assigned to a variety of ranks, occasionally to genus - `r nrow(gen)`, and species - `r nrow(spp)`, by consultation of the alpha-taxonomic literature (@sadeghian2015molecular, @sennikov2017phylogenetic, @rabeler2016new, @pusalkar2015taxonomic, @moore2003its,  @weber1998new).
-
-To determine at which level species in pollen loads could be detected the results of light microscopy were compared to the molecular results.
-The pollen samples contained three morphotypes which could readily be identified via microscopy.
-Two of these mapped to the clades (Boraginaceae & Heliantheae Alliance), and one to a Asteraceae less the Heliantheae Aliiance (Sunflower Tribe) (Boraginaceae grains `r toString(aa)`.
-Asteraceae type 1, non-helianthoids, `r toString(bb)` Asteraceae type 2, Helianthoids, `r toString(cc)`; however, Asteraceae `r toString(ff)`.
-Both morphotypes of Asteraceae pollen `r toString(dd)`.
-
-```{r clean sequence numbers from environment}
-rm( krak, brack, blast, loci, no_loci, aa, bb, cc, dd, fa_total_seqs, j, byhand,
-    sp_total_seqs, seqs_sp, seqs_ge, seqs_fa, pp_seqs_t, spp)
-```
-
-```{r Correlation between Number of Grains and Number of Sequences, fig.cap = 'Relationship between morphological count data and sequence reads', out.height="50%", out.width="50%"}
-
-pol_p <- pollen_props %>% 
-  mutate(sequenceProp = sequenceProp/100, 
-         adjSeqProp = sequenceProp * (100/Prcnt_Seqs_Class))
-
-ct <-cor.test(pol_p$countProp, pol_p$adjSeqProp, method = "kendall")
-# ct[["p.value"]] #too small to write out. tau > 0.4 strong relationship
-model <-  lm(adjSeqProp ~ countProp, data = pol_p)
-
-# png('../../thesis_talk/graphics/plots/CorrelationCountedSequences.png', )
-par(pty="s")
-plot(x = pol_p$countProp,  pol_p$adjSeqProp, 
-       xlim = c(0,1), ylim = c(0,1), axes = F, main = 'Correlation of Proportion Counted\nGrains and Sequence Reads',
-      xlab = 'Grains Counted', ylab = 'Classified Sequences',
-      pch=18, slty = 3, gap = 0.025)
-axis(1)
-axis(2)
-abline(model)
-text(0.3, 0.9, paste0("tau = ", round(ct[["estimate"]][["tau"]], 3)))
-text(0.3, 0.8, "p < 0.001") 
-# dev.off()
-
-# rm( newdat)
-```
-
-To detect whether the sequencing reads were semi-quantitative the subset of all pollen morphotypes distinguishable by microscopy were compared to the sequence reads.
-In all instances sequence reads were pooled to the highest taxonomic rank associated with the morphotype, e.g.
-if both species of *Mertensia* Huth, or one species and read only classified to genus were present in a sample, the reads were summed.
-The total percentage of the ten most abundant grains per sample were then were then relativized to constitute the entire sample.
-
-The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly linear, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated.
-This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing).
-There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).
-
-```{r}
-rm(pol_p, ct)
-```
-
-```{r Co-occurrence of splash drop sequences, eval = F}
-library(CooccurrenceAffinity)
-
-e <- blst %>% 
-  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
-                       'Celastraceae')) %>% 
-  select(sample.id, Genus) %>% 
-  mutate(Present = 1) %>% 
-  group_by(Genus) %>% 
-  distinct() %>% 
-  pivot_wider(names_from = Genus, values_from = Present) %>% 
-  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
-  column_to_rownames('sample.id')
-
-myout <- affinity(data = e, row.or.col = "col", squarematrix = c("all"))
-plotgg(data = myout, variable = "alpha_mle", legendlimit = "datarange")
-```
-
-A number of suspected false positives were returned by the sequence classified, in all instances these were species present in low abundances at the field station which were also returned with a common species in the same family.
-To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala Jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha).
-Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L.
-sp., *Thalictrum* L.
-spp., *Trollius* L.
-sp., *Aquilegia* L.
-spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz.
-were recorded.
-A similar relationship was observed in the Hydrophyllaceae R.Br.
-with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L.
-species were present.
-The flower size  and sparsity of *Nemophila breviflora* A.
-Gray make it unlikely to be visited by Bumble Bees, and it is a false positive.
-The floral morphology and orientation of flower of *Thalictrum*  spp.
-also makes them unlikely to be visited, and while evidence of visits to *Caltha* and *Trollius* are lacking, due to the association between the reads these results appear unlikely.
-
-## 3.6 | Integrated Observational, Molecular, and Palynological Results
-
-```{r compare results, out.width="50%", fig.cap= 'Comparision of Accuracy between the initial output data from BLAST, and these same data subjected to the post-classification process which removes surrogate, and temporally restricted species'}
-
-blast <- read.csv('../data/tally_blast_classifications.csv') %>% 
-  group_by(Sample) %>% 
-  slice_max(Prcnt_seqs, n = 10) %>% 
-  select(Sample, Taxon = taxid, Percent_Sample = Prcnt_seqs) %>% 
-  mutate(
-    Taxon = str_replace(Taxon, 'Arenaria', 'Eremogone'),
-    Genus = str_remove(str_extract(Taxon, '^.* '), ' '),
-    Method = 'BLAST')
-
-automated <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>% 
-  group_by(Sample, TAXON_NEW) %>%
-  mutate(Prcnt_seqs = sum(Prcnt_seqs)) %>% 
-  drop_na() %>% 
-  distinct(Sample, TAXON_NEW, .keep_all = T) %>% 
-  ungroup(TAXON_NEW) %>% 
-  slice_max(Prcnt_seqs, n = 10) %>% 
-  select(Sample, Genus = GENUS, Taxon = TAXON_NEW, Percent_Sample = Prcnt_seqs) %>% 
-  mutate(across(where(is.character), ~ str_replace(., 'Arenaria', 'Eremogone')),
-         Method = 'Automated')
-
-expert <- read.csv('../data/Fully_Integrated_Corbiculae.csv') %>% 
-  rename(Sample = sample.id) %>% 
-  mutate(across(where(is.character), ~ str_trim(.)), 
-         Method = 'Expert') %>% 
-  drop_na() %>% 
-  relocate()
-
-genera2query <- na.omit(unique(c(blast$Genus, automated$Genus, expert$Genus))) 
-genera2query <- genera2query[genera2query != 'Ericaceae'] 
-genera2query <- taxizedb::name2taxid(genera2query, out_type = 'summary')
-
-genera_out <- taxizedb::classification(genera2query$id, db='ncbi', rank = 'family') 
-genera_out <- map(genera_out, ~ 
-                    data.frame( 
-                      'Family' = filter(.x, rank == "family") %>% select(name), 
-                      'Genus' = filter(.x, rank == 'genus') %>%  select(name) 
-                      )) %>% 
-  bind_rows() %>% 
-  rename(Family = 1, Genus = 2)
-
-genera_out <- filter(genera_out, !Family %in% c('Mertensiidae', 'Hesperiidae'))
-
-all_classifications <- bind_rows(
-  automated, blast, expert
-) %>% 
-  left_join(., genera_out, by = 'Genus') %>% 
-  mutate(Rank = case_when(
-    is.na(Family) ~ 'Family',
-    Taxon == Family ~ 'Family',
-    Taxon == Genus ~ 'Genus',
-    TRUE ~ 'Species'
-  ), 
-  Family = if_else(Taxon == 'Ericaceae', 'Eriaceae', Family))
-
-rm(genera2query, genera_out, automated, blast, expert)
-
-expert <- all_classifications %>% filter(Method == 'Expert')
-blast <- all_classifications %>% 
-  filter(Method == 'BLAST' & Sample %in% expert$Sample)
-automated <- all_classifications %>% 
-  filter(Method == 'Automated' & Sample %in% expert$Sample)
-
-## now determine which proportion of records are correct/incorrect
-
-expert_g <- expert %>% 
-  select(Sample, Genus) %>% 
-  distinct()
-expert_f <- expert %>% 
-  select(Sample, Family) %>% 
-  distinct()
-
-# CORRECT RECORDS FIRST
-
-true_species_blast <- left_join(expert, blast, by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method.y) %>% 
-  summarize(Species_T = n())
-true_genus_blast <- left_join(expert_g, blast, by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_T = n())
-true_family_blast <- inner_join(expert_f, blast, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_T = n())
-
-blast_correct <- left_join(true_species_blast, true_genus_blast, by = 'Sample') %>% 
-  left_join(., true_family_blast, by = 'Sample')
-
-rm(true_species_blast, true_genus_blast, true_family_blast)
-
-true_species_computer <- left_join(expert, 
-                                    filter(automated, Rank == 'Species'),
-                                    by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method.y) %>% 
-  summarize(Species_T = n())
-
-true_genus_computer <- left_join(expert_g,  filter(automated, Rank != 'Family'),
-                                    by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_T = n())
-
-true_family_computer <- left_join(expert_f, automated, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  drop_na(Method) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_T = n())
-
-computer_correct <- left_join(true_species_computer, true_genus_computer, by = 'Sample') %>% 
-  left_join(., true_family_computer, by = 'Sample')
-
-rm(true_species_computer, true_genus_computer, true_family_computer)
-
-# FALSE records second
-
-false_species_blast <- anti_join(blast, expert,  by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  summarize(Species_F = n())
-
-false_genus_blast <- anti_join(blast, expert_g, by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_F = n())
-
-false_family_blast <-anti_join(blast, expert_f, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_F = n())
-
-blast_incorrect <- left_join(false_species_blast, false_genus_blast, by = 'Sample') %>% 
-  left_join(., false_family_blast, by = 'Sample')
-
-blast_results <- left_join(blast_correct, blast_incorrect, by = 'Sample') %>% 
-  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))
-
-rm(false_species_blast, false_genus_blast, false_family_blast, blast_correct, blast_incorrect)
-
-false_species_computer <- anti_join(automated, expert,  by = c('Sample', 'Taxon')) %>% 
-  group_by(Sample) %>% 
-  summarize(Species_F = n())
-
-false_genus_computer <- anti_join(automated, expert_g, by = c('Sample', 'Genus')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Genus) %>% 
-  summarize(Genera_F = n())
-
-false_family_computer <-anti_join(automated, expert_f, by = c('Sample', 'Family')) %>% 
-  group_by(Sample) %>% 
-  distinct(Sample, Family) %>% 
-  summarize(Family_F = n())
-
-computer_incorrect <- left_join(false_species_computer, false_genus_computer, by = 'Sample') %>% 
-  left_join(., false_family_computer, by = 'Sample')
-
-computer_results <- left_join(computer_correct, computer_incorrect, by = 'Sample') %>% 
-  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))
-
-rm(false_species_computer, false_genus_compuer, false_family_computer, 
-   computer_correct, computer_incorrect)
-
-### calculate accuracy
-
-# number of assignments
-
-automated_cnts <- automated %>% 
-  group_by(Sample) %>% 
-  count(Rank) %>% 
-  pivot_wider(names_from = Rank, values_from = n, names_prefix = 'NO_') %>% 
-  mutate(Method = 'COMP')
-blast_cnts <- blast %>% 
-  group_by(Sample) %>% 
-  count(Rank) %>% 
-  pivot_wider(names_from = Rank, values_from = n, names_prefix = 'NO_') %>% 
-  mutate(Method = 'BLAST')
-cnts <- bind_rows(automated_cnts, blast_cnts)
-
-spp_count <- expert %>% 
-  group_by(Sample) %>% 
-  count(name = 'Spp_No')
-gen_count <- expert_g %>% 
-  group_by(Sample) %>% 
-  count(name = 'Gen_No')
-fam_count <- expert_f %>% 
-  group_by(Sample) %>% 
-  count(name = 'Fam_No')
-
-ACC_sens_Spec <- bind_rows(
-  mutate(blast_results, Method = 'BLAST'),
-  mutate(computer_results, Method  = 'COMP')
-) %>% 
-  left_join(., spp_count) %>% 
-  left_join(., gen_count) %>% 
-  left_join(., fam_count) %>% 
-  left_join(., cnts, by = c('Sample', 'Method')) %>% 
-  rowwise() %>% 
-  mutate(
-    Species_false = NO_Species - Species_T,
-
-    Sp_Acc = (Species_T / NO_Species) * 100,
-    Gen_Acc = (Genera_T / sum(NO_Genus, NO_Species, na.rm = T) ) * 100,
-
-    Sp_Sens = (Species_T)  / (Species_T + (Spp_No - Species_T)) * 100
-  )
-
-# Spp_No - Species_match is my TRUE NEGATIVE
-# species_false is MY FALSE POSITIVE
-
-wt_res <- wilcox.test(
-  pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'COMP', 'Sp_Acc']) , 
-  pull(ACC_sens_Spec[ACC_sens_Spec$Method == 'BLAST', 'Sp_Acc']),
-  alternative = "greater"
-  )
-
-my_means <- ggpubr::compare_means(Sp_Acc ~ Method,  data = ACC_sens_Spec)
-my_comparisons <- my_means %>% 
-    nest(groups = c(group1, group2)) %>% 
-    pull(groups) %>% 
-    map(., as.character)
-sample_sizes <- ACC_sens_Spec %>% 
-    dplyr::group_by(Method) %>%  
-    dplyr::tally() %>% 
-    dplyr::mutate(n = paste0('n = ', n)) 
-min_v <- dplyr::summarise(ACC_sens_Spec, mean_mpg = floor(min(Sp_Acc))) |>
-    pull() |> min()
-
-ggplot(ACC_sens_Spec, aes(y = Sp_Acc,  x= Method, color = Method),  
-                        alpha = 0.5) + 
-  stat_boxplot(notch = T, notchwidth = 0.75, 
-               varwidth = T, 
-               outlier.shape = F, outlier.alpha = 0.8, outlier.colour = 'black') +
-  geom_jitter(width = 0.1, shape = 1) +
-
-  geom_text(data = sample_sizes,
-            aes(Method, Inf, label = n), color = 'black', 
-            vjust = "inward", size = 4, 
-            y = min_v * 0.5) +
-  ggpubr::stat_compare_means(comparisons = my_comparisons, 
-                             aes(label = ..p.signif..),
-                             tip.length = 0, vjust = 0.25, size = 4) +
-
-  expand_limits(y= min_v - 5) +
-  theme_bw() +
-  scale_colour_manual(values = c('#F4AFB4', '#94A89A')) +
-  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
-  scale_x_discrete(labels = c('BLAST', 'Post Classification')) + 
-  labs(y = 'Accuracy', title = 'Accuracy of "naive" BLAST and post classification process') +
-  theme(legend.position = 'none', 
-        plot.title = element_text(hjust = 0.5))
-
-```
-
-```{r Most Used Floral Resources Table}
-
-fam <- all_classifications %>% 
-  ungroup() %>% 
-  select(Family, Genus) %>% 
-  distinct()
-
-expert %>% 
-  group_by(Taxon) %>% 
-  add_count(name = 'No.Corb') %>% 
-  mutate(median_percent_corb = round( mean(Percent_Sample), 1), 
-         Taxon = str_remove(Taxon, '^.* '), 
-         Taxon = if_else(Taxon %in% c('Salix', 'Erigeron'), 'sp.', Taxon)) %>% 
-  select(Genus, Species = Taxon, Percent_Sample, No.Corb, median_percent_corb) %>% 
-  left_join(., fam, by = 'Genus') %>% 
-  select(-Percent_Sample) %>% 
-  relocate(Family, .before = Genus) %>% 
-  filter(No.Corb > 4) %>% 
-  distinct() %>% 
-  arrange(Family, Genus)  %>% 
-  knitr::kable(
-    escape = F,
-    align = c('l', 'l', 'l', 'c', 'c'),
-    caption = 'Plant species detected in five or more corbiculae loads',
-    col.names = c('Family', 'Genus', 'Species', 'No. Samples', 'Mean Prop.\\textsuperscript{a}')
-  ) %>% 
-  add_footnote(
-    'The mean only calculated across the samples where the species was detected') %>% 
-  collapse_rows()
-
-```
-
-```{r}
-rm(n_corbiculae, prop_obs)
-```
-
-# 4 | DISCUSSION
-
-We have demonstrated how the Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context (@johnson2019universal, @hollingsworth2016telling).
-This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history guided fundamental science and land management.
-The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g.
-*Mertensia*, *Lupinus* L.), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic).
-We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas.
-By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.
-
-  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area.
-However, we had difficulties in evaluating our predictions in an operational context.
-We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local).
-While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult.
-We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data.
-Further given the, size of the minimum spanning tree  which we extracted points to, a formal floristic inventory would still be a time intensive process.
-Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding.
-We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.
-
-In regards to the case study, our results indicate the overall information gleaned from observations of queen Bumble Bee foraging and analysis of pollen records are largely congruent.
-Relaxing concerns regarding differences between the broad insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence).
-In general when interaction networks are considered at coarse levels, such as the duration of a season, our perceptions regarding the generality of interactions at smaller time scales may be inflated relative to the actualized interactions within them, e.g.
-a week (@caradonna2020temporal).
-These results indicate a possibility that at even finer levels *Bombus* display high amounts of floral fidelity within foraging bouts, an observation which implies that part of the reason for the high efficiency of *Bombus* as a pollinator might partially be related to their lack of movement of hetero-specific pollen (@brosi2013single, @ashman2013toward, @galloni2008visitor, @brosi2016pollinator).
-The mechanisms behind this observed fidelity are likely related to pollen nutritional values, specifically high concentrations of protein, and the absence of particular amino acids required for larval development in other flower more commonly used by workers (@genissel2002influence, @tasei2008nutritive, @goulson2005causes, @goulson2008diet, @hanley2008breeding).
-
-```{r Table of Limitations Faced}
-read.csv('../data/Limitations_Table.csv') |>
-  knitr::kable(caption = 'Limitations Faced and Possible Solutions',
-             col.names = c('Method Component', 'Limitations', 'Paths Forward')) |>
-  kable_styling(full_width = TRUE)
-```
-
-Also regarding the case study, some foraging preferences of *Bombus*, at this field site and across several localities globally emerge.
-These suggest the need for land managers to maintain relatively high amounts of members of the plant families Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble).
-Numerous historic  practices reduce the ability of many landscapes to support stable populations of *Bombus*.
-Historic livestock grazing involved the removal of many species known to have compounds toxic to cattle.
-In particular, the removal of locoweeds (*Astragalus* & *Oxytropis*) and larkspurs (*Delphinium*) were common across public lands administered by the U.S.
-Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism).
-Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks, processes which lowered water tables and reduced the extent of stream-associated wetlands and the mesic meadows fringes which provide habitat for many species of 'tall' *Mertensia* (e.g.
-*M.
-ciliata*), to an extent *Delphinium barbeyi*  and many species of native *Trifolium* (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos).
-Fire suppression resulted in the succession of many Aspen (*Populus tremuloides*) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effecting habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading).
-Finally the effects of Nitrogen deposition, given the West's rapidly growing population still pose adverse effects on the abundance of species of Fabaceae at urban-rural interfaces (see @stevens2018atmospheric, @fenn2003ecological).
-Current solutions to the above issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing.
-The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias*, using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus*) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
-
-```{r}
-read.csv('../data/Errors_Table.csv') |>
-  knitr::kable(caption = 'Errors faced with components of the framework',
-             col.names = c('Component', 'Example', 'Correction')) 
-```
-
-  We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical).
-To reduce the effects of a low population density of botanists on the maintenance of and production of Flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists.
-In this proof-of-concept example we performed several iterations of modelling runs, and several approaches (i.e.
-the 'linear models', and the 'machine learning'), which took notable amounts of compute power.
-We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicit process to reduce spatial autocorrelation.
-However, given the time required to perform all aspects of a study, even our amount of computation was negligible.
-Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate.
-In certain scenarios modelling of predicted species via more formally tailored S(tacked)-SDM or J(oint)-SDM approaches may be beneficial (@wilkinson2021defining, @pinto2021predicting, @schmitt2017ssdm).
-
-  Tandem to the lack of continued expertise required to generate and maintain species lists, is the expertise required to continue tracking when major phenological events occur in many plant species at relatively fine scales or under novel climates.
-Knowledge of these events is currently limited to general time periods of only a handful of phenological events and groups of organisms (e.g.
-flowering initiation, or trees) (@prather2004implications, @li2016responses).
-While many programs and initiatives exist to collect phenological information on subsets of easily identifiable charismatic species to detect major trends in phenology, these capture only a subset of the extent diversity (@betancourt2005implementing, @havens2007chicago).
-In many instances it appears that while landscapes respond similarly to environmental variables which predict phenological responses, that individual species vary widely in their responses to similar environmental cues, or respond to different cues (@augspurger2020concordance, @xie2015deciduous, @xie2018predicting, @caradonna2014shifts).
-As can be seen here, predictions of when a single, major phenological event occurs is already data limited.
-A more promising approach for the tropics may lay in utilizing circular statistics (@park2022herbarium).
-
-  The nearly complete Plant and Fungal Tree of Life (PAFTOL) will provide a comprehensive phylogenetic backbone of the entire plant kingdom, and the inclusion of A353 probes with lineage specific probe sets is common in producing massive genetic datasets (@baker2021exploring).
-We predict that the A353 probes which it is utilizing to work nearly immediately for DNA barcoding of whole plant material, and that more elaborate validation studies in controlled metabarcoding settings, utilizing existing experimental designs, will have favorable results (@bell2017applying, @bell2019quantitative, @bell2021comparing, @lamb2019quantitative).
-In particular the harvesting of loci with more variation in certain lineages, and or with more variable flanking regions, will prove promising for identifying closely related plant material.
-We suspect that conserved reaches of genes resulted in the high amounts of reads in somewhat obscure species.
-Given that the A353 loci are nuclear, single copy, and a variety are present the possibility of identifying target loci for quantitative purposes is high, without continual PCR enrichment is possible; this would align with relatively high efficacy of WGS (@lang2019genome, @peel2019semi, @bell2021comparing).
-Recent evidence indicates that the potential for identifying nearly cryptic taxa and even infra-specific inference, of either whole plant material, and perhaps in metagenomic context are possible (@ottenlips2021resolving, @wenzell2021incomplete, Loke et al.
-in prep,  @slimp2021potential, @beck2021palmer).
-We further believe that in synthetic phylogenetic trees - with incorporation of NGS backbones - will allow in automatic reassignment of reads as a function of phylogenetic distance with measures of uncertainty (@hinchliff2015synthesis, @smith2018constructing, @baker2021PAFTOL).
-
-# 5 | CONCLUSION
-
-We believe that the combination of spatial and temporal models, united and guided by localized natural history knowledge, provides the essential components of a framework for approaching the coarse elucidation of ecological interactions using DNA Barcoding.
-Herein we crudely utilized this thinking via binary outcomes, should a species predicted be predicted present or not?
-Is it unequivocally flowering or not?
-Myriad data show biological systems and ecological interactions have more variance than can be reasonably discretely parsed.
-We expect that within a framework developed from our preliminary works studies of pollinator behavior may be enacted via this approach at a landscape level, e.g.
-the scale of an entire drainage basin such as the Gunnison which is quickly becoming one of the worlds few model ecosystems.
-We hope that the A353 probes as tools for metabarcoding play a role in these endeavors.
-
-**AUTHOR CONTRIBUTIONS:**  R.C.B conducted botanical collections, conducted all molecular lab work, lead all analyses, and writing.
-J.E.O conceived, designed, and conducted ecological fieldwork, assisted with analyses, and writing.
-P.J.C conducted ecological fieldwork, assisted with ecological analyses, and writing.
-E.J.W.
-prepared, imaged, and collected trait data on pollen reference slides, and assisted with analysis of trait data and writing a dichotomous key.
-S.T.
-assisted with spatial analyses and writing.
-J.B.F.
-conceived, and designed lab work, analyses, and integration of approaches, assisted with writing, and secured funding for molecular work.
-
-```{r}
-a <- pull(read.csv('../data/herbaria_4_acknowledgements.csv'))
-```
-
-**ACKNOWLEDGMENTS:** Nyree Zerega for assistance obtaining herbaria loans and accessioning our collections at CHIC.
-Pat Herendeen for assistance with virtually all aspects of preparing pollen vouchers and the identification process.
-Hilary Noble, Zoe Diaz-Martinez, Angela McDonnell, & Elena Loke for assistance with genomic library preparation.
-Ian Breckheimer for sharing the SDM predictor variables.
-We thank the curators at the following herbaria for supplying tissue: Ben Legler at Stillinger (ID), Charles (Rick) Williams at Ray J.
-Davis (IDS), (B)Ernie Nelson at Rocky Mountain (RM); and the collectors: D.
-Knoke, L.
-Brummer, J.
-Boyd, C.
-Davidson, I.
-Gilman, M.
-Kirkpatrick, S.
-McCauley, J.
-Smith, K.
-Taylor, & C.
-Williams.
-David Giblin & Mare Nazaire for sharing relevant sections of an advanced draft of FNA V.
-15.
-The Bureau of Land Management is thanked as many plant specimens were collected by R.C.B as a partner or contractor to the agency; Sarah Burnett and Lauren Price are thanked for sharing AIM data.
-Sanda and New England Biotech are gratefully acknowledged for technical support and generously sharing samples.
-T.C.H.
-Cole for sharing the Angiosperm Phylogeny 4 colour palette.
-The Program in Plant Biology and Conservation is thanked for funding.
-The holdings of the following herbaria were essential for this project: `r toString(a)`.
-
-**CONFLICT OF INTERESTS** The authors declare no conflicts of interest.
-
-**PEER REVIEW** The peer review history for this document is available at ...
-
-**DATA AVAILABILITY STATEMENT** The queries required to download all data used in this project are located in... All novel sequencing data are located at NCBI...
-
-**ORCID**
-
-Paul CaraDonna https://orcid.org/0000-0003-3517-9090  
-Jeremie Fant https://orcid.org/0000-0001-9276-1111  
-Jane Ogilvie https://orcid.org/0000-0001-8546-0417  
-Sophie Taddeo https://orcid.org/0000-0002-7789-1417
-
-# References
-
-# Supporting
+# Sp
 
@@
 \newpage
-
-APPENDIX 14 - Time Spent Generating Species Distribution Models
-
-```{r}
-knitr::include_graphics('../graphics/assorted/individual_model_runs.pdf')
-```
-
-\newpage
-
-APPENDIX 15 - Review Classified Reads and Reassign
-
-```{r}
-knitr::include_graphics('../graphics/tables/reclassify_reads_table.pdf')
-```
-
-\newpage
-
-APPENDIX 16 - Tips and tricks for implementing plant metagenomic sequencing
-
-```{r}
-knitr::include_graphics('../graphics/assorted/overall_metagenomic_workflow.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Morphological Workflow}
-knitr::include_graphics('../graphics/assorted/microscopy_workflow.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Molecular Workflow}
-knitr::include_graphics('../graphics/assorted/molecular_workflow.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Extraction Buffer}
-knitr::include_graphics('../graphics/assorted/pollen_ctab.pdf')
-```
-
-\newpage
-
-```{r Tips and tricks Extraction Buffer back}
-knitr::include_graphics('../graphics/assorted/pollen_ctab-1.pdf')
-```
-
-\newpage
-
-Appendix 17 Samples Per Bee
-
-```{r Samples per Bee}
-knitr::include_graphics('../graphics/tables/bombus_samples.pdf')
-```
-
-\newpage
-
-Appendix 18 - Overview of the Whole Process
-
-```{r Overview of entire process, out.height="90%"}
-knitr::include_graphics('../graphics/plots/flowchart.png')
-```
-
-\newpage
-
-# References

### 2024-02-11 script/bombus_summaries2015-2020.R

@@ -1,0 +1,52 @@
+setwd("~/Documents/ManuSCript/script")
+
+library(tidyverse)
+
+commonCrestedButteBees <- c('nevadensis', 'californicus', 'appositus', 'flavifrons', 'rufocinctus', 'mixtus', 'bifarius')
+
+dat <- read.csv('../data/Bombus_queen_observations_2015-2022.csv') %>% 
+  filter(year < 2021 & resource.coll != 'nectar', species %in% commonCrestedButteBees)
+
+dat %>% 
+  select(genus, species, plant.species) %>% 
+  distinct() %>% 
+  filter(species %in% c('nevadensis', 'californicus', 'appositus', 'flavifrons', 'rufocinctus', 'mixtus', 'bifarius')) %>% 
+  group_by(species) %>% 
+  count() %>% 
+  pull(n) %>% 
+  median()
+
+pl_stats <- dat %>% 
+  select(plant.species) %>% 
+  separate(plant.species, c('genus', 'species'), remove = F) 
+
+pl_stats %>% 
+  distinct(plant.species) %>% 
+  nrow()
+
+pl_stats %>%  # no of genera by Weber
+  distinct(genus) %>% 
+  nrow()
+
+pl_stats %>% 
+  distinct(genus)
+
+pl_stats %>% 
+  mutate(Family = case_when(
+    genus %in% c('Heracleum', 'Osmorhiza', 'Taraxacum') ~ 'Apiaceae',
+    genus %in% c('Dugaldia', 'Helianthella', 'Heliomeris', 'Erigeron', 'Senecio') ~ 'Asteraceae', 
+    genus %in% c('Mertensia') ~ 'Boraginaceae',
+    genus %in% c('Calochortus' ,'Erythronium') ~ 'Liliaceae', 
+    genus %in% c('Distegia') ~ 'Caprifoliaceae',
+    genus %in% c('Lathyrus', 'Lupinus', 'Vicia' ) ~ 'Fabaceae',
+    genus %in% c('Frasera') ~ 'Gentianaceae',
+    genus %in% c('Hydrophyllum', 'Phacelia') ~ 'Hydrophyllaceae',
+    genus %in% c('Adenolinum') ~ 'Linaceae',
+    genus %in% c('Claytona') ~ 'Montiaceae',
+    genus %in% c('Pedicularis') ~ 'Orobanchaceae', 
[…]

### 2024-02-11 script/collate_interaction_values.R

@@ -24,7 +24,7 @@ blst <- read.csv('../data/Post_Classified_BLAST_sqs.csv') %>%
 # use inner join to drop the unmatched records
 
 
-queen_obs <- read.csv('../data/Bombus_queen_observations_2015.csv') #%>% 
+queen_obs <- read.csv('../data/Bombus_queen_observations_2015.csv') %>% 
   select(Bee_spp = species, Site = site, Date = date,
          Plant_spp_morpho = plant.species ) %>% 
   mutate(Bee_spp = str_remove(Bee_spp, '[.]orange|[.]dark'), 

### 2024-02-18 script/PercentSpeciesFloweringWeekly.R
barplot with error bars supposedly of number of species flowering within ci

@@ -1,0 +1,72 @@
+setwd('~/Documents/ManuSCript/script')
+
+library(tidyverse)
+
+weibs <- bind_rows(
+  read.csv('../data/weibull_estimates.csv') %>% 
+  mutate(Estimate = 'SubalpineOnly'), 
+  read.csv('../data/weibull_estimates_mixed_montane.csv') %>% 
+  mutate(Estimate = 'SubalpineMontane')
+) %>% 
+  arrange(taxon) %>% 
+  filter(metric %in% c('tenth', 'ninety')) %>% 
+  select(event, taxon, doy = DOY)
+
+weibs <- split(weibs, f = list( weibs$taxon, weibs$event))
+weibs <- lapply(weibs, function(x){data.frame(
+  doy = as.matrix(
+    seq(x$doy[1], x$doy[2])))})  %>% 
+  bind_rows(.id = 'taxon')
+
+bees <- read.csv('../data/Bombus_queen_observations_2015.csv') %>% 
+  mutate(species = gsub('[.].*$', '', species)) %>% 
+  select(year, week, species) %>% 
+  group_by(species, week) %>% 
+  tally() %>% 
+  filter(! species %in% c('centralis', 'melanopygus', 'kirbiellus', 'unknown'))
+
+#ggplot(bees, aes(x = n, y = species)) + 
+#  geom_density_ridges()
+
+weeks <- data.frame(
+  week = rep(1:19, each = 7), 
+  doy = 
+    c(130:136, 137:143, 144:150, 151:157, 158:164, 165:171, 172:178, 179:185, 186:192, 
+      193:199, 200:206, 207:213, 214:220, 221:227, 228:234, 235:241, 242:248, 249:255, 256:262)
+    )
+
+test <- left_join(weibs, weeks) %>% 
+  distinct(taxon, week) %>% 
+  filter(week %in% 3:11) %>% 
+  separate(taxon,  c('taxon', 'estimate'), '[.]')
+
+wkkly <- test %>% 
+  group_by(week) %>% 
+  filter(estimate == 'estimate') %>% 
+  tally()
+wkkly$prop <- wkkly$n / nrow(distinct(test, taxon))
+
+wkklyLow <- test %>% 
+  group_by(week) %>% 
+  filter(estimate == 'low_ci') %>% 
+  tally()
+wkklyLow$prop <- wkklyLow$n / nrow(distinct(test, taxon))
+
+wkklyHigh <- test %>% 
[…]

### 2024-03-02 script/CoverLetter.Rmd
c

---
output: pdf_document
---

\begin{flushleft}
Dr. Aaron Ellison \newline
Executive Editor \newline
Methods in Ecology and Evolution \newline
\end{flushleft}

March 17^th^, 2024

Attached is our standard article submission "*Protocol to increase accuracy and fidelity of pollen meta-genomic barcoding using Angiosperms353: a case study using pollen loads from wild bumble bees*".

In this paper, we combine Hyb-Seq to generate *de novo* sequence databases, informed by museum records, and species distribution modelling. 
Further we use the results of phenological modelling to interpret our sequence alignment results to increase the likelihood of accurate results.

Given the rising popularity of metabarcoding and metagenomics, especially for projects over large spatial domains and with multiple seasonal time points, we believe that our method provides a framework for meaningfully assessing and contextualizing results. 
We have observed that many studies utilizing metabacoding generate spurious results, which are precluded due to spatial or temporal incongruities with the focal animal species and plants. 
However, a simple and intuitive framework for programmatically assessing and rectifying these issues are lacking.

We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and that several manuscripts utilizing them will be forthcoming to a variety of journals.

Our submission contains X supplementary materials, X figures and X tables. It unfortunately is currently at roughly 9,300 words. 
However *circa* 900 words are dedicated to a plain text overview of the methods, a section which we believe elucidates the integration of the various approaches, and clarifies adoption by subsequent users. 
Considerable length is dedicated to the methods, which contain five years of field work, and similarly thorough microscopy, and molecular work; but which we have cut by nearly 500 words.

We have no conflicts of interest to disclose. Additionally, this work - nor any part of it - has been submitted to any other journal.

Please address all correspondence regarding this manuscript to me at rbenkendorf@chicagobotanic.org

Thank you for considering this work.

\begin{flushleft}
Reed Clark Benkendorf \newline
Senior (Field) Botanist \& Senior Spatial Data Specialist  \newline
The Chicago Botanic Garden  \newline
rbenkendorf@chicagobotanic.org
\end{flushleft}

\begin{flushleft}
Jeremie B. Fant, Ph.D \newline
Medard and Elizabeth Welch Director of Conservation \newline
The Chicago Botanic Garden  \newline
jfant@chicagobotanic.org
\end{flushleft}

### 2024-03-02 script/PercentSpeciesFloweringWeekly.R

@@ -12,6 +12,8 @@ weibs <- bind_rows(
   filter(metric %in% c('tenth', 'ninety')) %>% 
   select(event, taxon, doy = DOY)
 
+n_tax_modelled <- nrow(distinct(weibs, taxon)) # 383
+
 weibs <- split(weibs, f = list( weibs$taxon, weibs$event))
 weibs <- lapply(weibs, function(x){data.frame(
   doy = as.matrix(
@@ -23,7 +25,8 @@ bees <- read.csv('../data/Bombus_queen_observations_2015.csv') %>%
   select(year, week, species) %>% 
   group_by(species, week) %>% 
   tally() %>% 
-  filter(! species %in% c('centralis', 'melanopygus', 'kirbiellus', 'unknown'))
+  filter(! species %in% c('centralis', 'melanopygus', 'kirbiellus', 
+                          'unknown', 'insularis', 'sylvicola'))
 
 #ggplot(bees, aes(x = n, y = species)) + 
 #  geom_density_ridges()
@@ -32,41 +35,37 @@ weeks <- data.frame(
   week = rep(1:19, each = 7), 
   doy = 
     c(130:136, 137:143, 144:150, 151:157, 158:164, 165:171, 172:178, 179:185, 186:192, 
-      193:199, 200:206, 207:213, 214:220, 221:227, 228:234, 235:241, 242:248, 249:255, 256:262)
+      193:199, 200:206, 207:213, 214:220, 221:227, 228:234, 235:241,
+      242:248, 249:255, 256:262)
     )
 
 test <- left_join(weibs, weeks) %>% 
   distinct(taxon, week) %>% 
   filter(week %in% 3:11) %>% 
-  separate(taxon,  c('taxon', 'estimate'), '[.]')
+  separate(taxon,  c('taxon', 'estimate'), '[.]') %>% 
+  filter(estimate == 'estimate')
+
+rm(weibs, weeks)
 
 wkkly <- test %>% 
   group_by(week) %>% 
   filter(estimate == 'estimate') %>% 
   tally()
-wkkly$prop <- wkkly$n / nrow(distinct(test, taxon))
+wkkly$prop <- wkkly$n / n_tax_modelled
 
-wkklyLow <- test %>% 
-  group_by(week) %>% 
-  filter(estimate == 'low_ci') %>% 
-  tally()
-wkklyLow$prop <- wkklyLow$n / nrow(distinct(test, taxon))
-
[…]

### 2024-03-02 script/methods_supplemental.Rmd
sent supplemental methods over from LS

--- 
title: Supplemental Methods
output:
  pdf_document: default
  toc: no
  word_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r Required Libraries, message=F, warning=F, results='hide'}
library(knitr)
```

To improve metabarcoding reliability and efficiency, we suggest creating a regional list of candidate species using digital collections gleaned from herbaria, survey work, and community science  (Figure 1). 
This list can further be refined using species distribution models and temporal filtering to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area. 
The final list is then used to inform collection of plant samples to create a library and inform metabarcoding.
We apply this methodological framework to the metabarcoding of corbiculae pollen loads of bumble bees and compare the accuracy of our metabarcoding approach both prior and after applying a spatial and temporal filtering to pollen identification conducted by experts and field observations.

### System background

To test the effectiveness of our methodological approach we applied it to identify the plant species
found in the corbiculae (pollen loads) of queen bumble bees (*Bombus* Latreille) collected from the Rocky
Mountain Biological Laboratory in Colorado, USA (RMBL; 38°57.5” N, 106°59.3” W (WGS 84), 2900 m.a.s.l.).
We collected pollen loads from wild foraging queen bees between May and July of 2015 at six permanent study sites (Ogilvie & 
CaraDonna 2022). 
To harvest the pollen loads, we captured queens in an insect net, transferred them
into a restraining device (Kearns et al. (2001)), collected a pollen load from one leg, and then released
them. 
We collected 64 corbiculae pollen loads from queens of several common wild bumble bee
species: *Bombus appositus*, *B. bifarius*, *B. californicus*, *B. flavifrons*, *B. nevadensis*, and *B. rufocinctus*
(Pyke 1982; Ogilvie & CaraDonna 2022). 
At the six study sites, bumble bee abundance and interactions with flowering plants were monitored for one hour at weekly intervals. 
The abundance of flowers visited by bumble bees within belt transects spread over the three vegetation types (0.5 x 40 m transects in each vegetation type, 60 m^2^ total area per site). 
We used six years (2015-2020) of observational data on *Bombus* flower visits to identify the plant taxa most frequently visited by queens across all years and to compare with metagenomic data.

### Survey database to generate a regional taxa list

We first generated a short list of potential candidate species. We downloaded, from the Botanical Information and Ecology Network ‘BIEN’ (@bien2022), all records adjacent to the field sites to develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study area. To reduce the number of species to include in the reference sequence databases, we then generated Species Distribution Models (SDMs) for these taxa to predict their distribution throughout the study area.

To minimize the number of species for which SDM’s were to be generated, BIEN was queried at a distance of up to 100km from our study area and all plant species records were downloaded. To account for the stochasticity of botanical collecting and offset the number of records associated with the research station, this data set was bootstrap re-sampled 250 times, with 90% of samples selected, to create a testing data set. The median of the logistic regression assessing the probability of occurrence of a species record as a function of distance from the study area was used as a threshold distance, under which, to include species as candidates for distribution modelling.

### Spatial filtering via species distribution modelling

To determine which clades to include in the reference sequence database we used Species Distribution Modelling. We used all occurrence records from BIEN (n = 23,919) within a 50km border of the ecoregion, Omernik level 3, which includes the study area *(No. 21 “Southern Rockies”)* to construct the species distribution model  (@omernik1987ecoregions). These records were copied into two, initially identical, sets, one for generating machine learning models (ML; Random Forest, and Boosted Regression Tree’s), and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM) (@barbet2012selecting). Ensembled predictions have been shown to outperform their constituent models, on average, and to reduce the ecological signal to the analytical noise of individual runs (@araujo2007ensemble). No single method of producing SDMs has been shown to universally outperform others when faced with a large and diverse number of applications, in our case a great number of species with different biology and ecology (@elith2006novel, @qiao2015no). In the spirit of these findings, multiple families of models, which can be generated together as they have similar requirements regarding the number and ratios of Presence to Absence records were ensembled together (@barbet2012selecting).

We then generated 4,029 absence points, locations where the focal taxon is anticipated missing, through a random stratification of 19% of the land cover in the area and included them in (@aim2019database). To achieve a larger absence data set, we generated 1,000 pseudo-absence records for each taxon by randomly selecting coordinates located at least 10km away from an occurrence record. For ML models, these pseudo-absences were reduced so that the ratio of presence to absence records were balanced (@barbet2012selecting). To achieve this, we removed absence records inside of 10% of the mean sample value of any predictor variable the presence records; the required number of absence records were then randomly sampled.

To predict the potential distribution of each species we used 26 environmental variables at 30m resolution, six related to climate, five soil, four topographic, four related to cloud cover, with the remaining reflecting assorted abiotic parameters (@wilson2016remotely, @wang2016locally, @hengl2017soilgrids250m, @robinson2014earthenv). These publicly available data sets, were selected as they pertain to a wide range of variables interacting with plant physiology. For linear regression models these predictors underwent both *vifstep* (theta = 10, max observations = 12,500) and *vifcor* (theta = 0.7, max observations = 12,500) to detect highly correlated variables, and collinear features were removed leaving 16 variables (@usdm2014).

Modelling: Random Forest and Boosted Regression Trees, were sub sampled with 30% test and two replicates each before weighted ensemble based on True Skill Statistics (tss) (@sdmPackage). Generalised linear models (GLM) and Generalised additive models (GAM) with 30% sub sampling and three replicates each were also ensembled using the tss (@sdmPackage). TSS was chosen as the ensemble criterion as it has been shown to work across a wide range of species occurrences prevalence (@allouche2006assessing). The results of these models were extracted on a cell-by-cell basis to a polygon feature derived from a minimum-spanning tree which encompasses the study sites, and species from either ensemble with greater than 50% mean habitat suitability across all cells were considered present for further purposes (@prim1957minimum).

A total of 535 species were modelled using Generalized Linear Models and Generalized Additive Models and 534 species were modelled using Random Forest and Boosted Regression Trees.  To evaluate the accuracy of the species distribution models, additional presence records from GBIF (n = 61,789), and AIM (n = 12,730) were used as test and training sets (n = 74,519) for logistic regression (@gbifDL2021sdms, @aim2019database). Additional novel absence records were generated from the AIM data set to create a data set where each species has balanced presence and absences. Eleven or more paired presence and absence records were required for this testing, resulting in 334 species being included in the logistic regression (Mdn = 110.0, $\bar{x}$ = 223.1, max = 1568 record pairs used) with a 70% test split (@caret).

#### Temporal Filtering of the species list

For assignment of reads to ecologically probabilistic species subsequent to BLAST, flowering time was used as a filter. To estimate the duration of dates in which plant species were flowering Weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical). Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not found in the nearest 5), and which were from herbarium records were included. To remove temporally irrelevant herbarium records, i.e. material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in fall. Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed. Species with > 10 records had their Weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering. These estimates were compared to a long-term observational study of flowering phenology 1974-2012 (@caradonna2014shifts), and the floral abundance data from 2015, using Kendall's tau.

### Microscopic pollen identification

To qualitatively identify, and quantitatively note, the plant species present in corbiculae loads microscopy was used.
A pollen reference library of fuchsin-jelly stained grains which may be present in corbiculae loads of slides was assembled from slides previously prepared by the authors (n = 21), and other researchers (n = 38) (@beattie1971technique, @brosi2013single).
Using five years of observational data on *Bombus* Queen Bee foraging at these studies sites (@ogilvie2022shifting), as well as the RMBL Vascular Plant Checklist (@fraser2007vpc), an additional 62 voucher slides for species were prepared and imaged at 400x (Leica DMLB, Leica MC170 HD Camera, Leica Application Suite V. 4.13.0) from non-accessioned herbarium collections to supplement the number of species and clades covered.

We used clustering techniques to supplement our subjective opinions of which plant taxa were distinguishable via light microscopy, and to develop a dichotomous key to pollen morphotypes. 
Ten readily discernible categorical traits were collected from each specimen in the image collection. 
These traits were transformed using Gower distances, and clustered using Divisive Hierarchical clustering techniques (@cluster2022). 
Using the cluster dendrogram, elbow plot, and heatmaps (@fpc2022), of these results morphological groups of pollen which could not be resolved via microscopy were delineated, and a dichotomous key was prepared. 
This key was then used to identify the pollen grains sampled from corbiculae loads to morphotypes in a consistent manner.

To prepare the pollen slides from corbiculae, all corbiculae loads were broken apart and rolled using dissection needlepoints to increase heterogeneity of samples. 
*Circa* 0.5mm^2^ of pollen was placed onto a ~4mm^2^ fuchsin jelly cube (@beattie1971technique) atop a graticulated microscope slide, with 20 transects and 20 rows (400 quadrants) (EMS, Hartfield, PA). 
The jelly was melted, with stirring, until pollen grains were homogeneously spread across the microscope slide. 
Slides were sealed with Canada Balsam (Rublev Colours, Willits, CA) followed by sealing with clear nail polish to prevent oxidation; all samples are noted in.

To identify the pollen present in corbiculae loads, light microscopy at 400x (Zeiss Axioscope A1) was used. 
In initial sampling in three transects, each pollen grain was identified to morphotype and counted; an additional two transects were scanned for morphotypes unique to that slide, if either transect contained a unique morphotype than all grains in that transect were also identified and counted. 
Subsequent to the first round of sampling, non-parametric species richness rarefaction curves (@vegans2022), and non-parametric species diversity rarefaction curves were used to assess the completeness of sampling (@inextArticle, @inextPackage). 
Slides not approaching the asymptote of the rarefaction curve were then re-sampled, and analysed iteratively for up to a total of seven transects.

#### Metagenomics: additional plant tissue collection and extraction

Using five years (2015-2020) of observational data on *Bombus* queen interactions with flowering plants at these studies sites, we identified the plant taxa most frequently visited by queens across all years. 
In order to capture more variability inherit in the 353 loci we sequenced the 12 most visited taxa twice using samples collected from one site within the Gunnison Basin River Drainage and one individual collected from another more distal population.
In addition we included a congener - or a species from a closely related genus to serve as an outgroup for all 12 taxa. 
We also sequenced another 15 taxa of plants commonly visited by *Bombus* workers, based on the abundances, and immediate access to plant tissue, in the aforementioned data set. 
Plant collections were identified typically using a combination, of dichotomous keys and primary literature as required (@flora1993flora, @hitchcock2018flora,  @ackerfield2015flora, @lesica2012manual,  @cronquist1977intermountain, @allred2012flora, @jepson2022online, @mohlenbrock2002vascular).

Plant genomic DNA was isolated from ~ 1 cm^2^ of leaf tissue from silica-gel dried or herbarium material using a modified cetyltrimethylammonium (CTAB) protocol (@doylesCTAB) that included two chloroform washes. DNA was quantified using a Nanodrop 2000 (Thermo Fisher Scientific, Waltham, Massachusetts, USA) and Qubit fluorometer (Thermo Fisher Scientific).

### Pollen DNA extraction

To extract genomic DNA from pollen a CTAB based protocol was modified from Lahlamgiahi et al. and Guertler et al. (2014, 2014).
A SDS extraction buffer (350µL , 100mM Tris-HCl, 50 mM EDTA, 50 mM NaCl, 10% SDS v/v., pH 7.5) was added to the sampled followed by vortexing to allow dissolution of corbiculae. 
Pollen grains were then macerated with Kontes Pellet Pestles, and the tip of these washed with 130 µL of the SDS extraction buffer, samples were then incubated for 1 hour at 30°C. 
This was followed by the addition of 10% CTAB solution (450ul, of 20 mM Tris-Cl pH. 8.0, 1.4 M NaCl, 10 mM EDTA pH 7.5, 10% CTAB, 5% PVP, ~85% Deionized water) and RNAse (10 uL of 10 mg/mL) and samples were incubated for 40 minutes at 37°C, on a heat block (Multi-Blok, Thermo Fisher Scientific, Waltham Massachusetts) set to 40°C. 
After 20 minutes incubation, Proteinase K (15 µL of 20mg/ml) and  DTT (12.5 µL of 1M in water) were added, and the samples were further incubated at 60°C for 1 hour. 
Samples were then incubated overnight at 40°C. 500 µL of Phenol-Chloroform-Isoamyl alcohol (25:24:1) were added, vortexed, and centrifuged at 10,000 rpm for 10 minutes and the aqueous phase was pipetted to a 1.5 ml centrifuge tube.

To precipitate the DNA, chilled Isopropyl alcohol & 3 mM Sodium acetate (5:1) equivalent to $\frac{2}{3}$ of the volume of sample were added, with 1 hour of chilling at -20°C, followed by 10 minutes of centrifuging at 13,000 rpm. 
The supernatant was pipetted to a new 1.5 ml centrifuge tube, and 70% EtOH (400 µL) were added before chilling at -20°C for 20 minutes followed by centrifugation at 13,000 rpm for 10 minutes. 
Both tubes were then washed with  75% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded, then washed with 95% EtOH (400 µL), inverted, centrifuged at 13,000 rpm for 4 minutes, and the solution discarded.
Pellets were dried at room temperature overnight before resuspension in nuclease free H~2~O. 
Extractions were assessed using a Nanodrop 2000 (Thermo Fisher Scientific) and Qubit fluorometer (Thermo Fisher Scientific). 
DNA extracts were then cleaned using 2:1 v./v. Sera-Mag beads (Cytiva, Little Chalfont, UK) to solute ratio following the manufacturer's protocol, eluted in 0.5x TE, and the eluent allowed to reduce by half volume in ambient conditions. 
DNA was quantified using a Qubit fluorometer.

### Library preparation & Bait capture (Barcoding)

Sequence library preparation was performed using the NEBNext Ultra II FS-DNA Library Prep Kit for Illumina (New England BioLabs, Ipswich, Massachusetts, USA) using slightly modified manufacturers recommendation. 
Fragmentation was performed at ½ volume of reagents and ¼ enzyme mix for 40 minutes at 37°C, with an input of 500 ng cleaned DNA.
Adapter Ligation and PCR enrichment were performed with ½ volumes, while cleanup of products was performed using SPRI beads (Beckman Coulter, Indianapolis, Indiana, USA) and recommended volumes of 80% v./v. ethanol washes. 
The exception was the herbarium specimens which were not fragmented and only end repaired, with similar library preparation of all samples. Products were analysed on 4% agarose gels, and a Qubit fluorometer. 
Libraries were pooled and enriched with the Angiosperms 353 probe kit V.4 (Arbor Biosciences myBaits Target Sequence Capture Kit) by following the manufacturer's protocol and Brewer et al. 2019. 
Sequencing was performed using an Illumina mi-Seq with 150-bp end reads, (NUSeq Core, Chicago, Illinois).

### Bioinformatics

Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950). 
Contigs generated were mapped to a reference with HybPiper with using target files created by M353 (@johnson2016hybpiper, @mclay2021new). 
A custom Kraken2 database was created by downloading representative species indicated as being present in the study area by the spatial analyses from the Sequence Read Archive (SRA) NCBI (@wood2019improved). 
These sequences were processed in the same manner as our novel sequences. The Kraken2 database was built using default parameters.
Kraken2 was run on sequences using default parameters. 
Following Kraken2, Bracken was used to classify sequences to terminal taxa (@lu2017bracken). 
Finally all reads which could be classified by these databases were passed to a local BLAST database.
A local NCBI database was built using the same processed novel and downloaded sequences as the previous database (@camacho2009blast).

\newpage

### 2024-03-03 script/CoverLetter.Rmd
making submission docs

@@
 Methods in Ecology and Evolution \newline
+Senior Research Fellow in Ecology \textit{Emeritus} \newline
+Harvard Forest \newline
 \end{flushleft}
@@
 
-In this paper, we combine Hyb-Seq to generate *de novo* sequence databases, informed by museum records, and species distribution modelling. 
-Further we use the results of phenological modelling to interpret our sequence alignment results to increase the likelihood of accurate results.
+In this paper we develop an intuitive approach to generate plant reference sequence databases tailored to specific areas for metabarcoding projects, as well as a lens to critically evaluate metagenomic results using phenological data.
+This spatio-temporal approach utilizes, among other sources - museum and citizen science records, species distribution modelling, and the weibull distribution, to create baseline environmental data in an approach which is globally applicable.
+Further we use bait-capture, via the popular Angiosperms 353 probes, as our method to generate sequence reads for metabarcoding.
 
@@
 
-We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and that several manuscripts utilizing them will be forthcoming to a variety of journals.
+We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and are aware of several manuscripts utilizing them which will be forthcoming in a variety of journals.
 
-Our submission contains X supplementary materials, X figures and X tables. It unfortunately is currently at roughly 9,300 words. 
+Our submission contains X supplementary materials, X figures and X tables. It unfortunately is currently at roughly 8,900 words. 
 However *circa* 900 words are dedicated to a plain text overview of the methods, a section which we believe elucidates the integration of the various approaches, and clarifies adoption by subsequent users.
-Considerable length is dedicated to the methods, which contain five years of field work, and similarly thorough microscopy, and molecular work; but which we have cut by nearly 500 words.
+Considerable length is dedicated to the methods, which contain five years of field work, and similarly thorough microscopy, and molecular work; but which we have cut by nearly 1000 words.
+The full details or our methods are provided as supplementary materials.
 
-We have no conflicts of interest to disclose. Additionally, this work - nor any part of it - has been submitted to any other journal.
+We have no conflicts of interest to disclose.
+Additionally, this work - nor any part of it - has been submitted to any other journal.
+Although, it is cited and discussed in our Grimes Review which is currently in review for *Journal of Ecology*.
 

### 2024-03-03 script/CoverPage.Rmd
making submission docs

---
title: "CoverPage"
author: "steppe"
date: "2024-03-03"
output: pdf_document
---

**Authors and Affiliations:**    
Reed Clark Benkendorf^1,2^       
Emily J. Woodworth^1,2^       
Paul J. CaraDonna^1,2,3^      
Jane E. Ogilvie^3^      
Sophie Taddeo^1,2,4^      
Jeremie B. Fant^1,2^

**Addresses:**      
^1^ Chicago Botanic Garden, Glencoe, Illinois 60022, USA    
^2^ Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA    
^3^ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, Colorado 81224, USA    
^4^ Department of Environmental and Ocean Sciences, University of San Diego, San Diego, USA

**Acknowledgements** The authors would like to sincerely thank N. Zerega, P. Herendeen, H. Noble, A. McDonnell, and E. Loke for their assistance with acquiring herbarium loans, pollen identification, and preparing the genomic libraries;; I. Breckheimer for sharing spatial data;; herbarium curators B. Legler, R. Williams, E. Nelson;; and J. Reitheel and staff at the the RMBL for access to field sites and logistical support.

**Author Contributions**  R.C.B, J.O., J.B.F, S.T. conceived the ideas and designed methodology; J.O., P.J.C., R.C.B. and E.J.W. collected the data; R.C.B. analysed the data. All authors contributed critically to the drafts and gave final approval for publication.

*Data Availability Statements* The queries required to download all data used in this project are located in… All novel sequencing data are located at NCBI…

            A concise and informative title. Do not include the authorities for taxonomic names.
            A list of all authors' names with names and addresses of Institutions.
            The name, address and e-mail address of the correspondence author.
            A running headline of not more than 45 characters.
            Acknowledgments, data availability, conflict of interest and author contribution statements.

### 2024-03-03 script/SpecimensSequencedTable.Rmd
making submission docs

---
output: pdf_document
classoption: landscape
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
header-includes:
  - \usepackage{titlesec}
  - \pagenumbering{gobble}
  - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
---

\vspace{-25truemm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r Required Libraries, message=F, warning=F, results='hide'}
library(knitr)
library(kableExtra)
library(tidyverse)
```

```{r}
dat <- read.csv('../data/Reference_library_for_appendix.csv') %>% 
  arrange(Family, Taxon) %>% 
  unite('Locality', Country:County, sep = ', ') %>% 
  mutate(Locality = str_replace(Locality, 'U[.]S[.]A[.]', 'USA')) %>% 
  select(-Coordinates) %>% 
  rename(Herbarium = 'Herb.', Preservation = 'Pres.')

kable(
  dat, 
  caption = 'Samples Used to Supplement the Sequence Reference Library',
  scaledown = T, 
  booktabs = TRUE, 
  format  = "latex") %>% 
  kableExtra::kable_styling(latex_options="scale_down")

```

### 2024-03-03 script/StatementOnInclusion.Rmd
making submission docs

---
title: "StatementOnInclusion"
output: pdf_document
---

'Statement on inclusion'

### 2024-03-03 script/SupplementaryInformation.Rmd
making submission docs

@@
 ```{r, include=FALSE}
-knitr::opts_chunk$set(
-  comment = "#>", echo = FALSE, fig.width = 6
-)
+knitr::opts_chunk$set(echo = F, warning = F, message = F)
 ```
@@
 
-# Table S1
 ```{r}
-knitr::kable(df, caption = info, col.names = c('Variable', 'Data Source', 'Geographic Extent'), booktabs = TRUE) |> 
+knitr::kable(
+  df, 
+  caption = info, 
+  col.names = c('Variable', 'Data Source', 'Geographic Extent'),
+  booktabs = TRUE, 
+  format  = "latex",
+  longtable = FALSE
+  ) |> 
   kableExtra::row_spec(0,bold=TRUE) |>
@@
 ```{r Voucher Reference Specimen table 1, echo = F}
-knitr::include_graphics('../graphics/tables/specimen_tables-1.pdf')
-```
-
-\newpage
-
-```{r Voucher Reference Specimen table 2, echo = F}
-knitr::include_graphics('../graphics/tables/specimen_tables-2.pdf')
+knitr::include_graphics('../graphics/tables/specimen_table_v2.png')
 ```

### 2024-03-29 script/CoverLetter.Rmd
appendix remade, similar to last...

@@
 ---
-output: pdf_document
+output:
+  pdf_document: default
+  word_document: default
 ---
@@
 Methods in Ecology and Evolution \newline
-Senior Research Fellow in Ecology \textit{Emeritus} \newline
-Harvard Forest \newline
 \end{flushleft}
 
-March 17^th^, 2024
+\hfill March 17^th^, 2024
 
-Attached is our standard article submission "*Protocol to increase accuracy and fidelity of pollen meta-genomic barcoding using Angiosperms353: a case study using pollen loads from wild bumble bees*".
+Dear Dr. Ellison,
 
+We are writing to propose an original research article entitled  "*Protocol to increase accuracy and fidelity of pollen meta-genomic barcoding using Angiosperms353: a case study using pollen loads from wild bumble bees*".
+
+Many meta-barcoding studies that include plant material are limited due to lack of reliable universal barcodes and the extensive potential candidate list (~350,000 plant species) that requires extensive computational and resource to identify best matches.
 In this paper we develop an intuitive approach to generate plant reference sequence databases tailored to specific areas for metabarcoding projects, as well as a lens to critically evaluate metagenomic results using phenological data.
-This spatio-temporal approach utilizes, among other sources - museum and citizen science records, species distribution modelling, and the weibull distribution, to create baseline environmental data in an approach which is globally applicable.
-Further we use bait-capture, via the popular Angiosperms 353 probes, as our method to generate sequence reads for metabarcoding.
+This spatio-temporal approach utilizes, among other sources - museum and citizen science records, species distribution modelling, and the weibull distribution -  to create baseline environmental data in an approach which is globally applicable
 
-Given the rising popularity of metabarcoding and metagenomics, especially for projects over large spatial domains and with multiple seasonal time points, we believe that our method provides a framework for meaningfully assessing and contextualizing results.
-We have observed that many studies utilizing metabacoding generate spurious results, which are precluded due to spatial or temporal incongruities with the focal animal species and plants. 
-However, a simple and intuitive framework for programmatically assessing and rectifying these issues are lacking.
-
+Further, we use bait-capture, via the popular Angiosperms 353 probes, to generate multiple sequence reads to increase the accuracy of species identification.
 We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and are aware of several manuscripts utilizing them which will be forthcoming in a variety of journals.
 
-Our submission contains X supplementary materials, X figures and X tables. It unfortunately is currently at roughly 8,900 words. 
-However *circa* 900 words are dedicated to a plain text overview of the methods, a section which we believe elucidates the integration of the various approaches, and clarifies adoption by subsequent users.
-Considerable length is dedicated to the methods, which contain five years of field work, and similarly thorough microscopy, and molecular work; but which we have cut by nearly 1000 words.
-The full details or our methods are provided as supplementary materials.
+Finally, many studies utilizing metabacoding generate spurious results, which are spatially or temporally incongruities with the focal animal species and plants/ There are few simple and intuitive framework for programmatically assessing and rectifying these issues are lacking.
+In this study, we also provide a framework for meaningfully assessing and contextualizing the meta-barcoding results.
 
 We have no conflicts of interest to disclose.
-Additionally, this work - nor any part of it - has been submitted to any other journal.
-Although, it is cited and discussed in our Grimes Review which is currently in review for *Journal of Ecology*.
-
-Please address all correspondence regarding this manuscript to me at rbenkendorf@chicagobotanic.org
-
-Thank you for considering this work.
-
-\begin{flushleft}
-Reed Clark Benkendorf \newline
-Senior (Field) Botanist \& Senior Spatial Data Specialist  \newline
-The Chicago Botanic Garden  \newline
-rbenkendorf@chicagobotanic.org
-\end{flushleft}
-
-\begin{flushleft}
-Jeremie B. Fant, Ph.D \newline
-Medard and Elizabeth Welch Director of Conservation \newline
-The Chicago Botanic Garden  \newline
-jfant@chicagobotanic.org
-\end{flushleft}

### 2024-03-29 script/StatementOnInclusion.Rmd
appendix remade, similar to last...

@@
 ---
-title: "StatementOnInclusion"
-output: pdf_document
+title: "Statement On Inclusion"
+output:
+  word_document: default
+  pdf_document: default
 ---
 
-'Statement on inclusion'
+Our studies focal area was within the Southern Rockies Mountains of Western North America and drew on natural history data collected throughout the region. 
+Two of our authors are *staff?affiliates?* at the research station which novel data was collected at. 
+Prior to matriculation our corresponding author worked at the largest institute for applied ecology in the Rocky Mountains, and has since worked as a contractor for the largest land management agency in the United States and is both based in and works in this geographic area.

### 2024-03-29 script/SupplementaryInformation.Rmd
appendix remade, similar to last...

@@
 author: "steppe"
-output: pdf_document
+output:
+  pdf_document: default
 ---
@@
 
-# Site Info
+Figure 1 - Site Info
 
@@
 
-# Table S2 - Models used for Species Distribution Model Ensembles
+Table 2 - Models used for Species Distribution Model Ensembles
 
@@
 
-Table S3 - Species Distribution Models Predictors
+Table 3 - Species Distribution Models Predictors
 
@@
 
-Appendix 3 - Molecular Reference Specimen Table
+Table 4 - Molecular Reference Specimen Table
 
@@
 
-Appendix 6 - Pollen Key
+Figure 2 - Pollen Key
 
@@
 
-Appendix 9 - All Species in the Sequence Databases
+Table 5 - All Species in the Sequence Databases
 
@@
 
-Appendix 11 - Reads Per Loci
+Figure 3 - Reads Per Loci
 
@@
 
+\newpage
+
+Figure 4 - Percent Loci Matched by Sample 
 ```{r READS PER LOCI - METAGENOMICS, echo = F, out.width = '98%'}
@@
 
-Appendix 12 - Comparison of Kraken2, Bracken, and BLAST
+Figure 5 - Comparison of Kraken2, Bracken, and BLAST
 
@@
 \newpage
+
+Figure 6 - Microscopy Rarefaction, Abundance
+
+```{r, echo = F, out.width="90%"}
+knitr::include_graphics('../graphics/plots/SppAbundance.png')
+```
+
+\newpage
+
+Figure 7 - Microscopy Rarefaction, Richness
+
+```{r, echo = F, out.width="90%"}
+knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
+```
+
+\newpage
+
+Table 6 - Binary reclassification process
+
+```{r, echo = F, out.width="90%"}
+knitr::include_graphics('../graphics/tables/reclassify_reads_table.pdf')
+```

### 2024-03-29 script/methods_supplemental.Rmd
appendix remade, similar to last...

@@
 
+### Comparision of Sequence classifications pre and post processing
+
+Using wilcox_effsize, with a one-sided hypothesis of greater, we have strong evidence of (p < 1e-04, wilcoxsign_test) an effect size of 0.732 95% CI [0.57, 0.84, n = 40, bootstrap replicates = 1000] (@kassambara2023rstatix, @Hothorn2006coin).
+
 \newpage

### 2024-03-29 script/PartsOfWhole.R

@@ -1,0 +1,38 @@
+library(tidyverse)
+library(ggraph)
+library(igraph)
+
+edges <- flare$edges
+vertices <- flare$vertices
+
+mygraph <- graph_from_data_frame( edges, vertices=vertices )
+
+ggraph(mygraph, layout = 'circlepack') + 
+  geom_node_circle() +
+  theme_void()
+
+edges1 <- data.frame(
+  from =  c('A353', 'A353.RPTL', 'A353.RPTL.SF', 
+            'A353-leaf', 'A353.RPTL-leaf','A353.RPTL.SF-leaf'),
+  to = c('A353.RPTL', 'A353.RPTL.SF', 'A353.RPTL.SF.TF', 
+         'A353.RPTL-leaf','A353.RPTL.SF-leaf', 'A353.RPTL.SF.TF-leaf') 
+)
+
+vertices1 <- data.frame(
+  name = 
+    c('A353-leaf', 'A353.RPTL-leaf','A353.RPTL.SF-leaf', 'A353.RPTL.SF.TF-leaf', 
+      'A353', 'A353.RPTL','A353.RPTL.SF', 'A353.RPTL.SF.TF'),
+  size = c(14000, 1295, 425, 346, 0, 0, 0, 0), 
+  shortName = rep(c('Angiosperms 353', 'Regional Plant Taxa List',
+                'Spatial Filtering', 'Temporal Filtering'), 2)
+#  Process = c('Of the 350,000+ known plant species, A353 data exist for 14,000+', 
+#              'Occurrence data create a list of candidate taxa likely to occur in the study area', 
+#              'Spatial filtering of Occurrence data retains species likely to be present in study sites', 
+#              'Temporal filtering of taxa via modelling informs metabarcoding results at multiple time points'
+#  )
+)
+
+mygraph1 <- graph_from_data_frame( edges1, vertices=vertices1 )
+ggraph(mygraph1, layout = 'circlepack') + 
+  geom_node_circle() +
+  theme_void()

### 2024-03-29 script/bombus_summaries2015-2020.R

@@ -1,52 +1,0 @@
-setwd("~/Documents/ManuSCript/script")
-
-library(tidyverse)
-
-commonCrestedButteBees <- c('nevadensis', 'californicus', 'appositus', 'flavifrons', 'rufocinctus', 'mixtus', 'bifarius')
-
-dat <- read.csv('../data/Bombus_queen_observations_2015-2022.csv') %>% 
-  filter(year < 2021 & resource.coll != 'nectar', species %in% commonCrestedButteBees)
-
-dat %>% 
-  select(genus, species, plant.species) %>% 
-  distinct() %>% 
-  filter(species %in% c('nevadensis', 'californicus', 'appositus', 'flavifrons', 'rufocinctus', 'mixtus', 'bifarius')) %>% 
-  group_by(species) %>% 
-  count() %>% 
-  pull(n) %>% 
-  median()
-
-pl_stats <- dat %>% 
-  select(plant.species) %>% 
-  separate(plant.species, c('genus', 'species'), remove = F) 
-
-pl_stats %>% 
-  distinct(plant.species) %>% 
-  nrow()
-
-pl_stats %>%  # no of genera by Weber
-  distinct(genus) %>% 
-  nrow()
-
-pl_stats %>% 
-  distinct(genus)
-
-pl_stats %>% 
-  mutate(Family = case_when(
-    genus %in% c('Heracleum', 'Osmorhiza', 'Taraxacum') ~ 'Apiaceae',
-    genus %in% c('Dugaldia', 'Helianthella', 'Heliomeris', 'Erigeron', 'Senecio') ~ 'Asteraceae', 
-    genus %in% c('Mertensia') ~ 'Boraginaceae',
-    genus %in% c('Calochortus' ,'Erythronium') ~ 'Liliaceae', 
-    genus %in% c('Distegia') ~ 'Caprifoliaceae',
-    genus %in% c('Lathyrus', 'Lupinus', 'Vicia' ) ~ 'Fabaceae',
-    genus %in% c('Frasera') ~ 'Gentianaceae',
-    genus %in% c('Hydrophyllum', 'Phacelia') ~ 'Hydrophyllaceae',
-    genus %in% c('Adenolinum') ~ 'Linaceae',
-    genus %in% c('Claytona') ~ 'Montiaceae',
-    genus %in% c('Pedicularis') ~ 'Orobanchaceae', 
[…]

### 2024-03-29 script/bombus_summaries2015-2022.R

@@ -1,0 +1,58 @@
+setwd("~/Documents/ManuSCript/script")
+
+library(tidyverse)
+
+commonCrestedButteBees <- c('nevadensis', 'californicus', 'appositus', 'flavifrons', 'rufocinctus', 'mixtus', 'bifarius')
+
+dat <- read.csv('../data/Bombus_queen_observations_2015-2022.csv') %>% 
+  filter(resource.coll != 'nectar', species %in% commonCrestedButteBees)
+
+plant_spp_by_bee <- dat %>% 
+  select(genus, species, plant.species) %>% 
+  distinct() %>% 
+  filter(species %in% c('nevadensis', 'californicus', 'appositus', 'flavifrons', 'rufocinctus', 'mixtus', 'bifarius')) %>% 
+  group_by(species) %>% 
+  count() %>% 
+  pull(n) 
+
+plant_spp_by_bee
+min(plant_spp_by_bee)
+mean(plant_spp_by_bee)
+median(plant_spp_by_bee)
+max(plant_spp_by_bee)
+
+pl_stats <- dat %>% 
+  select(plant.species) %>% 
+  separate(plant.species, c('genus', 'species'), remove = F) 
+
+pl_stats %>%  # number of plant species. 
+  distinct(plant.species) %>% 
+  nrow()
+
+pl_stats %>%  # no of genera by Weber
+  distinct(genus) %>% 
+  nrow()
+
+pl_stats %>%  # the names of the genera
+  distinct(genus)
+
+pl1 <- pl_stats %>% 
+  mutate(Family = case_when(
+    genus %in% c('Heracleum', 'Osmorhiza', 'Taraxacum') ~ 'Apiaceae',
+    genus %in% c('Dugaldia', 'Helianthella', 'Heliomeris', 'Erigeron', 'Senecio') ~ 'Asteraceae', 
+    genus %in% c('Mertensia') ~ 'Boraginaceae',
+    genus %in% c('Valeriana') ~ 'Valerianaceae',
+    genus %in% c('Calochortus' ,'Erythronium') ~ 'Liliaceae', 
+    genus %in% c('Distegia') ~ 'Caprifoliaceae',
+    genus %in% c('Lathyrus', 'Lupinus', 'Vicia' ) ~ 'Fabaceae',
+    genus %in% c('Frasera') ~ 'Gentianaceae',
[…]

### 2024-03-30 script/SpecimensSequencedTable.Rmd
ideally simplyfing repo for code submission

@@
   - \usepackage{titlesec}
+  - \usepackage{caption}
   - \pagenumbering{gobble}
@@
 
+\captionsetup[table]{labelformat=empty}
 \vspace{-25truemm}

### 2024-03-30 script/SupplementaryInformation.Rmd
ideally simplyfing repo for code submission

@@
 ---
-
 ```{r, include=FALSE}
@@
 ```{r Create Table 1}
-
 variable <- c(
@@
   kableExtra::kable_styling(latex_options="scale_down") |>
-  kableExtra::add_footnote('CONUS; CONtinental United States')
+  kableExtra::add_footnote('CONUS; CONtinental United States') |>
+  kableExtra::kable_styling(latex_options = "HOLD_position")
 
@@
 
-```{r Site Maps, echo = F}
+```{r Site Maps, echo = F, out.width="90%"}
 knitr::include_graphics('../graphics/plots/siteMaps.png')
@@
 
-```{r Voucher Reference Specimen table 1, echo = F}
+```{r Voucher Reference Specimen table 1, echo = F, out.height="90%"}
 knitr::include_graphics('../graphics/tables/specimen_table_v2.png')
@@
 
+\newpage
 Figure 2 - Pollen Key
@@
 
-Table 6 - Binary reclassification process
+Table 6 - Binary Reclassification Process
 

### 2024-03-30 script/pollen_countr3.Rmd
ideally simplyfing repo for code submission

@@
 title: "pollen_countR"
-author: "Reed & Emily"
+author: "steppe & EJW"
 date: "4/9/2021"

### 2024-03-30 script/Filter.R

@@ -1,104 +1,0 @@
-library(tidyverse)
-library(sf)
-library(terra)
-library(RColorBrewer)
-library(ggtext)
-library(ggnewscale)
-
-setwd('~/Documents/ManuSCript/script')
-
-# first we will create the shape of the filter itself. We will turn into an 
-# explicitly spatial object using the 'sf' package. We will do this so that we
-# are able to color the inside of the Filter shape with a gradient of colors, 
-# which is defined within a geom_raster, created by the Terra package. 
-
-filt <- st_polygon( 
-    list(
-      rbind(
-        c(43, 1), c(57, 15), # bottom portion right
-        c(57, 60), # stem portion right
-        c(85, 100), c(15, 100), # very top of filter
-        c(43, 60), # stem portion left
-        c(43, 1)  # bottom portion left
-      )
-    )
-  )  %>% 
-  st_geometry()  %>% 
-  st_as_sf(crs = 26913) %>% 
-  mutate(Attribute = 'Filter') %>% 
-  rename('geometry' = x) %>% 
-  st_as_sf()  
-  
-# we create an empty raster the size and extent of the filter, and mask it
-# to the shape to keep only pixels within the lines in; we are drawing between 
-# the lines! FINALLY after all these years!!
-
-filt_v <-   vect(filt)
-r <- rast(filt_v, nrow = 1000, ncol = 1000)
-r <- setValues(r, 1:1000000)
-r <- crop(r, filt_v, mask = T)
-r.df <- as.data.frame(r, xy = T)
-rm(r, filt_v)
-
-## we create labels and place these at positions to represent how many species 
-# exist at each stage of the filter 'head' (...the top part!!!) 
-
-filt_text <- data.frame(
-  'Value' = c(
-    'Angiosperms 353: 14,000+', 'Floristic Unit: 3226', 'Insect Pollinated: 2721', 
-    'Spatial Search: 501', 'SDM: 426', 'Survey: 117' ),
-  'Approach' = c(rep('Coarse', times = 3), 
-                 rep('Computational', times = 2), 
-                 rep('Survey', times = 1)),
-  y = c(96, 90, 84, 78, 72, 66),
-  x = c(50, 50, 50, 50, 50, 50)
-) 
-
-# we create this label seperately so that we can readily rotate it 90 degrees. 
-time <- data.frame(
-  'Value' = 'temporal gradient',
-  'Approach' = 'Computational/Survey',
-  x = 50, 
[…]

### 2024-03-30 script/bombus_summaries2015-2022.R

@@ -21,6 +21,11 @@ mean(plant_spp_by_bee)
 median(plant_spp_by_bee)
 max(plant_spp_by_bee)
 
+dat %>% 
+  separate(plant.species, c('pl_genus', 'pl_species'), remove = F) %>% 
+  select(pl_genus, species) %>% 
+  distinct() 
+
 pl_stats <- dat %>% 
   select(plant.species) %>% 
   separate(plant.species, c('genus', 'species'), remove = F) 

### 2024-03-30 script/predictors_tables.r
dropped another script

@@ -1,30 +1,0 @@
-
-Layer                   Description                            Source
-----    ----------------------------------------------  ---------------------
-1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
-2.           Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
-3.           Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
-4.           Cloudiness seasonality 3 - MODIS            Wilson et al. 2016
-5.          Beginning of the frost-free period               Wang et al.
-6.             Climatic moisture deficit                     Wang et al.
-7.             Degree-days above 5C from                     Wang et al.
-8.             Mean annual precipitation                     Wang et al.
-9.          Mean annual precipitation as snow                Wang et al.
-10.             Temperature seasonality                      Wang et al.
-11.       2015 Percent Grass/Herbaceous cover - MODIS         (MOD44B)
-12.      2015 Percent Tree cover from Landsat 7/8              (GLCF)
-13.       Soil probability of bedrock (R Horizon)             SoilGrids
-14.         Soil organic carbon (Tonnes / ha)                 SoilGrids
-15.              Surface soil pH in H2O                       SoilGrids
-16.             Surface soil percent sand                     SoilGrids
-17.                    Soil USDA class                        SoilGrids
-18.                Topographic elevation                    EarthEnv DEM
-19.          Topographic elevation, moving window           EarthEnv DEM
-20.            Topographic percent slope                    EarthEnv DEM 
-21.            Topographic wetness index                    EarthEnv DEM 
-22.        Topographic aspect from EarthEnv DEM
-23.      Annual potential solar radiation computed             r.sun.
-24.       Estimated actual (w/-cloud) solar radiation
-25.       Log-transformed distance to surface water   Global Surface Water Explorer
-26.             Percent surface water from            Global Surface Water Explorer 
-----    ----------------------------------------------  ---------------------
